### 2021-08-28 manuscript/Manuscript.tex
initializing manuscript via overleaf commit

\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers

\draftfalse

%% Enter journal name below.
%% Choose from this list of Journals:
%
% JGR: Atmospheres
% JGR: Biogeosciences
% JGR: Earth Surface
% JGR: Oceans
% JGR: Planets
% JGR: Solid Earth
% JGR: Space Physics
% Global Biogeochemical Cycles
% Geophysical Research Letters
% Paleoceanography and Paleoclimatology
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resources Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science
% Geohealth
%

\journalname{JGR: Solid Earth}

\begin{document}

\title{High geomagnetic field intensity in the late Mesoproterozoic recorded by Midcontinent Rift anorthosite xenoliths}

\authors{Yiming Zhang\affil{1}, Nicholas L. Swanson-Hysell\affil{1}, James D. Miller Jr.\affil{2}, Margaret S. Avery\affil{1,3}}

\affiliation{1}{Department of Earth and Planetary Science, University of California, Berkeley, CA, USA}
\affiliation{2}{Department of Earth and Environmental Sciences, University of Minnesota, Duluth, MN, USA}
\affiliation{3}{Geology, Minerals, Energy, and Geophysics Science Center, U.S. Geological Survey, Moffett Field, CA, USA}
\correspondingauthor{Yiming Zhang}{yimingzhang@berkeley.edu}

\begin{keypoints}
\item new ca. 1092 Ma Midcontinent Rift anorthosite record high paleointensity values
\item high VADM of the geomagnetic field during the late Mesoproterozoic is inconsistent with previous interpretation of long term decay
\item a strong geodynamo like today likely persisted throughout 25 myr during the Midcontinent Rift development
\end{keypoints}

A few days into the new year of 2020, I arrived at Twin Cities, Minnesota for the first time, with an exciting rock magnetic project to be conducted at the Institute for Rock Magnetism. My research was focused on developing paleomagnetic data from the 1.1 Ga Beaver River diabase and its anorthosite xenoliths of the Beaver Bay Complex in North Shore, Minnesota. Before this trip, I had been developing directional and intensity data at the paleomagnetism lab at UC Berkeley. With the help of rock magnetic experiments at the IRM, I was hoping to decipher the cause of a variety of paleointensity results of my samples.

Retrieving paleomagnetic information from Precambrian rocks can aid in our understanding of the long-term evolution of the geodynamo. A recent paleointensity study by \cite{Sprain2018a} reported a ca. 1.1 Ga paleomagnetic field strength similar to today based on the volcanics of the Midcontinent Rift System (MRS). That study suggested that this result is consistent with models wherein this time period of the Proterozoic is characterized by a strong rather than weak geomagnetic field. However, to further evaluating the evolution of the geomagnetic field during the Precambrian, more paleointensity data are needed.

%The MRS emplaced a vast amount of extrusive and intrusive rocks spanning a prolonged period of $\sim$25 Myrs. \cite{Swanson-Hysell2019a} published an extensive compilation of the MRS volcanics with paired high-precision CA-ID-TIMS U-Pb geochronology and high-quality paleomagnetic directions, revealing at least two dipole reversals and rapid differential plate motion of the ancient North American plate, Laurentia, during the emplacement of MRS. The synthesized North American apparent polar wander path (APWP) was interpreted to be composed primarily of rapid plate motion \citep{Swanson-Hysell2019a}. Through comparison between paleomagnetic pole positions and the APWP, such rapid plate motion and high-precision paleomagnetic and geochronological framework revealed by the MRS volcanics can provide insights into age constraints on units with poor or no geochronology data. This framework allows for further paleomagnetic contributions beyond the Midcontinent Rift volcanics.

  The hypabyssal intrusions of the ophitic Beaver River diabase (BRD) and the Ca-rich anorthosite xenoliths that it hosts are the targets of this study. The diabase contains abundant Fe-Ti oxides which are dominant carriers of paleomagnetic information. The nearly-pure anorthosite xenoliths have potential for paleomagnetic investigation because plagioclase hosts can protect the exsolved magnetic minerals from post-formation alteration \citep{Tarduno2005a}, and thus may faithfully record paleointensity information at the time of formation. A typical outcrop of anorthosite xenolith residing in the diabase and a hand sample of the anorthosite are shown in Fig. \ref{fig:field}(A, B). Preliminary petrographic analyses show that the anorthosites are dominantly monomineralic with recrystallization textures (Fig. \ref{fig:field}C). Crosscutting relationships together with high-precision geochronology dates from \cite{Swanson-Hysell2019a} bracket the age of the BRD to be between 1091.61 $\pm$ 0.14 Ma  (the age of the younger Silver Bay intrusions) and 1093.94 $\pm$ 0.28 Ma  (the age of the Palisade rhyolite).

\begin{figure}
\noindent\includegraphics[width=\textwidth]{Field_photo_petro_photo.pdf}
\caption{\small{(A) A typical outcrop of an anorthosite xenolith residing in the Beaver River diabase. (B) A hand sample of anorthosite. The large reflective face at the bottom center is a cleavage plane of a centimeter-size plagioclase crystal. The scale is 9 cm in total. (C) Cross-polarized light image of an anorthosite thin section. The second order birefringence of plagioclase indicates a high Ca content. The closely packed plagioclase crystals show recrystallization texture. (D) High-magnification SEM image of Fe-Ti oxides exsolved from a pyroxene crystal included in a plagioclase crystal. (E) Large, interstitial magnetite-ilmenite intergrowths between two plagioclase crystals. (F) A large Fe-Ti oxide grain with magnetite-ilmenite intergrowths next to a pyroxene in diabase. an: anorthite; px: pyroxene; mag: magnetite; ilm: ilmenite}}
\label{fig:field}
\end{figure}

  Most thermally demagnetized anorthosites and alternating field (AF) demagnetized diabase specimens have minimal secondary components and their magnetization is interpreted to be dominated by a primary thermal remanent magnetization. Characteristic magnetizations from both lithologies yield indistinguishable site-mean directions and the virtual geomagnetic poles (VGPs) fall close to the 1095 Ma pole from the synthesized APWP of \cite{Swanson-Hysell2019a}, which is in agreement with the geochronological constraints  (Fig. \ref{fig:pmag}).

I conducted a comparative study of modified IZZI paleointensity experiments on both the diabase and anorthosite, with a group treated with low-field AF demagnetization after each in-field step and another group without. Paleointensity experiments for most diabase specimens result in non-ideal Arai plots often with poor pTRM checks, making them difficult to interpret for paleointensity estimate (Fig. \ref{fig:pmag}). These results can be interpreted to be associated with observed color change of the diabase after heating, which suggests alterations of the oxides. The experiments conducted on anorthosite samples, on the other hand, had a high success rate of $>$ 50\%. Moreover, the resulting success rate for the anorthosite group with the AF treatment is even higher. The more straight Arai plots are likely due to the AF steps mitigating the exhibition of pTRM tails often associated with non-ideal behavior of multi-domain grains that are demagnetized by the pre-treatment step. However, as the nearly pure anorthosite samples did not display dramatic alterations that are easily detectable with common petrographic techniques, a question rises as why some of the compositionally similar anorthosites passed the paleointensity selection results and some did not.

\begin{figure}
\noindent\includegraphics[width=\textwidth]{pmag_plot_.pdf}
\caption{\small{(A) Site-mean directions of anorthosite and diabase plotted on an equal-area plot. (B) Calculated VGPs from the anorthosite and diabase plotted in context of a previously synthesized ca. 1.1 Ga Laurentia APWP from the Midcontinent Rift volcanics \citep{Swanson-Hysell2019a}. VGPs from both lithologies fall close to the 1095 Ma pole position. (C) Example Arai plots for diabase. Most diabase and many anorthosite specimens were rejected by selection criteria due to the zigzagging behavior shown in the diabase example plot. (D) Example Arai plots for anorthosite. A typical anorthosite specimen that passes selection shows straight Arai plot with a high paleointensity estimate before cooling rate correction. Note both specimens in (C) and (D) show dominantly single component magnetization in the inset orthogonal plots. The estimated field intensity is not cooling rate corrected.}}
\label{fig:pmag}
\end{figure}

At the IRM, I was seeking the answer to this question with the help of the vibrating sample magnetometer (VSM) systems, including a newly installed Lake Shore VSM which greatly helped improve measurement resolution on samples with weak magnetizations (Fig. \ref{fig:backfield}). Backfield demagnetization experiments were conducted and used to develop coercivity spectra (Fig. \ref{fig:backfield}). The spectra were subsequently modeled to fit for the distributions of different  populations of magnetic particles using similar procedure as \cite{Maxbauer2016a}. Most spectra can be well approximated with models with one component or two components with overlapping coercivity ranges (Fig. \ref{fig:backfield}). The dominance of single-component distributions from the unmixed coercivity spectra likely suggests minimal alterations or formation of secondary magnetic mineral within the samples. By compiling the values of median destructive field (MDF) from all measured specimens and categorizing them in terms of their sister specimens' paleointensity results, I found that the anorthosites samples that pass paleointensity selection criteria have distinctively higher MDF values than those did not (Fig. \ref{fig:backfield}), and all diabase specimens have low MDF values like the non-ideal anorthosites (Fig. \ref{fig:backfield}). The higher MDFs ($\sim$60 mT) of the anorthosites are similar to what is typical of stoichiometric, single-domain magnetite grains which are favored for paleointensity experiments. Furthermore, the low MDFs displayed in other specimens could be linked to a dominant population of large interstitial Fe-Ti oxides that are more likely to form multi-domain grains, which have been observed in both the anorthosite and diabase (Fig. \ref{fig:field}). In addition, preliminary tests on single anorthosite crystals might suggest that not all single crystals are necessarily better for paleointensity research when comparing to bulk samples, given their similar low MDF to the failed diabase and anorthosite specimens. However, more single crystal grains with paired paleointensity data and rock magnetic data may be needed before a general conclusion can be drawn.

Taken together, the rock magnetism experiment results support my paleointensity selection criteria in that rock specimens that produce straight Arai plots and have MDF similar to stoichiometric single-domain magnetite are preferentially selected. The preliminary paleointensity results yielded consistent estimates. The cooling rate corrected site-mean paleointensity estimate is about 40 $\mu$T, consistent with results from \cite{Sprain2018a} that the Earth's magnetic field strength at surface in the late Mesoproterozoic was close to that of today.

\begin{figure}
\begin{center}
	\noindent\includegraphics[width=0.5\textwidth]{backfield_example_.pdf}
\end{center}
\caption{\small{(A) Backfield measurement data acquired using a Princeton VSM on an anorthosite specimen. The noise level on coercivity plot is high and a heavy smoothing is needed to make subsequent coercivity modeling interpretable. (B) Backfield measurement data acquired using a Lake Shore VSM on the same specimen as that in (A). The inset coercivity unmixing plot shows that the much smoother initial backfield curve resulted in a tight single-component model fit. (C) Box plot of distribution of the median destructive field of all measured specimens. The anorthosite specimens that pass selection criteria have distinctly higher median destructive field (MDF) than other groups.}}
\label{fig:backfield}
\end{figure}

\begin{abstract}

\end{abstract}

\section*{Plain Language Summary}
[ enter your Plain Language Summary here or delete this section]

\section{Introduction}

\section{Method}

\section{Results}

\section{Discussion}

\section{Conclusion}

\acknowledgments
I thank Dario Bilardello, Peat Solheid, Mike Jackson, Josh Feinberg, and Bruce Moskowitz for their tremendous help of instrumental operations, data interpretations, and research guidance. I thank the IRM for the generous U.S. Visiting Student Fellowship. I thank Nicholas Swanson-Hysell at UC Berkeley for advising this research. Margaret Avery provided field assistance.

\bibliography{}

\end{document}

More Information and Advice:

### 2021-08-28 manuscript/Manuscript.tex
Updates from Overleaf

@@
 \begin{keypoints}
-\item new ca. 1092 Ma Midcontinent Rift anorthosite record high paleointensity values
-\item high VADM of the geomagnetic field during the late Mesoproterozoic is inconsistent with previous interpretation of long term decay
-\item a strong geodynamo like today likely persisted throughout 25 myr during the Midcontinent Rift development
+\item New ca. 1092 Ma Midcontinent Rift anorthosite record high paleointensity values
+\item High VADM of the geomagnetic field during the late Mesoproterozoic is inconsistent with previous interpretation of long term decay
+\item A strong geodynamo like today likely persisted throughout 25 myr during the Midcontinent Rift development
 \end{keypoints}
 
-A few days into the new year of 2020, I arrived at Twin Cities, Minnesota for the first time, with an exciting rock magnetic project to be conducted at the Institute for Rock Magnetism.
-My research was focused on developing paleomagnetic data from the 1.1 Ga Beaver River diabase and its anorthosite xenoliths of the Beaver Bay Complex in North Shore, Minnesota.
-Before this trip, I had been developing directional and intensity data at the paleomagnetism lab at UC Berkeley.
-With the help of rock magnetic experiments at the IRM, I was hoping to decipher the cause of a variety of paleointensity results of my samples.
+% keywords: absolute paleointensity, Laurentia, Midcontinent Rift, anorthosite,
 
+\begin{abstract}
+High-quality absolute paleointensity
+
+\end{abstract}
 Retrieving paleomagnetic information from Precambrian rocks can aid in our understanding of the long-term evolution of the geodynamo.
@@
 
-\begin{abstract}
+\section*{Introduction}
 
-\end{abstract}
+\section*{Method}
 
-\section*{Plain Language Summary}
-[ enter your Plain Language Summary here or delete this section]
+\section*{Results}
 
-\section{Introduction}
+\section*{Discussion}
 
-\section{Method}
-
-\section{Results}
-
-\section{Discussion}
-
-\section{Conclusion}
+\section*{Conclusion}
 

### 2021-08-28 manuscript/Manuscript.tex
manuscript text udpate

@@
 \item High VADM of the geomagnetic field during the late Mesoproterozoic is inconsistent with previous interpretation of long term decay
-\item A strong geodynamo like today likely persisted throughout 25 myr during the Midcontinent Rift development
+\item A strong geodynamo like today likely persisted for about 25 myr during the Midcontinent Rift development
 \end{keypoints}
 
-% keywords: absolute paleointensity, Laurentia, Midcontinent Rift, anorthosite,
+% keywords: absolute paleointensity, Laurentia, Midcontinent Rift, anorthosite, geodynamo
 
 \begin{abstract}
-High-quality absolute paleointensity
+New high-quality absolute paleointensity data from anorthosite xenoliths within the Beaver River diabase intrusions of the North American Midcontinent Rift (MCR) record a virtual axial dipole moment (VADM) estimate of $\sim$80 ZAm$^2$ ca.
+1092 Ma.
+This cooling rate corrected paleointensity estimate from an MCR intrusive unit is similar to previous results from extrusive units that span from ca 1108 Ma to ca.
+1085 Ma.
+Taken together, new and previously published paleointensity data from the Midcontinent Rift support that a geomagnetic field strength persisted through the Midcontinent Rift development and the data are inconsistent with the interpretation of a decay of Earth's geodynamo strength through the late Mesoproterozoic.
 
 \end{abstract}
-Retrieving paleomagnetic information from Precambrian rocks can aid in our understanding of the long-term evolution of the geodynamo.
+
+\section*{Introduction}
+%The question of the core and the importance of the observational data as complementary for numerical simulation work.
+
+That we have a solid inner core today is the result of secular cooling and differentiation of the layered Earth.
+The timing of inner core formation during Earth's thermal evolution is amongst the major unsolved questions in Earth's history.
+Due to the uncertain constraints on the thermal conductivity of the core \cite{Konopkova2016a, Gomi2013a, Ohta2016a}, numerical simulations have given drastically different predictions on the inner core nucleation age, ranging from about half a billion years ago to prior to the Archean \cite{Pozzo2012a, Gubbins2004a, Nimmo2015a}.
+Modeling has suggested that the emergence of the solid inner core could have expressed as changes of Earth's surface magnetic field strength due to the onset of compositional convection between the solid inner core and the liquid outer core \cite<e.g.>{Aubert2009a}.
+Thus, absolute paleointensity records preserved by magnetic minerals in igneous rocks may provide us with a probe into the birth of the inner core.
+Acquiring these observational data is crucial for advancing our understanding of Earth's thermal evolution and for testing existing models on the commencement of the inner core growth.
+However, due to the scarcity of well-preserved ancient rocks and low success rates typical of paleointensity experiments, only 9\% of the records in the current global paleointensity database (\href{http://earth.liv.ac.uk/pint/}{PINT, \cite{Biggin2010a}}) are from before the Cambrian and the existing records sparsely span over 3 billion years of time (Fig.
+\ref{fig:all_PINT_data}).
+
+\citeA{Bono2019a} filtered the available Precambrian paleointensity database (PINT) based on a set of quality criteria.
+A polynomial fit through the resulting paleointensity compilation from the Archean and the Proterozoic was used to interpret that a decrease trend of the estimated geomagnetic dipole strength occurred before the inner core started to form.
+In addition, \citeA{Bono2019a} recovered very low paleointensity values from ca.
+565 Ma single silicate crystals  that were interpreted to be consistent with it reflecting the weak geodynamo after progressive cooling of the core before the onset of the rigorous compositional convection driven by the nucleation of the solid inner core.
+Such decay trend is recently supported by new low paleointensity estimates derived from the ca.
+720 Ma Franklin Large Igneous Province \cite{Lloyd2021a}.
+
+However, previous paleointensity data from the MCR rock units \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record VADM values distinctly higher than what is predicted by the long term decay curve.
+These rock units notably are extrusive volcanics and span a tme period of about 25 myr during the development of the Midcontinent Rift magmatism.
+Many previous results from the Midcontinent Rift show non-ideal double-slope paleointensity behavior, which are difficult to interpret as they can yield two paleointensity estimates: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
+Nevertheless, single-slope, ideal behavior from the Lake Shore Traps \cite{Kulakov2013a} lava flows passed the selection criteria of \citeA{Bono2019a}, and a few lava flows yielded single-slope behavior from the Osler Volcanics \cite{Sprain2018a}.
+These data are inconsistent with the interpretation of a low geomagnetic field strength during the Midcontinent Rift because they yield significantly higher VADM estimates than that projected by the trend from \citeA{Bono2019a}.
+
+To further enrich the Precambrian paleointensity database, and to test the hypothesis that the Earth's geomagnetic dipole strength remained at high values throughout the Midcontinent Rift, in this study we perform ``IZZI" paleointensity experiments \cite{Tauxe2004a} on a suite of anorthosite xenoliths hosted by the intrusive Beaver River diabase of the Beaver Bay Complex within the Midcontinent Rift.
+
+%Retrieving paleomagnetic information from Precambrian rocks can aid in our understanding of the long-term evolution of the geodynamo.
 A recent paleointensity study by \cite{Sprain2018a} reported a ca.
@@
 
+\section*{Geologic Settings}
+The North American Midcontinent Rift (MCR) is a failed intracontinental rift where protracted magmatic activity lasted from \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}.
+Midcontinent Rift rocks extensively outcrop in today's Lake Superior region, with the total extent traceable by arcuate magnetic and gravity anomalies that extend to the southwest to Kansas, and to the southeast, to southern Michigan \cite{Hinze2020a}.
+Previous studies have divided magmatic activity in the rift into four stages based on interpreted changes in relative magmatic volume and the nature of magmatism: early ($\sim$1109–1104 Ma), latent ($\sim$1104–1098 Ma), main ($\sim$1098–1090 Ma) and late ($\sim$1090–1083 Ma) \cite{Vervoort2007a, Heaman2007a, Miller2013a}.
+In northeastern Minnesota, the Early Gabbro Series and the Felsic Series rocks of the Duluth Complex and reversed-polarity lavas of the lower North Shore Volcanic Group were emplaced during the early stage.
+The more voluminous Duluth Complex Layered Series and the plagioclase-rich Anorthositic Series, together with an associated $\sim$8 km thick extrusive volcanic sequences of the North Shore Volcanic Group (NSVG), were rapidly emplaced about 10 myr later at \textit{ca.} 1096 Ma during the main stage \cite{Paces1993a, Swanson-Hysell2020a}.
+
+The Beaver Bay Complex, which sits stratigraphically above the Duluth Complex, is another intrusive complex that resulted from main stage magmatism.
+The exposed area of the Beaver Bay Complex is $\sim$1000 km\textsuperscript{2} where it has been mapped along the northwestern shore of Lake Superior in northeastern Minnesota (Fig.
+\ref{fig:Geologic_map}).
+The Beaver Bay Complex is a multi-phase, composite intrusive complex that intrudes parts of the NSVG (Fig.
+\ref{fig:Geologic_map}; \citeA{Miller1997a, Swanson-Hysell2020a}).
+Distinct from the deep plutonic intrusions of the Duluth Complex, the majority of the Beaver Bay Complex is formed of hypabyssal intrusions that were emplaced as dikes and sills at shallow depths \cite{Miller1997a}.
+Most of the Beaver Bay Complex intrusions are dioritic to gabbroic in composition \cite{Miller1997a}.
+The main lithology of the Beaver River diabase dikes and sills network within the Beaver Bay Complex is an ophitic olivine gabbro (Fig.
+\ref{fig:Field_photo}), but in wider areas of dikes and the upper parts of thick sills, this rock type can abruptly transition into intergranular olivine oxide gabbro, then into subprismatic (and commonly foliated) ferrogabbro, and finally into granophyric monzodiorite.
+The more evolved and later emplaced components of the Beaver River diabase network are commonly distinguished as the Silver Bay intrusions in the southern Beaver Bay Complex (Fig.
+\ref{fig:Geologic_map}).
+Overall being intermediate in composition, the Silver Bay intrusions lithologies range from ophitic olivine gabbro to ferrogranite \cite{Shank1989a}.
+Field mapping by \citeA{Miller1994a} found intrusive relationships between the Silver Bay intrusions and the Beaver River diabase.
+Angular inclusions of the host Beaver River diabase within marginal zones of the Silver Bay intrusions led \citeA{Miller1997a} to interpret that the Silver Bay intrusions intruded after the diabase crystallized.
+
+One distinctive feature of the Beaver River diabase is its inclusions of anorthosite xenoliths.
+In the southern part of the Beaver Bay Complex, the Beaver River diabase occurs as dikes and sills, typically including anorthosites with various sizes ranging from centimeters to over 150 meters (Figs.
+\ref{fig:Geologic_map}, \ref{fig:Field_photo}; \citeA{Grout1939a, Morrison1983a}).
+The diabase in this region intrudes the Palisade rhyolite of the North Shore Volcanic Group (Fig.
+\ref{fig:Geologic_map}), which has a $^{206}$Pb/$^{238}$U date of 1093.94 $\pm$ 0.28 Ma (2$\sigma$ analytical uncertainty is presented for CA-ID-TIMS dates throughout this work; \citeA{Swanson-Hysell2019a}).
+The Beaver River diabase is locally intruded by the Silver Bay intrusions (Fig.
+\ref{fig:Geologic_map}).
+An aplite unit within the granophyre zone of one of these Silver Bay intrusions has a $^{206}$Pb/$^{238}$U date of 1091.61 $\pm$ 0.14 Ma \cite{Swanson-Hysell2019a}.
+Another arcuate, sill-like diabase body mapped as the Beaver River diabase outcrops along the eastern part of the complex (Fig.
+\ref{fig:Geologic_map}; \citeA{Miller1997a}).
+The diabase composition there is similar to that in the south and it also contains large anorthosite xenoliths with dimensions that exceed 100 meters at Carlton Peak (Fig.
+\ref{fig:Geologic_map}).
+The Beaver River diabase in the northern part of the complex, near the Houghtaling Creek area, typically forms narrow, near-vertical dikes instead of sheets in the southern and eastern regions (Fig.
+\ref{fig:Geologic_map}; \citeA{Miller1994a}).
+The diabase in this region only locally contains xenoliths of anorthosite.
+
+Hundreds of anorthosite xenoliths have been recognized and mapped within the Beaver River diabase (Fig.
+\ref{fig:Geologic_map}).
+Many hill tops in the Beaver Bay Complex, such as Carlton Peak and Britton Peak, are large anorthosite blocks (which lead \citeA{Lawson1893a} to erroneously conclude that they were relict Archean topography).
+Later work established the anorthosite blocks as xenoliths, which are now extensively documented through geologic mapping of the region (Fig.
+\ref{fig:Geologic_map}; \citeA{Miller2001a, Miller1988a, Miller1989a, Boerboom2004a, Boerboom2006a, Boerboom2006b, Boerboom2007a}) and outcrop-scale exposures (Fig.
+\ref{fig:Field_photo}).
+In the field, the anorthosites typically appear as subrounded to rounded, light-colored, translucent blocks that are in sharp contact with the hosting diabase (Fig.
+\ref{fig:Field_photo}).
+They also occur as exposures whose contact with the diabase is covered (Fig.
+\ref{fig:Field_photo}).
+\citeA{Grout1939a} suggested that the rounded anorthosites are the result of abrasion during transportation as they were entrained by the diabase (i.e.
+physical weathering within a magmatic system).
+While the Beaver River diabase is chilled against the North Shore Volcanic Group lithologies that it intrudes, the diabase is not chilled against the margin of the anorthosite xenoliths \cite{Morrison1983a, Miller1997a}.
+The lack of chilled contacts is consistent with the anorthosite being at elevated temperatures and cooling at the same time as the diabase magma (Fig.
+\ref{fig:thermal_history_model}).
+
+The anorthosite xenoliths are dominantly monomineralic plagioclase that has an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}.
+Interstitial pyroxene and olivine are present in minor concentrations in the xenoliths.
+Within the Carlton Peak anorthosite xenolith, up to 10 cm oikocrysts of olivine and pyroxene can occur.
+Nevertheless, the overall olivine content in the anorthosites is low.
+Interstitial titanomagnetite-ilmenite intergrowths that exceed 100 $\mu$m can be found with microscopy and $<$20 $\mu$m Fe-Ti oxide grains can be detected with scanning electron microscopy (Fig.
+\ref{fig:Field_photo}).
+Based on textural differences \citeA{Morrison1983a} divided the anorthosite xenoliths into four groups: one group which typically have well-developed granoblastic texture characterized by equigranular plagioclase crystals; another group which have interlocking, lath-shaped plagioclase crystals; an intermediate group which can have both granoblastic texture and interlocking plagioclase laths; and a brecciated group that have brittle deformation textures superposed on pre-existing textures.
+
 The hypabyssal intrusions of the ophitic Beaver River diabase (BRD) and the Ca-rich anorthosite xenoliths that it hosts are the targets of this study.
@@
 Crosscutting relationships together with high-precision geochronology dates from \cite{Swanson-Hysell2019a} bracket the age of the BRD to be between 1091.61 $\pm$ 0.14 Ma  (the age of the younger Silver Bay intrusions) and 1093.94 $\pm$ 0.28 Ma  (the age of the Palisade rhyolite).
+
+High-precision $^{206}$Pb/$^{238}$U zircon date of 1091.83 $\pm$ 0.21 from one anorthosite in Silver Bay area tightly constraint the age of the diabase intrusion to be 1091.7 $\pm$ 0.2 Ma.
+previous paleomagnetic directional data from \citeA{Zhang2021a} reveal that vast majority of the anorthosite xenoliths have minimal secondary remanence and single component characteristic remanence magnetizations carried by low-Ti magnetite.
+Step-wise demagnetization experiments show that the characteristic remanence often unblock sharply between temperature steps 500\textdegree C and 580\textdegree C.
+Moreover, the site mean paleodirection of the anorthosite xenoliths share a common mean with the directions of their host diabase.
+Together with thermal conduction modeling, these data are consistent with the interpretation that the anorthosite xenoliths within the Beaver River diabase acquired remanence during cooling with the diabase after their hypabyssal emplacement.
+A total of  xxx specimens from xx sites were used for paleointensity experiments.
+The locations of the xx sites are shown in Fig.
+1 and the corresponding site mean palaeomagnetic directions from \citeA{Zhang2021a} are shown in Fig.
+2(b).
+
+\begin{figure}
+\centering
+\noindent\includegraphics[width=4.75 in]{Figure/Field_photo.pdf}
+\caption{\footnotesize{Field photographs and petrographic images of the Beaver River diabase and the anorthosite xenoliths within it.
+(A) Centimeter-sized plagioclase megacrysts in the diabase.
+(B) Rounded anorthosite xenolith with a diameter of $\sim$7 meters fully enclosed within the diabase.
+(C) Exposure of a giant Carlton Peak anorthosite with a diameter $>$100 m.
+(D) 27.5 m diameter anorthosite xenolith sampled as paleomagnetic site AX16 and geochronology sample MS99033.
+(E) Cross polarized (XPL) image of the subophitic texture of diabase at site BD2 (pyroxene partially enclosing plagioclase).
+(F) XPL image of anorthosite geochronology sample MS99033.
+Plagioclase crystals exhibit both granoblastic texture and interlocking lath fabrics.
+(G) Backscattered electron (BSE) image of a large Fe-Ti oxide with titanomagnetite-ilmenite lamellae in Beaver River diabase site BD2.
+(H) BSE image of micron-sized Fe-Ti oxides exsolved from pyroxene between plagioclase crystals in anorthosite xenolith site AX4.
+an-plagioclase with $\sim$70\% anorthite; ilm-ilmenite; mag-magnetite; px-pyroxene.}}
+\label{fig:Field_photo}
+\end{figure}
 
@@
 
-\section*{Introduction}
-
 \section*{Method}
@@
 
-\bibliography{}
+\bibliography{YZ_ref}
 

### 2021-09-12 code/BiCEP_modules_and_scripts/BiCEP_functions.py
clean up PINT data, start BiCEP analyses

@@ -1,0 +1,1864 @@
+import pandas as pd
+from importlib import reload # allows reloading of modules
+import numpy as np
+import matplotlib.pyplot as plt
+import ipywidgets as widgets
+from IPython.display import display, clear_output
+import asyncio
+import pickle
+import pmagpy.pmag as pmag
+import pmagpy.ipmag as ipmag
+from matplotlib import cm
+from pmagpy import contribution_builder as cb
+import scipy as scipy
+from sklearn.decomposition import PCA
+from scipy.optimize import curve_fit
+import os
+THIS_FOLDER = os.path.dirname(os.path.abspath(__file__))
+fast_file = os.path.join(THIS_FOLDER, 'model_circle_fast.pkl')
+slow_file = os.path.join(THIS_FOLDER, 'model_circle_slow.pkl')
+
+model_circle_fast=pickle.load(open(fast_file,'rb'))
+model_circle_slow=pickle.load(open(slow_file,'rb'))
+
+def sufficient_statistics(ptrm, nrm):
+    """
+    inputs list of ptrm and nrm data and computes sufficent statistcs needed
+    for computations
+
+    Inputs
+    ------
+    ptrm: list
+    list of ptrm data
+
+    nrm: list
+    list of nrm data
+
+    Returns
+    -------
+    dict containing mean ptrm and nrm, and covariances in xx, xy and yy.
+    """
+
+    corr = np.cov( np.stack((ptrm, nrm), axis=0) )
+
+    return {'xbar': np.mean(ptrm), 'ybar': np.mean(nrm), 'S2xx': corr[0,0], 'S2yy': corr[1,1], 'S2xy': corr[0,1] }
+
+def TaubinSVD(x,y):
+    """
+    Function from PmagPy
+    algebraic circle fit
+    input: list [[x_1, y_1], [x_2, y_2], ....]
+    output: a, b, r.  a and b are the center of the fitting circle, and r is the radius
+
+     Algebraic circle fit by Taubin
+      G. Taubin, "Estimation Of Planar Curves, Surfaces And Nonplanar
+                  Space Curves Defined By Implicit Equations, With
+                  Applications To Edge And Range Image Segmentation",
+      IEEE Trans. PAMI, Vol. 13, pages 1115-1138, (1991)
+    """
+    X = np.array(list(map(float, x)))
+    Xprime=X
+    Y = np.array(list(map(float, y)))
+    Yprime=Y
[…]

### 2021-09-12 code/BiCEP_modules_and_scripts/BiCEP_functions.py
saving anorthosite PINT BiCEP analyses

@@ -2,6 +2,7 @@ import pandas as pd
 from importlib import reload # allows reloading of modules
 import numpy as np
 import matplotlib.pyplot as plt
+import matplotlib
 import ipywidgets as widgets
 from IPython.display import display, clear_output
 import asyncio
@@ -517,8 +518,8 @@ class SpecimenCollection():
         ax.set_xlabel('Intensity ($\mu$T)')
         ax.set_ylabel('Probability Density')
 
-        return
-        
+        return np.percentile(self.fit['int_site'],(2.5, 50, 97.5))
+
 class Specimen():
     """
     Specimen from a given site or sample SpecimenCollection object.

### 2021-09-13 manuscript/Manuscript.tex

@@
 
-\authors{Yiming Zhang\affil{1}, Nicholas L. Swanson-Hysell\affil{1}, James D. Miller Jr.\affil{2}, Margaret S. Avery\affil{1,3}}
+\authors{Yiming Zhang\affil{1}, Nicholas L. Swanson-Hysell\affil{1}, Margaret S. Avery\affil{1,3}}
 
@@
 \item New ca. 1092 Ma Midcontinent Rift anorthosite record high paleointensity values
-\item High VADM of the geomagnetic field during the late Mesoproterozoic is inconsistent with previous interpretation of long term decay
+\item High VADM of the geomagnetic field is inconsistent with previous interpretation of Proterozoic long term decay
 \item A strong geodynamo like today likely persisted for about 25 myr during the Midcontinent Rift development
@@
 \begin{abstract}
-New high-quality absolute paleointensity data from anorthosite xenoliths within the Beaver River diabase intrusions of the North American Midcontinent Rift (MCR) record a virtual axial dipole moment (VADM) estimate of $\sim$80 ZAm$^2$ ca.
-1092 Ma.
-This cooling rate corrected paleointensity estimate from an MCR intrusive unit is similar to previous results from extrusive units that span from ca 1108 Ma to ca.
-1085 Ma.
+New high-quality absolute paleointensity data from anorthosite xenoliths within the Beaver River diabase intrusions of the North American Midcontinent Rift (MCR) record an average virtual axial dipole moment (VADM) estimate of $\sim$80 ZAm$^2$ \textit{ca.} 1092 Ma.
+This cooling rate corrected paleointensity estimate from the MCR intrusive unit is similar to previous results from extrusive units that are dated to be \textit{ca.} 1108 Ma and \textit{ca.} 1085 Ma.
+Rock magnetic analyses and magnetic imaging reveal that the anorthosites can have abundant single-domain magnetic carriers that are ideal for paleointensity experiment, leading to a high success rate passing our selection criteria.
 Taken together, new and previously published paleointensity data from the Midcontinent Rift support that a geomagnetic field strength persisted through the Midcontinent Rift development and the data are inconsistent with the interpretation of a decay of Earth's geodynamo strength through the late Mesoproterozoic.
@@
 Due to the uncertain constraints on the thermal conductivity of the core \cite{Konopkova2016a, Gomi2013a, Ohta2016a}, numerical simulations have given drastically different predictions on the inner core nucleation age, ranging from about half a billion years ago to prior to the Archean \cite{Pozzo2012a, Gubbins2004a, Nimmo2015a}.
-Modeling has suggested that the emergence of the solid inner core could have expressed as changes of Earth's surface magnetic field strength due to the onset of compositional convection between the solid inner core and the liquid outer core \cite<e.g.>{Aubert2009a}.
-Thus, absolute paleointensity records preserved by magnetic minerals in igneous rocks may provide us with a probe into the birth of the inner core.
-Acquiring these observational data is crucial for advancing our understanding of Earth's thermal evolution and for testing existing models on the commencement of the inner core growth.
-However, due to the scarcity of well-preserved ancient rocks and low success rates typical of paleointensity experiments, only 9\% of the records in the current global paleointensity database (\href{http://earth.liv.ac.uk/pint/}{PINT, \cite{Biggin2010a}}) are from before the Cambrian and the existing records sparsely span over 3 billion years of time (Fig.
-\ref{fig:all_PINT_data}).
+Some models have suggested the thermal convection dominated liquid core motion experienced a long-term waning before the onset of the nucleation of the inner core, which kick-started the rigorous compositional convection between the solid inner core and liquid outer core \cite<e.g>{Davies2021a}.
+A corollary of such a prediction is that the geomagnetic axial dipole field, which is generated by the fluid motion in the liquid core, would have been decaying before the emergence of the solid inner core, when a significant increase in Earth's axial dipole field strength occurred \cite{Aubert2009a, Davies2021a}.
 
-\citeA{Bono2019a} filtered the available Precambrian paleointensity database (PINT) based on a set of quality criteria.
-A polynomial fit through the resulting paleointensity compilation from the Archean and the Proterozoic was used to interpret that a decrease trend of the estimated geomagnetic dipole strength occurred before the inner core started to form.
+Paleomagnetic data from ancient rocks are one of the few types of observational data that bear on the long-term evolution of Earth’s core and can be used to test the predictions from numerical simulations.
+Paleointensity studies have found very low axial dipole moment (VDM) estimates from rocks formed in the late Ediacaran \cite{Bono2019a, Shcherbakova2019a, Thallner2021b}, followed by a recovery of strength near the  Ediacaran-Cambrian boundary.
+Other studies focusing on paleomagnetic directions have also found that the late Ediacaran might have been a period of time when the geomagnetic field were experiencing frequent reversals and waning of dipole dominance.
+It has been interpreted that these findings are consistent with the results of previous geodynamo models using high core conductivity values that suggest the inner core nucleation age being near the Ediacaran-Cambrian boundary.
+\citeA{Bono2019a} filtered previous data compiled within the PINT database based on a set of quality criteria.
+A polynomial fit through the resulting paleointensity compilation from the Archean and the Proterozoic was used to interpret that a decrease trend of the estimated geomagnetic dipole strength persisted through the Proterozoic.
 In addition, \citeA{Bono2019a} recovered very low paleointensity values from ca.
-565 Ma single silicate crystals  that were interpreted to be consistent with it reflecting the weak geodynamo after progressive cooling of the core before the onset of the rigorous compositional convection driven by the nucleation of the solid inner core.
-Such decay trend is recently supported by new low paleointensity estimates derived from the ca.
+565 Ma single silicate crystals that were interpreted to be consistent with it reflecting the weak geodynamo near the Ediacaran-Cambrian boundary due to progressive cooling of the core before the onset of the rigorous compositional convection driven by the nucleation of the solid inner core.
+Such decay trend in the Neoproterozoic is recently supported by new low paleointensity estimates derived from the ca.
 720 Ma Franklin Large Igneous Province \cite{Lloyd2021a}.
 
-However, previous paleointensity data from the MCR rock units \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record VADM values distinctly higher than what is predicted by the long term decay curve.
-These rock units notably are extrusive volcanics and span a tme period of about 25 myr during the development of the Midcontinent Rift magmatism.
-Many previous results from the Midcontinent Rift show non-ideal double-slope paleointensity behavior, which are difficult to interpret as they can yield two paleointensity estimates: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
-Nevertheless, single-slope, ideal behavior from the Lake Shore Traps \cite{Kulakov2013a} lava flows passed the selection criteria of \citeA{Bono2019a}, and a few lava flows yielded single-slope behavior from the Osler Volcanics \cite{Sprain2018a}.
-These data are inconsistent with the interpretation of a low geomagnetic field strength during the Midcontinent Rift because they yield significantly higher VADM estimates than that projected by the trend from \citeA{Bono2019a}.
+However, previous paleointensity data from intrusive and extrusive rock units in the Mesoproterozoic Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record Earth's virtual axial dipole moment (VADM) values similar to that of today and are inconsistent with the prediction of the long term decay curve (Fig.
+\ref{}).
+\citeA{Pesonen1983b} obtained an average virtual dipole moment of 93 ZAm$^2$ after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using the ca.
+1108 Ma Osler Volcanics, Maimainse Point volcanics, intrusive Logan Sills, Thunder Bay dikes, and the Baraga and Marquette dikes, as well as the ca.
+1096 Ma North Shore Volcanic Group.
+\citeA{Sprain2018a} obtained more high-quality paleointensity data using the ca.
+1108 Ma Osler Volcanics and the ca.
+1100 Ma Mamainse Point volcanics.
+A calculated mean VADM value of 56 $\pm$ 21 ZAm$^2$ is similar to the average value of the past 300 million years.
+In addition, an average VADM of 59 $\pm$ 11 ZAm$^2$ obtained \citeA{Kulakov2013a} using the ca.
+1085 Ma Lake Shore Traps volcanics suggests that the present-day-like geomagnetic field strength likely persisted through the late stage of Midcontinent Rift magmatism.
+Taken together, these rock units that have been dated to span $\sim$25 myr during the development of the Midcontinent Rift magmatism \cite{Swanson-Hysell2019a} may be incompatible with the hypothesis of a ``Proterozoic dipole low" \cite{Biggin2009a}.
 
-To further enrich the Precambrian paleointensity database, and to test the hypothesis that the Earth's geomagnetic dipole strength remained at high values throughout the Midcontinent Rift, in this study we perform ``IZZI" paleointensity experiments \cite{Tauxe2004a} on a suite of anorthosite xenoliths hosted by the intrusive Beaver River diabase of the Beaver Bay Complex within the Midcontinent Rift.
+But the non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results.
+One typical failure of paleointensity experiment is the double-slope behavior, as two distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
+\citeA{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength whereas \citeA{Kulakov2013a} used the high-temperature slop.
+Such non-ideal behavior were interpreted as failure by \siteA{Sprain2018a} who applied stricter paleointensity selection criteria and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
+Similar concern on paleointensity result quality led \citeA{Bono2019a} to exclude all paleointensity records from the Midcontinent Rift but those from \citeA{Kulakov2013a}.
+Nevertheless, specimens that yield dominantly single-slope paleointensity results from the ca.
+1105 Ma Osler Volcanic Group, Mamainse Point volcanics \cite{Sprain2018a}, as well as the ca.
+1085 Ma Lake Shore Traps \cite{Kulakov2013a} suggest VADM values similar to that of the present-day values \cite{Sprain2018a}, inconsistent with the interpretation of a low geomagnetic field strength during the Midcontinent Rift (Fig.
+\ref{}).
+
+Although it is typical of paleointensity experiments to suffer from low success rates, quenched volcanic glass \cite{Tauxe2004a} and single silicate minerals \cite{Tarduno2005a} are found to be able to produce high success rates, likely due to the magnetic carriers being dominated by single-domain magnetite and being well shielded by their host against thermochemical alteration during paleointensity experiments.
+In this study, we present new paleointensity results from ``IZZI" absolute paleointensity experiments \cite{Yu2004a} on a suite of ca.
+1092 Ma anorthosite xenoliths and their diabase host rocks of the Beaver Bay Complex within the Midcontinent Rift.
+The anorthosite lithology yielded exceptionally high success rate thanks to the plagioclase crystals shielding the magnetic minerals within.
+With more single-slope, high quality paleointensity data we seek to compare paleointensity records from the Midcontinent Rift intrusives with those from the extrusives; to test the hypothesis that the Earth's geomagnetic dipole strength remained at high values throughout the Midcontinent Rift; and to further enrich the reliable paleointensity records during the $\sim$25 myr of Midcontinent Rift magmatic activity.
 
@@
 \section*{Geologic Settings}
+% a map of the lake superior including the pesonen units, the lake shore traps, the BRD, and the osler and mamainse, together with date compilation.
+
 The North American Midcontinent Rift (MCR) is a failed intracontinental rift where protracted magmatic activity lasted from \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}.
-Midcontinent Rift rocks extensively outcrop in today's Lake Superior region, with the total extent traceable by arcuate magnetic and gravity anomalies that extend to the southwest to Kansas, and to the southeast, to southern Michigan \cite{Hinze2020a}.
-Previous studies have divided magmatic activity in the rift into four stages based on interpreted changes in relative magmatic volume and the nature of magmatism: early ($\sim$1109–1104 Ma), latent ($\sim$1104–1098 Ma), main ($\sim$1098–1090 Ma) and late ($\sim$1090–1083 Ma) \cite{Vervoort2007a, Heaman2007a, Miller2013a}.
-In northeastern Minnesota, the Early Gabbro Series and the Felsic Series rocks of the Duluth Complex and reversed-polarity lavas of the lower North Shore Volcanic Group were emplaced during the early stage.
-The more voluminous Duluth Complex Layered Series and the plagioclase-rich Anorthositic Series, together with an associated $\sim$8 km thick extrusive volcanic sequences of the North Shore Volcanic Group (NSVG), were rapidly emplaced about 10 myr later at \textit{ca.} 1096 Ma during the main stage \cite{Paces1993a, Swanson-Hysell2020a}.
+Previous studies have divided magmatic activity in the rift into four stages based on interpreted changes in relative magmatic volume and the nature of magmatism: early ($\sim$1109–1104 Ma), latent ($\sim$1104–1098 Ma), main ($\sim$1098–1090 Ma) and late ($\sim$1090–1084 Ma) \cite{Vervoort2007a, Heaman2007a, Miller2013a}.
+In particular, the main stage magmatism is characterized by rapid and punctuated events which resulted in voluminous emplacement of intrusives as well as coeval outpouring of extrusives, as can be exampled by the massive ca.
+1096 Ma Duluth Complex and the North Shore Volcanic Group \cite{Swanson-Hysell2020a}, the ca.
+1092 Ma Beaver Bay Complex and the Portage Lake Volcanics (Fig.
+\ref{}; \citeA{Zhang2021a}.
+Excellent rock preservation thanks to minimal post-rift alteration in the Midcontinent Rift have yielded high-quality paleomagnetic directional data which have now been paired with high-precision zircon geochronology to reconstruct the well-resolved apparent polar wander path for Laurentia (cratonic North America) called the Keweenawan Track (Fig.
+\ref{}; \citeA{Swanson-Hysell2019a}).
 
@@
 \section*{Method}
+\subsection*{Sample collection}
+Sample cores were collected using a hand-held drill and were ori- ented using a magnetic compass as well as a sun compass when possible.
+Palaeohorizontal was determined from intercalated sedi- mentary layers and flow tops.
+6–10 cores were drilled from each lava flow in order to robustly determine an accurate site mean.
+More detailed information about the sites and their context along with directional results are reported in Swanson-Hysell et al.
+(2009, 2014a,b).
+For palaeointensity analysis, 5–8 samples per flow were chosen.
+Samples were chosen based on demagnetization behaviour with a preference for samples wherein relatively little remanence is held by haematite such that their magnetization is dominated by (titano)magnetite recording a primary thermal remanent mag- netization.
+Haematite can be a significant carrier of remanence in Mamainse Point basalts (see Swanson-Hysell et al.
+2011) and other successions around the Midcontinent Rift.
+Given that haematite can form at the expense of magmatic (titano)magnetite, such flows were deemed to not be appropriate targets for the palaeointensity experiments.
 
-\section*{Results}
+\subsection*{Paleointensity experiment}
+A total of 133 specimens from 21 selected sites underwent palaeoin- tensity experiments that followed the stepwise double-heating Thel- lier method (Thellier & Thellier 1959), using the IZZI protocol (Tauxe & Staudigel 2004).
+pTRM checks were performed systemati- cally throughout the experiment to test whether there was significant mineralogical alteration due to heating and were assessed using the SCAT parameter of Shaar & Tauxe (2013).
+All remanence measure- ments were made on a 2G Enterprises DC-SQUID superconduct- ing rock magnetometer equipped with an automated pick-and-place sample changer system at the UC Berkeley Paleomagnetism labora- tory.
+The magnetometer is housed inside a three-layer magnetostatic shield that maintains background fields of less than 500 nT.
+Heating steps were performed using an ASC TD-48SC thermal demagne- tizer with a controlled field coil that allows for a magnetic field to be generated in the oven in conjunction with a DC power sup-ply.
+The thermal demagnetizer was degaussed with an alternating field following ‘in-field’ steps such that residual fields were <10 nT during ‘zero-field’ steps.
+Samples were placed in the same location within the thermal demagnetizer for each heating step and were maintained in the same orientation with regard to the applied field.
+During each heating step, samples remained at peak temperatures for 20 min.
+An applied laboratory field of 30 μT was used for all in- field steps.
+All heating steps were performed in air.
+The temperature increments for the experiments were chosen to cover characteristic remanent magnetizations held by (titano)magnetite, with smaller increment temperature steps performed close to the expected un- blocking temperature of stoichiometric magnetite.
+Hysteresis mea- surements were conducted at the Institute for Rock Magnetism at the University of Minnesota.
+Major hysteresis loops were mea- sured at room temperature using a Micromag Princeton Measure- ments vibrating sample magnetometer with nominal sensitivity of 5 × 10−9 Am2.
 
-\section*{Discussion}
+A total of 59 specimens from 13 anorthosite sites and 46 specimens from 7 diabase sites underwent paleointensity experiments that followed the stepwise double-heating Thellier method with temperature steps up to 585 $^\circ$C \cite{Thellier1959a, Tauxe2004a}.
+On top of the IZZI modification of the Thellier experimental protocol, we also performed a comparative study where we added an extra step of 20 mT alternating field (AF) cleaning on half of the specimens after each in-field step.
+The purpose is to study whether the AF step could help improve specimens' behavior during paleointensity experiments.
+pTRM checks were performed systematically throughout the experiment to test whether there was significant mineralogical alteration due to heating and were assessed using the SCAT parameter of \citeA{Shaar2013a}.
+All remanence measurements were made on a 2G Enterprises DC-SQUID superconducting rock magnetometer equipped with an automated sample changer system in the paleomagnetism lab at UC Berkeley.
+Heating steps were performed using an ASC TD-48SC thermal demagnetizer with a controlled field coil that allows for a magnetic field to be generated in the oven in conjunction with a DC power supply.
+The thermal demagnetizer was degaussed with an alternating field following in-field steps such that residual fields were $<$10 nT during zero-field steps.
+Samples were placed in the same location within the thermal demagnetizer for each heating step and were maintained in the same orientation with regard to the applied field.
+During each heating step, samples remained at peak temperatures for 20 min.
+An applied laboratory field of 30 $\mu$T was used for all in-field steps.
+All heating steps were performed in air.
+The temperature increments for the experiments were chosen to cover characteristic remanent magnetizations held by (titano)magnetite, with smaller increment temperature steps performed close to the expected unblocking temperature of stoichiometric magnetite.
+
+\subsection*{Paleointensity result selection}
+The following criteria were used as quality filters on the palaeoin- tensity results: (1) a maximum angular deviation (MAD; Kirschvink 1980) of <20◦; (2) scatter parameter (β; Coe et al.
+1978) values of <15 per cent; (3) a deviation angle (DANG; Tanaka & Kobayashi 2003; Tauxe & Staudigel 2004) of < 5◦; (4) fraction of remanence (FRAC; Shaar & Tauxe 2013) >0.6; (5) scatter statistic (SCAT; Shaar & Tauxe 2013) = TRUE; (6) a maximum gap (GAP-Max; Shaar & Tauxe 2013) < 0.6; (7) number of pTRM checks > 2; (7) and number of measurements used for palaeointensity determi- nation ≥ 4; (Table 1)
+
+The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD; \citeA{Kirschvink1980a}) of $<$ 20; (2) scatter parameter ($\beta$; citeA{Coe1978a}) values of $<$ 15 \%; (3) a deviation angle (DANG; \citeA{Tanaka2003a}; \citeA{Tauxe2004a} of $<$ 5; (4) fraction of remanence (FRAC; \citeA{Shaar2013a}) $>$ 0.6; (5) scatter statistic (SCAT; \citeA{Shaar2013a}) = TRUE; (6) a maximum gap (GAP-Max; \citeA{Shaar2013a}) $<$ 0.6; (7) number of pTRM checks $>$ 2; (7) and number of measurements used for paleointensity determination $\geq$ 4; (Table \ref{tab:criteria}).
+The MAD measures the scatter about the best-fit line through NRM steps in the selected interval for which the intensity is defined.
+DANG, the deviation angle, is the angle between the best-fit direction that is free floating and the direction between the centre of mass of the data and the origin of the vector component diagram (\citeA{Tanaka2003a}; \citeA{Tauxe2004a}).
+Both MAD and DANG assess the directional variation of the NRM, with MAD measuring the scatter in the NRM directions and DANG assessing whether the component is trending toward the origin of the Zijderveld plot.
+$\beta$ is the ``scatter" parameter of \citeA{Coe1978a} and is the ratio of the standard error of the slope of the best-fit line of the selected NRM and pTRM points on an Arai plot to the absolute value of the slope.
+FRAC is the fraction of the NRM that is used in the best-fit line \cite{Shaar2013a}.
+The FRAC value was chosen to preferentially select samples with dominantly single-slope Arai plots.
+GAP-Max is the maximum gap between two points on the Arai plot determined by vector arithmetic.
+SCAT is a Boolean operator which uses the error on the best-fit slope of the selected data on the Arai plot to determine if the data are overly scattered.
+The parameter is used to assess pTRM checks in addition to assessing the degree to which IZZI steps are zigzagged.
+$\beta$, FRAC, GAP-Max and SCAT are all statistics to assess the behavior of Arai plots.
+See the Standard Paleointensity Definitions (\citeA{Paterson2014a}, https://earthref.org/PmagPy/SPD/home.html, last accessed 2018 March 17) for more details.
+Data analysis was conducted using Thellier GUI \cite{Shaar2013a} within the PmagPy software package \cite{Tauxe2016a}.
+
+\subsection*{Rock magnetic experiment}
+
+Hysteresis, backfield, and magnetic property measurements were conducted at the Institute for Rock Magnetism (IRM) at the University of Minnesota.
+Major hysteresis loops and backfield curves were measured at room temperature using a Micromag Princeton Measurements vibrating sample magnetometer with nominal sensitivity of $5 \times 10^{-9}$ Am\textsuperscript{2}.
+We used the magnetic property measurement system (MPMS) at the IRM to aid in the identification of magnetic mineral.
+The operational sequence follows an in-field cooling step in a 2.5 T field from room temperature to 10 K, a zero-field cooling from room temperature to 10 K, and a final cooling from room temperature to 10 K after an IRM acquisition at room temperature.
+
+Backfield demagnetization experiments were conducted and used to develop coercivity spectra (Fig.
+\ref{fig:backfield}).
+The spectra were subsequently modeled to fit for the distributions of different  populations of magnetic particles using similar procedure as \cite{Maxbauer2016a}.
+Most spectra can be well approximated with models with one component or two components with overlapping coercivity ranges (Fig.
+\ref{fig:backfield}).
+The dominance of single-component distributions from the unmixed coercivity spectra likely suggests minimal alterations or formation of secondary magnetic mineral within the samples.
+By compiling the values of median destructive field (MDF) from all measured specimens and categorizing them in terms of their sister specimens' paleointensity results, I found that the anorthosites samples that pass paleointensity selection criteria have distinctively higher MDF values than those did not (Fig.
+\ref{fig:backfield}), and all diabase specimens have low MDF values like the non-ideal anorthosites (Fig.
+\ref{fig:backfield}).
+The higher MDFs ($\sim$60 mT) of the anorthosites are similar to what is typical of stoichiometric, single-domain magnetite grains which are favored for paleointensity experiments.
+Furthermore, the low MDFs displayed in other specimens could be linked to a dominant population of large interstitial Fe-Ti oxides that are more likely to form multi-domain grains, which have been observed in both the anorthosite and diabase (Fig.
+\ref{fig:field}).
+In addition, preliminary tests on single anorthosite crystals might suggest that not all single crystals are necessarily better for paleointensity research when comparing to bulk samples, given their similar low MDF to the failed diabase and anorthosite specimens.
+However, more single crystal grains with paired paleointensity data and rock magnetic data may be needed before a general conclusion can be drawn.
+
+\section*{Results and Discussion}
+\subsection*{Paleomagnetism}
+The paleodirectional results of AF demagnetization of 7 diabase sites and thermal demagnetization of 16 anorthosite sites are shown in Table \ref{tab:}.
+In some cases, specimens from both lithologies show a minimal remanence component that was removed during very low AF field ($<$ 10 mT) for diabase and during low temperature demagnetization steps for anorthosites ($<$ 300 $^\circ$C).
+This component is likely a modern day field overprint, since its direction resembles the present-day field direction.
+The dominant components in both lithologies are northwest and down (Figure \ref{fig:directions}).
+This is likely the primary remanent magnetization recorded by the diabase and anorthosite during the emplacement of the diabase, since it is origin-trending and the principal component fitting yielded good specimen-level and site-level consistence.
+Common mean test (\cite{McFadden1990a}) results based on site-level directional comparison between anorthosites and their in situ diabase hosts are shown in Figure \ref{fig:directions}.
+Most anorthosite-diabase pairs have tightly clustered directions, and the three anorthosite sites that do no pass the statistical test with their diabase hosts nevertheless have within-site coherent directions that are close to the diabase directions (Figure \ref{fig:directions}).
+Putting the grand mean paleodirections and inferred virtual geomagnetic poles (VGPs) of both lithologies together in context of a previously synthesized apparent polar wander path (APWP) from the volcanics of the Midcontinent Rift \cite{Swanson-Hysell2019a}, we find the pole positions recovered from both the diabase and anorthosite to be lying close to the estimated VGP position around 1095 Ma, which is consistent with the geochronology constraints (between 1093.94 $\pm$ 0.28 Ma and 1091.61$\pm$ 0.14 Ma, \cite{Swanson-Hysell2019a}).
+
+Paleointensity experiment results show three types of dominant specimen behaviors on Arai plots (Figure \ref{fig:Arai_plot}): dominantly single-slope (--- out of --- specimens), double- slope (--- out of --- specimens) and zigzagging (--- out of --- specimens) (Figure \ref{fig:Arai_plot}).
+It is worth mentioning that almost all specimens have single direction demagnetization result as in Zijderveld plots of all zero-field steps.
+Upon applying the paleointensity result selection criteria (Table \ref{tab:}), we found a total of 31 out 59 anorthosite specimens and 4 out of 46 diabase specimens were accepted, yielding 7 anorthosite sites and 1 diabase sites with interpretable site-level paleointensity results (Table \ref{}).
+The selection criteria successfully rejected double-slope results (FRAC $>$ 0.6) and zigzagging results ($\beta$, DANG, and SCAT parameter).
+In addition, some rejected specimens have poor pTRM checks especially at high temperature steps (Figure \ref{fig:Arai_plot}).
+This may suggest a high degree of alteration in magnetic mineralogy during heating steps.
+The accepted anorthosite specimens do display straight Arai plots and the accepted fractions of temperature steps are mostly covering the primary remanence components (Figure \ref{fig:Arai_plot}).
+In terms of the diabase specimens, all but four of them failed the selection because double-sloping Arai plots, poor pTRM checks, and zigzagging behaviors (Figure \ref{fig:Arai_plot}).
+Even specimens in site BD2, the only site that passed selection, do not always display ideal Arai plots (Figure \ref{fig:Arai_plot}).
+Additionally, all accepted results from site BD2 are from low-temperature fractions of Arai plots (below 500 $^\circ$C), which are different than those used in anorthosite sites (typically above 500 $^\circ$).
+Although the lower temperature fraction of the Arai plots are straight and pTRM checks are consistent (Figure \ref{fig:Arai_plot}), we have to acknowledge the existence of second slopes at high temperature fractions (Figure \ref{fig:Arai_plot}).
+Nevertheless, the interpreted ancient geomagnetic field value from the accepted slopes is very similar to those reported by anorthosites (Figure \ref{fig:Arai_plot}).
+
+In this study, it is particularly important to assess the cooling rate effects, remanence anisotropy, and potential non-linear TRM or pTRM acquisition characteristics of diabase and anorthosites.
+Because the cooling duration of intrusive rocks could be much longer than the extrusive rocks, the slowly cooled diabase and anorthosite can bias laboratory paleointensity estimates toward an overestimate for SD grains \cite{Dodson1980a, Halgedahl1980a, Selkin2000a} or underestimation for MD grains \cite{McClelland-Brown1984a}.
+Further, preferred orientations of ferromagnetic fabrics along plagioclase crystallographic orientations in anorthosites could introduce a large degree of anisotropy in remanent magnetizations \cite{Selkin2007a, Biedermann2018a}.
+\citeA{Selkin2007a} reported that large, acicular SD grains can have profound non-linear acquisitions of TRM.
+According to their experiments, grains with such characteristics that are hosted within silicate crystals can saturate in fields as low as earth's field \cite{Selkin2007a}.
+
+Backfield experiment results suggest that the dominant remanence carrier in accepted anorthosite sites are likely SD magnetite (Figure \ref{fig:MDF}).
+The median destructive fields (MDF) of all accepted anorthosites are distinctively higher than those of the rejected anorthosites and the diabase.
+Moreover, the median of the MDFs of accepted anorthosites is about 50 mT, which is in the expected range of typical no-interactive SD magnetite grains.
+In addition, the anorthosites' narrow unblocking temperatures near the magnetite curie temperature also support that we idealize the remanence carriers in anorthosites to be SD magnetite.
+
+From the thermal history model we can also estimate the time duration which the diabase and anorthosite spent to block in the primary component of their remanence.
+We found the cooling time to be on the order of magnitude of $\sim$1 kyr from 580 to 500 $^\circ$C, which corresponds to a cooling rate of $1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
+In contrast, the lab cooling rate is much faster, at about $1.3\times10^{-1}$ $^\circ$C s$^{-1}$ if we assume it takes 10 minutes in lab condition to cool through the same temperature interval.
+As a result, this significant cooling rate difference leads to a prediction of 35.5\% overestimate of true field \cite{Halgedahl1980a}.
+We therefore correct our paleointensity results by a factor of 0.74.
+
+It has been well illustrated through experiments on anorthosites that anisotropic remanence acquisition can have a significant impact on paleointensity estimates \cite{Selkin2000a}.
+Following similar experiment protocol, we found that at least the bulk anorthosite specimens in this study are not very anisotropic.
+This result can be further supported by the fact that the recovered paleodirectional data from the anorthosite closely match those from the paired diabase hosts (Figure \ref{fig:directions}), the lithology of which is usually considered isotropic in terms of recording remanence magnetization.
+Both the diabase and anorthosite display consistent interpreted VGPs at the time of emplacement that matche with the expected position based on the volcanic APWP record (Figure \ref{fig:poles}, \citeA{Swanson-Hysell2019a}).
+We therefore do not correct our samples for anisotropy of remanence.
+
+We also performed non-linear full TRM checks following the experiment protocol proposed by \citeA{Selkin2007a}.
+We imparted full TRMs on our anorthosite specimens (after the paleointensity experiment) by heating them up to 600 $^\circ$C and cooling in field of 30, 50, 70, and 90 $\mu$T respectively.
+Unlike the anorthosites studied by \citeA{Selkin2007a}, none of our specimens show a saturating trend at progressively higher fields.
+More, the normalized magnetizations show nearly linear relationship with lab field strengths.
+There is thus no need for correcting non-linear thermal remanence acquisitions.
+
+Figure \ref{fig:PINT_plot} shows all accepted anorthosite and diabase specimen paleointensity results and their site-level means with one standard deviation envelopes.
+An anorthosite site mean paleointensity weighted by site standard deviations is plotted as a dashed orange line, and the standard deviation envelope as dashed green lines.
+The anorthosites are also grouped by their geographical vicinities \ref{fig:PINT_plot}.
+The site-level results are shown in Table \ref{tab:PINT_result}.
+Anorthosite sites AX10, 11, 12, 13, 16 yielded very high within-site coherence (Figure \ref{fig:Arai_plots, fig:PINT_plot}), which suggest high-confidence site-level paleointensity estimates.
+Clearly there are two populations of paleointensity results: a group of high estimates from anorthosite site 11 and 13 which have very high within-site consistency and are in agreement with the results from BRD site 2; and another group of low estimates from anorthosite site 6, 8, 10, 12, and 16, which also have tight agreement across sites.
+
+\subsection*{Rock magnetism}
+\begin{figure}[h!]
+\noindent\includegraphics[width=0.5\textwidth]{manuscript/Paleointensity_plot_cooling_corrected.pdf}
+\centering
+\caption{\small{ }}
+\label{fig:PINT_cooling_corrected}
+\end{figure}
+
+% that this study has robust data, discuss the reasoning using rock mag data that reflect the success and failure in PINT experiment
+
+% cooling rate correction problem - that the Nagy et al paper can suggest even less of a correction and higher of the estimated field.
+theoretical thermal modeling by \citeA{Zhang2021a} suggest tha the anorthosite xenoliths were likely heated to diabase magma temperature regardless of their sizes, and cooled through the typical characteristic unblocking temperature range from 580 to 500 on the order of 1000 years.
+Comparing to the typical lab cooling time of about 100 seconds duirng paleointensity experiments, the significantly slower cooling rate when the NRM was acquired as the anorthosite xenoliths cooled with the diabase led to overestimate of paleointensity by up to 40\%.
+Overall range of correction factor can be 1.1-1.4, in any case consistent with a high field strength
+
+% that this dataset also could have suffered from problem of averaging out secular variations, given that the mean pmag directions deviate from what is expected from the sythesized APWP, nevertheless the variations in estimate absolute paleointensity lead us to interpret that albeit close, the they intensity data were acquired at distinctly different times that are sufficiently far apart such that the surface geomagnetic field had already changed.
+
+% that \citeA{bono2019a} proposed an interesting trend but has problems in the directions, likely to be a short term feature of a low field during a reversal, instead of a low field as a result of long term cooling.
+
+% that Lloyd 2021 also presented low field that fits the trend, but they similarly have the issue of reversal, and that they do not have good control on secular variations
+
+% Together with previously published data, the Midcontinent Rift rocks are recording 1.1 Ga earth geomagnetic field that is stronger than the mean of the past 300 Ma (Fig. \ref{}).
 
 \section*{Conclusion}
+high success rate comparing to volcanics (more than half success sites), producing good consistent site level results, and across sites report present-day like field like today.
+
+AF pre-treatment is helpful in that it increase the specimens/sites passing the selecting criteria, suggesting that the method likely helped improve the MD ferromagnetic carriers.
+
+Rock mag and petrographic data might suggest differences in terms of the value of peak field on coercivity spectra between those passed and those didn't.
+
+IRM measured before and after an acid dunk in --- concentration acid on specimens after IZZI experiment show a difference (significant drop) maybe? between specs that passed and those did not.
+
+geographical association
+
+Anyways.
+anisotropy of full TRM experiment on accepted specs show low degree of bulk sample scale full TRM anisotropy, with a range of ---.
+Cooling rate correction based on \cite{Halgedahl1980a} suggest a lab over estimate of B ancient of 35 \%, assuming constant cooling rate in natural condition.
+The cooling rate corrected field is $\sim$40 $\mu$T which is similar to today's field strength at similar latitude.
 
 \acknowledgments
-I thank Dario Bilardello, Peat Solheid, Mike Jackson, Josh Feinberg, and Bruce Moskowitz for their tremendous help of instrumental operations, data interpretations, and research guidance.
-I thank the IRM for the generous U.S.
+Project research was supported by NSF CAREER grant EAR-1847277 to N.L.S.-H.
+and an Institute on Lake Superior Geology Student Research Fund grant to Y.Z.
+Permits for fieldwork and sampling from the Minnesota Department of Natural Resources are gratefully acknowledged.
+We thank James Pierce for assistance in the field.
+We thank Dario Bilardello, Peat Solheid, Mike Jackson, Josh Feinberg, and Bruce Moskowitz for their tremendous help of instrumental operations, data interpretations, and research guidance.
+Y.Z.
+thanks the IRM for the U.S.
 Visiting Student Fellowship.
-I thank Nicholas Swanson-Hysell at UC Berkeley for advising this research.
-Margaret Avery provided field assistance.
+Paleomagnetic data associated with this study are available within the MagIC database (\url{}; UPDATE WHEN DOI IS GENERATED) and all data are within a Github repository associated with this work (\url{}) that is also archived on Zenodo (INSERT URL AT TIME OF PROOFS).
+This repository also contains Python code related to calculations, visualizations and statistical tests discussed herein.
 
@@
 
-More Information and Advice:
+% potential reviewer: Peter Selkin, Jeff Gee, Lisa Tauxe, Roger Fu, Lauri Pesonen, Josh Feinberg Julie Bowles,

### 2021-09-13 manuscript/Manuscript.tex
draft intro

@@
 
+\begin{figure}
+\centering
+\noindent\includegraphics[width=4.75 in]{Geologic_map.pdf}
+\caption{\footnotesize{}}
+\label{fig:Geologic_map}
+\end{figure}
+
 The Beaver Bay Complex, which sits stratigraphically above the Duluth Complex, is another intrusive complex that resulted from main stage magmatism.
@@
 2(b).
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=4.75 in]{Figure/Field_photo.pdf}
-\caption{\footnotesize{Field photographs and petrographic images of the Beaver River diabase and the anorthosite xenoliths within it.
-(A) Centimeter-sized plagioclase megacrysts in the diabase.
-(B) Rounded anorthosite xenolith with a diameter of $\sim$7 meters fully enclosed within the diabase.
-(C) Exposure of a giant Carlton Peak anorthosite with a diameter $>$100 m.
-(D) 27.5 m diameter anorthosite xenolith sampled as paleomagnetic site AX16 and geochronology sample MS99033.
-(E) Cross polarized (XPL) image of the subophitic texture of diabase at site BD2 (pyroxene partially enclosing plagioclase).
-(F) XPL image of anorthosite geochronology sample MS99033.
-Plagioclase crystals exhibit both granoblastic texture and interlocking lath fabrics.
-(G) Backscattered electron (BSE) image of a large Fe-Ti oxide with titanomagnetite-ilmenite lamellae in Beaver River diabase site BD2.
-(H) BSE image of micron-sized Fe-Ti oxides exsolved from pyroxene between plagioclase crystals in anorthosite xenolith site AX4.
-an-plagioclase with $\sim$70\% anorthite; ilm-ilmenite; mag-magnetite; px-pyroxene.}}
-\label{fig:Field_photo}
-\end{figure}
 

### 2021-09-13 code/BiCEP_modules_and_scripts/BiCEP_functions.py
slight update of BiCEP notebook

@@ -518,7 +518,7 @@ class SpecimenCollection():
         ax.set_xlabel('Intensity ($\mu$T)')
         ax.set_ylabel('Probability Density')
 
-        return np.percentile(self.fit['int_site'],(2.5, 50, 97.5))
+        return np.percentile(self.fit['int_site'],(2.5, 50, 97.5)), self.fit['int_site']
 
 class Specimen():
     """

### 2021-09-28 manuscript/Manuscript.tex
manuscript update

--- 2021-08-28 manuscript/Manuscript.tex
@@
 
-\authors{Yiming Zhang\affil{1}, Nicholas L. Swanson-Hysell\affil{1}, James D. Miller Jr.\affil{2}, Margaret S. Avery\affil{1,3}}
+\authors{Yiming Zhang\affil{1}, Nicholas L. Swanson-Hysell\affil{1}, Margaret S. Avery\affil{1,3}, Roger R. Fu\affil{4}}
 
@@
 \affiliation{3}{Geology, Minerals, Energy, and Geophysics Science Center, U.S. Geological Survey, Moffett Field, CA, USA}
+\affiliation{4}{Department of Earth and Planetary Sciences, Harvard University, Cambridge, MA, USA}
 \correspondingauthor{Yiming Zhang}{yimingzhang@berkeley.edu}
@@
 \begin{keypoints}
-\item New ca. 1092 Ma Midcontinent Rift anorthosite record high paleointensity values
-\item High VADM of the geomagnetic field during the late Mesoproterozoic is inconsistent with previous interpretation of long term decay
+\item New \textit{ca.} 1092 Ma Midcontinent Rift anorthosite record high paleointensity values
+\item High VADM of the geomagnetic field is inconsistent with previous interpretation of Proterozoic long term decay
 \item A strong geodynamo like today likely persisted for about 25 myr during the Midcontinent Rift development
@@
 Due to the uncertain constraints on the thermal conductivity of the core \cite{Konopkova2016a, Gomi2013a, Ohta2016a}, numerical simulations have given drastically different predictions on the inner core nucleation age, ranging from about half a billion years ago to prior to the Archean \cite{Pozzo2012a, Gubbins2004a, Nimmo2015a}.
-Modeling has suggested that the emergence of the solid inner core could have expressed as changes of Earth's surface magnetic field strength due to the onset of compositional convection between the solid inner core and the liquid outer core \cite<e.g.>{Aubert2009a}.
-Thus, absolute paleointensity records preserved by magnetic minerals in igneous rocks may provide us with a probe into the birth of the inner core.
-Acquiring these observational data is crucial for advancing our understanding of Earth's thermal evolution and for testing existing models on the commencement of the inner core growth.
-However, due to the scarcity of well-preserved ancient rocks and low success rates typical of paleointensity experiments, only 9\% of the records in the current global paleointensity database (\href{http://earth.liv.ac.uk/pint/}{PINT, \cite{Biggin2010a}}) are from before the Cambrian and the existing records sparsely span over 3 billion years of time (Fig.
-\ref{fig:all_PINT_data}).
+Some models have suggested the thermal-convection-dominated liquid core motion experienced a long-term waning before the nucleation of the inner core, which kick-started the rigorous compositional convection between the solid inner core and liquid outer core \cite<e.g>{Davies2021a}.
+A corollary of such a prediction is that the geomagnetic axial dipole field, which is generated by the fluid motion in the liquid core, would have experienced a long-term decay before the emergence of the solid inner core, after which a significant increase in Earth's axial dipole field strength occurred \cite{Aubert2009a, Davies2021a}.
 
-\citeA{Bono2019a} filtered the available Precambrian paleointensity database (PINT) based on a set of quality criteria.
-A polynomial fit through the resulting paleointensity compilation from the Archean and the Proterozoic was used to interpret that a decrease trend of the estimated geomagnetic dipole strength occurred before the inner core started to form.
-In addition, \citeA{Bono2019a} recovered very low paleointensity values from ca.
-565 Ma single silicate crystals  that were interpreted to be consistent with it reflecting the weak geodynamo after progressive cooling of the core before the onset of the rigorous compositional convection driven by the nucleation of the solid inner core.
-Such decay trend is recently supported by new low paleointensity estimates derived from the ca.
-720 Ma Franklin Large Igneous Province \cite{Lloyd2021a}.
+Paleomagnetic data from ancient rocks are one of the few types of observational data that have the potential to record the evolution of Earth’s core and can be used to test the predictions from numerical simulations.
+Paleointensity studies have found very low axial dipole moment (VDM) estimates from rocks formed in the late Ediacaran \cite{Bono2019a, Shcherbakova2019a, Thallner2021b}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
+Other studies focusing on paleomagnetic directions have also found that the late Ediacaran might have been a period of time when the geomagnetic field were experiencing frequent reversals and waning of dipole dominance \cite{Bono2015a, Kodama2020a}.
+It has been interpreted that these findings are consistent with the results of geodynamo models using high core thermal conductivity values that suggest the inner core nucleation age being near the Ediacaran-Cambrian boundary \cite{Driscoll2016a, Davies2021a}.
+\citeA{Bono2019a} based on a polynomial fit through filtered Precambrian paleointensity database (PINT; \url{http://earth.liv.ac.uk/pint/}) to suggest the selected observational records show that the geodynamo power progressively decreased through the Proterozoic.
+They also recovered very low paleointensity values from \textit{ca.} 565 Ma single silicate crystals and interpreted them to be reflecting that the geodynamo reached its lowest near the Ediacaran-Cambrian boundary.
+Such decay trend in the Neoproterozoic is recently supported by new low paleointensity estimates derived from the \textit{ca.} 720 Ma Franklin Large Igneous Province \cite{Lloyd2021a}.
 
-However, previous paleointensity data from the MCR rock units \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record VADM values distinctly higher than what is predicted by the long term decay curve.
-These rock units notably are extrusive volcanics and span a tme period of about 25 myr during the development of the Midcontinent Rift magmatism.
-Many previous results from the Midcontinent Rift show non-ideal double-slope paleointensity behavior, which are difficult to interpret as they can yield two paleointensity estimates: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
-Nevertheless, single-slope, ideal behavior from the Lake Shore Traps \cite{Kulakov2013a} lava flows passed the selection criteria of \citeA{Bono2019a}, and a few lava flows yielded single-slope behavior from the Osler Volcanics \cite{Sprain2018a}.
-These data are inconsistent with the interpretation of a low geomagnetic field strength during the Midcontinent Rift because they yield significantly higher VADM estimates than that projected by the trend from \citeA{Bono2019a}.
+However, previous paleointensity data from intrusive and extrusive rock units in the Mesoproterozoic Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record Earth's virtual axial dipole moment (VADM) values similar to that of today, which are higher than the prediction of the long term decay curve (Fig.
+\ref{}).
+\citeA{Pesonen1983b} obtained an average virtual dipole moment of 93 ZAm$^2$ after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using the \textit{ca.} 1108 Ma Osler Volcanics, Maimainse Point volcanics, intrusive Logan Sills, Thunder Bay dikes, and the Baraga and Marquette dikes, as well as the \textit{ca.} 1096 Ma North Shore Volcanic Group.
+\citeA{Sprain2018a} obtained more high-quality paleointensity data using the \textit{ca.} 1108 Ma Osler Volcanics and the \textit{ca.} 1100 Ma Mamainse Point volcanics.
+A calculated mean VADM value of 56 $\pm$ 21 ZAm$^2$ is similar to the average value of the past 300 million years.
+In addition, an average VADM of 59 $\pm$ 11 ZAm$^2$ obtained \citeA{Kulakov2013a} using the \textit{ca.} 1085 Ma Lake Shore Traps volcanics suggests that the present-day-like geomagnetic field strength likely persisted through the late stage of Midcontinent Rift magmatism.
 
-To further enrich the Precambrian paleointensity database, and to test the hypothesis that the Earth's geomagnetic dipole strength remained at high values throughout the Midcontinent Rift, in this study we perform ``IZZI" paleointensity experiments \cite{Tauxe2004a} on a suite of anorthosite xenoliths hosted by the intrusive Beaver River diabase of the Beaver Bay Complex within the Midcontinent Rift.
+Non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results obtained from Midcontinent Rift rock units.
+One typical failure of paleointensity experiment is the double-slope behavior, as two distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
+\citeA{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength whereas \citeA{Kulakov2013a} used the high-temperature slope.
+Such non-ideal behavior were interpreted as failure by \siteA{Sprain2018a} who applied stricter paleointensity selection criteria and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
+Despite the low success rate of paleointensity experiments, specimens that yield dominantly single-slope paleointensity results from the \textit{ca.} 1105 Ma Osler Volcanic Group, Mamainse Point volcanics \cite{Sprain2018a}, as well as the \textit{ca.} 1085 Ma Lake Shore Traps \cite{Kulakov2013a} suggest VADM values similar to that of the present-day \cite{Sprain2018a}.
+Taken together, these rock units that have been dated to span $\sim$25 myr during the development of the Midcontinent Rift magmatism \cite{Swanson-Hysell2019a} may be incompatible with the hypothesis of a ``Proterozoic dipole low" \cite{Biggin2009a}.
+
+Although it is typical of paleointensity experiments to suffer from low success rates, bulk plagioclase-rich rocks \cite{Selkin2000a} and single plagioclase crystals \cite{Tarduno2005a} are found to be able to produce high quality paleointensity results.
+It has been suggested plagioclase crystals can protect magnetic minerals that exsolved within them against thermochemical alteration during paleointensity experiments.
+In this study, we present new paleointensity results from ``IZZI" absolute paleointensity experiments \cite{Yu2004a} on a suite of \textit{ca.} 1092 Ma bulk anorthosite xenoliths and their diabase host rocks of the Beaver Bay Complex within the Midcontinent Rift.
+The anorthosite lithology yielded exceptionally high success rate thanks to the plagioclase crystals shielding the magnetic minerals within.
+Thermal remanent magnetization (TRM) anisotropy test, full TRM acquisition test, together with magnetic imaging show that the bulk anorthosites from the Beaver Bay Complex have low TRM anisotropy, linear TRM acquisition property, and have dominant magnetic carriers interstitial to the plagioclase and are well shielded.
+With more single-slope, high quality paleointensity data, we seek to compare paleointensity records from the Midcontinent Rift intrusives with those from the extrusives; to test the hypothesis that the Earth's geomagnetic dipole strength remained at high values throughout the Midcontinent Rift; and to further enrich the reliable paleointensity records during the $\sim$25 myr of Midcontinent Rift magmatic activity.
 
 %Retrieving paleomagnetic information from Precambrian rocks can aid in our understanding of the long-term evolution of the geodynamo.
-A recent paleointensity study by \cite{Sprain2018a} reported a ca.
-1.1 Ga paleomagnetic field strength similar to today based on the volcanics of the Midcontinent Rift System (MRS).
+A recent paleointensity study by \cite{Sprain2018a} reported a \textit{ca.} 1.1 Ga paleomagnetic field strength similar to today based on the volcanics of the Midcontinent Rift System (MRS).
 That study suggested that this result is consistent with models wherein this time period of the Proterozoic is characterized by a strong rather than weak geomagnetic field.
@@
 \section*{Geologic Settings}
-The North American Midcontinent Rift (MCR) is a failed intracontinental rift where protracted magmatic activity lasted from \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}.
-Midcontinent Rift rocks extensively outcrop in today's Lake Superior region, with the total extent traceable by arcuate magnetic and gravity anomalies that extend to the southwest to Kansas, and to the southeast, to southern Michigan \cite{Hinze2020a}.
-Previous studies have divided magmatic activity in the rift into four stages based on interpreted changes in relative magmatic volume and the nature of magmatism: early ($\sim$1109–1104 Ma), latent ($\sim$1104–1098 Ma), main ($\sim$1098–1090 Ma) and late ($\sim$1090–1083 Ma) \cite{Vervoort2007a, Heaman2007a, Miller2013a}.
-In northeastern Minnesota, the Early Gabbro Series and the Felsic Series rocks of the Duluth Complex and reversed-polarity lavas of the lower North Shore Volcanic Group were emplaced during the early stage.
-The more voluminous Duluth Complex Layered Series and the plagioclase-rich Anorthositic Series, together with an associated $\sim$8 km thick extrusive volcanic sequences of the North Shore Volcanic Group (NSVG), were rapidly emplaced about 10 myr later at \textit{ca.} 1096 Ma during the main stage \cite{Paces1993a, Swanson-Hysell2020a}.
+% a map of the lake superior including the pesonen units, the lake shore traps, the BRD, and the osler and mamainse, together with date compilation.
 
-The Beaver Bay Complex, which sits stratigraphically above the Duluth Complex, is another intrusive complex that resulted from main stage magmatism.
-The exposed area of the Beaver Bay Complex is $\sim$1000 km\textsuperscript{2} where it has been mapped along the northwestern shore of Lake Superior in northeastern Minnesota (Fig.
-\ref{fig:Geologic_map}).
-The Beaver Bay Complex is a multi-phase, composite intrusive complex that intrudes parts of the NSVG (Fig.
-\ref{fig:Geologic_map}; \citeA{Miller1997a, Swanson-Hysell2020a}).
-Distinct from the deep plutonic intrusions of the Duluth Complex, the majority of the Beaver Bay Complex is formed of hypabyssal intrusions that were emplaced as dikes and sills at shallow depths \cite{Miller1997a}.
-Most of the Beaver Bay Complex intrusions are dioritic to gabbroic in composition \cite{Miller1997a}.
-The main lithology of the Beaver River diabase dikes and sills network within the Beaver Bay Complex is an ophitic olivine gabbro (Fig.
-\ref{fig:Field_photo}), but in wider areas of dikes and the upper parts of thick sills, this rock type can abruptly transition into intergranular olivine oxide gabbro, then into subprismatic (and commonly foliated) ferrogabbro, and finally into granophyric monzodiorite.
-The more evolved and later emplaced components of the Beaver River diabase network are commonly distinguished as the Silver Bay intrusions in the southern Beaver Bay Complex (Fig.
-\ref{fig:Geologic_map}).
-Overall being intermediate in composition, the Silver Bay intrusions lithologies range from ophitic olivine gabbro to ferrogranite \cite{Shank1989a}.
-Field mapping by \citeA{Miller1994a} found intrusive relationships between the Silver Bay intrusions and the Beaver River diabase.
-Angular inclusions of the host Beaver River diabase within marginal zones of the Silver Bay intrusions led \citeA{Miller1997a} to interpret that the Silver Bay intrusions intruded after the diabase crystallized.
+The Mesoproterozoic North American Midcontinent Rift (MCR) is an intracontinental rift that did not result in the break-up of Laurentia.
+Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; \citeA{Vervoort2007a, Miller2013a}) was punctuated with rapid and voluminous emplacement of intrusive rocks as well as extrusive volcanics.
+Based on high-precision $^{206}$Pb/$^{238}$U geochronology and paleomagnetic data, \citeA{Swanson-Hysell2021a} hypothesized that the massive Duluth Complex and the coeval North Shore Volcanic group (Figure \ref{fig:Geologic_map}) formed as a result of having upwelling mantled-derived plume material encountering Laurentia's thinned lithosphere due to rifting, following an ``upside-down” drainage" which rapidly transported melt to today's North Eastern Minnesota (Figure \ref{fig:Geologic_map}; \citeA{Sleep1997a, Swanson-Hysell2014a}).
+The plagioclase-phyric Anorthositic Series rocks of the Duluth Complex that outcrop along an arcuate shape along Northeastern Minnesota (Figure \ref{fig:Geologic_map}) are a result of such vigorous magmatic activity.
 
-One distinctive feature of the Beaver River diabase is its inclusions of anorthosite xenoliths.
-In the southern part of the Beaver Bay Complex, the Beaver River diabase occurs as dikes and sills, typically including anorthosites with various sizes ranging from centimeters to over 150 meters (Figs.
-\ref{fig:Geologic_map}, \ref{fig:Field_photo}; \citeA{Grout1939a, Morrison1983a}).
-The diabase in this region intrudes the Palisade rhyolite of the North Shore Volcanic Group (Fig.
-\ref{fig:Geologic_map}), which has a $^{206}$Pb/$^{238}$U date of 1093.94 $\pm$ 0.28 Ma (2$\sigma$ analytical uncertainty is presented for CA-ID-TIMS dates throughout this work; \citeA{Swanson-Hysell2019a}).
-The Beaver River diabase is locally intruded by the Silver Bay intrusions (Fig.
-\ref{fig:Geologic_map}).
-An aplite unit within the granophyre zone of one of these Silver Bay intrusions has a $^{206}$Pb/$^{238}$U date of 1091.61 $\pm$ 0.14 Ma \cite{Swanson-Hysell2019a}.
-Another arcuate, sill-like diabase body mapped as the Beaver River diabase outcrops along the eastern part of the complex (Fig.
-\ref{fig:Geologic_map}; \citeA{Miller1997a}).
-The diabase composition there is similar to that in the south and it also contains large anorthosite xenoliths with dimensions that exceed 100 meters at Carlton Peak (Fig.
-\ref{fig:Geologic_map}).
-The Beaver River diabase in the northern part of the complex, near the Houghtaling Creek area, typically forms narrow, near-vertical dikes instead of sheets in the southern and eastern regions (Fig.
-\ref{fig:Geologic_map}; \citeA{Miller1994a}).
-The diabase in this region only locally contains xenoliths of anorthosite.
+The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex punctuates another period of Midcontinent Rift main stage magmatism.
+$^{206}$Pb/$^{238}$U zircon dates and cross-cutting relationships tightly constraints the areally extensive Beaver River diabase of the Beaver Bay Complex to have been emplaced at 1091.7 $\pm$ 0.2 Ma \cite{Zhang2021b}.
+The Beaver River diabase intruded into hypabyssal depths via wide conduits that accommodated the entrainment of numerous anorthosite xenoliths that have short-axis diameters up to 180 meters.
+Geochronological, paleomagnetic, geochemical, and petrographic data are consistent with the hypothesis that the Beaver River diabase magma extruded to the surface, feeding the massive Greenstone Flow - one of the largest lava flows on Earth \cite{Doyle2016a, Zhang2021b}.
 
-Hundreds of anorthosite xenoliths have been recognized and mapped within the Beaver River diabase (Fig.
-\ref{fig:Geologic_map}).
-Many hill tops in the Beaver Bay Complex, such as Carlton Peak and Britton Peak, are large anorthosite blocks (which lead \citeA{Lawson1893a} to erroneously conclude that they were relict Archean topography).
-Later work established the anorthosite blocks as xenoliths, which are now extensively documented through geologic mapping of the region (Fig.
-\ref{fig:Geologic_map}; \citeA{Miller2001a, Miller1988a, Miller1989a, Boerboom2004a, Boerboom2006a, Boerboom2006b, Boerboom2007a}) and outcrop-scale exposures (Fig.
-\ref{fig:Field_photo}).
-In the field, the anorthosites typically appear as subrounded to rounded, light-colored, translucent blocks that are in sharp contact with the hosting diabase (Fig.
-\ref{fig:Field_photo}).
-They also occur as exposures whose contact with the diabase is covered (Fig.
-\ref{fig:Field_photo}).
-\citeA{Grout1939a} suggested that the rounded anorthosites are the result of abrasion during transportation as they were entrained by the diabase (i.e.
-physical weathering within a magmatic system).
-While the Beaver River diabase is chilled against the North Shore Volcanic Group lithologies that it intrudes, the diabase is not chilled against the margin of the anorthosite xenoliths \cite{Morrison1983a, Miller1997a}.
-The lack of chilled contacts is consistent with the anorthosite being at elevated temperatures and cooling at the same time as the diabase magma (Fig.
-\ref{fig:thermal_history_model}).
+% The anorthosite xenoliths that distinctly occur within the Beaver River diabase are not only recorders of the wide diabase conduit sizes,  of the geomagnetic field $\sim$1091.7 million years ago.
+Thermal modeling result and paleomagnetic directional data show the anorthosite xenoliths acquired thermal remanent magnetizations during cooling of the diabase \cite{Zhang2021b}.
+Step-wise thermal demagnetizations show many anorthosite xenoliths have minimal secondary remanence and single-component characteristic remanence magnetizations that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C.
 
-The anorthosite xenoliths are dominantly monomineralic plagioclase that has an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}.
-Interstitial pyroxene and olivine are present in minor concentrations in the xenoliths.
-Within the Carlton Peak anorthosite xenolith, up to 10 cm oikocrysts of olivine and pyroxene can occur.
-Nevertheless, the overall olivine content in the anorthosites is low.
-Interstitial titanomagnetite-ilmenite intergrowths that exceed 100 $\mu$m can be found with microscopy and $<$20 $\mu$m Fe-Ti oxide grains can be detected with scanning electron microscopy (Fig.
-\ref{fig:Field_photo}).
-Based on textural differences \citeA{Morrison1983a} divided the anorthosite xenoliths into four groups: one group which typically have well-developed granoblastic texture characterized by equigranular plagioclase crystals; another group which have interlocking, lath-shaped plagioclase crystals; an intermediate group which can have both granoblastic texture and interlocking plagioclase laths; and a brecciated group that have brittle deformation textures superposed on pre-existing textures.
+Despite that the anorthosite xenoliths of the Beaver River diabase were entrained as xenoliths and emplaced only $\sim$4 myr after the formation of the Anorthositic Series rocks of the Duluth Complex when massive plagioclase mushes were generated, $^{206}$Pb/$^{238}$U zircon geochronology data and zircon rare earth element diffusion modeling results suggest it is the most probable that the Beaver River diabase anorthosite xenoliths are entrained cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
+Moreover, the anorthosite xenoliths of the \textit{ca.} 1092 Ma Beaver River diabase has distinct mineralogy from that of the Anorthositic Series rocks of the Duluth Complex.
+The dominantly monomineralic anorthosite xenoliths have an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}, higher than the average of $\sim$60\% of the Anorthositic Series rocks of the Duluth Complex \cite{Miller1990a}.
+In addition, petrographic imaging support that the plagioclase from the Beaver River anorthosite formed under a magmatic conditions different from the Duluth Complex anorthositic series magma.
+While the plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture (Fig.
+\ref{Petro_QDM}), the more calcium-rich plagioclase of the Beaver River anorthosite often develop granoblastic texture characterized by equigranular plagioclase crystals (Fig, \ref{Petro_QDM}).
+Abundant magnetite and ilmenite needles that are typically 10 $\mu$m long can be found to have preferred orientations within Duluth Complex plagioclase.
+Vermicular titanomagnetite and ilmenite symplectic intergrowths also exist surrounding accessory olivine or pyroxene in the Duluth Complex Anorthositic Series rocks (Fig.
+\ref{Petro_QDM}).
 
-The hypabyssal intrusions of the ophitic Beaver River diabase (BRD) and the Ca-rich anorthosite xenoliths that it hosts are the targets of this study.
-The diabase contains abundant Fe-Ti oxides which are dominant carriers of paleomagnetic information.
-The nearly-pure anorthosite xenoliths have potential for paleomagnetic investigation because plagioclase hosts can protect the exsolved magnetic minerals from post-formation alteration \citep{Tarduno2005a}, and thus may faithfully record paleointensity information at the time of formation.
-A typical outcrop of anorthosite xenolith residing in the diabase and a hand sample of the anorthosite are shown in Fig.
-\ref{fig:field}(A, B).
-Preliminary petrographic analyses show that the anorthosites are dominantly monomineralic with recrystallization textures (Fig.
-\ref{fig:field}C).
-Crosscutting relationships together with high-precision geochronology dates from \cite{Swanson-Hysell2019a} bracket the age of the BRD to be between 1091.61 $\pm$ 0.14 Ma  (the age of the younger Silver Bay intrusions) and 1093.94 $\pm$ 0.28 Ma  (the age of the Palisade rhyolite).
+Similar occurrences of magnetite and ilmenite inclusions within plagioclase have been described by \citeA{Haggerty1976a} in slowly cooled, mafic intrusions, and \cite{Scofield1986a} hypothesized that the inclusions acted as a means of buffering the melt with respect to oxygen, forming either crystallites along preexisting plagioclase substrate or as iron incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures.
+Such a model requires that plagioclase, instead of other mafic minerals such as olivine, to be the first to crystallize from a magma and act as a sink for iron.
+The two-stage formation of the Duluth Complex Anorthositic Series fits this model \cite{Miller1990a}.
+During the first stage, high pressure at subcrustal depths (15-25 km depth) caused fractionation of Duluth Complex magma resulting in the formation of floating plagioclase mushes \cite{Kushiro1980a}.
+The second stage of crystallization then occurs at low pressures after the magma intruded into shallower chambers.
 
-High-precision $^{206}$Pb/$^{238}$U zircon date of 1091.83 $\pm$ 0.21 from one anorthosite in Silver Bay area tightly constraint the age of the diabase intrusion to be 1091.7 $\pm$ 0.2 Ma.
-previous paleomagnetic directional data from \citeA{Zhang2021a} reveal that vast majority of the anorthosite xenoliths have minimal secondary remanence and single component characteristic remanence magnetizations carried by low-Ti magnetite.
-Step-wise demagnetization experiments show that the characteristic remanence often unblock sharply between temperature steps 500\textdegree C and 580\textdegree C.
-Moreover, the site mean paleodirection of the anorthosite xenoliths share a common mean with the directions of their host diabase.
-Together with thermal conduction modeling, these data are consistent with the interpretation that the anorthosite xenoliths within the Beaver River diabase acquired remanence during cooling with the diabase after their hypabyssal emplacement.
-A total of  xxx specimens from xx sites were used for paleointensity experiments.
-The locations of the xx sites are shown in Fig.
-1 and the corresponding site mean palaeomagnetic directions from \citeA{Zhang2021a} are shown in Fig.
-2(b).
+This proposed magmatic evolution of the Duluth Complex is in distinct contrast to the emplacement of the Beaver River diabase which rapidly ascended and entrained the anorthosite xenoliths directly into shallow chambers \cite{Flower1980a, Miller1997a, Zhang2021b}.
+Particularly, \citeA{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
+This is consistent with the interpretation that these anorthosite were rapidly transported into shallow magma chambers without reaching structural equilibrium.
+Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Fig.
+\ref{Petro_QDM}).
+Instead, relatively sparse titano-magnetite    interstitial to the plagioclase are typical of Beaver River anorthosites (Fig, \ref{}).
+
+% Previous studies have divided magmatic activity in the rift into four stages based on interpreted changes in relative magmatic volume and the nature of magmatism: early ($\sim$1109–1104 Ma), latent ($\sim$1104–1098 Ma), main ($\sim$1098–1090 Ma) and late ($\sim$1090–1084 Ma)  In particular, the main stage magmatism is characterized by rapid and punctuated events which resulted in voluminous emplacement of intrusive rocks as well as coeval outpouring of extrusive lava flows.
+One example is the massive \textit{ca.} 1096 Ma Duluth Complex and the North Shore Volcanic Group \cite{Swanson-Hysell2020a}, the majority of which emplaced within $<$500 kyrs.
+Another one is the \textit{ca.} 1092 Ma Beaver Bay Complex.
+The Beaver River diabase intrusions of the complex formed synchronously with the Greenstone Flow of the Portage Lake Volcanics which is one of the largest lava flows known on Earth \cite{Zhang2021b}.
+Excellent rock preservation thanks to minimal post-rift alteration in the Midcontinent Rift have yielded high-quality paleomagnetic directional data which have now been paired with high-precision zircon geochronology to reconstruct the ``Keweenawan Track" - a well-resolved Laurentia's (cratonic North America) apparent polar wander path during the late Mesoproterozoic \cite{Swanson-Hysell2019a}.
 
@@
 \centering
-\noindent\includegraphics[width=4.75 in]{Figure/Field_photo.pdf}
-\caption{\footnotesize{Field photographs and petrographic images of the Beaver River diabase and the anorthosite xenoliths within it.
-(A) Centimeter-sized plagioclase megacrysts in the diabase.
-(B) Rounded anorthosite xenolith with a diameter of $\sim$7 meters fully enclosed within the diabase.
-(C) Exposure of a giant Carlton Peak anorthosite with a diameter $>$100 m.
-(D) 27.5 m diameter anorthosite xenolith sampled as paleomagnetic site AX16 and geochronology sample MS99033.
-(E) Cross polarized (XPL) image of the subophitic texture of diabase at site BD2 (pyroxene partially enclosing plagioclase).
-(F) XPL image of anorthosite geochronology sample MS99033.
-Plagioclase crystals exhibit both granoblastic texture and interlocking lath fabrics.
-(G) Backscattered electron (BSE) image of a large Fe-Ti oxide with titanomagnetite-ilmenite lamellae in Beaver River diabase site BD2.
-(H) BSE image of micron-sized Fe-Ti oxides exsolved from pyroxene between plagioclase crystals in anorthosite xenolith site AX4.
-an-plagioclase with $\sim$70\% anorthite; ilm-ilmenite; mag-magnetite; px-pyroxene.}}
-\label{fig:Field_photo}
+\noindent\includegraphics[width=\textwidth]{Geologic_map.pdf}
+\caption{\footnotesize{Simplified geologic map of Lake Superior region showing the distribution of rift and post-rift-related rocks including extrusive, intrusive, and late/post-rift sediments around Lake Superior.
+Red diamonds mark sites with paleointensity results that passed the selection criteria from this study.
+Paleomagnetic sites from \citeA{Pesonen1983a} are categorized by lithology.
+All sites from previous studies are color-coded by their ages.}}
+\label{fig:Geologic_map}
 \end{figure}
 
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{Field_photo_petro_photo.pdf}
-\caption{\small{(A) A typical outcrop of an anorthosite xenolith residing in the Beaver River diabase.
-(B) A hand sample of anorthosite.
-The large reflective face at the bottom center is a cleavage plane of a centimeter-size plagioclase crystal.
-The scale is 9 cm in total.
-(C) Cross-polarized light image of an anorthosite thin section.
-The second order birefringence of plagioclase indicates a high Ca content.
-The closely packed plagioclase crystals show recrystallization texture.
-(D) High-magnification SEM image of Fe-Ti oxides exsolved from a pyroxene crystal included in a plagioclase crystal.
-(E) Large, interstitial magnetite-ilmenite intergrowths between two plagioclase crystals.
-(F) A large Fe-Ti oxide grain with magnetite-ilmenite intergrowths next to a pyroxene in diabase.
-an: anorthite; px: pyroxene; mag: magnetite; ilm: ilmenite}}
-\label{fig:field}
+% \begin{figure}
+% \noindent\includegraphics[width=\textwidth]{}
+% \caption{\small{}
+% \label{fig:Petro_QDM}
+% \end{figure}
+
+\section*{Method}
+
+\subsection*{Sample collection and paleomagnetic directions}
+
+We collected paleomagnetic cores that are 2.5 cm in diameter along the southern and eastern Beaver Bay Complex with a particular focus on acquiring paired sites of anorthosite xenoliths and their local diabase hosts during summer field seasons in 2019 and 2020.
+Sample cores were collected using a hand-held gasoline-powered drill and were oriented using a magnetic compass as well as a sun compass when possible.
+Sun compass orientations were preferentially used for determining the sample azimuth.
+Specimens from a total of 7 diabase and 14 anorthosite sites were used for paleointensity experiments (Table \ref{tab:Pmag_site_data}).
+Sister specimens underwent step-wise alternating field (AF) or thermal demagnetization at the UC Berkeley Paleomagnetism Lab and the results are presented in \citeA{Zhang2021b}.
+Paleomagnetic direction measurement level data are available within the MagIC database (\url{https://earthref.org/MagIC/doi/10.1029/2021GC009909}).
+The close directional similarities between the anorthosite xenolith and their host diabase, the overall cooling unit mean directions between the two lithologies, and thermal modeling results are consistent with the interpretation that the Beaver River diabase and anorthosite xenoliths acquired primary thermal remanent magnetization \textit{ca.} 1092 Ma during cooling after the emplacement of the diabase \cite{Zhang2021b}.
+Based on the anorthosite thermal demagnetization results we selected sites whose unblocking temperature ranges are narrow and near 580\textdegree C.
+Beaver River diabase sites with minimal secondary remanence component were also selected for paleointensity experiments.
+
+\subsection*{Paleointensity experiment}
+A total of 69 specimens from 7 diabase sites and a total of 86 specimens from 14 anorthosite xenoliths underwent paleointensity experiments that followed the step-wise double-heating Thellier method \cite{Thellier1959a} using the IZZI protocol \cite{Tauxe2004a} with heating steps up to 585 \textdegree C.
+Partial thermal remanent magnetization (pTRM) checks were performed systematically throughout the experiment to test whether there was significant mineralogical alteration due to heating and were assessed using the SCAT parameter of \citeA{Shaar2013a}.
+On top of the IZZI modification of the Thellier experimental protocol, we also performed a comparative study where we added an extra step of 20 mT alternating field (AF) cleaning on some of the specimens after each in-field step.
+The purpose is to study whether the AF cleaning could help improve experiment success rate by removing the remanence component carried by materials such as multi-domain (MD) or vortex state grains that often contributes to non-ideal paleointensity behaviors.
+All remanence measurements were made on a 2G Enterprises DC-SQUID superconducting rock magnetometer equipped with an automated sample changer system at the UC Berkeley Paleomagnetism laboratory.
+The magnetometer is housed inside a three-layer magnetostatic shield that maintains background fields of less than 500 nT.
+Heating steps were performed using an ASC TD-48SC thermal demagnetizer with a controlled field coil that allows for a magnetic field to be generated in the oven in conjunction with a DC power supply.
+The thermal demagnetizer was degaussed with an alternating field in the axial orientation following each in-field step such that residual fields within the oven were $<$10 nT during zero-field steps.
+Samples were placed in the same location within the thermal demagnetizer for each heating step and were maintained in the same orientation with regard to the applied field.
+During each heating step, samples remained at peak temperatures for 20 min.
+An applied laboratory field of 30 μT was used for all in-field steps.
+All heating steps were performed in air.
+The temperature increments for the experiments were chosen to cover characteristic remanent magnetizations held by (titano)magnetite informed by the previous demagnetization data, with smaller increment temperature steps performed close to the expected unblocking temperature of stoichiometric magnetite \cite{Zhang2021b}.
+
+After the paleointensity experiment were finished, the same specimens underwent additional TRM anisotropy check with full TRMs given at 0, 30, 60, 90, 120, 150 degrees in the horizontal plane in specimen coordinates.
+After that, full TRM acquisition experiments were performed on the same specimens with applied fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axis.
+
+\subsection*{Paleointensity result selection}
+The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD; \citeA{Kirschvink1980a}) of <20\textdegree; (2) scatter parameter ($\beta$; \citeA{Coe1978a}) values of $<$15 per cent; (3) a deviation angle (DANG; \citeA{Tanaka2003a, Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC; \citeA{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT; \citeA{Shaar2013a}) = TRUE; (6) a maximum gap (GAP-Max; \citeA{Shaar2013a}) < 0.6; (7) number of pTRM checks > 2; (8) and number of measurements used for paleointensity determination $≥$ 4.
+The criteria are also summarized in Table \ref{tab:criteria}.
+The MAD measures the scatter about the best-fit line through NRM steps in the selected interval for which the intensity is defined.
+DANG, the deviation angle, is the angle between the best-fit direction that is free floating and the direction between the centre of mass of the data and the origin of the vector component diagram (\citeA{Tanaka2003a}; \citeA{Tauxe2004a}).
+Both MAD and DANG assess the directional variation of the NRM, with MAD measuring the scatter in the NRM directions and DANG assessing whether the component is trending toward the origin of the Zijderveld plot.
+$\beta$ is the ``scatter" parameter of \citeA{Coe1978a} and is the ratio of the standard error of the slope of the best-fit line of the selected NRM and pTRM points on an Arai plot to the absolute value of the slope.
+FRAC is the fraction of the NRM that is used in the best-fit line \cite{Shaar2013a}.
+The FRAC value was chosen to preferentially select samples with dominantly single-slope Arai plots.
+GAP-Max is the maximum gap between two points on the Arai plot determined by vector arithmetic.
+SCAT is a Boolean operator which uses the error on the best-fit slope of the selected data on the Arai plot to determine if the data are overly scattered.
+The parameter is used to assess pTRM checks in addition to assessing the degree to which IZZI steps are zigzagged.
+$\beta$, FRAC, GAP-Max and SCAT are all statistics to assess the behavior of Arai plots.
+See the Standard Paleointensity Definitions (\citeA{Paterson2014a}, \url{https://earthref.org/PmagPy/SPD/home.html}) for more details.
+Data analysis was conducted using Thellier GUI \cite{Shaar2013a} within the PmagPy software package \cite{Tauxe2016a}.
+
+\begin{table}
+\caption{Quality criteria used for accepting or rejecting paleointensity data. See the text for more detail.} 
+\centering
+\begin{tabular}{ccccccccc}
+\hline
+MAD (o) & Beta (per cent) & DANG (o) & FRAC& SCAT & GAP-MAX & NpTRM & N Arai\\
+\hline
+20 & 15 & 5 & 0.6 & TRUE & 0.6 & 2 & 4 \\
+\hline
+
+\end{tabular}
+\label{tab:criteria}
+\end{table}
+
+\subsection*{Rock magnetic experiment}
+
+In this study, we conduct rock magnetic experiments with a purpose of gaining magnetic mineralogy insight into the paleointensity results of the anorthosite and diabase.
+Backfield curves were measured at room temperature using a Micromag Princeton Measurements vibrating sample magnetometer (VSM) with nominal sensitivity of $5 \times 10^{-9}$ Am\textsuperscript{2}, and a Lake Shore 8600 series VSM with nominal sensitivity of xxxxxxxxxx.
+The calculated coercivity spectra were subsequently decomposed into one ore more components using skew-normal distributions following the method of \citeA{Maxbauer2016a}.
+The fitted components are then used to interpret the characteristics of different populations of magnetic particles within specimens.
+We also used the magnetic property measurement system (MPMS) at the IRM to aid in the identification of magnetic mineral.
+The operational sequence follows an in-field cooling step in a 2.5 T field from room temperature to 10 K, a zero-field cooling from room temperature to 10 K, and a final cooling from room temperature to 10 K after an IRM acquisition at room temperature.
+The magnetic moment transitions at critical temperatures revealed through MPMS experiments are used to identify magnetic minerals such as magnetite \cite{Verwey1939a} within specimens.
+% it would actually be nice to show the MPMS data (although it could be a bit redundant).
+The MPMS data nicely shows the BD and failed AX have the Verwey transition suppressed to lower temperatures and the successful AX have Verwey transition close to 120 K - the typical value for stochiometric magnetite.
+I don't think I will be using the hysteresis loops.
+
+\section*{Results and Discussion}
+\subsection*{Paleointensity}
+
+40 from a total of 86 anorthosite specimens and 9 out of a total of 69 diabase specimens passed our paleointensity selection criteria, yielding 7 anorthosite sites and 1 diabase sites with site-level paleointensity results that pass paleointensity selections.
+Example Arai plots are shown in Figure \ref{fig:IZZI_examples}.
+Summary specimen absolute paleointensity estimate and site-level weighted mean paleointensity values are shown in Figure \ref{fig:PINT_cooling_corrected}.
+All measurement-level paleointensity experiment data can be accessed through the MagIC database (\url{}; UPDATE WHEN DOI IS GENERATED).
+
+The Beaver River anorthosite specimens yielded a high pass rate ($\sim$46.5\%) through our selection criteria.
+In distinct contrast, the Beaver Rider diabase specimens were almost all rejected by our selection criteria.
+Only site BD2 has 9 specimens that passed selection ($\sim$13.0\% pass rate).
+Paleointensity experiment results from both lithologies show three types of dominant specimen behaviors on Arai plots (Figure \ref{fig:IZZI_examples}; \citeA{Arai1963a}): dominantly single-slope, double-slope, and zigzagging or sagging (Figure \ref{fig:IZZI_examples}).
+It is worth mentioning that almost all specimens have minimal secondary remanence component in thermal demagnetization results as shown in orthogonal plots of all zero-field steps (Figure \ref{fig:IZZI_examples}).
+Typical specimen paleointensity behavior of the anorthosites display straight, single-slope Arai plots and the accepted fractions of temperature steps span over the origin-trending, primary remanence components (Figure \ref{fig:IZZI_examples}).
+We accept all specimen- and site-level absolute paleointensity estimates from the anorthosite xenoliths.
+In terms of the diabase specimens, all but nine of them failed the selection because of double-slope behavior (fail FRAC selection), poor pTRM checks, and zigzagging (fail SCAT, DANG selection) Arai plots (Figure \ref{fig:IZZI_examples}).
+Even specimens in site BD2, the only site that passed selection, do not always display ideal Arai plots.
+As is shown in Figure \ref{fig:IZZI_examples}, specimen BD2-5b has a high-temperature second slope with zigzagging behavior, poor pTRM checks, and low NRM/TRM ratio, in addition to the straight slope with steep NRM/TRM ratio over the lower temperature range.
+Other accepted specimen results from site BD2 are from similar low-temperature fractions of Arai plots (below 500 $^\circ$C), which are different than those used in anorthosite sites (typically above 500 $^\circ$).
+Although the BD2 specimens' lower temperature fraction of the Arai plots are straight, the pTRM checks are consistent, and the specimen paleointensity results show within-site consistency and yield a weighted mean paleointensity estimate indistinguishable with that of anorthosite AX11 and AX13 (Figure \ref{fig:IZZI_examples}, \ref{fig:PINT_cooling_corrected}), specimens from BD2 give consistently higher paleointensity estimates than those from site AX6, AX8, and AX10, which are anorthosite xenoliths included within BD2 (Figure \ref{fig:Geologic_map}; \citeA{Zhang2021a}).
+This higher estimate from the diabase lithology could be attributed to the systematic bias of using the lower temperature slope of the Arai plots demonstrated in Figure \ref{fig:IZZI_examples}.
+
+Alternating field treatment after in-field heating steps likely improved the paleointensity behavior of diabase specimens such that they pass the selection criteria but the treatment did not result in significant success rate improvement or changes in paleointensity estimates for the anorthosites (Figure \ref{fig:PINT_cooling_corrected}).
+8 out of 13 specimens from site BD2 that were subjected to AF treatment have straight low temperature slopes that pass the selection criteria, whereas only 1 from 4 specimens that were not subjected to AF treatment passes.
+On the other hand, 27 out of 58 ($\sim$46.6\%) anorthosite specimens that had AF treatment passed selection and 13 out of 28 ($\sim$46.4\%) specimens without AF treatment passed selection.
+Paleointensity estimates from specimens with or without AF treatment in anorthosite AX12, AX13, AX16 do not show significant difference (Figure \ref{fig:PINT_cooling_corrected}).
+Specimen plots for AX11 particularly illustrates a case where all specimens within the site yield almost the same paleointensity estimate, regardless of the AF treatment.
+Such consistent behavior makes the results from this site of very high confidence.
+
+The different effects that alternating field treatment had upon anorthosite and diabase may reflect the magnetic mineralogy difference between lithologies.
+Diabase likely has significant populations of MD or vortex state grains which can have relatively low coercivities and result in non-ideal paleointensity behaviors, but whose remanence contribution could be significant to the whole rock natural remanent magnetization.
+Upon the application of the 20 mT peak AF treatment after each in-field heating step, the remanence acquired by these grains is likely significantly removed, therefore making a straight slope of NRM/TRM during temperatures below 500\textdegree C (Figure \ref{fig:backfield}).
+On the other hand, the anorthosite specimens with and without AF treatment have very similar paleointensity behaviors and consistent specimen level estimates.
+Together with the narrow range of unblocking temperatures of their sister specimens \cite{Zhang2021b}, these data are consistent with the interpretation that the anorthosites with successful paleointensity results have a magnetic mineral assemblage that likely contain dominantly single domain, low Ti magnetite and they do not contain a significant population of MD or vortext state grains.
+Some anorthosite specimens and many diabase specimens still failed in result selections even though AF treatment was applied.
+It could be that they have a very high dominance of MD or vortex state grains with relatively moderate coercivity such that an alternating field with 20 mT peak field did not significantly mitigate their non-ideal paleointensity behaviors.
+Another cause for their failure could be of thermochemical alterations that happened during heating in air.
+As can be exampled by the zigzagging high temperature slope with poor pTRM checks in paleointensity plot for specimen BD2-5b (Figure \ref{fig:IZZI_examples}), the decrease in NRM/TRM ratios during higher temperature steps likely resulted from the formation of new ferromagnetic material within the specimen due to oxidation.
+Such alteration can allow specimens to acquire additional remanence during in-field steps that may not be fully removed by an AF treatment.
+
+\subsection*{Rock magnetism}
+Rock magnetic data support the interpretation that the successful anorthosite specimens have dominant magnetic carriers with magnetic properties similar to stochiometric, non-interactive, single domain magnetite, whereas those anorthosites that failed the paleointensity result selection and all diabase have more pronounced non-ideal carriers.
+MPMS data show sister specimens to paleointensity specimens typically have dominant magnetic mineralogy of magnetite as can be evidenced through the Verwey transitions (Figure \ref{fig:MPMS}; \citeA{Verwey1939a}).
+However, specimens from sites that yield successful paleointensity results display Verwey transition near 120 Kelvin (K) as expected for stochiometric low-Ti magnetite (\cite{Ozdemir1993a}).
+On the other hand, the diabase specimens and anorthosite specimens that did not yielded accepted paleointensity results typically have suppressed Verwey transition toward lower temperatures, indicating that the titanomagnetite within the specimens likely experienced variable degrees of oxidation \cite{Ozdemir1993a}.
+
+Specimen coercivity spectra further support that our paleointensity selection criteria are successful in selecting experiment results.
+A compilation of median destructive field (MDF) in Figure \ref{fig:coercivity} based on all of our backfield experiments shows that successful anorthosites have distinctly higher average MDF range than others.
+Example single component skew-normal distribution fits of coercivity spectra based on backfield experiments on specimen chips sister to paleointensity specimens show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity distributions around 80 mT,
+very similar to that of a bulk sample composed of no-interactive single domain magnetite (Figure \ref{fig:coercivity}).
+In contrast, specimens from anorthosite that did not pass paleointensity selection, as well as diabase specimens, tend to have coercivity distributions over a significantly lower range ($\sim$30 mT).
+This is consistent with the interpretation that they have a more dominant population of high Ti content, multi-domain grains or vortex state grains.
+
+\begin{figure}[h!]
+\noindent\includegraphics[width=\textwidth]{MPMS.pdf}
+\centering
+\caption{\footnotesize{}}
+\label{fig:MPMS}
 \end{figure}
 
-Most thermally demagnetized anorthosites and alternating field (AF) demagnetized diabase specimens have minimal secondary components and their magnetization is interpreted to be dominated by a primary thermal remanent magnetization.
-Characteristic magnetizations from both lithologies yield indistinguishable site-mean directions and the virtual geomagnetic poles (VGPs) fall close to the 1095 Ma pole from the synthesized APWP of \cite{Swanson-Hysell2019a}, which is in agreement with the geochronological constraints  (Fig.
-\ref{fig:pmag}).
+\begin{figure}[h!]
+\noindent\includegraphics[width=\textwidth]{coercivity.pdf}
+\centering
+\caption{\footnotesize{Top: Example coercivity unmixing plots on anorthosites and diabase specimens from sites that pass or fail our paleointensity selection criteria.
+Bottom: Box plots summarize distributions of median destructive field (MDF) values for all anorthosite and diabase specimens that have single-component coercivity unmixing results.
+Both plots show that anorthosite specimens that pass our paleointensity selection criteria have distinctly higher MDF than other anorthosites that do not pass the selection, as well as all diabase specimens.
+The diabase site BD2 which is the only site that yielded specimen results that pass the selection, have similar MDF than other diabase sites.}}
+\label{fig:coercivity}
+\end{figure}
 
-I conducted a comparative study of modified IZZI paleointensity experiments on both the diabase and anorthosite, with a group treated with low-field AF demagnetization after each in-field step and another group without.
+\subsection*{Remanence anisotropy and nonlinear TRM acquisition check}
+Significant remanence anisotropy has been documented to exist within anorthositic rocks that form in layered intrusive rocks \cite{Selkin2000a, Feinberg2006a}.
+\cite{Selkin2000a} demonstrated that strong remanence anisotropy associated with the igneous foliation developed within an anorthosite sample from the Stillwater Complex can lead to paleointensity estimates that are different by three times.
+Depending on the orientation of the applied field during in-field steps and that of the magnetic anisotropy axes of a specimen, such strong preferred orientations of petrofabrics can result in significantly biased paleointensity estimates.
+To assess whether our paleointensity estimates are biased by remanence anisotropy, we first calculated the gamma statistic, which is the angular difference between the last pTRM step used for paleointensity determination and the applied field direction.
+The results show all anorthosite specimens used for paleointensity experiment have gamma values ranging from $\sim$0.9\textdegree-$\sim$11.9\textdegree, with a median value of 4.2\textdegree (Table \ref{}).
+These gamma values are similar to those of the Midcontinent Rift volcanics \cite{Sprain2018a} and they show that the Beaver River anorthosite bulk rocks do not have a significant directional remanence anisotropy.
+This result can be further supported by the fact that the recovered paleodirectional data from our anorthosite xenoliths closely match those from the paired Beaver River diabase hosts \cite{Zhang2021b}, the lithology of which is usually considered isotropic in terms of recording remanence magnetization.
+Further, we applied full TRMs to the anorthosite specimens that pass the paleointensity selection criteria after paleointensity experiment.
+With an applied field of 30 $\mu$T oriented toward the specimen at 0, 30, 60, 90, 120, and 150 degrees, the anorthosite specimens usually acquired consistent full TRMs.
+The most significant variations observed in acquired full TRMs reach to only XXX \%, much less than that has been reported to occurred within layerd anorthositic rocks in Stillwater (up to $\sim$60 \%, \citeA{Selkin2000a}).
+
+\cite{Selkin2007a} also noted that vortex state magnetic grains and certain single domain particles can produce biased paleointensity estimates due to their nonlinearity behavior of thermal remanence acquisition.
+We applied full TRMs to our anorthosite specimens with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axis.
+The results show that our anorthosite specimens did not saturate upon an applied field of 30 $\mu$T and there is thus no need for correcting nonlinear thermal remanence acquisitions.
+
+The reason that the Beaver River anorthosite xenoliths do not display strong remanence anisotropy, nor do they display nonlinear TRM acquisition problem, could be that they formed in a magmatic setting distinct from that of a layered mafic complex, such as the nearby Duluth Complex, where igneous foliation can occur (Figure \ref{fig:Geologic_map}).
+Petrographic images show that an equigranular texture where equidimensional, randomly oriented plagioclase often develop in the anorthosite xenoliths (Figure \ref{fig:Petro_QDM}; \citeA{Zhang2021b}).
+This is in contrast with the interlocking igneous foliation texture within the Anorthositic Series rocks of the \textit{ca.} 1096 Ma Duluth Complex (Figure \ref{fig:Petro_QDM}).
+Further, magnetic imaging also shows the Beaver River anorthosite xenoliths do not develop abundant magnetite and ilmenite needle exsolution within plagioclase crystals as the slowly-cooled anorthositic intrusions, in which exsolved magnetite needles can be the dominate remanence carriers (Figure \ref{fig:Petro_QDM}; \citeA{Scofield1986a, Feinberg2006a}).
+In addition, the aligned magnetic maps and the petrographic images in Figure \ref{fig:Petro_QDM} nicely illustrates that the magnetic dipole fields are exerted by the magnetite needles.
+The fact that all dipole features within a same plagioclase crystal have indistinguishable orientations is consistent with the interpretation of there being a magnetic remanence anisotropy resulting from preferred petrofabric alignment, giving rise to the remanence anisotropy and nonlinear TRM acquisition of layered anorthositic intrusions (Figure \ref{fig:Petro_QDM}).
+
+\subsection*{Cooling rate correction}
+Paleointensity estimates from igneous intrusions can also be biased toward higher values due to slow cooling rates \cite{Halgedahl1980a}.
+Because the cooling rates of intrusive rocks could be much longer than the extrusive rocks, the slowly cooled diabase and anorthosite can bias laboratory paleointensity estimates toward an overestimate for SD grains \cite{Dodson1980a, Halgedahl1980a, Selkin2000a}.
+
+From the thermal history model we can also estimate the time duration which the diabase and anorthosite spent to block in the primary component of their remanence.
+We found the cooling time to be on the order of magnitude of $\sim$1 kyr from 580 to 500 $^\circ$C, which corresponds to a cooling rate of $1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
+In contrast, the lab cooling rate is much faster, at about $1.3\times10^{-1}$ $^\circ$C s$^{-1}$ if we assume it takes 10 minutes in lab condition to cool through the same temperature interval.
+As a result, this significant cooling rate difference leads to a prediction of 35.5\% overestimate of true field \cite{Halgedahl1980a}.
+We therefore correct our paleointensity results by a factor of 0.74.
+
+\subsection*{Late Mesoproterozoic geomagnetic field and the Precambrian observation records}
+Figure \ref{fig:PINT_plot} shows all accepted anorthosite and diabase specimen paleointensity results and their site-level means with one standard deviation envelopes.
+An anorthosite site mean paleointensity weighted by site standard deviations is plotted as a dashed orange line, and the standard deviation envelope as dashed green lines.
+The anorthosites are also grouped by their geographical vicinities \ref{fig:PINT_plot}.
+The site-level results are shown in Table \ref{tab:PINT_result}.
+Anorthosite sites AX10, 11, 12, 13, 16 yielded very high within-site coherence (Figure \ref{fig:Arai_plots, fig:PINT_plot}), which suggest high-confidence site-level paleointensity estimates.
+Clearly there are two populations of paleointensity results: a group of high estimates from anorthosite site 11 and 13 which have very high within-site consistency and are in agreement with the results from BRD site 2; and another group of low estimates from anorthosite site 6, 8, 10, 12, and 16, which also have tight agreement across sites.
+
+All units have Q$_{PI}$ scores \cite{Biggin2014a} ranging between 6 and 7 (Table \ref{tab:QPI}).
+All sites passed AGE, TRM and ALT (pTRM checks passed and high-temperature steps that looked to be suffering from alteration were excluded), ACN and MD (no sig- nificant sagging or zigzagging).
+
+\begin{figure}[h!]
+\noindent\includegraphics[width=\textwidth]{IZZI_examples.pdf}
+\centering
+\caption{\footnotesize{Example results of paleointensity experiments displayed on Arai and Zijderveld plots (insets) for anorthosite and diabase specimens.
+Red (blue) circles indicate zero-field/in-field (in-field/zero-field) steps `ZI’ (`IZ’).
+Triangles mark partial thermal remanent (pTRM) checks.
+Blue and red squares in the Zijderveld plots are X–Y and X–Z projections, respectively, of the NRMs in specimen coordinates.
+Plots on the top row show successful specimen paleointensity results with straight, single-slope behaviors that pass our selection criteria.
+The green lines represent our fitting for the dominant single-slope component that passes our acceptance criteria with estimates of the ancient field (B$_anc$) being used to determine site means.
+The plots for anorthosite specimens AX1-1a and diabase BD5-1b on the bottom row show typical non-ideal sagging and zigzagging behaviour that fails our acceptance criteria.
+Plot for diabase specimen BD2-5a shows typical paleointensity behavior of this site - specimens typically have double-slope behaviors although the lower temperature slopes can pass the selection criteria, resulting in the high estimated B$_anc$ values shown in Figure \ref{fig:PINT_cooling_corrected}.
+Thus, the estimates of B$_anc$ for this specimen plot is illustrative, but were not used to calculate a site mean paleointensity value in Figure \ref{fig:PINT_cooling_corrected}.}}
+\label{fig:IZZI_examples}
+\end{figure}
+
+% need to export the figure with tight boundaries
+\begin{figure}[h!]
+\noindent\includegraphics[width=\textwidth]{Paleointensity_plot_cooling_corrected.pdf}
+\centering
+\caption{\small{ }}
+\label{fig:PINT_cooling_corrected}
+\end{figure}
+
+In addition, \cite{Cych2021a} developed a Bayesian probability method to estimate paleointensity - independent of the selection criteria, the Bayesian method fits for the curvature and generate probability distribution that guides credible interval for the paleoitnensity at a site level.
+The result distribution of Markov Chain Monte Carlo resampling are shown as violin plots in Figure \ref{}.
+The cooling rate corrected distributions show similar results as the frequentists' estimates based on the selection criteria derived results.
+
+Address the spacial groups and using direction and intensity to address the issue of secular variation.
+
+overall the PINT results are of high quality and the Qpi index average to be 6 to 7.
+
+The calculated paleomagnetic VADM values are shown in Figure \ref{} in context of the compilation of all Midcontinent Rift results.
+highlight the high quality of the results and the consistency of MCR rocks' high geomagnetic field values even considering the uncertainties of previous data.
+Supporting that the Mesoproterozoic likely experienced a period of persistent high geomagnetic strength in addition to the protracted magmatism of in the formation of the massive intracontinental rifting.
+
+% cooling rate correction problem - that the Nagy et al paper can suggest even less of a correction and higher of the estimated field.
+theoretical thermal modeling by \citeA{Zhang2021a} suggest tha the anorthosite xenoliths were likely heated to diabase magma temperature regardless of their sizes, and cooled through the typical characteristic unblocking temperature range from 580 to 500 on the order of 1000 years.
+Comparing to the typical lab cooling time of about 100 seconds duirng paleointensity experiments, the significantly slower cooling rate when the NRM was acquired as the anorthosite xenoliths cooled with the diabase led to overestimate of paleointensity by up to 40\%.
+Overall range of correction factor can be 1.1-1.4, in any case consistent with a high field strength
+
+% that this dataset also could have suffered from problem of averaging out secular variations, given that the mean pmag directions deviate from what is expected from the sythesized APWP, nevertheless the variations in estimate absolute paleointensity lead us to interpret that albeit close, the they intensity data were acquired at distinctly different times that are sufficiently far apart such that the surface geomagnetic field had already changed.
+
+% that \citeA{bono2019a} proposed an interesting trend but has problems in the directions, likely to be a short term feature of a low field during a reversal, instead of a low field as a result of long term cooling.
+
+% that Lloyd 2021 also presented low field that fits the trend, but they similarly have the issue of reversal, and that they do not have good control on secular variations
+
+% Together with previously published data, the Midcontinent Rift rocks are recording 1.1 Ga earth geomagnetic field that is stronger than the mean of the past 300 Ma (Fig. \ref{}).
+
+\section*{Conclusion}
+high success rate comparing to volcanics (more than half success sites), producing good consistent site level results, and across sites report present-day like field like today.
+
+AF pre-treatment is helpful in that it increase the specimens/sites passing the selecting criteria, suggesting that the method likely helped improve the MD ferromagnetic carriers.
+
+Rock mag and petrographic data might suggest differences in terms of the value of peak field on coercivity spectra between those passed and those didn't.
+
+IRM measured before and after an acid dunk in --- concentration acid on specimens after IZZI experiment show a difference (significant drop) maybe? between specs that passed and those did not.
+
+geographical association
+
+Anyways.
+anisotropy of full TRM experiment on accepted specs show low degree of bulk sample scale full TRM anisotropy, with a range of ---.
+Cooling rate correction based on \cite{Halgedahl1980a} suggest a lab over estimate of B ancient of 35 \%, assuming constant cooling rate in natural condition.
+The cooling rate corrected field is $\sim$40 $\mu$T which is similar to today's field strength at similar latitude.
+
+\acknowledgments
+Project research was supported by NSF CAREER grant EAR-1847277 to N.L.S.-H.
+and an Institute on Lake Superior Geology Student Research Fund grant to Y.Z.
+Permits for fieldwork and sampling from the Minnesota Department of Natural Resources are gratefully acknowledged.
+We thank James Pierce for assistance in the field.
+
+We thank Jim Miller for providing us with the Duluth Complex anorthosite thin section.
+
+We thank Dario Bilardello, Peat Solheid, Mike Jackson, Josh Feinberg, and Bruce Moskowitz for their tremendous help of instrumental operations, data interpretations, and research guidance.
+Y.Z.
+thanks the IRM for the U.S.
+Visiting Student Fellowship.
+Paleomagnetic data associated with this study are available within the MagIC database (\url{}; UPDATE WHEN DOI IS GENERATED) and all data are within a Github repository associated with this work (\url{}) that is also archived on Zenodo (INSERT URL AT TIME OF PROOFS).
+This repository also contains Python code related to calculations, visualizations and statistical tests discussed herein.
+
+\bibliography{YZ_ref}
+
+\end{document}
+
+% potential reviewer: Peter Selkin, Jeff Gee, Lisa Tauxe, Roger Fu, Lauri Pesonen, Josh Feinberg Julie Bowles,
+
+% I conducted a comparative study of modified IZZI paleointensity experiments on both the diabase and anorthosite, with a group treated with low-field AF demagnetization after each in-field step and another group without.
 Paleointensity experiments for most diabase specimens result in non-ideal Arai plots often with poor pTRM checks, making them difficult to interpret for paleointensity estimate (Fig.
@@
 
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{pmag_plot_.pdf}
-\caption{\small{(A) Site-mean directions of anorthosite and diabase plotted on an equal-area plot.
-(B) Calculated VGPs from the anorthosite and diabase plotted in context of a previously synthesized ca.
-1.1 Ga Laurentia APWP from the Midcontinent Rift volcanics \citep{Swanson-Hysell2019a}.
+% \begin{figure}
+% \noindent\includegraphics[width=\textwidth]{pmag_plot_.pdf}
+% \caption{\small{(A) Site-mean directions of anorthosite and diabase plotted on an equal-area plot.
+(B) Calculated VGPs from the anorthosite and diabase plotted in context of a previously synthesized \textit{ca.} 1.1 Ga Laurentia APWP from the Midcontinent Rift volcanics \citep{Swanson-Hysell2019a}.
 VGPs from both lithologies fall close to the 1095 Ma pole position.
@@
 The estimated field intensity is not cooling rate corrected.}}
-\label{fig:pmag}
-\end{figure}
+% \label{fig:pmag}
+% \end{figure}
 
-At the IRM, I was seeking the answer to this question with the help of the vibrating sample magnetometer (VSM) systems, including a newly installed Lake Shore VSM which greatly helped improve measurement resolution on samples with weak magnetizations (Fig.
+% At the IRM, I was seeking the answer to this question with the help of the vibrating sample magnetometer (VSM) systems, including a newly installed Lake Shore VSM which greatly helped improve measurement resolution on samples with weak magnetizations (Fig.
 \ref{fig:backfield}).
@@
 
-Taken together, the rock magnetism experiment results support my paleointensity selection criteria in that rock specimens that produce straight Arai plots and have MDF similar to stoichiometric single-domain magnetite are preferentially selected.
+% Taken together, the rock magnetism experiment results support my paleointensity selection criteria in that rock specimens that produce straight Arai plots and have MDF similar to stoichiometric single-domain magnetite are preferentially selected.
 The preliminary paleointensity results yielded consistent estimates.
@@
 
-\begin{figure}
-\begin{center}
-	\noindent\includegraphics[width=0.5\textwidth]{backfield_example_.pdf}
-\end{center}
-\caption{\small{(A) Backfield measurement data acquired using a Princeton VSM on an anorthosite specimen.
+% \begin{figure}
+% \begin{center}
+% 	\noindent\includegraphics[width=0.5\textwidth]{backfield_example_.pdf}
+% \end{center}
+% \caption{\small{(A) Backfield measurement data acquired using a Princeton VSM on an anorthosite specimen.
 The noise level on coercivity plot is high and a heavy smoothing is needed to make subsequent coercivity modeling interpretable.
@@
 The anorthosite specimens that pass selection criteria have distinctly higher median destructive field (MDF) than other groups.}}
-\label{fig:backfield}
-\end{figure}
-
-\section*{Method}
-
-\section*{Results}
-
-\section*{Discussion}
-
-\section*{Conclusion}
-
-\acknowledgments
-I thank Dario Bilardello, Peat Solheid, Mike Jackson, Josh Feinberg, and Bruce Moskowitz for their tremendous help of instrumental operations, data interpretations, and research guidance.
-I thank the IRM for the generous U.S.
-Visiting Student Fellowship.
-I thank Nicholas Swanson-Hysell at UC Berkeley for advising this research.
-Margaret Avery provided field assistance.
-
-\bibliography{YZ_ref}
-
-\end{document}
-
-More Information and Advice:
+% \label{fig:backfield}
+% \end{figure}

### 2021-09-29 manuscript/Manuscript.tex

@@
 \usepackage{soul}
+
 \linenumbers
@@
 
+\begin{figure}
+\centering
+\noindent\includegraphics[width=\textwidth]{Geologic_map.pdf}
+\caption{\footnotesize{Simplified geologic map of Lake Superior region showing the distribution of rift and post-rift-related rocks including extrusive, intrusive, and late/post-rift sediments around Lake Superior.
+Red diamonds mark sites with paleointensity results that passed the selection criteria from this study.
+Paleomagnetic sites from \citeA{Pesonen1983a} are categorized by lithology.
+All sites from previous studies are color-coded by their ages.}}
+\label{fig:Geologic_map}
+\end{figure}
+
 The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex punctuates another period of Midcontinent Rift main stage magmatism.
@@
 
+\begin{figure}
+\noindent\includegraphics[width=\textwidth]{Petro_QDM.png}
+\caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from Duluth Complex Anorthositic Series (left column) and anorthosite xenolith MS99033 from the Beaver River diabase (right column).
+Petrographic images (A-D) show Duluth Complex anorthosite have euhedral, elongate plagioclase crystals with interlocking texture and strong igneous foliation.
+The plagioclase can contain abundant Fe-Ti oxide needles (C) that have preferred orientations that are often parallel with the (010) planes of plagioclase.
+In contrast, plagioclase of the anorthosite xenolith MS99033 do not enclose oxide needles.
+Magnetic field maps acquired by the quantum diamond microscope (QDM) show that the oxide needles within Duluth plagioclase are strongly magnetic as evidenced through the overlapping magnetic dipole fields exerted above the oxides in (E).
+In comparison, the plagioclase of MS99033 is not magnetic.
+}}
+\label{fig:Petro_QDM}
+\end{figure}
+
 Similar occurrences of magnetite and ilmenite inclusions within plagioclase have been described by \citeA{Haggerty1976a} in slowly cooled, mafic intrusions, and \cite{Scofield1986a} hypothesized that the inclusions acted as a means of buffering the melt with respect to oxygen, forming either crystallites along preexisting plagioclase substrate or as iron incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures.
@@
 \ref{Petro_QDM}).
-Instead, relatively sparse titano-magnetite    interstitial to the plagioclase are typical of Beaver River anorthosites (Fig, \ref{}).
+Instead, relatively sparse titano-magnetite cinterstitial to the plagioclase are typical of Beaver River anorthosites (Fig, \ref{}).
 
@@
 Excellent rock preservation thanks to minimal post-rift alteration in the Midcontinent Rift have yielded high-quality paleomagnetic directional data which have now been paired with high-precision zircon geochronology to reconstruct the ``Keweenawan Track" - a well-resolved Laurentia's (cratonic North America) apparent polar wander path during the late Mesoproterozoic \cite{Swanson-Hysell2019a}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{Geologic_map.pdf}
-\caption{\footnotesize{Simplified geologic map of Lake Superior region showing the distribution of rift and post-rift-related rocks including extrusive, intrusive, and late/post-rift sediments around Lake Superior.
-Red diamonds mark sites with paleointensity results that passed the selection criteria from this study.
-Paleomagnetic sites from \citeA{Pesonen1983a} are categorized by lithology.
-All sites from previous studies are color-coded by their ages.}}
-\label{fig:Geologic_map}
-\end{figure}
-
-% \begin{figure}
-% \noindent\includegraphics[width=\textwidth]{}
-% \caption{\small{}
-% \label{fig:Petro_QDM}
-% \end{figure}
 
@@
 We also used the magnetic property measurement system (MPMS) at the IRM to aid in the identification of magnetic mineral.
-The operational sequence follows an in-field cooling step in a 2.5 T field from room temperature to 10 K, a zero-field cooling from room temperature to 10 K, and a final cooling from room temperature to 10 K after an IRM acquisition at room temperature.
+In the field-cooled (FC) experiments shown in Figure \ref{fig:MPMS}, the magnetization was measured upon warming following the specimen having cooled in an applied field of 2.5 T from 300 to 10 K.
+In the zero-field-cooled (ZFC) experiment, a low-temperature saturation isothermal remanence (LTSIRM) of 2.5 T was applied at 10 K after the specimen cooled in a (near-)zero field.
+In the room-temperature saturation isothermal remanence (RTSIRM) experiment, the sample was pulsed with a 2.5 T field at room temperature ($\sim$300 K) and then cooled to 10 K and warmed back to room temperature in a (near)zero field.
 The magnetic moment transitions at critical temperatures revealed through MPMS experiments are used to identify magnetic minerals such as magnetite \cite{Verwey1939a} within specimens.
@@
 All measurement-level paleointensity experiment data can be accessed through the MagIC database (\url{}; UPDATE WHEN DOI IS GENERATED).
+
+\begin{figure}[h!]
+\noindent\includegraphics[width=\textwidth]{IZZI_examples.pdf}
+\centering
+\caption{\footnotesize{Example results of paleointensity experiments displayed on Arai and Zijderveld plots (insets) for anorthosite and diabase specimens.
+Red (blue) circles indicate zero-field/in-field (in-field/zero-field) steps `ZI’ (`IZ’).
+Triangles mark partial thermal remanent (pTRM) checks.
+Blue and red squares in the Zijderveld plots are X–Y and X–Z projections, respectively, of the NRMs in specimen coordinates.
+Plots on the top row show successful specimen paleointensity results with straight, single-slope behaviors that pass our selection criteria.
+The green lines represent our fitting for the dominant single-slope component that passes our acceptance criteria with estimates of the ancient field (B$_anc$) being used to determine site means.
+The plots for anorthosite specimens AX1-1a and diabase BD5-1b on the bottom row show typical non-ideal sagging and zigzagging behaviour that fails our acceptance criteria.
+Plot for diabase specimen BD2-5a shows typical paleointensity behavior of this site - specimens typically have double-slope behaviors although the lower temperature slopes can pass the selection criteria, resulting in the high estimated B$_anc$ values shown in Figure \ref{fig:PINT_cooling_corrected}.
+Thus, the estimates of B$_anc$ for this specimen plot is illustrative, but were not used to calculate a site mean paleointensity value in Figure \ref{fig:PINT_cooling_corrected}.}}
+\label{fig:IZZI_examples}
+\end{figure}
 
@@
 
+\begin{table}[]
+\caption{\footnotesize{Summary statistics for the Q$_{PI}$ quality criteria of \citeA{Biggin2014a}.}}
+\centering
+\begin{tabular}{ccccccccccccc}
+\hline
+Site & N  & Age (Ma) & Method & AGE & STAT & TRM & ALT & MD & ACN & TECH & LITH & QPI \\ \hline
+AX6  & 3  & 1091.8   & T+     & 1   & 0    & 1   & 1   & 1  & 1   & 0    & 0    & 5   \\
+AX8  & 3  & 1091.8   & T+     & 1   & 0    & 1   & 1   & 1  & 1   & 0    & 0    & 5   \\
+AX10 & 3  & 1091.8   & T+     & 1   & 0    & 1   & 1   & 1  & 1   & 0    & 0    & 5   \\
+AX12 & 6  & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\
+AX16 & 11 & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\
+AX11 & 7  & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\
+AX13 & 7  & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\ \hline
+BD2  & 9  & 1091.7   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\ \hline
+\end{tabular}
+\label{tab:QPI}
+\end{table}
+
 \subsection*{Rock magnetism}
@@
 However, specimens from sites that yield successful paleointensity results display Verwey transition near 120 Kelvin (K) as expected for stochiometric low-Ti magnetite (\cite{Ozdemir1993a}).
-On the other hand, the diabase specimens and anorthosite specimens that did not yielded accepted paleointensity results typically have suppressed Verwey transition toward lower temperatures, indicating that the titanomagnetite within the specimens likely experienced variable degrees of oxidation \cite{Ozdemir1993a}.
-
-Specimen coercivity spectra further support that our paleointensity selection criteria are successful in selecting experiment results.
-A compilation of median destructive field (MDF) in Figure \ref{fig:coercivity} based on all of our backfield experiments shows that successful anorthosites have distinctly higher average MDF range than others.
-Example single component skew-normal distribution fits of coercivity spectra based on backfield experiments on specimen chips sister to paleointensity specimens show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity distributions around 80 mT,
-very similar to that of a bulk sample composed of no-interactive single domain magnetite (Figure \ref{fig:coercivity}).
-In contrast, specimens from anorthosite that did not pass paleointensity selection, as well as diabase specimens, tend to have coercivity distributions over a significantly lower range ($\sim$30 mT).
-This is consistent with the interpretation that they have a more dominant population of high Ti content, multi-domain grains or vortex state grains.
+On the other hand, the diabase specimens and anorthosite specimens that did not yielded accepted paleointensity results typically have suppressed Verwey transition toward lower temperatures (Figure \ref{fig:MPMS}), indicating that the specimens likely have a ferromagnetic mineralogy with relatively high Ti content and/or with variable degrees of oxidation \cite{Ozdemir1993a}.
 
@@
 \centering
-\caption{\footnotesize{}}
+\caption{\footnotesize{Low-temperature magnetic property measurement system (MPMS) experiment results.
+In the field-cooled (FC) experiments, the magnetization was measured upon warming following the specimen having cooled in an applied field of 2.5 T from 300 to 10 K.
+In the zero-field-cooled (ZFC) experiment, a low-temperature saturation isothermal remanence (LTSIRM) of 2.5 T was applied at 10 K after the specimen cooled in a (near-)zero field.
+In the room-temperature saturation isothermal remanence (RTSIRM) experiment, the sample was pulsed with a 2.5 T field at room temperature ($\sim$300 K) and then cooled to 10 K and warmed back to room temperature in a (near)zero field.
+Specimen AX6-1r is from anorthosite AX6 which passed our paleointensity selection.
+It has a well-defined Verwey transition $\sim$120 K.
+Specimens AX3-2r-ni and BD2-2r have prominent Verwey transition but the transition temperatures are suppressed below $\sim$120 K.
+}}
 \label{fig:MPMS}
 \end{figure}
+
+Specimen coercivity spectra further support that our paleointensity selection criteria are successful in selecting experiment results.
+A compilation of median destructive field (MDF) in Figure \ref{fig:coercivity} based on all of our backfield experiments shows that successful anorthosites have distinctly higher average MDF range than others.
+Example single component skew-normal distribution fits of coercivity spectra based on backfield experiments on specimen chips sister to paleointensity specimens show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity distributions around 80 mT, very similar to that of a bulk sample composed of no-interactive single domain magnetite (Figure \ref{fig:coercivity}).
+In contrast, specimens from anorthosite that did not pass paleointensity selection, as well as diabase specimens, tend to have coercivity distributions over a significantly lower range ($\sim$30 mT).
+This is consistent with the interpretation that they lack a dominance of single domain magnetite grains in their ferromagnetic mineral assemblages.
 
@@
 
-From the thermal history model we can also estimate the time duration which the diabase and anorthosite spent to block in the primary component of their remanence.
-We found the cooling time to be on the order of magnitude of $\sim$1 kyr from 580 to 500 $^\circ$C, which corresponds to a cooling rate of $1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
+From the thermal history model we can also estimate the duration which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully block in their characteristic natural remanence component ($\sim$500\textdegree C; \citeA{Zhang2021b}).
+We find the cooling time to be on the order of $\sim$1 kyr, which corresponds to a cooling rate of $1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
 In contrast, the lab cooling rate is much faster, at about $1.3\times10^{-1}$ $^\circ$C s$^{-1}$ if we assume it takes 10 minutes in lab condition to cool through the same temperature interval.
-As a result, this significant cooling rate difference leads to a prediction of 35.5\% overestimate of true field \cite{Halgedahl1980a}.
+As a result, this significant cooling rate difference leads to a prediction of 35.5\% overestimate of true ancient field \cite{Halgedahl1980a}.
 We therefore correct our paleointensity results by a factor of 0.74.
+The cooling rate corrected specimen paleointensity estimates together with specimen- and site-level weighted means are shown in Figure \ref{fig:PINT_cooling_corrected}.
+Taking all accepted anorthosite paleointensity results, they have a site mean absolute paleointensity estimates of 47.90 $\pm$ 9.32 $\mu$T (Figure \ref{fig:PINT_cooling_corrected}).
 
@@
 An anorthosite site mean paleointensity weighted by site standard deviations is plotted as a dashed orange line, and the standard deviation envelope as dashed green lines.
-The anorthosites are also grouped by their geographical vicinities \ref{fig:PINT_plot}.
+The anorthosites are also grouped by their geographical vicinity \ref{fig:PINT_plot}.
 The site-level results are shown in Table \ref{tab:PINT_result}.
@@
 
-All units have Q$_{PI}$ scores \cite{Biggin2014a} ranging between 6 and 7 (Table \ref{tab:QPI}).
-All sites passed AGE, TRM and ALT (pTRM checks passed and high-temperature steps that looked to be suffering from alteration were excluded), ACN and MD (no sig- nificant sagging or zigzagging).
-
-\begin{figure}[h!]
-\noindent\includegraphics[width=\textwidth]{IZZI_examples.pdf}
-\centering
-\caption{\footnotesize{Example results of paleointensity experiments displayed on Arai and Zijderveld plots (insets) for anorthosite and diabase specimens.
-Red (blue) circles indicate zero-field/in-field (in-field/zero-field) steps `ZI’ (`IZ’).
-Triangles mark partial thermal remanent (pTRM) checks.
-Blue and red squares in the Zijderveld plots are X–Y and X–Z projections, respectively, of the NRMs in specimen coordinates.
-Plots on the top row show successful specimen paleointensity results with straight, single-slope behaviors that pass our selection criteria.
-The green lines represent our fitting for the dominant single-slope component that passes our acceptance criteria with estimates of the ancient field (B$_anc$) being used to determine site means.
-The plots for anorthosite specimens AX1-1a and diabase BD5-1b on the bottom row show typical non-ideal sagging and zigzagging behaviour that fails our acceptance criteria.
-Plot for diabase specimen BD2-5a shows typical paleointensity behavior of this site - specimens typically have double-slope behaviors although the lower temperature slopes can pass the selection criteria, resulting in the high estimated B$_anc$ values shown in Figure \ref{fig:PINT_cooling_corrected}.
-Thus, the estimates of B$_anc$ for this specimen plot is illustrative, but were not used to calculate a site mean paleointensity value in Figure \ref{fig:PINT_cooling_corrected}.}}
-\label{fig:IZZI_examples}
-\end{figure}
+All units have Q$_{PI}$ scores \cite{Biggin2014a} ranging between 5 and 6 (Table \ref{tab:QPI}).
+All sites passed AGE, TRM and ALT (pTRM checks passed and high-temperature steps that looked to be suffering from alteration were excluded), ACN and MD (no significant sagging or zigzagging).
 
@@
 \centering
-\caption{\small{ }}
+\caption{\small{Summary plot of cooling rate corrected specimen absolute paleointensity results (square and diamond symbols) from this study and their weighted averages at site (grey bar) and locality level (orange bar).
+}}
 \label{fig:PINT_cooling_corrected}
 \end{figure}
+
+considering \citeA{Nagy2021a} developed a new cooling rate correction model for uniaxial single domain magnetite and suggest that the model of \citeA{Halgedahl1980a} tend to over-correct for cooling rates and lead to underestimating ancient fields.
+If our absolute paleointensity values are an underestimating,
 

### 2021-09-29 manuscript/Manuscript.tex

@@
 \noindent\includegraphics[width=\textwidth]{Petro_QDM.png}
-\caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from Duluth Complex Anorthositic Series (left column) and anorthosite xenolith MS99033 from the Beaver River diabase (right column).
+\caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from Duluth Complex Anorthositic Series (left column) and Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (right column; \citeA{Zhang2021b}).
 Petrographic images (A-D) show Duluth Complex anorthosite have euhedral, elongate plagioclase crystals with interlocking texture and strong igneous foliation.
 The plagioclase can contain abundant Fe-Ti oxide needles (C) that have preferred orientations that are often parallel with the (010) planes of plagioclase.
-In contrast, plagioclase of the anorthosite xenolith MS99033 do not enclose oxide needles.
+In contrast, plagioclase of sample MS99033 do not enclose oxide needles.
 Magnetic field maps acquired by the quantum diamond microscope (QDM) show that the oxide needles within Duluth plagioclase are strongly magnetic as evidenced through the overlapping magnetic dipole fields exerted above the oxides in (E).
@@
 \ref{Petro_QDM}).
-Instead, relatively sparse titano-magnetite cinterstitial to the plagioclase are typical of Beaver River anorthosites (Fig, \ref{}).
+Instead, relatively sparse titano-magnetite interstitial to the plagioclase are typical of Beaver River anorthosites (Fig, \ref{}).
 
@@
 
+% need to export the figure with tight boundaries
+\begin{figure}[h!]
+\noindent\includegraphics[width=\textwidth]{Paleointensity_plot_cooling_corrected.pdf}
+\centering
+\caption{\small{Summary plot of cooling rate corrected specimen absolute paleointensity results (square and diamond symbols) from this study and their weighted averages at site (grey bar) and locality level (orange bar).
+}}
+\label{fig:PINT_cooling_corrected}
+\end{figure}
+
 Alternating field treatment after in-field heating steps likely improved the paleointensity behavior of diabase specimens such that they pass the selection criteria but the treatment did not result in significant success rate improvement or changes in paleointensity estimates for the anorthosites (Figure \ref{fig:PINT_cooling_corrected}).
@@
 Clearly there are two populations of paleointensity results: a group of high estimates from anorthosite site 11 and 13 which have very high within-site consistency and are in agreement with the results from BRD site 2; and another group of low estimates from anorthosite site 6, 8, 10, 12, and 16, which also have tight agreement across sites.
-
-All units have Q$_{PI}$ scores \cite{Biggin2014a} ranging between 5 and 6 (Table \ref{tab:QPI}).
-All sites passed AGE, TRM and ALT (pTRM checks passed and high-temperature steps that looked to be suffering from alteration were excluded), ACN and MD (no significant sagging or zigzagging).
-
-% need to export the figure with tight boundaries
-\begin{figure}[h!]
-\noindent\includegraphics[width=\textwidth]{Paleointensity_plot_cooling_corrected.pdf}
-\centering
-\caption{\small{Summary plot of cooling rate corrected specimen absolute paleointensity results (square and diamond symbols) from this study and their weighted averages at site (grey bar) and locality level (orange bar).
-}}
-\label{fig:PINT_cooling_corrected}
-\end{figure}
 

### 2021-09-29 manuscript/Manuscript.tex

@@
 \centering
-\caption{\small{Summary plot of cooling rate corrected specimen absolute paleointensity results (square and diamond symbols) from this study and their weighted averages at site (grey bar) and locality level (orange bar).
+\caption{\small{Summary plot of cooling rate corrected individual specimen absolute paleointensity results (square and diamond symbols) from this study and their averages at site (grey bar) and locality level (weighted mean by site-level standard deviations; orange bar).
+The inset plot shows the graph of predicted paleointensity overestimate due to slow cooling of the intrusive Beaver River diabase relative to the cooling rate in laboratory following the model of \citeA{Halgedahl1980a}.
+Because the majority of the anorthosites have unblocking temperatures between 500\textdegree C and 580\textdegree C, we estimate that the slow cooling during natural remanence acquisition could have resulted in a maximum of 36\% overestimate.
+Thus a correction factor of 0.74 is applied at specimen level for the summary plot.
 }}
@@
 We find the cooling time to be on the order of $\sim$1 kyr, which corresponds to a cooling rate of $1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
-In contrast, the lab cooling rate is much faster, at about $1.3\times10^{-1}$ $^\circ$C s$^{-1}$ if we assume it takes 10 minutes in lab condition to cool through the same temperature interval.
-As a result, this significant cooling rate difference leads to a prediction of 35.5\% overestimate of true ancient field \cite{Halgedahl1980a}.
+In contrast, the lab cooling rate is much faster, at about $1.3\times10^{-1}$ $^\circ$C s$^{-1}$ if we assume it takes 1000s in lab condition to cool through the same temperature interval.
+As a result, this significant cooling rate difference leads to a prediction of $\sim$36\% overestimate of true ancient field \cite{Halgedahl1980a}.
 We therefore correct our paleointensity results by a factor of 0.74.
@@
 \subsection*{Late Mesoproterozoic geomagnetic field and the Precambrian observation records}
-Figure \ref{fig:PINT_plot} shows all accepted anorthosite and diabase specimen paleointensity results and their site-level means with one standard deviation envelopes.
-An anorthosite site mean paleointensity weighted by site standard deviations is plotted as a dashed orange line, and the standard deviation envelope as dashed green lines.
 The anorthosites are also grouped by their geographical vicinity \ref{fig:PINT_plot}.
@@
 Clearly there are two populations of paleointensity results: a group of high estimates from anorthosite site 11 and 13 which have very high within-site consistency and are in agreement with the results from BRD site 2; and another group of low estimates from anorthosite site 6, 8, 10, 12, and 16, which also have tight agreement across sites.
+
+\begin{figure}[h!]
+\noindent\includegraphics[width=\textwidth]{PINT_BiCEP.pdf}
+\centering
+\caption{\footnotesize{Violin plot of paleointensity estimates using the bias corrected estimation of paleointensity (BiCEP) method \cite{Cych2021a}. }}
+\label{fig:BiCEP}
+\end{figure}
+
+\begin{figure}[h!]
+\noindent\includegraphics[width=\textwidth]{MCR_VADM.pdf}
+\centering
+\caption{\footnotesize{Compilation of calculated virtual axial dipole moment (VADM) values from previous studies of the Midcontinent Rift rocks \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} and from the anorthosites that pass the selection criteria in this study.
+The median VADM value of all MCR rocks is higher than the median value of the past 300 myr.
+}}
+\label{fig:MCR_VADM}
+\end{figure}
 

### 2021-09-30 code/BiCEP_modules_and_scripts/BiCEP_functions.py
manuscript progress

@@ -518,7 +518,7 @@ class SpecimenCollection():
         ax.set_xlabel('Intensity ($\mu$T)')
         ax.set_ylabel('Probability Density')
 
-        return np.percentile(self.fit['int_site'],(2.5, 50, 97.5)), self.fit['int_site']
+        return np.percentile(self.fit['int_site'],(2.5, 50, 97.5)), np.array(self.fit['int_site'])
 
 class Specimen():
     """

### 2021-09-30 manuscript/Manuscript.tex

@@
 \begin{abstract}
-New high-quality absolute paleointensity data from anorthosite xenoliths within the Beaver River diabase intrusions of the North American Midcontinent Rift (MCR) record an average virtual axial dipole moment (VADM) estimate of $\sim$80 ZAm$^2$ \textit{ca.} 1092 Ma.
-This cooling rate corrected paleointensity estimate from the MCR intrusive unit is similar to previous results from extrusive units that are dated to be \textit{ca.} 1108 Ma and \textit{ca.} 1085 Ma.
-Rock magnetic analyses and magnetic imaging reveal that the anorthosites can have abundant single-domain magnetic carriers that are ideal for paleointensity experiment, leading to a high success rate passing our selection criteria.
-Taken together, new and previously published paleointensity data from the Midcontinent Rift support that a geomagnetic field strength persisted through the Midcontinent Rift development and the data are inconsistent with the interpretation of a decay of Earth's geodynamo strength through the late Mesoproterozoic.
+New high-quality absolute paleointensity data from \textit{ca.} 1092 Ma anorthosite xenoliths within the Beaver River diabase intrusions of the North American Midcontinent Rift record an average virtual axial dipole moment (VADM) estimate of $\sim$80 ZAm$^2$.
+This cooling rate corrected paleointensity estimate from Midcontinent Rift intrusions is similar to previous results from extrusive units that are dated to be between \textit{ca.} 1108 Ma and \textit{ca.} 1085 Ma.
+Rock magnetic analyses and magnetic imaging reveal that the anorthosite xenoliths can often have magnetic mineralogy dominated by grains with high coercivities, which can yield high success rates passing our selection criteria relative to their diabase hosts.
+Taken together, new and previously published paleointensity data from the Midcontinent Rift support that a strong geomagnetic field  persisted through 25 million years of the late Mesoproterozoic.
+These data are inconsistent with interpretations of continuous progressive decay of Earth's geodynamo strength through the Proterozoic Eon.
 
@@
 %The question of the core and the importance of the observational data as complementary for numerical simulation work.
+
+The strong geodynamo today is collectively powered by thermal convection and compositional convection in the liquid outer core.
+Studies have found that Earth may have sustained a geodynamo through the Archean \cite{Tarduno2014a}, and possibly as far back as the Hadean \cite{Tarduno2020a}.
+However, recent estimates of the thermal conductivity of the core suggest that the nucleation of the inner core nucleation could be $<$500 Ma, more recent than previous thought.
+Withouth the power of compositional convection provided by light elements being rejected into the outer core from the inner core, any identification of an earlier geodynamo would indicate that the core was stirred by other power sources, such as precipitation of Mg \cite{ORourcke2016a}, or perhaps
+that the dynamo was generated by exotic processes like a convecting basal magma ocean \cite{Ziegler2013a}.
+In any case, a geodynamo solely driven by thermal conduction through the core-mantle boundary (CMB) prior to the nucleation of the inner core is likely to decay persistently due to heat loss.
 
@@
 One typical failure of paleointensity experiment is the double-slope behavior, as two distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
-\citeA{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength whereas \citeA{Kulakov2013a} used the high-temperature slope.
-Such non-ideal behavior were interpreted as failure by \siteA{Sprain2018a} who applied stricter paleointensity selection criteria and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
-Despite the low success rate of paleointensity experiments, specimens that yield dominantly single-slope paleointensity results from the \textit{ca.} 1105 Ma Osler Volcanic Group, Mamainse Point volcanics \cite{Sprain2018a}, as well as the \textit{ca.} 1085 Ma Lake Shore Traps \cite{Kulakov2013a} suggest VADM values similar to that of the present-day \cite{Sprain2018a}.
+\citeA{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas \citeA{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
+Results with such non-ideal behavior were rejected by \siteA{Sprain2018a} who applied stricter paleointensity selection criteria and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
+Despite the low success rate of paleointensity experiments, specimens that yield dominantly single-slope paleointensity results from the \textit{ca.} 1105 Ma flows of the Osler Volcanic Group, ca.
+1095 Ma flows of the Mamainse Point volcanics \cite{Sprain2018a}, as well as the \textit{ca.} 1085 Ma flows of the Lake Shore Traps \cite{Kulakov2013a} suggest VADM values similar to that of the present-day geomagnetic field \cite{Sprain2018a}.
 Taken together, these rock units that have been dated to span $\sim$25 myr during the development of the Midcontinent Rift magmatism \cite{Swanson-Hysell2019a} may be incompatible with the hypothesis of a ``Proterozoic dipole low" \cite{Biggin2009a}.
@@
 It has been suggested plagioclase crystals can protect magnetic minerals that exsolved within them against thermochemical alteration during paleointensity experiments.
-In this study, we present new paleointensity results from ``IZZI" absolute paleointensity experiments \cite{Yu2004a} on a suite of \textit{ca.} 1092 Ma bulk anorthosite xenoliths and their diabase host rocks of the Beaver Bay Complex within the Midcontinent Rift.
-The anorthosite lithology yielded exceptionally high success rate thanks to the plagioclase crystals shielding the magnetic minerals within.
-Thermal remanent magnetization (TRM) anisotropy test, full TRM acquisition test, together with magnetic imaging show that the bulk anorthosites from the Beaver Bay Complex have low TRM anisotropy, linear TRM acquisition property, and have dominant magnetic carriers interstitial to the plagioclase and are well shielded.
+In this study, we present new paleointensity results from ``IZZI" absolute paleointensity experiments \cite{Yu2004a} on a suite of bulk samples from \textit{ca.} 1092 Ma anorthosite xenoliths and their Beaver River diabase host rocks of the Beaver Bay Complex within the Midcontinent Rift.
+Paleointensity experiments on the Beaver River anorthosite xenoliths has exceptionally high success rate.
+Thermal remanent magnetization (TRM) anisotropy and acquisition experiments reveal that the magnetite-dominated remanence of the anorthosite specimens has low TRM anisotropy and linear TRM acquisition.
+Magnetic imaging shows that the anorthosite specimens to have and dominant magnetic carriers interstitial to the plagioclase and are well shielded.
 With more single-slope, high quality paleointensity data, we seek to compare paleointensity records from the Midcontinent Rift intrusives with those from the extrusives; to test the hypothesis that the Earth's geomagnetic dipole strength remained at high values throughout the Midcontinent Rift; and to further enrich the reliable paleointensity records during the $\sim$25 myr of Midcontinent Rift magmatic activity.
@@
 \hline
-20 & 15 & 5 & 0.6 & TRUE & 0.6 & 2 & 4 \\
+20 & 15 & 5 & 0.6 & TRUE & 0.25 & 2 & 4 \\
 \hline
@@
 Specimens AX3-2r-ni and BD2-2r have prominent Verwey transition but the transition temperatures are suppressed below $\sim$120 K.
-}}
+Specimen BD2-2r has a consistently higher moment during the zero-field-cooled step than during the field-cooled step.
+This is consistent with the interpretation that multidomian magnetic carriers existed in significant quantity within this sepcimen.}}
 \label{fig:MPMS}
@@
 
-Anyways.
 anisotropy of full TRM experiment on accepted specs show low degree of bulk sample scale full TRM anisotropy, with a range of ---.
@@
 Permits for fieldwork and sampling from the Minnesota Department of Natural Resources are gratefully acknowledged.
-We thank James Pierce for assistance in the field.
-
+We thank James Pierce and Blake Hodgin for their assistance in the field.
 We thank Jim Miller for providing us with the Duluth Complex anorthosite thin section.
-
 We thank Dario Bilardello, Peat Solheid, Mike Jackson, Josh Feinberg, and Bruce Moskowitz for their tremendous help of instrumental operations, data interpretations, and research guidance.
@@
 This repository also contains Python code related to calculations, visualizations and statistical tests discussed herein.
-
 \bibliography{YZ_ref}
@@
 % potential reviewer: Peter Selkin, Jeff Gee, Lisa Tauxe, Roger Fu, Lauri Pesonen, Josh Feinberg Julie Bowles,
-
-% I conducted a comparative study of modified IZZI paleointensity experiments on both the diabase and anorthosite, with a group treated with low-field AF demagnetization after each in-field step and another group without.
-Paleointensity experiments for most diabase specimens result in non-ideal Arai plots often with poor pTRM checks, making them difficult to interpret for paleointensity estimate (Fig.
-\ref{fig:pmag}).
-These results can be interpreted to be associated with observed color change of the diabase after heating, which suggests alterations of the oxides.
-The experiments conducted on anorthosite samples, on the other hand, had a high success rate of $>$ 50\%.
-Moreover, the resulting success rate for the anorthosite group with the AF treatment is even higher.
-The more straight Arai plots are likely due to the AF steps mitigating the exhibition of pTRM tails often associated with non-ideal behavior of multi-domain grains that are demagnetized by the pre-treatment step.
-However, as the nearly pure anorthosite samples did not display dramatic alterations that are easily detectable with common petrographic techniques, a question rises as why some of the compositionally similar anorthosites passed the paleointensity selection results and some did not.
 

### 2021-10-07 manuscript/Manuscript.tex
Updates from Overleaf

@@
 Rock magnetic analyses and magnetic imaging reveal that the anorthosite xenoliths can often have magnetic mineralogy dominated by grains with high coercivities, which can yield high success rates passing our selection criteria relative to their diabase hosts.
-Taken together, new and previously published paleointensity data from the Midcontinent Rift support that a strong geomagnetic field  persisted through 25 million years of the late Mesoproterozoic.
-These data are inconsistent with interpretations of continuous progressive decay of Earth's geodynamo strength through the Proterozoic Eon.
+Taken together, new and previously published paleointensity data from the Midcontinent Rift support that a strong geomagnetic field likely persisted through 25 million years of the late Mesoproterozoic.
+These data are inconsistent with interpretations of a progressive decay of Earth's geodynamo strength through the Proterozoic Eon.
 
@@
 \section*{Introduction}
-%The question of the core and the importance of the observational data as complementary for numerical simulation work.
 
-The strong geodynamo today is collectively powered by thermal convection and compositional convection in the liquid outer core.
-Studies have found that Earth may have sustained a geodynamo through the Archean \cite{Tarduno2014a}, and possibly as far back as the Hadean \cite{Tarduno2020a}.
-However, recent estimates of the thermal conductivity of the core suggest that the nucleation of the inner core nucleation could be $<$500 Ma, more recent than previous thought.
-Withouth the power of compositional convection provided by light elements being rejected into the outer core from the inner core, any identification of an earlier geodynamo would indicate that the core was stirred by other power sources, such as precipitation of Mg \cite{ORourcke2016a}, or perhaps
-that the dynamo was generated by exotic processes like a convecting basal magma ocean \cite{Ziegler2013a}.
-In any case, a geodynamo solely driven by thermal conduction through the core-mantle boundary (CMB) prior to the nucleation of the inner core is likely to decay persistently due to heat loss.
-
-That we have a solid inner core today is the result of secular cooling and differentiation of the layered Earth.
-The timing of inner core formation during Earth's thermal evolution is amongst the major unsolved questions in Earth's history.
-Due to the uncertain constraints on the thermal conductivity of the core \cite{Konopkova2016a, Gomi2013a, Ohta2016a}, numerical simulations have given drastically different predictions on the inner core nucleation age, ranging from about half a billion years ago to prior to the Archean \cite{Pozzo2012a, Gubbins2004a, Nimmo2015a}.
-Some models have suggested the thermal-convection-dominated liquid core motion experienced a long-term waning before the nucleation of the inner core, which kick-started the rigorous compositional convection between the solid inner core and liquid outer core \cite<e.g>{Davies2021a}.
-A corollary of such a prediction is that the geomagnetic axial dipole field, which is generated by the fluid motion in the liquid core, would have experienced a long-term decay before the emergence of the solid inner core, after which a significant increase in Earth's axial dipole field strength occurred \cite{Aubert2009a, Davies2021a}.
+Earth's magnetic field is the result of convective flow of liquid iron-alloy in Earth's outer core.
+This self-sustaining geodynamo is collectively powered by the heat flow across the core-mantle boundary (CMB) and from the crystallization of the solid inner core from the liquid outer core which provides latent heat and compositional buoyancy.
+Paleomagnetic studies have found that a dynamo may have existed by the Archean \cite{Tarduno2014a}, and possibly as far back as the Hadean \cite{Tarduno2020a}.
+However, some estimates of the thermal conductivity of the core suggest that the nucleation of the inner core could be younger than 1 Ga, more recent than previous thought \cite{Pozzo2012a, Gomi2013a, Labrosse2015a, Ohta2016a}.
+Without the power of compositional convection provided by the process of light elements being rejected into the outer core from the crystallizing inner core, the geodynamo was likely dominantly driven by thermal conduction through the CMB, with perhaps additional stirring by other power sources such Mg precipitation \cite{ORourcke2016a}, or perhaps there were intervals early in Earth history when the dynamo was generated by exotic processes such as a convecting basal magma ocean \cite{Ziegler2013a}.
+Long-term secular cooling of Earth has the potential to have decreased the power available to the geodynamo through time prior to the nucleation of the inner core \cite<e.g>{Davies2021a}.
+As a result, it is possible that strength of the geomagnetic axial dipole field would have experienced a long-term decay before the emergence of the solid inner core after which it would have strengthened \cite{Aubert2009a, Davies2021a}.
+However, the estimates for thermal conductivity values for the core continue to vary\cite{Konopkova2016a, Ohta2016a}.
+Obtaining an observational record of the timing of inner core formation would constrain the thermal evolution of the core and mantle, in addition to improving our understanding of the history of Earth’s dynamo.
 
@@
 \ref{}).
-\citeA{Pesonen1983b} obtained an average virtual dipole moment of 93 ZAm$^2$ after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using the \textit{ca.} 1108 Ma Osler Volcanics, Maimainse Point volcanics, intrusive Logan Sills, Thunder Bay dikes, and the Baraga and Marquette dikes, as well as the \textit{ca.} 1096 Ma North Shore Volcanic Group.
-\citeA{Sprain2018a} obtained more high-quality paleointensity data using the \textit{ca.} 1108 Ma Osler Volcanics and the \textit{ca.} 1100 Ma Mamainse Point volcanics.
-A calculated mean VADM value of 56 $\pm$ 21 ZAm$^2$ is similar to the average value of the past 300 million years.
-In addition, an average VADM of 59 $\pm$ 11 ZAm$^2$ obtained \citeA{Kulakov2013a} using the \textit{ca.} 1085 Ma Lake Shore Traps volcanics suggests that the present-day-like geomagnetic field strength likely persisted through the late stage of Midcontinent Rift magmatism.
-
-Non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results obtained from Midcontinent Rift rock units.
+\citeA{Pesonen1983b} obtained an average virtual dipole moment of 93 ZAm$^2$ after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using both extrusive and intrusive rocks, including the \textit{ca.} 1108-1105 Ma Osler Volcanics, the Maimainse Point volcanics, intrusive Logan Sills, Thunder Bay dikes, the Baraga and Marquette dikes, as well as lava flow of the \textit{ca.} 1096 Ma North Shore Volcanic Group.
+\citeA{Kulakov2013a} obtained an average VADM of 59 $\pm$ 11 ZAm$^2$ using the \textit{ca.} 1085 Ma Lake Shore Traps volcanics and suggest that the present-day-like geomagnetic field strength likely existed through the late stage of Midcontinent Rift magmatism.
+However, non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results obtained by \citeA{Pesonen1983b, Kulakov2013a}.
 One typical failure of paleointensity experiment is the double-slope behavior, as two distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
 \citeA{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas \citeA{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
-Results with such non-ideal behavior were rejected by \siteA{Sprain2018a} who applied stricter paleointensity selection criteria and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
-Despite the low success rate of paleointensity experiments, specimens that yield dominantly single-slope paleointensity results from the \textit{ca.} 1105 Ma flows of the Osler Volcanic Group, ca.
-1095 Ma flows of the Mamainse Point volcanics \cite{Sprain2018a}, as well as the \textit{ca.} 1085 Ma flows of the Lake Shore Traps \cite{Kulakov2013a} suggest VADM values similar to that of the present-day geomagnetic field \cite{Sprain2018a}.
-Taken together, these rock units that have been dated to span $\sim$25 myr during the development of the Midcontinent Rift magmatism \cite{Swanson-Hysell2019a} may be incompatible with the hypothesis of a ``Proterozoic dipole low" \cite{Biggin2009a}.
+Results with such non-ideal behavior were rejected by \citeA{Sprain2018a} who applied stricter paleointensity selection criteria on new paleointensity results and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
+A calculated mean VADM value of 56 $\pm$ 21 ZAm$^2$ is similar to the average value of the past 300 million years \cite{Sprain2018a}.
+This high VADM value estimate from the Midcontinent Rift volcanics may be incompatible with the hypothesis of a ``Proterozoic dipole low" \cite{Biggin2009a}.
 
@@
 
+\subsection*{Bias Corrected Estimation of Paleointensity}
+While using a set of selection criteria for excluding specimens with non-ideal paleointensity results is one common way of performing data analyses, the selection criteria themselves are often user-determined according to interpretations of certain groups of magnetic mineralogy on a case-by-case basis.
+The operation of performing a binary selection (accept or reject) on specimen results thus has an underlying assumption that the magnetic grain assemblage of specimens that pass the selection have a determined type of magnetic mineralogy.
+However, natural samples often have complex assemblage of magnetic carriers that span a range of composition, grain sizes and shapes.
+Thus, imposing a threshold value on certain parameters based on the paleointensity results can introduce bias from the analyzer.
+Alternatively, \citeA{Cych2021a} developed a Bias Corrected Estimation of Paleointensity (BiCEP) method that can eliminate biased introduced by using a set of arbitrarily defined selection criteria by taking all specimens into account for site-level paleointensity estimates.
+The method assumes that the paleointensity recorded by each specimen is biased away from a true value by an amount that is dependent on a single metric of nonlinearity (the curvature parameter $|\vec{k}|$; \citeA{Paterson2011a}) on the Arai plot.
+Although this method has the advantage of mitigating paleointensity estimate bias, it does not account for specimens that experience detrimental magnetic mineralogy alteration during heating steps unless such specimens are omitted before the analyses.
+
+We apply the BiCEP method to all sites in this study and the cooling rate corrected site-level paleointensity estimates are shown in Figure \ref{fig:BiCEP}.
+For the anorthosite sites (e.g.
+AX11) that show good specimen-level consistency (Figure \ref{fig:PINT_cooling_corrected}) and straight specimen Arai plots (Figure \ref{fig:IZZI_examples}), the paleointensity estimates derived from the BiCEP method using all specimen data is indistinguishable with that derived from the site average value based on selected specimens (Figure \ref{fig:PINT_cooling_corrected}).
+For anorthosites that show rather large specimen-level dispersion (e.g.
+AX8), or diabase sites that have double-slop behaviors and significant MD component, the paleointensity estimates using the BiCEP method tend to have broader uncertainty ranges and the median values are lower than those from anorthosites that typically have more ideal paleointensity behaviors (Figure \ref{fig:PINT_cooling_corrected}).
+Overall, site mean value of the mean probability distribution of the BiCEP method results is consistent with the weighted mean value of the selection criteria-based estimates.
+In addition, the majority of the anorthosite sites give high probability estimate for there being a geomagnetic field value \textit{ca.} 1092 Ma at a paleolatitude of $\sim$30\textdegree similar to that of today.
+This result is consistent with the interpretation that the anorthosite xenoliths record a high geomagnetic field during the late Mesoproterozoic.
+
 \begin{figure}[h!]
-\noindent\includegraphics[width=\textwidth]{PINT_BiCEP.pdf}
+\noindent\includegraphics[width=\textwidth]{code/code_output/AX_BD_BiCEP.pdf}
 \centering
-\caption{\footnotesize{Violin plot of paleointensity estimates using the bias corrected estimation of paleointensity (BiCEP) method \cite{Cych2021a}. }}
+\caption{\footnotesize{Violin plot of paleointensity estimates from the anorthosite and diabase using the bias corrected estimation of paleointensity (BiCEP) method \cite{Cych2021a}.
+The results are grouped by individual cooling units following \citeA{Zhang2021b}.
+All estimates are corrected for cooling rate by a factor of 0.74.
+}}
 \label{fig:BiCEP}
@@
 
-% \begin{figure}
-% \noindent\includegraphics[width=\textwidth]{pmag_plot_.pdf}
-% \caption{\small{(A) Site-mean directions of anorthosite and diabase plotted on an equal-area plot.
-(B) Calculated VGPs from the anorthosite and diabase plotted in context of a previously synthesized \textit{ca.} 1.1 Ga Laurentia APWP from the Midcontinent Rift volcanics \citep{Swanson-Hysell2019a}.
-VGPs from both lithologies fall close to the 1095 Ma pole position.
-(C) Example Arai plots for diabase.
-Most diabase and many anorthosite specimens were rejected by selection criteria due to the zigzagging behavior shown in the diabase example plot.
-(D) Example Arai plots for anorthosite.
-A typical anorthosite specimen that passes selection shows straight Arai plot with a high paleointensity estimate before cooling rate correction.
-Note both specimens in (C) and (D) show dominantly single component magnetization in the inset orthogonal plots.
-The estimated field intensity is not cooling rate corrected.}}
-% \label{fig:pmag}
-% \end{figure}
-
 % At the IRM, I was seeking the answer to this question with the help of the vibrating sample magnetometer (VSM) systems, including a newly installed Lake Shore VSM which greatly helped improve measurement resolution on samples with weak magnetizations (Fig.

### 2021-10-11 manuscript/Manuscript.tex
manuscript update - need to finish the last discussion section and conclusion

@@
 \usepackage{soul}
-
+\usepackage{adjustbox}
 \linenumbers
@@
 This self-sustaining geodynamo is collectively powered by the heat flow across the core-mantle boundary (CMB) and from the crystallization of the solid inner core from the liquid outer core which provides latent heat and compositional buoyancy.
-Paleomagnetic studies have found that a dynamo may have existed by the Archean \cite{Tarduno2014a}, and possibly as far back as the Hadean \cite{Tarduno2020a}.
+Paleomagnetic studies have found that a dynamo may have existed by the Archean \cite{Selkin2007a, Tarduno2014a, Brenner2020a}.
 However, some estimates of the thermal conductivity of the core suggest that the nucleation of the inner core could be younger than 1 Ga, more recent than previous thought \cite{Pozzo2012a, Gomi2013a, Labrosse2015a, Ohta2016a}.
@@
 Long-term secular cooling of Earth has the potential to have decreased the power available to the geodynamo through time prior to the nucleation of the inner core \cite<e.g>{Davies2021a}.
-As a result, it is possible that strength of the geomagnetic axial dipole field would have experienced a long-term decay before the emergence of the solid inner core after which it would have strengthened \cite{Aubert2009a, Davies2021a}.
+As a result, it is possible that rigor of the geomagnetic axial dipole field would have experienced a long-term decay before the emergence of the solid inner core after which it would have strengthened \cite{Aubert2009a, Davies2021a}.
 However, the estimates for thermal conductivity values for the core continue to vary\cite{Konopkova2016a, Ohta2016a}.
@@
 Paleomagnetic data from ancient rocks are one of the few types of observational data that have the potential to record the evolution of Earth’s core and can be used to test the predictions from numerical simulations.
-Paleointensity studies have found very low axial dipole moment (VDM) estimates from rocks formed in the late Ediacaran \cite{Bono2019a, Shcherbakova2019a, Thallner2021b}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
+Some Precambrian absolute paleointensity records from the PINT database (\url{http://earth.liv.ac.uk/pint/}) show evidence of an existence of an appreciable geomagnetic field prior to the Ediacaran.
+On the other hand, studies have found very low axial dipole moment (VDM) estimates from rocks formed in the late Ediacaran \cite{Bono2019a, Shcherbakova2019a, Thallner2021b}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
 Other studies focusing on paleomagnetic directions have also found that the late Ediacaran might have been a period of time when the geomagnetic field were experiencing frequent reversals and waning of dipole dominance \cite{Bono2015a, Kodama2020a}.
 It has been interpreted that these findings are consistent with the results of geodynamo models using high core thermal conductivity values that suggest the inner core nucleation age being near the Ediacaran-Cambrian boundary \cite{Driscoll2016a, Davies2021a}.
-\citeA{Bono2019a} based on a polynomial fit through filtered Precambrian paleointensity database (PINT; \url{http://earth.liv.ac.uk/pint/}) to suggest the selected observational records show that the geodynamo power progressively decreased through the Proterozoic.
-They also recovered very low paleointensity values from \textit{ca.} 565 Ma single silicate crystals and interpreted them to be reflecting that the geodynamo reached its lowest near the Ediacaran-Cambrian boundary.
-Such decay trend in the Neoproterozoic is recently supported by new low paleointensity estimates derived from the \textit{ca.} 720 Ma Franklin Large Igneous Province \cite{Lloyd2021a}.
+\citeA{Bono2019a} based on a polynomial fit through filtered Precambrian paleointensity database to suggest the selected observational records show that the geodynamo power progressively decreased through the Proterozoic (Figure \ref{fig:MCR_VADM}).
+That study also recovered very low paleointensity values from \textit{ca.} 565 Ma single silicate crystals and interpreted them to be reflecting that the geodynamo reached its lowest near the Ediacaran-Cambrian boundary.
+New low paleointensity estimates derived from the \textit{ca.} 720 Ma Franklin Large Igneous Province are interpreted to support the existence of such decay trend in the Neoproterozoic \cite{Lloyd2021a}.
 
-However, previous paleointensity data from intrusive and extrusive rock units in the Mesoproterozoic Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record Earth's virtual axial dipole moment (VADM) values similar to that of today, which are higher than the prediction of the long term decay curve (Fig.
-\ref{}).
+However, previous paleointensity data from intrusive and extrusive rock units in the Mesoproterozoic Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record Earth's virtual axial dipole moment (VADM) values similar to that of today, which are higher than the prediction of the long term decay curve (Figure \ref{fig:MCR_VADM}).
 \citeA{Pesonen1983b} obtained an average virtual dipole moment of 93 ZAm$^2$ after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using both extrusive and intrusive rocks, including the \textit{ca.} 1108-1105 Ma Osler Volcanics, the Maimainse Point volcanics, intrusive Logan Sills, Thunder Bay dikes, the Baraga and Marquette dikes, as well as lava flow of the \textit{ca.} 1096 Ma North Shore Volcanic Group.
-\citeA{Kulakov2013a} obtained an average VADM of 59 $\pm$ 11 ZAm$^2$ using the \textit{ca.} 1085 Ma Lake Shore Traps volcanics and suggest that the present-day-like geomagnetic field strength likely existed through the late stage of Midcontinent Rift magmatism.
+\citeA{Kulakov2013a} obtained an average VADM of 59 $\pm$ 11 ZAm$^2$ using the \textit{ca.} 1085 Ma Lake Shore Traps volcanics and suggest that the present-day-like geomagnetic field strength likely existed through the late stage of Midcontinent Rift magmatism \cite{Vervoort2007a, Miller2013a}.
 However, non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results obtained by \citeA{Pesonen1983b, Kulakov2013a}.
@@
 Although it is typical of paleointensity experiments to suffer from low success rates, bulk plagioclase-rich rocks \cite{Selkin2000a} and single plagioclase crystals \cite{Tarduno2005a} are found to be able to produce high quality paleointensity results.
-It has been suggested plagioclase crystals can protect magnetic minerals that exsolved within them against thermochemical alteration during paleointensity experiments.
-In this study, we present new paleointensity results from ``IZZI" absolute paleointensity experiments \cite{Yu2004a} on a suite of bulk samples from \textit{ca.} 1092 Ma anorthosite xenoliths and their Beaver River diabase host rocks of the Beaver Bay Complex within the Midcontinent Rift.
-Paleointensity experiments on the Beaver River anorthosite xenoliths has exceptionally high success rate.
-Thermal remanent magnetization (TRM) anisotropy and acquisition experiments reveal that the magnetite-dominated remanence of the anorthosite specimens has low TRM anisotropy and linear TRM acquisition.
-Magnetic imaging shows that the anorthosite specimens to have and dominant magnetic carriers interstitial to the plagioclase and are well shielded.
+It has been suggested that plagioclase crystals can protect magnetic minerals that exsolved within them against post-formation weathering and thermochemical alteration during paleointensity experiments \cite{Tarduno2005a}.
+In this study, we present new paleointensity results from ``IZZI" modified absolute paleointensity experiments \cite{Yu2004a} on a suite of bulk samples from \textit{ca.} 1092 Ma anorthosite xenoliths and their Beaver River diabase host rocks of the Beaver Bay Complex within the Midcontinent Rift.
+Paleointensity experiments on the Beaver River anorthosite xenoliths have a very high success rate.
+The consistent specimen- and site-level paleointensity results reveal that the magnetite-dominated remanence of the anorthosite specimens has low thermal remanent magnetization (TRM) anisotropy and linear TRM acquisition.
+Magnetic imaging shows that the anorthosite specimens to have dominant magnetic carriers interstitial to and within plagioclase crystals and they do not display strong preferred orientations.
 With more single-slope, high quality paleointensity data, we seek to compare paleointensity records from the Midcontinent Rift intrusives with those from the extrusives; to test the hypothesis that the Earth's geomagnetic dipole strength remained at high values throughout the Midcontinent Rift; and to further enrich the reliable paleointensity records during the $\sim$25 myr of Midcontinent Rift magmatic activity.
@@
 The Mesoproterozoic North American Midcontinent Rift (MCR) is an intracontinental rift that did not result in the break-up of Laurentia.
-Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; \citeA{Vervoort2007a, Miller2013a}) was punctuated with rapid and voluminous emplacement of intrusive rocks as well as extrusive volcanics.
-Based on high-precision $^{206}$Pb/$^{238}$U geochronology and paleomagnetic data, \citeA{Swanson-Hysell2021a} hypothesized that the massive Duluth Complex and the coeval North Shore Volcanic group (Figure \ref{fig:Geologic_map}) formed as a result of having upwelling mantled-derived plume material encountering Laurentia's thinned lithosphere due to rifting, following an ``upside-down” drainage" which rapidly transported melt to today's North Eastern Minnesota (Figure \ref{fig:Geologic_map}; \citeA{Sleep1997a, Swanson-Hysell2014a}).
-The plagioclase-phyric Anorthositic Series rocks of the Duluth Complex that outcrop along an arcuate shape along Northeastern Minnesota (Figure \ref{fig:Geologic_map}) are a result of such vigorous magmatic activity.
+Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; \citeA{Vervoort2007a, Miller2013a}) was punctuated by rapid and voluminous emplacement of intrusive rocks as well as extrusive volcanics.
+Based on high-precision $^{206}$Pb/$^{238}$U geochronology and paleomagnetic data, \citeA{Swanson-Hysell2021a} hypothesized that the massive Duluth Complex and the coeval North Shore Volcanic group (Figure \ref{fig:Geologic_map}) formed as a result of having upwelling mantled-derived plume material encountering Laurentia's thinned lithosphere, following an ``upside-down” drainage" which rapidly funneled melt to today's North Eastern Minnesota (Figure \ref{fig:Geologic_map}; \citeA{Sleep1997a, Swanson-Hysell2014a}).
+The plagioclase-phyric Anorthositic Series rocks, the crystal-poor Layered Series intrusions of the Duluth Complex, and the coeval volcanics of the North Shore Volcanic Group (NSVG) that outcrop along Northeastern Minnesota (Figure \ref{fig:Geologic_map}) are interpreted to be a result of such vigorous magmatic activity.
 
@@
 
-The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex punctuates another period of Midcontinent Rift main stage magmatism.
+The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex punctuates another period of magmatism during the Midcontinent Rift main stage activity.
 $^{206}$Pb/$^{238}$U zircon dates and cross-cutting relationships tightly constraints the areally extensive Beaver River diabase of the Beaver Bay Complex to have been emplaced at 1091.7 $\pm$ 0.2 Ma \cite{Zhang2021b}.
-The Beaver River diabase intruded into hypabyssal depths via wide conduits that accommodated the entrainment of numerous anorthosite xenoliths that have short-axis diameters up to 180 meters.
+The Beaver River diabase intruded into hypabyssal depths via wide conduits that accommodated the entrainment of numerous anorthosite xenoliths that have short-axis diameters up to 180 meters \cite{Boerboom2004a, Boerboom2006b}.
 Geochronological, paleomagnetic, geochemical, and petrographic data are consistent with the hypothesis that the Beaver River diabase magma extruded to the surface, feeding the massive Greenstone Flow - one of the largest lava flows on Earth \cite{Doyle2016a, Zhang2021b}.
@@
 In addition, petrographic imaging support that the plagioclase from the Beaver River anorthosite formed under a magmatic conditions different from the Duluth Complex anorthositic series magma.
-While the plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture (Fig.
-\ref{Petro_QDM}), the more calcium-rich plagioclase of the Beaver River anorthosite often develop granoblastic texture characterized by equigranular plagioclase crystals (Fig, \ref{Petro_QDM}).
+While the plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture (Figure \ref{Petro_QDM}A), the more calcium-rich plagioclase of the Beaver River anorthosite often develop granoblastic texture characterized by equigranular plagioclase crystals (Figure \ref{Petro_QDM}B).
 Abundant magnetite and ilmenite needles that are typically 10 $\mu$m long can be found to have preferred orientations within Duluth Complex plagioclase.
-Vermicular titanomagnetite and ilmenite symplectic intergrowths also exist surrounding accessory olivine or pyroxene in the Duluth Complex Anorthositic Series rocks (Fig.
-\ref{Petro_QDM}).
+Interstitial vermicular titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex Anorthositic Series rocks \cite{Miller1990a}.
 
@@
 \caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from Duluth Complex Anorthositic Series (left column) and Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (right column; \citeA{Zhang2021b}).
-Petrographic images (A-D) show Duluth Complex anorthosite have euhedral, elongate plagioclase crystals with interlocking texture and strong igneous foliation.
-The plagioclase can contain abundant Fe-Ti oxide needles (C) that have preferred orientations that are often parallel with the (010) planes of plagioclase.
-In contrast, plagioclase of sample MS99033 do not enclose oxide needles.
-Magnetic field maps acquired by the quantum diamond microscope (QDM) show that the oxide needles within Duluth plagioclase are strongly magnetic as evidenced through the overlapping magnetic dipole fields exerted above the oxides in (E).
-In comparison, the plagioclase of MS99033 is not magnetic.
-}}
+Petrographic images (A-D) show Duluth Complex anorthosite have euhedral, elongate plagioclase crystals with interlocking texture and igneous foliation.
+(C) The plagioclase can contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the (010) planes of plagioclase.
+In contrast, plagioclase of sample MS99033 do not enclose large oxide needles.
+Magnetic field maps acquired by the quantum diamond microscope (QDM) show that the oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the overlapping magnetic dipole fields exerted above the oxides in (E).
+In comparison, relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite.
+The dipole-like features in (F) do not show strong preferred orientations.
+Magnetic field maps of (E) and (F) share the same spatial and magnetic scale.
+The unit of the axes are in $\mu$m.}}
 \label{fig:Petro_QDM}
@@
 
-Similar occurrences of magnetite and ilmenite inclusions within plagioclase have been described by \citeA{Haggerty1976a} in slowly cooled, mafic intrusions, and \cite{Scofield1986a} hypothesized that the inclusions acted as a means of buffering the melt with respect to oxygen, forming either crystallites along preexisting plagioclase substrate or as iron incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures.
+Occurrences of magnetite and ilmenite symplectites and needles within anorthositic rocks have been described by \citeA{Scofield1975a, Scofield1986a} in the Wichita Mountains and by \citeA{Haggerty1976a} in other slowly cooled, mafic intrusions.
+\citeA{Scofield1986a} hypothesized that the inclusions acted as a means of buffering the melt with respect to oxygen, forming either crystallites along preexisting plagioclase substrate or as iron incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures.
 Such a model requires that plagioclase, instead of other mafic minerals such as olivine, to be the first to crystallize from a magma and act as a sink for iron.
@@
 Particularly, \citeA{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
-This is consistent with the interpretation that these anorthosite were rapidly transported into shallow magma chambers without reaching structural equilibrium.
-Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Fig.
-\ref{Petro_QDM}).
-Instead, relatively sparse titano-magnetite interstitial to the plagioclase are typical of Beaver River anorthosites (Fig, \ref{}).
+This is consistent with the interpretation that these anorthosite were rapidly transported by upward currents into hypabyssal magma chambers without reaching structural equilibrium.
+Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Figure \ref{Petro_QDM}).
+Instead, relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Figure \ref{Petro_QDM}).
 
@@
 Sister specimens underwent step-wise alternating field (AF) or thermal demagnetization at the UC Berkeley Paleomagnetism Lab and the results are presented in \citeA{Zhang2021b}.
-Paleomagnetic direction measurement level data are available within the MagIC database (\url{https://earthref.org/MagIC/doi/10.1029/2021GC009909}).
-The close directional similarities between the anorthosite xenolith and their host diabase, the overall cooling unit mean directions between the two lithologies, and thermal modeling results are consistent with the interpretation that the Beaver River diabase and anorthosite xenoliths acquired primary thermal remanent magnetization \textit{ca.} 1092 Ma during cooling after the emplacement of the diabase \cite{Zhang2021b}.
-Based on the anorthosite thermal demagnetization results we selected sites whose unblocking temperature ranges are narrow and near 580\textdegree C.
-Beaver River diabase sites with minimal secondary remanence component were also selected for paleointensity experiments.
+The close directional similarities between the anorthosite xenoliths and their host diabase and thermal modeling results are consistent with the interpretation that the Beaver River diabase and anorthosite xenoliths acquired primary thermal remanent magnetization \textit{ca.} 1092 Ma during cooling after the emplacement of the diabase \cite{Zhang2021b}.
+Based on the anorthosite thermal demagnetization results we selected sites whose unblocking temperature ranges are narrow and near 580\textdegree C for paleointensity experiments.
+Beaver River diabase sites with minimal secondary remanence were also selected for paleointensity experiments.
 
 \subsection*{Paleointensity experiment}
-A total of 69 specimens from 7 diabase sites and a total of 86 specimens from 14 anorthosite xenoliths underwent paleointensity experiments that followed the step-wise double-heating Thellier method \cite{Thellier1959a} using the IZZI protocol \cite{Tauxe2004a} with heating steps up to 585 \textdegree C.
+A total of 86 specimens from 14 anorthosite xenoliths and a total of 69 specimens from 7 diabase sites underwent paleointensity experiments that followed the step-wise double-heating Thellier method \cite{Thellier1959a} using the IZZI protocol \cite{Tauxe2004a} with heating steps up to 585 \textdegree C.
 Partial thermal remanent magnetization (pTRM) checks were performed systematically throughout the experiment to test whether there was significant mineralogical alteration due to heating and were assessed using the SCAT parameter of \citeA{Shaar2013a}.
@@
 During each heating step, samples remained at peak temperatures for 20 min.
-An applied laboratory field of 30 μT was used for all in-field steps.
+An applied laboratory field of 30 $\mu$T was used for all in-field steps.
 All heating steps were performed in air.
-The temperature increments for the experiments were chosen to cover characteristic remanent magnetizations held by (titano)magnetite informed by the previous demagnetization data, with smaller increment temperature steps performed close to the expected unblocking temperature of stoichiometric magnetite \cite{Zhang2021b}.
+The temperature increments for the experiments were chosen to cover characteristic remanent magnetizations held by (titano)magnetite informed by the previous demagnetization data, with smaller increments performed close to the expected unblocking temperature of stoichiometric magnetite \cite{Zhang2021b}.
 
-After the paleointensity experiment were finished, the same specimens underwent additional TRM anisotropy check with full TRMs given at 0, 30, 60, 90, 120, 150 degrees in the horizontal plane in specimen coordinates.
+% After the paleointensity experiment were finished, the same specimens underwent additional TRM anisotropy checks with full TRMs given at 0, 30, 60, 90, 120, 150 degrees in the horizontal plane in specimen coordinates.
 After that, full TRM acquisition experiments were performed on the same specimens with applied fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axis.
@@
 \subsection*{Paleointensity result selection}
-The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD; \citeA{Kirschvink1980a}) of <20\textdegree; (2) scatter parameter ($\beta$; \citeA{Coe1978a}) values of $<$15 per cent; (3) a deviation angle (DANG; \citeA{Tanaka2003a, Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC; \citeA{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT; \citeA{Shaar2013a}) = TRUE; (6) a maximum gap (GAP-Max; \citeA{Shaar2013a}) < 0.6; (7) number of pTRM checks > 2; (8) and number of measurements used for paleointensity determination $≥$ 4.
+The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD; \citeA{Kirschvink1980a}) of $<$20\textdegree; (2) scatter parameter ($\beta$; \citeA{Coe1978a}) values of $<$15 per cent; (3) a deviation angle (DANG; \citeA{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC; \citeA{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT; \citeA{Shaar2013a}) = TRUE; (6) a maximum magnetic moment difference between adjacent zero-field steps (GAP-Max; \citeA{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
 The criteria are also summarized in Table \ref{tab:criteria}.
-The MAD measures the scatter about the best-fit line through NRM steps in the selected interval for which the intensity is defined.
-DANG, the deviation angle, is the angle between the best-fit direction that is free floating and the direction between the centre of mass of the data and the origin of the vector component diagram (\citeA{Tanaka2003a}; \citeA{Tauxe2004a}).
+The MAD measures the scatter about the best-fit line through the natural remanent magnetization (NRM) steps in the selected interval for which the intensity is defined.
+DANG, the deviation angle, is the angle between the best-fit direction that is free floating and the direction between the centre of mass of the data and the origin of the vector component diagram \cite{Tauxe2004a}.
 Both MAD and DANG assess the directional variation of the NRM, with MAD measuring the scatter in the NRM directions and DANG assessing whether the component is trending toward the origin of the Zijderveld plot.
@@
 $\beta$, FRAC, GAP-Max and SCAT are all statistics to assess the behavior of Arai plots.
-See the Standard Paleointensity Definitions (\citeA{Paterson2014a}, \url{https://earthref.org/PmagPy/SPD/home.html}) for more details.
+See the Standard Paleointensity Definitions (\citeA{Paterson2014a}; \url{https://earthref.org/PmagPy/SPD/home.html}) for more details.
 Data analysis was conducted using Thellier GUI \cite{Shaar2013a} within the PmagPy software package \cite{Tauxe2016a}.
@@
 \begin{table}
-\caption{Quality criteria used for accepting or rejecting paleointensity data. See the text for more detail.} 
+\caption{Quality criteria used for selecting paleointensity data. See the text for more detail.} 
 \centering
@@
 \hline
-MAD (o) & Beta (per cent) & DANG (o) & FRAC& SCAT & GAP-MAX & NpTRM & N Arai\\
+MAD (\textdegree) & Beta (per cent) & DANG (\textdegree) & FRAC& SCAT & GAP-MAX & NpTRM & N Arai\\
 \hline
@@
 
-In this study, we conduct rock magnetic experiments with a purpose of gaining magnetic mineralogy insight into the paleointensity results of the anorthosite and diabase.
-Backfield curves were measured at room temperature using a Micromag Princeton Measurements vibrating sample magnetometer (VSM) with nominal sensitivity of $5 \times 10^{-9}$ Am\textsuperscript{2}, and a Lake Shore 8600 series VSM with nominal sensitivity of xxxxxxxxxx.
+In this study, we conduct rock magnetic experiments with a purpose of gaining magnetic mineralogy insights into the paleointensity results of the anorthosite and diabase.
+Backfield curves were measured at room temperature using a Micromag Princeton Measurements vibrating sample magnetometer (VSM) and a Lake Shore 8600 series VSM.
+Specimen median destructive field (MDF) are calculated based on the backfield curves.
 The calculated coercivity spectra were subsequently decomposed into one ore more components using skew-normal distributions following the method of \citeA{Maxbauer2016a}.
@@
 We also used the magnetic property measurement system (MPMS) at the IRM to aid in the identification of magnetic mineral.
-In the field-cooled (FC) experiments shown in Figure \ref{fig:MPMS}, the magnetization was measured upon warming following the specimen having cooled in an applied field of 2.5 T from 300 to 10 K.
+In the field-cooled (FC) experiments, specimen magnetizations were measured upon warming following the specimen having cooled in an applied field of 2.5 T from 300 to 10 K.
 In the zero-field-cooled (ZFC) experiment, a low-temperature saturation isothermal remanence (LTSIRM) of 2.5 T was applied at 10 K after the specimen cooled in a (near-)zero field.
-In the room-temperature saturation isothermal remanence (RTSIRM) experiment, the sample was pulsed with a 2.5 T field at room temperature ($\sim$300 K) and then cooled to 10 K and warmed back to room temperature in a (near)zero field.
+In the room-temperature saturation isothermal remanence (RTSIRM) experiment, the sample was pulsed with a 2.5 T field at room temperature ($\sim$300 K) and then cooled to 10 K and warmed back to room temperature in a (near-)zero field.
 The magnetic moment transitions at critical temperatures revealed through MPMS experiments are used to identify magnetic minerals such as magnetite \cite{Verwey1939a} within specimens.
-% it would actually be nice to show the MPMS data (although it could be a bit redundant).
-The MPMS data nicely shows the BD and failed AX have the Verwey transition suppressed to lower temperatures and the successful AX have Verwey transition close to 120 K - the typical value for stochiometric magnetite.
-I don't think I will be using the hysteresis loops.
+
+To further identify the magnetic carriers within the Beaver River anorthosite xenoliths and compare them with the anorthosites of the Duluth Complex Anorthositic Series rocks, we used the quantum diamond microscope (QDM) at the UC Berkeley Paleomagnetism laboratory to image a thin section of sample MS99033 which is a hand sample from anorthosite xenolith AX16 (which yielded a $^{206}$Pb/$^{238}$U zircon date of 1091.83 $\pm$ 0.21 Ma; \citeA{Zhang2021b}), and a thin section of a Duluth Complex anorthosite (Figure \ref{fig:Petro_QDM}).
+We the QDM to image the magnetic field over the polished thin section surfaces at a sample-sensor distance of 5 $\mu$m in projective magnetic microscopy (PMM) mode with a spatial resolution of 4.7 μm per pixel and an instantaneous 0.9 mT bias field that is canceled during the course of measurement \cite{Glenn2017a}).
 
@@
 40 from a total of 86 anorthosite specimens and 9 out of a total of 69 diabase specimens passed our paleointensity selection criteria, yielding 7 anorthosite sites and 1 diabase sites with site-level paleointensity results that pass paleointensity selections.
-Example Arai plots are shown in Figure \ref{fig:IZZI_examples}.
-Summary specimen absolute paleointensity estimate and site-level weighted mean paleointensity values are shown in Figure \ref{fig:PINT_cooling_corrected}.
+Example NRM/TRM plots (Arai plots;
+\cite{Arai1963a}) are shown in Figure \ref{fig:IZZI_examples}.
+Summary specimen absolute paleointensity estimates and site-level weighted mean paleointensity values are shown in Table \ref{tab:PINT_result} and are plotted in Figure \ref{fig:PINT_cooling_corrected}.
 All measurement-level paleointensity experiment data can be accessed through the MagIC database (\url{}; UPDATE WHEN DOI IS GENERATED).
@@
 Plots on the top row show successful specimen paleointensity results with straight, single-slope behaviors that pass our selection criteria.
-The green lines represent our fitting for the dominant single-slope component that passes our acceptance criteria with estimates of the ancient field (B$_anc$) being used to determine site means.
+The green lines represent our fitting for the dominant single-slope component that passes our acceptance criteria with estimates of the ancient field (B$_{anc}$) being used to determine site means.
 The plots for anorthosite specimens AX1-1a and diabase BD5-1b on the bottom row show typical non-ideal sagging and zigzagging behaviour that fails our acceptance criteria.
-Plot for diabase specimen BD2-5a shows typical paleointensity behavior of this site - specimens typically have double-slope behaviors although the lower temperature slopes can pass the selection criteria, resulting in the high estimated B$_anc$ values shown in Figure \ref{fig:PINT_cooling_corrected}.
-Thus, the estimates of B$_anc$ for this specimen plot is illustrative, but were not used to calculate a site mean paleointensity value in Figure \ref{fig:PINT_cooling_corrected}.}}
+Plot for diabase specimen BD2-5a shows typical paleointensity behavior of this site - specimens typically have double-slope behaviors although the lower temperature slopes can pass the selection criteria, resulting in the high estimated B$_{anc}$ values shown in Figure \ref{fig:PINT_cooling_corrected}.
+Thus, the estimates of B$_{anc}$ for this specimen plot is illustrative, but were not used to calculate a site mean paleointensity value in Figure \ref{fig:PINT_cooling_corrected}.}}
 \label{fig:IZZI_examples}
@@
 The Beaver River anorthosite specimens yielded a high pass rate ($\sim$46.5\%) through our selection criteria.
-In distinct contrast, the Beaver Rider diabase specimens were almost all rejected by our selection criteria.
-Only site BD2 has 9 specimens that passed selection ($\sim$13.0\% pass rate).
+In distinct contrast, the Beaver River diabase specimens are all rejected by our selection criteria.
 Paleointensity experiment results from both lithologies show three types of dominant specimen behaviors on Arai plots (Figure \ref{fig:IZZI_examples}; \citeA{Arai1963a}): dominantly single-slope, double-slope, and zigzagging or sagging (Figure \ref{fig:IZZI_examples}).
-It is worth mentioning that almost all specimens have minimal secondary remanence component in thermal demagnetization results as shown in orthogonal plots of all zero-field steps (Figure \ref{fig:IZZI_examples}).
+Note that almost all specimens have minimal secondary remanence component in thermal demagnetization results as shown in example orthogonal plots of all zero-field steps (Figure \ref{fig:IZZI_examples}).
 Typical specimen paleointensity behavior of the anorthosites display straight, single-slope Arai plots and the accepted fractions of temperature steps span over the origin-trending, primary remanence components (Figure \ref{fig:IZZI_examples}).
-We accept all specimen- and site-level absolute paleointensity estimates from the anorthosite xenoliths.
-In terms of the diabase specimens, all but nine of them failed the selection because of double-slope behavior (fail FRAC selection), poor pTRM checks, and zigzagging (fail SCAT, DANG selection) Arai plots (Figure \ref{fig:IZZI_examples}).
-Even specimens in site BD2, the only site that passed selection, do not always display ideal Arai plots.
+We accept all specimen- and site-level absolute paleointensity estimates from the anorthosite xenoliths that pass the selection.
+In terms of the diabase specimens, all failed the selection because of double-slope behavior (fail FRAC selection), poor pTRM checks, and zigzagging (fail SCAT, DANG selection) Arai plots (Figure \ref{fig:IZZI_examples}).
 As is shown in Figure \ref{fig:IZZI_examples}, specimen BD2-5b has a high-temperature second slope with zigzagging behavior, poor pTRM checks, and low NRM/TRM ratio, in addition to the straight slope with steep NRM/TRM ratio over the lower temperature range.
-Other accepted specimen results from site BD2 are from similar low-temperature fractions of Arai plots (below 500 $^\circ$C), which are different than those used in anorthosite sites (typically above 500 $^\circ$).
-Although the BD2 specimens' lower temperature fraction of the Arai plots are straight, the pTRM checks are consistent, and the specimen paleointensity results show within-site consistency and yield a weighted mean paleointensity estimate indistinguishable with that of anorthosite AX11 and AX13 (Figure \ref{fig:IZZI_examples}, \ref{fig:PINT_cooling_corrected}), specimens from BD2 give consistently higher paleointensity estimates than those from site AX6, AX8, and AX10, which are anorthosite xenoliths included within BD2 (Figure \ref{fig:Geologic_map}; \citeA{Zhang2021a}).
+Although the BD2 specimens' lower temperature fraction of the Arai plots are straight andthe pTRM checks are consistent, lower temperature slopes from BD2 give consistently higher paleointensity estimates than those from site AX6, AX8, and AX10, which are anorthosite xenoliths included within BD2 (Figure \ref{fig:Geologic_map}; \citeA{Zhang2021a}).
 This higher estimate from the diabase lithology could be attributed to the systematic bias of using the lower temperature slope of the Arai plots demonstrated in Figure \ref{fig:IZZI_examples}.
+Thus we only accept the single-slope anorthosite specimens results for further analyses.
 
-% need to export the figure with tight boundaries
+% \begin{figure}[h!]
+% \noindent\includegraphics[width=\textwidth]{Cooling_rate_correction.pdf}
+% \centering
+% \caption{\small{}}
+% \label{fig:cooling_correction}
+% \end{figure}
+
 \begin{figure}[h!]
@@
 \caption{\small{Summary plot of cooling rate corrected individual specimen absolute paleointensity results (square and diamond symbols) from this study and their averages at site (grey bar) and locality level (weighted mean by site-level standard deviations; orange bar).
-The inset plot shows the graph of predicted paleointensity overestimate due to slow cooling of the intrusive Beaver River diabase relative to the cooling rate in laboratory following the model of \citeA{Halgedahl1980a}.
-Because the majority of the anorthosites have unblocking temperatures between 500\textdegree C and 580\textdegree C, we estimate that the slow cooling during natural remanence acquisition could have resulted in a maximum of 36\% overestimate.
-Thus a correction factor of 0.74 is applied at specimen level for the summary plot.
+The left plot shows the graph of predicted paleointensity overestimate due to slow cooling of the intrusive Beaver River diabase and anorthosite xenoliths relative to the cooling rate in laboratory following the model of \citeA{Halgedahl1980a}.
+Because the majority of the anorthosites have unblocking temperatures between 500\textdegree C and 580\textdegree C, we estimate that the slow cooling during natural remanence acquisition could have resulted in a 36\% overestimate.
+Thus a correction factor of 0.74 is applied at specimen level for the paleointensity summary plot.
 }}
@@
 
-Alternating field treatment after in-field heating steps likely improved the paleointensity behavior of diabase specimens such that they pass the selection criteria but the treatment did not result in significant success rate improvement or changes in paleointensity estimates for the anorthosites (Figure \ref{fig:PINT_cooling_corrected}).
-8 out of 13 specimens from site BD2 that were subjected to AF treatment have straight low temperature slopes that pass the selection criteria, whereas only 1 from 4 specimens that were not subjected to AF treatment passes.
+The 20 mT alternating field treatment after in-field heating steps did not make any diabase specimen to pass the selection criteria.
 On the other hand, 27 out of 58 ($\sim$46.6\%) anorthosite specimens that had AF treatment passed selection and 13 out of 28 ($\sim$46.4\%) specimens without AF treatment passed selection.
-Paleointensity estimates from specimens with or without AF treatment in anorthosite AX12, AX13, AX16 do not show significant difference (Figure \ref{fig:PINT_cooling_corrected}).
+Paleointensity estimates from specimens with or without AF treatment in anorthosite AX10, AX11, AX12, AX13, AX16 do not show significant difference (Figure \ref{fig:PINT_cooling_corrected}).
 Specimen plots for AX11 particularly illustrates a case where all specimens within the site yield almost the same paleointensity estimate, regardless of the AF treatment.
-Such consistent behavior makes the results from this site of very high confidence.
+Therefore, the alternating field demagnetization might have improve the specimen behavior for those anorthosites that were subjected to this treatment but it did not result in significant increase in specimen success rate and did not result in significant difference for the site-level paleointensity estimates.
+Together with the narrow range of unblocking temperatures of their sister specimens \cite{Zhang2021b}, these data are consistent with the interpretation that the anorthosites with successful paleointensity results have a magnetic mineral assemblage that likely contain dominantly single domain, low Ti magnetite and they do not contain a significant population of MD or vortex state grains.
+Some anorthosite specimens and all diabase specimens still failed in result selections even though AF treatment was applied.
+It could be that they have a very high dominance of MD or vortex state grains with relatively moderate coercivity such that an alternating field with 20 mT peak field did not significantly mitigate their non-ideal paleointensity behaviors.
+Another cause for their failure could be the thermochemical alterations that happened during heating in air.
+As can be exampled by the zigzagging high temperature slope with poor pTRM checks in paleointensity plot for specimen BD2-5b (Figure \ref{fig:IZZI_examples}), the decrease in NRM/TRM ratios during higher temperature steps likely resulted from the formation of new ferromagnetic material within the specimen due to oxidation.
+Such alterations can allow specimens to acquire additional remanence during in-field steps that may not be fully removed by an AF treatment.
+Overall, the consistent specimen-level behavior makes the successful results from the anorthosites of high confidence.
 
-The different effects that alternating field treatment had upon anorthosite and diabase may reflect the magnetic mineralogy difference between lithologies.
-Diabase likely has significant populations of MD or vortex state grains which can have relatively low coercivities and result in non-ideal paleointensity behaviors, but whose remanence contribution could be significant to the whole rock natural remanent magnetization.
-Upon the application of the 20 mT peak AF treatment after each in-field heating step, the remanence acquired by these grains is likely significantly removed, therefore making a straight slope of NRM/TRM during temperatures below 500\textdegree C (Figure \ref{fig:backfield}).
-On the other hand, the anorthosite specimens with and without AF treatment have very similar paleointensity behaviors and consistent specimen level estimates.
-Together with the narrow range of unblocking temperatures of their sister specimens \cite{Zhang2021b}, these data are consistent with the interpretation that the anorthosites with successful paleointensity results have a magnetic mineral assemblage that likely contain dominantly single domain, low Ti magnetite and they do not contain a significant population of MD or vortext state grains.
-Some anorthosite specimens and many diabase specimens still failed in result selections even though AF treatment was applied.
-It could be that they have a very high dominance of MD or vortex state grains with relatively moderate coercivity such that an alternating field with 20 mT peak field did not significantly mitigate their non-ideal paleointensity behaviors.
-Another cause for their failure could be of thermochemical alterations that happened during heating in air.
-As can be exampled by the zigzagging high temperature slope with poor pTRM checks in paleointensity plot for specimen BD2-5b (Figure \ref{fig:IZZI_examples}), the decrease in NRM/TRM ratios during higher temperature steps likely resulted from the formation of new ferromagnetic material within the specimen due to oxidation.
-Such alteration can allow specimens to acquire additional remanence during in-field steps that may not be fully removed by an AF treatment.
+In addition to applying paleointensity selection criteria to our experiment results, we applied the bias corrected estimation of paleointensity (BiCEP) method by \cite{Cych2021a} to all of our anorthosite and diabase specimens regardless of their Arai plot statistics.
+The BiCEP method is a Bayesian statistical method based on an assumption that all specimen paleointensity results from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the Arai plot \cite{Paterson2011a}.
+This method has the advantage of using as many specimens as available for estimating site-level paleointensities.
+The posterior distributions of paleointensity estimates given by the BiCEP method show that for the anorthosites that pass our paleointensity selection criteria, probabilities are very low for the case where these anorthosites are recording heavily biased paleointensity estimates from a true value below 10 $\mu$T (Supporting Information S1).
+Therefore, the anorthosite xenoliths record a high geomagnetic field \textit{ca.} 1092 Ma.
+
+\begin{sidewaystable}
+\caption{\footnotesize{Summary paleointensity results for specimens that passed our selection.
+Paleointensity results for specimens that passed quality criteria.
+B$_{anc}$ is the calculated ancient field intensity over the chosen temperature interval in $\mu$T.
+T$_{min}$ and T$_{max}$ indicate the temperature interval over which the best fit for paleointensity was defined.
+N is the number of steps used within the selected interval for paleointensity determination.
+FRAC is the fraction of remanence.
+NpTRM shows the number of pTRM checks within the selected interval for paleointensity determination.
+$\beta$ is the scatter parameter.
+GAP-MAX is the maximum magnetization gap between two adjacent steps.
+MAD is the maximum angle of deviation.
+DANG is the deviation angle.
+SCAT is the scatter parameter.
+Paleolatitude is calculated from the inclination values reported in \citeA{Zhang2021b}.
+$\gamma$ is the gamma statistic that measures the angle between the last pTRM step used for paleointensity determination and the applied field direction.
+V(A)DM is the virtual (axial) dipole moment reported in $10^{21}$Am$^2$ ((ZAm$^2$)).
+}}
+\centering
+\begin{tabular}{cccccccccccccccc}
+\hline
+Site & Specimen & B$_{anc}$ & T$_{min}$ & T$_{max}$ & N    & FRAC & NpTRM & $\beta$ & GAP-MAX & MAD ($^\circ$) & DANG ($^\circ$) & SCAT & Paleolatitude & $\gamma$ & VADM (ZAm$^2$) \\
+\hline
+AX6  & AX6-2a   & 18.25     & 400       & 585       & 18 & 0.7  & 10.0  & 0.04    & 0.12    & 3.44           & 3.43            & PASS & 15.61         & 2.7      & 31.69          \\
+AX6  & AX6-3a   & 25.39     & 400       & 585       & 18 & 0.76 & 10.0  & 0.04    & 0.1     & 4.28           & 2.88            & PASS & 15.61         & 3.2      & 44.08          \\
+AX6  & AX6-1a   & 49.13     & 425       & 585       & 17 & 0.71 & 10.0  & 0.03    & 0.14    & 4.45           & 0.96            & PASS & 15.61         & 2.0      & 85.3           \\
+AX8  & AX8-3a   & 41.3      & 400       & 580       & 17 & 0.75 & 9.0   & 0.03    & 0.14    & 4.38           & 2.22            & PASS & 17.16         & 11.2     & 70.45          \\
+AX8  & AX8-1a   & 66.18     & 450       & 570       & 13 & 0.68 & 8.0   & 0.06    & 0.18    & 4.54           & 2.37            & PASS & 17.16         & 7.0      & 112.88         \\
+AX10 & AX10-1a  & 45.84     & 425       & 585       & 17 & 0.78 & 10.0  & 0.06    & 0.24    & 5.6            & 2.62            & PASS & 14.41         & 4.7      & 80.64          \\
+AX10 & AX10-3a  & 48.01     & 425       & 585       & 17 & 0.69 & 10.0  & 0.04    & 0.24    & 4.02           & 1.56            & PASS & 14.41         & 5.8      & 84.45          \\
+AX10 & AX10-2a  & 48.23     & 425       & 585       & 17 & 0.67 & 10.0  & 0.05    & 0.19    & 5.3            & 1.93            & PASS & 14.41         & 7.8      & 84.84          \\
+AX11 & AX11-1a  & 73.29     & 425       & 560       & 10 & 0.67 & 6.0   & 0.08    & 0.21    & 5.28           & 2.54            & PASS & 12.27         & 5.9      & 131.75         \\
+AX11 & AX11-2a  & 73.29     & 400       & 560       & 11 & 0.65 & 6.0   & 0.07    & 0.21    & 3.94           & 1.51            & PASS & 12.27         & 9.9      & 131.75         \\
+AX11 & AX11-4a  & 73.31     & 500       & 570       & 11 & 0.66 & 8.0   & 0.09    & 0.2     & 1.53           & 3.27            & PASS & 12.27         & 5.6      & 131.78         \\
+AX11 & AX11-6a  & 73.32     & 200       & 562       & 14 & 0.69 & 7.0   & 0.07    & 0.17    & 4.0            & 1.33            & PASS & 12.27         & 3.0      & 131.8          \\
+AX11 & AX11-9a  & 73.34     & 100       & 562       & 15 & 0.67 & 7.0   & 0.06    & 0.17    & 5.34           & 1.13            & PASS & 12.27         & 11.9     & 131.84         \\
+AX11 & AX11-3a  & 73.41     & 400       & 560       & 11 & 0.66 & 6.0   & 0.08    & 0.21    & 3.95           & 2.11            & PASS & 12.27         & 8.2      & 131.96         \\
+AX11 & AX11-5a  & 73.44     & 400       & 566       & 14 & 0.73 & 8.0   & 0.05    & 0.18    & 2.93           & 0.85            & PASS & 12.27         & 1.9      & 132.02         \\
+AX12 & AX12-14a & 30.34     & 475       & 585       & 16 & 0.75 & 10.0  & 0.06    & 0.17    & 1.74           & 0.55            & PASS & 24.54         & 0.9      & 47.18          \\
+AX12 & AX12-1a  & 33.89     & 0         & 580       & 21 & 0.97 & 9.0   & 0.03    & 0.17    & 5.47           & 4.73            & PASS & 24.54         & 10.0     & 52.7           \\
+AX12 & AX12-6a  & 37.18     & 425       & 564       & 12 & 0.69 & 7.0   & 0.08    & 0.22    & 3.64           & 2.05            & PASS & 24.54         & 5.0      & 57.81          \\
+AX12 & AX12-8a  & 38.34     & 475       & 565       & 12 & 0.75 & 8.0   & 0.06    & 0.25    & 1.46           & 2.29            & PASS & 24.54         & 1.6      & 59.62          \\
+AX12 & AX12-4a  & 41.89     & 500       & 585       & 14 & 0.66 & 10.0  & 0.05    & 0.24    & 3.85           & 3.19            & PASS & 24.54         & 11.4     & 65.14          \\
+AX12 & AX12-2a  & 46.94     & 425       & 570       & 14 & 0.7  & 8.0   & 0.05    & 0.24    & 3.39           & 2.38            & PASS & 24.54         & 7.0      & 72.99          \\
+AX13 & AX13-3a  & 71.68     & 100       & 550       & 12 & 0.71 & 5.0   & 0.08    & 0.22    & 5.92           & 1.88            & PASS & 11.25         & 3.4      & 130.08         \\
+AX13 & AX13-4a  & 75.42     & 200       & 555       & 12 & 0.7  & 6.0   & 0.03    & 0.17    & 7.28           & 0.41            & PASS & 11.25         & 3.7      & 136.87         \\
+AX13 & AX13-6a  & 76.8      & 200       & 555       & 12 & 0.7  & 6.0   & 0.06    & 0.22    & 3.82           & 1.53            & PASS & 11.25         & 4.5      & 139.37         \\
+AX13 & AX13-2a  & 76.96     & 300       & 555       & 11 & 0.69 & 6.0   & 0.05    & 0.2     & 7.87           & 2.73            & PASS & 11.25         & 4.1      & 139.66         \\
+AX13 & AX13-9a  & 77.48     & 0         & 564       & 17 & 0.94 & 7.0   & 0.03    & 0.16    & 5.96           & 1.07            & PASS & 11.25         & 2.6      & 140.61         \\
+AX13 & AX13-8a  & 84.83     & 450       & 570       & 13 & 0.81 & 8.0   & 0.02    & 0.24    & 3.69           & 1.82            & PASS & 11.25         & 5.1      & 153.94   \\
+AX16 & AX16-13a & 41.68     & 400       & 570       & 15 & 0.72 & 8.0   & 0.07    & 0.2     & 3.47           & 1.85            & PASS & 22.11         & 1.1      & 66.88          \\
+AX16 & AX16-16a & 42.2      & 400       & 570       & 15 & 0.77 & 8.0   & 0.06    & 0.15    & 3.9            & 1.81            & PASS & 22.11         & 2.0      & 67.72          \\
+AX16 & AX16-14a & 43.57     & 400       & 570       & 15 & 0.71 & 8.0   & 0.06    & 0.13    & 3.25           & 1.79            & PASS & 22.11         & 6.0      & 69.91          \\
+AX16 & AX16-11a & 44.17     & 450       & 570       & 14 & 0.7  & 8.0   & 0.06    & 0.15    & 2.86           & 2.07            & PASS & 22.11         & 4.6      & 70.88          \\
+AX16 & AX16-5a  & 45.75     & 400       & 560       & 11 & 0.69 & 6.0   & 0.09    & 0.22    & 7.98           & 4.32            & PASS & 22.11         & 4.0      & 73.41          \\
+AX16 & AX16-4a  & 45.96     & 425       & 570       & 14 & 0.75 & 8.0   & 0.05    & 0.21    & 5.01           & 1.18            & PASS & 22.11         & 4.2      & 73.75          \\
+AX16 & AX16-2a  & 45.98     & 400       & 564       & 13 & 0.68 & 7.0   & 0.06    & 0.21    & 4.37           & 1.57            & PASS & 22.11         & 4.9      & 73.78          \\
+AX16 & AX16-1a  & 46.25     & 200       & 562       & 14 & 0.85 & 7.0   & 0.07    & 0.23    & 6.04           & 2.87            & PASS & 22.11         & 4.5      & 74.22          \\
+AX16 & AX16-9a  & 48.9      & 475       & 585       & 16 & 0.66 & 10.0  & 0.04    & 0.16    & 2.92           & 1.4             & PASS & 22.11         & 4.3      & 78.47          \\
+AX16 & AX16-10a & 54.02     & 500       & 585       & 15 & 0.66 & 10.0  & 0.04    & 0.18    & 2.78           & 1.15            & PASS & 22.11         & 5.5      & 86.68          \\
+\hline
+\end{tabular}
+\label{tab:PINT_result}
+\end{sidewaystable}
 
 \begin{table}[]
+
 \caption{\footnotesize{Summary statistics for the Q$_{PI}$ quality criteria of \citeA{Biggin2014a}.}}
@@
 AX10 & 3  & 1091.8   & T+     & 1   & 0    & 1   & 1   & 1  & 1   & 0    & 0    & 5   \\
+AX11 & 7  & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\
 AX12 & 6  & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\
-AX16 & 11 & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\
-AX11 & 7  & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\
-AX13 & 7  & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\ \hline
+AX13 & 7  & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\
+AX16 & 11 & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\ \hline
 BD2  & 9  & 1091.7   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\ \hline
@@
 \subsection*{Rock magnetism}
-Rock magnetic data support the interpretation that the successful anorthosite specimens have dominant magnetic carriers with magnetic properties similar to stochiometric, non-interactive, single domain magnetite, whereas those anorthosites that failed the paleointensity result selection and all diabase have more pronounced non-ideal carriers.
+Rock magnetic data support that the successful anorthosite specimens have dominant magnetic carriers with magnetic properties similar to stochiometric, non-interactive, single domain magnetite, whereas those anorthosites that failed the paleointensity result selection and all diabase have more pronounced populations of non-ideal carriers.
 MPMS data show sister specimens to paleointensity specimens typically have dominant magnetic mineralogy of magnetite as can be evidenced through the Verwey transitions (Figure \ref{fig:MPMS}; \citeA{Verwey1939a}).
-However, specimens from sites that yield successful paleointensity results display Verwey transition near 120 Kelvin (K) as expected for stochiometric low-Ti magnetite (\cite{Ozdemir1993a}).
-On the other hand, the diabase specimens and anorthosite specimens that did not yielded accepted paleointensity results typically have suppressed Verwey transition toward lower temperatures (Figure \ref{fig:MPMS}), indicating that the specimens likely have a ferromagnetic mineralogy with relatively high Ti content and/or with variable degrees of oxidation \cite{Ozdemir1993a}.
+However, specimens from sites that yield successful paleointensity results display Verwey transition near 120 Kelvin (K) as expected for stochiometric low-Ti magnetite \cite{Ozdemir1993a}.
+On the other hand, the diabase and anorthosite specimens that did not pass our paleointensity selection typically have suppressed Verwey transition toward lower temperatures (Figure \ref{fig:MPMS}), indicating that the specimens likely have a ferromagnetic mineralogy with variable degrees of oxidation \cite{Ozdemir1993a}.
 
@@
 In the zero-field-cooled (ZFC) experiment, a low-temperature saturation isothermal remanence (LTSIRM) of 2.5 T was applied at 10 K after the specimen cooled in a (near-)zero field.
-In the room-temperature saturation isothermal remanence (RTSIRM) experiment, the sample was pulsed with a 2.5 T field at room temperature ($\sim$300 K) and then cooled to 10 K and warmed back to room temperature in a (near)zero field.
+In the room-temperature saturation isothermal remanence (RTSIRM) experiment, the sample was pulsed with a 2.5 T field at room temperature ($\sim$300 K) and then cooled to 10 K and warmed back to room temperature in a (near-) zero field.
 Specimen AX6-1r is from anorthosite AX6 which passed our paleointensity selection.
-It has a well-defined Verwey transition $\sim$120 K.
-Specimens AX3-2r-ni and BD2-2r have prominent Verwey transition but the transition temperatures are suppressed below $\sim$120 K.
+It has a well-defined Verwey transition $\sim$120 K \cite{Verwey1939a}.
+Specimens AX3-2r-ni and BD2-2r show Verwey transition but the transition temperatures are suppressed below $\sim$120 K.
 Specimen BD2-2r has a consistently higher moment during the zero-field-cooled step than during the field-cooled step.
-This is consistent with the interpretation that multidomian magnetic carriers existed in significant quantity within this sepcimen.}}
+This is consistent with the interpretation that multidomian magnetic carriers exist in significant quantity in this specimen.}}
 \label{fig:MPMS}
@@
 Specimen coercivity spectra further support that our paleointensity selection criteria are successful in selecting experiment results.
-A compilation of median destructive field (MDF) in Figure \ref{fig:coercivity} based on all of our backfield experiments shows that successful anorthosites have distinctly higher average MDF range than others.
-Example single component skew-normal distribution fits of coercivity spectra based on backfield experiments on specimen chips sister to paleointensity specimens show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity distributions around 80 mT, very similar to that of a bulk sample composed of no-interactive single domain magnetite (Figure \ref{fig:coercivity}).
+A compilation of median destructive field (MDF) in Figure \ref{fig:coercivity} based on all of our backfield experiments shows that specimens from anorthosites that pass paleointensity selections have distinctly a higher average MDF range than other anorthosites and all diabase.
+Example single component skew-normal distribution fits of coercivity spectra based on backfield experiments on specimen chips sister to paleointensity specimens show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity distributions around 80 mT (Figure \ref{fig:coercivity}).
 In contrast, specimens from anorthosite that did not pass paleointensity selection, as well as diabase specimens, tend to have coercivity distributions over a significantly lower range ($\sim$30 mT).
@@
 Bottom: Box plots summarize distributions of median destructive field (MDF) values for all anorthosite and diabase specimens that have single-component coercivity unmixing results.
-Both plots show that anorthosite specimens that pass our paleointensity selection criteria have distinctly higher MDF than other anorthosites that do not pass the selection, as well as all diabase specimens.
-The diabase site BD2 which is the only site that yielded specimen results that pass the selection, have similar MDF than other diabase sites.}}
+Both plots show that anorthosite specimens that pass our paleointensity selection criteria have distinctly higher MDF than other anorthosites that do not pass the selection, as well as all diabase specimens.}}
 \label{fig:coercivity}
@@
 \subsection*{Remanence anisotropy and nonlinear TRM acquisition check}
-Significant remanence anisotropy has been documented to exist within anorthositic rocks that form in layered intrusive rocks \cite{Selkin2000a, Feinberg2006a}.
+Significant remanence anisotropy has been documented to exist within anorthositic rocks that form in layered intrusive complexes \cite{Selkin2000a, Feinberg2006a}.
 \cite{Selkin2000a} demonstrated that strong remanence anisotropy associated with the igneous foliation developed within an anorthosite sample from the Stillwater Complex can lead to paleointensity estimates that are different by three times.
 Depending on the orientation of the applied field during in-field steps and that of the magnetic anisotropy axes of a specimen, such strong preferred orientations of petrofabrics can result in significantly biased paleointensity estimates.
-To assess whether our paleointensity estimates are biased by remanence anisotropy, we first calculated the gamma statistic, which is the angular difference between the last pTRM step used for paleointensity determination and the applied field direction.
-The results show all anorthosite specimens used for paleointensity experiment have gamma values ranging from $\sim$0.9\textdegree-$\sim$11.9\textdegree, with a median value of 4.2\textdegree (Table \ref{}).
-These gamma values are similar to those of the Midcontinent Rift volcanics \cite{Sprain2018a} and they show that the Beaver River anorthosite bulk rocks do not have a significant directional remanence anisotropy.
-This result can be further supported by the fact that the recovered paleodirectional data from our anorthosite xenoliths closely match those from the paired Beaver River diabase hosts \cite{Zhang2021b}, the lithology of which is usually considered isotropic in terms of recording remanence magnetization.
-Further, we applied full TRMs to the anorthosite specimens that pass the paleointensity selection criteria after paleointensity experiment.
+To assess whether our paleointensity estimates are biased by remanence anisotropy, we calculated the gamma statistic, which is the angular difference between the last pTRM step used for paleointensity determination and the applied field direction (Table \ref{tab:PINT_result}).
+The results show all anorthosite specimens used for paleointensity experiment have gamma values ranging from $\sim$0.9\textdegree-$\sim$11.9\textdegree, with a median value of 4.2\textdegree (Table \ref{tab:PINT_result}).
+These gamma values are similar to those of the Midcontinent Rift volcanics \cite{Sprain2018a} and they show that the Beaver River anorthosite bulk rocks do not have a significant remanence anisotropy.
+This result can be further supported by the fact that the recovered paleodirectional data from our anorthosite xenoliths closely match those from the paired Beaver River diabase hosts \cite{Zhang2021b}.
+\cite{Selkin2007a} also noted that vortex state magnetic grains and certain single domain particles can produce biased paleointensity estimates due to their acquisition of saturation remanence during laboratory in-field steps.
+According to the numerical modeling of \citeA{Selkin2007a}, thermoremanent magnetization acquired by a randomly-oriented population of elongate single-domian-sized magnetite particles are unlikely to saturate upon our applied field of 30 $\mu$T.
+Therefore, in this study we do not apply any anisotropy or nonlinear TRM correction to our specimens.
+
+% We applied full TRMs to our anorthosite specimens with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axis.
+The results show that our anorthosite specimens did not saturate upon an applied field of 30 $\mu$T and there is thus no need for correcting nonlinear thermal remanence acquisitions.
+% Further, we applied full TRMs to the anorthosite specimens that pass the paleointensity selection criteria after paleointensity experiment.
 With an applied field of 30 $\mu$T oriented toward the specimen at 0, 30, 60, 90, 120, and 150 degrees, the anorthosite specimens usually acquired consistent full TRMs.
@@
 
-\cite{Selkin2007a} also noted that vortex state magnetic grains and certain single domain particles can produce biased paleointensity estimates due to their nonlinearity behavior of thermal remanence acquisition.
-We applied full TRMs to our anorthosite specimens with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axis.
-The results show that our anorthosite specimens did not saturate upon an applied field of 30 $\mu$T and there is thus no need for correcting nonlinear thermal remanence acquisitions.
-
 The reason that the Beaver River anorthosite xenoliths do not display strong remanence anisotropy, nor do they display nonlinear TRM acquisition problem, could be that they formed in a magmatic setting distinct from that of a layered mafic complex, such as the nearby Duluth Complex, where igneous foliation can occur (Figure \ref{fig:Geologic_map}).
 Petrographic images show that an equigranular texture where equidimensional, randomly oriented plagioclase often develop in the anorthosite xenoliths (Figure \ref{fig:Petro_QDM}; \citeA{Zhang2021b}).
-This is in contrast with the interlocking igneous foliation texture within the Anorthositic Series rocks of the \textit{ca.} 1096 Ma Duluth Complex (Figure \ref{fig:Petro_QDM}).
-Further, magnetic imaging also shows the Beaver River anorthosite xenoliths do not develop abundant magnetite and ilmenite needle exsolution within plagioclase crystals as the slowly-cooled anorthositic intrusions, in which exsolved magnetite needles can be the dominate remanence carriers (Figure \ref{fig:Petro_QDM}; \citeA{Scofield1986a, Feinberg2006a}).
-In addition, the aligned magnetic maps and the petrographic images in Figure \ref{fig:Petro_QDM} nicely illustrates that the magnetic dipole fields are exerted by the magnetite needles.
-The fact that all dipole features within a same plagioclase crystal have indistinguishable orientations is consistent with the interpretation of there being a magnetic remanence anisotropy resulting from preferred petrofabric alignment, giving rise to the remanence anisotropy and nonlinear TRM acquisition of layered anorthositic intrusions (Figure \ref{fig:Petro_QDM}).
+This is in contrast to the interlocking igneous foliation texture within the Anorthositic Series rocks of the \textit{ca.} 1096 Ma Duluth Complex (Figure \ref{fig:Petro_QDM}).
+Further, petrographic and magnetic imaging also shows the Beaver River anorthosite xenoliths do not develop abundant exsolved magnetite and ilmenite needles within plagioclase crystals as the slowly-cooled anorthositic intrusions do, in which exsolved magnetite needles can be the dominate remanence carriers (Figure \ref{fig:Petro_QDM}; \citeA{Scofield1986a, Feinberg2006a}).
+The fact that all dipole features within a same Duluth Complex plagioclase crystal have indistinguishable orientations is consistent with the interpretation of there being a magnetic remanence anisotropy resulting from preferred petrofabric alignment, which could give rise to the remanence anisotropy and nonlinear TRM acquisition of layered anorthositic intrusions (Figure \ref{fig:Petro_QDM}).
 
@@
 \subsection*{Late Mesoproterozoic geomagnetic field and the Precambrian observation records}
-The anorthosites are also grouped by their geographical vicinity \ref{fig:PINT_plot}.
-The site-level results are shown in Table \ref{tab:PINT_result}.
-Anorthosite sites AX10, 11, 12, 13, 16 yielded very high within-site coherence (Figure \ref{fig:Arai_plots, fig:PINT_plot}), which suggest high-confidence site-level paleointensity estimates.
-Clearly there are two populations of paleointensity results: a group of high estimates from anorthosite site 11 and 13 which have very high within-site consistency and are in agreement with the results from BRD site 2; and another group of low estimates from anorthosite site 6, 8, 10, 12, and 16, which also have tight agreement across sites.
-
-\subsection*{Bias Corrected Estimation of Paleointensity}
-While using a set of selection criteria for excluding specimens with non-ideal paleointensity results is one common way of performing data analyses, the selection criteria themselves are often user-determined according to interpretations of certain groups of magnetic mineralogy on a case-by-case basis.
-The operation of performing a binary selection (accept or reject) on specimen results thus has an underlying assumption that the magnetic grain assemblage of specimens that pass the selection have a determined type of magnetic mineralogy.
-However, natural samples often have complex assemblage of magnetic carriers that span a range of composition, grain sizes and shapes.
-Thus, imposing a threshold value on certain parameters based on the paleointensity results can introduce bias from the analyzer.
-Alternatively, \citeA{Cych2021a} developed a Bias Corrected Estimation of Paleointensity (BiCEP) method that can eliminate biased introduced by using a set of arbitrarily defined selection criteria by taking all specimens into account for site-level paleointensity estimates.
-The method assumes that the paleointensity recorded by each specimen is biased away from a true value by an amount that is dependent on a single metric of nonlinearity (the curvature parameter $|\vec{k}|$; \citeA{Paterson2011a}) on the Arai plot.
-Although this method has the advantage of mitigating paleointensity estimate bias, it does not account for specimens that experience detrimental magnetic mineralogy alteration during heating steps unless such specimens are omitted before the analyses.
-
-We apply the BiCEP method to all sites in this study and the cooling rate corrected site-level paleointensity estimates are shown in Figure \ref{fig:BiCEP}.
-For the anorthosite sites (e.g.
-AX11) that show good specimen-level consistency (Figure \ref{fig:PINT_cooling_corrected}) and straight specimen Arai plots (Figure \ref{fig:IZZI_examples}), the paleointensity estimates derived from the BiCEP method using all specimen data is indistinguishable with that derived from the site average value based on selected specimens (Figure \ref{fig:PINT_cooling_corrected}).
-For anorthosites that show rather large specimen-level dispersion (e.g.
-AX8), or diabase sites that have double-slop behaviors and significant MD component, the paleointensity estimates using the BiCEP method tend to have broader uncertainty ranges and the median values are lower than those from anorthosites that typically have more ideal paleointensity behaviors (Figure \ref{fig:PINT_cooling_corrected}).
-Overall, site mean value of the mean probability distribution of the BiCEP method results is consistent with the weighted mean value of the selection criteria-based estimates.
-In addition, the majority of the anorthosite sites give high probability estimate for there being a geomagnetic field value \textit{ca.} 1092 Ma at a paleolatitude of $\sim$30\textdegree similar to that of today.
-This result is consistent with the interpretation that the anorthosite xenoliths record a high geomagnetic field during the late Mesoproterozoic.
-
-\begin{figure}[h!]
-\noindent\includegraphics[width=\textwidth]{code/code_output/AX_BD_BiCEP.pdf}
-\centering
-\caption{\footnotesize{Violin plot of paleointensity estimates from the anorthosite and diabase using the bias corrected estimation of paleointensity (BiCEP) method \cite{Cych2021a}.
-The results are grouped by individual cooling units following \citeA{Zhang2021b}.
-All estimates are corrected for cooling rate by a factor of 0.74.
-}}
-\label{fig:BiCEP}
-\end{figure}
 
@@
 
-% Together with previously published data, the Midcontinent Rift rocks are recording 1.1 Ga earth geomagnetic field that is stronger than the mean of the past 300 Ma (Fig. \ref{}).
+% Together with previously published data, the Midcontinent Rift rocks are recording 1.1 Ga earth geomagnetic field that is stronger than the mean of the past 300 Ma (Figure \ref{}).
 
@@
 
-% At the IRM, I was seeking the answer to this question with the help of the vibrating sample magnetometer (VSM) systems, including a newly installed Lake Shore VSM which greatly helped improve measurement resolution on samples with weak magnetizations (Fig.
-\ref{fig:backfield}).
-Backfield demagnetization experiments were conducted and used to develop coercivity spectra (Fig.
-\ref{fig:backfield}).
+% At the IRM, I was seeking the answer to this question with the help of the vibrating sample magnetometer (VSM) systems, including a newly installed Lake Shore VSM which greatly helped improve measurement resolution on samples with weak magnetizations (Figure \ref{fig:backfield}).
+Backfield demagnetization experiments were conducted and used to develop coercivity spectra (Figure \ref{fig:backfield}).
 The spectra were subsequently modeled to fit for the distributions of different  populations of magnetic particles using similar procedure as \cite{Maxbauer2016a}.
-Most spectra can be well approximated with models with one component or two components with overlapping coercivity ranges (Fig.
-\ref{fig:backfield}).
+Most spectra can be well approximated with models with one component or two components with overlapping coercivity ranges (Figure \ref{fig:backfield}).
 The dominance of single-component distributions from the unmixed coercivity spectra likely suggests minimal alterations or formation of secondary magnetic mineral within the samples.
-By compiling the values of median destructive field (MDF) from all measured specimens and categorizing them in terms of their sister specimens' paleointensity results, I found that the anorthosites samples that pass paleointensity selection criteria have distinctively higher MDF values than those did not (Fig.
-\ref{fig:backfield}), and all diabase specimens have low MDF values like the non-ideal anorthosites (Fig.
-\ref{fig:backfield}).
+By compiling the values of median destructive field (MDF) from all measured specimens and categorizing them in terms of their sister specimens' paleointensity results, I found that the anorthosites samples that pass paleointensity selection criteria have distinctively higher MDF values than those did not (Figure \ref{fig:backfield}), and all diabase specimens have low MDF values like the non-ideal anorthosites (Figure \ref{fig:backfield}).
 The higher MDFs ($\sim$60 mT) of the anorthosites are similar to what is typical of stoichiometric, single-domain magnetite grains which are favored for paleointensity experiments.
-Furthermore, the low MDFs displayed in other specimens could be linked to a dominant population of large interstitial Fe-Ti oxides that are more likely to form multi-domain grains, which have been observed in both the anorthosite and diabase (Fig.
-\ref{fig:field}).
+Furthermore, the low MDFs displayed in other specimens could be linked to a dominant population of large interstitial Fe-Ti oxides that are more likely to form multi-domain grains, which have been observed in both the anorthosite and diabase (Figure \ref{fig:field}).
 In addition, preliminary tests on single anorthosite crystals might suggest that not all single crystals are necessarily better for paleointensity research when comparing to bulk samples, given their similar low MDF to the failed diabase and anorthosite specimens.

### 2021-10-11 manuscript/supporting_info.tex

\documentclass[11pt,letterpaper]{article}
\usepackage{xcolor}
\usepackage{textcomp,marvosym}
\usepackage{amsmath,amssymb}
\usepackage[left]{lineno}
\usepackage{changepage}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{sidecap}
\usepackage{pdfpages}
\usepackage{longtable}
\usepackage{url}
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
%\doublespacing

\raggedright
\textwidth = 6.5 in
\textheight = 8.5 in
\oddsidemargin = 0.0 in
\evensidemargin = 0.0 in
\topmargin = -0.5 in
\headheight = 0.0 in
\headsep = 0.5 in
\parskip = 0.0 in
\parindent = 0.2 in

\begin{document}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\subsection*{Supporting Information for ``Synchronous emplacement of the anorthosite xenolith-bearing Beaver River diabase and one of the largest lava flows on Earth"}
Yiming Zhang, Nicholas L. Swanson-Hysell, Mark D. Schmitz, James D. Miller Jr., Margaret S. Avery

\subsection*{Bias Corrected Estimation of Paleointensity}
While using a set of selection criteria for excluding specimens with non-ideal paleointensity results is one common way of performing data analyses, the selection criteria themselves are often user-determined according to interpretations of certain groups of magnetic mineralogy on a case-by-case basis. The operation of performing a binary selection (accept or reject) on specimen results thus has an underlying assumption that the magnetic grain assemblage of specimens that pass the selection have a determined type of magnetic mineralogy. However, natural samples often have complex assemblage of magnetic carriers that span a range of composition, grain sizes and shapes. Thus, imposing a threshold value on certain parameters based on the paleointensity results can introduce bias from the analyzer. Alternatively, \citeA{Cych2021a} developed a Bias Corrected Estimation of Paleointensity (BiCEP) method that can eliminate biased introduced by using a set of arbitrarily defined selection criteria by taking all specimens into account for site-level paleointensity estimates. The method assumes that the paleointensity recorded by each specimen is biased away from a true value by an amount that is dependent on a single metric of nonlinearity (the curvature parameter $|\vec{k}|$; \citeA{Paterson2011a}) on the Arai plot. Although this method has the advantage of mitigating paleointensity estimate bias, it does not account for specimens that experience detrimental magnetic mineralogy alteration during heating steps unless such specimens are omitted before the analyses.

We apply the BiCEP method to all sites in this study and the cooling rate corrected site-level paleointensity estimates are shown in Figure \ref{fig:BiCEP}. For the anorthosite sites (e.g. AX11) that show good specimen-level consistency (Figure \ref{fig:PINT_cooling_corrected}) and straight specimen Arai plots (Figure \ref{fig:IZZI_examples}), the paleointensity estimates derived from the BiCEP method using all specimen data is indistinguishable with that derived from the site average value based on selected specimens (Figure \ref{fig:PINT_cooling_corrected}). For anorthosites that show rather large specimen-level dispersion (e.g. AX8), or diabase sites that have double-slop behaviors and significant MD component, the paleointensity estimates using the BiCEP method tend to have broader uncertainty ranges and the median values are lower than those from anorthosites that typically have more ideal paleointensity behaviors (Figure \ref{fig:PINT_cooling_corrected}). Overall, site mean value of the mean probability distribution of the BiCEP method results is consistent with the weighted mean value of the selection criteria-based estimates. In addition, the majority of the anorthosite sites give high probability estimate for there being a geomagnetic field value \textit{ca.} 1092 Ma at a paleolatitude of $\sim$30\textdegree similar to that of today. This result is consistent with the interpretation that the anorthosite xenoliths record a high geomagnetic field during the late Mesoproterozoic.

\begin{figure}[h!]
\noindent\includegraphics[width=\textwidth]{code/code_output/AX_BD_BiCEP.pdf}
\centering
\caption{\footnotesize{Violin plot of paleointensity estimates from the anorthosite and diabase using the bias corrected estimation of paleointensity (BiCEP) method \cite{Cych2021a}. The results are grouped by individual cooling units following \citeA{Zhang2021b}. All estimates are corrected for cooling rate by a factor of 0.74. }}
\label{fig:BiCEP}
\end{figure}

\clearpage

\bibliographystyle{gsabull}
\bibliography{YZ_ref}

\end{document}

### 2021-10-12 manuscript/Manuscript.tex
manuscript update

@@
 
-40 from a total of 86 anorthosite specimens and 9 out of a total of 69 diabase specimens passed our paleointensity selection criteria, yielding 7 anorthosite sites and 1 diabase sites with site-level paleointensity results that pass paleointensity selections.
+40 from a total of 86 anorthosite specimens and 0 out of a total of 69 diabase specimens passed our paleointensity selection criteria, yielding 7 anorthosite sites and no diabase sites with site-level paleointensity results that pass paleointensity selections.
 Example NRM/TRM plots (Arai plots;
@@
 Although the BD2 specimens' lower temperature fraction of the Arai plots are straight andthe pTRM checks are consistent, lower temperature slopes from BD2 give consistently higher paleointensity estimates than those from site AX6, AX8, and AX10, which are anorthosite xenoliths included within BD2 (Figure \ref{fig:Geologic_map}; \citeA{Zhang2021a}).
-This higher estimate from the diabase lithology could be attributed to the systematic bias of using the lower temperature slope of the Arai plots demonstrated in Figure \ref{fig:IZZI_examples}.
+This higher estimate from the diabase lithology could be attributed to the systematic bias of using the lower temperature slope of the Arai plots demonstrated in Figure \ref{fig:IZZI_examples} and specimens from site BD2 all fail the criteria of having a maximum magnetizatiom gap between adjacent steps of 25\% of the total NRM.
 Thus we only accept the single-slope anorthosite specimens results for further analyses.
@@
 Overall, the consistent specimen-level behavior makes the successful results from the anorthosites of high confidence.
+The paleointensity quality index (Q$_{PI}$) for the anorthosite sites are all above 3 and the summary statistics are shown in Table \ref{tab:QPI}.
 
@@
 $\gamma$ is the gamma statistic that measures the angle between the last pTRM step used for paleointensity determination and the applied field direction.
-V(A)DM is the virtual (axial) dipole moment reported in $10^{21}$Am$^2$ ((ZAm$^2$)).
+V(A)DM is the virtual (axial) dipole moment reported in $10^{21}$Am$^2$ (ZAm$^2$).
 }}
@@
 AX6  & 3  & 1091.8   & T+     & 1   & 0    & 1   & 1   & 1  & 1   & 0    & 0    & 5   \\
-AX8  & 3  & 1091.8   & T+     & 1   & 0    & 1   & 1   & 1  & 1   & 0    & 0    & 5   \\
+AX8  & 2  & 1091.8   & T+     & 1   & 0    & 1   & 1   & 1  & 1   & 0    & 0    & 5   \\
 AX10 & 3  & 1091.8   & T+     & 1   & 0    & 1   & 1   & 1  & 1   & 0    & 0    & 5   \\
@@
 AX16 & 11 & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\ \hline
-BD2  & 9  & 1091.7   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\ \hline
 \end{tabular}
@@
 
-\subsection*{Late Mesoproterozoic geomagnetic field and the Precambrian observation records}
+\subsection*{Late Mesoproterozoic geodynamo }
+To best characterize the geomagnetic axial dipole field intensity during a certain time period requires that a paleointensity dataset to cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged and the geocentric axial dipole hypothesis (GAD) can be assumed.
+Thermal modeling results from \citeA{Zhang2021b} suggest that the anorthosite xenoliths were likely heated to tholeiitic magma temperature and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a few thousand years.
+In addition, the closely located sites AX12 and AX16 have very similar paleointensity estimates that are distinct from those of sites AX11 and AX13 which were emplaced $\sim$10 km away (Figure \ref{fig:Geologic_map}).
+The consistency within small regions but variations between regions could be the result of differential cooling history due to heterogeneity of the timing and depths of the Beaver River diabase intrusions, causing the anorthosites to capture paleosecular variations \textit{ca.} 1092 Ma.
+Overall, none of the anorthosite xenoliths record a V(A)DM value smaller than 40 ZAm$^2$ (Figure \ref{fig:PINT_cooling_corrected}).
+
+Such a high estimate for the late Mesoproterozoic may invoke more careful examination of the geodynamo evolution through the Proterozoic and the timing for the inner core nucleation.
+Paleomagnetic observational records near the Ediacaran-Cambrian boundary have been interpreted to be consistent with a trend of decreasing axial dipole strength toward a very low value \cite{Lloyd2021a, Bono2019a}, followed by a rapid recovery (a few tens of million years) toward higher values soon after \cite{Thallner2021a}.
+Such a transition near this time can be consistent with the dynamo numerical simulations which adopt a rather high thermal conductivity of the core \cite{Davies2021a}.
+In particular, some models have suggested that the kick-start of rigorous compositional convection between the inner core and the outer core soon after the onset of inner core nucleation can result in almost an order of magnitude increase in Earth's axial dipole moment \cite{Davies2021a}.
+
+Although more recent data
+
+using paleomagnetic observational records is guided by the predictions from numerical simulations that a cooling liquid core with only thermal convection power to drive a dynamo would experience a progressive decay in its magnetic dipole field strength, and experience a rapid recover after the onset of inner core nucleation \cite{Davies2021a}.
+Previous studies has pointed to the Ediacaran-Cambrian boundary for such a weakening field behavior and recent studies has claim to found a rapid increase after the single crystal time which fits in the numerical models.
+However, such a recover may not be unique given the strong geodynamo estimate from the 1.1 Ga Midcontinent Rift.
+One could point to the increase from low values from the ca.
+1140 Ma abitibi dikes to the ca.
+1092 Ma anorthosites to be the result of recovery; another line of disagreement is between the Bangemall sills and the low field from Sept Ils.
+MCR being an anomaly?
+or the time scale we are looking at is too long?
+
+Regardless these several lines of increase/decrease trend needs further explanations/considerations.
+
+use a combined method of direction and intensity research for further examinations.
+Re-evaluation of the spatial and temporal variability captured by the existing Precambrian database.
+More data are needed for the Precambrian in order to better characterize the geodynamo evolution.
 
 \begin{figure}[h!]
-\noindent\includegraphics[width=\textwidth]{MCR_VADM.pdf}
+\noindent\includegraphics[width=\textwidth]{PINT_compilation.pdf}
 \centering
-\caption{\footnotesize{Compilation of calculated virtual axial dipole moment (VADM) values from previous studies of the Midcontinent Rift rocks \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} and from the anorthosites that pass the selection criteria in this study.
-The median VADM value of all MCR rocks is higher than the median value of the past 300 myr.
-}}
+\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment (VADM) values from the PINT database (v.2015.05; \url{http://earth.liv.ac.uk/ pint/}; \citeA{Biggin2015a}) with additional Mesoproterozoic data from \citeA{Sprain2018a, Kodama2019a}, as well as those from the Neoproterozoic \cite{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}.
+Paleointensity estimates from \citeA{Pesonen1983a} and \citeA{Kulakov2013a} are not included in the compilation due to the specimen-level double-slope behavior as discussed in the text.
+Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \citeA{Bono2019a} which is based on an interpretation of there being a progressive decay of the geodynamo through the Proterozoic.}}
 \label{fig:MCR_VADM}
 \end{figure}
-
-considering \citeA{Nagy2021a} developed a new cooling rate correction model for uniaxial single domain magnetite and suggest that the model of \citeA{Halgedahl1980a} tend to over-correct for cooling rates and lead to underestimating ancient fields.
-If our absolute paleointensity values are an underestimating,
-
-In addition, \cite{Cych2021a} developed a Bayesian probability method to estimate paleointensity - independent of the selection criteria, the Bayesian method fits for the curvature and generate probability distribution that guides credible interval for the paleoitnensity at a site level.
-The result distribution of Markov Chain Monte Carlo resampling are shown as violin plots in Figure \ref{}.
-The cooling rate corrected distributions show similar results as the frequentists' estimates based on the selection criteria derived results.
-
-Address the spacial groups and using direction and intensity to address the issue of secular variation.
-
-overall the PINT results are of high quality and the Qpi index average to be 6 to 7.
 

### 2021-10-12 manuscript/Manuscript.tex
update on manuscript

@@
 \item High VADM of the geomagnetic field is inconsistent with previous interpretation of Proterozoic long term decay
-\item A strong geodynamo like today likely persisted for about 25 myr during the Midcontinent Rift development
+\item A strong geodynamo like today likely persisted for about 25 Myr during the Midcontinent Rift development
 \end{keypoints}
@@
 It has been interpreted that these findings are consistent with the results of geodynamo models using high core thermal conductivity values that suggest the inner core nucleation age being near the Ediacaran-Cambrian boundary \cite{Driscoll2016a, Davies2021a}.
-\citeA{Bono2019a} based on a polynomial fit through filtered Precambrian paleointensity database to suggest the selected observational records show that the geodynamo power progressively decreased through the Proterozoic (Figure \ref{fig:MCR_VADM}).
+\citeA{Bono2019a} based on a polynomial fit through filtered Precambrian paleointensity database to suggest the selected observational records show that the geodynamo power progressively decreased through the Proterozoic (Figure \ref{fig:PINT_compilation}).
 That study also recovered very low paleointensity values from \textit{ca.} 565 Ma single silicate crystals and interpreted them to be reflecting that the geodynamo reached its lowest near the Ediacaran-Cambrian boundary.
@@
 
-However, previous paleointensity data from intrusive and extrusive rock units in the Mesoproterozoic Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record Earth's virtual axial dipole moment (VADM) values similar to that of today, which are higher than the prediction of the long term decay curve (Figure \ref{fig:MCR_VADM}).
+However, previous paleointensity data from intrusive and extrusive rock units in the Mesoproterozoic Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record Earth's virtual axial dipole moment (VADM) values similar to that of today, which are higher than the prediction of the long term decay curve (Figure \ref{fig:PINT_compilation}).
 \citeA{Pesonen1983b} obtained an average virtual dipole moment of 93 ZAm$^2$ after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using both extrusive and intrusive rocks, including the \textit{ca.} 1108-1105 Ma Osler Volcanics, the Maimainse Point volcanics, intrusive Logan Sills, Thunder Bay dikes, the Baraga and Marquette dikes, as well as lava flow of the \textit{ca.} 1096 Ma North Shore Volcanic Group.
@@
 Magnetic imaging shows that the anorthosite specimens to have dominant magnetic carriers interstitial to and within plagioclase crystals and they do not display strong preferred orientations.
-With more single-slope, high quality paleointensity data, we seek to compare paleointensity records from the Midcontinent Rift intrusives with those from the extrusives; to test the hypothesis that the Earth's geomagnetic dipole strength remained at high values throughout the Midcontinent Rift; and to further enrich the reliable paleointensity records during the $\sim$25 myr of Midcontinent Rift magmatic activity.
+With more single-slope, high quality paleointensity data, we seek to compare paleointensity records from the Midcontinent Rift intrusives with those from the extrusives; to test the hypothesis that the Earth's geomagnetic dipole strength remained at high values throughout the Midcontinent Rift; and to further enrich the reliable paleointensity records during the $\sim$25 Myr of Midcontinent Rift magmatic activity.
 
@@
 
-Despite that the anorthosite xenoliths of the Beaver River diabase were entrained as xenoliths and emplaced only $\sim$4 myr after the formation of the Anorthositic Series rocks of the Duluth Complex when massive plagioclase mushes were generated, $^{206}$Pb/$^{238}$U zircon geochronology data and zircon rare earth element diffusion modeling results suggest it is the most probable that the Beaver River diabase anorthosite xenoliths are entrained cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
+Despite that the anorthosite xenoliths of the Beaver River diabase were entrained as xenoliths and emplaced only $\sim$4 Myr after the formation of the Anorthositic Series rocks of the Duluth Complex when massive plagioclase mushes were generated, $^{206}$Pb/$^{238}$U zircon geochronology data and zircon rare earth element diffusion modeling results suggest it is the most probable that the Beaver River diabase anorthosite xenoliths are entrained cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
 Moreover, the anorthosite xenoliths of the \textit{ca.} 1092 Ma Beaver River diabase has distinct mineralogy from that of the Anorthositic Series rocks of the Duluth Complex.
@@
 The consistency within small regions but variations between regions could be the result of differential cooling history due to heterogeneity of the timing and depths of the Beaver River diabase intrusions, causing the anorthosites to capture paleosecular variations \textit{ca.} 1092 Ma.
-Overall, none of the anorthosite xenoliths record a V(A)DM value smaller than 40 ZAm$^2$ (Figure \ref{fig:PINT_cooling_corrected}).
+Overall, none of the anorthosite xenoliths record a V(A)DM value smaller than 40 ZAm$^2$ (Figure \ref{fig:PINT_compilation}).
 
-Such a high estimate for the late Mesoproterozoic may invoke more careful examination of the geodynamo evolution through the Proterozoic and the timing for the inner core nucleation.
 Paleomagnetic observational records near the Ediacaran-Cambrian boundary have been interpreted to be consistent with a trend of decreasing axial dipole strength toward a very low value \cite{Lloyd2021a, Bono2019a}, followed by a rapid recovery (a few tens of million years) toward higher values soon after \cite{Thallner2021a}.
-Such a transition near this time can be consistent with the dynamo numerical simulations which adopt a rather high thermal conductivity of the core \cite{Davies2021a}.
-In particular, some models have suggested that the kick-start of rigorous compositional convection between the inner core and the outer core soon after the onset of inner core nucleation can result in almost an order of magnitude increase in Earth's axial dipole moment \cite{Davies2021a}.
+Such a decrease-increase transition can be consistent with the dynamo numerical simulations which adopt a rather high thermal conductivity of the core and projects an inner core birth near this time \cite{Gomi2013a, Davies2015a, Labrosse2015a, Ohta2016a}.
+In particular, some models have suggested that the kick-start of rigorous compositional convection between the inner core and the outer core soon after the onset of inner core nucleation can result in almost an order of magnitude increase in Earth's axial dipole moment \cite[e.g.]{Davies2021a}.
+The enriching observational records from the Ediacaran and Cambrian have been suggested to indicate that a geodynamo driven solely by thermal convection was waning to its brink of collapse near the end of Ediacaran \cite{Bono2019a, Thallner2021b} and was recovering to ward modern day strengths soon after \cite{Thallner2021a}.
 
-Although more recent data
+However, the updated paleointensity records from the Midcontinent Rift together with previous results from other Mesoproterozoic rocks suggest that the a dramatic increase in geomagnetic dipole moment similar in scale to that modeled for after the onset of inner core nucleation may not be unique to the Edicaran-Cambrian boundary before 500 Ma (Figure \ref{fig:PINT_compilation}).
+The \textit{ca.} 1140 Ma Abitibi dikes \cite{Macouin2003a} record an average virtual dipole moment of $\sim$13 ZAm$^2$, similar to the low values of $\sim$7 ZAm$^2$ obtained from the single silicate crystals of the \textit{ca.} 565 Ma Sept-\hat{I}les intrusions \cite{Bono2019a}.
+However, the relatively high virtual dipole moments of $\sim$72 ZAm$^2$ from the \textit{ca.} 1105 Ma Osler Volcanic Group \cite{Sprain2018a} and $\sim$82 ZAm$^2$ from the \textit{ca.} 1092 Ma Beaver River anorthosite xenoliths (Figure \ref{fig:PINT_compilation} are about six times the values from the \textit{ca.} 1140 Ma Abitibi dikes.
+In addition, the increase in absolute dipole moment estimates recorded by the Midcontinent Rift rocks are much higher than that in the Ediacaran (Figure \ref{fig:PINT_compilation}).
 
-using paleomagnetic observational records is guided by the predictions from numerical simulations that a cooling liquid core with only thermal convection power to drive a dynamo would experience a progressive decay in its magnetic dipole field strength, and experience a rapid recover after the onset of inner core nucleation \cite{Davies2021a}.
+Furthermore, the high geomagnetic field estimates from the Midcontinent Rift volcanics and intrusives are not the only suite of rocks that bear records at odds with the hypothesis of a progressive decay trend during the Proterozoic \cite{Biggin2009a, Bono2019a}.
+The $\sim$60 ZAm$^2$ dipole moment estimate from the \textit{ca.} 755 Ma Mundine Well dikes are discordant with the long-term decrease trend \cite{Bono2019a} and is about three times the estimates from the \textit{ca.} 720 Ma dolerite dikes of the Franklin Large Igneous Province \cite{Lloyd2021a}.
+Similarly, an apparent rapid decrease in geomagnetic dipole moment from \textit{ca.} 1092 Ma to \textit{ca.} 1070 Ma can be evidenced by the contrasting high values from the Midcontinent Rift anorthosite xenoliths and the low values from the Mundine Well dikes (Figure \ref{PINT_Compilation}).
+However, a sparse record in a short time period may not sufficiently average out paleosecular variations.
+Given that a five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years (5 and 95 percentiles), when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed, we cannot conclude on whether an apparent rise or decay within a narrow time period ($\sim$35 Ma) during the Proterozoic is associated with transitions in the core fluid regime or is a result of short-term variations of the geomagnetic field that may be common through much of the evolution of Earth's core.
+In addition, the observations of geomagnetic dipole moment fluctuations in more recent time scales may suggest that evaluations of long-term evolution of the Precambrian paleointensity records should be accompanied by development of higher-resolution data in more short-term scales.
+
+Overall, the existing absolute paleointensity records in the PINT database remains sparse in terms of spatial and temporal coverage as compared to more geologically recent times.
+(Figure \ref{PINT_compilation}).
+More data are needed in order to better characterize the evolution of Earth's dynamo and thermal evolution.
+
+% 1.
+whether septils was recording the time averaged field or the low values are from the hyperactive state which is another signal prior to the onset of inner core?
+indeed it is interesting that no high values from the Ediacaran has been recovered
+
+% 2. where is the recovery, how long was the recovery, how strong was the recovery is at question for he Ediacaran boundary.
+
+% 3. How should one explain the Mundine Well dikes high values?
+
+% 4. How should one explain the Bungemall sills low values in 1070 Ma following the high values of the MCR?
+
+% 5. We need more spatial and temporal coverage of the Precambrian observational data.
+
+% using paleomagnetic observational records is guided by the predictions from numerical simulations that a cooling liquid core with only thermal convection power to drive a dynamo would experience a progressive decay in its magnetic dipole field strength, and experience a rapid recover after the onset of inner core nucleation \cite{Davies2021a}.
 Previous studies has pointed to the Ediacaran-Cambrian boundary for such a weakening field behavior and recent studies has claim to found a rapid increase after the single crystal time which fits in the numerical models.
@@
 
-Regardless these several lines of increase/decrease trend needs further explanations/considerations.
-
-use a combined method of direction and intensity research for further examinations.
+% use a combined method of direction and intensity research for further examinations.
 Re-evaluation of the spatial and temporal variability captured by the existing Precambrian database.
@@
 Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \citeA{Bono2019a} which is based on an interpretation of there being a progressive decay of the geodynamo through the Proterozoic.}}
-\label{fig:MCR_VADM}
+\label{fig:PINT_compilation}
 \end{figure}
-
-The calculated paleomagnetic VADM values are shown in Figure \ref{} in context of the compilation of all Midcontinent Rift results.
-highlight the high quality of the results and the consistency of MCR rocks' high geomagnetic field values even considering the uncertainties of previous data.
-Supporting that the Mesoproterozoic likely experienced a period of persistent high geomagnetic strength in addition to the protracted magmatism of in the formation of the massive intracontinental rifting.
 
@@
 Overall range of correction factor can be 1.1-1.4, in any case consistent with a high field strength
-
-% that this dataset also could have suffered from problem of averaging out secular variations, given that the mean pmag directions deviate from what is expected from the sythesized APWP, nevertheless the variations in estimate absolute paleointensity lead us to interpret that albeit close, the they intensity data were acquired at distinctly different times that are sufficiently far apart such that the surface geomagnetic field had already changed.
 
@@
 
+% Regardless, a strong geomagnetic field has been recorded by the \textit{ca.} 1092 Ma Beaver River anorthosite xenoliths and a strong geodynamo might have persisted through the protracted Midcontinent Rift magmatism.
+
 \section*{Conclusion}
-high success rate comparing to volcanics (more than half success sites), producing good consistent site level results, and across sites report present-day like field like today.
 
-AF pre-treatment is helpful in that it increase the specimens/sites passing the selecting criteria, suggesting that the method likely helped improve the MD ferromagnetic carriers.
+The paucity of Precambrian paleointensity estimates, as well as non-ideal results within the existing database, present challenges when seeking to interpret long-term trends in the strength of the geomagnetic field.
+The Midcontinent Rift of North America presents an opportunity to develop robust paleointensity estimates spanning a $\sim$25 Myr interval of the late Mesoproterozoic.
+Rock magnetic data support that the \textit{ca.} 1092 Ma anorthosite xenoliths of the Midcontinent Rift diabase can have dominantly high coercivity magnetic mineralogy with minimal anisotropic or nonlinear thermal remanence acquisition, producing high quality paleointensity results.
+Their overall high estimates of the late Mesoproterozoic virtual dipole moment together with previous Midcontinent Rift records suggest that a strong geomagnetic field like today's might have persisted through the rifting.
 
-Rock mag and petrographic data might suggest differences in terms of the value of peak field on coercivity spectra between those passed and those didn't.
-
-IRM measured before and after an acid dunk in --- concentration acid on specimens after IZZI experiment show a difference (significant drop) maybe? between specs that passed and those did not.
-
-geographical association
-
-anisotropy of full TRM experiment on accepted specs show low degree of bulk sample scale full TRM anisotropy, with a range of ---.
-Cooling rate correction based on \cite{Halgedahl1980a} suggest a lab over estimate of B ancient of 35 \%, assuming constant cooling rate in natural condition.
-The cooling rate corrected field is $\sim$40 $\mu$T which is similar to today's field strength at similar latitude.
+The new data in context of the current Precambrian paleointensity database suggest that a strong increase in geomagnetic dipole moment which is thought to be indicative of the onset of compositional convection caused by the inner core nucleation may not be unique during the Precambrian.
+More spatial and temporal coverage of the observational records is needed for further analyses of the long-term evolution of the geodynamo.
 

### 2021-10-12 manuscript/Manuscript.tex

@@
 However, the updated paleointensity records from the Midcontinent Rift together with previous results from other Mesoproterozoic rocks suggest that the a dramatic increase in geomagnetic dipole moment similar in scale to that modeled for after the onset of inner core nucleation may not be unique to the Edicaran-Cambrian boundary before 500 Ma (Figure \ref{fig:PINT_compilation}).
-The \textit{ca.} 1140 Ma Abitibi dikes \cite{Macouin2003a} record an average virtual dipole moment of $\sim$13 ZAm$^2$, similar to the low values of $\sim$7 ZAm$^2$ obtained from the single silicate crystals of the \textit{ca.} 565 Ma Sept-\hat{I}les intrusions \cite{Bono2019a}.
+The \textit{ca.} 1140 Ma Abitibi dikes \cite{Macouin2003a} record an average virtual dipole moment of $\sim$13 ZAm$^2$, similar to the low values of $\sim$7 ZAm$^2$ obtained from the single silicate crystals of the \textit{ca.} 565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a}.
 However, the relatively high virtual dipole moments of $\sim$72 ZAm$^2$ from the \textit{ca.} 1105 Ma Osler Volcanic Group \cite{Sprain2018a} and $\sim$82 ZAm$^2$ from the \textit{ca.} 1092 Ma Beaver River anorthosite xenoliths (Figure \ref{fig:PINT_compilation} are about six times the values from the \textit{ca.} 1140 Ma Abitibi dikes.
@@
 The $\sim$60 ZAm$^2$ dipole moment estimate from the \textit{ca.} 755 Ma Mundine Well dikes are discordant with the long-term decrease trend \cite{Bono2019a} and is about three times the estimates from the \textit{ca.} 720 Ma dolerite dikes of the Franklin Large Igneous Province \cite{Lloyd2021a}.
-Similarly, an apparent rapid decrease in geomagnetic dipole moment from \textit{ca.} 1092 Ma to \textit{ca.} 1070 Ma can be evidenced by the contrasting high values from the Midcontinent Rift anorthosite xenoliths and the low values from the Mundine Well dikes (Figure \ref{PINT_Compilation}).
+Similarly, an apparent rapid decrease in geomagnetic dipole moment from \textit{ca.} 1092 Ma to \textit{ca.} 1070 Ma can be evidenced by the contrasting high values from the Midcontinent Rift anorthosite xenoliths and the low values from the Mundine Well dikes (Figure \ref{fig:PINT_compilation}).
 However, a sparse record in a short time period may not sufficiently average out paleosecular variations.
@@
 Overall, the existing absolute paleointensity records in the PINT database remains sparse in terms of spatial and temporal coverage as compared to more geologically recent times.
-(Figure \ref{PINT_compilation}).
+(Figure \ref{fig:PINT_compilation}).
 More data are needed in order to better characterize the evolution of Earth's dynamo and thermal evolution.
@@
 More data are needed for the Precambrian in order to better characterize the geodynamo evolution.
+
+% question to ask about the paleointensity database:
 

### 2021-11-06 PNAS_manuscript/PNAS_manuscript.tex
Updates from Overleaf

--- 2021-10-12 manuscript/Manuscript.tex
@@
-\documentclass[draft]{agujournal2019}
-\usepackage{url} %this package should fix any errors with URLs in refs.
-\usepackage{lineno}
-\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
-\usepackage{soul}
-\usepackage{adjustbox}
-\linenumbers
+\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+% Use the lineno option to display guide line numbers if required.
 
-\draftfalse
+\templatetype{pnasresearcharticle} 
+% {pnasresearcharticle} = Template for a two-column research article
+% {pnasmathematics} %= Template for a one-column mathematics article
+% {pnasinvited} %= Template for a PNAS invited submission
 
-%% Enter journal name below.
-%% Choose from this list of Journals:
-%
-% JGR: Atmospheres
-% JGR: Biogeosciences
-% JGR: Earth Surface
-% JGR: Oceans
-% JGR: Planets
-% JGR: Solid Earth
-% JGR: Space Physics
-% Global Biogeochemical Cycles
-% Geophysical Research Letters
-% Paleoceanography and Paleoclimatology
-% Radio Science
-% Reviews of Geophysics
-% Tectonics
-% Space Weather
-% Water Resources Research
-% Geochemistry, Geophysics, Geosystems
-% Journal of Advances in Modeling Earth Systems (JAMES)
-% Earth's Future
-% Earth and Space Science
-% Geohealth
-%
+\title{High geomagnetic field intensity recorded by anorthosite xenoliths requires a vigorous late Mesoproterozoic geodynamo}
 
-\journalname{JGR: Solid Earth}
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a,1]{Yiming Zhang}
+\author[a,1]{Nicholas L. Swanson-Hysell} 
+\author[a,b]{Margaret S. Avery}
+\author[c]{Roger R. Fu}
+
+\affil[a]{Department of Earth and Planetary Science, University of California, Berkeley, CA, 94720}
+\affil[b]{Geology, Minerals, Energy, and Geophysics Science Center, U.S. Geological Survey, Moffett Field, CA, 94025}
+\affil[c]{Department of Earth and Planetary Sciences, Harvard University, Cambridge, MA, 02138}
+
+% Please give the surname of the lead author for the running footer
+\leadauthor{Zhang}
+
+% Please add a significance statement to explain the relevance of your work
+\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
+The primary goal of the significance statement is to explain the relevance of the work in broad context to a broad readership.
+The significance statement appears in the paper itself and is required for all research papers.}
+
+% Please include corresponding author, author contribution and author declaration information
+\authorcontributions{Author contributions: Y.Z.
+and N.L.S.-H.
+designed research; Y.Z., N.L.S.-H.
+and M.S.A studied and sampled the rock cores; Y.Z.
+and N.L.S.-H.
+conducted and analyzed rock magnetic experiments; Y.Z.
+performed petrography and electron microscopy; R.R.F.
+contributed to magnetic imaging data analyses; and Y.Z., N.L.S.-H., M.S.A, and R.R.F.
+wrote the paper.}
+
+\authordeclaration{The authors declare no conflict of interest.}
+% \equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
+\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: yimingzhang\@berkeley.edu}
+
+% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
+\keywords{absolute paleointensity $|$ Laurentia $|$ Midcontinent Rift $|$ anorthosite $|$ geodynamo $|$}
+
+\begin{abstract}
+Obtaining estimates of Earth's magnetic field strength in deep time is complicated by non-ideal rock magnetic behavior in many igneous rocks.
+In this study, we target \textit{ca.} 1092 Ma anorthosite xenoliths that cooled and acquired their magnetization within the Beaver River diabase intrusions of the North American Midcontinent Rift.
+In contrast to the diabase which
+
+These anorthosite xenoliths cooled and acquired their magnetization at the same time as their diabase hosts, but their magnetic
+
+New high-quality absolute paleointensity data from  record virtual axial dipole moments (VADM) estimates of $\sim$40 to 130 ZAm$^2$.
+These cooling rate corrected paleointensity estimates are similar to and higher than previous results from extrusive units that are dated to be between \textit{ca.} 1108 Ma and \textit{ca.} 1085 Ma.
+Rock magnetic analyses and magnetic imaging reveal that the anorthosite xenoliths can often have magnetic mineralogy dominated by grains with high coercivities, which can yield high success rates passing our selection criteria relative to their diabase hosts.
+Taken together, new and previously published paleointensity data from the Midcontinent Rift support that a strong geomagnetic field likely persisted through 25 million years of the late Mesoproterozoic.
+These data are inconsistent with interpretations of a continuous progressive decay of Earth's geodynamo strength through the Proterozoic Eon.
+\end{abstract}
+
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
 
@@
 
-\title{High geomagnetic field intensity in the late Mesoproterozoic recorded by Midcontinent Rift anorthosite xenoliths}
+\maketitle
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
 
-\authors{Yiming Zhang\affil{1}, Nicholas L. Swanson-Hysell\affil{1}, Margaret S. Avery\affil{1,3}, Roger R. Fu\affil{4}}
-
-\affiliation{1}{Department of Earth and Planetary Science, University of California, Berkeley, CA, USA}
-\affiliation{2}{Department of Earth and Environmental Sciences, University of Minnesota, Duluth, MN, USA}
-\affiliation{3}{Geology, Minerals, Energy, and Geophysics Science Center, U.S. Geological Survey, Moffett Field, CA, USA}
-\affiliation{4}{Department of Earth and Planetary Sciences, Harvard University, Cambridge, MA, USA}
-\correspondingauthor{Yiming Zhang}{yimingzhang@berkeley.edu}
-
-\begin{keypoints}
-\item New \textit{ca.} 1092 Ma Midcontinent Rift anorthosite record high paleointensity values
-\item High VADM of the geomagnetic field is inconsistent with previous interpretation of Proterozoic long term decay
-\item A strong geodynamo like today likely persisted for about 25 Myr during the Midcontinent Rift development
-\end{keypoints}
-
-% keywords: absolute paleointensity, Laurentia, Midcontinent Rift, anorthosite, geodynamo
-
-\begin{abstract}
-New high-quality absolute paleointensity data from \textit{ca.} 1092 Ma anorthosite xenoliths within the Beaver River diabase intrusions of the North American Midcontinent Rift record an average virtual axial dipole moment (VADM) estimate of $\sim$80 ZAm$^2$.
-This cooling rate corrected paleointensity estimate from Midcontinent Rift intrusions is similar to previous results from extrusive units that are dated to be between \textit{ca.} 1108 Ma and \textit{ca.} 1085 Ma.
-Rock magnetic analyses and magnetic imaging reveal that the anorthosite xenoliths can often have magnetic mineralogy dominated by grains with high coercivities, which can yield high success rates passing our selection criteria relative to their diabase hosts.
-Taken together, new and previously published paleointensity data from the Midcontinent Rift support that a strong geomagnetic field likely persisted through 25 million years of the late Mesoproterozoic.
-These data are inconsistent with interpretations of a progressive decay of Earth's geodynamo strength through the Proterozoic Eon.
-
-\end{abstract}
-
-\section*{Introduction}
-
-Earth's magnetic field is the result of convective flow of liquid iron-alloy in Earth's outer core.
-This self-sustaining geodynamo is collectively powered by the heat flow across the core-mantle boundary (CMB) and from the crystallization of the solid inner core from the liquid outer core which provides latent heat and compositional buoyancy.
-Paleomagnetic studies have found that a dynamo may have existed by the Archean \cite{Selkin2007a, Tarduno2014a, Brenner2020a}.
-However, some estimates of the thermal conductivity of the core suggest that the nucleation of the inner core could be younger than 1 Ga, more recent than previous thought \cite{Pozzo2012a, Gomi2013a, Labrosse2015a, Ohta2016a}.
-Without the power of compositional convection provided by the process of light elements being rejected into the outer core from the crystallizing inner core, the geodynamo was likely dominantly driven by thermal conduction through the CMB, with perhaps additional stirring by other power sources such Mg precipitation \cite{ORourcke2016a}, or perhaps there were intervals early in Earth history when the dynamo was generated by exotic processes such as a convecting basal magma ocean \cite{Ziegler2013a}.
+% If your first paragraph (i.e.
+with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
+If this is the case, add \parshape=0 to the end of the list environment.
+\dropcap{E}arth's magnetic field is the result of convective flow of liquid iron-alloy in Earth's outer core.
+At present day, the geodynamo is collectively powered by heat flow across the core-mantle boundary (CMB) and from the crystallization of the solid inner core from the liquid outer core which provides latent heat and compositional buoyancy.
+However,  while paleomagnetic studies have found that a dynamo field has existed since at least 3.4 billion years ago in the Archean \cite{Selkin2007a, Tarduno2014a, Brenner2020a}, Earth's inner core likely crystallized more recently.
+Without an inner core, Earth's magnetic field would have been generated solely through CMB heat flow.
+Estimates of the timing of the initial crystallization of Earth's inner core are interconnected with estimates for the core's thermal conductivity.
+Higher conductivity values imply faster cooling rates which in turn imply that the threshold for the freezing of the inner core happened more recently \cite{Davies2015a}.
+While estimates of these values were once consistent with a ca.
+3.5 Ga age for the inner core \cite{Gubbins2004a}, revised estimates based on first-principles calculations implied higher conductivities \cite{Koker2012a,Pozzo2012a} and an age for the inner core that is less than 1 Ga \cite{Pozzo2012a, Gomi2013a, Labrosse2015a, Ohta2016a}.
+Without the power of compositional convection provided by the process of light elements being rejected into the outer core from the crystallizing inner core, the geodynamo was likely dominantly driven by thermal conduction through the CMB, with perhaps additional stirring by other power sources such Mg precipitation \cite{ORourke2016a}, or perhaps there were intervals early in Earth history when the dynamo was generated by exotic processes such as a convecting basal magma ocean \cite{Ziegler2013a}.
 Long-term secular cooling of Earth has the potential to have decreased the power available to the geodynamo through time prior to the nucleation of the inner core \cite<e.g>{Davies2021a}.
-As a result, it is possible that rigor of the geomagnetic axial dipole field would have experienced a long-term decay before the emergence of the solid inner core after which it would have strengthened \cite{Aubert2009a, Davies2021a}.
-However, the estimates for thermal conductivity values for the core continue to vary\cite{Konopkova2016a, Ohta2016a}.
+As a result, it is possible that intensity of the geomagnetic axial dipole field would have experienced a long-term decay before the emergence of the solid inner core after which it would have strengthened \cite{Aubert2009a, Davies2021a}.
+However, the estimates for thermal conductivity values for the core continue to vary \cite{Konopkova2016a, Ohta2016a}.
 Obtaining an observational record of the timing of inner core formation would constrain the thermal evolution of the core and mantle, in addition to improving our understanding of the history of Earth’s dynamo.
@@
 
-%Retrieving paleomagnetic information from Precambrian rocks can aid in our understanding of the long-term evolution of the geodynamo.
-A recent paleointensity study by \cite{Sprain2018a} reported a \textit{ca.} 1.1 Ga paleomagnetic field strength similar to today based on the volcanics of the Midcontinent Rift System (MRS).
-That study suggested that this result is consistent with models wherein this time period of the Proterozoic is characterized by a strong rather than weak geomagnetic field.
-However, to further evaluating the evolution of the geomagnetic field during the Precambrian, more paleointensity data are needed.
+\section*{Results and Discussion}
 
-%The MRS emplaced a vast amount of extrusive and intrusive rocks spanning a prolonged period of $\sim$25 Myrs.
-\cite{Swanson-Hysell2019a} published an extensive compilation of the MRS volcanics with paired high-precision CA-ID-TIMS U-Pb geochronology and high-quality paleomagnetic directions, revealing at least two dipole reversals and rapid differential plate motion of the ancient North American plate, Laurentia, during the emplacement of MRS.
-The synthesized North American apparent polar wander path (APWP) was interpreted to be composed primarily of rapid plate motion \citep{Swanson-Hysell2019a}.
-Through comparison between paleomagnetic pole positions and the APWP, such rapid plate motion and high-precision paleomagnetic and geochronological framework revealed by the MRS volcanics can provide insights into age constraints on units with poor or no geochronology data.
-This framework allows for further paleomagnetic contributions beyond the Midcontinent Rift volcanics.
-
-\section*{Geologic Settings}
-% a map of the lake superior including the pesonen units, the lake shore traps, the BRD, and the osler and mamainse, together with date compilation.
-
-The Mesoproterozoic North American Midcontinent Rift (MCR) is an intracontinental rift that did not result in the break-up of Laurentia.
-Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; \citeA{Vervoort2007a, Miller2013a}) was punctuated by rapid and voluminous emplacement of intrusive rocks as well as extrusive volcanics.
-Based on high-precision $^{206}$Pb/$^{238}$U geochronology and paleomagnetic data, \citeA{Swanson-Hysell2021a} hypothesized that the massive Duluth Complex and the coeval North Shore Volcanic group (Figure \ref{fig:Geologic_map}) formed as a result of having upwelling mantled-derived plume material encountering Laurentia's thinned lithosphere, following an ``upside-down” drainage" which rapidly funneled melt to today's North Eastern Minnesota (Figure \ref{fig:Geologic_map}; \citeA{Sleep1997a, Swanson-Hysell2014a}).
-The plagioclase-phyric Anorthositic Series rocks, the crystal-poor Layered Series intrusions of the Duluth Complex, and the coeval volcanics of the North Shore Volcanic Group (NSVG) that outcrop along Northeastern Minnesota (Figure \ref{fig:Geologic_map}) are interpreted to be a result of such vigorous magmatic activity.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{Geologic_map.pdf}
-\caption{\footnotesize{Simplified geologic map of Lake Superior region showing the distribution of rift and post-rift-related rocks including extrusive, intrusive, and late/post-rift sediments around Lake Superior.
-Red diamonds mark sites with paleointensity results that passed the selection criteria from this study.
-Paleomagnetic sites from \citeA{Pesonen1983a} are categorized by lithology.
-All sites from previous studies are color-coded by their ages.}}
-\label{fig:Geologic_map}
-\end{figure}
-
-The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex punctuates another period of magmatism during the Midcontinent Rift main stage activity.
-$^{206}$Pb/$^{238}$U zircon dates and cross-cutting relationships tightly constraints the areally extensive Beaver River diabase of the Beaver Bay Complex to have been emplaced at 1091.7 $\pm$ 0.2 Ma \cite{Zhang2021b}.
-The Beaver River diabase intruded into hypabyssal depths via wide conduits that accommodated the entrainment of numerous anorthosite xenoliths that have short-axis diameters up to 180 meters \cite{Boerboom2004a, Boerboom2006b}.
-Geochronological, paleomagnetic, geochemical, and petrographic data are consistent with the hypothesis that the Beaver River diabase magma extruded to the surface, feeding the massive Greenstone Flow - one of the largest lava flows on Earth \cite{Doyle2016a, Zhang2021b}.
-
-% The anorthosite xenoliths that distinctly occur within the Beaver River diabase are not only recorders of the wide diabase conduit sizes,  of the geomagnetic field $\sim$1091.7 million years ago.
-Thermal modeling result and paleomagnetic directional data show the anorthosite xenoliths acquired thermal remanent magnetizations during cooling of the diabase \cite{Zhang2021b}.
-Step-wise thermal demagnetizations show many anorthosite xenoliths have minimal secondary remanence and single-component characteristic remanence magnetizations that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C.
-
-Despite that the anorthosite xenoliths of the Beaver River diabase were entrained as xenoliths and emplaced only $\sim$4 Myr after the formation of the Anorthositic Series rocks of the Duluth Complex when massive plagioclase mushes were generated, $^{206}$Pb/$^{238}$U zircon geochronology data and zircon rare earth element diffusion modeling results suggest it is the most probable that the Beaver River diabase anorthosite xenoliths are entrained cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
-Moreover, the anorthosite xenoliths of the \textit{ca.} 1092 Ma Beaver River diabase has distinct mineralogy from that of the Anorthositic Series rocks of the Duluth Complex.
-The dominantly monomineralic anorthosite xenoliths have an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}, higher than the average of $\sim$60\% of the Anorthositic Series rocks of the Duluth Complex \cite{Miller1990a}.
-In addition, petrographic imaging support that the plagioclase from the Beaver River anorthosite formed under a magmatic conditions different from the Duluth Complex anorthositic series magma.
-While the plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture (Figure \ref{Petro_QDM}A), the more calcium-rich plagioclase of the Beaver River anorthosite often develop granoblastic texture characterized by equigranular plagioclase crystals (Figure \ref{Petro_QDM}B).
-Abundant magnetite and ilmenite needles that are typically 10 $\mu$m long can be found to have preferred orientations within Duluth Complex plagioclase.
-Interstitial vermicular titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex Anorthositic Series rocks \cite{Miller1990a}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{Petro_QDM.png}
-\caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from Duluth Complex Anorthositic Series (left column) and Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (right column; \citeA{Zhang2021b}).
-Petrographic images (A-D) show Duluth Complex anorthosite have euhedral, elongate plagioclase crystals with interlocking texture and igneous foliation.
-(C) The plagioclase can contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the (010) planes of plagioclase.
-In contrast, plagioclase of sample MS99033 do not enclose large oxide needles.
-Magnetic field maps acquired by the quantum diamond microscope (QDM) show that the oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the overlapping magnetic dipole fields exerted above the oxides in (E).
-In comparison, relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite.
-The dipole-like features in (F) do not show strong preferred orientations.
-Magnetic field maps of (E) and (F) share the same spatial and magnetic scale.
-The unit of the axes are in $\mu$m.}}
-\label{fig:Petro_QDM}
-\end{figure}
-
-Occurrences of magnetite and ilmenite symplectites and needles within anorthositic rocks have been described by \citeA{Scofield1975a, Scofield1986a} in the Wichita Mountains and by \citeA{Haggerty1976a} in other slowly cooled, mafic intrusions.
-\citeA{Scofield1986a} hypothesized that the inclusions acted as a means of buffering the melt with respect to oxygen, forming either crystallites along preexisting plagioclase substrate or as iron incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures.
-Such a model requires that plagioclase, instead of other mafic minerals such as olivine, to be the first to crystallize from a magma and act as a sink for iron.
-The two-stage formation of the Duluth Complex Anorthositic Series fits this model \cite{Miller1990a}.
-During the first stage, high pressure at subcrustal depths (15-25 km depth) caused fractionation of Duluth Complex magma resulting in the formation of floating plagioclase mushes \cite{Kushiro1980a}.
-The second stage of crystallization then occurs at low pressures after the magma intruded into shallower chambers.
-
-This proposed magmatic evolution of the Duluth Complex is in distinct contrast to the emplacement of the Beaver River diabase which rapidly ascended and entrained the anorthosite xenoliths directly into shallow chambers \cite{Flower1980a, Miller1997a, Zhang2021b}.
-Particularly, \citeA{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
-This is consistent with the interpretation that these anorthosite were rapidly transported by upward currents into hypabyssal magma chambers without reaching structural equilibrium.
-Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Figure \ref{Petro_QDM}).
-Instead, relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Figure \ref{Petro_QDM}).
-
-% Previous studies have divided magmatic activity in the rift into four stages based on interpreted changes in relative magmatic volume and the nature of magmatism: early ($\sim$1109–1104 Ma), latent ($\sim$1104–1098 Ma), main ($\sim$1098–1090 Ma) and late ($\sim$1090–1084 Ma)  In particular, the main stage magmatism is characterized by rapid and punctuated events which resulted in voluminous emplacement of intrusive rocks as well as coeval outpouring of extrusive lava flows.
-One example is the massive \textit{ca.} 1096 Ma Duluth Complex and the North Shore Volcanic Group \cite{Swanson-Hysell2020a}, the majority of which emplaced within $<$500 kyrs.
-Another one is the \textit{ca.} 1092 Ma Beaver Bay Complex.
-The Beaver River diabase intrusions of the complex formed synchronously with the Greenstone Flow of the Portage Lake Volcanics which is one of the largest lava flows known on Earth \cite{Zhang2021b}.
-Excellent rock preservation thanks to minimal post-rift alteration in the Midcontinent Rift have yielded high-quality paleomagnetic directional data which have now been paired with high-precision zircon geochronology to reconstruct the ``Keweenawan Track" - a well-resolved Laurentia's (cratonic North America) apparent polar wander path during the late Mesoproterozoic \cite{Swanson-Hysell2019a}.
-
-\section*{Method}
-
-\subsection*{Sample collection and paleomagnetic directions}
-
-We collected paleomagnetic cores that are 2.5 cm in diameter along the southern and eastern Beaver Bay Complex with a particular focus on acquiring paired sites of anorthosite xenoliths and their local diabase hosts during summer field seasons in 2019 and 2020.
-Sample cores were collected using a hand-held gasoline-powered drill and were oriented using a magnetic compass as well as a sun compass when possible.
-Sun compass orientations were preferentially used for determining the sample azimuth.
-Specimens from a total of 7 diabase and 14 anorthosite sites were used for paleointensity experiments (Table \ref{tab:Pmag_site_data}).
-Sister specimens underwent step-wise alternating field (AF) or thermal demagnetization at the UC Berkeley Paleomagnetism Lab and the results are presented in \citeA{Zhang2021b}.
-The close directional similarities between the anorthosite xenoliths and their host diabase and thermal modeling results are consistent with the interpretation that the Beaver River diabase and anorthosite xenoliths acquired primary thermal remanent magnetization \textit{ca.} 1092 Ma during cooling after the emplacement of the diabase \cite{Zhang2021b}.
-Based on the anorthosite thermal demagnetization results we selected sites whose unblocking temperature ranges are narrow and near 580\textdegree C for paleointensity experiments.
-Beaver River diabase sites with minimal secondary remanence were also selected for paleointensity experiments.
-
-\subsection*{Paleointensity experiment}
-A total of 86 specimens from 14 anorthosite xenoliths and a total of 69 specimens from 7 diabase sites underwent paleointensity experiments that followed the step-wise double-heating Thellier method \cite{Thellier1959a} using the IZZI protocol \cite{Tauxe2004a} with heating steps up to 585 \textdegree C.
-Partial thermal remanent magnetization (pTRM) checks were performed systematically throughout the experiment to test whether there was significant mineralogical alteration due to heating and were assessed using the SCAT parameter of \citeA{Shaar2013a}.
-On top of the IZZI modification of the Thellier experimental protocol, we also performed a comparative study where we added an extra step of 20 mT alternating field (AF) cleaning on some of the specimens after each in-field step.
-The purpose is to study whether the AF cleaning could help improve experiment success rate by removing the remanence component carried by materials such as multi-domain (MD) or vortex state grains that often contributes to non-ideal paleointensity behaviors.
-All remanence measurements were made on a 2G Enterprises DC-SQUID superconducting rock magnetometer equipped with an automated sample changer system at the UC Berkeley Paleomagnetism laboratory.
-The magnetometer is housed inside a three-layer magnetostatic shield that maintains background fields of less than 500 nT.
-Heating steps were performed using an ASC TD-48SC thermal demagnetizer with a controlled field coil that allows for a magnetic field to be generated in the oven in conjunction with a DC power supply.
-The thermal demagnetizer was degaussed with an alternating field in the axial orientation following each in-field step such that residual fields within the oven were $<$10 nT during zero-field steps.
-Samples were placed in the same location within the thermal demagnetizer for each heating step and were maintained in the same orientation with regard to the applied field.
-During each heating step, samples remained at peak temperatures for 20 min.
-An applied laboratory field of 30 $\mu$T was used for all in-field steps.
-All heating steps were performed in air.
-The temperature increments for the experiments were chosen to cover characteristic remanent magnetizations held by (titano)magnetite informed by the previous demagnetization data, with smaller increments performed close to the expected unblocking temperature of stoichiometric magnetite \cite{Zhang2021b}.
-
-% After the paleointensity experiment were finished, the same specimens underwent additional TRM anisotropy checks with full TRMs given at 0, 30, 60, 90, 120, 150 degrees in the horizontal plane in specimen coordinates.
-After that, full TRM acquisition experiments were performed on the same specimens with applied fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axis.
-
-\subsection*{Paleointensity result selection}
-The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD; \citeA{Kirschvink1980a}) of $<$20\textdegree; (2) scatter parameter ($\beta$; \citeA{Coe1978a}) values of $<$15 per cent; (3) a deviation angle (DANG; \citeA{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC; \citeA{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT; \citeA{Shaar2013a}) = TRUE; (6) a maximum magnetic moment difference between adjacent zero-field steps (GAP-Max; \citeA{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
-The criteria are also summarized in Table \ref{tab:criteria}.
-The MAD measures the scatter about the best-fit line through the natural remanent magnetization (NRM) steps in the selected interval for which the intensity is defined.
-DANG, the deviation angle, is the angle between the best-fit direction that is free floating and the direction between the centre of mass of the data and the origin of the vector component diagram \cite{Tauxe2004a}.
-Both MAD and DANG assess the directional variation of the NRM, with MAD measuring the scatter in the NRM directions and DANG assessing whether the component is trending toward the origin of the Zijderveld plot.
-$\beta$ is the ``scatter" parameter of \citeA{Coe1978a} and is the ratio of the standard error of the slope of the best-fit line of the selected NRM and pTRM points on an Arai plot to the absolute value of the slope.
-FRAC is the fraction of the NRM that is used in the best-fit line \cite{Shaar2013a}.
-The FRAC value was chosen to preferentially select samples with dominantly single-slope Arai plots.
-GAP-Max is the maximum gap between two points on the Arai plot determined by vector arithmetic.
-SCAT is a Boolean operator which uses the error on the best-fit slope of the selected data on the Arai plot to determine if the data are overly scattered.
-The parameter is used to assess pTRM checks in addition to assessing the degree to which IZZI steps are zigzagged.
-$\beta$, FRAC, GAP-Max and SCAT are all statistics to assess the behavior of Arai plots.
-See the Standard Paleointensity Definitions (\citeA{Paterson2014a}; \url{https://earthref.org/PmagPy/SPD/home.html}) for more details.
-Data analysis was conducted using Thellier GUI \cite{Shaar2013a} within the PmagPy software package \cite{Tauxe2016a}.
-
-\begin{table}
-\caption{Quality criteria used for selecting paleointensity data. See the text for more detail.} 
-\centering
-\begin{tabular}{ccccccccc}
-\hline
-MAD (\textdegree) & Beta (per cent) & DANG (\textdegree) & FRAC& SCAT & GAP-MAX & NpTRM & N Arai\\
-\hline
-20 & 15 & 5 & 0.6 & TRUE & 0.25 & 2 & 4 \\
-\hline
-
-\end{tabular}
-\label{tab:criteria}
-\end{table}
-
-\subsection*{Rock magnetic experiment}
-
-In this study, we conduct rock magnetic experiments with a purpose of gaining magnetic mineralogy insights into the paleointensity results of the anorthosite and diabase.
-Backfield curves were measured at room temperature using a Micromag Princeton Measurements vibrating sample magnetometer (VSM) and a Lake Shore 8600 series VSM.
-Specimen median destructive field (MDF) are calculated based on the backfield curves.
-The calculated coercivity spectra were subsequently decomposed into one ore more components using skew-normal distributions following the method of \citeA{Maxbauer2016a}.
-The fitted components are then used to interpret the characteristics of different populations of magnetic particles within specimens.
-We also used the magnetic property measurement system (MPMS) at the IRM to aid in the identification of magnetic mineral.
-In the field-cooled (FC) experiments, specimen magnetizations were measured upon warming following the specimen having cooled in an applied field of 2.5 T from 300 to 10 K.
-In the zero-field-cooled (ZFC) experiment, a low-temperature saturation isothermal remanence (LTSIRM) of 2.5 T was applied at 10 K after the specimen cooled in a (near-)zero field.
-In the room-temperature saturation isothermal remanence (RTSIRM) experiment, the sample was pulsed with a 2.5 T field at room temperature ($\sim$300 K) and then cooled to 10 K and warmed back to room temperature in a (near-)zero field.
-The magnetic moment transitions at critical temperatures revealed through MPMS experiments are used to identify magnetic minerals such as magnetite \cite{Verwey1939a} within specimens.
-
-To further identify the magnetic carriers within the Beaver River anorthosite xenoliths and compare them with the anorthosites of the Duluth Complex Anorthositic Series rocks, we used the quantum diamond microscope (QDM) at the UC Berkeley Paleomagnetism laboratory to image a thin section of sample MS99033 which is a hand sample from anorthosite xenolith AX16 (which yielded a $^{206}$Pb/$^{238}$U zircon date of 1091.83 $\pm$ 0.21 Ma; \citeA{Zhang2021b}), and a thin section of a Duluth Complex anorthosite (Figure \ref{fig:Petro_QDM}).
-We the QDM to image the magnetic field over the polished thin section surfaces at a sample-sensor distance of 5 $\mu$m in projective magnetic microscopy (PMM) mode with a spatial resolution of 4.7 μm per pixel and an instantaneous 0.9 mT bias field that is canceled during the course of measurement \cite{Glenn2017a}).
-
-\section*{Results and Discussion}
 \subsection*{Paleointensity}
@@
 Therefore, the anorthosite xenoliths record a high geomagnetic field \textit{ca.} 1092 Ma.
+
+Many authors find it useful to organize their manuscripts with the following order of sections;  title, author line and affiliations, keywords, abstract, significance statement, introduction, results, discussion, materials and methods, acknowledgments, and references.
+Other orders and headings are permitted.
 
@@
 
-\subsection*{Late Mesoproterozoic geodynamo }
+\section*{Late Mesoproterozoic geodynamo }
 To best characterize the geomagnetic axial dipole field intensity during a certain time period requires that a paleointensity dataset to cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged and the geocentric axial dipole hypothesis (GAD) can be assumed.
@@
 
-% 1.
-whether septils was recording the time averaged field or the low values are from the hyperactive state which is another signal prior to the onset of inner core?
-indeed it is interesting that no high values from the Ediacaran has been recovered
-
-% 2. where is the recovery, how long was the recovery, how strong was the recovery is at question for he Ediacaran boundary.
-
-% 3. How should one explain the Mundine Well dikes high values?
-
-% 4. How should one explain the Bungemall sills low values in 1070 Ma following the high values of the MCR?
-
-% 5. We need more spatial and temporal coverage of the Precambrian observational data.
-
-% using paleomagnetic observational records is guided by the predictions from numerical simulations that a cooling liquid core with only thermal convection power to drive a dynamo would experience a progressive decay in its magnetic dipole field strength, and experience a rapid recover after the onset of inner core nucleation \cite{Davies2021a}.
-Previous studies has pointed to the Ediacaran-Cambrian boundary for such a weakening field behavior and recent studies has claim to found a rapid increase after the single crystal time which fits in the numerical models.
-However, such a recover may not be unique given the strong geodynamo estimate from the 1.1 Ga Midcontinent Rift.
-One could point to the increase from low values from the ca.
-1140 Ma abitibi dikes to the ca.
-1092 Ma anorthosites to be the result of recovery; another line of disagreement is between the Bangemall sills and the low field from Sept Ils.
-MCR being an anomaly?
-or the time scale we are looking at is too long?
-
-% use a combined method of direction and intensity research for further examinations.
-Re-evaluation of the spatial and temporal variability captured by the existing Precambrian database.
-More data are needed for the Precambrian in order to better characterize the geodynamo evolution.
-
-% question to ask about the paleointensity database:
-
 \begin{figure}[h!]
@@
 \end{figure}
-
-% cooling rate correction problem - that the Nagy et al paper can suggest even less of a correction and higher of the estimated field.
-theoretical thermal modeling by \citeA{Zhang2021a} suggest tha the anorthosite xenoliths were likely heated to diabase magma temperature regardless of their sizes, and cooled through the typical characteristic unblocking temperature range from 580 to 500 on the order of 1000 years.
-Comparing to the typical lab cooling time of about 100 seconds duirng paleointensity experiments, the significantly slower cooling rate when the NRM was acquired as the anorthosite xenoliths cooled with the diabase led to overestimate of paleointensity by up to 40\%.
-Overall range of correction factor can be 1.1-1.4, in any case consistent with a high field strength
-
-% that \citeA{bono2019a} proposed an interesting trend but has problems in the directions, likely to be a short term feature of a low field during a reversal, instead of a low field as a result of long term cooling.
-
-% that Lloyd 2021 also presented low field that fits the trend, but they similarly have the issue of reversal, and that they do not have good control on secular variations
-
-% Together with previously published data, the Midcontinent Rift rocks are recording 1.1 Ga earth geomagnetic field that is stronger than the mean of the past 300 Ma (Figure \ref{}).
-
-% Regardless, a strong geomagnetic field has been recorded by the \textit{ca.} 1092 Ma Beaver River anorthosite xenoliths and a strong geodynamo might have persisted through the protracted Midcontinent Rift magmatism.
 
@@
 
-\acknowledgments
-Project research was supported by NSF CAREER grant EAR-1847277 to N.L.S.-H.
+\matmethods{
+\subsection*{Sample collection and paleomagnetic directions}
+
+We collected paleomagnetic cores that are 2.5 cm in diameter along the southern and eastern Beaver Bay Complex with a particular focus on acquiring paired sites of anorthosite xenoliths and their local diabase hosts during summer field seasons in 2019 and 2020.
+Sample cores were collected using a hand-held gasoline-powered drill and were oriented using a magnetic compass as well as a sun compass when possible.
+Sun compass orientations were preferentially used for determining the sample azimuth.
+Specimens from a total of 7 diabase and 14 anorthosite sites were used for paleointensity experiments (Table \ref{tab:Pmag_site_data}).
+Sister specimens underwent step-wise alternating field (AF) or thermal demagnetization at the UC Berkeley Paleomagnetism Lab and the results are presented in \citeA{Zhang2021b}.
+The close directional similarities between the anorthosite xenoliths and their host diabase and thermal modeling results are consistent with the interpretation that the Beaver River diabase and anorthosite xenoliths acquired primary thermal remanent magnetization \textit{ca.} 1092 Ma during cooling after the emplacement of the diabase \cite{Zhang2021b}.
+Based on the anorthosite thermal demagnetization results we selected sites whose unblocking temperature ranges are narrow and near 580\textdegree C for paleointensity experiments.
+Beaver River diabase sites with minimal secondary remanence were also selected for paleointensity experiments.
+
+\subsection*{Paleointensity experiment}
+A total of 86 specimens from 14 anorthosite xenoliths and a total of 69 specimens from 7 diabase sites underwent paleointensity experiments that followed the step-wise double-heating Thellier method \cite{Thellier1959a} using the IZZI protocol \cite{Tauxe2004a} with heating steps up to 585 \textdegree C.
+Partial thermal remanent magnetization (pTRM) checks were performed systematically throughout the experiment to test whether there was significant mineralogical alteration due to heating and were assessed using the SCAT parameter of \citeA{Shaar2013a}.
+On top of the IZZI modification of the Thellier experimental protocol, we also performed a comparative study where we added an extra step of 20 mT alternating field (AF) cleaning on some of the specimens after each in-field step.
+The purpose is to study whether the AF cleaning could help improve experiment success rate by removing the remanence component carried by materials such as multi-domain (MD) or vortex state grains that often contributes to non-ideal paleointensity behaviors.
+All remanence measurements were made on a 2G Enterprises DC-SQUID superconducting rock magnetometer equipped with an automated sample changer system at the UC Berkeley Paleomagnetism laboratory.
+The magnetometer is housed inside a three-layer magnetostatic shield that maintains background fields of less than 500 nT.
+Heating steps were performed using an ASC TD-48SC thermal demagnetizer with a controlled field coil that allows for a magnetic field to be generated in the oven in conjunction with a DC power supply.
+The thermal demagnetizer was degaussed with an alternating field in the axial orientation following each in-field step such that residual fields within the oven were $<$10 nT during zero-field steps.
+Samples were placed in the same location within the thermal demagnetizer for each heating step and were maintained in the same orientation with regard to the applied field.
+During each heating step, samples remained at peak temperatures for 20 min.
+An applied laboratory field of 30 $\mu$T was used for all in-field steps.
+All heating steps were performed in air.
+The temperature increments for the experiments were chosen to cover characteristic remanent magnetizations held by (titano)magnetite informed by the previous demagnetization data, with smaller increments performed close to the expected unblocking temperature of stoichiometric magnetite \cite{Zhang2021b}.
+
+% After the paleointensity experiment were finished, the same specimens underwent additional TRM anisotropy checks with full TRMs given at 0, 30, 60, 90, 120, 150 degrees in the horizontal plane in specimen coordinates.
+After that, full TRM acquisition experiments were performed on the same specimens with applied fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axis.
+
+\subsection*{Paleointensity result selection}
+The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD; \citeA{Kirschvink1980a}) of $<$20\textdegree; (2) scatter parameter ($\beta$; \citeA{Coe1978a}) values of $<$15 per cent; (3) a deviation angle (DANG; \citeA{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC; \citeA{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT; \citeA{Shaar2013a}) = TRUE; (6) a maximum magnetic moment difference between adjacent zero-field steps (GAP-Max; \citeA{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
+The criteria are also summarized in Table \ref{tab:criteria}.
+The MAD measures the scatter about the best-fit line through the natural remanent magnetization (NRM) steps in the selected interval for which the intensity is defined.
+DANG, the deviation angle, is the angle between the best-fit direction that is free floating and the direction between the centre of mass of the data and the origin of the vector component diagram \cite{Tauxe2004a}.
+Both MAD and DANG assess the directional variation of the NRM, with MAD measuring the scatter in the NRM directions and DANG assessing whether the component is trending toward the origin of the Zijderveld plot.
+$\beta$ is the ``scatter" parameter of \citeA{Coe1978a} and is the ratio of the standard error of the slope of the best-fit line of the selected NRM and pTRM points on an Arai plot to the absolute value of the slope.
+FRAC is the fraction of the NRM that is used in the best-fit line \cite{Shaar2013a}.
+The FRAC value was chosen to preferentially select samples with dominantly single-slope Arai plots.
+GAP-Max is the maximum gap between two points on the Arai plot determined by vector arithmetic.
+SCAT is a Boolean operator which uses the error on the best-fit slope of the selected data on the Arai plot to determine if the data are overly scattered.
+The parameter is used to assess pTRM checks in addition to assessing the degree to which IZZI steps are zigzagged.
+$\beta$, FRAC, GAP-Max and SCAT are all statistics to assess the behavior of Arai plots.
+See the Standard Paleointensity Definitions (\citeA{Paterson2014a}; \url{https://earthref.org/PmagPy/SPD/home.html}) for more details.
+Data analysis was conducted using Thellier GUI \cite{Shaar2013a} within the PmagPy software package \cite{Tauxe2016a}.
+
+\begin{table}
+\caption{Quality criteria used for selecting paleointensity data. See the text for more detail.} 
+\centering
+\begin{tabular}{ccccccccc}
+\hline
+MAD (\textdegree) & Beta (per cent) & DANG (\textdegree) & FRAC& SCAT & GAP-MAX & NpTRM & N Arai\\
+\hline
+20 & 15 & 5 & 0.6 & TRUE & 0.25 & 2 & 4 \\
+\hline
+
+\end{tabular}
+\label{tab:criteria}
+\end{table}
+
+\subsection*{Rock magnetic experiment}
+
+In this study, we conduct rock magnetic experiments with a purpose of gaining magnetic mineralogy insights into the paleointensity results of the anorthosite and diabase.
+Backfield curves were measured at room temperature using a Micromag Princeton Measurements vibrating sample magnetometer (VSM) and a Lake Shore 8600 series VSM.
+Specimen median destructive field (MDF) are calculated based on the backfield curves.
+The calculated coercivity spectra were subsequently decomposed into one ore more components using skew-normal distributions following the method of \citeA{Maxbauer2016a}.
+The fitted components are then used to interpret the characteristics of different populations of magnetic particles within specimens.
+We also used the magnetic property measurement system (MPMS) at the IRM to aid in the identification of magnetic mineral.
+In the field-cooled (FC) experiments, specimen magnetizations were measured upon warming following the specimen having cooled in an applied field of 2.5 T from 300 to 10 K.
+In the zero-field-cooled (ZFC) experiment, a low-temperature saturation isothermal remanence (LTSIRM) of 2.5 T was applied at 10 K after the specimen cooled in a (near-)zero field.
+In the room-temperature saturation isothermal remanence (RTSIRM) experiment, the sample was pulsed with a 2.5 T field at room temperature ($\sim$300 K) and then cooled to 10 K and warmed back to room temperature in a (near-)zero field.
+The magnetic moment transitions at critical temperatures revealed through MPMS experiments are used to identify magnetic minerals such as magnetite \cite{Verwey1939a} within specimens.
+
+To further identify the magnetic carriers within the Beaver River anorthosite xenoliths and compare them with the anorthosites of the Duluth Complex Anorthositic Series rocks, we used the quantum diamond microscope (QDM) at the UC Berkeley Paleomagnetism laboratory to image a thin section of sample MS99033 which is a hand sample from anorthosite xenolith AX16 (which yielded a $^{206}$Pb/$^{238}$U zircon date of 1091.83 $\pm$ 0.21 Ma; \citeA{Zhang2021b}), and a thin section of a Duluth Complex anorthosite (Figure \ref{fig:Petro_QDM}).
+We the QDM to image the magnetic field over the polished thin section surfaces at a sample-sensor distance of 5 $\mu$m in projective magnetic microscopy (PMM) mode with a spatial resolution of 4.7 μm per pixel and an instantaneous 0.9 mT bias field that is canceled during the course of measurement \cite{Glenn2017a}).
+}
+
+\showmatmethods{} % Display the Materials and Methods section
+
+\acknow{Project research was supported by NSF CAREER grant EAR-1847277 to N.L.S.-H.
 and an Institute on Lake Superior Geology Student Research Fund grant to Y.Z.
@@
 Paleomagnetic data associated with this study are available within the MagIC database (\url{}; UPDATE WHEN DOI IS GENERATED) and all data are within a Github repository associated with this work (\url{}) that is also archived on Zenodo (INSERT URL AT TIME OF PROOFS).
-This repository also contains Python code related to calculations, visualizations and statistical tests discussed herein.
+This repository also contains Python code related to calculations, visualizations and statistical tests discussed herein.}
+
+\showacknow{} % Display the acknowledgments section
+
+% Bibliography
 \bibliography{YZ_ref}
@@
 
-% potential reviewer: Peter Selkin, Jeff Gee, Lisa Tauxe, Roger Fu, Lauri Pesonen, Josh Feinberg Julie Bowles,
+% \subsection*{Manuscript Length}
 
-% At the IRM, I was seeking the answer to this question with the help of the vibrating sample magnetometer (VSM) systems, including a newly installed Lake Shore VSM which greatly helped improve measurement resolution on samples with weak magnetizations (Figure \ref{fig:backfield}).
-Backfield demagnetization experiments were conducted and used to develop coercivity spectra (Figure \ref{fig:backfield}).
-The spectra were subsequently modeled to fit for the distributions of different  populations of magnetic particles using similar procedure as \cite{Maxbauer2016a}.
-Most spectra can be well approximated with models with one component or two components with overlapping coercivity ranges (Figure \ref{fig:backfield}).
-The dominance of single-component distributions from the unmixed coercivity spectra likely suggests minimal alterations or formation of secondary magnetic mineral within the samples.
-By compiling the values of median destructive field (MDF) from all measured specimens and categorizing them in terms of their sister specimens' paleointensity results, I found that the anorthosites samples that pass paleointensity selection criteria have distinctively higher MDF values than those did not (Figure \ref{fig:backfield}), and all diabase specimens have low MDF values like the non-ideal anorthosites (Figure \ref{fig:backfield}).
-The higher MDFs ($\sim$60 mT) of the anorthosites are similar to what is typical of stoichiometric, single-domain magnetite grains which are favored for paleointensity experiments.
-Furthermore, the low MDFs displayed in other specimens could be linked to a dominant population of large interstitial Fe-Ti oxides that are more likely to form multi-domain grains, which have been observed in both the anorthosite and diabase (Figure \ref{fig:field}).
-In addition, preliminary tests on single anorthosite crystals might suggest that not all single crystals are necessarily better for paleointensity research when comparing to bulk samples, given their similar low MDF to the failed diabase and anorthosite specimens.
-However, more single crystal grains with paired paleointensity data and rock magnetic data may be needed before a general conclusion can be drawn.
+% A standard 6-page article is approximately 4,000 words, 50 references, and 4 medium-size graphical elements (i.e., figures and tables).
+The preferred length of articles remains at 6 pages, but PNAS will allow articles up to a maximum of 12 pages.
 
-% Taken together, the rock magnetism experiment results support my paleointensity selection criteria in that rock specimens that produce straight Arai plots and have MDF similar to stoichiometric single-domain magnetite are preferentially selected.
-The preliminary paleointensity results yielded consistent estimates.
-The cooling rate corrected site-mean paleointensity estimate is about 40 $\mu$T, consistent with results from \cite{Sprain2018a} that the Earth's magnetic field strength at surface in the late Mesoproterozoic was close to that of today.
+% \subsection*{References}
 
-% \begin{figure}
-% \begin{center}
-% 	\noindent\includegraphics[width=0.5\textwidth]{backfield_example_.pdf}
-% \end{center}
-% \caption{\small{(A) Backfield measurement data acquired using a Princeton VSM on an anorthosite specimen.
-The noise level on coercivity plot is high and a heavy smoothing is needed to make subsequent coercivity modeling interpretable.
-(B) Backfield measurement data acquired using a Lake Shore VSM on the same specimen as that in (A).
-The inset coercivity unmixing plot shows that the much smoother initial backfield curve resulted in a tight single-component model fit.
-(C) Box plot of distribution of the median destructive field of all measured specimens.
-The anorthosite specimens that pass selection criteria have distinctly higher median destructive field (MDF) than other groups.}}
-% \label{fig:backfield}
-% \end{figure}
+% References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g.
+\cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric}.
+All references cited in the main text should be included in the main manuscript file.
+
+% \subsection*{Data Archival}
+
+% PNAS must be able to archive the data essential to a published article.
+Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the \href{https://www.pnas.org/page/authors/journal-policies#xi}{Information for Authors}, is acceptable.
+
+% \begin{SCfigure*}[\sidecaptionrelwidth][t]
+% \centering
+% \includegraphics[width=11.4cm,height=11.4cm]{frog}
+% \caption{This legend would be placed at the side of the figure, rather than below it.}
+% \label{fig:side}
+% \end{SCfigure*}
+
+% \subsection*{Digital Figures}
+
+% EPS, high-resolution PDF, and PowerPoint are preferred formats for figures that will be used in the main manuscript.
+Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.
+Color images must be in RGB (red, green, blue) mode.
+Include the font files for any text.
+
+% Images must be provided at final size, preferably 1 column width (8.7cm).
+Figures wider than 1 column should be sized to 11.4cm or 17.8cm wide.
+Numbers, letters, and symbols should be no smaller than 6 points (2mm) and no larger than 12 points (6mm) after reduction and must be consistent.
+
+% Figures and tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.
+
+% Figure \ref{fig:frog} shows an example of how to insert a column-wide figure.
+To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment.
+Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
+Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side legends.
+
+% \subsection*{Tables}
+% Tables should be included in the main manuscript file and should not be uploaded separately.
+
+% \begin{table}%[tbhp]
+% \centering
+% \caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
+% \begin{tabular}{lrrr}
+% Species & CBS & CV & G3 \\
+% \midrule
+% 1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
+% 2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
+% 3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
+% \bottomrule
+% \end{tabular}
+
+% \addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
+% \end{table}
+
+% \subsection*{Supporting Information Appendix (SI)}
+
+% Authors should submit SI as a single separate SI Appendix PDF file, combining all text, figures, tables, movie legends, and SI references.
+SI will be published as provided by the authors; it will not be edited or composed.
+Additional details can be found in the \href{https://www.pnas.org/authors/submitting-your-manuscript#manuscript-formatting-guidelines}{PNAS Author Center}.
+The PNAS Overleaf SI template can be found \href{https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd}{here}.
+Refer to the SI Appendix in the manuscript at an appropriate point in the text.
+Number supporting figures and tables starting with S1, S2, etc.
+
+% Authors who place detailed materials and methods in an SI Appendix must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the SI methods.
+If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.
+
+% \subsubsection*{SI Datasets}
+
+% Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be published in raw format and will not be edited or composed.

### 2021-11-07 PNAS_manuscript/PNAS_manuscript.tex

@@
 In this study, we target \textit{ca.} 1092 Ma anorthosite xenoliths that cooled and acquired their magnetization within the Beaver River diabase intrusions of the North American Midcontinent Rift.
-In contrast to the diabase which
-
-These anorthosite xenoliths cooled and acquired their magnetization at the same time as their diabase hosts, but their magnetic
-
-New high-quality absolute paleointensity data from  record virtual axial dipole moments (VADM) estimates of $\sim$40 to 130 ZAm$^2$.
-These cooling rate corrected paleointensity estimates are similar to and higher than previous results from extrusive units that are dated to be between \textit{ca.} 1108 Ma and \textit{ca.} 1085 Ma.
-Rock magnetic analyses and magnetic imaging reveal that the anorthosite xenoliths can often have magnetic mineralogy dominated by grains with high coercivities, which can yield high success rates passing our selection criteria relative to their diabase hosts.
-Taken together, new and previously published paleointensity data from the Midcontinent Rift support that a strong geomagnetic field likely persisted through 25 million years of the late Mesoproterozoic.
-These data are inconsistent with interpretations of a continuous progressive decay of Earth's geodynamo strength through the Proterozoic Eon.
+In contrast to the diabase which has non-ideal paleointensity behaviors leading to its failure in providing reliable paleointensity estimates, the anorthosite xenoliths often yield high-quality, single-slope paleointensity results that are consistent at specimen and site levels.
+The high paleointensity estimates from the anorthosites require an existence of a vigorous late Mesoproterozoic geodynamo.
+Together with previous paleointensity data from the Midcontinent Rift, our new results suggest that a strong geodynamo likely persisted through $\sim$25 Myr at ca.
+1.1 Ga, inconsistent with the interpretation of a progressive decay of Earth's geodynamo strength through the Proterozoic Eon.
 \end{abstract}
@@
 However, while paleomagnetic studies have found that a dynamo field has existed since at least 3.4 billion years ago in the Archean \cite{Selkin2007a, Tarduno2014a, Brenner2020a}, Earth's inner core likely crystallized more recently.
-Without an inner core, Earth's magnetic field would have been generated solely through CMB heat flow.
 Estimates of the timing of the initial crystallization of Earth's inner core are interconnected with estimates for the core's thermal conductivity.
 Higher conductivity values imply faster cooling rates which in turn imply that the threshold for the freezing of the inner core happened more recently \cite{Davies2015a}.
-While estimates of these values were once consistent with a ca.
-3.5 Ga age for the inner core \cite{Gubbins2004a}, revised estimates based on first-principles calculations implied higher conductivities \cite{Koker2012a,Pozzo2012a} and an age for the inner core that is less than 1 Ga \cite{Pozzo2012a, Gomi2013a, Labrosse2015a, Ohta2016a}.
-Without the power of compositional convection provided by the process of light elements being rejected into the outer core from the crystallizing inner core, the geodynamo was likely dominantly driven by thermal conduction through the CMB, with perhaps additional stirring by other power sources such Mg precipitation \cite{ORourke2016a}, or perhaps there were intervals early in Earth history when the dynamo was generated by exotic processes such as a convecting basal magma ocean \cite{Ziegler2013a}.
-Long-term secular cooling of Earth has the potential to have decreased the power available to the geodynamo through time prior to the nucleation of the inner core \cite<e.g>{Davies2021a}.
+While some estimates of these values are consistent with an inner core age $>$ 3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates based on first-principles calculations implied higher conductivities \cite{Koker2012a,Pozzo2012a} and an age for the inner core that is less than 1.3 Ga \cite{Pozzo2012a, Gomi2013a, Labrosse2015a, Ohta2016a, Zhang2020b}.
+With the estimates for the core thermal conductivity continue to be debated, it is crucial to use observational records as an independent constraint on the thermal evolution of the core and mantle.
+
+% Without an inner core, Earth's magnetic field would have been generated solely through CMB heat flow, with perhaps additional stirring by other power sources such as Mg precipitation \cite{ORourke2016a}, or by exotic processes such as a convecting basal magma ocean \cite{Ziegler2013a}.
+Long-term secular cooling of Earth has the potential to have decreased the power available to the geodynamo through time prior to the nucleation of the inner core \cite[e.g][]{Davies2021a}.
 As a result, it is possible that intensity of the geomagnetic axial dipole field would have experienced a long-term decay before the emergence of the solid inner core after which it would have strengthened \cite{Aubert2009a, Davies2021a}.
-However, the estimates for thermal conductivity values for the core continue to vary \cite{Konopkova2016a, Ohta2016a}.
+However, the estimates for thermal conductivity values for the core continue to vary \cite{Konopkova2016a, Ohta2016a, Zhang2020a}.
 Obtaining an observational record of the timing of inner core formation would constrain the thermal evolution of the core and mantle, in addition to improving our understanding of the history of Earth’s dynamo.
@@
 Paleomagnetic data from ancient rocks are one of the few types of observational data that have the potential to record the evolution of Earth’s core and can be used to test the predictions from numerical simulations.
-Some Precambrian absolute paleointensity records from the PINT database (\url{http://earth.liv.ac.uk/pint/}) show evidence of an existence of an appreciable geomagnetic field prior to the Ediacaran.
-On the other hand, studies have found very low axial dipole moment (VDM) estimates from rocks formed in the late Ediacaran \cite{Bono2019a, Shcherbakova2019a, Thallner2021b}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
+Some Precambrian absolute paleointensity records from the PINT database (\url{http://earth.liv.ac.uk/pint/}) show evidence of an existence of an appreciable geomagnetic field prior to the Ediacaran (Figure \ref{fig:PINT_compilation}).
+On the other hand, studies have found very low virtual dipole moment (VDM) estimates from rocks formed in the late Ediacaran \cite{Bono2019a, Shcherbakova2019a, Thallner2021b}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
 Other studies focusing on paleomagnetic directions have also found that the late Ediacaran might have been a period of time when the geomagnetic field were experiencing frequent reversals and waning of dipole dominance \cite{Bono2015a, Kodama2020a}.
 It has been interpreted that these findings are consistent with the results of geodynamo models using high core thermal conductivity values that suggest the inner core nucleation age being near the Ediacaran-Cambrian boundary \cite{Driscoll2016a, Davies2021a}.
-\citeA{Bono2019a} based on a polynomial fit through filtered Precambrian paleointensity database to suggest the selected observational records show that the geodynamo power progressively decreased through the Proterozoic (Figure \ref{fig:PINT_compilation}).
-That study also recovered very low paleointensity values from \textit{ca.} 565 Ma single silicate crystals and interpreted them to be reflecting that the geodynamo reached its lowest near the Ediacaran-Cambrian boundary.
+\citeA{Bono2019a} adopted a billion-year-scale view of the Precambrian paleointensity records and based on a polynomial fit through selected data to suggest that the geodynamo power progressively decreased through the Proterozoic (Figure \ref{fig:PINT_compilation}).
+That study also recovered very low paleointensity values from \textit{ca.} 565 Ma single silicate crystals and interpreted them to be reflecting the geodynamo to have reached its lowest near the Ediacaran-Cambrian boundary.
 New low paleointensity estimates derived from the \textit{ca.} 720 Ma Franklin Large Igneous Province are interpreted to support the existence of such decay trend in the Neoproterozoic \cite{Lloyd2021a}.
@@
 
+\section*{Anorthosite xenoliths of the Midcontinent diabase}
+
+The Mesoproterozoic North American Midcontinent Rift (MCR) is an intracontinental rift that did not result in the break-up of Laurentia.
+Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; \citeA{Vervoort2007a, Miller2013a}) was punctuated by rapid and voluminous emplacement of intrusive rocks as well as extrusive volcanics.
+Based on high-precision $^{206}$Pb/$^{238}$U geochronology and paleomagnetic data, \citeA{Swanson-Hysell2021a} hypothesized that the massive Duluth Complex and the coeval North Shore Volcanic group (Figure \ref{fig:Geologic_map}) formed as a result of having upwelling mantled-derived plume material encountering Laurentia's thinned lithosphere, following an ``upside-down” drainage" which rapidly funneled melt to today's North Eastern Minnesota (Figure \ref{fig:Geologic_map}; \citeA{Sleep1997a, Swanson-Hysell2014a}).
+The plagioclase-phyric Anorthositic Series rocks, the crystal-poor Layered Series intrusions of the Duluth Complex, and the coeval volcanics of the North Shore Volcanic Group (NSVG) that outcrop along Northeastern Minnesota (Figure \ref{fig:Geologic_map}) are interpreted to be a result of such vigorous magmatic activity.
+
+\begin{figure*}
+\centering
+\noindent\includegraphics[width=11.4 cm]{Geologic_map.pdf}
+\caption{\footnotesize{Simplified geologic map of Lake Superior region showing the distribution of rift and post-rift-related rocks including extrusive, intrusive, and late/post-rift sediments around Lake Superior.
+Red diamonds mark sites with paleointensity results that passed the selection criteria from this study.
+Paleomagnetic sites from \citeA{Pesonen1983a} are categorized by lithology.
+All sites from previous studies are color-coded by their ages.}}
+\label{fig:Geologic_map}
+\end{figure*}
+
+The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex punctuates another period of magmatism during the Midcontinent Rift main stage activity.
+$^{206}$Pb/$^{238}$U zircon dates and cross-cutting relationships tightly constraints the areally extensive Beaver River diabase of the Beaver Bay Complex to have been emplaced at 1091.7 $\pm$ 0.2 Ma \cite{Zhang2021b}.
+The Beaver River diabase intruded into hypabyssal depths via wide conduits that accommodated the entrainment of numerous anorthosite xenoliths that have short-axis diameters up to 180 meters \cite{Boerboom2004a, Boerboom2006b}.
+Geochronological, paleomagnetic, geochemical, and petrographic data are consistent with the hypothesis that the Beaver River diabase magma extruded to the surface, feeding the massive Greenstone Flow - one of the largest lava flows on Earth \cite{Doyle2016a, Zhang2021b}.
+
+% The anorthosite xenoliths that distinctly occur within the Beaver River diabase are not only recorders of the wide diabase conduit sizes,  of the geomagnetic field $\sim$1091.7 million years ago.
+Thermal modeling result and paleomagnetic directional data show the anorthosite xenoliths acquired thermal remanent magnetizations during cooling of the diabase \cite{Zhang2021b}.
+Step-wise thermal demagnetizations show many anorthosite xenoliths have minimal secondary remanence and single-component characteristic remanence magnetizations that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C.
+
+Despite that the anorthosite xenoliths of the Beaver River diabase were entrained as xenoliths and emplaced only $\sim$4 Myr after the formation of the Anorthositic Series rocks of the Duluth Complex when massive plagioclase mushes were generated, $^{206}$Pb/$^{238}$U zircon geochronology data and zircon rare earth element diffusion modeling results suggest it is the most probable that the Beaver River diabase anorthosite xenoliths are entrained cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
+The anorthosite xenoliths of the \textit{ca.} 1092 Ma Beaver River diabase has distinct mineralogy from that of the Anorthositic Series rocks of the Duluth Complex.
+The dominantly monomineralic anorthosite xenoliths have an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}, higher than the average of $\sim$60\% of the Anorthositic Series rocks of the Duluth Complex \cite{Miller1990a}.
+In addition, petrographic imaging support that the plagioclase from the Beaver River anorthosite formed under a magmatic conditions different from the Duluth Complex anorthositic series magma.
+While the plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture (Figure \ref{Petro_QDM}A), the more calcium-rich plagioclase of the Beaver River anorthosite often develop granoblastic texture characterized by equigranular plagioclase crystals (Figure \ref{Petro_QDM}B).
+Abundant magnetite and ilmenite needles that are typically 10 $\mu$m long can be found to have preferred orientations within Duluth Complex plagioclase.
+Interstitial vermicular titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex Anorthositic Series rocks \cite{Miller1990a}.
+
+\begin{SCfigure*}[\sidecaptionrelwidth][t]
+\centering
+\noindent\includegraphics[width=11.4 cm]{Petro_QDM.png}
+\caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from Duluth Complex Anorthositic Series (left column) and Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (right column; \citeA{Zhang2021b}).
+Petrographic images (A-D) show Duluth Complex anorthosite have euhedral, elongate plagioclase crystals with interlocking texture and igneous foliation.
+(C) The plagioclase can contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the (010) planes of plagioclase.
+In contrast, plagioclase of sample MS99033 do not enclose large oxide needles.
+Magnetic field maps acquired by the quantum diamond microscope (QDM) show that the oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the overlapping magnetic dipole fields exerted above the oxides in (E).
+In comparison, relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite.
+The dipole-like features in (F) do not show strong preferred orientations.
+Magnetic field maps of (E) and (F) share the same spatial and magnetic scale.
+The unit of the axes are in $\mu$m.}}
+\label{fig:Petro_QDM}
+\end{SCfigure*}
+
+Occurrences of magnetite and ilmenite symplectites and needles within anorthositic rocks have been described by \citeA{Scofield1975a, Scofield1986a} in the Wichita Mountains and by \citeA{Haggerty1976a} in other slowly cooled, mafic intrusions.
+\citeA{Scofield1986a} hypothesized that the inclusions acted as a means of buffering the melt with respect to oxygen, forming either crystallites along preexisting plagioclase substrate or as iron incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures.
+Such a model requires that plagioclase, instead of other mafic minerals such as olivine, to be the first to crystallize from a magma and act as a sink for iron.
+The two-stage formation of the Duluth Complex Anorthositic Series fits this model \cite{Miller1990a}.
+During the first stage, high pressure at subcrustal depths (15-25 km depth) caused fractionation of Duluth Complex magma resulting in the formation of floating plagioclase mushes \cite{Kushiro1980a}.
+The second stage of crystallization then occurs at low pressures after the magma intruded into shallower chambers.
+
+This proposed magmatic evolution of the Duluth Complex is in distinct contrast to the emplacement of the Beaver River diabase which rapidly ascended and entrained the anorthosite xenoliths directly into shallow chambers \cite{Flower1980a, Miller1997a, Zhang2021b}.
+Particularly, \citeA{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
+This is consistent with the interpretation that these anorthosite were rapidly transported by upward currents into hypabyssal magma chambers without reaching structural equilibrium.
+Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Figure \ref{Petro_QDM}).
+Instead, relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Figure \ref{Petro_QDM}).
+
 \section*{Results and Discussion}
@@
 
-\begin{figure}[h!]
-\noindent\includegraphics[width=\textwidth]{IZZI_examples.pdf}
+\begin{figure*}[h!]
+\noindent\includegraphics[width=17.8cm]{IZZI_examples.pdf}
 \centering
@@
 \label{fig:IZZI_examples}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[h!]
-\noindent\includegraphics[width=\textwidth]{Paleointensity_plot_cooling_corrected.pdf}
+\begin{figure*}[h!]
+\noindent\includegraphics[width=17.8 cm]{Paleointensity_plot_cooling_corrected.pdf}
 \centering
@@
 \label{fig:PINT_cooling_corrected}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{sidewaystable}
+\begin{table*}
 \caption{\footnotesize{Summary paleointensity results for specimens that passed our selection.
@@
 \centering
-\begin{tabular}{cccccccccccccccc}
+\begin{tabular}[width=17.8cm]{cccccccccccccccc}
 \hline
@@
 \label{tab:PINT_result}
-\end{sidewaystable}
+\end{table*}
 
-\begin{table}[]
+\begin{table*}[]
 
@@
 \label{tab:QPI}
-\end{table}
+\end{table*}
 
@@
 
-\begin{figure}[h!]
-\noindent\includegraphics[width=\textwidth]{MPMS.pdf}
-\centering
-\caption{\footnotesize{Low-temperature magnetic property measurement system (MPMS) experiment results.
+% \begin{figure*}
+% \noindent\includegraphics[width=11.4 cm]{MPMS.pdf}
+% \centering
+% \caption{\footnotesize{Low-temperature magnetic property measurement system (MPMS) experiment results.
 In the field-cooled (FC) experiments, the magnetization was measured upon warming following the specimen having cooled in an applied field of 2.5 T from 300 to 10 K.
@@
 This is consistent with the interpretation that multidomian magnetic carriers exist in significant quantity in this specimen.}}
-\label{fig:MPMS}
-\end{figure}
+% \label{fig:MPMS}
+% \end{figure*}
 
@@
 
-\begin{figure}[h!]
-\noindent\includegraphics[width=\textwidth]{coercivity.pdf}
+\begin{figure}
+\noindent\includegraphics[width=\linewidth]{coercivity.pdf}
 \centering
@@
 
-\begin{figure}[h!]
-\noindent\includegraphics[width=\textwidth]{PINT_compilation.pdf}
+\begin{figure*}[h!]
+\noindent\includegraphics[width=17.8 cm]{PINT_compilation.pdf}
 \centering
@@
 \label{fig:PINT_compilation}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{table}
+\begin{table*}
 \caption{Quality criteria used for selecting paleointensity data. See the text for more detail.} 
@@
 \label{tab:criteria}
-\end{table}
+\end{table*}
 

### 2021-11-07 PNAS_manuscript/Supporting Info.tex
update manuscript

--- 2021-10-12 manuscript/Manuscript.tex
@@
-\documentclass[draft]{agujournal2019}
-\usepackage{url} %this package should fix any errors with URLs in refs.
-\usepackage{lineno}
-\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
-\usepackage{soul}
-\usepackage{adjustbox}
-\linenumbers
+\documentclass[9pt,twoside,lineno]{pnas-new}
+% Use the lineno option to display guide line numbers if required.
+\usepackage{rotating}
+\templatetype{pnassupportinginfo}
 
-\draftfalse
-
-%% Enter journal name below.
-%% Choose from this list of Journals:
-%
-% JGR: Atmospheres
-% JGR: Biogeosciences
-% JGR: Earth Surface
-% JGR: Oceans
-% JGR: Planets
-% JGR: Solid Earth
-% JGR: Space Physics
-% Global Biogeochemical Cycles
-% Geophysical Research Letters
-% Paleoceanography and Paleoclimatology
-% Radio Science
-% Reviews of Geophysics
-% Tectonics
-% Space Weather
-% Water Resources Research
-% Geochemistry, Geophysics, Geosystems
-% Journal of Advances in Modeling Earth Systems (JAMES)
-% Earth's Future
-% Earth and Space Science
-% Geohealth
-%
-
-\journalname{JGR: Solid Earth}
+\title{High geomagnetic field intensity recorded by anorthosite xenoliths requires a vigorous late Mesoproterozoic geodynamo}
+\author{Yiming Zhang, Nicholas L. Swanson-Hysell, Margaret S. Avery, Roger R. Fu}
+\correspondingauthor{Yiming Zhang\\E-mail: yimingzhang@berkeley.edu}
 
@@
 
-\title{High geomagnetic field intensity in the late Mesoproterozoic recorded by Midcontinent Rift anorthosite xenoliths}
+\maketitle
 
-\authors{Yiming Zhang\affil{1}, Nicholas L. Swanson-Hysell\affil{1}, Margaret S. Avery\affil{1,3}, Roger R. Fu\affil{4}}
+%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
+% \SItext
 
-\affiliation{1}{Department of Earth and Planetary Science, University of California, Berkeley, CA, USA}
-\affiliation{2}{Department of Earth and Environmental Sciences, University of Minnesota, Duluth, MN, USA}
-\affiliation{3}{Geology, Minerals, Energy, and Geophysics Science Center, U.S. Geological Survey, Moffett Field, CA, USA}
-\affiliation{4}{Department of Earth and Planetary Sciences, Harvard University, Cambridge, MA, USA}
-\correspondingauthor{Yiming Zhang}{yimingzhang@berkeley.edu}
+% \subsection*{Subhead}
+% Type or paste text here. This should be additional explanatory text such as an extended technical description of results, full details of mathematical models, etc.
 
-\begin{keypoints}
-\item New \textit{ca.} 1092 Ma Midcontinent Rift anorthosite record high paleointensity values
-\item High VADM of the geomagnetic field is inconsistent with previous interpretation of Proterozoic long term decay
-\item A strong geodynamo like today likely persisted for about 25 myr during the Midcontinent Rift development
-\end{keypoints}
+% \section*{Heading}
+% \subsection*{Subhead}
+% Type or paste text here. You may break this section up into subheads as needed (e.g., one section on ``Materials'' and one on ``Methods'').
 
-% keywords: absolute paleointensity, Laurentia, Midcontinent Rift, anorthosite, geodynamo
+% \subsection*{Materials}
+% Add a materials subsection if you need to.
 
-\begin{abstract}
-New high-quality absolute paleointensity data from \textit{ca.} 1092 Ma anorthosite xenoliths within the Beaver River diabase intrusions of the North American Midcontinent Rift record an average virtual axial dipole moment (VADM) estimate of $\sim$80 ZAm$^2$.
-This cooling rate corrected paleointensity estimate from Midcontinent Rift intrusions is similar to previous results from extrusive units that are dated to be between \textit{ca.} 1108 Ma and \textit{ca.} 1085 Ma.
-Rock magnetic analyses and magnetic imaging reveal that the anorthosite xenoliths can often have magnetic mineralogy dominated by grains with high coercivities, which can yield high success rates passing our selection criteria relative to their diabase hosts.
-Taken together, new and previously published paleointensity data from the Midcontinent Rift support that a strong geomagnetic field likely persisted through 25 million years of the late Mesoproterozoic.
-These data are inconsistent with interpretations of a progressive decay of Earth's geodynamo strength through the Proterozoic Eon.
-
-\end{abstract}
-
-\section*{Introduction}
-
-Earth's magnetic field is the result of convective flow of liquid iron-alloy in Earth's outer core.
-This self-sustaining geodynamo is collectively powered by the heat flow across the core-mantle boundary (CMB) and from the crystallization of the solid inner core from the liquid outer core which provides latent heat and compositional buoyancy.
-Paleomagnetic studies have found that a dynamo may have existed by the Archean \cite{Selkin2007a, Tarduno2014a, Brenner2020a}.
-However, some estimates of the thermal conductivity of the core suggest that the nucleation of the inner core could be younger than 1 Ga, more recent than previous thought \cite{Pozzo2012a, Gomi2013a, Labrosse2015a, Ohta2016a}.
-Without the power of compositional convection provided by the process of light elements being rejected into the outer core from the crystallizing inner core, the geodynamo was likely dominantly driven by thermal conduction through the CMB, with perhaps additional stirring by other power sources such Mg precipitation \cite{ORourcke2016a}, or perhaps there were intervals early in Earth history when the dynamo was generated by exotic processes such as a convecting basal magma ocean \cite{Ziegler2013a}.
-Long-term secular cooling of Earth has the potential to have decreased the power available to the geodynamo through time prior to the nucleation of the inner core \cite<e.g>{Davies2021a}.
-As a result, it is possible that rigor of the geomagnetic axial dipole field would have experienced a long-term decay before the emergence of the solid inner core after which it would have strengthened \cite{Aubert2009a, Davies2021a}.
-However, the estimates for thermal conductivity values for the core continue to vary\cite{Konopkova2016a, Ohta2016a}.
-Obtaining an observational record of the timing of inner core formation would constrain the thermal evolution of the core and mantle, in addition to improving our understanding of the history of Earth’s dynamo.
-
-Paleomagnetic data from ancient rocks are one of the few types of observational data that have the potential to record the evolution of Earth’s core and can be used to test the predictions from numerical simulations.
-Some Precambrian absolute paleointensity records from the PINT database (\url{http://earth.liv.ac.uk/pint/}) show evidence of an existence of an appreciable geomagnetic field prior to the Ediacaran.
-On the other hand, studies have found very low axial dipole moment (VDM) estimates from rocks formed in the late Ediacaran \cite{Bono2019a, Shcherbakova2019a, Thallner2021b}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
-Other studies focusing on paleomagnetic directions have also found that the late Ediacaran might have been a period of time when the geomagnetic field were experiencing frequent reversals and waning of dipole dominance \cite{Bono2015a, Kodama2020a}.
-It has been interpreted that these findings are consistent with the results of geodynamo models using high core thermal conductivity values that suggest the inner core nucleation age being near the Ediacaran-Cambrian boundary \cite{Driscoll2016a, Davies2021a}.
-\citeA{Bono2019a} based on a polynomial fit through filtered Precambrian paleointensity database to suggest the selected observational records show that the geodynamo power progressively decreased through the Proterozoic (Figure \ref{fig:MCR_VADM}).
-That study also recovered very low paleointensity values from \textit{ca.} 565 Ma single silicate crystals and interpreted them to be reflecting that the geodynamo reached its lowest near the Ediacaran-Cambrian boundary.
-New low paleointensity estimates derived from the \textit{ca.} 720 Ma Franklin Large Igneous Province are interpreted to support the existence of such decay trend in the Neoproterozoic \cite{Lloyd2021a}.
-
-However, previous paleointensity data from intrusive and extrusive rock units in the Mesoproterozoic Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record Earth's virtual axial dipole moment (VADM) values similar to that of today, which are higher than the prediction of the long term decay curve (Figure \ref{fig:MCR_VADM}).
-\citeA{Pesonen1983b} obtained an average virtual dipole moment of 93 ZAm$^2$ after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using both extrusive and intrusive rocks, including the \textit{ca.} 1108-1105 Ma Osler Volcanics, the Maimainse Point volcanics, intrusive Logan Sills, Thunder Bay dikes, the Baraga and Marquette dikes, as well as lava flow of the \textit{ca.} 1096 Ma North Shore Volcanic Group.
-\citeA{Kulakov2013a} obtained an average VADM of 59 $\pm$ 11 ZAm$^2$ using the \textit{ca.} 1085 Ma Lake Shore Traps volcanics and suggest that the present-day-like geomagnetic field strength likely existed through the late stage of Midcontinent Rift magmatism \cite{Vervoort2007a, Miller2013a}.
-However, non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results obtained by \citeA{Pesonen1983b, Kulakov2013a}.
-One typical failure of paleointensity experiment is the double-slope behavior, as two distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
-\citeA{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas \citeA{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
-Results with such non-ideal behavior were rejected by \citeA{Sprain2018a} who applied stricter paleointensity selection criteria on new paleointensity results and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
-A calculated mean VADM value of 56 $\pm$ 21 ZAm$^2$ is similar to the average value of the past 300 million years \cite{Sprain2018a}.
-This high VADM value estimate from the Midcontinent Rift volcanics may be incompatible with the hypothesis of a ``Proterozoic dipole low" \cite{Biggin2009a}.
-
-Although it is typical of paleointensity experiments to suffer from low success rates, bulk plagioclase-rich rocks \cite{Selkin2000a} and single plagioclase crystals \cite{Tarduno2005a} are found to be able to produce high quality paleointensity results.
-It has been suggested that plagioclase crystals can protect magnetic minerals that exsolved within them against post-formation weathering and thermochemical alteration during paleointensity experiments \cite{Tarduno2005a}.
-In this study, we present new paleointensity results from ``IZZI" modified absolute paleointensity experiments \cite{Yu2004a} on a suite of bulk samples from \textit{ca.} 1092 Ma anorthosite xenoliths and their Beaver River diabase host rocks of the Beaver Bay Complex within the Midcontinent Rift.
-Paleointensity experiments on the Beaver River anorthosite xenoliths have a very high success rate.
-The consistent specimen- and site-level paleointensity results reveal that the magnetite-dominated remanence of the anorthosite specimens has low thermal remanent magnetization (TRM) anisotropy and linear TRM acquisition.
-Magnetic imaging shows that the anorthosite specimens to have dominant magnetic carriers interstitial to and within plagioclase crystals and they do not display strong preferred orientations.
-With more single-slope, high quality paleointensity data, we seek to compare paleointensity records from the Midcontinent Rift intrusives with those from the extrusives; to test the hypothesis that the Earth's geomagnetic dipole strength remained at high values throughout the Midcontinent Rift; and to further enrich the reliable paleointensity records during the $\sim$25 myr of Midcontinent Rift magmatic activity.
-
-%Retrieving paleomagnetic information from Precambrian rocks can aid in our understanding of the long-term evolution of the geodynamo.
-A recent paleointensity study by \cite{Sprain2018a} reported a \textit{ca.} 1.1 Ga paleomagnetic field strength similar to today based on the volcanics of the Midcontinent Rift System (MRS).
-That study suggested that this result is consistent with models wherein this time period of the Proterozoic is characterized by a strong rather than weak geomagnetic field.
-However, to further evaluating the evolution of the geomagnetic field during the Precambrian, more paleointensity data are needed.
-
-%The MRS emplaced a vast amount of extrusive and intrusive rocks spanning a prolonged period of $\sim$25 Myrs.
-\cite{Swanson-Hysell2019a} published an extensive compilation of the MRS volcanics with paired high-precision CA-ID-TIMS U-Pb geochronology and high-quality paleomagnetic directions, revealing at least two dipole reversals and rapid differential plate motion of the ancient North American plate, Laurentia, during the emplacement of MRS.
-The synthesized North American apparent polar wander path (APWP) was interpreted to be composed primarily of rapid plate motion \citep{Swanson-Hysell2019a}.
-Through comparison between paleomagnetic pole positions and the APWP, such rapid plate motion and high-precision paleomagnetic and geochronological framework revealed by the MRS volcanics can provide insights into age constraints on units with poor or no geochronology data.
-This framework allows for further paleomagnetic contributions beyond the Midcontinent Rift volcanics.
-
-\section*{Geologic Settings}
-% a map of the lake superior including the pesonen units, the lake shore traps, the BRD, and the osler and mamainse, together with date compilation.
-
-The Mesoproterozoic North American Midcontinent Rift (MCR) is an intracontinental rift that did not result in the break-up of Laurentia.
-Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; \citeA{Vervoort2007a, Miller2013a}) was punctuated by rapid and voluminous emplacement of intrusive rocks as well as extrusive volcanics.
-Based on high-precision $^{206}$Pb/$^{238}$U geochronology and paleomagnetic data, \citeA{Swanson-Hysell2021a} hypothesized that the massive Duluth Complex and the coeval North Shore Volcanic group (Figure \ref{fig:Geologic_map}) formed as a result of having upwelling mantled-derived plume material encountering Laurentia's thinned lithosphere, following an ``upside-down” drainage" which rapidly funneled melt to today's North Eastern Minnesota (Figure \ref{fig:Geologic_map}; \citeA{Sleep1997a, Swanson-Hysell2014a}).
-The plagioclase-phyric Anorthositic Series rocks, the crystal-poor Layered Series intrusions of the Duluth Complex, and the coeval volcanics of the North Shore Volcanic Group (NSVG) that outcrop along Northeastern Minnesota (Figure \ref{fig:Geologic_map}) are interpreted to be a result of such vigorous magmatic activity.
+% \subsection*{Methods}
+% Add a methods subsection if you need to.
 
 \begin{figure}
+\noindent\includegraphics[width=\textwidth]{MPMS.pdf}
 \centering
-\noindent\includegraphics[width=\textwidth]{Geologic_map.pdf}
-\caption{\footnotesize{Simplified geologic map of Lake Superior region showing the distribution of rift and post-rift-related rocks including extrusive, intrusive, and late/post-rift sediments around Lake Superior.
-Red diamonds mark sites with paleointensity results that passed the selection criteria from this study.
-Paleomagnetic sites from \citeA{Pesonen1983a} are categorized by lithology.
-All sites from previous studies are color-coded by their ages.}}
-\label{fig:Geologic_map}
+\caption{{Low-temperature magnetic property measurement system (MPMS) experiment results.
+In the field-cooled (FC) experiments, the magnetization was measured upon warming following the specimen having cooled in an applied field of 2.5 T from 300 to 10 K.
+In the zero-field-cooled (ZFC) experiment, a low-temperature saturation isothermal remanence (LTSIRM) of 2.5 T was applied at 10 K after the specimen cooled in a (near-)zero field.
+In the room-temperature saturation isothermal remanence (RTSIRM) experiment, the sample was pulsed with a 2.5 T field at room temperature ($\sim$300 K) and then cooled to 10 K and warmed back to room temperature in a (near-) zero field.
+Specimen AX6-1r is from anorthosite AX6 which passed our paleointensity selection.
+It has a well-defined Verwey transition $\sim$120 K \cite{Verwey1939a}.
+Specimens AX3-2r-ni and BD2-2r show Verwey transition but the transition temperatures are suppressed below $\sim$120 K.
+Specimen BD2-2r has a consistently higher moment during the zero-field-cooled step than during the field-cooled step.
+This is consistent with the interpretation that multidomian magnetic carriers exist in significant quantity in this specimen.}}
+\label{fig:MPMS}
 \end{figure}
 
-The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex punctuates another period of magmatism during the Midcontinent Rift main stage activity.
-$^{206}$Pb/$^{238}$U zircon dates and cross-cutting relationships tightly constraints the areally extensive Beaver River diabase of the Beaver Bay Complex to have been emplaced at 1091.7 $\pm$ 0.2 Ma \cite{Zhang2021b}.
-The Beaver River diabase intruded into hypabyssal depths via wide conduits that accommodated the entrainment of numerous anorthosite xenoliths that have short-axis diameters up to 180 meters \cite{Boerboom2004a, Boerboom2006b}.
-Geochronological, paleomagnetic, geochemical, and petrographic data are consistent with the hypothesis that the Beaver River diabase magma extruded to the surface, feeding the massive Greenstone Flow - one of the largest lava flows on Earth \cite{Doyle2016a, Zhang2021b}.
+\clearpage
 
-% The anorthosite xenoliths that distinctly occur within the Beaver River diabase are not only recorders of the wide diabase conduit sizes,  of the geomagnetic field $\sim$1091.7 million years ago.
-Thermal modeling result and paleomagnetic directional data show the anorthosite xenoliths acquired thermal remanent magnetizations during cooling of the diabase \cite{Zhang2021b}.
-Step-wise thermal demagnetizations show many anorthosite xenoliths have minimal secondary remanence and single-component characteristic remanence magnetizations that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C.
-
-Despite that the anorthosite xenoliths of the Beaver River diabase were entrained as xenoliths and emplaced only $\sim$4 myr after the formation of the Anorthositic Series rocks of the Duluth Complex when massive plagioclase mushes were generated, $^{206}$Pb/$^{238}$U zircon geochronology data and zircon rare earth element diffusion modeling results suggest it is the most probable that the Beaver River diabase anorthosite xenoliths are entrained cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
-Moreover, the anorthosite xenoliths of the \textit{ca.} 1092 Ma Beaver River diabase has distinct mineralogy from that of the Anorthositic Series rocks of the Duluth Complex.
-The dominantly monomineralic anorthosite xenoliths have an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}, higher than the average of $\sim$60\% of the Anorthositic Series rocks of the Duluth Complex \cite{Miller1990a}.
-In addition, petrographic imaging support that the plagioclase from the Beaver River anorthosite formed under a magmatic conditions different from the Duluth Complex anorthositic series magma.
-While the plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture (Figure \ref{Petro_QDM}A), the more calcium-rich plagioclase of the Beaver River anorthosite often develop granoblastic texture characterized by equigranular plagioclase crystals (Figure \ref{Petro_QDM}B).
-Abundant magnetite and ilmenite needles that are typically 10 $\mu$m long can be found to have preferred orientations within Duluth Complex plagioclase.
-Interstitial vermicular titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex Anorthositic Series rocks \cite{Miller1990a}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{Petro_QDM.png}
-\caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from Duluth Complex Anorthositic Series (left column) and Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (right column; \citeA{Zhang2021b}).
-Petrographic images (A-D) show Duluth Complex anorthosite have euhedral, elongate plagioclase crystals with interlocking texture and igneous foliation.
-(C) The plagioclase can contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the (010) planes of plagioclase.
-In contrast, plagioclase of sample MS99033 do not enclose large oxide needles.
-Magnetic field maps acquired by the quantum diamond microscope (QDM) show that the oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the overlapping magnetic dipole fields exerted above the oxides in (E).
-In comparison, relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite.
-The dipole-like features in (F) do not show strong preferred orientations.
-Magnetic field maps of (E) and (F) share the same spatial and magnetic scale.
-The unit of the axes are in $\mu$m.}}
-\label{fig:Petro_QDM}
-\end{figure}
-
-Occurrences of magnetite and ilmenite symplectites and needles within anorthositic rocks have been described by \citeA{Scofield1975a, Scofield1986a} in the Wichita Mountains and by \citeA{Haggerty1976a} in other slowly cooled, mafic intrusions.
-\citeA{Scofield1986a} hypothesized that the inclusions acted as a means of buffering the melt with respect to oxygen, forming either crystallites along preexisting plagioclase substrate or as iron incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures.
-Such a model requires that plagioclase, instead of other mafic minerals such as olivine, to be the first to crystallize from a magma and act as a sink for iron.
-The two-stage formation of the Duluth Complex Anorthositic Series fits this model \cite{Miller1990a}.
-During the first stage, high pressure at subcrustal depths (15-25 km depth) caused fractionation of Duluth Complex magma resulting in the formation of floating plagioclase mushes \cite{Kushiro1980a}.
-The second stage of crystallization then occurs at low pressures after the magma intruded into shallower chambers.
-
-This proposed magmatic evolution of the Duluth Complex is in distinct contrast to the emplacement of the Beaver River diabase which rapidly ascended and entrained the anorthosite xenoliths directly into shallow chambers \cite{Flower1980a, Miller1997a, Zhang2021b}.
-Particularly, \citeA{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
-This is consistent with the interpretation that these anorthosite were rapidly transported by upward currents into hypabyssal magma chambers without reaching structural equilibrium.
-Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Figure \ref{Petro_QDM}).
-Instead, relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Figure \ref{Petro_QDM}).
-
-% Previous studies have divided magmatic activity in the rift into four stages based on interpreted changes in relative magmatic volume and the nature of magmatism: early ($\sim$1109–1104 Ma), latent ($\sim$1104–1098 Ma), main ($\sim$1098–1090 Ma) and late ($\sim$1090–1084 Ma)  In particular, the main stage magmatism is characterized by rapid and punctuated events which resulted in voluminous emplacement of intrusive rocks as well as coeval outpouring of extrusive lava flows.
-One example is the massive \textit{ca.} 1096 Ma Duluth Complex and the North Shore Volcanic Group \cite{Swanson-Hysell2020a}, the majority of which emplaced within $<$500 kyrs.
-Another one is the \textit{ca.} 1092 Ma Beaver Bay Complex.
-The Beaver River diabase intrusions of the complex formed synchronously with the Greenstone Flow of the Portage Lake Volcanics which is one of the largest lava flows known on Earth \cite{Zhang2021b}.
-Excellent rock preservation thanks to minimal post-rift alteration in the Midcontinent Rift have yielded high-quality paleomagnetic directional data which have now been paired with high-precision zircon geochronology to reconstruct the ``Keweenawan Track" - a well-resolved Laurentia's (cratonic North America) apparent polar wander path during the late Mesoproterozoic \cite{Swanson-Hysell2019a}.
-
-\section*{Method}
-
-\subsection*{Sample collection and paleomagnetic directions}
-
-We collected paleomagnetic cores that are 2.5 cm in diameter along the southern and eastern Beaver Bay Complex with a particular focus on acquiring paired sites of anorthosite xenoliths and their local diabase hosts during summer field seasons in 2019 and 2020.
-Sample cores were collected using a hand-held gasoline-powered drill and were oriented using a magnetic compass as well as a sun compass when possible.
-Sun compass orientations were preferentially used for determining the sample azimuth.
-Specimens from a total of 7 diabase and 14 anorthosite sites were used for paleointensity experiments (Table \ref{tab:Pmag_site_data}).
-Sister specimens underwent step-wise alternating field (AF) or thermal demagnetization at the UC Berkeley Paleomagnetism Lab and the results are presented in \citeA{Zhang2021b}.
-The close directional similarities between the anorthosite xenoliths and their host diabase and thermal modeling results are consistent with the interpretation that the Beaver River diabase and anorthosite xenoliths acquired primary thermal remanent magnetization \textit{ca.} 1092 Ma during cooling after the emplacement of the diabase \cite{Zhang2021b}.
-Based on the anorthosite thermal demagnetization results we selected sites whose unblocking temperature ranges are narrow and near 580\textdegree C for paleointensity experiments.
-Beaver River diabase sites with minimal secondary remanence were also selected for paleointensity experiments.
-
-\subsection*{Paleointensity experiment}
-A total of 86 specimens from 14 anorthosite xenoliths and a total of 69 specimens from 7 diabase sites underwent paleointensity experiments that followed the step-wise double-heating Thellier method \cite{Thellier1959a} using the IZZI protocol \cite{Tauxe2004a} with heating steps up to 585 \textdegree C.
-Partial thermal remanent magnetization (pTRM) checks were performed systematically throughout the experiment to test whether there was significant mineralogical alteration due to heating and were assessed using the SCAT parameter of \citeA{Shaar2013a}.
-On top of the IZZI modification of the Thellier experimental protocol, we also performed a comparative study where we added an extra step of 20 mT alternating field (AF) cleaning on some of the specimens after each in-field step.
-The purpose is to study whether the AF cleaning could help improve experiment success rate by removing the remanence component carried by materials such as multi-domain (MD) or vortex state grains that often contributes to non-ideal paleointensity behaviors.
-All remanence measurements were made on a 2G Enterprises DC-SQUID superconducting rock magnetometer equipped with an automated sample changer system at the UC Berkeley Paleomagnetism laboratory.
-The magnetometer is housed inside a three-layer magnetostatic shield that maintains background fields of less than 500 nT.
-Heating steps were performed using an ASC TD-48SC thermal demagnetizer with a controlled field coil that allows for a magnetic field to be generated in the oven in conjunction with a DC power supply.
-The thermal demagnetizer was degaussed with an alternating field in the axial orientation following each in-field step such that residual fields within the oven were $<$10 nT during zero-field steps.
-Samples were placed in the same location within the thermal demagnetizer for each heating step and were maintained in the same orientation with regard to the applied field.
-During each heating step, samples remained at peak temperatures for 20 min.
-An applied laboratory field of 30 $\mu$T was used for all in-field steps.
-All heating steps were performed in air.
-The temperature increments for the experiments were chosen to cover characteristic remanent magnetizations held by (titano)magnetite informed by the previous demagnetization data, with smaller increments performed close to the expected unblocking temperature of stoichiometric magnetite \cite{Zhang2021b}.
-
-% After the paleointensity experiment were finished, the same specimens underwent additional TRM anisotropy checks with full TRMs given at 0, 30, 60, 90, 120, 150 degrees in the horizontal plane in specimen coordinates.
-After that, full TRM acquisition experiments were performed on the same specimens with applied fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axis.
-
-\subsection*{Paleointensity result selection}
-The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD; \citeA{Kirschvink1980a}) of $<$20\textdegree; (2) scatter parameter ($\beta$; \citeA{Coe1978a}) values of $<$15 per cent; (3) a deviation angle (DANG; \citeA{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC; \citeA{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT; \citeA{Shaar2013a}) = TRUE; (6) a maximum magnetic moment difference between adjacent zero-field steps (GAP-Max; \citeA{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
-The criteria are also summarized in Table \ref{tab:criteria}.
-The MAD measures the scatter about the best-fit line through the natural remanent magnetization (NRM) steps in the selected interval for which the intensity is defined.
-DANG, the deviation angle, is the angle between the best-fit direction that is free floating and the direction between the centre of mass of the data and the origin of the vector component diagram \cite{Tauxe2004a}.
-Both MAD and DANG assess the directional variation of the NRM, with MAD measuring the scatter in the NRM directions and DANG assessing whether the component is trending toward the origin of the Zijderveld plot.
-$\beta$ is the ``scatter" parameter of \citeA{Coe1978a} and is the ratio of the standard error of the slope of the best-fit line of the selected NRM and pTRM points on an Arai plot to the absolute value of the slope.
-FRAC is the fraction of the NRM that is used in the best-fit line \cite{Shaar2013a}.
-The FRAC value was chosen to preferentially select samples with dominantly single-slope Arai plots.
-GAP-Max is the maximum gap between two points on the Arai plot determined by vector arithmetic.
-SCAT is a Boolean operator which uses the error on the best-fit slope of the selected data on the Arai plot to determine if the data are overly scattered.
-The parameter is used to assess pTRM checks in addition to assessing the degree to which IZZI steps are zigzagged.
-$\beta$, FRAC, GAP-Max and SCAT are all statistics to assess the behavior of Arai plots.
-See the Standard Paleointensity Definitions (\citeA{Paterson2014a}; \url{https://earthref.org/PmagPy/SPD/home.html}) for more details.
-Data analysis was conducted using Thellier GUI \cite{Shaar2013a} within the PmagPy software package \cite{Tauxe2016a}.
-
-\begin{table}
-\caption{Quality criteria used for selecting paleointensity data. See the text for more detail.} 
+\begin{figure*}[h!]
+\noindent\includegraphics[width=17.8 cm]{manuscript/Cooling_rate_correction.pdf}
 \centering
-\begin{tabular}{ccccccccc}
-\hline
-MAD (\textdegree) & Beta (per cent) & DANG (\textdegree) & FRAC& SCAT & GAP-MAX & NpTRM & N Arai\\
-\hline
-20 & 15 & 5 & 0.6 & TRUE & 0.25 & 2 & 4 \\
-\hline
-
-\end{tabular}
-\label{tab:criteria}
-\end{table}
-
-\subsection*{Rock magnetic experiment}
-
-In this study, we conduct rock magnetic experiments with a purpose of gaining magnetic mineralogy insights into the paleointensity results of the anorthosite and diabase.
-Backfield curves were measured at room temperature using a Micromag Princeton Measurements vibrating sample magnetometer (VSM) and a Lake Shore 8600 series VSM.
-Specimen median destructive field (MDF) are calculated based on the backfield curves.
-The calculated coercivity spectra were subsequently decomposed into one ore more components using skew-normal distributions following the method of \citeA{Maxbauer2016a}.
-The fitted components are then used to interpret the characteristics of different populations of magnetic particles within specimens.
-We also used the magnetic property measurement system (MPMS) at the IRM to aid in the identification of magnetic mineral.
-In the field-cooled (FC) experiments, specimen magnetizations were measured upon warming following the specimen having cooled in an applied field of 2.5 T from 300 to 10 K.
-In the zero-field-cooled (ZFC) experiment, a low-temperature saturation isothermal remanence (LTSIRM) of 2.5 T was applied at 10 K after the specimen cooled in a (near-)zero field.
-In the room-temperature saturation isothermal remanence (RTSIRM) experiment, the sample was pulsed with a 2.5 T field at room temperature ($\sim$300 K) and then cooled to 10 K and warmed back to room temperature in a (near-)zero field.
-The magnetic moment transitions at critical temperatures revealed through MPMS experiments are used to identify magnetic minerals such as magnetite \cite{Verwey1939a} within specimens.
-
-To further identify the magnetic carriers within the Beaver River anorthosite xenoliths and compare them with the anorthosites of the Duluth Complex Anorthositic Series rocks, we used the quantum diamond microscope (QDM) at the UC Berkeley Paleomagnetism laboratory to image a thin section of sample MS99033 which is a hand sample from anorthosite xenolith AX16 (which yielded a $^{206}$Pb/$^{238}$U zircon date of 1091.83 $\pm$ 0.21 Ma; \citeA{Zhang2021b}), and a thin section of a Duluth Complex anorthosite (Figure \ref{fig:Petro_QDM}).
-We the QDM to image the magnetic field over the polished thin section surfaces at a sample-sensor distance of 5 $\mu$m in projective magnetic microscopy (PMM) mode with a spatial resolution of 4.7 μm per pixel and an instantaneous 0.9 mT bias field that is canceled during the course of measurement \cite{Glenn2017a}).
-
-\section*{Results and Discussion}
-\subsection*{Paleointensity}
-
-40 from a total of 86 anorthosite specimens and 0 out of a total of 69 diabase specimens passed our paleointensity selection criteria, yielding 7 anorthosite sites and no diabase sites with site-level paleointensity results that pass paleointensity selections.
-Example NRM/TRM plots (Arai plots;
-\cite{Arai1963a}) are shown in Figure \ref{fig:IZZI_examples}.
-Summary specimen absolute paleointensity estimates and site-level weighted mean paleointensity values are shown in Table \ref{tab:PINT_result} and are plotted in Figure \ref{fig:PINT_cooling_corrected}.
-All measurement-level paleointensity experiment data can be accessed through the MagIC database (\url{}; UPDATE WHEN DOI IS GENERATED).
-
-\begin{figure}[h!]
-\noindent\includegraphics[width=\textwidth]{IZZI_examples.pdf}
-\centering
-\caption{\footnotesize{Example results of paleointensity experiments displayed on Arai and Zijderveld plots (insets) for anorthosite and diabase specimens.
-Red (blue) circles indicate zero-field/in-field (in-field/zero-field) steps `ZI’ (`IZ’).
-Triangles mark partial thermal remanent (pTRM) checks.
-Blue and red squares in the Zijderveld plots are X–Y and X–Z projections, respectively, of the NRMs in specimen coordinates.
-Plots on the top row show successful specimen paleointensity results with straight, single-slope behaviors that pass our selection criteria.
-The green lines represent our fitting for the dominant single-slope component that passes our acceptance criteria with estimates of the ancient field (B$_{anc}$) being used to determine site means.
-The plots for anorthosite specimens AX1-1a and diabase BD5-1b on the bottom row show typical non-ideal sagging and zigzagging behaviour that fails our acceptance criteria.
-Plot for diabase specimen BD2-5a shows typical paleointensity behavior of this site - specimens typically have double-slope behaviors although the lower temperature slopes can pass the selection criteria, resulting in the high estimated B$_{anc}$ values shown in Figure \ref{fig:PINT_cooling_corrected}.
-Thus, the estimates of B$_{anc}$ for this specimen plot is illustrative, but were not used to calculate a site mean paleointensity value in Figure \ref{fig:PINT_cooling_corrected}.}}
-\label{fig:IZZI_examples}
-\end{figure}
-
-The Beaver River anorthosite specimens yielded a high pass rate ($\sim$46.5\%) through our selection criteria.
-In distinct contrast, the Beaver River diabase specimens are all rejected by our selection criteria.
-Paleointensity experiment results from both lithologies show three types of dominant specimen behaviors on Arai plots (Figure \ref{fig:IZZI_examples}; \citeA{Arai1963a}): dominantly single-slope, double-slope, and zigzagging or sagging (Figure \ref{fig:IZZI_examples}).
-Note that almost all specimens have minimal secondary remanence component in thermal demagnetization results as shown in example orthogonal plots of all zero-field steps (Figure \ref{fig:IZZI_examples}).
-Typical specimen paleointensity behavior of the anorthosites display straight, single-slope Arai plots and the accepted fractions of temperature steps span over the origin-trending, primary remanence components (Figure \ref{fig:IZZI_examples}).
-We accept all specimen- and site-level absolute paleointensity estimates from the anorthosite xenoliths that pass the selection.
-In terms of the diabase specimens, all failed the selection because of double-slope behavior (fail FRAC selection), poor pTRM checks, and zigzagging (fail SCAT, DANG selection) Arai plots (Figure \ref{fig:IZZI_examples}).
-As is shown in Figure \ref{fig:IZZI_examples}, specimen BD2-5b has a high-temperature second slope with zigzagging behavior, poor pTRM checks, and low NRM/TRM ratio, in addition to the straight slope with steep NRM/TRM ratio over the lower temperature range.
-Although the BD2 specimens' lower temperature fraction of the Arai plots are straight andthe pTRM checks are consistent, lower temperature slopes from BD2 give consistently higher paleointensity estimates than those from site AX6, AX8, and AX10, which are anorthosite xenoliths included within BD2 (Figure \ref{fig:Geologic_map}; \citeA{Zhang2021a}).
-This higher estimate from the diabase lithology could be attributed to the systematic bias of using the lower temperature slope of the Arai plots demonstrated in Figure \ref{fig:IZZI_examples} and specimens from site BD2 all fail the criteria of having a maximum magnetizatiom gap between adjacent steps of 25\% of the total NRM.
-Thus we only accept the single-slope anorthosite specimens results for further analyses.
-
-% \begin{figure}[h!]
-% \noindent\includegraphics[width=\textwidth]{Cooling_rate_correction.pdf}
-% \centering
-% \caption{\small{}}
-% \label{fig:cooling_correction}
-% \end{figure}
-
-\begin{figure}[h!]
-\noindent\includegraphics[width=\textwidth]{Paleointensity_plot_cooling_corrected.pdf}
-\centering
-\caption{\small{Summary plot of cooling rate corrected individual specimen absolute paleointensity results (square and diamond symbols) from this study and their averages at site (grey bar) and locality level (weighted mean by site-level standard deviations; orange bar).
-The left plot shows the graph of predicted paleointensity overestimate due to slow cooling of the intrusive Beaver River diabase and anorthosite xenoliths relative to the cooling rate in laboratory following the model of \citeA{Halgedahl1980a}.
+\caption{{Graph of predicted paleointensity overestimate due to slow cooling of the intrusive Beaver River diabase and anorthosite xenoliths relative to the cooling rate in laboratory following the model of \citeA{Halgedahl1980a}.
 Because the majority of the anorthosites have unblocking temperatures between 500\textdegree C and 580\textdegree C, we estimate that the slow cooling during natural remanence acquisition could have resulted in a 36\% overestimate.
-Thus a correction factor of 0.74 is applied at specimen level for the paleointensity summary plot.
+Thus, a correction factor of 0.74 is applied at specimen level for the paleointensity summary plot.
 }}
 \label{fig:PINT_cooling_corrected}
-\end{figure}
+\end{figure*}
 
-The 20 mT alternating field treatment after in-field heating steps did not make any diabase specimen to pass the selection criteria.
-On the other hand, 27 out of 58 ($\sim$46.6\%) anorthosite specimens that had AF treatment passed selection and 13 out of 28 ($\sim$46.4\%) specimens without AF treatment passed selection.
-Paleointensity estimates from specimens with or without AF treatment in anorthosite AX10, AX11, AX12, AX13, AX16 do not show significant difference (Figure \ref{fig:PINT_cooling_corrected}).
-Specimen plots for AX11 particularly illustrates a case where all specimens within the site yield almost the same paleointensity estimate, regardless of the AF treatment.
-Therefore, the alternating field demagnetization might have improve the specimen behavior for those anorthosites that were subjected to this treatment but it did not result in significant increase in specimen success rate and did not result in significant difference for the site-level paleointensity estimates.
-Together with the narrow range of unblocking temperatures of their sister specimens \cite{Zhang2021b}, these data are consistent with the interpretation that the anorthosites with successful paleointensity results have a magnetic mineral assemblage that likely contain dominantly single domain, low Ti magnetite and they do not contain a significant population of MD or vortex state grains.
-Some anorthosite specimens and all diabase specimens still failed in result selections even though AF treatment was applied.
-It could be that they have a very high dominance of MD or vortex state grains with relatively moderate coercivity such that an alternating field with 20 mT peak field did not significantly mitigate their non-ideal paleointensity behaviors.
-Another cause for their failure could be the thermochemical alterations that happened during heating in air.
-As can be exampled by the zigzagging high temperature slope with poor pTRM checks in paleointensity plot for specimen BD2-5b (Figure \ref{fig:IZZI_examples}), the decrease in NRM/TRM ratios during higher temperature steps likely resulted from the formation of new ferromagnetic material within the specimen due to oxidation.
-Such alterations can allow specimens to acquire additional remanence during in-field steps that may not be fully removed by an AF treatment.
-Overall, the consistent specimen-level behavior makes the successful results from the anorthosites of high confidence.
-The paleointensity quality index (Q$_{PI}$) for the anorthosite sites are all above 3 and the summary statistics are shown in Table \ref{tab:QPI}.
-
-In addition to applying paleointensity selection criteria to our experiment results, we applied the bias corrected estimation of paleointensity (BiCEP) method by \cite{Cych2021a} to all of our anorthosite and diabase specimens regardless of their Arai plot statistics.
-The BiCEP method is a Bayesian statistical method based on an assumption that all specimen paleointensity results from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the Arai plot \cite{Paterson2011a}.
-This method has the advantage of using as many specimens as available for estimating site-level paleointensities.
-The posterior distributions of paleointensity estimates given by the BiCEP method show that for the anorthosites that pass our paleointensity selection criteria, probabilities are very low for the case where these anorthosites are recording heavily biased paleointensity estimates from a true value below 10 $\mu$T (Supporting Information S1).
-Therefore, the anorthosite xenoliths record a high geomagnetic field \textit{ca.} 1092 Ma.
+\clearpage
 
@@
 SCAT is the scatter parameter.
-Paleolatitude is calculated from the inclination values reported in \citeA{Zhang2021b}.
+Paleolatitude is calculated from the inclination values reported in \cite{Zhang2021b}.
 $\gamma$ is the gamma statistic that measures the angle between the last pTRM step used for paleointensity determination and the applied field direction.
@@
 
-\begin{table}[]
+\begin{table}
 
-\caption{\footnotesize{Summary statistics for the Q$_{PI}$ quality criteria of \citeA{Biggin2014a}.}}
+\caption{\footnotesize{Summary statistics for the Q$_{PI}$ quality criteria of \cite{Biggin2014a}.}}
 \centering
@@
 
-\subsection*{Rock magnetism}
-Rock magnetic data support that the successful anorthosite specimens have dominant magnetic carriers with magnetic properties similar to stochiometric, non-interactive, single domain magnetite, whereas those anorthosites that failed the paleointensity result selection and all diabase have more pronounced populations of non-ideal carriers.
-MPMS data show sister specimens to paleointensity specimens typically have dominant magnetic mineralogy of magnetite as can be evidenced through the Verwey transitions (Figure \ref{fig:MPMS}; \citeA{Verwey1939a}).
-However, specimens from sites that yield successful paleointensity results display Verwey transition near 120 Kelvin (K) as expected for stochiometric low-Ti magnetite \cite{Ozdemir1993a}.
-On the other hand, the diabase and anorthosite specimens that did not pass our paleointensity selection typically have suppressed Verwey transition toward lower temperatures (Figure \ref{fig:MPMS}), indicating that the specimens likely have a ferromagnetic mineralogy with variable degrees of oxidation \cite{Ozdemir1993a}.
+% \begin{table}\centering
+% \caption{This is a table}
 
-\begin{figure}[h!]
-\noindent\includegraphics[width=\textwidth]{MPMS.pdf}
-\centering
-\caption{\footnotesize{Low-temperature magnetic property measurement system (MPMS) experiment results.
-In the field-cooled (FC) experiments, the magnetization was measured upon warming following the specimen having cooled in an applied field of 2.5 T from 300 to 10 K.
-In the zero-field-cooled (ZFC) experiment, a low-temperature saturation isothermal remanence (LTSIRM) of 2.5 T was applied at 10 K after the specimen cooled in a (near-)zero field.
-In the room-temperature saturation isothermal remanence (RTSIRM) experiment, the sample was pulsed with a 2.5 T field at room temperature ($\sim$300 K) and then cooled to 10 K and warmed back to room temperature in a (near-) zero field.
-Specimen AX6-1r is from anorthosite AX6 which passed our paleointensity selection.
-It has a well-defined Verwey transition $\sim$120 K \cite{Verwey1939a}.
-Specimens AX3-2r-ni and BD2-2r show Verwey transition but the transition temperatures are suppressed below $\sim$120 K.
-Specimen BD2-2r has a consistently higher moment during the zero-field-cooled step than during the field-cooled step.
-This is consistent with the interpretation that multidomian magnetic carriers exist in significant quantity in this specimen.}}
-\label{fig:MPMS}
-\end{figure}
+% \begin{tabular}{lrrr}s
+% Species & CBS & CV & G3 \\
+% \midrule
+% 1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
+% 2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
+% 3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
+% \bottomrule
+% \end{tabular}
+% \end{table}
 
-Specimen coercivity spectra further support that our paleointensity selection criteria are successful in selecting experiment results.
-A compilation of median destructive field (MDF) in Figure \ref{fig:coercivity} based on all of our backfield experiments shows that specimens from anorthosites that pass paleointensity selections have distinctly a higher average MDF range than other anorthosites and all diabase.
-Example single component skew-normal distribution fits of coercivity spectra based on backfield experiments on specimen chips sister to paleointensity specimens show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity distributions around 80 mT (Figure \ref{fig:coercivity}).
-In contrast, specimens from anorthosite that did not pass paleointensity selection, as well as diabase specimens, tend to have coercivity distributions over a significantly lower range ($\sim$30 mT).
-This is consistent with the interpretation that they lack a dominance of single domain magnetite grains in their ferromagnetic mineral assemblages.
+%%% Add this line AFTER all your figures and tables
+\FloatBarrier
 
-\begin{figure}[h!]
-\noindent\includegraphics[width=\textwidth]{coercivity.pdf}
-\centering
-\caption{\footnotesize{Top: Example coercivity unmixing plots on anorthosites and diabase specimens from sites that pass or fail our paleointensity selection criteria.
-Bottom: Box plots summarize distributions of median destructive field (MDF) values for all anorthosite and diabase specimens that have single-component coercivity unmixing results.
-Both plots show that anorthosite specimens that pass our paleointensity selection criteria have distinctly higher MDF than other anorthosites that do not pass the selection, as well as all diabase specimens.}}
-\label{fig:coercivity}
-\end{figure}
+% \movie{Type legend for the movie here.}
 
-\subsection*{Remanence anisotropy and nonlinear TRM acquisition check}
-Significant remanence anisotropy has been documented to exist within anorthositic rocks that form in layered intrusive complexes \cite{Selkin2000a, Feinberg2006a}.
-\cite{Selkin2000a} demonstrated that strong remanence anisotropy associated with the igneous foliation developed within an anorthosite sample from the Stillwater Complex can lead to paleointensity estimates that are different by three times.
-Depending on the orientation of the applied field during in-field steps and that of the magnetic anisotropy axes of a specimen, such strong preferred orientations of petrofabrics can result in significantly biased paleointensity estimates.
-To assess whether our paleointensity estimates are biased by remanence anisotropy, we calculated the gamma statistic, which is the angular difference between the last pTRM step used for paleointensity determination and the applied field direction (Table \ref{tab:PINT_result}).
-The results show all anorthosite specimens used for paleointensity experiment have gamma values ranging from $\sim$0.9\textdegree-$\sim$11.9\textdegree, with a median value of 4.2\textdegree (Table \ref{tab:PINT_result}).
-These gamma values are similar to those of the Midcontinent Rift volcanics \cite{Sprain2018a} and they show that the Beaver River anorthosite bulk rocks do not have a significant remanence anisotropy.
-This result can be further supported by the fact that the recovered paleodirectional data from our anorthosite xenoliths closely match those from the paired Beaver River diabase hosts \cite{Zhang2021b}.
-\cite{Selkin2007a} also noted that vortex state magnetic grains and certain single domain particles can produce biased paleointensity estimates due to their acquisition of saturation remanence during laboratory in-field steps.
-According to the numerical modeling of \citeA{Selkin2007a}, thermoremanent magnetization acquired by a randomly-oriented population of elongate single-domian-sized magnetite particles are unlikely to saturate upon our applied field of 30 $\mu$T.
-Therefore, in this study we do not apply any anisotropy or nonlinear TRM correction to our specimens.
+% \movie{Type legend for the other movie here. Adding longer text to show what happens, to decide on alignment and/or indentations.}
 
-% We applied full TRMs to our anorthosite specimens with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axis.
-The results show that our anorthosite specimens did not saturate upon an applied field of 30 $\mu$T and there is thus no need for correcting nonlinear thermal remanence acquisitions.
-% Further, we applied full TRMs to the anorthosite specimens that pass the paleointensity selection criteria after paleointensity experiment.
-With an applied field of 30 $\mu$T oriented toward the specimen at 0, 30, 60, 90, 120, and 150 degrees, the anorthosite specimens usually acquired consistent full TRMs.
-The most significant variations observed in acquired full TRMs reach to only XXX \%, much less than that has been reported to occurred within layerd anorthositic rocks in Stillwater (up to $\sim$60 \%, \citeA{Selkin2000a}).
+% \movie{A third movie, just for kicks.}
 
-The reason that the Beaver River anorthosite xenoliths do not display strong remanence anisotropy, nor do they display nonlinear TRM acquisition problem, could be that they formed in a magmatic setting distinct from that of a layered mafic complex, such as the nearby Duluth Complex, where igneous foliation can occur (Figure \ref{fig:Geologic_map}).
-Petrographic images show that an equigranular texture where equidimensional, randomly oriented plagioclase often develop in the anorthosite xenoliths (Figure \ref{fig:Petro_QDM}; \citeA{Zhang2021b}).
-This is in contrast to the interlocking igneous foliation texture within the Anorthositic Series rocks of the \textit{ca.} 1096 Ma Duluth Complex (Figure \ref{fig:Petro_QDM}).
-Further, petrographic and magnetic imaging also shows the Beaver River anorthosite xenoliths do not develop abundant exsolved magnetite and ilmenite needles within plagioclase crystals as the slowly-cooled anorthositic intrusions do, in which exsolved magnetite needles can be the dominate remanence carriers (Figure \ref{fig:Petro_QDM}; \citeA{Scofield1986a, Feinberg2006a}).
-The fact that all dipole features within a same Duluth Complex plagioclase crystal have indistinguishable orientations is consistent with the interpretation of there being a magnetic remanence anisotropy resulting from preferred petrofabric alignment, which could give rise to the remanence anisotropy and nonlinear TRM acquisition of layered anorthositic intrusions (Figure \ref{fig:Petro_QDM}).
+% \dataset{dataset_one.txt}{Type or paste legend here.}
 
-\subsection*{Cooling rate correction}
-Paleointensity estimates from igneous intrusions can also be biased toward higher values due to slow cooling rates \cite{Halgedahl1980a}.
-Because the cooling rates of intrusive rocks could be much longer than the extrusive rocks, the slowly cooled diabase and anorthosite can bias laboratory paleointensity estimates toward an overestimate for SD grains \cite{Dodson1980a, Halgedahl1980a, Selkin2000a}.
+% \dataset{dataset_two.txt}{Type or paste legend here. Adding longer text to show what happens, to decide on alignment and/or indentations for multi-line or paragraph captions.}
 
-From the thermal history model we can also estimate the duration which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully block in their characteristic natural remanence component ($\sim$500\textdegree C; \citeA{Zhang2021b}).
-We find the cooling time to be on the order of $\sim$1 kyr, which corresponds to a cooling rate of $1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
-In contrast, the lab cooling rate is much faster, at about $1.3\times10^{-1}$ $^\circ$C s$^{-1}$ if we assume it takes 1000s in lab condition to cool through the same temperature interval.
-As a result, this significant cooling rate difference leads to a prediction of $\sim$36\% overestimate of true ancient field \cite{Halgedahl1980a}.
-We therefore correct our paleointensity results by a factor of 0.74.
-The cooling rate corrected specimen paleointensity estimates together with specimen- and site-level weighted means are shown in Figure \ref{fig:PINT_cooling_corrected}.
-Taking all accepted anorthosite paleointensity results, they have a site mean absolute paleointensity estimates of 47.90 $\pm$ 9.32 $\mu$T (Figure \ref{fig:PINT_cooling_corrected}).
-
-\subsection*{Late Mesoproterozoic geodynamo }
-To best characterize the geomagnetic axial dipole field intensity during a certain time period requires that a paleointensity dataset to cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged and the geocentric axial dipole hypothesis (GAD) can be assumed.
-Thermal modeling results from \citeA{Zhang2021b} suggest that the anorthosite xenoliths were likely heated to tholeiitic magma temperature and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a few thousand years.
-In addition, the closely located sites AX12 and AX16 have very similar paleointensity estimates that are distinct from those of sites AX11 and AX13 which were emplaced $\sim$10 km away (Figure \ref{fig:Geologic_map}).
-The consistency within small regions but variations between regions could be the result of differential cooling history due to heterogeneity of the timing and depths of the Beaver River diabase intrusions, causing the anorthosites to capture paleosecular variations \textit{ca.} 1092 Ma.
-Overall, none of the anorthosite xenoliths record a V(A)DM value smaller than 40 ZAm$^2$ (Figure \ref{fig:PINT_cooling_corrected}).
-
-Such a high estimate for the late Mesoproterozoic may invoke more careful examination of the geodynamo evolution through the Proterozoic and the timing for the inner core nucleation.
-Paleomagnetic observational records near the Ediacaran-Cambrian boundary have been interpreted to be consistent with a trend of decreasing axial dipole strength toward a very low value \cite{Lloyd2021a, Bono2019a}, followed by a rapid recovery (a few tens of million years) toward higher values soon after \cite{Thallner2021a}.
-Such a transition near this time can be consistent with the dynamo numerical simulations which adopt a rather high thermal conductivity of the core \cite{Davies2021a}.
-In particular, some models have suggested that the kick-start of rigorous compositional convection between the inner core and the outer core soon after the onset of inner core nucleation can result in almost an order of magnitude increase in Earth's axial dipole moment \cite{Davies2021a}.
-
-Although more recent data
-
-using paleomagnetic observational records is guided by the predictions from numerical simulations that a cooling liquid core with only thermal convection power to drive a dynamo would experience a progressive decay in its magnetic dipole field strength, and experience a rapid recover after the onset of inner core nucleation \cite{Davies2021a}.
-Previous studies has pointed to the Ediacaran-Cambrian boundary for such a weakening field behavior and recent studies has claim to found a rapid increase after the single crystal time which fits in the numerical models.
-However, such a recover may not be unique given the strong geodynamo estimate from the 1.1 Ga Midcontinent Rift.
-One could point to the increase from low values from the ca.
-1140 Ma abitibi dikes to the ca.
-1092 Ma anorthosites to be the result of recovery; another line of disagreement is between the Bangemall sills and the low field from Sept Ils.
-MCR being an anomaly?
-or the time scale we are looking at is too long?
-
-Regardless these several lines of increase/decrease trend needs further explanations/considerations.
-
-use a combined method of direction and intensity research for further examinations.
-Re-evaluation of the spatial and temporal variability captured by the existing Precambrian database.
-More data are needed for the Precambrian in order to better characterize the geodynamo evolution.
-
-\begin{figure}[h!]
-\noindent\includegraphics[width=\textwidth]{PINT_compilation.pdf}
-\centering
-\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment (VADM) values from the PINT database (v.2015.05; \url{http://earth.liv.ac.uk/ pint/}; \citeA{Biggin2015a}) with additional Mesoproterozoic data from \citeA{Sprain2018a, Kodama2019a}, as well as those from the Neoproterozoic \cite{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}.
-Paleointensity estimates from \citeA{Pesonen1983a} and \citeA{Kulakov2013a} are not included in the compilation due to the specimen-level double-slope behavior as discussed in the text.
-Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \citeA{Bono2019a} which is based on an interpretation of there being a progressive decay of the geodynamo through the Proterozoic.}}
-\label{fig:MCR_VADM}
-\end{figure}
-
-The calculated paleomagnetic VADM values are shown in Figure \ref{} in context of the compilation of all Midcontinent Rift results.
-highlight the high quality of the results and the consistency of MCR rocks' high geomagnetic field values even considering the uncertainties of previous data.
-Supporting that the Mesoproterozoic likely experienced a period of persistent high geomagnetic strength in addition to the protracted magmatism of in the formation of the massive intracontinental rifting.
-
-% cooling rate correction problem - that the Nagy et al paper can suggest even less of a correction and higher of the estimated field.
-theoretical thermal modeling by \citeA{Zhang2021a} suggest tha the anorthosite xenoliths were likely heated to diabase magma temperature regardless of their sizes, and cooled through the typical characteristic unblocking temperature range from 580 to 500 on the order of 1000 years.
-Comparing to the typical lab cooling time of about 100 seconds duirng paleointensity experiments, the significantly slower cooling rate when the NRM was acquired as the anorthosite xenoliths cooled with the diabase led to overestimate of paleointensity by up to 40\%.
-Overall range of correction factor can be 1.1-1.4, in any case consistent with a high field strength
-
-% that this dataset also could have suffered from problem of averaging out secular variations, given that the mean pmag directions deviate from what is expected from the sythesized APWP, nevertheless the variations in estimate absolute paleointensity lead us to interpret that albeit close, the they intensity data were acquired at distinctly different times that are sufficiently far apart such that the surface geomagnetic field had already changed.
-
-% that \citeA{bono2019a} proposed an interesting trend but has problems in the directions, likely to be a short term feature of a low field during a reversal, instead of a low field as a result of long term cooling.
-
-% that Lloyd 2021 also presented low field that fits the trend, but they similarly have the issue of reversal, and that they do not have good control on secular variations
-
-% Together with previously published data, the Midcontinent Rift rocks are recording 1.1 Ga earth geomagnetic field that is stronger than the mean of the past 300 Ma (Figure \ref{}).
-
-\section*{Conclusion}
-high success rate comparing to volcanics (more than half success sites), producing good consistent site level results, and across sites report present-day like field like today.
-
-AF pre-treatment is helpful in that it increase the specimens/sites passing the selecting criteria, suggesting that the method likely helped improve the MD ferromagnetic carriers.
-
-Rock mag and petrographic data might suggest differences in terms of the value of peak field on coercivity spectra between those passed and those didn't.
-
-IRM measured before and after an acid dunk in --- concentration acid on specimens after IZZI experiment show a difference (significant drop) maybe? between specs that passed and those did not.
-
-geographical association
-
-anisotropy of full TRM experiment on accepted specs show low degree of bulk sample scale full TRM anisotropy, with a range of ---.
-Cooling rate correction based on \cite{Halgedahl1980a} suggest a lab over estimate of B ancient of 35 \%, assuming constant cooling rate in natural condition.
-The cooling rate corrected field is $\sim$40 $\mu$T which is similar to today's field strength at similar latitude.
-
-\acknowledgments
-Project research was supported by NSF CAREER grant EAR-1847277 to N.L.S.-H.
-and an Institute on Lake Superior Geology Student Research Fund grant to Y.Z.
-Permits for fieldwork and sampling from the Minnesota Department of Natural Resources are gratefully acknowledged.
-We thank James Pierce and Blake Hodgin for their assistance in the field.
-We thank Jim Miller for providing us with the Duluth Complex anorthosite thin section.
-We thank Dario Bilardello, Peat Solheid, Mike Jackson, Josh Feinberg, and Bruce Moskowitz for their tremendous help of instrumental operations, data interpretations, and research guidance.
-Y.Z.
-thanks the IRM for the U.S.
-Visiting Student Fellowship.
-Paleomagnetic data associated with this study are available within the MagIC database (\url{}; UPDATE WHEN DOI IS GENERATED) and all data are within a Github repository associated with this work (\url{}) that is also archived on Zenodo (INSERT URL AT TIME OF PROOFS).
-This repository also contains Python code related to calculations, visualizations and statistical tests discussed herein.
 \bibliography{YZ_ref}
@@
 \end{document}
-
-% potential reviewer: Peter Selkin, Jeff Gee, Lisa Tauxe, Roger Fu, Lauri Pesonen, Josh Feinberg Julie Bowles,
-
-% At the IRM, I was seeking the answer to this question with the help of the vibrating sample magnetometer (VSM) systems, including a newly installed Lake Shore VSM which greatly helped improve measurement resolution on samples with weak magnetizations (Figure \ref{fig:backfield}).
-Backfield demagnetization experiments were conducted and used to develop coercivity spectra (Figure \ref{fig:backfield}).
-The spectra were subsequently modeled to fit for the distributions of different  populations of magnetic particles using similar procedure as \cite{Maxbauer2016a}.
-Most spectra can be well approximated with models with one component or two components with overlapping coercivity ranges (Figure \ref{fig:backfield}).
-The dominance of single-component distributions from the unmixed coercivity spectra likely suggests minimal alterations or formation of secondary magnetic mineral within the samples.
-By compiling the values of median destructive field (MDF) from all measured specimens and categorizing them in terms of their sister specimens' paleointensity results, I found that the anorthosites samples that pass paleointensity selection criteria have distinctively higher MDF values than those did not (Figure \ref{fig:backfield}), and all diabase specimens have low MDF values like the non-ideal anorthosites (Figure \ref{fig:backfield}).
-The higher MDFs ($\sim$60 mT) of the anorthosites are similar to what is typical of stoichiometric, single-domain magnetite grains which are favored for paleointensity experiments.
-Furthermore, the low MDFs displayed in other specimens could be linked to a dominant population of large interstitial Fe-Ti oxides that are more likely to form multi-domain grains, which have been observed in both the anorthosite and diabase (Figure \ref{fig:field}).
-In addition, preliminary tests on single anorthosite crystals might suggest that not all single crystals are necessarily better for paleointensity research when comparing to bulk samples, given their similar low MDF to the failed diabase and anorthosite specimens.
-However, more single crystal grains with paired paleointensity data and rock magnetic data may be needed before a general conclusion can be drawn.
-
-% Taken together, the rock magnetism experiment results support my paleointensity selection criteria in that rock specimens that produce straight Arai plots and have MDF similar to stoichiometric single-domain magnetite are preferentially selected.
-The preliminary paleointensity results yielded consistent estimates.
-The cooling rate corrected site-mean paleointensity estimate is about 40 $\mu$T, consistent with results from \cite{Sprain2018a} that the Earth's magnetic field strength at surface in the late Mesoproterozoic was close to that of today.
-
-% \begin{figure}
-% \begin{center}
-% 	\noindent\includegraphics[width=0.5\textwidth]{backfield_example_.pdf}
-% \end{center}
-% \caption{\small{(A) Backfield measurement data acquired using a Princeton VSM on an anorthosite specimen.
-The noise level on coercivity plot is high and a heavy smoothing is needed to make subsequent coercivity modeling interpretable.
-(B) Backfield measurement data acquired using a Lake Shore VSM on the same specimen as that in (A).
-The inset coercivity unmixing plot shows that the much smoother initial backfield curve resulted in a tight single-component model fit.
-(C) Box plot of distribution of the median destructive field of all measured specimens.
-The anorthosite specimens that pass selection criteria have distinctly higher median destructive field (MDF) than other groups.}}
-% \label{fig:backfield}
-% \end{figure}

### 2021-11-07 PNAS_manuscript/PNAS_manuscript.tex

@@
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
-The primary goal of the significance statement is to explain the relevance of the work in broad context to a broad readership.
-The significance statement appears in the paper itself and is required for all research papers.}
+\significancestatement{Acquiring high-fidelity observational records of ancient magnetic field intensity from remanent magnetization of rocks is crucial for constraining the evolution history of Earth's core.
+However, such records are often difficult to recover due to the alterations associated with weathering after the formation of rocks.
+In this study, we use plagioclase cumulates from the North American Midcontinent Rift that has experienced minimal post-formation alterations to obtain a record of $\sim$89 ZAm$^2$ virtual dipole moment 1.1 billion years ago.
+The results indicate the existence of strong power sources for Earth's dynamo during the late Mesoproterozoic.
+}
 
@@
 In contrast to the diabase which has non-ideal paleointensity behaviors leading to its failure in providing reliable paleointensity estimates, the anorthosite xenoliths often yield high-quality, single-slope paleointensity results that are consistent at specimen and site levels.
-The high paleointensity estimates from the anorthosites require an existence of a vigorous late Mesoproterozoic geodynamo.
-Together with previous paleointensity data from the Midcontinent Rift, our new results suggest that a strong geodynamo likely persisted through $\sim$25 Myr at ca.
-1.1 Ga, inconsistent with the interpretation of a progressive decay of Earth's geodynamo strength through the Proterozoic Eon.
+An average value of $\sim$89 ZAm$^2$ for the virtual dipole moment estimates from the anorthosites is higher than that of the dipole component of Earth's magnetic field today.
+Such a strong dipole strength recorded by the anorthosites require an existence of a vigorous late Mesoproterozoic geodynamo.
+Together with previous paleointensity data from other Midcontinent Rift rocks, our new results suggest that a dynamo with strong power sources likely persisted through $\sim$25 Myr at \textit{ca.} 1.1 Ga.
+These observations are inconsistent with there being a progressive decay of Earth's dynamo strength through the Proterozoic Eon.
 \end{abstract}
@@
 \dropcap{E}arth's magnetic field is the result of convective flow of liquid iron-alloy in Earth's outer core.
-At present day, the geodynamo is collectively powered by heat flow across the core-mantle boundary (CMB) and from the crystallization of the solid inner core from the liquid outer core which provides latent heat and compositional buoyancy.
+At present day, the geodynamo is collectively powered by heat flow across the core-mantle boundary (CMB), and from the crystallization of the solid inner core from the liquid outer core which provides latent heat and compositional buoyancy.
 However, while paleomagnetic studies have found that a dynamo field has existed since at least 3.4 billion years ago in the Archean \cite{Selkin2007a, Tarduno2014a, Brenner2020a}, Earth's inner core likely crystallized more recently.
 Estimates of the timing of the initial crystallization of Earth's inner core are interconnected with estimates for the core's thermal conductivity.
-Higher conductivity values imply faster cooling rates which in turn imply that the threshold for the freezing of the inner core happened more recently \cite{Davies2015a}.
-While some estimates of these values are consistent with an inner core age $>$ 3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates based on first-principles calculations implied higher conductivities \cite{Koker2012a,Pozzo2012a} and an age for the inner core that is less than 1.3 Ga \cite{Pozzo2012a, Gomi2013a, Labrosse2015a, Ohta2016a, Zhang2020b}.
+Higher conductivity values imply faster cooling rates, which in turn imply that the threshold for the freezing of the inner core happened more recently \cite{Davies2015a}.
+While some estimates of these values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates based on first-principles calculations implied higher thermal conductivity values and an age for the inner core that is less than 1.3 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Labrosse2015a, Ohta2016a, Zhang2020b}.
 With the estimates for the core thermal conductivity continue to be debated, it is crucial to use observational records as an independent constraint on the thermal evolution of the core and mantle.
@@
 Paleomagnetic data from ancient rocks are one of the few types of observational data that have the potential to record the evolution of Earth’s core and can be used to test the predictions from numerical simulations.
-Some Precambrian absolute paleointensity records from the PINT database (\url{http://earth.liv.ac.uk/pint/}) show evidence of an existence of an appreciable geomagnetic field prior to the Ediacaran (Figure \ref{fig:PINT_compilation}).
+Some Precambrian absolute paleointensity records from the PINT database (\url{http://earth.liv.ac.uk/pint/}) show evidence that an appreciable geomagnetic field existed prior to the Ediacaran (Fig.
+\ref{fig:PINT_compilation}).
 On the other hand, studies have found very low virtual dipole moment (VDM) estimates from rocks formed in the late Ediacaran \cite{Bono2019a, Shcherbakova2019a, Thallner2021b}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
@@
 It has been interpreted that these findings are consistent with the results of geodynamo models using high core thermal conductivity values that suggest the inner core nucleation age being near the Ediacaran-Cambrian boundary \cite{Driscoll2016a, Davies2021a}.
-\citeA{Bono2019a} adopted a billion-year-scale view of the Precambrian paleointensity records and based on a polynomial fit through selected data to suggest that the geodynamo power progressively decreased through the Proterozoic (Figure \ref{fig:PINT_compilation}).
+\cite{Bono2019a} adopted a billion-year-scale view of the Precambrian paleointensity records and based on a polynomial fit through selected data to suggest that the geodynamo power progressively decreased through the Proterozoic (Fig.
+\ref{fig:PINT_compilation}).
 That study also recovered very low paleointensity values from \textit{ca.} 565 Ma single silicate crystals and interpreted them to be reflecting the geodynamo to have reached its lowest near the Ediacaran-Cambrian boundary.
@@
 
-However, previous paleointensity data from intrusive and extrusive rock units in the Mesoproterozoic Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record Earth's virtual axial dipole moment (VADM) values similar to that of today, which are higher than the prediction of the long term decay curve (Figure \ref{fig:PINT_compilation}).
-\citeA{Pesonen1983b} obtained an average virtual dipole moment of 93 ZAm$^2$ after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using both extrusive and intrusive rocks, including the \textit{ca.} 1108-1105 Ma Osler Volcanics, the Maimainse Point volcanics, intrusive Logan Sills, Thunder Bay dikes, the Baraga and Marquette dikes, as well as lava flow of the \textit{ca.} 1096 Ma North Shore Volcanic Group.
-\citeA{Kulakov2013a} obtained an average VADM of 59 $\pm$ 11 ZAm$^2$ using the \textit{ca.} 1085 Ma Lake Shore Traps volcanics and suggest that the present-day-like geomagnetic field strength likely existed through the late stage of Midcontinent Rift magmatism \cite{Vervoort2007a, Miller2013a}.
-However, non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results obtained by \citeA{Pesonen1983b, Kulakov2013a}.
+However, previous paleointensity data from intrusive and extrusive rocks in the Mesoproterozoic Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record Earth's virtual dipole moment (VDM) values similar to that of today, which are higher than the prediction of the long term decay curve (Fig.
+\ref{fig:PINT_compilation}).
+\cite{Pesonen1983b} obtained an average virtual dipole moment of 93 $\times 10^21$ Am$^2$ (ZAm$^2$) after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using both extrusive and intrusive rocks, including the \textit{ca.} 1108-1105 Ma Osler Volcanics, the \textit{ca.} 1100 Ma Maimainse Point volcanics, \textit{ca.} 1107 Ma intrusive Logan Sills, Thunder Bay dikes, the Baraga and Marquette dikes, as well as lava flow of the \textit{ca.} 1096 Ma North Shore Volcanic Group (Fig.
+\ref{fig:Geologic_map}).
+\cite{Kulakov2013a} obtained an average VADM of 59 $\pm$ 11 ZAm$^2$ using the \textit{ca.} 1085 Ma Lake Shore Traps volcanics and suggest that the present-day-like geomagnetic field strength likely existed through the late stage of Midcontinent Rift magmatism \cite{Vervoort2007a, Miller2013a}.
+However, non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results obtained by \cite{Pesonen1983b, Kulakov2013a}.
 One typical failure of paleointensity experiment is the double-slope behavior, as two distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
-\citeA{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas \citeA{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
-Results with such non-ideal behavior were rejected by \citeA{Sprain2018a} who applied stricter paleointensity selection criteria on new paleointensity results and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
+\cite{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas \cite{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
+Results with such non-ideal behavior were rejected by \cite{Sprain2018a} who applied stricter paleointensity selection criteria on new paleointensity results and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
 A calculated mean VADM value of 56 $\pm$ 21 ZAm$^2$ is similar to the average value of the past 300 million years \cite{Sprain2018a}.
@@
 The Mesoproterozoic North American Midcontinent Rift (MCR) is an intracontinental rift that did not result in the break-up of Laurentia.
-Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; \citeA{Vervoort2007a, Miller2013a}) was punctuated by rapid and voluminous emplacement of intrusive rocks as well as extrusive volcanics.
-Based on high-precision $^{206}$Pb/$^{238}$U geochronology and paleomagnetic data, \citeA{Swanson-Hysell2021a} hypothesized that the massive Duluth Complex and the coeval North Shore Volcanic group (Figure \ref{fig:Geologic_map}) formed as a result of having upwelling mantled-derived plume material encountering Laurentia's thinned lithosphere, following an ``upside-down” drainage" which rapidly funneled melt to today's North Eastern Minnesota (Figure \ref{fig:Geologic_map}; \citeA{Sleep1997a, Swanson-Hysell2014a}).
-The plagioclase-phyric Anorthositic Series rocks, the crystal-poor Layered Series intrusions of the Duluth Complex, and the coeval volcanics of the North Shore Volcanic Group (NSVG) that outcrop along Northeastern Minnesota (Figure \ref{fig:Geologic_map}) are interpreted to be a result of such vigorous magmatic activity.
+Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; \cite{Vervoort2007a, Miller2013a}) was punctuated by rapid and voluminous emplacement of intrusive rocks as well as extrusive volcanics.
+Based on high-precision $^{206}$Pb/$^{238}$U geochronology and paleomagnetic data, \cite{Swanson-Hysell2021a} hypothesized that the massive Duluth Complex and the coeval North Shore Volcanic group (Fig.
+\ref{fig:Geologic_map}) formed as a result of upwelling mantled-derived plume material encountering Laurentia's thinned lithosphere, following an ``upside-down” drainage" which rapidly funneled melt to today's northeastern Minnesota (Fig.
+\ref{fig:Geologic_map}) \cite{Sleep1997a, Swanson-Hysell2014a}).
+The plagioclase-phyric Anorthositic Series rocks, the crystal-poor Layered Series intrusions of the Duluth Complex, and the coeval volcanics of the North Shore Volcanic Group (NSVG) that outcrop along Northeastern Minnesota (Fig.
+\ref{fig:Geologic_map}) are interpreted to be a result of such vigorous magmatic activity.
 
@@
 Red diamonds mark sites with paleointensity results that passed the selection criteria from this study.
-Paleomagnetic sites from \citeA{Pesonen1983a} are categorized by lithology.
-All sites from previous studies are color-coded by their ages.}}
+Paleomagnetic sites from \cite{Pesonen1983a} are categorized by lithology.
+All sites from \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} are color-coded by their ages.}}
 \label{fig:Geologic_map}
@@
 The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex punctuates another period of magmatism during the Midcontinent Rift main stage activity.
-$^{206}$Pb/$^{238}$U zircon dates and cross-cutting relationships tightly constraints the areally extensive Beaver River diabase of the Beaver Bay Complex to have been emplaced at 1091.7 $\pm$ 0.2 Ma \cite{Zhang2021b}.
+$^{206}$Pb/$^{238}$U zircon dates and cross-cutting relationships tightly constrain the areally extensive Beaver River diabase of the Beaver Bay Complex to have been emplaced at 1091.7 $\pm$ 0.2 Ma \cite{Zhang2021b}.
 The Beaver River diabase intruded into hypabyssal depths via wide conduits that accommodated the entrainment of numerous anorthosite xenoliths that have short-axis diameters up to 180 meters \cite{Boerboom2004a, Boerboom2006b}.
@@
 In addition, petrographic imaging support that the plagioclase from the Beaver River anorthosite formed under a magmatic conditions different from the Duluth Complex anorthositic series magma.
-While the plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture (Figure \ref{Petro_QDM}A), the more calcium-rich plagioclase of the Beaver River anorthosite often develop granoblastic texture characterized by equigranular plagioclase crystals (Figure \ref{Petro_QDM}B).
+While the plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture (Fig.
+\ref{Petro_QDM}A), the more calcium-rich plagioclase of the Beaver River anorthosite often develop granoblastic texture characterized by equigranular plagioclase crystals (Fig.
+\ref{Petro_QDM}B).
 Abundant magnetite and ilmenite needles that are typically 10 $\mu$m long can be found to have preferred orientations within Duluth Complex plagioclase.
@@
 
-\begin{SCfigure*}[\sidecaptionrelwidth][t]
+\begin{figure*}
 \centering
 \noindent\includegraphics[width=11.4 cm]{Petro_QDM.png}
-\caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from Duluth Complex Anorthositic Series (left column) and Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (right column; \citeA{Zhang2021b}).
-Petrographic images (A-D) show Duluth Complex anorthosite have euhedral, elongate plagioclase crystals with interlocking texture and igneous foliation.
-(C) The plagioclase can contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the (010) planes of plagioclase.
+\caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from Duluth Complex Anorthositic Series (left column) and Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (right column; \cite{Zhang2021b}).
+Petrographic image (A) shows Duluth Complex anorthosite have euhedral, interlocking plagioclase crystals showing igneous foliation.
+(B) Beaver River anorthosite have plagioclase with a granoblastic texture.
+(C) Duluth Complex plagioclase can contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the (010) planes of plagioclase.
 In contrast, plagioclase of sample MS99033 do not enclose large oxide needles.
-Magnetic field maps acquired by the quantum diamond microscope (QDM) show that the oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the overlapping magnetic dipole fields exerted above the oxides in (E).
+Magnetic field maps acquired by the quantum diamond microscope (QDM) at UC Berkeley Paleomagnetism Lab show that the oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the complex overlapping magnetic fields exerted above the oxides in (E).
 In comparison, relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite.
@@
 \label{fig:Petro_QDM}
-\end{SCfigure*}
+\end{figure*}
 
-Occurrences of magnetite and ilmenite symplectites and needles within anorthositic rocks have been described by \citeA{Scofield1975a, Scofield1986a} in the Wichita Mountains and by \citeA{Haggerty1976a} in other slowly cooled, mafic intrusions.
-\citeA{Scofield1986a} hypothesized that the inclusions acted as a means of buffering the melt with respect to oxygen, forming either crystallites along preexisting plagioclase substrate or as iron incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures.
+Occurrences of magnetite and ilmenite symplectites and needles within anorthositic rocks have been described by \cite{Scofield1975a, Scofield1986a} in the Wichita Mountains and by \cite{Haggerty1976a} in other slowly cooled, mafic intrusions.
+\cite{Scofield1986a} hypothesized that the inclusions acted as a means of buffering the melt with respect to oxygen, forming either crystallites along preexisting plagioclase substrate or as iron incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures.
 Such a model requires that plagioclase, instead of other mafic minerals such as olivine, to be the first to crystallize from a magma and act as a sink for iron.
@@
 This proposed magmatic evolution of the Duluth Complex is in distinct contrast to the emplacement of the Beaver River diabase which rapidly ascended and entrained the anorthosite xenoliths directly into shallow chambers \cite{Flower1980a, Miller1997a, Zhang2021b}.
-Particularly, \citeA{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
+Particularly, \cite{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
 This is consistent with the interpretation that these anorthosite were rapidly transported by upward currents into hypabyssal magma chambers without reaching structural equilibrium.
-Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Figure \ref{Petro_QDM}).
-Instead, relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Figure \ref{Petro_QDM}).
+Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Fig.
+\ref{Petro_QDM}).
+Instead, relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
+\ref{Petro_QDM}).
 
@@
 40 from a total of 86 anorthosite specimens and 0 out of a total of 69 diabase specimens passed our paleointensity selection criteria, yielding 7 anorthosite sites and no diabase sites with site-level paleointensity results that pass paleointensity selections.
-Example NRM/TRM plots (Arai plots;
-\cite{Arai1963a}) are shown in Figure \ref{fig:IZZI_examples}.
-Summary specimen absolute paleointensity estimates and site-level weighted mean paleointensity values are shown in Table \ref{tab:PINT_result} and are plotted in Figure \ref{fig:PINT_cooling_corrected}.
+Example NRM/TRM plots
+\cite{Arai1963a} are shown in Fig.
+\ref{fig:IZZI_examples}.
+Summary specimen absolute paleointensity estimates and site-level weighted mean paleointensity values are shown in Table \ref{tab:PINT_result} and are plotted in Fig.
+\ref{fig:PINT_cooling_corrected}.
 All measurement-level paleointensity experiment data can be accessed through the MagIC database (\url{}; UPDATE WHEN DOI IS GENERATED).
 
-\begin{figure*}[h!]
+\begin{figure*}
 \noindent\includegraphics[width=17.8cm]{IZZI_examples.pdf}
@@
 The plots for anorthosite specimens AX1-1a and diabase BD5-1b on the bottom row show typical non-ideal sagging and zigzagging behaviour that fails our acceptance criteria.
-Plot for diabase specimen BD2-5a shows typical paleointensity behavior of this site - specimens typically have double-slope behaviors although the lower temperature slopes can pass the selection criteria, resulting in the high estimated B$_{anc}$ values shown in Figure \ref{fig:PINT_cooling_corrected}.
-Thus, the estimates of B$_{anc}$ for this specimen plot is illustrative, but were not used to calculate a site mean paleointensity value in Figure \ref{fig:PINT_cooling_corrected}.}}
+Plot for diabase specimen BD2-5a shows typical paleointensity behavior of this site - specimens typically have double-slope behaviors although the lower temperature slopes can pass the selection criteria, resulting in the high estimated B$_{anc}$ values shown in Fig.
+\ref{fig:PINT_cooling_corrected}.
+Thus, the estimates of B$_{anc}$ for this specimen plot is illustrative, but were not used to calculate a site mean paleointensity value in Fig.
+\ref{fig:PINT_cooling_corrected}.}}
 \label{fig:IZZI_examples}
@@
 The Beaver River anorthosite specimens yielded a high pass rate ($\sim$46.5\%) through our selection criteria.
-In distinct contrast, the Beaver River diabase specimens are all rejected by our selection criteria.
-Paleointensity experiment results from both lithologies show three types of dominant specimen behaviors on Arai plots (Figure \ref{fig:IZZI_examples}; \citeA{Arai1963a}): dominantly single-slope, double-slope, and zigzagging or sagging (Figure \ref{fig:IZZI_examples}).
-Note that almost all specimens have minimal secondary remanence component in thermal demagnetization results as shown in example orthogonal plots of all zero-field steps (Figure \ref{fig:IZZI_examples}).
-Typical specimen paleointensity behavior of the anorthosites display straight, single-slope Arai plots and the accepted fractions of temperature steps span over the origin-trending, primary remanence components (Figure \ref{fig:IZZI_examples}).
+In distinct contrast, the Beaver River diabase specimens are all rejected.
+Paleointensity experiment results from both lithologies show three types of dominant specimen behaviors on Arai plots (Fig.
+\ref{fig:IZZI_examples}): dominantly single-slope, double-slope, and zigzagging or sagging.
+Note that almost all specimens have minimal secondary remanence component in thermal demagnetization results as shown in example orthogonal plots of all zero-field steps (Fig.
+\ref{fig:IZZI_examples}).
+This reflects that the rocks have experienced mild alteration after their formation.
+Typical specimen paleointensity behavior of the anorthosites display straight, single-slope Arai plots and the accepted fractions of temperature steps span over the origin-trending, primary remanence components (Fig.
+\ref{fig:IZZI_examples}).
 We accept all specimen- and site-level absolute paleointensity estimates from the anorthosite xenoliths that pass the selection.
-In terms of the diabase specimens, all failed the selection because of double-slope behavior (fail FRAC selection), poor pTRM checks, and zigzagging (fail SCAT, DANG selection) Arai plots (Figure \ref{fig:IZZI_examples}).
-As is shown in Figure \ref{fig:IZZI_examples}, specimen BD2-5b has a high-temperature second slope with zigzagging behavior, poor pTRM checks, and low NRM/TRM ratio, in addition to the straight slope with steep NRM/TRM ratio over the lower temperature range.
-Although the BD2 specimens' lower temperature fraction of the Arai plots are straight andthe pTRM checks are consistent, lower temperature slopes from BD2 give consistently higher paleointensity estimates than those from site AX6, AX8, and AX10, which are anorthosite xenoliths included within BD2 (Figure \ref{fig:Geologic_map}; \citeA{Zhang2021a}).
-This higher estimate from the diabase lithology could be attributed to the systematic bias of using the lower temperature slope of the Arai plots demonstrated in Figure \ref{fig:IZZI_examples} and specimens from site BD2 all fail the criteria of having a maximum magnetizatiom gap between adjacent steps of 25\% of the total NRM.
-Thus we only accept the single-slope anorthosite specimens results for further analyses.
+All diabase specimens failed the selection because of double-slope behavior (fail FRAC selection; see methods), poor pTRM checks, and zigzagging (fail SCAT, DANG selection; see methods) Arai plots (Fig.
+\ref{fig:IZZI_examples}).
+As is shown in Fig.
+\ref{fig:IZZI_examples}, specimen BD2-5b has a high-temperature slope with zigzagging behavior, poor pTRM checks, and low NRM/TRM ratio, in addition to a low-temperature slope with a steep NRM/TRM ratio.
+Although the BD2 specimens' lower temperature fraction of the Arai plots are straight and the pTRM checks are consistent, lower temperature slopes from BD2 give consistently higher paleointensity estimates than those from site AX6, AX8, and AX10, which are anorthosite xenoliths included within BD2 (Fig.
+\ref{fig:Geologic_map}) \cite{Zhang2021a}.
+This higher estimate from the diabase lithology could be attributed to the systematic bias of using the lower temperature slope of the Arai plots demonstrated in Fig.
+\ref{fig:IZZI_examples} and specimens from site BD2 all fail the criteria of having a maximum magnetization gap between adjacent steps of 25\% of the total NRM.
+Eventually, we only accept the single-slope anorthosite specimens results for further analyses.
 
@@
 \caption{\small{Summary plot of cooling rate corrected individual specimen absolute paleointensity results (square and diamond symbols) from this study and their averages at site (grey bar) and locality level (weighted mean by site-level standard deviations; orange bar).
-The left plot shows the graph of predicted paleointensity overestimate due to slow cooling of the intrusive Beaver River diabase and anorthosite xenoliths relative to the cooling rate in laboratory following the model of \citeA{Halgedahl1980a}.
-Because the majority of the anorthosites have unblocking temperatures between 500\textdegree C and 580\textdegree C, we estimate that the slow cooling during natural remanence acquisition could have resulted in a 36\% overestimate.
-Thus a correction factor of 0.74 is applied at specimen level for the paleointensity summary plot.
 }}
@@
 On the other hand, 27 out of 58 ($\sim$46.6\%) anorthosite specimens that had AF treatment passed selection and 13 out of 28 ($\sim$46.4\%) specimens without AF treatment passed selection.
-Paleointensity estimates from specimens with or without AF treatment in anorthosite AX10, AX11, AX12, AX13, AX16 do not show significant difference (Figure \ref{fig:PINT_cooling_corrected}).
+Paleointensity estimates from specimens with or without AF treatment in anorthosite AX10, AX11, AX12, AX13, AX16 do not show significant difference (Fig.
+\ref{fig:PINT_cooling_corrected}).
 Specimen plots for AX11 particularly illustrates a case where all specimens within the site yield almost the same paleointensity estimate, regardless of the AF treatment.
@@
 Another cause for their failure could be the thermochemical alterations that happened during heating in air.
-As can be exampled by the zigzagging high temperature slope with poor pTRM checks in paleointensity plot for specimen BD2-5b (Figure \ref{fig:IZZI_examples}), the decrease in NRM/TRM ratios during higher temperature steps likely resulted from the formation of new ferromagnetic material within the specimen due to oxidation.
+As can be exampled by the zigzagging high temperature slope with poor pTRM checks in paleointensity plot for specimen BD2-5b (Fig.
+\ref{fig:IZZI_examples}), the decrease in NRM/TRM ratios during higher temperature steps likely resulted from the formation of new ferromagnetic material within the specimen due to oxidation.
 Such alterations can allow specimens to acquire additional remanence during in-field steps that may not be fully removed by an AF treatment.
@@
 
-Many authors find it useful to organize their manuscripts with the following order of sections;  title, author line and affiliations, keywords, abstract, significance statement, introduction, results, discussion, materials and methods, acknowledgments, and references.
-Other orders and headings are permitted.
-
-\begin{table*}
-\caption{\footnotesize{Summary paleointensity results for specimens that passed our selection.
-Paleointensity results for specimens that passed quality criteria.
-B$_{anc}$ is the calculated ancient field intensity over the chosen temperature interval in $\mu$T.
-T$_{min}$ and T$_{max}$ indicate the temperature interval over which the best fit for paleointensity was defined.
-N is the number of steps used within the selected interval for paleointensity determination.
-FRAC is the fraction of remanence.
-NpTRM shows the number of pTRM checks within the selected interval for paleointensity determination.
-$\beta$ is the scatter parameter.
-GAP-MAX is the maximum magnetization gap between two adjacent steps.
-MAD is the maximum angle of deviation.
-DANG is the deviation angle.
-SCAT is the scatter parameter.
-Paleolatitude is calculated from the inclination values reported in \citeA{Zhang2021b}.
-$\gamma$ is the gamma statistic that measures the angle between the last pTRM step used for paleointensity determination and the applied field direction.
-V(A)DM is the virtual (axial) dipole moment reported in $10^{21}$Am$^2$ (ZAm$^2$).
-}}
-\centering
-\begin{tabular}[width=17.8cm]{cccccccccccccccc}
-\hline
-Site & Specimen & B$_{anc}$ & T$_{min}$ & T$_{max}$ & N    & FRAC & NpTRM & $\beta$ & GAP-MAX & MAD ($^\circ$) & DANG ($^\circ$) & SCAT & Paleolatitude & $\gamma$ & VADM (ZAm$^2$) \\
-\hline
-AX6  & AX6-2a   & 18.25     & 400       & 585       & 18 & 0.7  & 10.0  & 0.04    & 0.12    & 3.44           & 3.43            & PASS & 15.61         & 2.7      & 31.69          \\
-AX6  & AX6-3a   & 25.39     & 400       & 585       & 18 & 0.76 & 10.0  & 0.04    & 0.1     & 4.28           & 2.88            & PASS & 15.61         & 3.2      & 44.08          \\
-AX6  & AX6-1a   & 49.13     & 425       & 585       & 17 & 0.71 & 10.0  & 0.03    & 0.14    & 4.45           & 0.96            & PASS & 15.61         & 2.0      & 85.3           \\
-AX8  & AX8-3a   & 41.3      & 400       & 580       & 17 & 0.75 & 9.0   & 0.03    & 0.14    & 4.38           & 2.22            & PASS & 17.16         & 11.2     & 70.45          \\
-AX8  & AX8-1a   & 66.18     & 450       & 570       & 13 & 0.68 & 8.0   & 0.06    & 0.18    & 4.54           & 2.37            & PASS & 17.16         & 7.0      & 112.88         \\
-AX10 & AX10-1a  & 45.84     & 425       & 585       & 17 & 0.78 & 10.0  & 0.06    & 0.24    & 5.6            & 2.62            & PASS & 14.41         & 4.7      & 80.64          \\
-AX10 & AX10-3a  & 48.01     & 425       & 585       & 17 & 0.69 & 10.0  & 0.04    & 0.24    & 4.02           & 1.56            & PASS & 14.41         & 5.8      & 84.45          \\
-AX10 & AX10-2a  & 48.23     & 425       & 585       & 17 & 0.67 & 10.0  & 0.05    & 0.19    & 5.3            & 1.93            & PASS & 14.41         & 7.8      & 84.84          \\
-AX11 & AX11-1a  & 73.29     & 425       & 560       & 10 & 0.67 & 6.0   & 0.08    & 0.21    & 5.28           & 2.54            & PASS & 12.27         & 5.9      & 131.75         \\
-AX11 & AX11-2a  & 73.29     & 400       & 560       & 11 & 0.65 & 6.0   & 0.07    & 0.21    & 3.94           & 1.51            & PASS & 12.27         & 9.9      & 131.75         \\
-AX11 & AX11-4a  & 73.31     & 500       & 570       & 11 & 0.66 & 8.0   & 0.09    & 0.2     & 1.53           & 3.27            & PASS & 12.27         & 5.6      & 131.78         \\
-AX11 & AX11-6a  & 73.32     & 200       & 562       & 14 & 0.69 & 7.0   & 0.07    & 0.17    & 4.0            & 1.33            & PASS & 12.27         & 3.0      & 131.8          \\
-AX11 & AX11-9a  & 73.34     & 100       & 562       & 15 & 0.67 & 7.0   & 0.06    & 0.17    & 5.34           & 1.13            & PASS & 12.27         & 11.9     & 131.84         \\
-AX11 & AX11-3a  & 73.41     & 400       & 560       & 11 & 0.66 & 6.0   & 0.08    & 0.21    & 3.95           & 2.11            & PASS & 12.27         & 8.2      & 131.96         \\
-AX11 & AX11-5a  & 73.44     & 400       & 566       & 14 & 0.73 & 8.0   & 0.05    & 0.18    & 2.93           & 0.85            & PASS & 12.27         & 1.9      & 132.02         \\
-AX12 & AX12-14a & 30.34     & 475       & 585       & 16 & 0.75 & 10.0  & 0.06    & 0.17    & 1.74           & 0.55            & PASS & 24.54         & 0.9      & 47.18          \\
-AX12 & AX12-1a  & 33.89     & 0         & 580       & 21 & 0.97 & 9.0   & 0.03    & 0.17    & 5.47           & 4.73            & PASS & 24.54         & 10.0     & 52.7           \\
-AX12 & AX12-6a  & 37.18     & 425       & 564       & 12 & 0.69 & 7.0   & 0.08    & 0.22    & 3.64           & 2.05            & PASS & 24.54         & 5.0      & 57.81          \\
-AX12 & AX12-8a  & 38.34     & 475       & 565       & 12 & 0.75 & 8.0   & 0.06    & 0.25    & 1.46           & 2.29            & PASS & 24.54         & 1.6      & 59.62          \\
-AX12 & AX12-4a  & 41.89     & 500       & 585       & 14 & 0.66 & 10.0  & 0.05    & 0.24    & 3.85           & 3.19            & PASS & 24.54         & 11.4     & 65.14          \\
-AX12 & AX12-2a  & 46.94     & 425       & 570       & 14 & 0.7  & 8.0   & 0.05    & 0.24    & 3.39           & 2.38            & PASS & 24.54         & 7.0      & 72.99          \\
-AX13 & AX13-3a  & 71.68     & 100       & 550       & 12 & 0.71 & 5.0   & 0.08    & 0.22    & 5.92           & 1.88            & PASS & 11.25         & 3.4      & 130.08         \\
-AX13 & AX13-4a  & 75.42     & 200       & 555       & 12 & 0.7  & 6.0   & 0.03    & 0.17    & 7.28           & 0.41            & PASS & 11.25         & 3.7      & 136.87         \\
-AX13 & AX13-6a  & 76.8      & 200       & 555       & 12 & 0.7  & 6.0   & 0.06    & 0.22    & 3.82           & 1.53            & PASS & 11.25         & 4.5      & 139.37         \\
-AX13 & AX13-2a  & 76.96     & 300       & 555       & 11 & 0.69 & 6.0   & 0.05    & 0.2     & 7.87           & 2.73            & PASS & 11.25         & 4.1      & 139.66         \\
-AX13 & AX13-9a  & 77.48     & 0         & 564       & 17 & 0.94 & 7.0   & 0.03    & 0.16    & 5.96           & 1.07            & PASS & 11.25         & 2.6      & 140.61         \\
-AX13 & AX13-8a  & 84.83     & 450       & 570       & 13 & 0.81 & 8.0   & 0.02    & 0.24    & 3.69           & 1.82            & PASS & 11.25         & 5.1      & 153.94   \\
-AX16 & AX16-13a & 41.68     & 400       & 570       & 15 & 0.72 & 8.0   & 0.07    & 0.2     & 3.47           & 1.85            & PASS & 22.11         & 1.1      & 66.88          \\
-AX16 & AX16-16a & 42.2      & 400       & 570       & 15 & 0.77 & 8.0   & 0.06    & 0.15    & 3.9            & 1.81            & PASS & 22.11         & 2.0      & 67.72          \\
-AX16 & AX16-14a & 43.57     & 400       & 570       & 15 & 0.71 & 8.0   & 0.06    & 0.13    & 3.25           & 1.79            & PASS & 22.11         & 6.0      & 69.91          \\
-AX16 & AX16-11a & 44.17     & 450       & 570       & 14 & 0.7  & 8.0   & 0.06    & 0.15    & 2.86           & 2.07            & PASS & 22.11         & 4.6      & 70.88          \\
-AX16 & AX16-5a  & 45.75     & 400       & 560       & 11 & 0.69 & 6.0   & 0.09    & 0.22    & 7.98           & 4.32            & PASS & 22.11         & 4.0      & 73.41          \\
-AX16 & AX16-4a  & 45.96     & 425       & 570       & 14 & 0.75 & 8.0   & 0.05    & 0.21    & 5.01           & 1.18            & PASS & 22.11         & 4.2      & 73.75          \\
-AX16 & AX16-2a  & 45.98     & 400       & 564       & 13 & 0.68 & 7.0   & 0.06    & 0.21    & 4.37           & 1.57            & PASS & 22.11         & 4.9      & 73.78          \\
-AX16 & AX16-1a  & 46.25     & 200       & 562       & 14 & 0.85 & 7.0   & 0.07    & 0.23    & 6.04           & 2.87            & PASS & 22.11         & 4.5      & 74.22          \\
-AX16 & AX16-9a  & 48.9      & 475       & 585       & 16 & 0.66 & 10.0  & 0.04    & 0.16    & 2.92           & 1.4             & PASS & 22.11         & 4.3      & 78.47          \\
-AX16 & AX16-10a & 54.02     & 500       & 585       & 15 & 0.66 & 10.0  & 0.04    & 0.18    & 2.78           & 1.15            & PASS & 22.11         & 5.5      & 86.68          \\
-\hline
-\end{tabular}
-\label{tab:PINT_result}
-\end{table*}
-
-\begin{table*}[]
-
-\caption{\footnotesize{Summary statistics for the Q$_{PI}$ quality criteria of \citeA{Biggin2014a}.}}
-\centering
-\begin{tabular}{ccccccccccccc}
-\hline
-Site & N  & Age (Ma) & Method & AGE & STAT & TRM & ALT & MD & ACN & TECH & LITH & QPI \\ \hline
-AX6  & 3  & 1091.8   & T+     & 1   & 0    & 1   & 1   & 1  & 1   & 0    & 0    & 5   \\
-AX8  & 2  & 1091.8   & T+     & 1   & 0    & 1   & 1   & 1  & 1   & 0    & 0    & 5   \\
-AX10 & 3  & 1091.8   & T+     & 1   & 0    & 1   & 1   & 1  & 1   & 0    & 0    & 5   \\
-AX11 & 7  & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\
-AX12 & 6  & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\
-AX13 & 7  & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\
-AX16 & 11 & 1091.8   & T+     & 1   & 1    & 1   & 1   & 1  & 1   & 0    & 0    & 6   \\ \hline
-\end{tabular}
-\label{tab:QPI}
-\end{table*}
-
 \subsection*{Rock magnetism}
 Rock magnetic data support that the successful anorthosite specimens have dominant magnetic carriers with magnetic properties similar to stochiometric, non-interactive, single domain magnetite, whereas those anorthosites that failed the paleointensity result selection and all diabase have more pronounced populations of non-ideal carriers.
-MPMS data show sister specimens to paleointensity specimens typically have dominant magnetic mineralogy of magnetite as can be evidenced through the Verwey transitions (Figure \ref{fig:MPMS}; \citeA{Verwey1939a}).
+MPMS data show sister specimens to paleointensity specimens typically have dominant magnetic mineralogy of magnetite as can be evidenced through the Verwey transitions (Fig.
+\ref{fig:MPMS}; \cite{Verwey1939a}).
 However, specimens from sites that yield successful paleointensity results display Verwey transition near 120 Kelvin (K) as expected for stochiometric low-Ti magnetite \cite{Ozdemir1993a}.
-On the other hand, the diabase and anorthosite specimens that did not pass our paleointensity selection typically have suppressed Verwey transition toward lower temperatures (Figure \ref{fig:MPMS}), indicating that the specimens likely have a ferromagnetic mineralogy with variable degrees of oxidation \cite{Ozdemir1993a}.
-
-% \begin{figure*}
-% \noindent\includegraphics[width=11.4 cm]{MPMS.pdf}
-% \centering
-% \caption{\footnotesize{Low-temperature magnetic property measurement system (MPMS) experiment results.
-In the field-cooled (FC) experiments, the magnetization was measured upon warming following the specimen having cooled in an applied field of 2.5 T from 300 to 10 K.
-In the zero-field-cooled (ZFC) experiment, a low-temperature saturation isothermal remanence (LTSIRM) of 2.5 T was applied at 10 K after the specimen cooled in a (near-)zero field.
-In the room-temperature saturation isothermal remanence (RTSIRM) experiment, the sample was pulsed with a 2.5 T field at room temperature ($\sim$300 K) and then cooled to 10 K and warmed back to room temperature in a (near-) zero field.
-Specimen AX6-1r is from anorthosite AX6 which passed our paleointensity selection.
-It has a well-defined Verwey transition $\sim$120 K \cite{Verwey1939a}.
-Specimens AX3-2r-ni and BD2-2r show Verwey transition but the transition temperatures are suppressed below $\sim$120 K.
-Specimen BD2-2r has a consistently higher moment during the zero-field-cooled step than during the field-cooled step.
-This is consistent with the interpretation that multidomian magnetic carriers exist in significant quantity in this specimen.}}
-% \label{fig:MPMS}
-% \end{figure*}
+On the other hand, the diabase and anorthosite specimens that did not pass our paleointensity selection typically have suppressed Verwey transition toward lower temperatures (Fig.
+\ref{fig:MPMS}), indicating that the specimens likely have a ferromagnetic mineralogy with variable degrees of oxidation \cite{Ozdemir1993a}.
 
 Specimen coercivity spectra further support that our paleointensity selection criteria are successful in selecting experiment results.
-A compilation of median destructive field (MDF) in Figure \ref{fig:coercivity} based on all of our backfield experiments shows that specimens from anorthosites that pass paleointensity selections have distinctly a higher average MDF range than other anorthosites and all diabase.
-Example single component skew-normal distribution fits of coercivity spectra based on backfield experiments on specimen chips sister to paleointensity specimens show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity distributions around 80 mT (Figure \ref{fig:coercivity}).
+A compilation of median destructive field (MDF) in Fig.
+\ref{fig:coercivity} based on all of our backfield experiments shows that specimens from anorthosites that pass paleointensity selections have distinctly a higher average MDF range than other anorthosites and all diabase.
+Example single component skew-normal distribution fits of coercivity spectra based on backfield experiments on specimen chips sister to paleointensity specimens show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity distributions around 80 mT (Fig.
+\ref{fig:coercivity}).
 In contrast, specimens from anorthosite that did not pass paleointensity selection, as well as diabase specimens, tend to have coercivity distributions over a significantly lower range ($\sim$30 mT).
@@
 \cite{Selkin2007a} also noted that vortex state magnetic grains and certain single domain particles can produce biased paleointensity estimates due to their acquisition of saturation remanence during laboratory in-field steps.
-According to the numerical modeling of \citeA{Selkin2007a}, thermoremanent magnetization acquired by a randomly-oriented population of elongate single-domian-sized magnetite particles are unlikely to saturate upon our applied field of 30 $\mu$T.
+According to the numerical modeling of \cite{Selkin2007a}, thermoremanent magnetization acquired by a randomly-oriented population of elongate single-domian-sized magnetite particles are unlikely to saturate upon our applied field of 30 $\mu$T.
 Therefore, in this study we do not apply any anisotropy or nonlinear TRM correction to our specimens.
@@
 With an applied field of 30 $\mu$T oriented toward the specimen at 0, 30, 60, 90, 120, and 150 degrees, the anorthosite specimens usually acquired consistent full TRMs.
-The most significant variations observed in acquired full TRMs reach to only XXX \%, much less than that has been reported to occurred within layerd anorthositic rocks in Stillwater (up to $\sim$60 \%, \citeA{Selkin2000a}).
+The most significant variations observed in acquired full TRMs reach to only XXX \%, much less than that has been reported to occurred within layerd anorthositic rocks in Stillwater (up to $\sim$60 \%, \cite{Selkin2000a}).
 
-The reason that the Beaver River anorthosite xenoliths do not display strong remanence anisotropy, nor do they display nonlinear TRM acquisition problem, could be that they formed in a magmatic setting distinct from that of a layered mafic complex, such as the nearby Duluth Complex, where igneous foliation can occur (Figure \ref{fig:Geologic_map}).
-Petrographic images show that an equigranular texture where equidimensional, randomly oriented plagioclase often develop in the anorthosite xenoliths (Figure \ref{fig:Petro_QDM}; \citeA{Zhang2021b}).
-This is in contrast to the interlocking igneous foliation texture within the Anorthositic Series rocks of the \textit{ca.} 1096 Ma Duluth Complex (Figure \ref{fig:Petro_QDM}).
-Further, petrographic and magnetic imaging also shows the Beaver River anorthosite xenoliths do not develop abundant exsolved magnetite and ilmenite needles within plagioclase crystals as the slowly-cooled anorthositic intrusions do, in which exsolved magnetite needles can be the dominate remanence carriers (Figure \ref{fig:Petro_QDM}; \citeA{Scofield1986a, Feinberg2006a}).
-The fact that all dipole features within a same Duluth Complex plagioclase crystal have indistinguishable orientations is consistent with the interpretation of there being a magnetic remanence anisotropy resulting from preferred petrofabric alignment, which could give rise to the remanence anisotropy and nonlinear TRM acquisition of layered anorthositic intrusions (Figure \ref{fig:Petro_QDM}).
+The reason that the Beaver River anorthosite xenoliths do not display strong remanence anisotropy, nor do they display nonlinear TRM acquisition problem, could be that they formed in a magmatic setting distinct from that of a layered mafic complex, such as the nearby Duluth Complex, where igneous foliation can occur (Fig.
+\ref{fig:Geologic_map}).
+Petrographic images show that an equigranular texture where equidimensional, randomly oriented plagioclase often develop in the anorthosite xenoliths (Fig.
+\ref{fig:Petro_QDM}; \cite{Zhang2021b}).
+This is in contrast to the interlocking igneous foliation texture within the Anorthositic Series rocks of the \textit{ca.} 1096 Ma Duluth Complex (Fig.
+\ref{fig:Petro_QDM}).
+Further, petrographic and magnetic imaging also shows the Beaver River anorthosite xenoliths do not develop abundant exsolved magnetite and ilmenite needles within plagioclase crystals as the slowly-cooled anorthositic intrusions do, in which exsolved magnetite needles can be the dominate remanence carriers (Fig.
+\ref{fig:Petro_QDM}; \cite{Scofield1986a, Feinberg2006a}).
+The fact that all dipole features within a same Duluth Complex plagioclase crystal have indistinguishable orientations is consistent with the interpretation of there being a magnetic remanence anisotropy resulting from preferred petrofabric alignment, which could give rise to the remanence anisotropy and nonlinear TRM acquisition of layered anorthositic intrusions (Fig.
+\ref{fig:Petro_QDM}).
 
@@
 
-From the thermal history model we can also estimate the duration which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully block in their characteristic natural remanence component ($\sim$500\textdegree C; \citeA{Zhang2021b}).
+From the thermal history model we can also estimate the duration which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully block in their characteristic natural remanence component ($\sim$500\textdegree C; \cite{Zhang2021b}).
 We find the cooling time to be on the order of $\sim$1 kyr, which corresponds to a cooling rate of $1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
@@
 We therefore correct our paleointensity results by a factor of 0.74.
-The cooling rate corrected specimen paleointensity estimates together with specimen- and site-level weighted means are shown in Figure \ref{fig:PINT_cooling_corrected}.
-Taking all accepted anorthosite paleointensity results, they have a site mean absolute paleointensity estimates of 47.90 $\pm$ 9.32 $\mu$T (Figure \ref{fig:PINT_cooling_corrected}).
+The cooling rate corrected specimen paleointensity estimates together with specimen- and site-level weighted means are shown in Fig.
+\ref{fig:PINT_cooling_corrected}.
+Taking all accepted anorthosite paleointensity results, they have a site mean absolute paleointensity estimates of 47.90 $\pm$ 9.32 $\mu$T (Fig.
+\ref{fig:PINT_cooling_corrected}).
 
@@
 To best characterize the geomagnetic axial dipole field intensity during a certain time period requires that a paleointensity dataset to cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged and the geocentric axial dipole hypothesis (GAD) can be assumed.
-Thermal modeling results from \citeA{Zhang2021b} suggest that the anorthosite xenoliths were likely heated to tholeiitic magma temperature and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a few thousand years.
-In addition, the closely located sites AX12 and AX16 have very similar paleointensity estimates that are distinct from those of sites AX11 and AX13 which were emplaced $\sim$10 km away (Figure \ref{fig:Geologic_map}).
+Thermal modeling results from \cite{Zhang2021b} suggest that the anorthosite xenoliths were likely heated to tholeiitic magma temperature and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a few thousand years.
+In addition, the closely located sites AX12 and AX16 have very similar paleointensity estimates that are distinct from those of sites AX11 and AX13 which were emplaced $\sim$10 km away (Fig.
+\ref{fig:Geologic_map}).
 The consistency within small regions but variations between regions could be the result of differential cooling history due to heterogeneity of the timing and depths of the Beaver River diabase intrusions, causing the anorthosites to capture paleosecular variations \textit{ca.} 1092 Ma.
-Overall, none of the anorthosite xenoliths record a V(A)DM value smaller than 40 ZAm$^2$ (Figure \ref{fig:PINT_compilation}).
+Overall, none of the anorthosite xenoliths record a V(A)DM value smaller than 40 ZAm$^2$ (Fig.
+\ref{fig:PINT_compilation}).
 
@@
 
-However, the updated paleointensity records from the Midcontinent Rift together with previous results from other Mesoproterozoic rocks suggest that the a dramatic increase in geomagnetic dipole moment similar in scale to that modeled for after the onset of inner core nucleation may not be unique to the Edicaran-Cambrian boundary before 500 Ma (Figure \ref{fig:PINT_compilation}).
+However, the updated paleointensity records from the Midcontinent Rift together with previous results from other Mesoproterozoic rocks suggest that the a dramatic increase in geomagnetic dipole moment similar in scale to that modeled for after the onset of inner core nucleation may not be unique to the Edicaran-Cambrian boundary before 500 Ma (Fig.
+\ref{fig:PINT_compilation}).
 The \textit{ca.} 1140 Ma Abitibi dikes \cite{Macouin2003a} record an average virtual dipole moment of $\sim$13 ZAm$^2$, similar to the low values of $\sim$7 ZAm$^2$ obtained from the single silicate crystals of the \textit{ca.} 565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a}.
-However, the relatively high virtual dipole moments of $\sim$72 ZAm$^2$ from the \textit{ca.} 1105 Ma Osler Volcanic Group \cite{Sprain2018a} and $\sim$82 ZAm$^2$ from the \textit{ca.} 1092 Ma Beaver River anorthosite xenoliths (Figure \ref{fig:PINT_compilation} are about six times the values from the \textit{ca.} 1140 Ma Abitibi dikes.
-In addition, the increase in absolute dipole moment estimates recorded by the Midcontinent Rift rocks are much higher than that in the Ediacaran (Figure \ref{fig:PINT_compilation}).
+However, the relatively high virtual dipole moments of $\sim$72 ZAm$^2$ from the \textit{ca.} 1105 Ma Osler Volcanic Group \cite{Sprain2018a} and $\sim$82 ZAm$^2$ from the \textit{ca.} 1092 Ma Beaver River anorthosite xenoliths (Fig.
+\ref{fig:PINT_compilation} are about six times the values from the \textit{ca.} 1140 Ma Abitibi dikes.
+In addition, the increase in absolute dipole moment estimates recorded by the Midcontinent Rift rocks are much higher than that in the Ediacaran (Fig.
+\ref{fig:PINT_compilation}).
 
@@
 The $\sim$60 ZAm$^2$ dipole moment estimate from the \textit{ca.} 755 Ma Mundine Well dikes are discordant with the long-term decrease trend \cite{Bono2019a} and is about three times the estimates from the \textit{ca.} 720 Ma dolerite dikes of the Franklin Large Igneous Province \cite{Lloyd2021a}.
-Similarly, an apparent rapid decrease in geomagnetic dipole moment from \textit{ca.} 1092 Ma to \textit{ca.} 1070 Ma can be evidenced by the contrasting high values from the Midcontinent Rift anorthosite xenoliths and the low values from the Mundine Well dikes (Figure \ref{fig:PINT_compilation}).
+Similarly, an apparent rapid decrease in geomagnetic dipole moment from \textit{ca.} 1092 Ma to \textit{ca.} 1070 Ma can be evidenced by the contrasting high values from the Midcontinent Rift anorthosite xenoliths and the low values from the Mundine Well dikes (Fig.
+\ref{fig:PINT_compilation}).
 However, a sparse record in a short time period may not sufficiently average out paleosecular variations.
@@
 Overall, the existing absolute paleointensity records in the PINT database remains sparse in terms of spatial and temporal coverage as compared to more geologically recent times.
-(Figure \ref{fig:PINT_compilation}).
+(Fig.
+\ref{fig:PINT_compilation}).
 More data are needed in order to better characterize the evolution of Earth's dynamo and thermal evolution.
@@
 \centering
-\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment (VADM) values from the PINT database (v.2015.05; \url{http://earth.liv.ac.uk/ pint/}; \citeA{Biggin2015a}) with additional Mesoproterozoic data from \citeA{Sprain2018a, Kodama2019a}, as well as those from the Neoproterozoic \cite{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}.
-Paleointensity estimates from \citeA{Pesonen1983a} and \citeA{Kulakov2013a} are not included in the compilation due to the specimen-level double-slope behavior as discussed in the text.
-Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \citeA{Bono2019a} which is based on an interpretation of there being a progressive decay of the geodynamo through the Proterozoic.}}
+\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment (VADM) values from the PINT database (v.2015.05; \url{http://earth.liv.ac.uk/ pint/}; \cite{Biggin2015a}) with additional Mesoproterozoic data from \cite{Sprain2018a, Kodama2019a}, as well as those from the Neoproterozoic \cite{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}.
+Paleointensity estimates from \cite{Pesonen1983a} and \cite{Kulakov2013a} are not included in the compilation due to the specimen-level double-slope behavior as discussed in the text.
+Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \cite{Bono2019a} which is based on an interpretation of there being a progressive decay of the geodynamo through the Proterozoic.}}
 \label{fig:PINT_compilation}
@@
 Specimens from a total of 7 diabase and 14 anorthosite sites were used for paleointensity experiments (Table \ref{tab:Pmag_site_data}).
-Sister specimens underwent step-wise alternating field (AF) or thermal demagnetization at the UC Berkeley Paleomagnetism Lab and the results are presented in \citeA{Zhang2021b}.
+Sister specimens underwent step-wise alternating field (AF) or thermal demagnetization at the UC Berkeley Paleomagnetism Lab and the results are presented in \cite{Zhang2021b}.
 The close directional similarities between the anorthosite xenoliths and their host diabase and thermal modeling results are consistent with the interpretation that the Beaver River diabase and anorthosite xenoliths acquired primary thermal remanent magnetization \textit{ca.} 1092 Ma during cooling after the emplacement of the diabase \cite{Zhang2021b}.
@@
 A total of 86 specimens from 14 anorthosite xenoliths and a total of 69 specimens from 7 diabase sites underwent paleointensity experiments that followed the step-wise double-heating Thellier method \cite{Thellier1959a} using the IZZI protocol \cite{Tauxe2004a} with heating steps up to 585 \textdegree C.
-Partial thermal remanent magnetization (pTRM) checks were performed systematically throughout the experiment to test whether there was significant mineralogical alteration due to heating and were assessed using the SCAT parameter of \citeA{Shaar2013a}.
+Partial thermal remanent magnetization (pTRM) checks were performed systematically throughout the experiment to test whether there was significant mineralogical alteration due to heating and were assessed using the SCAT parameter of \cite{Shaar2013a}.
 On top of the IZZI modification of the Thellier experimental protocol, we also performed a comparative study where we added an extra step of 20 mT alternating field (AF) cleaning on some of the specimens after each in-field step.
@@
 \subsection*{Paleointensity result selection}
-The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD; \citeA{Kirschvink1980a}) of $<$20\textdegree; (2) scatter parameter ($\beta$; \citeA{Coe1978a}) values of $<$15 per cent; (3) a deviation angle (DANG; \citeA{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC; \citeA{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT; \citeA{Shaar2013a}) = TRUE; (6) a maximum magnetic moment difference between adjacent zero-field steps (GAP-Max; \citeA{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
+The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD; \cite{Kirschvink1980a}) of $<$20\textdegree; (2) scatter parameter ($\beta$; \cite{Coe1978a}) values of $<$15 per cent; (3) a deviation angle (DANG; \cite{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC; \cite{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT; \cite{Shaar2013a}) = TRUE; (6) a maximum magnetic moment difference between adjacent zero-field steps (GAP-Max; \cite{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
 The criteria are also summarized in Table \ref{tab:criteria}.
@@
 Both MAD and DANG assess the directional variation of the NRM, with MAD measuring the scatter in the NRM directions and DANG assessing whether the component is trending toward the origin of the Zijderveld plot.
-$\beta$ is the ``scatter" parameter of \citeA{Coe1978a} and is the ratio of the standard error of the slope of the best-fit line of the selected NRM and pTRM points on an Arai plot to the absolute value of the slope.
+$\beta$ is the ``scatter" parameter of \cite{Coe1978a} and is the ratio of the standard error of the slope of the best-fit line of the selected NRM and pTRM points on an Arai plot to the absolute value of the slope.
 FRAC is the fraction of the NRM that is used in the best-fit line \cite{Shaar2013a}.
@@
 $\beta$, FRAC, GAP-Max and SCAT are all statistics to assess the behavior of Arai plots.
-See the Standard Paleointensity Definitions (\citeA{Paterson2014a}; \url{https://earthref.org/PmagPy/SPD/home.html}) for more details.
+See the Standard Paleointensity Definitions (\cite{Paterson2014a}; \url{https://earthref.org/PmagPy/SPD/home.html}) for more details.
 Data analysis was conducted using Thellier GUI \cite{Shaar2013a} within the PmagPy software package \cite{Tauxe2016a}.
@@
 Specimen median destructive field (MDF) are calculated based on the backfield curves.
-The calculated coercivity spectra were subsequently decomposed into one ore more components using skew-normal distributions following the method of \citeA{Maxbauer2016a}.
+The calculated coercivity spectra were subsequently decomposed into one ore more components using skew-normal distributions following the method of \cite{Maxbauer2016a}.
 The fitted components are then used to interpret the characteristics of different populations of magnetic particles within specimens.
@@
 
-To further identify the magnetic carriers within the Beaver River anorthosite xenoliths and compare them with the anorthosites of the Duluth Complex Anorthositic Series rocks, we used the quantum diamond microscope (QDM) at the UC Berkeley Paleomagnetism laboratory to image a thin section of sample MS99033 which is a hand sample from anorthosite xenolith AX16 (which yielded a $^{206}$Pb/$^{238}$U zircon date of 1091.83 $\pm$ 0.21 Ma; \citeA{Zhang2021b}), and a thin section of a Duluth Complex anorthosite (Figure \ref{fig:Petro_QDM}).
+To further identify the magnetic carriers within the Beaver River anorthosite xenoliths and compare them with the anorthosites of the Duluth Complex Anorthositic Series rocks, we used the quantum diamond microscope (QDM) at the UC Berkeley Paleomagnetism laboratory to image a thin section of sample MS99033 which is a hand sample from anorthosite xenolith AX16 (which yielded a $^{206}$Pb/$^{238}$U zircon date of 1091.83 $\pm$ 0.21 Ma; \cite{Zhang2021b}), and a thin section of a Duluth Complex anorthosite (Fig.
+\ref{fig:Petro_QDM}).
 We the QDM to image the magnetic field over the polished thin section surfaces at a sample-sensor distance of 5 $\mu$m in projective magnetic microscopy (PMM) mode with a spatial resolution of 4.7 μm per pixel and an instantaneous 0.9 mT bias field that is canceled during the course of measurement \cite{Glenn2017a}).
@@
 
-% Figure \ref{fig:frog} shows an example of how to insert a column-wide figure.
-To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment.
+% Fig.
+\ref{fig:frog} shows an example of how to insert a column-wide figure.
+To insert a Fig.
+wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment.
 Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
-Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side legends.
+Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide Fig.
+with side legends.
 

### 2021-11-08 PNAS_manuscript/PNAS_manuscript.tex
update manuscript

@@
 In contrast to the diabase which has non-ideal paleointensity behaviors leading to its failure in providing reliable paleointensity estimates, the anorthosite xenoliths often yield high-quality, single-slope paleointensity results that are consistent at specimen and site levels.
-An average value of $\sim$89 ZAm$^2$ for the virtual dipole moment estimates from the anorthosites is higher than that of the dipole component of Earth's magnetic field today.
-Such a strong dipole strength recorded by the anorthosites require an existence of a vigorous late Mesoproterozoic geodynamo.
+An average value of $\sim$89 ZAm$^2$ for the virtual dipole moment from the anorthosites is higher than that of the dipole component of Earth's magnetic field today.
+Such a strong late Mesoproterozoic dipole recorded by the anorthosites require the existence of a vigorous geodynamo at the time.
 Together with previous paleointensity data from other Midcontinent Rift rocks, our new results suggest that a dynamo with strong power sources likely persisted through $\sim$25 Myr at \textit{ca.} 1.1 Ga.
@@
 Higher conductivity values imply faster cooling rates, which in turn imply that the threshold for the freezing of the inner core happened more recently \cite{Davies2015a}.
-While some estimates of these values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates based on first-principles calculations implied higher thermal conductivity values and an age for the inner core that is less than 1.3 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Labrosse2015a, Ohta2016a, Zhang2020b}.
-With the estimates for the core thermal conductivity continue to be debated, it is crucial to use observational records as an independent constraint on the thermal evolution of the core and mantle.
+While some estimates of these values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates based on first principles calculations implied higher thermal conductivity values and an age for the inner core that is less than 1.3 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b}, with some suggesting even younger ages ($<$ 700 Ma) \cite{Labrosse2015a, Ohta2016a}.
+With the estimates for the core's thermal conductivity continue to be debated, it is crucial to use observational records as an independent constraint on the thermal evolution of the core and mantle.
 
@@
 Paleomagnetic data from ancient rocks are one of the few types of observational data that have the potential to record the evolution of Earth’s core and can be used to test the predictions from numerical simulations.
-Some Precambrian absolute paleointensity records from the PINT database (\url{http://earth.liv.ac.uk/pint/}) show evidence that an appreciable geomagnetic field existed prior to the Ediacaran (Fig.
+Previous interpretations of the current Precambrian absolute paleointensity records (compiled in the PINT database; \url{http://earth.liv.ac.uk/pint/}) suggest that an appreciable geomagnetic field existed in the Archean and progressively decayed through the Proterozoic till it reached its weakest in the Ediacaran when very low virtual dipole moment (VDM) estimates have been obtained \cite{Bono2019a, Shcherbakova2019a, Thallner2021b, Lloyd2021a}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
+Bono et al \cite{Bono2019a} adopted a billion-year-scale view of such observations of the Precambrian geomagnetic field intensities to suggest that these evidence are consistent with models that use high core thermal conductivity values where the nucleation of the inner core began to provide compositional convection power for driving the geodynamo during the Ediacaran-Cambrian after experiencing a progressive decay when the geodynamo was dominantly driven by thermal conduction earlier in the Proterozoic \cite{Driscoll2016a, Davies2021a}.
+
+% Other studies focusing on paleomagnetic directions have also found that the late Ediacaran might have been a period of time when the geomagnetic field were experiencing frequent reversals and waning of dipole dominance \cite{Bono2015a, Kodama2020a}.
+
+However, previous paleointensity data from intrusive and extrusive rocks in the \textit{ca.} 1.1 Ga Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record VDM values similar to that of today, which are higher than the values predicted by the long term decay curve proposed by Bono et al \cite{Bono2019a} (Fig.
 \ref{fig:PINT_compilation}).
-On the other hand, studies have found very low virtual dipole moment (VDM) estimates from rocks formed in the late Ediacaran \cite{Bono2019a, Shcherbakova2019a, Thallner2021b}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
-Other studies focusing on paleomagnetic directions have also found that the late Ediacaran might have been a period of time when the geomagnetic field were experiencing frequent reversals and waning of dipole dominance \cite{Bono2015a, Kodama2020a}.
-It has been interpreted that these findings are consistent with the results of geodynamo models using high core thermal conductivity values that suggest the inner core nucleation age being near the Ediacaran-Cambrian boundary \cite{Driscoll2016a, Davies2021a}.
-\cite{Bono2019a} adopted a billion-year-scale view of the Precambrian paleointensity records and based on a polynomial fit through selected data to suggest that the geodynamo power progressively decreased through the Proterozoic (Fig.
-\ref{fig:PINT_compilation}).
-That study also recovered very low paleointensity values from \textit{ca.} 565 Ma single silicate crystals and interpreted them to be reflecting the geodynamo to have reached its lowest near the Ediacaran-Cambrian boundary.
-New low paleointensity estimates derived from the \textit{ca.} 720 Ma Franklin Large Igneous Province are interpreted to support the existence of such decay trend in the Neoproterozoic \cite{Lloyd2021a}.
-
-However, previous paleointensity data from intrusive and extrusive rocks in the Mesoproterozoic Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record Earth's virtual dipole moment (VDM) values similar to that of today, which are higher than the prediction of the long term decay curve (Fig.
-\ref{fig:PINT_compilation}).
-\cite{Pesonen1983b} obtained an average virtual dipole moment of 93 $\times 10^21$ Am$^2$ (ZAm$^2$) after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using both extrusive and intrusive rocks, including the \textit{ca.} 1108-1105 Ma Osler Volcanics, the \textit{ca.} 1100 Ma Maimainse Point volcanics, \textit{ca.} 1107 Ma intrusive Logan Sills, Thunder Bay dikes, the Baraga and Marquette dikes, as well as lava flow of the \textit{ca.} 1096 Ma North Shore Volcanic Group (Fig.
+Pesonen and Halls \cite{Pesonen1983b} obtained an average virtual dipole moment of 93 $\times 10^21$ Am$^2$ (ZAm$^2$) after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using both extrusive and intrusive rocks from the Midcontinent Rift, including the \textit{ca.} 1108-1105 Ma Osler Volcanics, the \textit{ca.} 1100 Ma Maimainse Point volcanics, \textit{ca.} 1107 Ma intrusive Logan Sills, Thunder Bay dikes, the Baraga and Marquette dikes, as well as lava flows of the \textit{ca.} 1096 Ma North Shore Volcanic Group (Fig.
 \ref{fig:Geologic_map}).
-\cite{Kulakov2013a} obtained an average VADM of 59 $\pm$ 11 ZAm$^2$ using the \textit{ca.} 1085 Ma Lake Shore Traps volcanics and suggest that the present-day-like geomagnetic field strength likely existed through the late stage of Midcontinent Rift magmatism \cite{Vervoort2007a, Miller2013a}.
-However, non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results obtained by \cite{Pesonen1983b, Kulakov2013a}.
+Kulakov et al \cite{Kulakov2013a} obtained an average VDM of 59 $\pm$ 11 ZAm$^2$ using the \textit{ca.} 1085 Ma Lake Shore Traps volcanics and suggest that the present-day-like geomagnetic field strength likely existed through the late stage of Midcontinent Rift magmatism \cite{Vervoort2007a, Miller2013a}.
+However, non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results obtained by ref.
+\citealp{Pesonen1983b, Kulakov2013a}.
 One typical failure of paleointensity experiment is the double-slope behavior, as two distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
-\cite{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas \cite{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
-Results with such non-ideal behavior were rejected by \cite{Sprain2018a} who applied stricter paleointensity selection criteria on new paleointensity results and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
-A calculated mean VADM value of 56 $\pm$ 21 ZAm$^2$ is similar to the average value of the past 300 million years \cite{Sprain2018a}.
-This high VADM value estimate from the Midcontinent Rift volcanics may be incompatible with the hypothesis of a ``Proterozoic dipole low" \cite{Biggin2009a}.
+Pesonen and Halls \cite{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas Kulakov et al \cite{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
+Results with such non-ideal behavior were rejected by Sprain et al \cite{Sprain2018a} who applied stricter paleointensity selection criteria on new paleointensity results and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
+A calculated mean VDM value of 56 $\pm$ 21 ZAm$^2$ is similar to the average value of the past 300 million years \cite{Sprain2018a}.
+This high value from the Midcontinent Rift volcanics may be incompatible with the hypothesis of a ``Proterozoic dipole low" \cite{Biggin2009a}.
 
@@
 In this study, we present new paleointensity results from ``IZZI" modified absolute paleointensity experiments \cite{Yu2004a} on a suite of bulk samples from \textit{ca.} 1092 Ma anorthosite xenoliths and their Beaver River diabase host rocks of the Beaver Bay Complex within the Midcontinent Rift.
-Paleointensity experiments on the Beaver River anorthosite xenoliths have a very high success rate.
-The consistent specimen- and site-level paleointensity results reveal that the magnetite-dominated remanence of the anorthosite specimens has low thermal remanent magnetization (TRM) anisotropy and linear TRM acquisition.
-Magnetic imaging shows that the anorthosite specimens to have dominant magnetic carriers interstitial to and within plagioclase crystals and they do not display strong preferred orientations.
-With more single-slope, high quality paleointensity data, we seek to compare paleointensity records from the Midcontinent Rift intrusives with those from the extrusives; to test the hypothesis that the Earth's geomagnetic dipole strength remained at high values throughout the Midcontinent Rift; and to further enrich the reliable paleointensity records during the $\sim$25 Myr of Midcontinent Rift magmatic activity.
+Paleointensity experiments on the anorthosite xenoliths have a very high success rate.
+Their consistent specimen- and site-level paleointensity results together with rock magnetic data reveal that the anorthosite xenoliths have low anisotropy of thermal remanent magnetization (TRM) and can acquire TRM linearly.
+Magnetic imaging shows the anorthosite specimens have dominant magnetic carriers interstitial to and within plagioclase crystals and they do not display strong preferred orientations.
+With more single-slope, high quality paleointensity data, further enrich the reliable paleointensity records during the $\sim$25 Myr of Midcontinent Rift magmatic activity.
+Our new data from the Midcontinent Rift intrusive rocks support previous data from the volcanics that there was a strong late Mesoproterozoic geodynamo.
 
@@
 
-The Mesoproterozoic North American Midcontinent Rift (MCR) is an intracontinental rift that did not result in the break-up of Laurentia.
-Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; \cite{Vervoort2007a, Miller2013a}) was punctuated by rapid and voluminous emplacement of intrusive rocks as well as extrusive volcanics.
-Based on high-precision $^{206}$Pb/$^{238}$U geochronology and paleomagnetic data, \cite{Swanson-Hysell2021a} hypothesized that the massive Duluth Complex and the coeval North Shore Volcanic group (Fig.
-\ref{fig:Geologic_map}) formed as a result of upwelling mantled-derived plume material encountering Laurentia's thinned lithosphere, following an ``upside-down” drainage" which rapidly funneled melt to today's northeastern Minnesota (Fig.
+The Mesoproterozoic North American Midcontinent Rift is an intracontinental rift that did not result in the break-up of Laurentia.
+Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; \cite{Vervoort2007a, Miller2013a}) was characterized by rapid and voluminous emplacement of intrusive rocks and extrusive volcanics.
+The $<$500 Kyr emplacement of the \textit{ca.} 1096 Ma Duluth Complex and the $~$8 km thick volcanics associated with it is one manifestation \cite{Swanson-Hysell2021a}.
+The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex punctuates another such period of magmatism during the Midcontinent Rift main stage activity.
+High precision $^{206}$Pb/$^{238}$U zircon dates from one anorthosite xenolith and units that have cross-cutting relationships with the Beaver River diabase tightly constrain the timing of diabase emplacement to be at 1091.7 $\pm$ 0.2 Ma \cite{Zhang2021b}.
+The xenolith-bearing diabase dikes and sills intruded into hypabyssal depths via wide conduits that accommodated the entrainment of numerous anorthosite xenoliths that have short-axis diameters ranging from centimeters to up to 180 meters \cite{Boerboom2004a, Boerboom2006b}.
+
+Previous geochronology data and rare earth element diffusion modeling suggest that the anorthosite xenoliths of the Beaver River diabase are cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
+Petrological textures and geochemical attributes imply that the Beaver River anorthosite xenoliths formed in a lower crustal magmatic condition distinct from that of the Duluth Complex which is $\sim$4 Myr older and also generated massive anorthositic cumulates that led to the formation of the Anorthositic Series \cite{Zhang2021b, Swanson-Hysell2021a}.
+The dominantly monomineralic anorthosite xenoliths of the Beaver River diabase have an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}, higher than the average of $\sim$60\% of the Anorthositic Series rocks of the Duluth Complex \cite{Miller1990a}.
+The plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture showing igneous foliations (Fig.
+\ref{Petro_QDM}A), but the more calcium-rich plagioclase of the Beaver River anorthosites often develop granoblastic texture characterized by equigranular crystals with poor alignments (Fig.
+\ref{Petro_QDM}B).
+Abundant magnetite and ilmenite needles that are typically 10 $\mu$m long can be found to have preferred orientations within Duluth Complex plagioclase.
+Interstitial symplectic titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex Anorthositic Series rocks \cite{Miller1990a}.
+Such a high concentration of Fe oxides carrying magnetic remanence in the slowly cooled Duluth Complex anorthositic rocks reflect that iron in the plagioclase mush in the magma chamber acted as a means of buffering the melt with respect to oxygen, forming crystallites along preexisting plagioclase substrate or as iron oxides incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures \cite{Scofield1986a}.
+In contrast, titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent but relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
+\ref{Petro_QDM}).
+This likely reflects that the Beaver River diabase magma had a different magma oxidation history than that of the Duluth Complex.
+The plagioclase and oxide mineral textures of the Beaver River anorthosites also distinguish them from plagioclase cumulates of other layered mafic complex where magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
+
+% the Beaver River diabase rapidly ascended and entrained the anorthosite xenoliths directly into shallow chambers and likely did not reach  \cite{Flower1980a, Miller1997a, Zhang2021b}.
+Particularly, \cite{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
+This is consistent with the interpretation that these anorthosite were rapidly transported by upward currents into hypabyssal magma chambers without reaching structural equilibrium.
+Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Fig.
+\ref{Petro_QDM}).
+Instead, relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
+\ref{Petro_QDM}).
+
+% In the Beaver River anorthosite xenoliths, relatively sparse titano-magnetite occur within or interstitial to the plagioclase (Fig. \ref{Petro_QDM}).
+
+% Occurrences of magnetite and ilmenite symplectites and needles within anorthositic rocks have been described by \cite{Scofield1975a, Scofield1986a} in the Wichita Mountains and by \cite{Haggerty1976a} in other slowly cooled, mafic intrusions.
+\cite{Scofield1986a} hypothesized that the inclusions acted as a means of buffering the melt with respect to oxygen, forming either crystallites along preexisting plagioclase substrate or as iron incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures.
+Such a model requires that plagioclase, instead of other mafic minerals such as olivine, to be the first to crystallize from a magma and act as a sink for iron.
+The two-stage formation of the Duluth Complex Anorthositic Series fits this model \cite{Miller1990a}.
+During the first stage, high pressure at subcrustal depths (15-25 km depth) caused fractionation of Duluth Complex magma resulting in the formation of floating plagioclase mushes \cite{Kushiro1980a}.
+The second stage of crystallization then occurs at low pressures after the magma intruded into shallower chambers.
+
+% This proposed magmatic evolution of the Duluth Complex is in distinct contrast to the emplacement of the Beaver River diabase which rapidly ascended and entrained the anorthosite xenoliths directly into shallow chambers \cite{Flower1980a, Miller1997a, Zhang2021b}.
+Particularly, \cite{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
+This is consistent with the interpretation that these anorthosite were rapidly transported by upward currents into hypabyssal magma chambers without reaching structural equilibrium.
+Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Fig.
+\ref{Petro_QDM}).
+Instead, relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
+\ref{Petro_QDM}).
+
+% it is the most probable that the anorthosite xenoliths of the Beaver River diabase are cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
+Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths acquired thermal remanent magnetizations during cooling of the diabase \cite{Zhang2021b}.
+Step-wise thermal demagnetizations show many anorthosite xenoliths have minimal secondary remanence and single-component characteristic remanence magnetizations that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C.
+
+% Geochronological, paleomagnetic, geochemical, and petrologic data are consistent with the hypothesis that the Beaver River diabase magma extruded to the surface, feeding the massive Greenstone Flow - one of the largest lava flows on Earth \cite{Doyle2016a, Zhang2021b}.
+
+% Based on high-precision $^{206}$Pb/$^{238}$U geochronology and paleomagnetic data, \cite{Swanson-Hysell2021a} hypothesized that the massive Duluth Complex and the coeval North Shore Volcanic group (Fig.
+\ref{fig:Geologic_map}) formed as a result of upwelling mantled-derived plume encountering Laurentia's thinned lithosphere, following an ``upside-down” drainage" which rapidly funneled melt to today's northeastern Minnesota (Fig.
 \ref{fig:Geologic_map}) \cite{Sleep1997a, Swanson-Hysell2014a}).
@@
 
-The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex punctuates another period of magmatism during the Midcontinent Rift main stage activity.
-$^{206}$Pb/$^{238}$U zircon dates and cross-cutting relationships tightly constrain the areally extensive Beaver River diabase of the Beaver Bay Complex to have been emplaced at 1091.7 $\pm$ 0.2 Ma \cite{Zhang2021b}.
-The Beaver River diabase intruded into hypabyssal depths via wide conduits that accommodated the entrainment of numerous anorthosite xenoliths that have short-axis diameters up to 180 meters \cite{Boerboom2004a, Boerboom2006b}.
-Geochronological, paleomagnetic, geochemical, and petrographic data are consistent with the hypothesis that the Beaver River diabase magma extruded to the surface, feeding the massive Greenstone Flow - one of the largest lava flows on Earth \cite{Doyle2016a, Zhang2021b}.
-
-% The anorthosite xenoliths that distinctly occur within the Beaver River diabase are not only recorders of the wide diabase conduit sizes,  of the geomagnetic field $\sim$1091.7 million years ago.
-Thermal modeling result and paleomagnetic directional data show the anorthosite xenoliths acquired thermal remanent magnetizations during cooling of the diabase \cite{Zhang2021b}.
-Step-wise thermal demagnetizations show many anorthosite xenoliths have minimal secondary remanence and single-component characteristic remanence magnetizations that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C.
-
-Despite that the anorthosite xenoliths of the Beaver River diabase were entrained as xenoliths and emplaced only $\sim$4 Myr after the formation of the Anorthositic Series rocks of the Duluth Complex when massive plagioclase mushes were generated, $^{206}$Pb/$^{238}$U zircon geochronology data and zircon rare earth element diffusion modeling results suggest it is the most probable that the Beaver River diabase anorthosite xenoliths are entrained cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
+% Despite that the anorthosite xenoliths of the Beaver River diabase were entrained as xenoliths and emplaced only $\sim$4 Myr after the formation of the Anorthositic Series rocks of the Duluth Complex when massive plagioclase mushes were generated, $^{206}$Pb/$^{238}$U zircon geochronology data and zircon rare earth element diffusion modeling results suggest it is the most probable that the Beaver River diabase anorthosite xenoliths are entrained cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
 The anorthosite xenoliths of the \textit{ca.} 1092 Ma Beaver River diabase has distinct mineralogy from that of the Anorthositic Series rocks of the Duluth Complex.
@@
 
-Occurrences of magnetite and ilmenite symplectites and needles within anorthositic rocks have been described by \cite{Scofield1975a, Scofield1986a} in the Wichita Mountains and by \cite{Haggerty1976a} in other slowly cooled, mafic intrusions.
-\cite{Scofield1986a} hypothesized that the inclusions acted as a means of buffering the melt with respect to oxygen, forming either crystallites along preexisting plagioclase substrate or as iron incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures.
-Such a model requires that plagioclase, instead of other mafic minerals such as olivine, to be the first to crystallize from a magma and act as a sink for iron.
-The two-stage formation of the Duluth Complex Anorthositic Series fits this model \cite{Miller1990a}.
-During the first stage, high pressure at subcrustal depths (15-25 km depth) caused fractionation of Duluth Complex magma resulting in the formation of floating plagioclase mushes \cite{Kushiro1980a}.
-The second stage of crystallization then occurs at low pressures after the magma intruded into shallower chambers.
-
-This proposed magmatic evolution of the Duluth Complex is in distinct contrast to the emplacement of the Beaver River diabase which rapidly ascended and entrained the anorthosite xenoliths directly into shallow chambers \cite{Flower1980a, Miller1997a, Zhang2021b}.
-Particularly, \cite{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
-This is consistent with the interpretation that these anorthosite were rapidly transported by upward currents into hypabyssal magma chambers without reaching structural equilibrium.
-Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Fig.
-\ref{Petro_QDM}).
-Instead, relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
-\ref{Petro_QDM}).
-
 \section*{Results and Discussion}
@@
 \ref{fig:IZZI_examples}.
-Summary specimen absolute paleointensity estimates and site-level weighted mean paleointensity values are shown in Table \ref{tab:PINT_result} and are plotted in Fig.
+Summary specimen absolute paleointensity estimates and site-level weighted mean paleointensity values are shown in Table S1 and are plotted in Fig.
 \ref{fig:PINT_cooling_corrected}.
@@
 The Beaver River anorthosite specimens yielded a high pass rate ($\sim$46.5\%) through our selection criteria.
-In distinct contrast, the Beaver River diabase specimens are all rejected.
+In contrast, the Beaver River diabase specimens are all rejected.
 Paleointensity experiment results from both lithologies show three types of dominant specimen behaviors on Arai plots (Fig.
@@
 \centering
-\caption{\small{Summary plot of cooling rate corrected individual specimen absolute paleointensity results (square and diamond symbols) from this study and their averages at site (grey bar) and locality level (weighted mean by site-level standard deviations; orange bar).
+\caption{\small{Summary plot of individual specimen absolute paleointensity results (square and diamond symbols) from this study and their averages at site (grey bar) and locality level (weighted mean by site-level standard deviations; orange bar).
+All results are corrected for cooling rate.
 }}
@@
 
-The 20 mT alternating field treatment after in-field heating steps did not make any diabase specimen to pass the selection criteria.
+A 20 mT alternating field treatment after in-field heating steps were applied to some specimens in an attempt to mitigate non-ideal behavior contributed by multidomain or non-uniformly magnetized magnetic carriers.
+The treatment did not make any diabase specimen to pass the selection criteria.
 On the other hand, 27 out of 58 ($\sim$46.6\%) anorthosite specimens that had AF treatment passed selection and 13 out of 28 ($\sim$46.4\%) specimens without AF treatment passed selection.
@@
 Overall, the consistent specimen-level behavior makes the successful results from the anorthosites of high confidence.
-The paleointensity quality index (Q$_{PI}$) for the anorthosite sites are all above 3 and the summary statistics are shown in Table \ref{tab:QPI}.
+The paleointensity quality index (Q$_{PI}$) for the anorthosite sites are all above 3 and the summary statistics are shown in Table S2.
 
-In addition to applying paleointensity selection criteria to our experiment results, we applied the bias corrected estimation of paleointensity (BiCEP) method by \cite{Cych2021a} to all of our anorthosite and diabase specimens regardless of their Arai plot statistics.
-The BiCEP method is a Bayesian statistical method based on an assumption that all specimen paleointensity results from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the Arai plot \cite{Paterson2011a}.
+In addition to applying paleointensity selection criteria to our experiment results, we applied the bias corrected estimation of paleointensity (BiCEP) method by Cych et al \cite{Cych2021a} to all of our anorthosite and diabase specimens regardless of their Arai plot statistics.
+The BiCEP method is a Bayesian statistical method based on an assumption that all specimen paleointensity results from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the Arai plot \cite{Arai1963a, Paterson2011a}.
 This method has the advantage of using as many specimens as available for estimating site-level paleointensities.
-The posterior distributions of paleointensity estimates given by the BiCEP method show that for the anorthosites that pass our paleointensity selection criteria, probabilities are very low for the case where these anorthosites are recording heavily biased paleointensity estimates from a true value below 10 $\mu$T (Supporting Information S1).
-Therefore, the anorthosite xenoliths record a high geomagnetic field \textit{ca.} 1092 Ma.
+The posterior distributions of paleointensity estimates given by the BiCEP method show that for the anorthosites that pass our paleointensity selection criteria, probabilities are very low for the case where these anorthosites are recording heavily biased paleointensity estimates from a true value below 10 $\mu$T (Fig.
+S3).
+Instead, posterior probabilities for sites with high-quality specimen-level data are in agreement with site-level weighted-mean results, consistent with the interpretation that the anorthosite xenoliths record a high geomagnetic field \textit{ca.} 1092 Ma.
 
@@
 Rock magnetic data support that the successful anorthosite specimens have dominant magnetic carriers with magnetic properties similar to stochiometric, non-interactive, single domain magnetite, whereas those anorthosites that failed the paleointensity result selection and all diabase have more pronounced populations of non-ideal carriers.
-MPMS data show sister specimens to paleointensity specimens typically have dominant magnetic mineralogy of magnetite as can be evidenced through the Verwey transitions (Fig.
-\ref{fig:MPMS}; \cite{Verwey1939a}).
+Magnetic property measurement system (MPMS) data show sister specimens to paleointensity specimens typically have dominant magnetic mineralogy of magnetite as can be evidenced through the Verwey transitions (Fig.
+S1; \citealp{Verwey1939a}).
 However, specimens from sites that yield successful paleointensity results display Verwey transition near 120 Kelvin (K) as expected for stochiometric low-Ti magnetite \cite{Ozdemir1993a}.
 On the other hand, the diabase and anorthosite specimens that did not pass our paleointensity selection typically have suppressed Verwey transition toward lower temperatures (Fig.
-\ref{fig:MPMS}), indicating that the specimens likely have a ferromagnetic mineralogy with variable degrees of oxidation \cite{Ozdemir1993a}.
+S1), indicating that the specimens likely have a ferromagnetic mineralogy with variable degrees of oxidation \cite{Ozdemir1993a}.
 
@@
 A compilation of median destructive field (MDF) in Fig.
-\ref{fig:coercivity} based on all of our backfield experiments shows that specimens from anorthosites that pass paleointensity selections have distinctly a higher average MDF range than other anorthosites and all diabase.
-Example single component skew-normal distribution fits of coercivity spectra based on backfield experiments on specimen chips sister to paleointensity specimens show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity distributions around 80 mT (Fig.
+\ref{fig:coercivity} based on our backfield experiments shows that specimens from anorthosites that pass paleointensity selections have a distinctly higher average MDF range than other anorthosites and all diabase.
+Example single component skew-normal distribution fits of coercivity spectra based on backfield experiments on rock chips sister to paleointensity specimens show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity distributions around 80 mT (Fig.
 \ref{fig:coercivity}).
@@
 \cite{Selkin2000a} demonstrated that strong remanence anisotropy associated with the igneous foliation developed within an anorthosite sample from the Stillwater Complex can lead to paleointensity estimates that are different by three times.
-Depending on the orientation of the applied field during in-field steps and that of the magnetic anisotropy axes of a specimen, such strong preferred orientations of petrofabrics can result in significantly biased paleointensity estimates.
+Depending on the orientation of the applied field during in-field steps and that of the magnetic anisotropy axes of a specimen, such strong magnetic fabrics associated with petrofabrics can result in significantly biased paleointensity estimates.
 To assess whether our paleointensity estimates are biased by remanence anisotropy, we calculated the gamma statistic, which is the angular difference between the last pTRM step used for paleointensity determination and the applied field direction (Table \ref{tab:PINT_result}).
@@
 These gamma values are similar to those of the Midcontinent Rift volcanics \cite{Sprain2018a} and they show that the Beaver River anorthosite bulk rocks do not have a significant remanence anisotropy.
-This result can be further supported by the fact that the recovered paleodirectional data from our anorthosite xenoliths closely match those from the paired Beaver River diabase hosts \cite{Zhang2021b}.
+This result can be further supported by the fact that the recovered paleodirectional data from our anorthosite xenoliths closely match those from the paired Beaver River diabase hosts \cite{Zhang2021b} and that no strong magnetic anisotropy was observed through magnetic imaging.
 \cite{Selkin2007a} also noted that vortex state magnetic grains and certain single domain particles can produce biased paleointensity estimates due to their acquisition of saturation remanence during laboratory in-field steps.
-According to the numerical modeling of \cite{Selkin2007a}, thermoremanent magnetization acquired by a randomly-oriented population of elongate single-domian-sized magnetite particles are unlikely to saturate upon our applied field of 30 $\mu$T.
+According to the numerical modeling of \cite{Selkin2007a}, thermoremanent magnetization acquired by a randomly-oriented population of elongate single-domian magnetite particles are unlikely to saturate upon our applied field of 30 $\mu$T.
 Therefore, in this study we do not apply any anisotropy or nonlinear TRM correction to our specimens.
@@
 
-The reason that the Beaver River anorthosite xenoliths do not display strong remanence anisotropy, nor do they display nonlinear TRM acquisition problem, could be that they formed in a magmatic setting distinct from that of a layered mafic complex, such as the nearby Duluth Complex, where igneous foliation can occur (Fig.
+% The reason that the Beaver River anorthosite xenoliths do not display strong remanence anisotropy, nor do they display nonlinear TRM acquisition problem, could be that they formed in a magmatic setting distinct from that of a layered mafic complex such as the nearby Duluth Complex, where igneous foliation can occur (Fig.
 \ref{fig:Geologic_map}).
@@
 Paleointensity estimates from igneous intrusions can also be biased toward higher values due to slow cooling rates \cite{Halgedahl1980a}.
-Because the cooling rates of intrusive rocks could be much longer than the extrusive rocks, the slowly cooled diabase and anorthosite can bias laboratory paleointensity estimates toward an overestimate for SD grains \cite{Dodson1980a, Halgedahl1980a, Selkin2000a}.
+Because the cooling rates of intrusive rocks could be much longer than the extrusive rocks, the slowly cooled Beaver River diabase and anorthosite xenoliths therein can bias laboratory paleointensity estimates toward an overestimate for single domain grains \cite{Dodson1980a, Halgedahl1980a, Selkin2000a}.
 
-From the thermal history model we can also estimate the duration which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully block in their characteristic natural remanence component ($\sim$500\textdegree C; \cite{Zhang2021b}).
+From the thermal history model of ref.
+\citealp{Zhang2021b} we can also estimate the duration which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully block in their characteristic natural remanence component ($\sim$500\textdegree C; \cite{Zhang2021b}).
 We find the cooling time to be on the order of $\sim$1 kyr, which corresponds to a cooling rate of $1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
-In contrast, the lab cooling rate is much faster, at about $1.3\times10^{-1}$ $^\circ$C s$^{-1}$ if we assume it takes 1000s in lab condition to cool through the same temperature interval.
-As a result, this significant cooling rate difference leads to a prediction of $\sim$36\% overestimate of true ancient field \cite{Halgedahl1980a}.
+In contrast, the lab cooling rate is much faster, at about $1.3\times10^{-1}$ $^\circ$C s$^{-1}$, assuming that it takes 1000s in lab condition to cool through the same temperature interval.
+This significant cooling rate difference leads to a predicted $\sim$36\% overestimate of true ancient field following the model of ref.
+\citealp{Halgedahl1980a} (Fig.
+S2).
 We therefore correct our paleointensity results by a factor of 0.74.
@@
 To best characterize the geomagnetic axial dipole field intensity during a certain time period requires that a paleointensity dataset to cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged and the geocentric axial dipole hypothesis (GAD) can be assumed.
-Thermal modeling results from \cite{Zhang2021b} suggest that the anorthosite xenoliths were likely heated to tholeiitic magma temperature and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a few thousand years.
+Thermal modeling results from ref.
+\citealp{Zhang2021b} suggest that the anorthosite xenoliths were likely heated to tholeiitic magma temperature and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a few thousand years.
 In addition, the closely located sites AX12 and AX16 have very similar paleointensity estimates that are distinct from those of sites AX11 and AX13 which were emplaced $\sim$10 km away (Fig.
@@
 The consistency within small regions but variations between regions could be the result of differential cooling history due to heterogeneity of the timing and depths of the Beaver River diabase intrusions, causing the anorthosites to capture paleosecular variations \textit{ca.} 1092 Ma.
-Overall, none of the anorthosite xenoliths record a V(A)DM value smaller than 40 ZAm$^2$ (Fig.
+Overall, none of the anorthosite xenoliths record a site mean VDM value lower than 40 ZAm$^2$ (Fig.
 \ref{fig:PINT_compilation}).
@@
 In particular, some models have suggested that the kick-start of rigorous compositional convection between the inner core and the outer core soon after the onset of inner core nucleation can result in almost an order of magnitude increase in Earth's axial dipole moment \cite[e.g.]{Davies2021a}.
-The enriching observational records from the Ediacaran and Cambrian have been suggested to indicate that a geodynamo driven solely by thermal convection was waning to its brink of collapse near the end of Ediacaran \cite{Bono2019a, Thallner2021b} and was recovering to ward modern day strengths soon after \cite{Thallner2021a}.
+The enriching observational records from the Ediacaran and Cambrian have been suggested to indicate that a geodynamo driven solely by thermal convection was waning to its brink of collapse near the end of Ediacaran \cite{Bono2019a, Thallner2021b} and was recovering toward modern-day strengths soon after \cite{Thallner2021a}.
 
-However, the updated paleointensity records from the Midcontinent Rift together with previous results from other Mesoproterozoic rocks suggest that the a dramatic increase in geomagnetic dipole moment similar in scale to that modeled for after the onset of inner core nucleation may not be unique to the Edicaran-Cambrian boundary before 500 Ma (Fig.
+However, the updated paleointensity records from the Midcontinent Rift together with previous results from other Mesoproterozoic rocks suggest that the a dramatic increase in geomagnetic dipole moment similar in scale to that modeled for after the onset of inner core nucleation may not be unique to the Ediacaran-Cambrian boundary (Fig.
 \ref{fig:PINT_compilation}).
 The \textit{ca.} 1140 Ma Abitibi dikes \cite{Macouin2003a} record an average virtual dipole moment of $\sim$13 ZAm$^2$, similar to the low values of $\sim$7 ZAm$^2$ obtained from the single silicate crystals of the \textit{ca.} 565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a}.
-However, the relatively high virtual dipole moments of $\sim$72 ZAm$^2$ from the \textit{ca.} 1105 Ma Osler Volcanic Group \cite{Sprain2018a} and $\sim$82 ZAm$^2$ from the \textit{ca.} 1092 Ma Beaver River anorthosite xenoliths (Fig.
-\ref{fig:PINT_compilation} are about six times the values from the \textit{ca.} 1140 Ma Abitibi dikes.
-In addition, the increase in absolute dipole moment estimates recorded by the Midcontinent Rift rocks are much higher than that in the Ediacaran (Fig.
+However, the relatively high moments of $\sim$72 ZAm$^2$ from the \textit{ca.} 1105 Ma Osler Volcanic Group \cite{Sprain2018a} and $\sim$89 ZAm$^2$ from the \textit{ca.} 1092 Ma Beaver River anorthosite xenoliths (Fig.
+\ref{fig:PINT_compilation} are about six times the values from the Abitibi dikes.
+This increase in absolute dipole moment estimates during $\sim$50 Myr recorded by the Midcontinent Rift rocks are much higher than that in the Ediacaran (Fig.
 \ref{fig:PINT_compilation}).
 
-Furthermore, the high geomagnetic field estimates from the Midcontinent Rift volcanics and intrusives are not the only suite of rocks that bear records at odds with the hypothesis of a progressive decay trend during the Proterozoic \cite{Biggin2009a, Bono2019a}.
+The high geomagnetic field estimates from the Midcontinent Rift volcanics and intrusives are not the only suite of rocks that bear records at odds with the hypothesis of a progressive decay trend during the entire Proterozoic Eon \cite{Biggin2009a, Bono2019a}.
 The $\sim$60 ZAm$^2$ dipole moment estimate from the \textit{ca.} 755 Ma Mundine Well dikes are discordant with the long-term decrease trend \cite{Bono2019a} and is about three times the estimates from the \textit{ca.} 720 Ma dolerite dikes of the Franklin Large Igneous Province \cite{Lloyd2021a}.
@@
 \ref{fig:PINT_compilation}).
-However, a sparse record in a short time period may not sufficiently average out paleosecular variations.
+These data suggest that an observation of a strong increase or decrease in geomagnetic dipole moment may not be unique during the Precambrian.
+A sparse record in a short time period may not sufficiently average out paleosecular variations and more spatial and temporal coverage of the observational records is needed for further analyses of the long-term evolution of the geodynamo.
 Given that a five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years (5 and 95 percentiles), when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed, we cannot conclude on whether an apparent rise or decay within a narrow time period ($\sim$35 Ma) during the Proterozoic is associated with transitions in the core fluid regime or is a result of short-term variations of the geomagnetic field that may be common through much of the evolution of Earth's core.
-In addition, the observations of geomagnetic dipole moment fluctuations in more recent time scales may suggest that evaluations of long-term evolution of the Precambrian paleointensity records should be accompanied by development of higher-resolution data in more short-term scales.
+Regardless, the high geomagnetic dipole field moment estimate from the Beaver River anorthosite xenoliths necessitates that there being a vigorous late Mesoproterozoic geodynamo.
 
-Overall, the existing absolute paleointensity records in the PINT database remains sparse in terms of spatial and temporal coverage as compared to more geologically recent times.
+% Overall, the existing absolute paleointensity records in the PINT database remains sparse in terms of spatial and temporal coverage as compared to more geologically recent times.
 (Fig.
@@
 
-\section*{Conclusion}
+% \section*{Conclusion}
 
-The paucity of Precambrian paleointensity estimates, as well as non-ideal results within the existing database, present challenges when seeking to interpret long-term trends in the strength of the geomagnetic field.
+% The paucity of Precambrian paleointensity estimates, as well as non-ideal results within the existing database, present challenges when seeking to interpret long-term trends in the strength of the geomagnetic field.
 The Midcontinent Rift of North America presents an opportunity to develop robust paleointensity estimates spanning a $\sim$25 Myr interval of the late Mesoproterozoic.
@@
 
-The new data in context of the current Precambrian paleointensity database suggest that a strong increase in geomagnetic dipole moment which is thought to be indicative of the onset of compositional convection caused by the inner core nucleation may not be unique during the Precambrian.
+% The new data in context of the current Precambrian paleointensity database suggest that an observation of a strong increase in geomagnetic dipole moment may not be unique during the Precambrian.
 More spatial and temporal coverage of the observational records is needed for further analyses of the long-term evolution of the geodynamo.

### 2021-11-09 PNAS_manuscript/PNAS_manuscript.tex
manuscript update

@@
 \significancestatement{Acquiring high-fidelity observational records of ancient magnetic field intensity from remanent magnetization of rocks is crucial for constraining the evolution history of Earth's core.
-However, such records are often difficult to recover due to the alterations associated with weathering after the formation of rocks.
-In this study, we use plagioclase cumulates from the North American Midcontinent Rift that has experienced minimal post-formation alterations to obtain a record of $\sim$89 ZAm$^2$ virtual dipole moment 1.1 billion years ago.
+However, such records are often difficult to recover due to the alterations associated with weathering after the formation of rocks and heating during experiments.
+In this study, we use plagioclase cumulates from the North American Midcontinent Rift that has experienced minimal post-formation alterations to obtain a present-day-like virtual dipole moment 1.1 billion years ago.
 The results indicate the existence of strong power sources for Earth's dynamo during the late Mesoproterozoic.
@@
 % \equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
-\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: yimingzhang\@berkeley.edu}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: yimingzhang\@berkeley.edu.}
 
 % At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
-\keywords{absolute paleointensity $|$ Laurentia $|$ Midcontinent Rift $|$ anorthosite $|$ geodynamo $|$}
+\keywords{absolute paleointensity $|$ Laurentia $|$ Midcontinent Rift $|$ anorthosite $|$ geodynamo $|$ Mesoproterozoic $|$ inner core}
 
@@
 Higher conductivity values imply faster cooling rates, which in turn imply that the threshold for the freezing of the inner core happened more recently \cite{Davies2015a}.
-While some estimates of these values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates based on first principles calculations implied higher thermal conductivity values and an age for the inner core that is less than 1.3 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b}, with some suggesting even younger ages ($<$ 700 Ma) \cite{Labrosse2015a, Ohta2016a}.
-With the estimates for the core's thermal conductivity continue to be debated, it is crucial to use observational records as an independent constraint on the thermal evolution of the core and mantle.
+While some estimates of these values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates based on first principles calculations implied higher thermal conductivity values and an age for the inner core that is less than 1.3 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b}, with some suggesting even younger ages ($<$ 700 Ma; ref.s \citealp{Labrosse2015a, Ohta2016a}.
+With the estimates for the core's thermal conductivity continue to be debated, it is crucial to use observational records as an independent constraint on the thermal evolution of Earth's core and mantle.
 
@@
 
-Paleomagnetic data from ancient rocks are one of the few types of observational data that have the potential to record the evolution of Earth’s core and can be used to test the predictions from numerical simulations.
-Previous interpretations of the current Precambrian absolute paleointensity records (compiled in the PINT database; \url{http://earth.liv.ac.uk/pint/}) suggest that an appreciable geomagnetic field existed in the Archean and progressively decayed through the Proterozoic till it reached its weakest in the Ediacaran when very low virtual dipole moment (VDM) estimates have been obtained \cite{Bono2019a, Shcherbakova2019a, Thallner2021b, Lloyd2021a}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
-Bono et al \cite{Bono2019a} adopted a billion-year-scale view of such observations of the Precambrian geomagnetic field intensities to suggest that these evidence are consistent with models that use high core thermal conductivity values where the nucleation of the inner core began to provide compositional convection power for driving the geodynamo during the Ediacaran-Cambrian after experiencing a progressive decay when the geodynamo was dominantly driven by thermal conduction earlier in the Proterozoic \cite{Driscoll2016a, Davies2021a}.
+Paleomagnetic records from ancient rocks are one of the few types of observational data that have the potential to record the evolution of Earth’s core and can be used to test the predictions from numerical simulations.
+Interpretations of selected records of the current Precambrian absolute paleointensity database (compiled in the PINT database; \url{http://earth.liv.ac.uk/pint/}) suggest that an appreciable geomagnetic field existed in the Archean but reached its weakest in the Ediacaran when very low virtual dipole moment (VDM) estimates have been obtained \cite{Bono2019a, Shcherbakova2019a, Thallner2021b, Lloyd2021a}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
+Ref.
+\citealp{Bono2019a} adopted a billion-year-scale view of such observations to suggest that these data are consistent with thermal evolution models that use high core thermal conductivity values, where the nucleation of the inner core did not begin to provide compositional convection power for driving the geodynamo until the Ediacaran-Cambrian, suggesting that the Proterozoic dynamo might have experienced a progressive decay when compositional convection was lacking and thermal convection was decreasing \cite{Driscoll2016a, Davies2021a}.
 
@@
 
-However, previous paleointensity data from intrusive and extrusive rocks in the \textit{ca.} 1.1 Ga Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record VDM values similar to that of today, which are higher than the values predicted by the long term decay curve proposed by Bono et al \cite{Bono2019a} (Fig.
+However, previous paleointensity data from intrusive and extrusive rocks in the \textit{ca.} 1.1 Ga Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record VDM values similar to that of today, which are higher than the values predicted by the long term decay curve proposed by ref.
+\citealp{Bono2019a} (Fig.
 \ref{fig:PINT_compilation}).
-Pesonen and Halls \cite{Pesonen1983b} obtained an average virtual dipole moment of 93 $\times 10^21$ Am$^2$ (ZAm$^2$) after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using both extrusive and intrusive rocks from the Midcontinent Rift, including the \textit{ca.} 1108-1105 Ma Osler Volcanics, the \textit{ca.} 1100 Ma Maimainse Point volcanics, \textit{ca.} 1107 Ma intrusive Logan Sills, Thunder Bay dikes, the Baraga and Marquette dikes, as well as lava flows of the \textit{ca.} 1096 Ma North Shore Volcanic Group (Fig.
+Ref.
+\citealp{Pesonen1983a} obtained an average virtual dipole moment of 93 $\times 10^{21}$ Am$^2$ (ZAm$^2$) after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using both extrusive and intrusive rocks from the Midcontinent Rift, including the \textit{ca.} 1108-1105 Ma Osler Volcanics, the \textit{ca.} 1100 Ma Maimainse Point volcanics, the \textit{ca.} 1107 Ma intrusive Logan Sills, Thunder Bay dikes, the Baraga and Marquette dikes, as well as lava flows of the \textit{ca.} 1096 Ma North Shore Volcanic Group (Fig.
 \ref{fig:Geologic_map}).
-Kulakov et al \cite{Kulakov2013a} obtained an average VDM of 59 $\pm$ 11 ZAm$^2$ using the \textit{ca.} 1085 Ma Lake Shore Traps volcanics and suggest that the present-day-like geomagnetic field strength likely existed through the late stage of Midcontinent Rift magmatism \cite{Vervoort2007a, Miller2013a}.
+Ref.
+\citealp{Kulakov2013a} obtained an average VDM of 59 $\pm$ 11 ZAm$^2$ using the \textit{ca.} 1085 Ma Lake Shore Traps volcanics and suggest that the present-day-like geomagnetic field strength likely existed through the late stage of Midcontinent Rift magmatism \cite{Vervoort2007a, Miller2013a}.
 However, non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results obtained by ref.
-\citealp{Pesonen1983b, Kulakov2013a}.
+\citealp{Pesonen1983a, Kulakov2013a}.
 One typical failure of paleointensity experiment is the double-slope behavior, as two distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
-Pesonen and Halls \cite{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas Kulakov et al \cite{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
-Results with such non-ideal behavior were rejected by Sprain et al \cite{Sprain2018a} who applied stricter paleointensity selection criteria on new paleointensity results and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
+Ref \citealp{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas ref.
+\citealp{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
+Results with such non-ideal behavior were rejected by ref.
+\citealp{Sprain2018a} who applied stricter paleointensity selection criteria on new paleointensity results and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
 A calculated mean VDM value of 56 $\pm$ 21 ZAm$^2$ is similar to the average value of the past 300 million years \cite{Sprain2018a}.
-This high value from the Midcontinent Rift volcanics may be incompatible with the hypothesis of a ``Proterozoic dipole low" \cite{Biggin2009a}.
 
-Although it is typical of paleointensity experiments to suffer from low success rates, bulk plagioclase-rich rocks \cite{Selkin2000a} and single plagioclase crystals \cite{Tarduno2005a} are found to be able to produce high quality paleointensity results.
+Although it is typical of paleointensity experiments to have low success rates, bulk plagioclase-rich rocks \cite{Selkin2000a} and single plagioclase crystals \cite{Tarduno2005a} are found to be able to produce high quality paleointensity results.
 It has been suggested that plagioclase crystals can protect magnetic minerals that exsolved within them against post-formation weathering and thermochemical alteration during paleointensity experiments \cite{Tarduno2005a}.
-In this study, we present new paleointensity results from ``IZZI" modified absolute paleointensity experiments \cite{Yu2004a} on a suite of bulk samples from \textit{ca.} 1092 Ma anorthosite xenoliths and their Beaver River diabase host rocks of the Beaver Bay Complex within the Midcontinent Rift.
+In this study, we present new paleointensity results from IZZ modified absolute paleointensity experiments \cite{Yu2004a} on a suite of bulk samples from \textit{ca.} 1092 Ma anorthosite xenoliths and their Beaver River diabase host rocks of the Beaver Bay Complex within the Midcontinent Rift.
 Paleointensity experiments on the anorthosite xenoliths have a very high success rate.
@@
 Magnetic imaging shows the anorthosite specimens have dominant magnetic carriers interstitial to and within plagioclase crystals and they do not display strong preferred orientations.
-With more single-slope, high quality paleointensity data, further enrich the reliable paleointensity records during the $\sim$25 Myr of Midcontinent Rift magmatic activity.
+Our single-slope, high quality paleointensity data add to the reliable paleointensity records during the $\sim$25 Myr of Midcontinent Rift magmatic activity.
 Our new data from the Midcontinent Rift intrusive rocks support previous data from the volcanics that there was a strong late Mesoproterozoic geodynamo.
@@
 The Mesoproterozoic North American Midcontinent Rift is an intracontinental rift that did not result in the break-up of Laurentia.
-Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; \cite{Vervoort2007a, Miller2013a}) was characterized by rapid and voluminous emplacement of intrusive rocks and extrusive volcanics.
-The $<$500 Kyr emplacement of the \textit{ca.} 1096 Ma Duluth Complex and the $~$8 km thick volcanics associated with it is one manifestation \cite{Swanson-Hysell2021a}.
-The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex punctuates another such period of magmatism during the Midcontinent Rift main stage activity.
-High precision $^{206}$Pb/$^{238}$U zircon dates from one anorthosite xenolith and units that have cross-cutting relationships with the Beaver River diabase tightly constrain the timing of diabase emplacement to be at 1091.7 $\pm$ 0.2 Ma \cite{Zhang2021b}.
-The xenolith-bearing diabase dikes and sills intruded into hypabyssal depths via wide conduits that accommodated the entrainment of numerous anorthosite xenoliths that have short-axis diameters ranging from centimeters to up to 180 meters \cite{Boerboom2004a, Boerboom2006b}.
+Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; refs.
+\citealp{Vervoort2007a, Miller2013a}) was characterized by rapid and voluminous emplacement of intrusive rocks and extrusive volcanics.
+The $<$500 Kyr emplacement of the \textit{ca.} 1096 Ma Duluth Complex and the associated $\sim$8 km thick extrusive volcanics is one manifestation \cite{Swanson-Hysell2021a}.
+The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex in northern Minnesota punctuates another such period of magmatism during the Midcontinent Rift main stage activity.
+The hypabyssal intrusion of the 1091.7 $\pm$ 0.2 Ma Beaver River diabase dikes and sills of the Beaver Bay Complex transported numerous anorthosite xenoliths that have short-axis diameters up to 180 meters via wide conduits \cite{Boerboom2004a, Boerboom2006b}.
+Geochronology, geochemical, and paleomagnetic data support the hypothesis that the diabase extruded to the surface and fed the coeval voluminous Greenstone Flow of the Portage Lake volcanics \cite{Doyle2016a, Zhang2021b}.
 
-Previous geochronology data and rare earth element diffusion modeling suggest that the anorthosite xenoliths of the Beaver River diabase are cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
-Petrological textures and geochemical attributes imply that the Beaver River anorthosite xenoliths formed in a lower crustal magmatic condition distinct from that of the Duluth Complex which is $\sim$4 Myr older and also generated massive anorthositic cumulates that led to the formation of the Anorthositic Series \cite{Zhang2021b, Swanson-Hysell2021a}.
+Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths acquired thermal remanent magnetizations during cooling within the Beaver River diabase \cite{Zhang2021b}.
+Step-wise thermal demagnetizations show the anorthosite xenoliths have minimal secondary remanence and single-component characteristic remanence magnetizations that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C.
+Geochronology data and rare earth element diffusion modeling further suggest that the anorthosite xenoliths are cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
+Petrological textures and geochemical attributes imply that the Beaver River anorthosite xenoliths formed in a lower crustal magmatic conditionm \cite{Zhang2021b, Miller1997a} distinct from that of layered mafic intrusions, such as the Duluth Complex which is $\sim$4 Myr older and generated massive anorthositic cumulates that led to the formation of the Anorthositic Series \cite{Swanson-Hysell2021a, Miller1990a}.
 The dominantly monomineralic anorthosite xenoliths of the Beaver River diabase have an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}, higher than the average of $\sim$60\% of the Anorthositic Series rocks of the Duluth Complex \cite{Miller1990a}.
 The plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture showing igneous foliations (Fig.
-\ref{Petro_QDM}A), but the more calcium-rich plagioclase of the Beaver River anorthosites often develop granoblastic texture characterized by equigranular crystals with poor alignments (Fig.
-\ref{Petro_QDM}B).
+\ref{fig:Petro_QDM}A), but the more calcium-rich plagioclase of the Beaver River anorthosites often develop granoblastic texture characterized by equigranular crystals with poor alignments (Fig.
+\ref{fig:Petro_QDM}B).
 Abundant magnetite and ilmenite needles that are typically 10 $\mu$m long can be found to have preferred orientations within Duluth Complex plagioclase.
-Interstitial symplectic titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex Anorthositic Series rocks \cite{Miller1990a}.
-Such a high concentration of Fe oxides carrying magnetic remanence in the slowly cooled Duluth Complex anorthositic rocks reflect that iron in the plagioclase mush in the magma chamber acted as a means of buffering the melt with respect to oxygen, forming crystallites along preexisting plagioclase substrate or as iron oxides incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures \cite{Scofield1986a}.
+Interstitial titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex plagioclase \cite{Miller1990a}.
 In contrast, titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent but relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
-\ref{Petro_QDM}).
-This likely reflects that the Beaver River diabase magma had a different magma oxidation history than that of the Duluth Complex.
+\ref{fig:Petro_QDM}).
+This reflects that the Beaver River diabase magma had a different oxidation history than that of the Duluth Complex.
 The plagioclase and oxide mineral textures of the Beaver River anorthosites also distinguish them from plagioclase cumulates of other layered mafic complex where magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
 
+% Such a high concentration of Fe oxides carrying magnetic remanence in the slowly cooled Duluth Complex anorthositic rocks reflect that iron in the plagioclase mush in the magma chamber acted as a means of buffering the melt with respect to oxygen, forming crystallites along preexisting plagioclase substrate or as iron oxides incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures \cite{Scofield1986a}.
+In contrast, titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent but relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
+\ref{fig:Petro_QDM}).
 % the Beaver River diabase rapidly ascended and entrained the anorthosite xenoliths directly into shallow chambers and likely did not reach  \cite{Flower1980a, Miller1997a, Zhang2021b}.
@@
 Red diamonds mark sites with paleointensity results that passed the selection criteria from this study.
-Paleomagnetic sites from \cite{Pesonen1983a} are categorized by lithology.
-All sites from \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} are color-coded by their ages.}}
+Paleomagnetic sites from ref.
+\citealp{Pesonen1983a} are categorized by lithology.
+All sites from refs.
+\citealp{Pesonen1983a, Kulakov2013a, Sprain2018a} are color-coded by their ages.}}
 \label{fig:Geologic_map}
@@
 \noindent\includegraphics[width=11.4 cm]{Petro_QDM.png}
-\caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from Duluth Complex Anorthositic Series (left column) and Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (right column; \cite{Zhang2021b}).
+\caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from Duluth Complex Anorthositic Series (left column) and Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (right column; ref.
+\citealp{Zhang2021b}).
 Petrographic image (A) shows Duluth Complex anorthosite have euhedral, interlocking plagioclase crystals showing igneous foliation.
@@
 
-40 from a total of 86 anorthosite specimens and 0 out of a total of 69 diabase specimens passed our paleointensity selection criteria, yielding 7 anorthosite sites and no diabase sites with site-level paleointensity results that pass paleointensity selections.
-Example NRM/TRM plots
-\cite{Arai1963a} are shown in Fig.
+40 from a total of 86 anorthosite specimens and 0 out of a total of 69 diabase specimens passed our paleointensity selection criteria, resulting in 7 anorthosite sites and no diabase sites with site-level paleointensity results that pass paleointensity selections.
+Example NRM/TRM plots are shown in Fig.
 \ref{fig:IZZI_examples}.
@@
 \centering
-\caption{\footnotesize{Example results of paleointensity experiments displayed on Arai and Zijderveld plots (insets) for anorthosite and diabase specimens.
+\caption{\footnotesize{Example results of paleointensity experiments displayed on NRM/TRM plots and zero-field heating results are shown on orthogonal plots (insets) for anorthosite and diabase specimens.
 Red (blue) circles indicate zero-field/in-field (in-field/zero-field) steps `ZI’ (`IZ’).
@@
 
-The Beaver River anorthosite specimens yielded a high pass rate ($\sim$46.5\%) through our selection criteria.
-In contrast, the Beaver River diabase specimens are all rejected.
-Paleointensity experiment results from both lithologies show three types of dominant specimen behaviors on Arai plots (Fig.
+The Beaver River anorthosite specimens yielded a high pass rate of $\sim$46.5\% through our selection criteria.
+In contrast, the diabase specimens are all rejected.
+Paleointensity experiment results from both lithologies show three types of dominant specimen behaviors on NRM/TRM plots (Fig.
 \ref{fig:IZZI_examples}): dominantly single-slope, double-slope, and zigzagging or sagging.
-Note that almost all specimens have minimal secondary remanence component in thermal demagnetization results as shown in example orthogonal plots of all zero-field steps (Fig.
+% Note that almost all specimens have minimal secondary remanence component in thermal demagnetization results as shown in example orthogonal plots of all zero-field steps (Fig.
 \ref{fig:IZZI_examples}).
-This reflects that the rocks have experienced mild alteration after their formation.
-Typical specimen paleointensity behavior of the anorthosites display straight, single-slope Arai plots and the accepted fractions of temperature steps span over the origin-trending, primary remanence components (Fig.
+This reflects that the rocks have experienced mild alterations after their formation.
+Typical specimen paleointensity behavior of the anorthosites display straight, single-slope NRM/TRM plots and the accepted fractions of temperature steps span over the origin-trending, primary remanence components (Fig.
 \ref{fig:IZZI_examples}).
 We accept all specimen- and site-level absolute paleointensity estimates from the anorthosite xenoliths that pass the selection.
-All diabase specimens failed the selection because of double-slope behavior (fail FRAC selection; see methods), poor pTRM checks, and zigzagging (fail SCAT, DANG selection; see methods) Arai plots (Fig.
+All diabase specimens failed the selection because of double-slope behavior (fail FRAC selection; see methods), poor pTRM checks, and zigzagging (fail SCAT, DANG selection; see methods) plots (Fig.
 \ref{fig:IZZI_examples}).
@@
 \ref{fig:IZZI_examples}, specimen BD2-5b has a high-temperature slope with zigzagging behavior, poor pTRM checks, and low NRM/TRM ratio, in addition to a low-temperature slope with a steep NRM/TRM ratio.
-Although the BD2 specimens' lower temperature fraction of the Arai plots are straight and the pTRM checks are consistent, lower temperature slopes from BD2 give consistently higher paleointensity estimates than those from site AX6, AX8, and AX10, which are anorthosite xenoliths included within BD2 (Fig.
-\ref{fig:Geologic_map}) \cite{Zhang2021a}.
-This higher estimate from the diabase lithology could be attributed to the systematic bias of using the lower temperature slope of the Arai plots demonstrated in Fig.
+Although the BD2 specimens' lower temperature fraction of the plots are straight and the pTRM checks are consistent, lower temperature slopes from BD2 give consistently higher paleointensity estimates than those from site AX6, AX8, and AX10, which are anorthosite xenoliths included within BD2 (Fig.
+\ref{fig:Geologic_map}; ref.
+\citealp{Zhang2021a}.
+This higher estimate from the diabase lithology could be attributed to the systematic bias of using the lower temperature slope of the NRM/TRM plots demonstrated in Fig.
 \ref{fig:IZZI_examples} and specimens from site BD2 all fail the criteria of having a maximum magnetization gap between adjacent steps of 25\% of the total NRM.
@@
 \caption{\small{Summary plot of individual specimen absolute paleointensity results (square and diamond symbols) from this study and their averages at site (grey bar) and locality level (weighted mean by site-level standard deviations; orange bar).
-All results are corrected for cooling rate.
+All results are corrected for cooling rate with a factor of 0.74.
 }}
@@
 \ref{fig:PINT_cooling_corrected}).
-Specimen plots for AX11 particularly illustrates a case where all specimens within the site yield almost the same paleointensity estimate, regardless of the AF treatment.
-Therefore, the alternating field demagnetization might have improve the specimen behavior for those anorthosites that were subjected to this treatment but it did not result in significant increase in specimen success rate and did not result in significant difference for the site-level paleointensity estimates.
-Together with the narrow range of unblocking temperatures of their sister specimens \cite{Zhang2021b}, these data are consistent with the interpretation that the anorthosites with successful paleointensity results have a magnetic mineral assemblage that likely contain dominantly single domain, low Ti magnetite and they do not contain a significant population of MD or vortex state grains.
-Some anorthosite specimens and all diabase specimens still failed in result selections even though AF treatment was applied.
-It could be that they have a very high dominance of MD or vortex state grains with relatively moderate coercivity such that an alternating field with 20 mT peak field did not significantly mitigate their non-ideal paleointensity behaviors.
+Specimen results for AX11 particularly illustrates a case where all specimens within the site yield almost the same paleointensity estimate, regardless of the AF treatment.
+Therefore, the alternating field demagnetization did not result in significant increase in specimen success rate for the anorthosites although it might have improved the specimen behavior for those anorthosites that were subjected to this treatment.
+Together with the narrow range of unblocking temperatures of their sister specimens \cite{Zhang2021b}, these data are consistent with the interpretation that the anorthosites with successful paleointensity results have a magnetic mineral assemblage that is likely dominated by single domain, low-Ti magnetite.
+Some anorthosite specimens and all diabase specimens still failed during result selections even though AF treatment was applied.
+It could be that they have a very high dominance of multidomain (MD) or vortex state grains with relatively moderate coercivity such that an alternating field with 20 mT peak field did not significantly mitigate their non-ideal paleointensity behaviors.
 Another cause for their failure could be the thermochemical alterations that happened during heating in air.
@@
 Overall, the consistent specimen-level behavior makes the successful results from the anorthosites of high confidence.
-The paleointensity quality index (Q$_{PI}$) for the anorthosite sites are all above 3 and the summary statistics are shown in Table S2.
+The paleointensity quality index (Q$_{PI}$) for the anorthosite sites are all above 4 and the summary statistics are shown in Table S2.
 
-In addition to applying paleointensity selection criteria to our experiment results, we applied the bias corrected estimation of paleointensity (BiCEP) method by Cych et al \cite{Cych2021a} to all of our anorthosite and diabase specimens regardless of their Arai plot statistics.
-The BiCEP method is a Bayesian statistical method based on an assumption that all specimen paleointensity results from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the Arai plot \cite{Arai1963a, Paterson2011a}.
+In addition to applying paleointensity selection criteria to our experiment results, we applied the bias corrected estimation of paleointensity (BiCEP) method by ref.
+\citealp{Cych2021a} to all of our anorthosite and diabase specimens regardless of their NRM/TRM plot statistics.
+The BiCEP method is a Bayesian statistical method based on an assumption that all specimen paleointensity results from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Arai1963a, Paterson2011a}.
 This method has the advantage of using as many specimens as available for estimating site-level paleointensities.
-The posterior distributions of paleointensity estimates given by the BiCEP method show that for the anorthosites that pass our paleointensity selection criteria, probabilities are very low for the case where these anorthosites are recording heavily biased paleointensity estimates from a true value below 10 $\mu$T (Fig.
+The posterior distributions of paleointensity estimates given by the BiCEP method address the high quality of the paleointensity estimates from the anorthosite xenoliths, in that the posterior probabilities are very low for the case where these anorthosites are providing high paleointensity values that are heavily biased from a true value below 10 $\mu$T (Fig.
 S3).
@@
 \subsection*{Rock magnetism}
-Rock magnetic data support that the successful anorthosite specimens have dominant magnetic carriers with magnetic properties similar to stochiometric, non-interactive, single domain magnetite, whereas those anorthosites that failed the paleointensity result selection and all diabase have more pronounced populations of non-ideal carriers.
+Rock magnetic data support that the successful anorthosite specimens have dominant magnetic remanence carriers with magnetic properties similar to stochiometric, non-interactive, single domain magnetite, whereas those anorthosites that failed the paleointensity result selection and all diabase have more pronounced populations of non-ideal carriers.
 Magnetic property measurement system (MPMS) data show sister specimens to paleointensity specimens typically have dominant magnetic mineralogy of magnetite as can be evidenced through the Verwey transitions (Fig.
-S1; \citealp{Verwey1939a}).
-However, specimens from sites that yield successful paleointensity results display Verwey transition near 120 Kelvin (K) as expected for stochiometric low-Ti magnetite \cite{Ozdemir1993a}.
-On the other hand, the diabase and anorthosite specimens that did not pass our paleointensity selection typically have suppressed Verwey transition toward lower temperatures (Fig.
+S1; ref.
+\citealp{Verwey1939a}).
+More, specimens from sites that yield successful paleointensity results have Verwey transition temperatures near 120 Kelvin (K) as expected for stochiometric low-Ti magnetite \cite{Ozdemir1993a}.
+However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have suppressed Verwey transition toward lower temperatures (Fig.
 S1), indicating that the specimens likely have a ferromagnetic mineralogy with variable degrees of oxidation \cite{Ozdemir1993a}.
@@
 \ref{fig:coercivity} based on our backfield experiments shows that specimens from anorthosites that pass paleointensity selections have a distinctly higher average MDF range than other anorthosites and all diabase.
-Example single component skew-normal distribution fits of coercivity spectra based on backfield experiments on rock chips sister to paleointensity specimens show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity distributions around 80 mT (Fig.
+Example single component skew-normal distribution fits of coercivity spectra \cite{Maxbauer2016a} based on backfield experiments on rock chips sister to paleointensity specimens show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity distributions around 80 mT (Fig.
 \ref{fig:coercivity}).
@@
 
-\subsection*{Remanence anisotropy and nonlinear TRM acquisition check}
+\subsection*{TRM anisotropy and nonlinear acquisition check}
 Significant remanence anisotropy has been documented to exist within anorthositic rocks that form in layered intrusive complexes \cite{Selkin2000a, Feinberg2006a}.
-\cite{Selkin2000a} demonstrated that strong remanence anisotropy associated with the igneous foliation developed within an anorthosite sample from the Stillwater Complex can lead to paleointensity estimates that are different by three times.
-Depending on the orientation of the applied field during in-field steps and that of the magnetic anisotropy axes of a specimen, such strong magnetic fabrics associated with petrofabrics can result in significantly biased paleointensity estimates.
-To assess whether our paleointensity estimates are biased by remanence anisotropy, we calculated the gamma statistic, which is the angular difference between the last pTRM step used for paleointensity determination and the applied field direction (Table \ref{tab:PINT_result}).
-The results show all anorthosite specimens used for paleointensity experiment have gamma values ranging from $\sim$0.9\textdegree-$\sim$11.9\textdegree, with a median value of 4.2\textdegree (Table \ref{tab:PINT_result}).
+Ref.
+\citealp{Selkin2000a} demonstrated that strong remanence anisotropy associated with the igneous foliation developed within an anorthosite sample from the Stillwater Complex can lead to significant overestimate or underestimate of paleointensity values depending on the relative orientations between the external applied field magnetic anisotropy axes of a specimen.
+To assess whether our paleointensity estimates are biased by remanence anisotropy, we calculated the gamma statistic, which is the angular difference between the last pTRM step of paleointensity experiment and the applied field direction.
+The results show that the anorthosite specimens used for paleointensity experiment have gamma values ranging from $\sim$0.9\textdegree to $\sim$11.9\textdegree, with a median value of 4.2\textdegree (Table S1).
 These gamma values are similar to those of the Midcontinent Rift volcanics \cite{Sprain2018a} and they show that the Beaver River anorthosite bulk rocks do not have a significant remanence anisotropy.
-This result can be further supported by the fact that the recovered paleodirectional data from our anorthosite xenoliths closely match those from the paired Beaver River diabase hosts \cite{Zhang2021b} and that no strong magnetic anisotropy was observed through magnetic imaging.
-\cite{Selkin2007a} also noted that vortex state magnetic grains and certain single domain particles can produce biased paleointensity estimates due to their acquisition of saturation remanence during laboratory in-field steps.
-According to the numerical modeling of \cite{Selkin2007a}, thermoremanent magnetization acquired by a randomly-oriented population of elongate single-domian magnetite particles are unlikely to saturate upon our applied field of 30 $\mu$T.
-Therefore, in this study we do not apply any anisotropy or nonlinear TRM correction to our specimens.
+Paleodirectional data from our anorthosite xenoliths further support that they have minimal remanence anisotropy as their site mean directions closely match those of the Beaver River diabase hosts \cite{Zhang2021b}.
+In addition, no strong magnetic fabric was observed through magnetic imaging (Fig.
+\ref{fig:Petro_QDM}).
+Ref.
+\citealp{Selkin2007a} also noted that vortex state magnetic grains and certain single domain particles can produce biased paleointensity estimates due to their acquisition of saturation remanence during laboratory in-field steps.
+According to the numerical modeling of ref.
+\citealp{Selkin2007a}, thermoremanent magnetization acquired by a population of randomly-oriented, elongate, single-domain magnetite particles are unlikely to saturate upon our applied field of 30 $\mu$T.
+Therefore, in this study we do not correct for anisotropy or nonlinear TRM to our specimens.
 
@@
 Paleointensity estimates from igneous intrusions can also be biased toward higher values due to slow cooling rates \cite{Halgedahl1980a}.
-Because the cooling rates of intrusive rocks could be much longer than the extrusive rocks, the slowly cooled Beaver River diabase and anorthosite xenoliths therein can bias laboratory paleointensity estimates toward an overestimate for single domain grains \cite{Dodson1980a, Halgedahl1980a, Selkin2000a}.
-
+It has been shown that large differences in cooling rates between acquisition of an NRM versus a laboratory pTRM can result in overestimated paleointensities for SD grains \cite{Dodson1980a, Halgedahl1980a}.
 From the thermal history model of ref.
-\citealp{Zhang2021b} we can also estimate the duration which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully block in their characteristic natural remanence component ($\sim$500\textdegree C; \cite{Zhang2021b}).
+\citealp{Zhang2021b} we can estimate the duration which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully block in their characteristic natural remanence component ($\sim$500\textdegree C; ref.
+\citealp{Zhang2021b}).
 We find the cooling time to be on the order of $\sim$1 kyr, which corresponds to a cooling rate of $1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
@@
 We therefore correct our paleointensity results by a factor of 0.74.
-The cooling rate corrected specimen paleointensity estimates together with specimen- and site-level weighted means are shown in Fig.
+The cooling rate-corrected specimen paleointensity estimates together with specimen- and site-level weighted means are shown in Fig.
 \ref{fig:PINT_cooling_corrected}.
-Taking all accepted anorthosite paleointensity results, they have a site mean absolute paleointensity estimates of 47.90 $\pm$ 9.32 $\mu$T (Fig.
+Taking all accepted anorthosite paleointensity results, they give a site mean absolute paleointensity estimate of 47.90 $\pm$ 9.32 $\mu$T (Fig.
 \ref{fig:PINT_cooling_corrected}).
 
-\section*{Late Mesoproterozoic geodynamo }
-To best characterize the geomagnetic axial dipole field intensity during a certain time period requires that a paleointensity dataset to cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged and the geocentric axial dipole hypothesis (GAD) can be assumed.
+\section*{Proterozoic geodynamo}
+To best characterize the geomagnetic axial dipole field intensity during a certain time period requires that a paleointensity dataset to cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged and the geocentric axial dipole hypothesis can be assumed.
 Thermal modeling results from ref.
@@
 \ref{fig:Geologic_map}).
-The consistency within small regions but variations between regions could be the result of differential cooling history due to heterogeneity of the timing and depths of the Beaver River diabase intrusions, causing the anorthosites to capture paleosecular variations \textit{ca.} 1092 Ma.
+The consistency within small regions but variations between regions is consistent with the interpretation that differential cooling history due to heterogeneity of the timing and depths of the Beaver River diabase intrusions caused the anorthosites to capture paleosecular variations with their magnetic remanence \textit{ca.} 1092 Ma.
 Overall, none of the anorthosite xenoliths record a site mean VDM value lower than 40 ZAm$^2$ (Fig.
@@
 
-Paleomagnetic observational records near the Ediacaran-Cambrian boundary have been interpreted to be consistent with a trend of decreasing axial dipole strength toward a very low value \cite{Lloyd2021a, Bono2019a}, followed by a rapid recovery (a few tens of million years) toward higher values soon after \cite{Thallner2021a}.
-Such a decrease-increase transition can be consistent with the dynamo numerical simulations which adopt a rather high thermal conductivity of the core and projects an inner core birth near this time \cite{Gomi2013a, Davies2015a, Labrosse2015a, Ohta2016a}.
-In particular, some models have suggested that the kick-start of rigorous compositional convection between the inner core and the outer core soon after the onset of inner core nucleation can result in almost an order of magnitude increase in Earth's axial dipole moment \cite[e.g.]{Davies2021a}.
-The enriching observational records from the Ediacaran and Cambrian have been suggested to indicate that a geodynamo driven solely by thermal convection was waning to its brink of collapse near the end of Ediacaran \cite{Bono2019a, Thallner2021b} and was recovering toward modern-day strengths soon after \cite{Thallner2021a}.
+The new paleointensity records from the Midcontinent Rift anorthosite xenoliths together with previous results from other Mesoproterozoic rocks suggest that a dramatic six-fold increase in geomagnetic dipole moment happened within $\sim$50 Myr during the late Mesoproterozoic (Fig.
+\ref{fig:PINT_compilation}).
+The scale of the increase is similar to the increasing trend observed near the Ediacaran-Cambrian which has been suggested to evidence the onset of Earth's inner core nucleation \cite{Thallner2021a}.
+The \textit{ca.} 1140 Ma Abitibi dikes \cite{Macouin2003a} record an average virtual dipole moment of $\sim$13 ZAm$^2$, similar to the low values of $\sim$7 ZAm$^2$ obtained from the single silicate crystals of the \textit{ca.} 565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a}.
+However, the relatively high moments of $\sim$72 ZAm$^2$ from the \textit{ca.} 1105 Ma Osler Volcanic Group \cite{Sprain2018a} and $\sim$89 ZAm$^2$ from the \textit{ca.} 1092 Ma Beaver River anorthosite xenoliths (Fig.
+\ref{fig:PINT_compilation}) are about six times the values from the Abitibi dikes.
 
-However, the updated paleointensity records from the Midcontinent Rift together with previous results from other Mesoproterozoic rocks suggest that the a dramatic increase in geomagnetic dipole moment similar in scale to that modeled for after the onset of inner core nucleation may not be unique to the Ediacaran-Cambrian boundary (Fig.
+The high geomagnetic field estimates from the Midcontinent Rift volcanics and intrusives are not the only suite of records at odds with the hypothesis of a progressive decay trend during the entire Proterozoic Eon \cite{Biggin2009a, Bono2019a}.
+The $\sim$60 ZAm$^2$ dipole moment estimates from the \textit{ca.} 755 Ma Mundine Well dikes are discordant with the long-term decrease trend \cite{Bono2019a} and are about three times the estimates from the \textit{ca.} 720 Ma dolerite dikes of the Franklin Large Igneous Province \cite{Lloyd2021a}.
+Similarly, an apparent rapid decrease in geomagnetic dipole moment from \textit{ca.} 1092 Ma to \textit{ca.} 1070 Ma can be evidenced by the contrasting high values from the Midcontinent Rift anorthosite xenoliths and the low values from the Bangemall Sills (Fig.
+\ref{fig:PINT_compilation}; ref.
+\citealp{Lloyd2021b}).
+These data suggest that an apparent transition of virtual dipole moment similar in scale with what is predicted for a dynamo powered with or without compositional convection in the core may not be unique during the Precambrian.
+In addition, given that a five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years, when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed, we cannot conclude on whether an apparent rise or decay within a narrow time period ($\sim$35 Ma) during the Proterozoic is associated with transitions in the core phase transition or is a result of short-term excursion of the geomagnetic field that may be common through much of the evolution of Earth's core.
+
+Regardless of whether the Abitibi dikes recorded low virtual dipole moments of a geodynamo with waning strength or an excursion at \textit{ca.} 1140 Ma, the high values from the Midcontinent Rift intrusions and extrusive volcanics are inconsistent with the prediction of there being a progressive decay at the time (Fig.
+\ref{fig:PINT_compilation}; ref.
+\citealp{Bono2019a}) and necessitates that there being a vigorous late Mesoproterozoic geodynamo.
+
+% Paleomagnetic observational records near the Ediacaran-Cambrian boundary have been interpreted to be consistent with a trend of decreasing axial dipole strength toward a very low value \cite{Lloyd2021a, Bono2019a}, followed by a rapid recovery (a few tens of million years) toward higher values soon after \cite{Thallner2021a}.
+Such a decrease-increase transition can be consistent with dynamo numerical simulations which adopt a rather high thermal conductivity of the core and projects an inner core birth near this time \cite{Gomi2013a, Davies2015a, Labrosse2015a, Ohta2016a}.
+In particular, some models have suggested that the kick-start of rigorous compositional convection between the inner core and the outer core due to the onset of inner core nucleation can result in almost an order of magnitude increase in Earth's axial dipole moment \cite[e.g.][]{Davies2021a}.
+The enriching observational records from the Ediacaran and Cambrian have been suggested to indicate that a geodynamo driven solely by thermal convection was waning to its brink of collapse near the end of Ediacaran \cite{Bono2019a, Thallner2021b} and was recovering toward modern-day strengths in the Cambrian \cite{Thallner2021a}.
+
+% However, the updated paleointensity records from the Midcontinent Rift together with previous results from other Mesoproterozoic rocks suggest that the a dramatic increase in geomagnetic dipole moment similar in scale to that modeled for after the onset of inner core nucleation may not be unique to the Ediacaran-Cambrian boundary (Fig.
 \ref{fig:PINT_compilation}).
@@
 However, the relatively high moments of $\sim$72 ZAm$^2$ from the \textit{ca.} 1105 Ma Osler Volcanic Group \cite{Sprain2018a} and $\sim$89 ZAm$^2$ from the \textit{ca.} 1092 Ma Beaver River anorthosite xenoliths (Fig.
-\ref{fig:PINT_compilation} are about six times the values from the Abitibi dikes.
+\ref{fig:PINT_compilation}) are about six times the values from the Abitibi dikes.
 This increase in absolute dipole moment estimates during $\sim$50 Myr recorded by the Midcontinent Rift rocks are much higher than that in the Ediacaran (Fig.
@@
 
-The high geomagnetic field estimates from the Midcontinent Rift volcanics and intrusives are not the only suite of rocks that bear records at odds with the hypothesis of a progressive decay trend during the entire Proterozoic Eon \cite{Biggin2009a, Bono2019a}.
-The $\sim$60 ZAm$^2$ dipole moment estimate from the \textit{ca.} 755 Ma Mundine Well dikes are discordant with the long-term decrease trend \cite{Bono2019a} and is about three times the estimates from the \textit{ca.} 720 Ma dolerite dikes of the Franklin Large Igneous Province \cite{Lloyd2021a}.
-Similarly, an apparent rapid decrease in geomagnetic dipole moment from \textit{ca.} 1092 Ma to \textit{ca.} 1070 Ma can be evidenced by the contrasting high values from the Midcontinent Rift anorthosite xenoliths and the low values from the Mundine Well dikes (Fig.
-\ref{fig:PINT_compilation}).
-These data suggest that an observation of a strong increase or decrease in geomagnetic dipole moment may not be unique during the Precambrian.
-A sparse record in a short time period may not sufficiently average out paleosecular variations and more spatial and temporal coverage of the observational records is needed for further analyses of the long-term evolution of the geodynamo.
+% observation of an increase or a decrease in virtual dipole moment may not be unique during the Precambrian.
+A sparse spatial and temporal coverage of the Precambrian paleointensity records may not be sufficient for further analyses of the long-term evolution of the geodynamo.
 Given that a five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years (5 and 95 percentiles), when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed, we cannot conclude on whether an apparent rise or decay within a narrow time period ($\sim$35 Ma) during the Proterozoic is associated with transitions in the core fluid regime or is a result of short-term variations of the geomagnetic field that may be common through much of the evolution of Earth's core.
@@
 Sun compass orientations were preferentially used for determining the sample azimuth.
-Specimens from a total of 7 diabase and 14 anorthosite sites were used for paleointensity experiments (Table \ref{tab:Pmag_site_data}).
+Specimens from a total of 7 S1).
 Sister specimens underwent step-wise alternating field (AF) or thermal demagnetization at the UC Berkeley Paleomagnetism Lab and the results are presented in \cite{Zhang2021b}.
@@
 A total of 86 specimens from 14 anorthosite xenoliths and a total of 69 specimens from 7 diabase sites underwent paleointensity experiments that followed the step-wise double-heating Thellier method \cite{Thellier1959a} using the IZZI protocol \cite{Tauxe2004a} with heating steps up to 585 \textdegree C.
-Partial thermal remanent magnetization (pTRM) checks were performed systematically throughout the experiment to test whether there was significant mineralogical alteration due to heating and were assessed using the SCAT parameter of \cite{Shaar2013a}.
+Partial thermal remanent magnetization (pTRM) checks were performed systematically throughout the experiment to test whether there was significant mineralogical alteration due to heating and were assessed using the SCAT parameter of ref.
+\citealp{Shaar2013a}.
 On top of the IZZI modification of the Thellier experimental protocol, we also performed a comparative study where we added an extra step of 20 mT alternating field (AF) cleaning on some of the specimens after each in-field step.
@@
 \subsection*{Paleointensity result selection}
-The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD; \cite{Kirschvink1980a}) of $<$20\textdegree; (2) scatter parameter ($\beta$; \cite{Coe1978a}) values of $<$15 per cent; (3) a deviation angle (DANG; \cite{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC; \cite{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT; \cite{Shaar2013a}) = TRUE; (6) a maximum magnetic moment difference between adjacent zero-field steps (GAP-Max; \cite{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
+The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD; ref.
+\citealp{Kirschvink1980a}) of $<$20\textdegree; (2) scatter parameter ($\beta$; ref.
+\citealp{Coe1978a}) values of $<$15 per cent; (3) a deviation angle (DANG; ref.
+\citealp{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC; ref.
+\cite{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT; ref.
+\citealp{Shaar2013a}) = TRUE; (6) a maximum magnetic moment difference between adjacent zero-field steps (GAP-Max; ref.
+\citealp{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
 The criteria are also summarized in Table \ref{tab:criteria}.
@@
 Both MAD and DANG assess the directional variation of the NRM, with MAD measuring the scatter in the NRM directions and DANG assessing whether the component is trending toward the origin of the Zijderveld plot.
-$\beta$ is the ``scatter" parameter of \cite{Coe1978a} and is the ratio of the standard error of the slope of the best-fit line of the selected NRM and pTRM points on an Arai plot to the absolute value of the slope.
+$\beta$ is the ``scatter" parameter of ref.
+\cite{Coe1978a} and is the ratio of the standard error of the slope of the best-fit line of the selected NRM and pTRM points on an NRM/TRM plot to the absolute value of the slope.
 FRAC is the fraction of the NRM that is used in the best-fit line \cite{Shaar2013a}.
-The FRAC value was chosen to preferentially select samples with dominantly single-slope Arai plots.
-GAP-Max is the maximum gap between two points on the Arai plot determined by vector arithmetic.
-SCAT is a Boolean operator which uses the error on the best-fit slope of the selected data on the Arai plot to determine if the data are overly scattered.
+The FRAC value was chosen to preferentially select samples with dominantly single-slope NRM/TRM plots.
+GAP-Max is the maximum gap between two points on the NRM/TRM plot determined by vector arithmetic.
+SCAT is a Boolean operator which uses the error on the best-fit slope of the selected data on the NRM/TRM plot to determine if the data are overly scattered.
 The parameter is used to assess pTRM checks in addition to assessing the degree to which IZZI steps are zigzagged.
-$\beta$, FRAC, GAP-Max and SCAT are all statistics to assess the behavior of Arai plots.
-See the Standard Paleointensity Definitions (\cite{Paterson2014a}; \url{https://earthref.org/PmagPy/SPD/home.html}) for more details.
+$\beta$, FRAC, GAP-Max and SCAT are all statistics to assess the behavior of NRM/TRM plots.
+See the Standard Paleointensity Definitions (ref.
+\citealp{Paterson2014a}; \url{https://earthref.org/PmagPy/SPD/home.html}) for more details.
 Data analysis was conducted using Thellier GUI \cite{Shaar2013a} within the PmagPy software package \cite{Tauxe2016a}.
 
-\begin{table*}
-\caption{Quality criteria used for selecting paleointensity data. See the text for more detail.} 
-\centering
-\begin{tabular}{ccccccccc}
-\hline
-MAD (\textdegree) & Beta (per cent) & DANG (\textdegree) & FRAC& SCAT & GAP-MAX & NpTRM & N Arai\\
-\hline
-20 & 15 & 5 & 0.6 & TRUE & 0.25 & 2 & 4 \\
-\hline
+% \begin{table*}
+% \caption{Quality criteria used for selecting paleointensity data. See the text for more detail.} 
+% \centering
+% \begin{tabular}{ccccccccc}
+% \hline
+% MAD (\textdegree) & Beta (per cent) & DANG (\textdegree) & FRAC& SCAT & GAP-MAX & NpTRM & N Arai\\
+% \hline
+% 20 & 15 & 5 & 0.6 & TRUE & 0.25 & 2 & 4 \\
+% \hline
 
-\end{tabular}
-\label{tab:criteria}
-\end{table*}
+% \end{tabular}
+% \label{tab:criteria}
+% \end{table*}
 
@@
 Specimen median destructive field (MDF) are calculated based on the backfield curves.
-The calculated coercivity spectra were subsequently decomposed into one ore more components using skew-normal distributions following the method of \cite{Maxbauer2016a}.
+The calculated coercivity spectra were subsequently decomposed into one ore more components using skew-normal distributions following the method of ref.
+\citealp{Maxbauer2016a}.
 The fitted components are then used to interpret the characteristics of different populations of magnetic particles within specimens.
@@
 
-To further identify the magnetic carriers within the Beaver River anorthosite xenoliths and compare them with the anorthosites of the Duluth Complex Anorthositic Series rocks, we used the quantum diamond microscope (QDM) at the UC Berkeley Paleomagnetism laboratory to image a thin section of sample MS99033 which is a hand sample from anorthosite xenolith AX16 (which yielded a $^{206}$Pb/$^{238}$U zircon date of 1091.83 $\pm$ 0.21 Ma; \cite{Zhang2021b}), and a thin section of a Duluth Complex anorthosite (Fig.
+To further identify the magnetic carriers within the Beaver River anorthosite xenoliths and compare them with the anorthosites of the Duluth Complex Anorthositic Series rocks, we used the quantum diamond microscope (QDM) at the UC Berkeley Paleomagnetism laboratory to image a thin section of sample MS99033 from anorthosite xenolith AX16 (which yielded a $^{206}$Pb/$^{238}$U zircon date of 1091.83 $\pm$ 0.21 Ma; ref.
+\citealp{Zhang2021b}), and a thin section of a Duluth Complex anorthosite (Fig.
 \ref{fig:Petro_QDM}).
-We the QDM to image the magnetic field over the polished thin section surfaces at a sample-sensor distance of 5 $\mu$m in projective magnetic microscopy (PMM) mode with a spatial resolution of 4.7 μm per pixel and an instantaneous 0.9 mT bias field that is canceled during the course of measurement \cite{Glenn2017a}).
+We use the QDM to image the magnetic field over the polished thin section surfaces at a sample-sensor distance of 5 $\mu$m in projective magnetic microscopy (PMM) mode with a spatial resolution of 4.7 μm per pixel and an instantaneous 0.9 mT bias field that is canceled during the course of measurement \cite{Glenn2017a}).
 }

### 2021-12-18 PNAS_manuscript/PNAS_manuscript.tex
combine figure, code and text updates

@@
 \author[a,1]{Yiming Zhang}
-\author[a,1]{Nicholas L. Swanson-Hysell} 
+\author[a]{Nicholas L. Swanson-Hysell} 
 \author[a,b]{Margaret S. Avery}
@@
 \significancestatement{Acquiring high-fidelity observational records of ancient magnetic field intensity from remanent magnetization of rocks is crucial for constraining the evolution history of Earth's core.
-However, such records are often difficult to recover due to the alterations associated with weathering after the formation of rocks and heating during experiments.
-In this study, we use plagioclase cumulates from the North American Midcontinent Rift that has experienced minimal post-formation alterations to obtain a present-day-like virtual dipole moment 1.1 billion years ago.
+However, robust estimates of ancient field strengths are often difficult to recover due to alterations or non-ideal rock magnetic behaviors.
+In this study, we use plagioclase cumulates that formed in the deeper crust and were brought to the near surface by magma where they cooled and acquired thermal remanent magnetizations.
+These plagioclase from the North American Midcontinent Rift have experienced minimal post-formation alterations and record a present-day-like virtual dipole moment 1.1 billion years ago.
 The results indicate the existence of strong power sources for Earth's dynamo during the late Mesoproterozoic.
@@
 designed research; Y.Z., N.L.S.-H.
-and M.S.A studied and sampled the rock cores; Y.Z.
+and M.S.A conducted fieldwork and sampling; Y.Z.
 and N.L.S.-H.
@@
 performed petrography and electron microscopy; R.R.F.
-contributed to magnetic imaging data analyses; and Y.Z., N.L.S.-H., M.S.A, and R.R.F.
+contributed to magnetic imaging data analyses; Y.Z.
+and N.L.S.-H.
 wrote the paper.}
@@
 Obtaining estimates of Earth's magnetic field strength in deep time is complicated by non-ideal rock magnetic behavior in many igneous rocks.
-In this study, we target \textit{ca.} 1092 Ma anorthosite xenoliths that cooled and acquired their magnetization within the Beaver River diabase intrusions of the North American Midcontinent Rift.
-In contrast to the diabase which has non-ideal paleointensity behaviors leading to its failure in providing reliable paleointensity estimates, the anorthosite xenoliths often yield high-quality, single-slope paleointensity results that are consistent at specimen and site levels.
+In this study, we target \textit{ca.} 1092 Ma anorthosite xenoliths that cooled and acquired their magnetization within Beaver River diabase intrusions of the North American Midcontinent Rift.
+In contrast to the diabase which has non-ideal paleointensity behaviors and fails to provide reliable paleointensity estimates, the anorthosite xenoliths are unusually high fidelity recorders yielding high-quality, single-slope paleointensity results that are consistent at specimen and site levels.
 An average value of $\sim$89 ZAm$^2$ for the virtual dipole moment from the anorthosites is higher than that of the dipole component of Earth's magnetic field today.
-Such a strong late Mesoproterozoic dipole recorded by the anorthosites require the existence of a vigorous geodynamo at the time.
-Together with previous paleointensity data from other Midcontinent Rift rocks, our new results suggest that a dynamo with strong power sources likely persisted through $\sim$25 Myr at \textit{ca.} 1.1 Ga.
-These observations are inconsistent with there being a progressive decay of Earth's dynamo strength through the Proterozoic Eon.
+Such a strong late Mesoproterozoic dipole recorded by the anorthosites requires the existence of a vigorous geodynamo at the time.
+Together with previous paleointensity data from other Midcontinent Rift rocks, our new results indicate that a dynamo with strong power sources persisted through $\sim$25 Myr at \textit{ca.} 1.1 Ga.
+These observations are inconsistent with there being a progressive monotonic decay of Earth's dynamo strength through the Proterozoic Eon.
 \end{abstract}
@@
 At present day, the geodynamo is collectively powered by heat flow across the core-mantle boundary (CMB), and from the crystallization of the solid inner core from the liquid outer core which provides latent heat and compositional buoyancy.
-However, while paleomagnetic studies have found that a dynamo field has existed since at least 3.4 billion years ago in the Archean \cite{Selkin2007a, Tarduno2014a, Brenner2020a}, Earth's inner core likely crystallized more recently.
+However, while paleomagnetic studies have found that a dynamo field has existed since at least 3.4 billion years ago \cite{Selkin2007a, Tarduno2014a, Brenner2020a}, Earth's inner core likely crystallized more recently.
 Estimates of the timing of the initial crystallization of Earth's inner core are interconnected with estimates for the core's thermal conductivity.
 Higher conductivity values imply faster cooling rates, which in turn imply that the threshold for the freezing of the inner core happened more recently \cite{Davies2015a}.
-While some estimates of these values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates based on first principles calculations implied higher thermal conductivity values and an age for the inner core that is less than 1.3 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b}, with some suggesting even younger ages ($<$ 700 Ma; ref.s \citealp{Labrosse2015a, Ohta2016a}.
-With the estimates for the core's thermal conductivity continue to be debated, it is crucial to use observational records as an independent constraint on the thermal evolution of Earth's core and mantle.
+While some estimates of these values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates have implied higher thermal conductivity values and an age for the inner core that is less than 1.3 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b}, with some suggesting even younger ages ($<$ 700 Ma; ref.
+\citealp{Labrosse2015a, Ohta2016a}).
+Given that estimates for the core's thermal conductivity continue to be debated, it is crucial to use observational records as an independent constraint on the thermal evolution of Earth's core and mantle.
 
@@
 
-Paleomagnetic records from ancient rocks are one of the few types of observational data that have the potential to record the evolution of Earth’s core and can be used to test the predictions from numerical simulations.
-Interpretations of selected records of the current Precambrian absolute paleointensity database (compiled in the PINT database; \url{http://earth.liv.ac.uk/pint/}) suggest that an appreciable geomagnetic field existed in the Archean but reached its weakest in the Ediacaran when very low virtual dipole moment (VDM) estimates have been obtained \cite{Bono2019a, Shcherbakova2019a, Thallner2021b, Lloyd2021a}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
+Paleomagnetic records from ancient rocks are one of the few types of observational data that have the potential to record the thermal evolution of Earth’s core.
+Evidence for a persistent magnetic field through the Proterozoic, for example, necessitates the existence of plate tectonics that sustained core-mantle boundary heat flow.
+However, strikingly low estimates of geomagnetic field strengths have been obtained in the Ediacaran Period (ref.
+\citealp{Bono2019a}; \textit{ca.} 565 Ma), potentially indicating an unusual period of geodynamo regime at the time.
+
+% Those paleointensity data has been interpreted to be consistent with thermal evolution models that use high core thermal conductivity values, where the nucleation of the inner core did not begin to provide compositional convection power for driving the geodynamo until after \textit{ca.} 565 Ma, suggesting that the Proterozoic dynamo might have experienced a monotonic decay when compositional convection was lacking and thermal convection was progressively decreasing \cite{Driscoll2016a, Davies2021a}.
+
+Determinations of ancient magnetic field strengths rely on igneous rocks that acquire thermal remanent magnetizations as they cool in Earth's magnetic field.
+These magnetizations need to be unmodified by subsequent heating or chemical alterations in order to be able to faithfully record ancient geomagnetic field record through time.
+Intraplate magmatic events are therefore an important target as they can be well-preserved within continental interiors at distance from mountain-building events along continental margins.
+However, magmatic events associated with large igneous provinces are typically of geologically short duration with the bulk of magmatic products emplaced within $<$1 Myr.
+The Midcontinent Rift is an exception as magmatism lasted $\sim$25 Myr from 1109 Ma to 1084 Ma.
+Additionally, extension ceased in the Midcontinent Rift prior to lithosphere separation, preserving volcanics, intrusives, and sediments of the rift within the continental interior.
+As a result, rocks of the rift have simple paleomagnetic behavior and form a central record of Mesoproterozoic paleogeography \cite{Swanson-Hysell2021c}.
+This simple magnetization and the long duration of magnetization holds the potential for a well-preserved extended window into the intensity of Earth's magnetic field from 1109 to 1084 Ma.
+
+Despite the exquisite preservation of the rocks, non-ideal paleointensity behaviors and overall low success rates have often led to high failure rates and challenged the interpretation of many previous paleointensity results from the Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a}.
+One typical failure of paleointensity experiment is the double-slope behavior, as two distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
+Ref \citealp{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas ref.
+\citealp{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
+Results with such non-ideal behavior were rejected by ref.
+\citealp{Sprain2018a} who applied stricter paleointensity selection criteria.
+
+In this study, we target a unique rock type - high-purity anorthosite xenoliths.
+These anorthosite xenoliths were brought to the near surface with magma that formed sills of the Beaver River diabase.
+They cooled and acquired their magnetizations in conjunction with the host diabase.
+Paleointensity experiments on the anorthosite xenoliths have a very high success rate, yielding consistent specimen- and site-level paleointensity results.
+Rock magnetic data reveal that the anorthosite xenoliths have low anisotropy of thermal remanent magnetization (TRM) and can acquire TRM linearly.
+Magnetic imaging shows the anorthosite specimens have dominant magnetic carriers interstitial to and within plagioclase crystals and they do not display strong preferred orientations.
+Our single-slope, high quality paleointensity data add to the reliable paleointensity records during the $\sim$25 Myr of Midcontinent Rift magmatic activity.
+Our new data from the Midcontinent Rift intrusive rocks support previous data from the volcanics that there was a strong late Mesoproterozoic geodynamo.
+
+% Interpretations of selected records of the current Precambrian absolute paleointensity database (compiled in the PINT database; \url{http://earth.liv.ac.uk/pint/}) suggest that an appreciable geomagnetic field existed in the Archean but reached its weakest in the Ediacaran when very low virtual dipole moment (VDM) estimates have been obtained \cite{Bono2019a, Shcherbakova2019a, Thallner2021b, Lloyd2021a}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
 Ref.
-\citealp{Bono2019a} adopted a billion-year-scale view of such observations to suggest that these data are consistent with thermal evolution models that use high core thermal conductivity values, where the nucleation of the inner core did not begin to provide compositional convection power for driving the geodynamo until the Ediacaran-Cambrian, suggesting that the Proterozoic dynamo might have experienced a progressive decay when compositional convection was lacking and thermal convection was decreasing \cite{Driscoll2016a, Davies2021a}.
+\citealp{Bono2019a} adopted a billion-year-scale view of such observations to suggest that these data are consistent with thermal evolution models that use high core thermal conductivity values, where the nucleation of the inner core did not begin to provide compositional convection power for driving the geodynamo until the Ediacaran-Cambrian, suggesting that the Proterozoic dynamo might have experienced a monotonic decay when compositional convection was lacking and thermal convection was progressively decreasing \cite{Driscoll2016a, Davies2021a}.
 
@@
 
-However, previous paleointensity data from intrusive and extrusive rocks in the \textit{ca.} 1.1 Ga Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record VDM values similar to that of today, which are higher than the values predicted by the long term decay curve proposed by ref.
+% However, previous paleointensity data from intrusive and extrusive rocks in the \textit{ca.} 1.1 Ga Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record VDM values similar to that of today, which are higher than the values predicted by the long term decay curve proposed by ref.
 \citealp{Bono2019a} (Fig.
@@
 
-Although it is typical of paleointensity experiments to have low success rates, bulk plagioclase-rich rocks \cite{Selkin2000a} and single plagioclase crystals \cite{Tarduno2005a} are found to be able to produce high quality paleointensity results.
+% Although it is typical of paleointensity experiments to have low success rates, bulk plagioclase-rich rocks \cite{Selkin2000a} and single plagioclase crystals \cite{Tarduno2005a} are found to be able to produce high quality paleointensity results.
 It has been suggested that plagioclase crystals can protect magnetic minerals that exsolved within them against post-formation weathering and thermochemical alteration during paleointensity experiments \cite{Tarduno2005a}.
@@
 
-The Mesoproterozoic North American Midcontinent Rift is an intracontinental rift that did not result in the break-up of Laurentia.
+While magmatic activity within the Midcontinent Rift was protracted, there were volcanism and intrusions characterized by particularly rapid and voluminous emplacement.
+The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex in northern Minnesota punctuates one such period of magmatism during the Midcontinent Rift main stage activity.
+The magma that formed the 1091.7 $\pm$ 0.2 Ma Beaver River diabase dikes and sills of the Beaver Bay Complex transported numerous anorthosite xenoliths that have short-axis diameters up to 180 meters via wide conduits \cite{Boerboom2004a, Boerboom2006b}.
+The anorthosite xenoliths are plagioclase cumulates that formed comagmatically with the host diabase in the lower crust - an interpretation confirmed by U-Pb zircon geochronology \cite{Zhang2021b}.
+
+% The Mesoproterozoic North American Midcontinent Rift is an intracontinental rift that did not result in the break-up of Laurentia.
 Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; refs.
@@
 
-Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths acquired thermal remanent magnetizations during cooling within the Beaver River diabase \cite{Zhang2021b}.
-Step-wise thermal demagnetizations show the anorthosite xenoliths have minimal secondary remanence and single-component characteristic remanence magnetizations that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C.
-Geochronology data and rare earth element diffusion modeling further suggest that the anorthosite xenoliths are cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
-Petrological textures and geochemical attributes imply that the Beaver River anorthosite xenoliths formed in a lower crustal magmatic conditionm \cite{Zhang2021b, Miller1997a} distinct from that of layered mafic intrusions, such as the Duluth Complex which is $\sim$4 Myr older and generated massive anorthositic cumulates that led to the formation of the Anorthositic Series \cite{Swanson-Hysell2021a, Miller1990a}.
-The dominantly monomineralic anorthosite xenoliths of the Beaver River diabase have an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}, higher than the average of $\sim$60\% of the Anorthositic Series rocks of the Duluth Complex \cite{Miller1990a}.
+Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths to have acquired thermal remanent magnetizations while cooing with the Beaver River diabase \cite{Zhang2021b}.
+Step-wise thermal demagnetizations show the anorthosite xenoliths have dominantly single-component magnetization that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C.
+
+\section*{Petrography and magnetic imaging of anorthosite xenoliths}
+
+Petrographic imaging shows the dominantly monomineralic anorthosite xenoliths of the Beaver River diabase often develop granoblastic texture characterized by equigranular crystals with poor development of petrofabrics (Fig.
+\ref{fig:Petro_QDM}A).
+In addition, the plagioclase crystals of the anorthosite xenoliths are often free of large inclusions.
+Magnetic imaging using the quantum diamond microscope (QDM) shows that the dominant magnetic remanence carriers often exist within and sometimes interstitial to plagioclase crystals (Fig.
+\ref{fig:Petro_QDM}E).
+
+These features of the Beaver River anorthosite xenoliths are distinct from other anorthosites such as the anorthosite xenoliths within the Duluth Complex Anorthositic Series rocks.
+The plagioclase of the Duluth Complex anorthosites typically develop interlocking textures that display strong igneous foliations (Fig.
+\ref{fig:Petro_QDM}B) and associated magnetic mineral fabrics characterized by having aligned titanomagnetite needles within plagioclase crystals (Fig.
+\ref{fig:Petro_QDM}D).
+Magnetic imaging confirms that the titanomagnetite needles within the plagioclase crystals are often strongly magnetic with magnetic fabrics oriented along the long axes of the needles (Fig.
+\ref{fig:Petro_QDM}F).
+Titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex plagioclase \cite{Miller1990a}.
+
+The lack of symplecites and exsolved titanomagnetite needles within the Beaver River anorthosite xenoliths reflects that the Beaver River diabase magma had a different oxidation history than that of the Duluth Complex.
+The plagioclase and oxide mineral textures of the Beaver River anorthosites also distinguish them from plagioclase cumulates of other layered mafic complex where magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
+
+% The dominantly monomineralic anorthosite xenoliths of the Beaver River diabase have an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}, higher than the average of $\sim$60\% of the Anorthositic Series rocks of the Duluth Complex \cite{Miller1990a}.
 The plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture showing igneous foliations (Fig.
@@
 \citealp{Pesonen1983a} are categorized by lithology.
-All sites from refs.
+All sites from ref.
 \citealp{Pesonen1983a, Kulakov2013a, Sprain2018a} are color-coded by their ages.}}
@@
 \noindent\includegraphics[width=11.4 cm]{Petro_QDM.png}
-\caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from Duluth Complex Anorthositic Series (left column) and Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (right column; ref.
-\citealp{Zhang2021b}).
-Petrographic image (A) shows Duluth Complex anorthosite have euhedral, interlocking plagioclase crystals showing igneous foliation.
-(B) Beaver River anorthosite have plagioclase with a granoblastic texture.
-(C) Duluth Complex plagioclase can contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the (010) planes of plagioclase.
-In contrast, plagioclase of sample MS99033 do not enclose large oxide needles.
-Magnetic field maps acquired by the quantum diamond microscope (QDM) at UC Berkeley Paleomagnetism Lab show that the oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the complex overlapping magnetic fields exerted above the oxides in (E).
-In comparison, relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite.
-The dipole-like features in (F) do not show strong preferred orientations.
+\caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from the Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (left column; ref.
+\citealp{Zhang2021b}) and and distinct anorthosite from the Duluth Complex Anorthositic Series (right column).
+Cross-polarized petrographic images (A) Beaver River anorthosite have plagioclase with a granoblastic texture; (B) Plagioclase of Duluth Complex anorthosite have euhedral, interlocking crystals showing igneous foliation; (C) zoomed-in image of plagioclase from MS99033 showing the crystals are largely free of large opaque inclusions; (D) Duluth Complex plagioclase can contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the (010) planes of plagioclase.
+Magnetic field maps acquired by the quantum diamond microscope (QDM) show that the relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite in (E) while large oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the complex overlapping magnetic fields above the oxides in (F).
 Magnetic field maps of (E) and (F) share the same spatial and magnetic scale.
@@
 
-40 from a total of 86 anorthosite specimens and 0 out of a total of 69 diabase specimens passed our paleointensity selection criteria, resulting in 7 anorthosite sites and no diabase sites with site-level paleointensity results that pass paleointensity selections.
+Following IZZI-modified paleointensity experiments, 40 from a total of 86 anorthosite specimens and 0 out of a total of 69 diabase specimens passed our paleointensity result selection criteria of having (1) a maximum angular deviation (MAD; ref.
+\citealp{Kirschvink1980a}) of $<$20\textdegree; (2) scatter parameter ($\beta$; ref.
+\citealp{Coe1978a}) values of $<$15 per cent; (3) a deviation angle (DANG; ref.
+\citealp{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC; ref.
+\cite{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT; ref.
+\citealp{Shaar2013a}) = TRUE; (6) a maximum magnetic moment difference between adjacent zero-field steps (GAP-Max; ref.
+\citealp{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
+7 anorthosite sites and no diabase sites with site-level paleointensity results pass our selections.
 Example NRM/TRM plots are shown in Fig.
 \ref{fig:IZZI_examples}.
-Summary specimen absolute paleointensity estimates and site-level weighted mean paleointensity values are shown in Table S1 and are plotted in Fig.
-\ref{fig:PINT_cooling_corrected}.
-All measurement-level paleointensity experiment data can be accessed through the MagIC database (\url{}; UPDATE WHEN DOI IS GENERATED).
+The cooling rate corrected absolute paleointensity estimates from all anorthosite sites have a weighted mean of 52.72 $\pm $5.72 $\mu$T.
+This corresponds to a calculated virtual dipole moment of $\sim$89 ZAm$^2$ \textit{ca.} 1092 Ma.
+Summary specimen absolute paleointensity estimates and site-level weighted mean paleointensity values are plotted in Fig.
+\ref{fig:PINT_cooling_corrected} (and provided in Table S1).
+All measurement-level paleointensity experiment data is available within the MagIC database (\url{}; UPDATE WHEN DOI IS GENERATED).
 
@@
 
-The Beaver River anorthosite specimens yielded a high pass rate of $\sim$46.5\% through our selection criteria.
-In contrast, the diabase specimens are all rejected.
-Paleointensity experiment results from both lithologies show three types of dominant specimen behaviors on NRM/TRM plots (Fig.
-\ref{fig:IZZI_examples}): dominantly single-slope, double-slope, and zigzagging or sagging.
-% Note that almost all specimens have minimal secondary remanence component in thermal demagnetization results as shown in example orthogonal plots of all zero-field steps (Fig.
-\ref{fig:IZZI_examples}).
-This reflects that the rocks have experienced mild alterations after their formation.
+Paleointensity experiment results from both lithologies show three general types of dominant specimen behaviors on NRM/TRM plots (Fig.
+\ref{fig:IZZI_examples}): dominantly single-slope, double-slope or sagging, and zigzagging.
 Typical specimen paleointensity behavior of the anorthosites display straight, single-slope NRM/TRM plots and the accepted fractions of temperature steps span over the origin-trending, primary remanence components (Fig.
 \ref{fig:IZZI_examples}).
-We accept all specimen- and site-level absolute paleointensity estimates from the anorthosite xenoliths that pass the selection.
-All diabase specimens failed the selection because of double-slope behavior (fail FRAC selection; see methods), poor pTRM checks, and zigzagging (fail SCAT, DANG selection; see methods) plots (Fig.
+We accept all specimen- and site-level absolute paleointensity results from the anorthosite xenoliths that pass the selection.
+Diabase specimens failed the selection because of double-slope or sagging behavior (fail FRAC selection; see methods), poor pTRM checks, and zigzagging (fail SCAT, DANG selection; see methods) plots (Fig.
 \ref{fig:IZZI_examples}).
-As is shown in Fig.
+A 20 mT alternating field treatment after in-field heating steps was applied to some specimens but it did not result in significant changes in paleointensity results from the anorthosite xenoliths and the diabase (Fig.
+\ref{fig:PINT_cooling_corrected}; Supporting Information).
+
+In addition to estimating paleointensities by introducing a set of selection criteria to filter our experiment results, we apply an independent method of ref.
+\citealp{Cych2021a} which uses all experimental data regardless of their NRM/TRM plot statistics to perform bias corrected estimation of paleointensity (BiCEP).
+This BiCEP method is a Bayesian probabilistic method based on an assumption that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Arai1963a, Paterson2011a}.
+The posterior paleointensity distributions for sites with high-quality specimen-level data are in agreement with site-level weighted-mean results (Fig.
+S3), consistent with the interpretation that the anorthosite xenoliths record a high geomagnetic field \textit{ca.} 1092 Ma.
+
+Overall, the consistent specimen- and site-level behavior makes the paleointensity results from the anorthosites of high quality.
+The paleointensity quality index (Q$_{PI}$; ref.
+\citealp{Biggin2014a}) for the anorthosite sites are all above 4 (Table S2).
+
+% As is shown in Fig.
 \ref{fig:IZZI_examples}, specimen BD2-5b has a high-temperature slope with zigzagging behavior, poor pTRM checks, and low NRM/TRM ratio, in addition to a low-temperature slope with a steep NRM/TRM ratio.
@@
 
-A 20 mT alternating field treatment after in-field heating steps were applied to some specimens in an attempt to mitigate non-ideal behavior contributed by multidomain or non-uniformly magnetized magnetic carriers.
+% A 20 mT alternating field treatment after in-field heating steps were applied to some specimens in an attempt to mitigate non-ideal behavior contributed by multidomain or non-uniformly magnetized magnetic carriers.
 The treatment did not make any diabase specimen to pass the selection criteria.
@@
 Specimen results for AX11 particularly illustrates a case where all specimens within the site yield almost the same paleointensity estimate, regardless of the AF treatment.
-Therefore, the alternating field demagnetization did not result in significant increase in specimen success rate for the anorthosites although it might have improved the specimen behavior for those anorthosites that were subjected to this treatment.
+
+% Therefore, the alternating field demagnetization did not result in significant increase in specimen success rate for the anorthosites although it might have improved the specimen behavior for those anorthosites that were subjected to this treatment.
 Together with the narrow range of unblocking temperatures of their sister specimens \cite{Zhang2021b}, these data are consistent with the interpretation that the anorthosites with successful paleointensity results have a magnetic mineral assemblage that is likely dominated by single domain, low-Ti magnetite.
@@
 Such alterations can allow specimens to acquire additional remanence during in-field steps that may not be fully removed by an AF treatment.
-Overall, the consistent specimen-level behavior makes the successful results from the anorthosites of high confidence.
-The paleointensity quality index (Q$_{PI}$) for the anorthosite sites are all above 4 and the summary statistics are shown in Table S2.
-
-In addition to applying paleointensity selection criteria to our experiment results, we applied the bias corrected estimation of paleointensity (BiCEP) method by ref.
-\citealp{Cych2021a} to all of our anorthosite and diabase specimens regardless of their NRM/TRM plot statistics.
-The BiCEP method is a Bayesian statistical method based on an assumption that all specimen paleointensity results from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Arai1963a, Paterson2011a}.
-This method has the advantage of using as many specimens as available for estimating site-level paleointensities.
-The posterior distributions of paleointensity estimates given by the BiCEP method address the high quality of the paleointensity estimates from the anorthosite xenoliths, in that the posterior probabilities are very low for the case where these anorthosites are providing high paleointensity values that are heavily biased from a true value below 10 $\mu$T (Fig.
-S3).
-Instead, posterior probabilities for sites with high-quality specimen-level data are in agreement with site-level weighted-mean results, consistent with the interpretation that the anorthosite xenoliths record a high geomagnetic field \textit{ca.} 1092 Ma.
 
 \subsection*{Rock magnetism}
-Rock magnetic data support that the successful anorthosite specimens have dominant magnetic remanence carriers with magnetic properties similar to stochiometric, non-interactive, single domain magnetite, whereas those anorthosites that failed the paleointensity result selection and all diabase have more pronounced populations of non-ideal carriers.
-Magnetic property measurement system (MPMS) data show sister specimens to paleointensity specimens typically have dominant magnetic mineralogy of magnetite as can be evidenced through the Verwey transitions (Fig.
+
+Rock magnetic data support that the successful anorthosite specimens have dominant magnetic remanence carriers with magnetic properties similar to stochiometric, non-interacting, single domain magnetite, whereas anorthosite samples that failed the paleointensity result selection and all diabase samples have more pronounced populations of non-ideal carriers.
+Magnetic property measurement system (MPMS) data show both diabase and anorthosite have dominant magnetic mineralogy of magnetite as evidenced through the Verwey transitions (Fig.
 S1; ref.
 \citealp{Verwey1939a}).
-More, specimens from sites that yield successful paleointensity results have Verwey transition temperatures near 120 Kelvin (K) as expected for stochiometric low-Ti magnetite \cite{Ozdemir1993a}.
+Anorthosite chips from sites that yield successful paleointensity results have Verwey transition temperatures near 120 Kelvin (K) as expected for stochiometric low-Ti magnetite \cite{Ozdemir1993a}.
 However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have suppressed Verwey transition toward lower temperatures (Fig.
-S1), indicating that the specimens likely have a ferromagnetic mineralogy with variable degrees of oxidation \cite{Ozdemir1993a}.
+S1), indicating that the specimens likely have a magnetic mineralogy with variable degrees of oxidation or relatively high Ti content \cite{Ozdemir1993a}.
 
-Specimen coercivity spectra further support that our paleointensity selection criteria are successful in selecting experiment results.
+Coercivity spectra from backfield experiments further support the interpretation that rocks that pass or fail the selection criteria have distinct magnetic mineralogy.
 A compilation of median destructive field (MDF) in Fig.
-\ref{fig:coercivity} based on our backfield experiments shows that specimens from anorthosites that pass paleointensity selections have a distinctly higher average MDF range than other anorthosites and all diabase.
-Example single component skew-normal distribution fits of coercivity spectra \cite{Maxbauer2016a} based on backfield experiments on rock chips sister to paleointensity specimens show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity distributions around 80 mT (Fig.
+\ref{fig:coercivity} based on our backfield experiments shows that specimens from anorthosites that pass paleointensity selections have a distinctly higher average MDF range than other anorthosites and diabase.
+Example single component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity around 80 mT (Fig.
 \ref{fig:coercivity}).
-In contrast, specimens from anorthosite that did not pass paleointensity selection, as well as diabase specimens, tend to have coercivity distributions over a significantly lower range ($\sim$30 mT).
-This is consistent with the interpretation that they lack a dominance of single domain magnetite grains in their ferromagnetic mineral assemblages.
+In contrast, other anorthosite and diabase tend to have lower coercivities ($\sim$30 mT).
+This is consistent with the interpretation that they lack a dominance of single domain magnetite grains in their magnetic mineral assemblages.
 
@@
 \centering
-\caption{\footnotesize{Top: Example coercivity unmixing plots on anorthosites and diabase specimens from sites that pass or fail our paleointensity selection criteria.
-Bottom: Box plots summarize distributions of median destructive field (MDF) values for all anorthosite and diabase specimens that have single-component coercivity unmixing results.
+\caption{\footnotesize{Top: Example coercivity unmixing plots of anorthosite and diabase specimens from sites that pass or fail our paleointensity selection criteria.
+Bottom: Box plots of median destructive field (MDF) values for all anorthosite and diabase specimens with single-component coercivity unmixing results.
 Both plots show that anorthosite specimens that pass our paleointensity selection criteria have distinctly higher MDF than other anorthosites that do not pass the selection, as well as all diabase specimens.}}
@@
 The results show that the anorthosite specimens used for paleointensity experiment have gamma values ranging from $\sim$0.9\textdegree  to $\sim$11.9\textdegree, with a median value of 4.2\textdegree (Table S1).
-These gamma values are similar to those of the Midcontinent Rift volcanics \cite{Sprain2018a} and they show that the Beaver River anorthosite bulk rocks do not have a significant remanence anisotropy.
+These gamma values are similar to those of the Midcontinent Rift volcanics \cite{Sprain2018a}.
+Therefore, the Beaver River anorthosite bulk samples do not have a significant remanence anisotropy.
 Paleodirectional data from our anorthosite xenoliths further support that they have minimal remanence anisotropy as their site mean directions closely match those of the Beaver River diabase hosts \cite{Zhang2021b}.
@@
 \ref{fig:Petro_QDM}).
-Ref.
+
+We applied full TRMs to our anorthosite specimens that pass the paleointensity selection criteria with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axes.
+The results show that our anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions (Fig.
+S4).
+
+% Further, we applied full TRMs to the anorthosite specimens that pass the paleointensity selection criteria after paleointensity experiment.
+With an applied field of 30 $\mu$T oriented toward the specimen at 0, 30, 60, 90, 120, and 150 degrees, the anorthosite specimens usually acquired consistent full TRMs.
+The most significant variations observed in acquired full TRMs reach to only XXX \%, much less than that has been reported to occurred within layerd anorthositic rocks in Stillwater (up to $\sim$60 \%, \cite{Selkin2000a}).
+
+% Ref.
 \citealp{Selkin2007a} also noted that vortex state magnetic grains and certain single domain particles can produce biased paleointensity estimates due to their acquisition of saturation remanence during laboratory in-field steps.
@@
 Paleointensity estimates from igneous intrusions can also be biased toward higher values due to slow cooling rates \cite{Halgedahl1980a}.
-It has been shown that large differences in cooling rates between acquisition of an NRM versus a laboratory pTRM can result in overestimated paleointensities for SD grains \cite{Dodson1980a, Halgedahl1980a}.
+It has been shown that large differences in cooling rates between acquisition of an NRM versus a laboratory pTRM can result in overestimated paleointensities for SD grains \cite{Dodson1980a, Halgedahl1980a, Nagy2021a}.
 From the thermal history model of ref.
-\citealp{Zhang2021b} we can estimate the duration which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully block in their characteristic natural remanence component ($\sim$500\textdegree C; ref.
+\citealp{Zhang2021b} we can estimate the duration which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully block in their characteristic natural remanence magnetization ($\sim$500\textdegree C; ref.
 \citealp{Zhang2021b}).
-We find the cooling time to be on the order of $\sim$1 kyr, which corresponds to a cooling rate of $1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
-In contrast, the lab cooling rate is much faster, at about $1.3\times10^{-1}$ $^\circ$C s$^{-1}$, assuming that it takes 1000s in lab condition to cool through the same temperature interval.
+We find the cooling time to be on the order of $\sim$1 kyr, which corresponds to a cooling rate of $\sim 1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
+In contrast, the lab cooling rate is much faster, at $\sim 1.3\times10^{-1}$ $^\circ$C s$^{-1}$, assuming that it takes 1000s in lab condition to cool through the same temperature interval.
 This significant cooling rate difference leads to a predicted $\sim$36\% overestimate of true ancient field following the model of ref.
@@
 S2).
+This estimate on cooling rate effect is similar to that derived from the model of ref.
+\citealp{Nagy2021a}.
 We therefore correct our paleointensity results by a factor of 0.74.
@@
 \ref{fig:PINT_cooling_corrected}.
-Taking all accepted anorthosite paleointensity results, they give a site mean absolute paleointensity estimate of 47.90 $\pm$ 9.32 $\mu$T (Fig.
+Taking all accepted anorthosite paleointensity results, they give a site mean absolute paleointensity estimate of 52.72 $\pm$ 5.72 $\mu$T (Fig.
 \ref{fig:PINT_cooling_corrected}).
@@
 \section*{Proterozoic geodynamo}
-To best characterize the geomagnetic axial dipole field intensity during a certain time period requires that a paleointensity dataset to cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged and the geocentric axial dipole hypothesis can be assumed.
+To best characterize the geomagnetic axial dipole field intensity during a certain time period requires a paleointensity dataset to cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged and the geocentric axial dipole hypothesis can be assumed.
 Thermal modeling results from ref.
-\citealp{Zhang2021b} suggest that the anorthosite xenoliths were likely heated to tholeiitic magma temperature and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a few thousand years.
-In addition, the closely located sites AX12 and AX16 have very similar paleointensity estimates that are distinct from those of sites AX11 and AX13 which were emplaced $\sim$10 km away (Fig.
+\citealp{Zhang2021b} suggest that the Beaver River anorthosite xenoliths were likely heated to tholeiitic magma temperature and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a few thousand years.
+In addition, anorthosite xenoliths have paleointensity results that are consistent within small regions but variations between localities.
+Anorthosite AX12 and AX16 have very similar paleointensity estimates that are distinct from those of sites AX11 and AX13 which were emplaced $\sim$10 km away (Fig.
 \ref{fig:Geologic_map}).
-The consistency within small regions but variations between regions is consistent with the interpretation that differential cooling history due to heterogeneity of the timing and depths of the Beaver River diabase intrusions caused the anorthosites to capture paleosecular variations with their magnetic remanence \textit{ca.} 1092 Ma.
-Overall, none of the anorthosite xenoliths record a site mean VDM value lower than 40 ZAm$^2$ (Fig.
+This is consistent with the interpretation that differential cooling history due to heterogeneity of the timing and depths of the Beaver River diabase intrusions caused the anorthosites to capture paleosecular variations \textit{ca.} 1092 Ma.
+% Overall, none of the anorthosite xenoliths record a site mean VDM value lower than 40 ZAm$^2$ (Fig. \ref{fig:PINT_compilation}).
+
+The new paleointensity records from the \textit{ca.} 1092 Ma Midcontinent Rift anorthosite xenoliths together with previous results from the \textit{ca.} 1140 Ma rocks of the Midcontinent Rift region indicate that a dramatic six-fold increase in geomagnetic dipole moment happened within $\sim$50 Myr during the late Mesoproterozoic (Fig.
 \ref{fig:PINT_compilation}).
+The Abitibi dikes \cite{Macouin2003a} record a low average virtual dipole moment of $\sim$13 ZAm$^2$, similar to the values of $\sim$7 ZAm$^2$ obtained from the single silicate crystals of the \textit{ca.} 565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a}.
+However, the relatively high moments of $\sim$72 ZAm$^2$ from the \textit{ca.} 1105 Ma Osler Volcanic Group \cite{Sprain2018a} and $\sim$89 ZAm$^2$ from the \textit{ca.} 1092 Ma Beaver River anorthosite xenoliths (Fig.
+\ref{fig:PINT_compilation}) are about six times the values of the Abitibi dikes.
+The scale of this increase is similar to the increasing trend observed near the Ediacaran-Cambrian which has been suggested to evidence the onset of Earth's inner core nucleation \cite{Thallner2021a}.
 
-The new paleointensity records from the Midcontinent Rift anorthosite xenoliths together with previous results from other Mesoproterozoic rocks suggest that a dramatic six-fold increase in geomagnetic dipole moment happened within $\sim$50 Myr during the late Mesoproterozoic (Fig.
-\ref{fig:PINT_compilation}).
-The scale of the increase is similar to the increasing trend observed near the Ediacaran-Cambrian which has been suggested to evidence the onset of Earth's inner core nucleation \cite{Thallner2021a}.
-The \textit{ca.} 1140 Ma Abitibi dikes \cite{Macouin2003a} record an average virtual dipole moment of $\sim$13 ZAm$^2$, similar to the low values of $\sim$7 ZAm$^2$ obtained from the single silicate crystals of the \textit{ca.} 565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a}.
-However, the relatively high moments of $\sim$72 ZAm$^2$ from the \textit{ca.} 1105 Ma Osler Volcanic Group \cite{Sprain2018a} and $\sim$89 ZAm$^2$ from the \textit{ca.} 1092 Ma Beaver River anorthosite xenoliths (Fig.
-\ref{fig:PINT_compilation}) are about six times the values from the Abitibi dikes.
+The high geomagnetic field estimates from the Midcontinent Rift volcanics and intrusives are not the only suite of high paleointensity records at odds with the hypothesis of a monotonically decaying geodynamo through the entire Proterozoic Eon \cite{Biggin2009a, Bono2019a}.
+The $\sim$60 ZAm$^2$ dipole moment estimates from the \textit{ca.} 755 Ma Mundine Well dikes are discordant with the long-term decrease trend \cite{Bono2019a} and are about three times the estimates from the \textit{ca.} 720 Ma dolerite dikes of the Franklin Large Igneous Province \cite{Lloyd2021a}.
+On the other hand, the low geomagnetic field estimates from the \textit{ca.} 565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a} appears to be not a unique occurrence during Earth's history.
+Besides the low values recorded by the Abitibi dikes, low paleointensity estimates has been obtained from the \textit{ca.} 1070 Ma Bangemall Sills \cite{Lloyd2021b} and Devonian-aged rocks from the Minusa Basin, Kola Peninsula, and Buribay volcanic complex (Fig.
+\ref{fig:PINT_compilation}; ref.
+\citealp{Shcherbakova2017a, Shcherbakova2021a}).
+In addition, a five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years, when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed.
+Therefore, paleointensity dataset alone may not be sufficient for interpreting Earth's inner core evolution history.
+This makes the paleomagnetic directional anomalies during the Ediacaran and Devonian \cite{Bono2015a, Kodama2021a, Shcherbakova2019a} intriguing aspects to keep investigating.
+Regardless, the high values from the Midcontinent Rift intrusions and extrusive volcanics are inconsistent with the prediction of there being a progressive decay at the time (Fig.
+\ref{fig:PINT_compilation}; ref.
+\citealp{Bono2019a}) and necessitates that there being a vigorous late Mesoproterozoic geodynamo.
 
-The high geomagnetic field estimates from the Midcontinent Rift volcanics and intrusives are not the only suite of records at odds with the hypothesis of a progressive decay trend during the entire Proterozoic Eon \cite{Biggin2009a, Bono2019a}.
-The $\sim$60 ZAm$^2$ dipole moment estimates from the \textit{ca.} 755 Ma Mundine Well dikes are discordant with the long-term decrease trend \cite{Bono2019a} and are about three times the estimates from the \textit{ca.} 720 Ma dolerite dikes of the Franklin Large Igneous Province \cite{Lloyd2021a}.
-Similarly, an apparent rapid decrease in geomagnetic dipole moment from \textit{ca.} 1092 Ma to \textit{ca.} 1070 Ma can be evidenced by the contrasting high values from the Midcontinent Rift anorthosite xenoliths and the low values from the Bangemall Sills (Fig.
+% conclude on whether an apparent rise or decay within a narrow time period ($\sim$35 Ma) during the Proterozoic is associated with transitions in the core phase transition or is a result of short-term excursion of the geomagnetic field that may be common through much of the evolution of Earth's core.
+
+% an apparent rapid decrease in geomagnetic dipole moment from \textit{ca.} 1092 Ma to \textit{ca.} 1070 Ma can be evidenced by the contrasting high values from the Midcontinent Rift anorthosite xenoliths and the low values from the Bangemall Sills (Fig.
 \ref{fig:PINT_compilation}; ref.
@@
 In addition, given that a five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years, when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed, we cannot conclude on whether an apparent rise or decay within a narrow time period ($\sim$35 Ma) during the Proterozoic is associated with transitions in the core phase transition or is a result of short-term excursion of the geomagnetic field that may be common through much of the evolution of Earth's core.
+Recently, ultra-low paleointensities (2.01–7.07) $\mu$T from the Devonian has been found by ref.
+\cite{Shcherbakova2021a}.
+These data suggest that significant excursion of low geomagnetic field intensities may occur regardless of the inner core formation.
 
-Regardless of whether the Abitibi dikes recorded low virtual dipole moments of a geodynamo with waning strength or an excursion at \textit{ca.} 1140 Ma, the high values from the Midcontinent Rift intrusions and extrusive volcanics are inconsistent with the prediction of there being a progressive decay at the time (Fig.
-\ref{fig:PINT_compilation}; ref.
-\citealp{Bono2019a}) and necessitates that there being a vigorous late Mesoproterozoic geodynamo.
+% These data suggest that an apparent transition of virtual dipole moment similar in scale with what is predicted for a dynamo powered with or without compositional convection in the core may not be unique during the Precambrian.
+In addition, given that a five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years, when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed, we cannot conclude on whether an apparent rise or decay within a narrow time period ($\sim$35 Ma) during the Proterozoic is associated with transitions in the core phase transition or is a result of short-term excursion of the geomagnetic field that may be common through much of the evolution of Earth's core.
 
@@
 \citealp{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
-The criteria are also summarized in Table \ref{tab:criteria}.
 The MAD measures the scatter about the best-fit line through the natural remanent magnetization (NRM) steps in the selected interval for which the intensity is defined.

### 2021-12-20 PNAS_manuscript/PNAS_manuscript.tex

@@
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{Acquiring high-fidelity observational records of ancient magnetic field intensity from remanent magnetization of rocks is crucial for constraining the evolution history of Earth's core.
+\significancestatement{Acquiring high-fidelity observational records of ancient magnetic field intensity from remanent magnetization of rocks is crucial for constraining the evolution of Earth's core.
 However, robust estimates of ancient field strengths are often difficult to recover due to alterations or non-ideal rock magnetic behaviors.
-In this study, we use plagioclase cumulates that formed in the deeper crust and were brought to the near surface by magma where they cooled and acquired thermal remanent magnetizations.
+In this study, we use plagioclase cumulates that formed in the deep crust and were brought to the near surface by magma where they cooled and acquired thermal remanent magnetizations.
 These plagioclase from the North American Midcontinent Rift have experienced minimal post-formation alterations and record a present-day-like virtual dipole moment 1.1 billion years ago.
@@
 In this study, we target \textit{ca.} 1092 Ma anorthosite xenoliths that cooled and acquired their magnetization within Beaver River diabase intrusions of the North American Midcontinent Rift.
-In contrast to the diabase which has non-ideal paleointensity behaviors and fails to provide reliable paleointensity estimates, the anorthosite xenoliths are unusually high fidelity recorders yielding high-quality, single-slope paleointensity results that are consistent at specimen and site levels.
+In contrast to the diabase which has non-ideal paleointensity behaviors and fails to provide reliable paleointensity estimates, the anorthosite xenoliths are unusually high-fidelity recorders yielding high-quality, single-slope paleointensity results that are consistent at specimen and site levels.
 An average value of $\sim$89 ZAm$^2$ for the virtual dipole moment from the anorthosites is higher than that of the dipole component of Earth's magnetic field today.
@@
 Evidence for a persistent magnetic field through the Proterozoic, for example, necessitates the existence of plate tectonics that sustained core-mantle boundary heat flow.
-However, strikingly low estimates of geomagnetic field strengths have been obtained in the Ediacaran Period (ref.
-\citealp{Bono2019a}; \textit{ca.} 565 Ma), potentially indicating an unusual period of geodynamo regime at the time.
+However, strikingly low estimates of geomagnetic field strengths have been obtained in the Ediacaran Period \cite{Bono2019a, Thallner2021b}, potentially indicating an unusual period of core dynamo regime at the time.
 
@@
 Determinations of ancient magnetic field strengths rely on igneous rocks that acquire thermal remanent magnetizations as they cool in Earth's magnetic field.
-These magnetizations need to be unmodified by subsequent heating or chemical alterations in order to be able to faithfully record ancient geomagnetic field record through time.
+These magnetizations need to be unmodified by subsequent heating or chemical alterations in order to be able to faithfully record ancient geomagnetic field through time.
 Intraplate magmatic events are therefore an important target as they can be well-preserved within continental interiors at distance from mountain-building events along continental margins.
-However, magmatic events associated with large igneous provinces are typically of geologically short duration with the bulk of magmatic products emplaced within $<$1 Myr.
+However, magmatic events such as the formation of large igneous provinces that are often associated with intraplate magmatism are typically of geologically short duration with the bulk of magmatic products emplaced within 1 Myr.
 The Midcontinent Rift is an exception as magmatism lasted $\sim$25 Myr from 1109 Ma to 1084 Ma.
@@
 As a result, rocks of the rift have simple paleomagnetic behavior and form a central record of Mesoproterozoic paleogeography \cite{Swanson-Hysell2021c}.
-This simple magnetization and the long duration of magnetization holds the potential for a well-preserved extended window into the intensity of Earth's magnetic field from 1109 to 1084 Ma.
+The simple magnetizations held by igneous rocks that formed through an extended duration from 1109 to 1084 Ma that is longer than the Neogene Period have the potential for a well-preserved window into the intensity of Earth's magnetic field .
 
@@
 One typical failure of paleointensity experiment is the double-slope behavior, as two distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
-Ref \citealp{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas ref.
+Ref.
+\citealp{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas ref.
 \citealp{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
@@
 In this study, we target a unique rock type - high-purity anorthosite xenoliths.
-These anorthosite xenoliths were brought to the near surface with magma that formed sills of the Beaver River diabase.
+These anorthosite xenoliths were brought to the near surface with magma that formed sills of the Beaver River diabase \cite{Zhang2021b}.
 They cooled and acquired their magnetizations in conjunction with the host diabase.
@@
 While magmatic activity within the Midcontinent Rift was protracted, there were volcanism and intrusions characterized by particularly rapid and voluminous emplacement.
-The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex in northern Minnesota punctuates one such period of magmatism during the Midcontinent Rift main stage activity.
+The \textit{ca.} 1092 Ma Beaver Bay Complex in northern Minnesota punctuates one such period of magmatism during the Midcontinent Rift main stage activity.
 The magma that formed the 1091.7 $\pm$ 0.2 Ma Beaver River diabase dikes and sills of the Beaver Bay Complex transported numerous anorthosite xenoliths that have short-axis diameters up to 180 meters via wide conduits \cite{Boerboom2004a, Boerboom2006b}.
@@
 
-Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths to have acquired thermal remanent magnetizations while cooing with the Beaver River diabase \cite{Zhang2021b}.
-Step-wise thermal demagnetizations show the anorthosite xenoliths have dominantly single-component magnetization that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C.
+Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths to have acquired thermal remanent magnetizations while cooling with the Beaver River diabase \cite{Zhang2021b}.
+Step-wise thermal demagnetizations show the anorthosite xenoliths have dominantly single-component magnetization that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C \cite{Zhang2021b}.
 
@@
 Petrographic imaging shows the dominantly monomineralic anorthosite xenoliths of the Beaver River diabase often develop granoblastic texture characterized by equigranular crystals with poor development of petrofabrics (Fig.
-\ref{fig:Petro_QDM}A).
+\ref{fig:Petro_QDM}A,C).
 In addition, the plagioclase crystals of the anorthosite xenoliths are often free of large inclusions.
@@
 The plagioclase of the Duluth Complex anorthosites typically develop interlocking textures that display strong igneous foliations (Fig.
-\ref{fig:Petro_QDM}B) and associated magnetic mineral fabrics characterized by having aligned titanomagnetite needles within plagioclase crystals (Fig.
+\ref{fig:Petro_QDM}B) and associated titanomagnetite mineral fabrics characterized by having aligned needles within plagioclase crystals (Fig.
 \ref{fig:Petro_QDM}D).
-Magnetic imaging confirms that the titanomagnetite needles within the plagioclase crystals are often strongly magnetic with magnetic fabrics oriented along the long axes of the needles (Fig.
-\ref{fig:Petro_QDM}F).
+Magnetic imaging confirms that the titanomagnetite needles within the plagioclase crystals are often strongly magnetic with magnetic moments of individual needles oriented along their long axes (Fig.
+\ref{fig:Petro_QDM}D,F).
 Titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex plagioclase \cite{Miller1990a}.
 
-The lack of symplecites and exsolved titanomagnetite needles within the Beaver River anorthosite xenoliths reflects that the Beaver River diabase magma had a different oxidation history than that of the Duluth Complex.
-The plagioclase and oxide mineral textures of the Beaver River anorthosites also distinguish them from plagioclase cumulates of other layered mafic complex where magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
+The lack of symplecites and exsolved titanomagnetite needles within the Beaver River anorthosite xenoliths distinguish them from plagioclase cumulates of the Duluth Complex and other layered mafic complex where anisotropic magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
 
@@
 
-Following IZZI-modified paleointensity experiments, 40 from a total of 86 anorthosite specimens and 0 out of a total of 69 diabase specimens passed our paleointensity result selection criteria of having (1) a maximum angular deviation (MAD; ref.
+Following IZZI-modified paleointensity experiments \cite{Yu2004a}, 40 from a total of 86 anorthosite specimens and 0 out of a total of 69 diabase specimens passed our paleointensity result selection criteria of having (1) a maximum angular deviation (MAD; ref.
 \citealp{Kirschvink1980a}) of $<$20\textdegree; (2) scatter parameter ($\beta$; ref.
@@
 \citealp{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
-7 anorthosite sites and no diabase sites with site-level paleointensity results pass our selections.
+7 anorthosite sites and no diabase site-level results pass our selections.
 Example NRM/TRM plots are shown in Fig.
 \ref{fig:IZZI_examples}.
-The cooling rate corrected absolute paleointensity estimates from all anorthosite sites have a weighted mean of 52.72 $\pm $5.72 $\mu$T.
-This corresponds to a calculated virtual dipole moment of $\sim$89 ZAm$^2$ \textit{ca.} 1092 Ma.
 Summary specimen absolute paleointensity estimates and site-level weighted mean paleointensity values are plotted in Fig.
 \ref{fig:PINT_cooling_corrected} (and provided in Table S1).
+The cooling rate-corrected absolute paleointensity estimates from all anorthosite sites have a weighted mean of 52.72 $\pm $5.72 $\mu$T.
+This corresponds to a calculated virtual dipole moment of $\sim$89 ZAm$^2$ ($10^{21} Am^2$) \textit{ca.} 1092 Ma.
 All measurement-level paleointensity experiment data is available within the MagIC database (\url{}; UPDATE WHEN DOI IS GENERATED).
@@
 
-Paleointensity experiment results from both lithologies show three general types of dominant specimen behaviors on NRM/TRM plots (Fig.
+Paleointensity experiment results from the anorthosite xenoliths and their host diabase show three general types of dominant specimen behaviors on NRM/TRM plots (Fig.
 \ref{fig:IZZI_examples}): dominantly single-slope, double-slope or sagging, and zigzagging.
@@
 We accept all specimen- and site-level absolute paleointensity results from the anorthosite xenoliths that pass the selection.
-Diabase specimens failed the selection because of double-slope or sagging behavior (fail FRAC selection; see methods), poor pTRM checks, and zigzagging (fail SCAT, DANG selection; see methods) plots (Fig.
+Diabase specimens failed the selection because of double-slope or sagging behavior (fail FRAC selection; see methods), poor pTRM checks, and zigzagging plots (fail SCAT, DANG selection; see methods; Fig.
 \ref{fig:IZZI_examples}).
-A 20 mT alternating field treatment after in-field heating steps was applied to some specimens but it did not result in significant changes in paleointensity results from the anorthosite xenoliths and the diabase (Fig.
+A 20 mT alternating field treatment after in-field heating steps was applied to some specimens but it did not result in significant changes in paleointensity behaviors of the anorthosite xenoliths or the diabase (Fig.
 \ref{fig:PINT_cooling_corrected}; Supporting Information).
@@
 \citealp{Cych2021a} which uses all experimental data regardless of their NRM/TRM plot statistics to perform bias corrected estimation of paleointensity (BiCEP).
-This BiCEP method is a Bayesian probabilistic method based on an assumption that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Arai1963a, Paterson2011a}.
-The posterior paleointensity distributions for sites with high-quality specimen-level data are in agreement with site-level weighted-mean results (Fig.
-S3), consistent with the interpretation that the anorthosite xenoliths record a high geomagnetic field \textit{ca.} 1092 Ma.
+This Bayesian probabilistic method is based on an assumption that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Arai1963a, Paterson2011a}.
+The posterior paleointensity distributions for sites with high-quality specimen-level data (Fig.
+\ref{fig:PINT_cooling_corrected}) are in agreement with site-level weighted-mean results (Fig.
+S3).
 
-Overall, the consistent specimen- and site-level behavior makes the paleointensity results from the anorthosites of high quality.
+Overall, the high-quality experiment results from the anorthosite xenoliths of the Beaver River diabase and the consistent paleointensity estimates based on independent methods confirm that the anorthosite xenoliths record a high geomagnetic field \textit{ca.} 1092 Ma.
 The paleointensity quality index (Q$_{PI}$; ref.
-\citealp{Biggin2014a}) for the anorthosite sites are all above 4 (Table S2).
+\citealp{Biggin2014a}) for the anorthosite xenoliths are all above 4 (Table S2).
 
@@
 
-Rock magnetic data support that the successful anorthosite specimens have dominant magnetic remanence carriers with magnetic properties similar to stochiometric, non-interacting, single domain magnetite, whereas anorthosite samples that failed the paleointensity result selection and all diabase samples have more pronounced populations of non-ideal carriers.
-Magnetic property measurement system (MPMS) data show both diabase and anorthosite have dominant magnetic mineralogy of magnetite as evidenced through the Verwey transitions (Fig.
+Rock magnetic data support that the anorthosite specimens that pass the paleointensity selection criteria have dominant magnetic remanence carriers with magnetic properties similar to stochiometric, non-interacting, single domain magnetite, whereas anorthosite samples that failed the paleointensity result selection and all diabase samples have more pronounced populations of non-ideal carriers.
+Magnetic property measurement system (MPMS) data show both diabase and anorthosite have dominant magnetic mineralogy of magnetite as evidenced through the Verwey transition (Fig.
 S1; ref.
 \citealp{Verwey1939a}).
-Anorthosite chips from sites that yield successful paleointensity results have Verwey transition temperatures near 120 Kelvin (K) as expected for stochiometric low-Ti magnetite \cite{Ozdemir1993a}.
+More, anorthosite chips from sites that yield successful paleointensity results have Verwey transition temperatures near 120 Kelvin (K) as expected for stochiometric low-Ti magnetite \cite{Ozdemir1993a}.
 However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have suppressed Verwey transition toward lower temperatures (Fig.
@@
 A compilation of median destructive field (MDF) in Fig.
-\ref{fig:coercivity} based on our backfield experiments shows that specimens from anorthosites that pass paleointensity selections have a distinctly higher average MDF range than other anorthosites and diabase.
-Example single component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity around 80 mT (Fig.
+\ref{fig:coercivity} shows that specimens from anorthosites that pass paleointensity selections have a distinctly higher average MDF range ($\sim$60 mT) than other anorthosites and diabase ($\sim$30 mT).
+Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity around 80 mT (Fig.
 \ref{fig:coercivity}).
 In contrast, other anorthosite and diabase tend to have lower coercivities ($\sim$30 mT).
-This is consistent with the interpretation that they lack a dominance of single domain magnetite grains in their magnetic mineral assemblages.
+This is consistent with the interpretation that they lack a dominance of single domain magnetite grains in their magnetic mineral assemblages and thus have non-ideal paleointensity behaviors.
 
@@
 Bottom: Box plots of median destructive field (MDF) values for all anorthosite and diabase specimens with single-component coercivity unmixing results.
-Both plots show that anorthosite specimens that pass our paleointensity selection criteria have distinctly higher MDF than other anorthosites that do not pass the selection, as well as all diabase specimens.}}
+Both plots show that anorthosite specimens that pass paleointensity selection criteria have distinctly higher portion of single-domain-like magnetic mineralogy than other anorthosite and diabase.}}
 \label{fig:coercivity}
@@
 Ref.
-\citealp{Selkin2000a} demonstrated that strong remanence anisotropy associated with the igneous foliation developed within an anorthosite sample from the Stillwater Complex can lead to significant overestimate or underestimate of paleointensity values depending on the relative orientations between the external applied field magnetic anisotropy axes of a specimen.
+\citealp{Selkin2000a} demonstrated that strong remanence anisotropy associated with the igneous foliations developed within an anorthosite sample from the Stillwater Complex can lead to significant overestimate or underestimate of paleointensity values depending on the relative orientations between the external applied field magnetic anisotropy axes of a specimen.
 To assess whether our paleointensity estimates are biased by remanence anisotropy, we calculated the gamma statistic, which is the angular difference between the last pTRM step of paleointensity experiment and the applied field direction.
@@
 These gamma values are similar to those of the Midcontinent Rift volcanics \cite{Sprain2018a}.
-Therefore, the Beaver River anorthosite bulk samples do not have a significant remanence anisotropy.
+Therefore, the Beaver River anorthosite xenolith bulk samples do not have a significant remanence anisotropy.
 Paleodirectional data from our anorthosite xenoliths further support that they have minimal remanence anisotropy as their site mean directions closely match those of the Beaver River diabase hosts \cite{Zhang2021b}.
-In addition, no strong magnetic fabric was observed through magnetic imaging (Fig.
+In addition, no strong magnetic fabric was observed in the anorthosite xenoliths through magnetic imaging (Fig.
 \ref{fig:Petro_QDM}).
 
-We applied full TRMs to our anorthosite specimens that pass the paleointensity selection criteria with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axes.
+We applied full TRMs to a set of our anorthosite specimens that pass the paleointensity selection criteria with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axes.
 The results show that our anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions (Fig.
@@
 \ref{fig:PINT_cooling_corrected}.
-Taking all accepted anorthosite paleointensity results, they give a site mean absolute paleointensity estimate of 52.72 $\pm$ 5.72 $\mu$T (Fig.
-\ref{fig:PINT_cooling_corrected}).
 
@@
 \citealp{Zhang2021b} suggest that the Beaver River anorthosite xenoliths were likely heated to tholeiitic magma temperature and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a few thousand years.
-In addition, anorthosite xenoliths have paleointensity results that are consistent within small regions but variations between localities.
+In addition, anorthosite xenoliths have paleointensity results that are consistent within small regions but vary between localities.
 Anorthosite AX12 and AX16 have very similar paleointensity estimates that are distinct from those of sites AX11 and AX13 which were emplaced $\sim$10 km away (Fig.
@@
 
-The new paleointensity records from the \textit{ca.} 1092 Ma Midcontinent Rift anorthosite xenoliths together with previous results from the \textit{ca.} 1140 Ma rocks of the Midcontinent Rift region indicate that a dramatic six-fold increase in geomagnetic dipole moment happened within $\sim$50 Myr during the late Mesoproterozoic (Fig.
+The new paleointensity records from the \textit{ca.} 1092 Ma Midcontinent Rift anorthosite xenoliths together with previous results from the \textit{ca.} 1140 Ma rocks of the Midcontinent Rift region indicate that a six-fold increase in geomagnetic dipole moment happened within $\sim$50 Myr during the late Mesoproterozoic (Fig.
 \ref{fig:PINT_compilation}).
@@
 However, the relatively high moments of $\sim$72 ZAm$^2$ from the \textit{ca.} 1105 Ma Osler Volcanic Group \cite{Sprain2018a} and $\sim$89 ZAm$^2$ from the \textit{ca.} 1092 Ma Beaver River anorthosite xenoliths (Fig.
-\ref{fig:PINT_compilation}) are about six times the values of the Abitibi dikes.
+\ref{fig:PINT_compilation}) are strikingly higher than the values of the Abitibi dikes.
 The scale of this increase is similar to the increasing trend observed near the Ediacaran-Cambrian which has been suggested to evidence the onset of Earth's inner core nucleation \cite{Thallner2021a}.
@@
 The high geomagnetic field estimates from the Midcontinent Rift volcanics and intrusives are not the only suite of high paleointensity records at odds with the hypothesis of a monotonically decaying geodynamo through the entire Proterozoic Eon \cite{Biggin2009a, Bono2019a}.
-The $\sim$60 ZAm$^2$ dipole moment estimates from the \textit{ca.} 755 Ma Mundine Well dikes are discordant with the long-term decrease trend \cite{Bono2019a} and are about three times the estimates from the \textit{ca.} 720 Ma dolerite dikes of the Franklin Large Igneous Province \cite{Lloyd2021a}.
-On the other hand, the low geomagnetic field estimates from the \textit{ca.} 565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a} appears to be not a unique occurrence during Earth's history.
-Besides the low values recorded by the Abitibi dikes, low paleointensity estimates has been obtained from the \textit{ca.} 1070 Ma Bangemall Sills \cite{Lloyd2021b} and Devonian-aged rocks from the Minusa Basin, Kola Peninsula, and Buribay volcanic complex (Fig.
+The $\sim$60 ZAm$^2$ virtual dipole moment estimates from the \textit{ca.} 755 Ma Mundine Well dikes are much higher than the values predicted by the monotonic decrease model of \cite{Bono2019a}.
+Although one may consider these data capturing relatively short term fluctuations of the geomagnetic field, they nevertheless indicate that the geodynamo had strong power sources during the Proterozoic.
+On the other hand, besides the low paleointensity values recorded by the \textit{ca.} 1140 Ma Abitibi dikes, low estimates have been obtained from the \textit{ca.} 1070 Ma Bangemall Sills \cite{Lloyd2021b}, the \textit{ca.} 720 Ma dolerite dikes of the Franklin Large Igneous Province \cite{Lloyd2021a}, as well as Devonian-aged rocks from the Minusa Basin, Kola Peninsula, and Buribay volcanic complex (Fig.
 \ref{fig:PINT_compilation}; ref.
 \citealp{Shcherbakova2017a, Shcherbakova2021a}).
-In addition, a five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years, when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed.
-Therefore, paleointensity dataset alone may not be sufficient for interpreting Earth's inner core evolution history.
-This makes the paleomagnetic directional anomalies during the Ediacaran and Devonian \cite{Bono2015a, Kodama2021a, Shcherbakova2019a} intriguing aspects to keep investigating.
+The multiple occurrences of low geomagnetic field intensities during the Proterozoic and the Paleozoic indicate that a waning thermal convection regime in Earth's before the emergence of the solid inner core may not be the solely possible mechanism for triggering an weak dipole field.
+
+In addition, a five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years \cite{Bono2021a}, when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed.
+The observation of weak fields before and after the Cambrian indicate that the geomagnetic dipole may reach particularly low strengths regardless of the existence of a solid inner core.
+Therefore, low paleointensity values or short-term fluctuations may not be sufficient for interpreting Earth's inner core status through history.
+
+Recent advancements in high-pressure, high-temperature experiments for measuring thermal conductivity of iron alloys under conditions similar to Earth's core may shed new lights on the age of the inner core.
+However, large uncertainties still remain regarding the applicability Wiedemann-Franz law of extrapolating thermal conductivity values based on electrical resistivity measurements by ref.
+\citealp{Ohta2016a}, and regarding the thermal modeling parameters used in the direct thermal conduction measurement experiment performed by ref.
+\citealp{Konopkova2016a}.
+Until further experiments and theory that can explain these contrasting results are developed, our knowledge on the age of Earth's inner core remain largely uncertain.
+
+In the meantime, paleomagnetic directional anomalies may be another suite of intriguing observational records that is worth investigating.
+The anomalous directional fluctuations during the Ediacaran and Devonian \cite{Bono2015a, Kodama2021a, Shcherbakova2019a} perhaps can be consistent with there being a weak geodynamo regime when dipole component of Earth's field was largely diminished.
+
 Regardless, the high values from the Midcontinent Rift intrusions and extrusive volcanics are inconsistent with the prediction of there being a progressive decay at the time (Fig.
@@
 \centering
-\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment (VADM) values from the PINT database (v.2015.05; \url{http://earth.liv.ac.uk/ pint/}; \cite{Biggin2015a}) with additional Mesoproterozoic data from \cite{Sprain2018a, Kodama2019a}, as well as those from the Neoproterozoic \cite{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}.
+\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment values from the PINT database (PINT v8.0.0; \url{http://www.pintdb.org/}; ref.
+\citealp{Bono2021a}) with additional Mesoproterozoic data from \cite{Sprain2018a, Kodama2019a}, as well as those from the Neoproterozoic \cite{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}.
 Paleointensity estimates from \cite{Pesonen1983a} and \cite{Kulakov2013a} are not included in the compilation due to the specimen-level double-slope behavior as discussed in the text.
@@
 Sun compass orientations were preferentially used for determining the sample azimuth.
-Specimens from a total of 7 S1).
 Sister specimens underwent step-wise alternating field (AF) or thermal demagnetization at the UC Berkeley Paleomagnetism Lab and the results are presented in \cite{Zhang2021b}.
-The close directional similarities between the anorthosite xenoliths and their host diabase and thermal modeling results are consistent with the interpretation that the Beaver River diabase and anorthosite xenoliths acquired primary thermal remanent magnetization \textit{ca.} 1092 Ma during cooling after the emplacement of the diabase \cite{Zhang2021b}.
 Based on the anorthosite thermal demagnetization results we selected sites whose unblocking temperature ranges are narrow and near 580\textdegree C for paleointensity experiments.
@@
 \subsection*{Paleointensity experiment}
-A total of 86 specimens from 14 anorthosite xenoliths and a total of 69 specimens from 7 diabase sites underwent paleointensity experiments that followed the step-wise double-heating Thellier method \cite{Thellier1959a} using the IZZI protocol \cite{Tauxe2004a} with heating steps up to 585 \textdegree C.
+A total of 86 specimens from 14 anorthosite xenoliths and a total of 69 specimens from 7 diabase sites underwent paleointensity experiments that followed the step-wise double-heating Thellier method \cite{Thellier1959a} using the IZZI protocol \cite{Yu2004a} with heating steps up to 585 \textdegree C.
 Partial thermal remanent magnetization (pTRM) checks were performed systematically throughout the experiment to test whether there was significant mineralogical alteration due to heating and were assessed using the SCAT parameter of ref.
 \citealp{Shaar2013a}.
-On top of the IZZI modification of the Thellier experimental protocol, we also performed a comparative study where we added an extra step of 20 mT alternating field (AF) cleaning on some of the specimens after each in-field step.
-The purpose is to study whether the AF cleaning could help improve experiment success rate by removing the remanence component carried by materials such as multi-domain (MD) or vortex state grains that often contributes to non-ideal paleointensity behaviors.
+On top of the IZZI-Thellier experiment protocol, we also performed a comparative study where we added an extra step of 20 mT alternating field (AF) cleaning on some of the specimens after each in-field step.
+The purpose is to study whether the AF cleaning could help improve experiment success rate by removing the remanence component carried by materials such as multi-domain (MD) grains that contribute to non-ideal paleointensity behaviors.
 All remanence measurements were made on a 2G Enterprises DC-SQUID superconducting rock magnetometer equipped with an automated sample changer system at the UC Berkeley Paleomagnetism laboratory.
@@
 Samples were placed in the same location within the thermal demagnetizer for each heating step and were maintained in the same orientation with regard to the applied field.
-During each heating step, samples remained at peak temperatures for 20 min.
+During each heating step, the oven remained at peak temperatures for 20 min to make sure each specimen reached the target temperature.
 An applied laboratory field of 30 $\mu$T was used for all in-field steps.
 All heating steps were performed in air.
-The temperature increments for the experiments were chosen to cover characteristic remanent magnetizations held by (titano)magnetite informed by the previous demagnetization data, with smaller increments performed close to the expected unblocking temperature of stoichiometric magnetite \cite{Zhang2021b}.
+The temperature increments for the experiments were chosen to isolate magnetizations held by (titano)magnetite informed by the previous demagnetization data, with smaller increments performed close to the expected unblocking temperature of $\sim$580\textdegreeC \cite{Zhang2021b}.
 
@@
 Backfield curves were measured at room temperature using a Micromag Princeton Measurements vibrating sample magnetometer (VSM) and a Lake Shore 8600 series VSM.
-Specimen median destructive field (MDF) are calculated based on the backfield curves.
-The calculated coercivity spectra were subsequently decomposed into one ore more components using skew-normal distributions following the method of ref.
+Specimen median destructive fields (MDF) are calculated based on the backfield curves.
+The calculated coercivity spectra were subsequently decomposed into one or more components using skew-normal distributions following the method of ref.
 \citealp{Maxbauer2016a}.
 The fitted components are then used to interpret the characteristics of different populations of magnetic particles within specimens.
-We also used the magnetic property measurement system (MPMS) at the IRM to aid in the identification of magnetic mineral.
+We also used the magnetic property measurement system (MPMS) at the IRM to aid in the identification of magnetic minerals.
 In the field-cooled (FC) experiments, specimen magnetizations were measured upon warming following the specimen having cooled in an applied field of 2.5 T from 300 to 10 K.
@@
 \ref{fig:Petro_QDM}).
-We use the QDM to image the magnetic field over the polished thin section surfaces at a sample-sensor distance of 5 $\mu$m in projective magnetic microscopy (PMM) mode with a spatial resolution of 4.7 μm per pixel and an instantaneous 0.9 mT bias field that is canceled during the course of measurement \cite{Glenn2017a}).
+We use the QDM to image the magnetic field over the polished thin section surfaces with a sample-sensor distance of 5 $\mu$m in projective magnetic microscopy (PMM) mode with a spatial resolution of 4.7 μm per pixel and an instantaneous 0.9 mT bias field that is canceled during the course of measurement \cite{Glenn2017a}.
 }

### 2021-12-20 PNAS_manuscript/PNAS_manuscript.tex

@@
 \citealp{Cych2021a} which uses all experimental data regardless of their NRM/TRM plot statistics to perform bias corrected estimation of paleointensity (BiCEP).
-This Bayesian probabilistic method is based on an assumption that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Arai1963a, Paterson2011a}.
-The posterior paleointensity distributions for sites with high-quality specimen-level data (Fig.
-\ref{fig:PINT_cooling_corrected}) are in agreement with site-level weighted-mean results (Fig.
-S3).
+This Bayesian probabilistic method is based on an assumption that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Paterson2011a}.
+The posterior paleointensity distributions for sites with high-quality specimen-level data are in agreement with site-level weighted-mean results (Fig.
+\ref{fig:PINT_cooling_corrected}; Fig.
+S1).
 
@@
 Magnetic property measurement system (MPMS) data show both diabase and anorthosite have dominant magnetic mineralogy of magnetite as evidenced through the Verwey transition (Fig.
-S1; ref.
+S2; ref.
 \citealp{Verwey1939a}).
@@
 However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have suppressed Verwey transition toward lower temperatures (Fig.
-S1), indicating that the specimens likely have a magnetic mineralogy with variable degrees of oxidation or relatively high Ti content \cite{Ozdemir1993a}.
+S2), indicating that the specimens likely have a magnetic mineralogy with variable degrees of oxidation or relatively high Ti content \cite{Ozdemir1993a}.
 
@@
 \citealp{Halgedahl1980a} (Fig.
-S2).
+S3).
 This estimate on cooling rate effect is similar to that derived from the model of ref.
@@
 Although one may consider these data capturing relatively short term fluctuations of the geomagnetic field, they nevertheless indicate that the geodynamo had strong power sources during the Proterozoic.
-On the other hand, besides the low paleointensity values recorded by the \textit{ca.} 1140 Ma Abitibi dikes, low estimates have been obtained from the \textit{ca.} 1070 Ma Bangemall Sills \cite{Lloyd2021b}, the \textit{ca.} 720 Ma dolerite dikes of the Franklin Large Igneous Province \cite{Lloyd2021a}, as well as Devonian-aged rocks from the Minusa Basin, Kola Peninsula, and Buribay volcanic complex (Fig.
+On the other hand, besides the low paleointensity values recorded by the \textit{ca.} 1140 Ma Abitibi dikes, low estimates have been obtained from the \textit{ca.} 1070 Ma Bangemall Sills \cite{Lloyd2021b}, the \textit{ca.} 720 Ma dolerite dikes of the Franklin Large Igneous Province \cite{Lloyd2021a}, as well as Devonian-aged rocks from the Minusa Basin, Kola Peninsula, Buribay volcanic complex and volcanic rocks from the East European Craton (Fig.
 \ref{fig:PINT_compilation}; ref.
-\citealp{Shcherbakova2017a, Shcherbakova2021a}).
+\citealp{Shcherbakova2017a, Shcherbakova2019a, Shcherbakova2021a}).
 The multiple occurrences of low geomagnetic field intensities during the Proterozoic and the Paleozoic indicate that a waning thermal convection regime in Earth's before the emergence of the solid inner core may not be the solely possible mechanism for triggering an weak dipole field.
@@
 All heating steps were performed in air.
-The temperature increments for the experiments were chosen to isolate magnetizations held by (titano)magnetite informed by the previous demagnetization data, with smaller increments performed close to the expected unblocking temperature of $\sim$580\textdegreeC \cite{Zhang2021b}.
+The temperature increments for the experiments were chosen to isolate magnetizations held by (titano)magnetite informed by the previous demagnetization data, with smaller increments performed close to the expected unblocking temperature of $\sim$580\textdegree C \cite{Zhang2021b}.
 

### 2021-12-20 PNAS_manuscript/Supporting Info.tex
update manuscript

@@
 
+\begin{figure*}[h!]
+\noindent\includegraphics[width=17.8 cm]{manuscript/PINT_BiCEP.pdf}
+\centering
+\caption{{Violin plots of site-level posterior paleointensity distributions estimated using the bias corrected estimation of paleointensity (BiCEP) method developed by ref.
+\ref{Cych2021a}.
+Assuming that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Arai1963a, Paterson2011a}, the method uses all paleointensity measurement-level data without applying selection criteria.
+For comparison of results from this independent method with those based on our selection (as show in Fig.
+4 in manuscript), we highlight the anorthosite sites that pass our paleointensity selection and make other anorthosite and diabase transparent.
+The results from the BiCEP method address the uncertainties associated with anorthosite AX6 and AX8.
+This is associated with the relatively variable specimen behaviors within these two sites.
+But for sites AX10, AX11, AX13, AX12, and AX16, the posterior probability distributions have very narrow bounds, consistent with the interpretation that these anorthosites are faithful paleointensity recorders that have high-quality paleointensity behaviors.
+}}
+\label{fig:PINT_BiCEP}
+\end{figure*}
+
+\clearpage
+
 \begin{figure}
@@
 \begin{figure*}[h!]
-\noindent\includegraphics[width=17.8 cm]{manuscript/Cooling_rate_correction.pdf}
+\noindent\includegraphics[width=17.8 cm]{Cooling_rate_correction.pdf}
 \centering
@@
 \clearpage
+
+% \begin{figure*}[h!]
+% \noindent\includegraphics[width=17.8 cm]{manuscript/nonlinear.pdf}
+% \centering
+% \caption{{}}
+% \label{fig:nonlinear_check}
+% \end{figure*}
+
+% \clearpage
 

### 2022-01-13 PNAS_manuscript/PNAS_manuscript.tex
edit through Cooling rate correction and secular variations section

@@
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{Acquiring high-fidelity observational records of ancient magnetic field intensity from remanent magnetization of rocks is crucial for constraining the evolution of Earth's core.
-However, robust estimates of ancient field strengths are often difficult to recover due to alterations or non-ideal rock magnetic behaviors.
-In this study, we use plagioclase cumulates that formed in the deep crust and were brought to the near surface by magma where they cooled and acquired thermal remanent magnetizations.
-These plagioclase from the North American Midcontinent Rift have experienced minimal post-formation alterations and record a present-day-like virtual dipole moment 1.1 billion years ago.
-The results indicate the existence of strong power sources for Earth's dynamo during the late Mesoproterozoic.
+\significancestatement{Acquiring high-fidelity observational records of ancient magnetic field intensity from the remanent magnetization of rocks is crucial for constraining the evolution of Earth's core.
+However, robust estimates of ancient field strengths are often difficult to recover due to alteration or non-ideal rock magnetic behavior.
+In this study, we use plagioclase cumulates (anorthosite) that formed in the deep crust and were brought to the near surface by magma where they cooled and acquired thermal remanent magnetizations.
+These anorthosite xenoliths from the North American Midcontinent Rift have experienced minimal alteration since their emplacement and yield high quality paleointensity estimates.
+These data record a strong present-day-like virtual dipole moment 1.1 billion years ago.
+A strong field persisted over a 25 million year or greater interval indicating the existence of appreciable power sources for Earth's dynamo during the late Mesoproterozoic.
 }
@@
 Obtaining estimates of Earth's magnetic field strength in deep time is complicated by non-ideal rock magnetic behavior in many igneous rocks.
-In this study, we target \textit{ca.} 1092 Ma anorthosite xenoliths that cooled and acquired their magnetization within Beaver River diabase intrusions of the North American Midcontinent Rift.
-In contrast to the diabase which has non-ideal paleointensity behaviors and fails to provide reliable paleointensity estimates, the anorthosite xenoliths are unusually high-fidelity recorders yielding high-quality, single-slope paleointensity results that are consistent at specimen and site levels.
+In this study, we target anorthosite xenoliths that cooled and acquired their magnetization within \textit{ca.} 1092 Ma shallowly emplaced diabase intrusions of the North American Midcontinent Rift.
+In contrast to the diabase which fails to provide reliable paleointensity estimates, the anorthosite xenoliths are unusually high-fidelity recorders yielding high-quality, single-slope paleointensity results that are consistent at specimen and site levels.
 An average value of $\sim$89 ZAm$^2$ for the virtual dipole moment from the anorthosites is higher than that of the dipole component of Earth's magnetic field today.
-Such a strong late Mesoproterozoic dipole recorded by the anorthosites requires the existence of a vigorous geodynamo at the time.
-Together with previous paleointensity data from other Midcontinent Rift rocks, our new results indicate that a dynamo with strong power sources persisted through $\sim$25 Myr at \textit{ca.} 1.1 Ga.
-These observations are inconsistent with there being a progressive monotonic decay of Earth's dynamo strength through the Proterozoic Eon.
+Such a strong late Mesoproterozoic dipole recorded by the anorthosite xenoliths requires the existence of a vigorous geodynamo at the time.
+Together with previous paleointensity data from other Midcontinent Rift rocks, our new results indicate that a dynamo with strong power sources persisted for at least $\sim$25 million years \textit{ca.} 1.1 Ga.
+These data are inconsistent with there being a progressive monotonic decay of Earth's dynamo strength through the Proterozoic Eon.
 \end{abstract}
@@
 Paleomagnetic records from ancient rocks are one of the few types of observational data that have the potential to record the thermal evolution of Earth’s core.
-Evidence for a persistent magnetic field through the Proterozoic, for example, necessitates the existence of plate tectonics that sustained core-mantle boundary heat flow.
-However, strikingly low estimates of geomagnetic field strengths have been obtained in the Ediacaran Period \cite{Bono2019a, Thallner2021b}, potentially indicating an unusual period of core dynamo regime at the time.
+Evidence for a persistent magnetic field through the Proterozoic, for example, necessitates the existence of plate tectonics that sustained core-mantle boundary heat flow \cite{Swanson-Hysell2021c}.
+However, strikingly low estimates of geomagnetic field strengths have been obtained in the Ediacaran Period \cite{Bono2019a, Shcherbakova2019a,Thallner2021b} and the Devonian Period \cite{Shcherbakova2017a, Shcherbakova2021a}, potentially indicating unusual periods of core dynamo regime at those times.
+% These data have been interpreted to indicate that there was a progressively decaying field up to that time that was followed soon after by initial crystallization of the inner core \cite{Bono2019a}.
 
-% Those paleointensity data has been interpreted to be consistent with thermal evolution models that use high core thermal conductivity values, where the nucleation of the inner core did not begin to provide compositional convection power for driving the geodynamo until after \textit{ca.} 565 Ma, suggesting that the Proterozoic dynamo might have experienced a monotonic decay when compositional convection was lacking and thermal convection was progressively decreasing \cite{Driscoll2016a, Davies2021a}.
+%\begin{figure*}
+\begin{SCfigure*}
+\centering
+\noindent\includegraphics[width=11.4 cm]{Geologic_map.pdf}
+\caption{\footnotesize{Simplified geologic map of the Lake Superior region showing the distribution of rift and post-rift-related rocks including extrusive, intrusive, and late/post-rift sediments around Lake Superior.
+Purple stars mark sites with paleointensity results that passed the selection criteria from this study.
+Paleomagnetic sites from ref.
+\citealp{Pesonen1983a} are categorized by lithology.
+All sites from ref.
+\citealp{Pesonen1983a, Kulakov2013a, Sprain2018a} are color-coded by their ages.}}
+\label{fig:Geologic_map}
+%\end{figure*}
+\end{SCfigure*}
 
-Determinations of ancient magnetic field strengths rely on igneous rocks that acquire thermal remanent magnetizations as they cool in Earth's magnetic field.
-These magnetizations need to be unmodified by subsequent heating or chemical alterations in order to be able to faithfully record ancient geomagnetic field through time.
-Intraplate magmatic events are therefore an important target as they can be well-preserved within continental interiors at distance from mountain-building events along continental margins.
-However, magmatic events such as the formation of large igneous provinces that are often associated with intraplate magmatism are typically of geologically short duration with the bulk of magmatic products emplaced within 1 Myr.
-The Midcontinent Rift is an exception as magmatism lasted $\sim$25 Myr from 1109 Ma to 1084 Ma.
-Additionally, extension ceased in the Midcontinent Rift prior to lithosphere separation, preserving volcanics, intrusives, and sediments of the rift within the continental interior.
-As a result, rocks of the rift have simple paleomagnetic behavior and form a central record of Mesoproterozoic paleogeography \cite{Swanson-Hysell2021c}.
-The simple magnetizations held by igneous rocks that formed through an extended duration from 1109 to 1084 Ma that is longer than the Neogene Period have the potential for a well-preserved window into the intensity of Earth's magnetic field .
+Determinations of the absolute value of the the ancient geomagnetic field strength rely on igneous rocks that acquire thermal remanent magnetizations as they cool.
+These magnetizations need to be unmodified by subsequent heating or chemical alteration in order to maintain a record of the ancient geomagnetic field from the time of cooling.
+Intracontinental magmatic events are therefore an important target for determination of ancient paleointensity as they can be well-preserved within continental interiors.
+This interior position results in them typically being distant from tectonic events along continental margins that can drive alteration through heat and fluids.
+However, large igneous provinces that are often associated with intraplate magmatism are typically of geologically short duration with the bulk of magmatic products emplaced within 1 Myr or less \cite{Kasbohm2021a}.
+The Midcontinent Rift is an exception as it is a large igneous province where magmatism lasted $\sim$25 Myr from \texit{ca.} 1109 Ma to 1084 Ma \cite{Swanson-Hysell2019a}.
+Additionally, extension ceased in the Midcontinent Rift prior to lithospheric separation, preserving volcanic, intrusive, and sedimentary rocks of the rift within the continental interior.
+As a result, rocks of the rift have unusually simple paleomagnetic behavior for their greater than one billion-year-old age and form a central record of Mesoproterozoic paleogeography \cite{Swanson-Hysell2021c}.
+The duration of magmatic activity within the Midcontinent Rift is longer than the entire 20.4 Myr long Neogene Period such that it enables an extended well-preserved window into the intensity of Earth's magnetic field in the late Mesoproterozoic.
 
-Despite the exquisite preservation of the rocks, non-ideal paleointensity behaviors and overall low success rates have often led to high failure rates and challenged the interpretation of many previous paleointensity results from the Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a}.
-One typical failure of paleointensity experiment is the double-slope behavior, as two distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
-Ref.
+Despite the excellent preservation of the rocks, non-ideal paleointensity behaviors and overall low success rates have often led to high failure rates and challenged the interpretation of many previous paleointensity results from the Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a}.
+The most trusted type of paleointensity estimate is that obtained through experiments in which the primary natural thermal remanent magnetization is progressively replaced by a laboratory magnetization that is imparted in a known field with internal consistency checks (such as in the IZZI-style Thellier experiments; ref.
+\citealp{Yu2004a}).
+In such Thellier paleointensity experiments, one typical departure from ideal behavior is sagging or double-slopes as visualized in plots that show thermal remanent magnetization (TRM) acquired versus natural remanent magnetization (NRM) lost.
+For such data, distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope.
+Typically, such non-ideal behavior would result in a higher paleointensity estimate from the steeper-sloped low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
+For example, in data from the Midcontinent Rift, ref.
 \citealp{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas ref.
@@
 Results with such non-ideal behavior were rejected by ref.
-\citealp{Sprain2018a} who applied stricter paleointensity selection criteria.
+\citealp{Sprain2018a} who applied stricter paleointensity selection criteria, but as a result had few accepted sites.
 
-In this study, we target a unique rock type - high-purity anorthosite xenoliths.
-These anorthosite xenoliths were brought to the near surface with magma that formed sills of the Beaver River diabase \cite{Zhang2021b}.
+In this study, we target a unique rock type---high-purity anorthosite xenoliths.
+Anorthosites are cumulate rocks composed dominantly of the mineral plagioclase.
+They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud minute magnetic inclusions from alteration with alteration of the plagioclase crystals not resulting in the formation of secondary iron oxides.
+The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal intrusions of the Beaver River diabase \cite{Zhang2021b}.
 They cooled and acquired their magnetizations in conjunction with the host diabase.
 Paleointensity experiments on the anorthosite xenoliths have a very high success rate, yielding consistent specimen- and site-level paleointensity results.
-Rock magnetic data reveal that the anorthosite xenoliths have low anisotropy of thermal remanent magnetization (TRM) and can acquire TRM linearly.
+Rock magnetic data reveal that the anorthosite xenoliths have low anisotropy of thermal remanent magnetization (TRM) and can acquire TRM linearly within relevant field strengths.
 Magnetic imaging shows the anorthosite specimens have dominant magnetic carriers interstitial to and within plagioclase crystals and they do not display strong preferred orientations.
-Our single-slope, high quality paleointensity data add to the reliable paleointensity records during the $\sim$25 Myr of Midcontinent Rift magmatic activity.
-Our new data from the Midcontinent Rift intrusive rocks support previous data from the volcanics that there was a strong late Mesoproterozoic geodynamo.
+Our single-slope, high quality paleointensity data add to the reliable paleointensity records during $\sim$25 Myr of Midcontinent Rift magmatic activity and require that there was a strong late Mesoproterozoic geodynamo.
 
@@
 
-While magmatic activity within the Midcontinent Rift was protracted, there were volcanism and intrusions characterized by particularly rapid and voluminous emplacement.
-The \textit{ca.} 1092 Ma Beaver Bay Complex in northern Minnesota punctuates one such period of magmatism during the Midcontinent Rift main stage activity.
+While magmatic activity within the Midcontinent Rift was protracted, there were intervals of particularly rapid volcanism and voluminous emplacement of intrusions.
+The \textit{ca.} 1092 Ma Beaver Bay Complex in northern Minnesota punctuates one such period of magmatism during the main stage of Midcontinent Rift activity.
 The magma that formed the 1091.7 $\pm$ 0.2 Ma Beaver River diabase dikes and sills of the Beaver Bay Complex transported numerous anorthosite xenoliths that have short-axis diameters up to 180 meters via wide conduits \cite{Boerboom2004a, Boerboom2006b}.
-The anorthosite xenoliths are plagioclase cumulates that formed comagmatically with the host diabase in the lower crust - an interpretation confirmed by U-Pb zircon geochronology \cite{Zhang2021b}.
+These anorthosite xenoliths are plagioclase cumulates that formed comagmatically with the host diabase in the lower crust---an interpretation confirmed by U-Pb zircon geochronology \cite{Zhang2021b}.
 
@@
 Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths to have acquired thermal remanent magnetizations while cooling with the Beaver River diabase \cite{Zhang2021b}.
-Step-wise thermal demagnetizations show the anorthosite xenoliths have dominantly single-component magnetization that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C \cite{Zhang2021b}.
+Step-wise thermal demagnetization data show the anorthosite xenoliths have dominantly single-component magnetization that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C, consistent with them being held by low titanium titanomagnetite \cite{Zhang2021b}.
 
-\section*{Petrography and magnetic imaging of anorthosite xenoliths}
+%\begin{figure*}
+\begin{SCfigure*}
+\centering
+\noindent\includegraphics[width=10.5 cm]{Petro_QDM.png}
+\caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from the Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (left column; ref.
+\citealp{Zhang2021b}) and distinct anorthosite from the Duluth Complex Anorthositic Series (right column).
+The Duluth Complex anorthosites were not targeted for paleointensity experiments in this study given the complexities associated with more pronounced fabrics.
+Cross-polarized petrographic images (A) Beaver River anorthosite have plagioclase with a granoblastic texture; (B) Plagioclase of Duluth Complex anorthosite have euhedral, interlocking crystals showing igneous foliation; (C) zoomed-in image of plagioclase from MS99033 showing the crystals to be largely free of large opaque inclusions; (D) Duluth Complex plagioclase can contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the [010] planes of the plagioclase crystals.
+Magnetic field maps acquired by the quantum diamond microscope (QDM) show that the relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite in (E) while large oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the overlapping magnetic fields above the oxides in (F).
+Magnetic field maps of (E) and (F) share the same spatial and magnetic scale.
+The unit of the axes are in $\mu$m.}}
+\label{fig:Petro_QDM}
+%\end{figure*}
+\end{SCfigure*}
 
-Petrographic imaging shows the dominantly monomineralic anorthosite xenoliths of the Beaver River diabase often develop granoblastic texture characterized by equigranular crystals with poor development of petrofabrics (Fig.
+\section*{Results and Discussion}
+
+\subsection*{Petrography and magnetic imaging of anorthosite xenoliths}
+
+Petrographic imaging shows the dominantly monomineralic anorthosite xenoliths within the Beaver River diabase often have granoblastic texture characterized by equigranular crystals with weakly developed petrofabrics (Fig.
 \ref{fig:Petro_QDM}A,C).
-In addition, the plagioclase crystals of the anorthosite xenoliths are often free of large inclusions.
-Magnetic imaging using the quantum diamond microscope (QDM) shows that the dominant magnetic remanence carriers often exist within and sometimes interstitial to plagioclase crystals (Fig.
+In addition, the plagioclase crystals of the anorthosite xenoliths are typically free of large oxide inclusions.
+Magnetic imaging using the quantum diamond microscope (QDM) shows that the dominant magnetic remanence carriers often exist within, and sometimes interstitial to, plagioclase crystals (Fig.
 \ref{fig:Petro_QDM}E).
 
-These features of the Beaver River anorthosite xenoliths are distinct from other anorthosites such as the anorthosite xenoliths within the Duluth Complex Anorthositic Series rocks.
-The plagioclase of the Duluth Complex anorthosites typically develop interlocking textures that display strong igneous foliations (Fig.
-\ref{fig:Petro_QDM}B) and associated titanomagnetite mineral fabrics characterized by having aligned needles within plagioclase crystals (Fig.
+These features of the Beaver River anorthosite xenoliths are distinct from other anorthosites such as the anorthosite xenoliths within the Duluth Complex Anorthositic Series rocks---older intrusions within the Midcontinent Rift.
+To illustrate these difference, we present petrographic and magnetic imaging data from a sample of Duluth Complex anorthosite that was not targeted for paleointensity experiments.
+The plagioclase of the Duluth Complex anorthosites typically develop interlocking textures that display strong igneous foliation (Fig.
+\ref{fig:Petro_QDM}B).
+In addition, there are abundant Fe-Ti oxide needles that are typically aligned with the [010] planes of plagioclase crystals (Fig.
 \ref{fig:Petro_QDM}D).
-Magnetic imaging confirms that the titanomagnetite needles within the plagioclase crystals are often strongly magnetic with magnetic moments of individual needles oriented along their long axes (Fig.
+Magnetic imaging confirms that those strongly magnetic needles within the plagioclase crystals often have magnetic moments oriented along their long axes (Fig.
 \ref{fig:Petro_QDM}D,F).
@@
 
-The lack of symplecites and exsolved titanomagnetite needles within the Beaver River anorthosite xenoliths distinguish them from plagioclase cumulates of the Duluth Complex and other layered mafic complex where anisotropic magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
+The lack of symplecites and the lack of large plagioclase-hosted titanomagnetite needles within the Beaver River anorthosite xenoliths distinguish them from plagioclase cumulates of the Duluth Complex and other layered mafic complex where anisotropic magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
+The relative lack of fabric and associated preferred orientation makes the Beaver River anorthosite xenoliths a particularly compelling target for paleointensity experiments.
 
@@
 
-\begin{figure*}
-\centering
-\noindent\includegraphics[width=11.4 cm]{Geologic_map.pdf}
-\caption{\footnotesize{Simplified geologic map of Lake Superior region showing the distribution of rift and post-rift-related rocks including extrusive, intrusive, and late/post-rift sediments around Lake Superior.
-Red diamonds mark sites with paleointensity results that passed the selection criteria from this study.
-Paleomagnetic sites from ref.
-\citealp{Pesonen1983a} are categorized by lithology.
-All sites from ref.
-\citealp{Pesonen1983a, Kulakov2013a, Sprain2018a} are color-coded by their ages.}}
-\label{fig:Geologic_map}
-\end{figure*}
-
 % Despite that the anorthosite xenoliths of the Beaver River diabase were entrained as xenoliths and emplaced only $\sim$4 Myr after the formation of the Anorthositic Series rocks of the Duluth Complex when massive plagioclase mushes were generated, $^{206}$Pb/$^{238}$U zircon geochronology data and zircon rare earth element diffusion modeling results suggest it is the most probable that the Beaver River diabase anorthosite xenoliths are entrained cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
@@
 While the plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture (Fig.
-\ref{Petro_QDM}A), the more calcium-rich plagioclase of the Beaver River anorthosite often develop granoblastic texture characterized by equigranular plagioclase crystals (Fig.
+\ref{Petro_QDM}A), the more calcium-rich plagioclase of the Beaver River anorthosite often develop granoblastic texsture characterized by equigranular plagioclase crystals (Fig.
 \ref{Petro_QDM}B).
@@
 
-\begin{figure*}
-\centering
-\noindent\includegraphics[width=11.4 cm]{Petro_QDM.png}
-\caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from the Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (left column; ref.
-\citealp{Zhang2021b}) and and distinct anorthosite from the Duluth Complex Anorthositic Series (right column).
-Cross-polarized petrographic images (A) Beaver River anorthosite have plagioclase with a granoblastic texture; (B) Plagioclase of Duluth Complex anorthosite have euhedral, interlocking crystals showing igneous foliation; (C) zoomed-in image of plagioclase from MS99033 showing the crystals are largely free of large opaque inclusions; (D) Duluth Complex plagioclase can contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the (010) planes of plagioclase.
-Magnetic field maps acquired by the quantum diamond microscope (QDM) show that the relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite in (E) while large oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the complex overlapping magnetic fields above the oxides in (F).
-Magnetic field maps of (E) and (F) share the same spatial and magnetic scale.
-The unit of the axes are in $\mu$m.}}
-\label{fig:Petro_QDM}
-\end{figure*}
-
-\section*{Results and Discussion}
-
 \subsection*{Paleointensity}
 
-Following IZZI-modified paleointensity experiments \cite{Yu2004a}, 40 from a total of 86 anorthosite specimens and 0 out of a total of 69 diabase specimens passed our paleointensity result selection criteria of having (1) a maximum angular deviation (MAD; ref.
-\citealp{Kirschvink1980a}) of $<$20\textdegree; (2) scatter parameter ($\beta$; ref.
-\citealp{Coe1978a}) values of $<$15 per cent; (3) a deviation angle (DANG; ref.
-\citealp{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC; ref.
-\cite{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT; ref.
-\citealp{Shaar2013a}) = TRUE; (6) a maximum magnetic moment difference between adjacent zero-field steps (GAP-Max; ref.
-\citealp{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
-7 anorthosite sites and no diabase site-level results pass our selections.
-Example NRM/TRM plots are shown in Fig.
+Following IZZI-modified paleointensity experiments \cite{Yu2004a}, 40 from a total of 86 anorthosite specimens and 0 out of a total of 69 diabase specimens passed our paleointensity result selection criteria (see Materials and Methods section).
+7 anorthosite sites and no diabase site-level results pass these selection criteria.
+Example NRM/TRM (Arai) plots are shown in Fig.
 \ref{fig:IZZI_examples}.
@@
 \ref{fig:PINT_cooling_corrected} (and provided in Table S1).
+The paleointensity quality index (Q$_{PI}$; ref.
+\citealp{Biggin2014a}) for the anorthosite xenoliths are all above 4 (Table S2).
 The cooling rate-corrected absolute paleointensity estimates from all anorthosite sites have a weighted mean of 52.72 $\pm $5.72 $\mu$T.
 This corresponds to a calculated virtual dipole moment of $\sim$89 ZAm$^2$ ($10^{21} Am^2$) \textit{ca.} 1092 Ma.
-All measurement-level paleointensity experiment data is available within the MagIC database (\url{}; UPDATE WHEN DOI IS GENERATED).
+All measurement-level paleointensity experiment data is available within the MagIC database (\url{}; ATTACHED IS A LINK FOR REVIEWING PURPOSESS, THE URL WILL BE UPDATED WHEN DOI IS GENERATED).
 
@@
 \centering
-\caption{\footnotesize{Example results of paleointensity experiments displayed on NRM/TRM plots and zero-field heating results are shown on orthogonal plots (insets) for anorthosite and diabase specimens.
+\caption{\footnotesize{Example results of paleointensity experiments are displayed on NRM/TRM plots and zero-field heating results are shown on inset orthogonal plots (Zijderveld plots) for anorthosite and diabase specimens.
 Red (blue) circles indicate zero-field/in-field (in-field/zero-field) steps `ZI’ (`IZ’).
@@
 Plots on the top row show successful specimen paleointensity results with straight, single-slope behaviors that pass our selection criteria.
-The green lines represent our fitting for the dominant single-slope component that passes our acceptance criteria with estimates of the ancient field (B$_{anc}$) being used to determine site means.
+The green lines represent fits for the dominant single-slope component that passes the acceptance criteria and gives an estimate of the ancient field strength (B$_{anc}$).
 The plots for anorthosite specimens AX1-1a and diabase BD5-1b on the bottom row show typical non-ideal sagging and zigzagging behaviour that fails our acceptance criteria.
-Plot for diabase specimen BD2-5a shows typical paleointensity behavior of this site - specimens typically have double-slope behaviors although the lower temperature slopes can pass the selection criteria, resulting in the high estimated B$_{anc}$ values shown in Fig.
-\ref{fig:PINT_cooling_corrected}.
-Thus, the estimates of B$_{anc}$ for this specimen plot is illustrative, but were not used to calculate a site mean paleointensity value in Fig.
-\ref{fig:PINT_cooling_corrected}.}}
+Specimen BD2-5a is an example where the data appear linear with distinct slopes in the low and high temperature ranges such that it could pass less restrictive selection criteria, particularly if a more narrow temperature range was used for the experiment.}}
 \label{fig:IZZI_examples}
@@
 
-Paleointensity experiment results from the anorthosite xenoliths and their host diabase show three general types of dominant specimen behaviors on NRM/TRM plots (Fig.
+% Paleointensity experiment results from the anorthosite xenoliths and their host diabase show three general types of dominant specimen behaviors on NRM/TRM plots (Fig.
 \ref{fig:IZZI_examples}): dominantly single-slope, double-slope or sagging, and zigzagging.
-Typical specimen paleointensity behavior of the anorthosites display straight, single-slope NRM/TRM plots and the accepted fractions of temperature steps span over the origin-trending, primary remanence components (Fig.
+Typical paleointensity experimental data of the anorthosite specimens have straight, single-slope NRM/TRM plots and the accepted fractions of temperature steps span over the origin-trending, primary remanence components (Fig.
 \ref{fig:IZZI_examples}).
-We accept all specimen- and site-level absolute paleointensity results from the anorthosite xenoliths that pass the selection.
-Diabase specimens failed the selection because of double-slope or sagging behavior (fail FRAC selection; see methods), poor pTRM checks, and zigzagging plots (fail SCAT, DANG selection; see methods; Fig.
+We accept specimen- and site-level absolute paleointensity results from those anorthosite xenoliths that pass the selection criteria.
+Other anorthosite xenoliths and diabase specimens failed the selection criteria largely because of double-slope or sagging behavior (fail FRAC selection; see methods), poor pTRM checks, and sometimes zigzagging behaviors superimposed on top (fail SCAT, DANG selection; see methods; Fig.
 \ref{fig:IZZI_examples}).
-A 20 mT alternating field treatment after in-field heating steps was applied to some specimens but it did not result in significant changes in paleointensity behaviors of the anorthosite xenoliths or the diabase (Fig.
+A 20 mT alternating field treatment after in-field heating steps was applied to some specimens, but it did not result in significant changes in the experimental results for the anorthosite xenolith or diabase specimens (Fig.
 \ref{fig:PINT_cooling_corrected}; Supporting Information).
 
-In addition to estimating paleointensities by introducing a set of selection criteria to filter our experiment results, we apply an independent method of ref.
+\begin{figure*}[h!]
+\noindent\includegraphics[width=17.8 cm]{Paleointensity_plot_cooling_corrected.pdf}
+\centering
+\caption{\small{Summary plot of individual specimen absolute paleointensity results (square and diamond symbols) from this study and their averages at site (grey bar) and locality level (weighted mean by site-level standard deviations; orange bar).
+All results are corrected for cooling rate with a factor of 0.74.
+}}
+\label{fig:PINT_cooling_corrected}
+\end{figure*}
+
+In addition to estimating paleointensity values by introducing a set of selection criteria to filter our experiment results, we apply an independent method of ref.
 \citealp{Cych2021a} which uses all experimental data regardless of their NRM/TRM plot statistics to perform bias corrected estimation of paleointensity (BiCEP).
 This Bayesian probabilistic method is based on an assumption that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Paterson2011a}.
-The posterior paleointensity distributions for sites with high-quality specimen-level data are in agreement with site-level weighted-mean results (Fig.
+The posterior paleointensity distributions from these sites with high-quality specimen-level data are in agreement with the site-level weighted-mean results developed using the selection criteria approach (Fig.
 \ref{fig:PINT_cooling_corrected}; Fig.
@@
 Overall, the high-quality experiment results from the anorthosite xenoliths of the Beaver River diabase and the consistent paleointensity estimates based on independent methods confirm that the anorthosite xenoliths record a high geomagnetic field \textit{ca.} 1092 Ma.
-The paleointensity quality index (Q$_{PI}$; ref.
-\citealp{Biggin2014a}) for the anorthosite xenoliths are all above 4 (Table S2).
 
@@
 % \end{figure}
-
-\begin{figure*}[h!]
-\noindent\includegraphics[width=17.8 cm]{Paleointensity_plot_cooling_corrected.pdf}
-\centering
-\caption{\small{Summary plot of individual specimen absolute paleointensity results (square and diamond symbols) from this study and their averages at site (grey bar) and locality level (weighted mean by site-level standard deviations; orange bar).
-All results are corrected for cooling rate with a factor of 0.74.
-}}
-\label{fig:PINT_cooling_corrected}
-\end{figure*}
 
@@
 Rock magnetic data support that the anorthosite specimens that pass the paleointensity selection criteria have dominant magnetic remanence carriers with magnetic properties similar to stochiometric, non-interacting, single domain magnetite, whereas anorthosite samples that failed the paleointensity result selection and all diabase samples have more pronounced populations of non-ideal carriers.
-Magnetic property measurement system (MPMS) data show both diabase and anorthosite have dominant magnetic mineralogy of magnetite as evidenced through the Verwey transition (Fig.
+Magnetic property measurement system (MPMS) data show both diabase and anorthosite contain (titano)magnetite as evidenced through the presence of the Verwey transition (Fig.
 S2; ref.
-\citealp{Verwey1939a}).
-More, anorthosite chips from sites that yield successful paleointensity results have Verwey transition temperatures near 120 Kelvin (K) as expected for stochiometric low-Ti magnetite \cite{Ozdemir1993a}.
-However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have suppressed Verwey transition toward lower temperatures (Fig.
-S2), indicating that the specimens likely have a magnetic mineralogy with variable degrees of oxidation or relatively high Ti content \cite{Ozdemir1993a}.
+\citealp{Verwey1939a, Feinberg2015a}).
+Anorthosite chips from sites that yield successful paleointensity results have Verwey transition temperatures near 120 Kelvin (K) as expected for stochiometric magnetite with minimal Ti \cite{Ozdemir1993a}.
+However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have Verwey transitions that are suppressed toward lower temperatures (Fig.
+S2), indicating that the specimens either has a relatively higher Ti content or has been partially oxidized \cite{Ozdemir1993a}.
 
 Coercivity spectra from backfield experiments further support the interpretation that rocks that pass or fail the selection criteria have distinct magnetic mineralogy.
-A compilation of median destructive field (MDF) in Fig.
-\ref{fig:coercivity} shows that specimens from anorthosites that pass paleointensity selections have a distinctly higher average MDF range ($\sim$60 mT) than other anorthosites and diabase ($\sim$30 mT).
+A compilation of median destructive field (MDF) values in Fig.
+\ref{fig:coercivity} shows that specimens from anorthosites that pass paleointensity selections have a distinctly higher average MDF values ($\sim$60 mT) than other anorthosites and diabase ($\sim$30 mT).
 Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity around 80 mT (Fig.
@@
 In contrast, other anorthosite and diabase tend to have lower coercivities ($\sim$30 mT).
-This is consistent with the interpretation that they lack a dominance of single domain magnetite grains in their magnetic mineral assemblages and thus have non-ideal paleointensity behaviors.
+This result is consistent with the interpretation that they lack a dominance of single domain magnetite grains in their magnetic mineral assemblages and thus have non-ideal paleointensity behaviors.
 
@@
 \centering
-\caption{\footnotesize{Top: Example coercivity unmixing plots of anorthosite and diabase specimens from sites that pass or fail our paleointensity selection criteria.
+\caption{\footnotesize{Top: Example coercivity spectra of anorthosite and diabase specimens from sites that pass or fail our paleointensity selection criteria.
 Bottom: Box plots of median destructive field (MDF) values for all anorthosite and diabase specimens with single-component coercivity unmixing results.
-Both plots show that anorthosite specimens that pass paleointensity selection criteria have distinctly higher portion of single-domain-like magnetic mineralogy than other anorthosite and diabase.}}
+Both plots show that anorthosite specimens that pass paleointensity selection criteria have higher coercivities consistent with a higher portion of single-domain-like magnetite grains than the other anorthosite specimens and the diabase.}}
 \label{fig:coercivity}
@@
 \subsection*{TRM anisotropy and nonlinear acquisition check}
-Significant remanence anisotropy has been documented to exist within anorthositic rocks that form in layered intrusive complexes \cite{Selkin2000a, Feinberg2006a}.
-Ref.
-\citealp{Selkin2000a} demonstrated that strong remanence anisotropy associated with the igneous foliations developed within an anorthosite sample from the Stillwater Complex can lead to significant overestimate or underestimate of paleointensity values depending on the relative orientations between the external applied field magnetic anisotropy axes of a specimen.
+Significant remanence anisotropy has been documented to exist within certain anorthositic rocks that form in layered intrusive complexes \cite{Selkin2000a, Feinberg2006a}.
+Strong remanence anisotropy associated with the igneous foliation developed within anorthosite from the Stillwater Complex has been shown to lead to significant overestimate or underestimate of paleointensity values depending on the relative orientations between the fabrics and an applied magnetic field \cite{Selkin2000a}.
 To assess whether our paleointensity estimates are biased by remanence anisotropy, we calculated the gamma statistic, which is the angular difference between the last pTRM step of paleointensity experiment and the applied field direction.
-The results show that the anorthosite specimens used for paleointensity experiment have gamma values ranging from $\sim$0.9\textdegree to $\sim$11.9\textdegree, with a median value of 4.2\textdegree (Table S1).
-These gamma values are similar to those of the Midcontinent Rift volcanics \cite{Sprain2018a}.
+The results show that the anorthosite specimens used for paleointensity experiment have low gamma values ranging from $\sim$0.9\textdegree to $\sim$11.9\textdegree, with a median value of 4.2\textdegree (Table S1).
+These gamma values are similar to those of Midcontinent Rift volcanics \cite{Sprain2018a}.
 Therefore, the Beaver River anorthosite xenolith bulk samples do not have a significant remanence anisotropy.
-Paleodirectional data from our anorthosite xenoliths further support that they have minimal remanence anisotropy as their site mean directions closely match those of the Beaver River diabase hosts \cite{Zhang2021b}.
-In addition, no strong magnetic fabric was observed in the anorthosite xenoliths through magnetic imaging (Fig.
+Paleodirectional data from our anorthosite xenoliths further support that they have minimal remanence anisotropy as their site mean directions closely match those of the Beaver River diabase hosts without deviating due to a fabric \cite{Zhang2021b}.
+In addition, no resolvable magnetic fabric was observed in the anorthosite xenoliths through petrographic and magnetic imaging (Fig.
 \ref{fig:Petro_QDM}).
@@
 We applied full TRMs to a set of our anorthosite specimens that pass the paleointensity selection criteria with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axes.
-The results show that our anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions (Fig.
+The results show that our anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions (ref.
+\citealp{Selkin2007a}; Fig.
 S4).
@@
 
-\subsection*{Cooling rate correction}
-Paleointensity estimates from igneous intrusions can also be biased toward higher values due to slow cooling rates \cite{Halgedahl1980a}.
-It has been shown that large differences in cooling rates between acquisition of an NRM versus a laboratory pTRM can result in overestimated paleointensities for SD grains \cite{Dodson1980a, Halgedahl1980a, Nagy2021a}.
+\subsection*{Cooling rate correction and secular variations}
+Paleointensity estimates from slowly cooled igneous intrusions can also be biased toward higher values due to the orders of magnitude difference in natural and laboratory cooling rates \cite{Halgedahl1980a}.
+It has been shown that large differences in cooling rates between acquisition of an NRM in nature versus a pTRM in the lab can result in overestimated paleointensities for SD grains \cite{Dodson1980a, Halgedahl1980a, Nagy2021a}.
 From the thermal history model of ref.
-\citealp{Zhang2021b} we can estimate the duration which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully block in their characteristic natural remanence magnetization ($\sim$500\textdegree C; ref.
+\citealp{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully blocked in their characteristic natural remanence magnetization ($\sim$500\textdegree C; ref.
 \citealp{Zhang2021b}).
 We find the cooling time to be on the order of $\sim$1 kyr, which corresponds to a cooling rate of $\sim 1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
-In contrast, the lab cooling rate is much faster, at $\sim 1.3\times10^{-1}$ $^\circ$C s$^{-1}$, assuming that it takes 1000s in lab condition to cool through the same temperature interval.
+In contrast, the lab cooling rate is much faster, at $\sim 1.3\times10^{-1}$ $^\circ$C s$^{-1}$, assuming that it takes $\sim$1000s in lab condition to cool through the same temperature interval.
 This significant cooling rate difference leads to a predicted $\sim$36\% overestimate of true ancient field following the model of ref.
@@
 S3).
-This estimate on cooling rate effect is similar to that derived from the model of ref.
+This estimate on cooling rate effect is similar to the value of $\sim$35\% overestimate derived from the model of ref.
 \citealp{Nagy2021a}.
@@
 \ref{fig:PINT_cooling_corrected}.
+
+\begin{figure*}[h!]
+\noindent\includegraphics[width=17.8 cm]{PINT_compilation.pdf}
+\centering
+\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment values from the PINT database (PINT v8.0.0; \url{http://www.pintdb.org/}; ref.
+\citealp{Bono2021a}) with additional Mesoproterozoic data from \cite{Sprain2018a, Kodama2019a}, as well as those from the Neoproterozoic \cite{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}.
+Paleointensity estimates from \cite{Pesonen1983a} and \cite{Kulakov2013a} are not included in the compilation due to the specimen-level double-slope behavior as discussed in the text.
+Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \cite{Bono2019a} which is based on an interpretation of there being a progressive decay of the geodynamo through the Proterozoic.}}
+\label{fig:PINT_compilation}
+\end{figure*}
 
@@
 
-Recent advancements in high-pressure, high-temperature experiments for measuring thermal conductivity of iron alloys under conditions similar to Earth's core may shed new lights on the age of the inner core.
+Recent advancements in high-pressure, high-temperature experiments for measuring thermal conductivity of iron alloys under conditions similar to Earth's core may shed new light on the age of the inner core.
 However, large uncertainties still remain regarding the applicability Wiedemann-Franz law of extrapolating thermal conductivity values based on electrical resistivity measurements by ref.
@@
 
-\begin{figure*}[h!]
-\noindent\includegraphics[width=17.8 cm]{PINT_compilation.pdf}
-\centering
-\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment values from the PINT database (PINT v8.0.0; \url{http://www.pintdb.org/}; ref.
-\citealp{Bono2021a}) with additional Mesoproterozoic data from \cite{Sprain2018a, Kodama2019a}, as well as those from the Neoproterozoic \cite{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}.
-Paleointensity estimates from \cite{Pesonen1983a} and \cite{Kulakov2013a} are not included in the compilation due to the specimen-level double-slope behavior as discussed in the text.
-Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \cite{Bono2019a} which is based on an interpretation of there being a progressive decay of the geodynamo through the Proterozoic.}}
-\label{fig:PINT_compilation}
-\end{figure*}
-
 % \section*{Conclusion}
@@
 Sun compass orientations were preferentially used for determining the sample azimuth.
-Sister specimens underwent step-wise alternating field (AF) or thermal demagnetization at the UC Berkeley Paleomagnetism Lab and the results are presented in \cite{Zhang2021b}.
-Based on the anorthosite thermal demagnetization results we selected sites whose unblocking temperature ranges are narrow and near 580\textdegree C for paleointensity experiments.
+Sister specimens underwent step-wise alternating field (AF) or thermal demagnetization at the UC Berkeley Paleomagnetism Lab to isolate paleomagnetic directions (data presented in \citeA{Zhang2021b}).
+Based on the anorthosite thermal demagnetization results, we selected sites whose unblocking temperature ranges are narrow and near 580\textdegree C for paleointensity experiments.
 Beaver River diabase sites with minimal secondary remanence were also selected for paleointensity experiments.
@@
 The purpose is to study whether the AF cleaning could help improve experiment success rate by removing the remanence component carried by materials such as multi-domain (MD) grains that contribute to non-ideal paleointensity behaviors.
+The results were similar when this step was applied without an observed change in experimental success rate.
 All remanence measurements were made on a 2G Enterprises DC-SQUID superconducting rock magnetometer equipped with an automated sample changer system at the UC Berkeley Paleomagnetism laboratory.
@@
 
-\subsection*{Rock magnetic experiment}
+\subsection*{Rock magnetic experiments}
 
-In this study, we conduct rock magnetic experiments with a purpose of gaining magnetic mineralogy insights into the paleointensity results of the anorthosite and diabase.
-Backfield curves were measured at room temperature using a Micromag Princeton Measurements vibrating sample magnetometer (VSM) and a Lake Shore 8600 series VSM.
+We conducted rock magnetic experiments with a purpose of gaining magnetic mineralogy insights into the paleointensity results of the anorthosite and diabase.
+Backfield curves were measured at room temperature using a Micromag Princeton Measurements vibrating sample magnetometer (VSM) and a Lake Shore 8600 series VSM at the Institute for Rock Magnetism.
 Specimen median destructive fields (MDF) are calculated based on the backfield curves.
 The calculated coercivity spectra were subsequently decomposed into one or more components using skew-normal distributions following the method of ref.
-\citealp{Maxbauer2016a}.
-The fitted components are then used to interpret the characteristics of different populations of magnetic particles within specimens.
-We also used the magnetic property measurement system (MPMS) at the IRM to aid in the identification of magnetic minerals.
+\citealp{Maxbauer2016a} examples of which are shown in Fig.
+\ref{fig:coercivity}.
+We also used a magnetic property measurement system (MPMS) at the Institute for Rock Magnetism to aid in the identification of magnetic minerals.
 In the field-cooled (FC) experiments, specimen magnetizations were measured upon warming following the specimen having cooled in an applied field of 2.5 T from 300 to 10 K.
@@
 In the room-temperature saturation isothermal remanence (RTSIRM) experiment, the sample was pulsed with a 2.5 T field at room temperature ($\sim$300 K) and then cooled to 10 K and warmed back to room temperature in a (near-)zero field.
-The magnetic moment transitions at critical temperatures revealed through MPMS experiments are used to identify magnetic minerals such as magnetite \cite{Verwey1939a} within specimens.
+The magnetic moment transitions at critical temperatures revealed through MPMS experiments are used to identify magnetic minerals such as magnetite within specimens \cite{Feinberg2015a}.
 

### 2022-01-14 PNAS_manuscript/PNAS_manuscript.tex

@@
 
-\subsection*{Cooling rate correction and secular variations}
-Paleointensity estimates from slowly cooled igneous intrusions can also be biased toward higher values due to the orders of magnitude difference in natural and laboratory cooling rates \cite{Halgedahl1980a}.
+\subsection*{Averaging secular variations and applying cooling rate correction}
+To best characterize the geomagnetic axial dipole field intensity during a certain time period requires a paleointensity dataset to cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged and the geocentric axial dipole hypothesis can be assumed.
+Thermal modeling results from ref.
+\citealp{Zhang2021b} suggest that the Beaver River anorthosite xenoliths were likely heated to tholeiitic magma temperature and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a few thousand years.
+In addition, anorthosite xenoliths have paleointensity results that are consistent within small regions but vary between localities.
+Anorthosite AX12 and AX16 were emplaced only $\sim$450 meters apart and have very similar paleointensity estimates.
+$\sim$10 km to the north, anorthosite AX11 and AX13 were also emplaced closely at a distance of $\sim$125 meters but yielding lower paleointensity estimates from those of AX12 and AX16 (Fig.
+\ref{fig:Geologic_map}).
+This is consistent with the interpretation that differential cooling history due to heterogeneity of the timing and depths of the Beaver River diabase intrusions caused the anorthosites to capture paleosecular variations \textit{ca.} 1092 Ma.
+
+Another consequence of having the diabase and anorthosite cooled slowly during emplacement is its effect on biasing the paleointensity estimates toward higher values.
 It has been shown that large differences in cooling rates between acquisition of an NRM in nature versus a pTRM in the lab can result in overestimated paleointensities for SD grains \cite{Dodson1980a, Halgedahl1980a, Nagy2021a}.
@@
 \caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment values from the PINT database (PINT v8.0.0; \url{http://www.pintdb.org/}; ref.
-\citealp{Bono2021a}) with additional Mesoproterozoic data from \cite{Sprain2018a, Kodama2019a}, as well as those from the Neoproterozoic \cite{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}.
+\citealp{Bono2021a}) with additional Neoproterozoic data from refs.
+\citealp{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}.
 Paleointensity estimates from \cite{Pesonen1983a} and \cite{Kulakov2013a} are not included in the compilation due to the specimen-level double-slope behavior as discussed in the text.
-Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \cite{Bono2019a} which is based on an interpretation of there being a progressive decay of the geodynamo through the Proterozoic.}}
+Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \cite{Bono2019a} which is based on an interpretation of there being a monotonic decay of the geodynamo through the Proterozoic.}}
 \label{fig:PINT_compilation}
@@
 \section*{Proterozoic geodynamo}
-To best characterize the geomagnetic axial dipole field intensity during a certain time period requires a paleointensity dataset to cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged and the geocentric axial dipole hypothesis can be assumed.
-Thermal modeling results from ref.
-\citealp{Zhang2021b} suggest that the Beaver River anorthosite xenoliths were likely heated to tholeiitic magma temperature and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a few thousand years.
-In addition, anorthosite xenoliths have paleointensity results that are consistent within small regions but vary between localities.
-Anorthosite AX12 and AX16 have very similar paleointensity estimates that are distinct from those of sites AX11 and AX13 which were emplaced $\sim$10 km away (Fig.
-\ref{fig:Geologic_map}).
-This is consistent with the interpretation that differential cooling history due to heterogeneity of the timing and depths of the Beaver River diabase intrusions caused the anorthosites to capture paleosecular variations \textit{ca.} 1092 Ma.
-% Overall, none of the anorthosite xenoliths record a site mean VDM value lower than 40 ZAm$^2$ (Fig. \ref{fig:PINT_compilation}).
 
-The new paleointensity records from the \textit{ca.} 1092 Ma Midcontinent Rift anorthosite xenoliths together with previous results from the \textit{ca.} 1140 Ma rocks of the Midcontinent Rift region indicate that a six-fold increase in geomagnetic dipole moment happened within $\sim$50 Myr during the late Mesoproterozoic (Fig.
+The long-term evolution of the core's thermal regime and geodynamo has been an outstanding question in the study of Earth history.
+Progress has been made in several fronts including constraining iron and its alloy's thermal conductivity values in high-temperature, high-pressure conditions similar to that in Earth's core; constructing numerical geodynamo models to simulate geomagnetic field behaviors under various thermal regimes; and obtaining and compiling observational paleomagnetic intensity records to test hypotheses put forward by numerical models.
+
+Some studies interpret that the records of the Proterozoic paleointensities hint a progressive, monotonic decay trend, indicating that the geodynamo had a progressively waning power source \cite[e.g.]{Bono2019a}.
+This trend was interpreted to support the hypothesis that Earth's inner core had not solidified and thermal convection was dissipating away its rigor during the Proterozoic.
+With ultra-low paleointensity observations obtained from the ca.
+565 Ma Sept-$\hat{I}$les intrusions, ref.
+\citealp{Bono2019a} hypothesized that the inner core did not begin to nucleate before at least the late Neoproterozoic and the thermal conductivity of the core must be near the high end of viable values obtained through experiments \cite{Ohta2016a}.
+Following this hypothesis, numerical modeling with particularly high core thermal conductivity values show that a decaying geodynamo through the Proterozoic would reach its lowest strength before the emergence of the solid inner core during the Ediacaran-Cambrian \cite{Davies2021a}.
+Modeling also predicts that a rapid (within a few 10s of Myr) recovery in geomagnetic field intensity occurred soon after the onset of inner core nucleation, as compositional convection started to provide additional power to drive the geodynamo \cite{Driscoll2016a}.
+
+However, the high-quality, consistent paleointensity estimates from the well-preserved ca.
+1092 Ma anorthosite xenoliths of the intracontinental Midcontinent Rift confirm the existence of a high geomagnetic field in the late Mesoproterozoic.
+Together with previous records obtained from the ca.
+1106 Ma Osler Volcanics of the Midcontinent Rift \cite{Sprain2018a}, these data indicate an appreciable power source for Earth's dynamo could have persisted through at least 14 Myr during the Mesoproterozoic (Fig.
 \ref{fig:PINT_compilation}).
-The Abitibi dikes \cite{Macouin2003a} record a low average virtual dipole moment of $\sim$13 ZAm$^2$, similar to the values of $\sim$7 ZAm$^2$ obtained from the single silicate crystals of the \textit{ca.} 565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a}.
-However, the relatively high moments of $\sim$72 ZAm$^2$ from the \textit{ca.} 1105 Ma Osler Volcanic Group \cite{Sprain2018a} and $\sim$89 ZAm$^2$ from the \textit{ca.} 1092 Ma Beaver River anorthosite xenoliths (Fig.
-\ref{fig:PINT_compilation}) are strikingly higher than the values of the Abitibi dikes.
-The scale of this increase is similar to the increasing trend observed near the Ediacaran-Cambrian which has been suggested to evidence the onset of Earth's inner core nucleation \cite{Thallner2021a}.
+Besides the high geomagnetic fields recorded by the Midcontinent Rift rocks, the ca.
+755 Ma Mundine well dikes also record a strong geomagnetic field in the Neoproterozoic with an average virtual dipole field of $\sim$60 ZAm$^2$ (Fig.
+\ref{fig:PINT_compilation}).
+These observations are inconsistent with there being a monotonic decay during the Proterozoic as predicted by magnetic scaling and thermal evolution models using high core thermal conductivity values \cite{Aubert2009a, Davies2021a}.
 
-The high geomagnetic field estimates from the Midcontinent Rift volcanics and intrusives are not the only suite of high paleointensity records at odds with the hypothesis of a monotonically decaying geodynamo through the entire Proterozoic Eon \cite{Biggin2009a, Bono2019a}.
+Numerical simulations that apply magnetic scaling law also predict that there was a strong increase in true geomagnetic dipole moment during the early stages of inner core growth \cite{Davies2021a}.
+The updated paleointensity data compilation shows that rapid transitions from low to high field intensities do not occur on an unique case during Earth history (Fig.
+\ref{fig:PINT_compilation}).
+A two-fold increase in Earth's virtual dipole moment in the Ediacaran has been recorded by the ca.
+565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a} and the ca.
+550 Ma volcanics rocks of the Skinner Cove Formation \cite{Thallner2021a} (Fig.
+\ref{fig:PINT_compilation}).
+That transition has been interpreted to be caused by the initiation of the nucleation of the solid inner core \cite{Thallner2021a} .
+However, in the Mesoproterozoic, a six-fold increase in geomagnetic dipole moment is observed within a short period ($\sim$40 Myr) from a very low average virtual dipole moment of $\sim$13 ZAm$^2$ recorded by the ca.
+1140 Ma Abitibi dikes which yielded straight Arai plots \cite{Macouin2003a}, to a high moment of $\sim$70 ZAm$^2$ recorded by the ca.
+1106 Ma Osler Volcanics, which might have lasted through ca.
+1092 Ma when the Beaver River anorthosite xenoliths recorded a virtual dipole moment of $\sim$80 ZAm$^2$.
+This increase in absolute value of virtual dipole moment dwarfs all other existing records of geomagnetic field fluctuations in the Precambrian, including those in the Ediacaran (Fig.
+\ref{fig:PINT_compilation}).
+The tempo and scale of this field intensity transition may also be a better match with that predicted by numerical simulations upon the onset of inner core nucleation \cite{Davies2021a} but this similarity could be coincidental.
+
+On the other hand, particularly low paleointensity values similar to those observed in the Ediacaran has been obtained at multiple times before and after the Ediacaran-Cambrian.
+In the late Mesoproterozoic, besides the the \textit{ca.} low intensities recorded by the ca.
+1140 Ma Abitibi dikes \cite{Macouin2003a}, one ca.
+1070 Ma Bangemall Sill records an extremely low virtual dipole moment of $\sim$6.4 ZAm$^2$ (Fig.
+\ref{fig:PINT_compilation}; ref.
+\citealp{Lloyd2021b}).
+In the Neoproterozoic, ca.
+720 Ma dolerite dikes of the Franklin Large Igneous Province record a mean virtual dipole moment of 11 ZAm$^2$ (Fig.
+\ref{fig:PINT_compilation}; ref.
+\citealp{Lloyd2021a}).
+In the Phanerazoic, Devonian-aged rocks from the Strathmore lava flows \cite{Hawkins2021a}, Minusa Basin, Kola Peninsula \cite{Shcherbakova2017a}, volcanic rocks from the East European Craton \cite{Shcherbakova2019a}, as well as the Buribay volcanic complex \cite{Shcherbakova2021a} show low paleointensities through an extended period of time (Fig.
+\ref{fig:PINT_compilation}).
+Accompanying the low paleointensity values, paleodirectional data from the Ediacaran-Cambrian \cite{Bono2015a, Halls2015b, Kodama2021a} and the Devonian show anomalous directional fluctuations \cite{Shcherbakova2019a} which could be consistent with there being a largely diminished dipole component of Earth's magnetic field.
+If the weak paleointensities in the Ediacaran or the Devonian are attributed to the geodynamo being on the brink of collapse just prior to the nucleation of the solid inner core, such data would suggest that the thermal conductivity of the core must be extremely high \cite{Davies2015a, Ohta2016a}.
+Alternatively, these Proterozoic and Paleozoic records could be a manifestation of cyclic forcing of the geodynamo across the core–mantle boundary with a time scale on the order of 150 to 200 Myr \cite{Courtillot1987a, Hawkins2021a}, a feature that has been proposed to be associated with the Wilson cycle and magnetic superchrons, and perhaps postdate the emergence of the solid inner core \cite{Courtillot1987a, Courtillot2007a}.
+
+Another alternative explanation for the low paleointensities in the Paleozoic and Proterozoic is that they could be a result of short-term geomagnetic field fluctuations.
+A five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years \cite{Bono2021a}, when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed.
+The observation of weak fields before and after the Cambrian indicate that the geomagnetic field may excurse toward particularly low strengths on narrow time scales regardless of the existence of a solid inner core.
+Therefore, low paleointensity values or short-term fluctuations may not be sufficient for interpreting Earth's inner core status through history.
+
+% The high geomagnetic field estimates from the Midcontinent Rift volcanics and intrusives are not the only suite of high paleointensity records at odds with the hypothesis of a monotonically decaying geodynamo through the entire Proterozoic Eon \cite{Biggin2009a, Bono2019a}.
 The $\sim$60 ZAm$^2$ virtual dipole moment estimates from the \textit{ca.} 755 Ma Mundine Well dikes are much higher than the values predicted by the monotonic decrease model of \cite{Bono2019a}.
@@
 
-In addition, a five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years \cite{Bono2021a}, when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed.
-The observation of weak fields before and after the Cambrian indicate that the geomagnetic dipole may reach particularly low strengths regardless of the existence of a solid inner core.
-Therefore, low paleointensity values or short-term fluctuations may not be sufficient for interpreting Earth's inner core status through history.
+Constraining the age of the inner core remains to be an open question given the uncertainties in measuring the thermal conductivity values of iron alloys in lab using different approaches \cite{Konopkova2016a, Ohta2016a}.
+Experimental challenges such as measuring sample thicknesses under high-pressure, high-temperature conditions, together with theoretical difficulties of the validity of applying the Wiedemann-Franz law to extrapolate thermal conductivity values based on electrical resistivity measurements \cite{Ohta2016a}, and modeling uncertainties associated with parameters used in the direct thermal conduction measurement experiment \citea{Konopkova2016a} cast uncertainty in the age of the inner core by hundreds of millions of years.
+Until further experiments and theory that can explain these contrasting results are developed, our knowledge on the age of Earth's thermal evolution remain largely uncertain.
 
-Recent advancements in high-pressure, high-temperature experiments for measuring thermal conductivity of iron alloys under conditions similar to Earth's core may shed new light on the age of the inner core.
-However, large uncertainties still remain regarding the applicability Wiedemann-Franz law of extrapolating thermal conductivity values based on electrical resistivity measurements by ref.
-\citealp{Ohta2016a}, and regarding the thermal modeling parameters used in the direct thermal conduction measurement experiment performed by ref.
-\citealp{Konopkova2016a}.
-Until further experiments and theory that can explain these contrasting results are developed, our knowledge on the age of Earth's inner core remain largely uncertain.
-
-In the meantime, paleomagnetic directional anomalies may be another suite of intriguing observational records that is worth investigating.
+% In the meantime, paleomagnetic directional anomalies may be another suite of intriguing observational records that is worth investigating.
 The anomalous directional fluctuations during the Ediacaran and Devonian \cite{Bono2015a, Kodama2021a, Shcherbakova2019a} perhaps can be consistent with there being a weak geodynamo regime when dipole component of Earth's field was largely diminished.
 
-Regardless, the high values from the Midcontinent Rift intrusions and extrusive volcanics are inconsistent with the prediction of there being a progressive decay at the time (Fig.
-\ref{fig:PINT_compilation}; ref.
-\citealp{Bono2019a}) and necessitates that there being a vigorous late Mesoproterozoic geodynamo.
+Overall, the high-fidelity paleointensity recorders of the Beaver River anorthosite xenoliths in the well-preserved Midcontinent Rift record high paleointensities values and necessitates that there was a vigorous late Mesoproterozoic geodynamo.
 

### 2022-01-25 PNAS_manuscript/PNAS_manuscript.tex

@@
 These data record a strong present-day-like virtual dipole moment 1.1 billion years ago.
-A strong field persisted over a 25 million year or greater interval indicating the existence of appreciable power sources for Earth's dynamo during the late Mesoproterozoic.
-}
+A strong field persisted over a 25 million year or greater interval indicating the existence of appreciable power sources for Earth's dynamo during the late Mesoproterozoic.}
 
@@
 % At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
-\keywords{absolute paleointensity $|$ Laurentia $|$ Midcontinent Rift $|$ anorthosite $|$ geodynamo $|$ Mesoproterozoic $|$ inner core}
+\keywords{absolute paleointensity $|$ Midcontinent Rift $|$ anorthosite $|$ geodynamo $|$ Mesoproterozoic $|$ inner core}
 
@@
 Such a strong late Mesoproterozoic dipole recorded by the anorthosite xenoliths requires the existence of a vigorous geodynamo at the time.
-Together with previous paleointensity data from other Midcontinent Rift rocks, our new results indicate that a dynamo with strong power sources persisted for at least $\sim$25 million years \textit{ca.} 1.1 Ga.
+Together with previous paleointensity data from other Midcontinent Rift rocks, our new results indicate that a dynamo with strong power sources persisted for more than 20 million years \textit{ca.} 1.1 Ga.
 These data are inconsistent with there being a progressive monotonic decay of Earth's dynamo strength through the Proterozoic Eon.
@@
 \dropcap{E}arth's magnetic field is the result of convective flow of liquid iron-alloy in Earth's outer core.
-At present day, the geodynamo is collectively powered by heat flow across the core-mantle boundary (CMB), and from the crystallization of the solid inner core from the liquid outer core which provides latent heat and compositional buoyancy.
+At present day, the geodynamo is collectively powered by heat flow across the core-mantle boundary (CMB) and from the crystallization of the solid inner core from the liquid outer core which provides latent heat and compositional buoyancy \cite{Buffett2000a}.
 However, while paleomagnetic studies have found that a dynamo field has existed since at least 3.4 billion years ago \cite{Selkin2007a, Tarduno2014a, Brenner2020a}, Earth's inner core likely crystallized more recently.
@@
 
-% Without an inner core, Earth's magnetic field would have been generated solely through CMB heat flow, with perhaps additional stirring by other power sources such as Mg precipitation \cite{ORourke2016a}, or by exotic processes such as a convecting basal magma ocean \cite{Ziegler2013a}.
-Long-term secular cooling of Earth has the potential to have decreased the power available to the geodynamo through time prior to the nucleation of the inner core \cite[e.g][]{Davies2021a}.
-As a result, it is possible that intensity of the geomagnetic axial dipole field would have experienced a long-term decay before the emergence of the solid inner core after which it would have strengthened \cite{Aubert2009a, Davies2021a}.
-However, the estimates for thermal conductivity values for the core continue to vary \cite{Konopkova2016a, Ohta2016a, Zhang2020a}.
-Obtaining an observational record of the timing of inner core formation would constrain the thermal evolution of the core and mantle, in addition to improving our understanding of the history of Earth’s dynamo.
-
-Paleomagnetic records from ancient rocks are one of the few types of observational data that have the potential to record the thermal evolution of Earth’s core.
+Paleomagnetic records from ancient rocks are one of the few types of observational data that have the potential to provide constraints on the thermal evolution of Earth’s core.
 Evidence for a persistent magnetic field through the Proterozoic, for example, necessitates the existence of plate tectonics that sustained core-mantle boundary heat flow \cite{Swanson-Hysell2021c}.
-However, strikingly low estimates of geomagnetic field strengths have been obtained in the Ediacaran Period \cite{Bono2019a, Shcherbakova2019a,Thallner2021b} and the Devonian Period \cite{Shcherbakova2017a, Shcherbakova2021a}, potentially indicating unusual periods of core dynamo regime at those times.
-% These data have been interpreted to indicate that there was a progressively decaying field up to that time that was followed soon after by initial crystallization of the inner core \cite{Bono2019a}.
+However, strikingly low estimates of geomagnetic field strengths have been obtained during the 635 to 539 Ma Ediacaran Period \cite{Bono2019a, Shcherbakova2019a,Thallner2021b} and the 419 to 359 Ma Devonian Period \cite{Shcherbakova2017a, Shcherbakova2021a}, potentially indicating unusual periods of core dynamo regime at those times.
+The Ediacaran data have been interpreted to indicate that there was a progressively decaying field up to that time that was followed soon after by initial crystallization of the inner core \cite{Bono2019a}.
 
@@
 
-Determinations of the absolute value of the the ancient geomagnetic field strength rely on igneous rocks that acquire thermal remanent magnetizations as they cool.
+Determinations of the absolute value of the ancient geomagnetic field strength rely on igneous rocks that acquire thermal remanent magnetizations as they cool.
 These magnetizations need to be unmodified by subsequent heating or chemical alteration in order to maintain a record of the ancient geomagnetic field from the time of cooling.
@@
 This interior position results in them typically being distant from tectonic events along continental margins that can drive alteration through heat and fluids.
-However, large igneous provinces that are often associated with intraplate magmatism are typically of geologically short duration with the bulk of magmatic products emplaced within 1 Myr or less \cite{Kasbohm2021a}.
+However, intraplate magmatism associated with large igneous provinces is typically of geologically short duration with the bulk of magmatic products emplaced within 1 Myr or less \cite{Kasbohm2021a}.
 The Midcontinent Rift is an exception as it is a large igneous province where magmatism lasted $\sim$25 Myr from \texit{ca.} 1109 Ma to 1084 Ma \cite{Swanson-Hysell2019a}.
 Additionally, extension ceased in the Midcontinent Rift prior to lithospheric separation, preserving volcanic, intrusive, and sedimentary rocks of the rift within the continental interior.
-As a result, rocks of the rift have unusually simple paleomagnetic behavior for their greater than one billion-year-old age and form a central record of Mesoproterozoic paleogeography \cite{Swanson-Hysell2021c}.
+As a result, rocks of the rift have unusually simple paleomagnetic behavior for their greater than one billion-year-old age and paleomagnetic data from rift rocks forms a central record of Mesoproterozoic paleogeography \cite{Swanson-Hysell2021c}.
 The duration of magmatic activity within the Midcontinent Rift is longer than the entire 20.4 Myr long Neogene Period such that it enables an extended well-preserved window into the intensity of Earth's magnetic field in the late Mesoproterozoic.
 
-Despite the excellent preservation of the rocks, non-ideal paleointensity behaviors and overall low success rates have often led to high failure rates and challenged the interpretation of many previous paleointensity results from the Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a}.
-The most trusted type of paleointensity estimate is that obtained through experiments in which the primary natural thermal remanent magnetization is progressively replaced by a laboratory magnetization that is imparted in a known field with internal consistency checks (such as in the IZZI-style Thellier experiments; ref.
+Despite the excellent preservation of the rocks, non-ideal paleointensity behaviors have challenged the interpretation of many previous paleointensity results from the Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a}.
+The most trusted type of paleointensity estimate is that obtained through experiments in which the primary natural thermal remanent magnetization is progressively replaced by a laboratory magnetization that is imparted in a known field with internal consistency checks (such as in IZZI-style Thellier experiments; ref.
 \citealp{Yu2004a}).
@@
 Anorthosites are cumulate rocks composed dominantly of the mineral plagioclase.
-They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud minute magnetic inclusions from alteration with alteration of the plagioclase crystals not resulting in the formation of secondary iron oxides.
+They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud minute magnetic inclusions from alteration with minor alteration of the plagioclase crystals not resulting in the formation of secondary iron oxides in contrast with mafic minerals such as olivine.
 The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal intrusions of the Beaver River diabase \cite{Zhang2021b}.
@@
 Magnetic imaging shows the anorthosite specimens have dominant magnetic carriers interstitial to and within plagioclase crystals and they do not display strong preferred orientations.
-Our single-slope, high quality paleointensity data add to the reliable paleointensity records during $\sim$25 Myr of Midcontinent Rift magmatic activity and require that there was a strong late Mesoproterozoic geodynamo.
-
-% Interpretations of selected records of the current Precambrian absolute paleointensity database (compiled in the PINT database; \url{http://earth.liv.ac.uk/pint/}) suggest that an appreciable geomagnetic field existed in the Archean but reached its weakest in the Ediacaran when very low virtual dipole moment (VDM) estimates have been obtained \cite{Bono2019a, Shcherbakova2019a, Thallner2021b, Lloyd2021a}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
-Ref.
-\citealp{Bono2019a} adopted a billion-year-scale view of such observations to suggest that these data are consistent with thermal evolution models that use high core thermal conductivity values, where the nucleation of the inner core did not begin to provide compositional convection power for driving the geodynamo until the Ediacaran-Cambrian, suggesting that the Proterozoic dynamo might have experienced a monotonic decay when compositional convection was lacking and thermal convection was progressively decreasing \cite{Driscoll2016a, Davies2021a}.
-
-% Other studies focusing on paleomagnetic directions have also found that the late Ediacaran might have been a period of time when the geomagnetic field were experiencing frequent reversals and waning of dipole dominance \cite{Bono2015a, Kodama2020a}.
-
-% However, previous paleointensity data from intrusive and extrusive rocks in the \textit{ca.} 1.1 Ga Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record VDM values similar to that of today, which are higher than the values predicted by the long term decay curve proposed by ref.
-\citealp{Bono2019a} (Fig.
-\ref{fig:PINT_compilation}).
-Ref.
-\citealp{Pesonen1983a} obtained an average virtual dipole moment of 93 $\times 10^{21}$ Am$^2$ (ZAm$^2$) after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using both extrusive and intrusive rocks from the Midcontinent Rift, including the \textit{ca.} 1108-1105 Ma Osler Volcanics, the \textit{ca.} 1100 Ma Maimainse Point volcanics, the \textit{ca.} 1107 Ma intrusive Logan Sills, Thunder Bay dikes, the Baraga and Marquette dikes, as well as lava flows of the \textit{ca.} 1096 Ma North Shore Volcanic Group (Fig.
-\ref{fig:Geologic_map}).
-Ref.
-\citealp{Kulakov2013a} obtained an average VDM of 59 $\pm$ 11 ZAm$^2$ using the \textit{ca.} 1085 Ma Lake Shore Traps volcanics and suggest that the present-day-like geomagnetic field strength likely existed through the late stage of Midcontinent Rift magmatism \cite{Vervoort2007a, Miller2013a}.
-However, non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results obtained by ref.
-\citealp{Pesonen1983a, Kulakov2013a}.
-One typical failure of paleointensity experiment is the double-slope behavior, as two distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
-Ref \citealp{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas ref.
-\citealp{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
-Results with such non-ideal behavior were rejected by ref.
-\citealp{Sprain2018a} who applied stricter paleointensity selection criteria on new paleointensity results and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
-A calculated mean VDM value of 56 $\pm$ 21 ZAm$^2$ is similar to the average value of the past 300 million years \cite{Sprain2018a}.
-
-% Although it is typical of paleointensity experiments to have low success rates, bulk plagioclase-rich rocks \cite{Selkin2000a} and single plagioclase crystals \cite{Tarduno2005a} are found to be able to produce high quality paleointensity results.
-It has been suggested that plagioclase crystals can protect magnetic minerals that exsolved within them against post-formation weathering and thermochemical alteration during paleointensity experiments \cite{Tarduno2005a}.
-In this study, we present new paleointensity results from IZZ modified absolute paleointensity experiments \cite{Yu2004a} on a suite of bulk samples from \textit{ca.} 1092 Ma anorthosite xenoliths and their Beaver River diabase host rocks of the Beaver Bay Complex within the Midcontinent Rift.
-Paleointensity experiments on the anorthosite xenoliths have a very high success rate.
-Their consistent specimen- and site-level paleointensity results together with rock magnetic data reveal that the anorthosite xenoliths have low anisotropy of thermal remanent magnetization (TRM) and can acquire TRM linearly.
-Magnetic imaging shows the anorthosite specimens have dominant magnetic carriers interstitial to and within plagioclase crystals and they do not display strong preferred orientations.
-Our single-slope, high quality paleointensity data add to the reliable paleointensity records during the $\sim$25 Myr of Midcontinent Rift magmatic activity.
-Our new data from the Midcontinent Rift intrusive rocks support previous data from the volcanics that there was a strong late Mesoproterozoic geodynamo.
+These single-slope, high quality paleointensity data, together with other high-quality paleointensity records during $\sim$25 Myr of Midcontinent Rift magmatic activity, require that there was a strong late Mesoproterozoic geodynamo.
 
@@
 
-% The Mesoproterozoic North American Midcontinent Rift is an intracontinental rift that did not result in the break-up of Laurentia.
-Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; refs.
-\citealp{Vervoort2007a, Miller2013a}) was characterized by rapid and voluminous emplacement of intrusive rocks and extrusive volcanics.
-The $<$500 Kyr emplacement of the \textit{ca.} 1096 Ma Duluth Complex and the associated $\sim$8 km thick extrusive volcanics is one manifestation \cite{Swanson-Hysell2021a}.
-The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex in northern Minnesota punctuates another such period of magmatism during the Midcontinent Rift main stage activity.
-The hypabyssal intrusion of the 1091.7 $\pm$ 0.2 Ma Beaver River diabase dikes and sills of the Beaver Bay Complex transported numerous anorthosite xenoliths that have short-axis diameters up to 180 meters via wide conduits \cite{Boerboom2004a, Boerboom2006b}.
-Geochronology, geochemical, and paleomagnetic data support the hypothesis that the diabase extruded to the surface and fed the coeval voluminous Greenstone Flow of the Portage Lake volcanics \cite{Doyle2016a, Zhang2021b}.
-
 Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths to have acquired thermal remanent magnetizations while cooling with the Beaver River diabase \cite{Zhang2021b}.
@@
 \centering
-\noindent\includegraphics[width=10.5 cm]{Petro_QDM.png}
-\caption{\footnotesize{Thin section petrographic and magnetic images of anorthosites from the Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (left column; ref.
-\citealp{Zhang2021b}) and distinct anorthosite from the Duluth Complex Anorthositic Series (right column).
+\noindent\includegraphics[width=13 cm]{Petro_QDM.png}
+\caption{\footnotesize{Thin section petrographic images (A,B,C,D) and magnetic maps (E,F) of anorthosite from the Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (left column; ref.
+\citealp{Zhang2021b}) and a distinct anorthosite from the Duluth Complex Anorthositic Series (right column).
 The Duluth Complex anorthosites were not targeted for paleointensity experiments in this study given the complexities associated with more pronounced fabrics.
-Cross-polarized petrographic images (A) Beaver River anorthosite have plagioclase with a granoblastic texture; (B) Plagioclase of Duluth Complex anorthosite have euhedral, interlocking crystals showing igneous foliation; (C) zoomed-in image of plagioclase from MS99033 showing the crystals to be largely free of large opaque inclusions; (D) Duluth Complex plagioclase can contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the [010] planes of the plagioclase crystals.
+Cross-polarized petrographic images of the Beaver River anorthosite (A,C) reveal plagioclase with a granoblastic texture with platioclase crystals to are largely free of large opaque inclusions.
+In contrast, plagioclase crystals in the Duluth Complex anorthosite have euhedral, interlocking crystals with an igneous foliation (B) and the plagioclase crystals contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the [010] planes of the crystals.
 Magnetic field maps acquired by the quantum diamond microscope (QDM) show that the relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite in (E) while large oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the overlapping magnetic fields above the oxides in (F).
@@
 
-\section*{Results and Discussion}
+\section*{Results and Interpretations}
 
@@
 
-Petrographic imaging shows the dominantly monomineralic anorthosite xenoliths within the Beaver River diabase often have granoblastic texture characterized by equigranular crystals with weakly developed petrofabrics (Fig.
+The dominantly monomineralic anorthosite xenoliths within the Beaver River diabase often have granoblastic texture characterized by equigranular crystals with weakly developed petrofabrics (Fig.
 \ref{fig:Petro_QDM}A,C).
@@
 \ref{fig:Petro_QDM}B).
-In addition, there are abundant Fe-Ti oxide needles that are typically aligned with the [010] planes of plagioclase crystals (Fig.
+In addition, there are abundant Fe-Ti oxide needles that are typically aligned with the [010] planes of the plagioclase crystals (Fig.
 \ref{fig:Petro_QDM}D).
@@
 The relative lack of fabric and associated preferred orientation makes the Beaver River anorthosite xenoliths a particularly compelling target for paleointensity experiments.
-
-% The dominantly monomineralic anorthosite xenoliths of the Beaver River diabase have an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}, higher than the average of $\sim$60\% of the Anorthositic Series rocks of the Duluth Complex \cite{Miller1990a}.
-The plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture showing igneous foliations (Fig.
-\ref{fig:Petro_QDM}A), but the more calcium-rich plagioclase of the Beaver River anorthosites often develop granoblastic texture characterized by equigranular crystals with poor alignments (Fig.
-\ref{fig:Petro_QDM}B).
-Abundant magnetite and ilmenite needles that are typically 10 $\mu$m long can be found to have preferred orientations within Duluth Complex plagioclase.
-Interstitial titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex plagioclase \cite{Miller1990a}.
-In contrast, titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent but relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
-\ref{fig:Petro_QDM}).
-This reflects that the Beaver River diabase magma had a different oxidation history than that of the Duluth Complex.
-The plagioclase and oxide mineral textures of the Beaver River anorthosites also distinguish them from plagioclase cumulates of other layered mafic complex where magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
-
-% Such a high concentration of Fe oxides carrying magnetic remanence in the slowly cooled Duluth Complex anorthositic rocks reflect that iron in the plagioclase mush in the magma chamber acted as a means of buffering the melt with respect to oxygen, forming crystallites along preexisting plagioclase substrate or as iron oxides incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures \cite{Scofield1986a}.
-In contrast, titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent but relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
-\ref{fig:Petro_QDM}).
-% the Beaver River diabase rapidly ascended and entrained the anorthosite xenoliths directly into shallow chambers and likely did not reach  \cite{Flower1980a, Miller1997a, Zhang2021b}.
-Particularly, \cite{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
-This is consistent with the interpretation that these anorthosite were rapidly transported by upward currents into hypabyssal magma chambers without reaching structural equilibrium.
-Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Fig.
-\ref{Petro_QDM}).
-Instead, relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
-\ref{Petro_QDM}).
-
-% In the Beaver River anorthosite xenoliths, relatively sparse titano-magnetite occur within or interstitial to the plagioclase (Fig. \ref{Petro_QDM}).
-
-% Occurrences of magnetite and ilmenite symplectites and needles within anorthositic rocks have been described by \cite{Scofield1975a, Scofield1986a} in the Wichita Mountains and by \cite{Haggerty1976a} in other slowly cooled, mafic intrusions.
-\cite{Scofield1986a} hypothesized that the inclusions acted as a means of buffering the melt with respect to oxygen, forming either crystallites along preexisting plagioclase substrate or as iron incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures.
-Such a model requires that plagioclase, instead of other mafic minerals such as olivine, to be the first to crystallize from a magma and act as a sink for iron.
-The two-stage formation of the Duluth Complex Anorthositic Series fits this model \cite{Miller1990a}.
-During the first stage, high pressure at subcrustal depths (15-25 km depth) caused fractionation of Duluth Complex magma resulting in the formation of floating plagioclase mushes \cite{Kushiro1980a}.
-The second stage of crystallization then occurs at low pressures after the magma intruded into shallower chambers.
-
-% This proposed magmatic evolution of the Duluth Complex is in distinct contrast to the emplacement of the Beaver River diabase which rapidly ascended and entrained the anorthosite xenoliths directly into shallow chambers \cite{Flower1980a, Miller1997a, Zhang2021b}.
-Particularly, \cite{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
-This is consistent with the interpretation that these anorthosite were rapidly transported by upward currents into hypabyssal magma chambers without reaching structural equilibrium.
-Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Fig.
-\ref{Petro_QDM}).
-Instead, relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
-\ref{Petro_QDM}).
-
-% it is the most probable that the anorthosite xenoliths of the Beaver River diabase are cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
-Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths acquired thermal remanent magnetizations during cooling of the diabase \cite{Zhang2021b}.
-Step-wise thermal demagnetizations show many anorthosite xenoliths have minimal secondary remanence and single-component characteristic remanence magnetizations that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C.
-
-% Geochronological, paleomagnetic, geochemical, and petrologic data are consistent with the hypothesis that the Beaver River diabase magma extruded to the surface, feeding the massive Greenstone Flow - one of the largest lava flows on Earth \cite{Doyle2016a, Zhang2021b}.
-
-% Based on high-precision $^{206}$Pb/$^{238}$U geochronology and paleomagnetic data, \cite{Swanson-Hysell2021a} hypothesized that the massive Duluth Complex and the coeval North Shore Volcanic group (Fig.
-\ref{fig:Geologic_map}) formed as a result of upwelling mantled-derived plume encountering Laurentia's thinned lithosphere, following an ``upside-down” drainage" which rapidly funneled melt to today's northeastern Minnesota (Fig.
-\ref{fig:Geologic_map}) \cite{Sleep1997a, Swanson-Hysell2014a}).
-The plagioclase-phyric Anorthositic Series rocks, the crystal-poor Layered Series intrusions of the Duluth Complex, and the coeval volcanics of the North Shore Volcanic Group (NSVG) that outcrop along Northeastern Minnesota (Fig.
-\ref{fig:Geologic_map}) are interpreted to be a result of such vigorous magmatic activity.
-
-% Despite that the anorthosite xenoliths of the Beaver River diabase were entrained as xenoliths and emplaced only $\sim$4 Myr after the formation of the Anorthositic Series rocks of the Duluth Complex when massive plagioclase mushes were generated, $^{206}$Pb/$^{238}$U zircon geochronology data and zircon rare earth element diffusion modeling results suggest it is the most probable that the Beaver River diabase anorthosite xenoliths are entrained cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
-The anorthosite xenoliths of the \textit{ca.} 1092 Ma Beaver River diabase has distinct mineralogy from that of the Anorthositic Series rocks of the Duluth Complex.
-The dominantly monomineralic anorthosite xenoliths have an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}, higher than the average of $\sim$60\% of the Anorthositic Series rocks of the Duluth Complex \cite{Miller1990a}.
-In addition, petrographic imaging support that the plagioclase from the Beaver River anorthosite formed under a magmatic conditions different from the Duluth Complex anorthositic series magma.
-While the plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture (Fig.
-\ref{Petro_QDM}A), the more calcium-rich plagioclase of the Beaver River anorthosite often develop granoblastic texsture characterized by equigranular plagioclase crystals (Fig.
-\ref{Petro_QDM}B).
-Abundant magnetite and ilmenite needles that are typically 10 $\mu$m long can be found to have preferred orientations within Duluth Complex plagioclase.
-Interstitial vermicular titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex Anorthositic Series rocks \cite{Miller1990a}.
 
@@
 \citealp{Biggin2014a}) for the anorthosite xenoliths are all above 4 (Table S2).
-The cooling rate-corrected absolute paleointensity estimates from all anorthosite sites have a weighted mean of 52.72 $\pm $5.72 $\mu$T.
+The cooling rate-corrected absolute paleointensity estimates from all anorthosite sites have a weighted mean of 52.72 $\pm$ 5.72 $\mu$T.
 This corresponds to a calculated virtual dipole moment of $\sim$89 ZAm$^2$ ($10^{21} Am^2$) \textit{ca.} 1092 Ma.
-All measurement-level paleointensity experiment data is available within the MagIC database (\url{}; ATTACHED IS A LINK FOR REVIEWING PURPOSESS, THE URL WILL BE UPDATED WHEN DOI IS GENERATED).
+All measurement-level paleointensity experiment data is available within the MagIC database (\url{https://earthref.org/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}; \textit{this private link is provided for the purpose of review; the url will be updated when a doi is generated for this manuscript}).
 
@@
 
-% Paleointensity experiment results from the anorthosite xenoliths and their host diabase show three general types of dominant specimen behaviors on NRM/TRM plots (Fig.
-\ref{fig:IZZI_examples}): dominantly single-slope, double-slope or sagging, and zigzagging.
 Typical paleointensity experimental data of the anorthosite specimens have straight, single-slope NRM/TRM plots and the accepted fractions of temperature steps span over the origin-trending, primary remanence components (Fig.
@@
 We accept specimen- and site-level absolute paleointensity results from those anorthosite xenoliths that pass the selection criteria.
-Other anorthosite xenoliths and diabase specimens failed the selection criteria largely because of double-slope or sagging behavior (fail FRAC selection; see methods), poor pTRM checks, and sometimes zigzagging behaviors superimposed on top (fail SCAT, DANG selection; see methods; Fig.
+Other anorthosite xenoliths and diabase specimens failed the selection criteria largely because of double-slope or sagging behavior (fail FRAC selection; see methods), poor pTRM checks, and sometimes zigzagging behaviors superimposed on top of sagging behavior (fail SCAT, DANG selection; see methods; Fig.
 \ref{fig:IZZI_examples}).
@@
 
-% As is shown in Fig.
-\ref{fig:IZZI_examples}, specimen BD2-5b has a high-temperature slope with zigzagging behavior, poor pTRM checks, and low NRM/TRM ratio, in addition to a low-temperature slope with a steep NRM/TRM ratio.
-Although the BD2 specimens' lower temperature fraction of the plots are straight and the pTRM checks are consistent, lower temperature slopes from BD2 give consistently higher paleointensity estimates than those from site AX6, AX8, and AX10, which are anorthosite xenoliths included within BD2 (Fig.
-\ref{fig:Geologic_map}; ref.
-\citealp{Zhang2021a}.
-This higher estimate from the diabase lithology could be attributed to the systematic bias of using the lower temperature slope of the NRM/TRM plots demonstrated in Fig.
-\ref{fig:IZZI_examples} and specimens from site BD2 all fail the criteria of having a maximum magnetization gap between adjacent steps of 25\% of the total NRM.
-Eventually, we only accept the single-slope anorthosite specimens results for further analyses.
-
-% \begin{figure}[h!]
-% \noindent\includegraphics[width=\textwidth]{Cooling_rate_correction.pdf}
-% \centering
-% \caption{\small{}}
-% \label{fig:cooling_correction}
-% \end{figure}
-
-% A 20 mT alternating field treatment after in-field heating steps were applied to some specimens in an attempt to mitigate non-ideal behavior contributed by multidomain or non-uniformly magnetized magnetic carriers.
-The treatment did not make any diabase specimen to pass the selection criteria.
-On the other hand, 27 out of 58 ($\sim$46.6\%) anorthosite specimens that had AF treatment passed selection and 13 out of 28 ($\sim$46.4\%) specimens without AF treatment passed selection.
-Paleointensity estimates from specimens with or without AF treatment in anorthosite AX10, AX11, AX12, AX13, AX16 do not show significant difference (Fig.
-\ref{fig:PINT_cooling_corrected}).
-Specimen results for AX11 particularly illustrates a case where all specimens within the site yield almost the same paleointensity estimate, regardless of the AF treatment.
-
-% Therefore, the alternating field demagnetization did not result in significant increase in specimen success rate for the anorthosites although it might have improved the specimen behavior for those anorthosites that were subjected to this treatment.
-Together with the narrow range of unblocking temperatures of their sister specimens \cite{Zhang2021b}, these data are consistent with the interpretation that the anorthosites with successful paleointensity results have a magnetic mineral assemblage that is likely dominated by single domain, low-Ti magnetite.
-Some anorthosite specimens and all diabase specimens still failed during result selections even though AF treatment was applied.
-It could be that they have a very high dominance of multidomain (MD) or vortex state grains with relatively moderate coercivity such that an alternating field with 20 mT peak field did not significantly mitigate their non-ideal paleointensity behaviors.
-Another cause for their failure could be the thermochemical alterations that happened during heating in air.
-As can be exampled by the zigzagging high temperature slope with poor pTRM checks in paleointensity plot for specimen BD2-5b (Fig.
-\ref{fig:IZZI_examples}), the decrease in NRM/TRM ratios during higher temperature steps likely resulted from the formation of new ferromagnetic material within the specimen due to oxidation.
-Such alterations can allow specimens to acquire additional remanence during in-field steps that may not be fully removed by an AF treatment.
-
 \subsection*{Rock magnetism}
@@
 \citealp{Verwey1939a, Feinberg2015a}).
-Anorthosite chips from sites that yield successful paleointensity results have Verwey transition temperatures near 120 Kelvin (K) as expected for stochiometric magnetite with minimal Ti \cite{Ozdemir1993a}.
+Anorthosite chips from sites that yield successful paleointensity results have Verwey transition temperatures near 120 K as expected for stochiometric magnetite with minimal Ti \cite{Ozdemir1993a}.
 However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have Verwey transitions that are suppressed toward lower temperatures (Fig.
-S2), indicating that the specimens either has a relatively higher Ti content or has been partially oxidized \cite{Ozdemir1993a}.
+S2), indicating that the magnetite in the specimens either has relatively higher Ti content or has been partially oxidized \cite{Ozdemir1993a}.
 
-Coercivity spectra from backfield experiments further support the interpretation that rocks that pass or fail the selection criteria have distinct magnetic mineralogy.
+Coercivity spectra from backfield experiments further support the interpretation that rocks that pass or fail the selection criteria have distinct magnetic mineralogy (Fig.
+\ref{fig:coercivity}).
 A compilation of median destructive field (MDF) values in Fig.
-\ref{fig:coercivity} shows that specimens from anorthosites that pass paleointensity selections have a distinctly higher average MDF values ($\sim$60 mT) than other anorthosites and diabase ($\sim$30 mT).
+\ref{fig:coercivity} shows that specimens from anorthosites that pass paleointensity selections have a distinctly higher average MDF values than other anorthosites and diabase.
 Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity around 80 mT (Fig.
 \ref{fig:coercivity}).
-In contrast, other anorthosite and diabase tend to have lower coercivities ($\sim$30 mT).
-This result is consistent with the interpretation that they lack a dominance of single domain magnetite grains in their magnetic mineral assemblages and thus have non-ideal paleointensity behaviors.
+In contrast, other anorthosite and diabase tend to have lower peak coercivities ($\sim$30 mT).
+This result is consistent with an interpretation that a population of magnetic grains with more multidomain-like behavior is responsible for the non-ideal paleointensity behaviors during experiments on such specimens.
 
@@
 To assess whether our paleointensity estimates are biased by remanence anisotropy, we calculated the gamma statistic, which is the angular difference between the last pTRM step of paleointensity experiment and the applied field direction.
-The results show that the anorthosite specimens used for paleointensity experiment have low gamma values ranging from $\sim$0.9\textdegree to $\sim$11.9\textdegree, with a median value of 4.2\textdegree (Table S1).
+The results show that the anorthosite specimens used for paleointensity experiment have low gamma values ranging from $\sim$0.9\textdegree to $\sim$11.9\textdegree, with a median value of 4.2\textdegree$\;$(Table S1).
 These gamma values are similar to those of Midcontinent Rift volcanics \cite{Sprain2018a}.
@@
 
-We applied full TRMs to a set of our anorthosite specimens that pass the paleointensity selection criteria with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axes.
+We applied full TRMs to a set of anorthosite specimens with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axes.
 The results show that our anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions (ref.
@@
 
-% Further, we applied full TRMs to the anorthosite specimens that pass the paleointensity selection criteria after paleointensity experiment.
-With an applied field of 30 $\mu$T oriented toward the specimen at 0, 30, 60, 90, 120, and 150 degrees, the anorthosite specimens usually acquired consistent full TRMs.
-The most significant variations observed in acquired full TRMs reach to only XXX \%, much less than that has been reported to occurred within layerd anorthositic rocks in Stillwater (up to $\sim$60 \%, \cite{Selkin2000a}).
+\subsection*{Averaging secular variations and applying cooling rate correction}
+To best characterize the geomagnetic axial dipole field intensity during a certain time period a paleointensity dataset should cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged.
+Thermal modeling results from ref.
+\citealp{Zhang2021b} suggest that the Beaver River anorthosite xenoliths were likely heated to tholeiitic magma temperature ($\sim$1100\textdegree C) and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a thousand years, partially averaging secular variation within single samples.
+Anorthosite xenoliths have paleointensity results that are consistent within small regions, but vary between localities.
+Anorthosite AX12 and AX16 were emplaced $\sim$450 meters apart and have very similar paleointensity estimates.
+$\sim$10 km to the north, anorthosite xenoliths AX11 and AX13 were also emplaced closely at a distance of $\sim$125 meters and yield consistent and distinct paleointensity estimates from those of AX12 and AX16 (Fig.
+\ref{fig:Geologic_map}).
+The data therefore capture different intervals of time during the emplacement and cooling of the Beaver River diabase sills \textit{ca.} 1092 Ma.
 
-% Ref.
-\citealp{Selkin2007a} also noted that vortex state magnetic grains and certain single domain particles can produce biased paleointensity estimates due to their acquisition of saturation remanence during laboratory in-field steps.
-According to the numerical modeling of ref.
-\citealp{Selkin2007a}, thermoremanent magnetization acquired by a population of randomly-oriented, elongate, single-domain magnetite particles are unlikely to saturate upon our applied field of 30 $\mu$T.
-Therefore, in this study we do not correct for anisotropy or nonlinear TRM to our specimens.
-
-% We applied full TRMs to our anorthosite specimens with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axis.
-The results show that our anorthosite specimens did not saturate upon an applied field of 30 $\mu$T and there is thus no need for correcting nonlinear thermal remanence acquisitions.
-% Further, we applied full TRMs to the anorthosite specimens that pass the paleointensity selection criteria after paleointensity experiment.
-With an applied field of 30 $\mu$T oriented toward the specimen at 0, 30, 60, 90, 120, and 150 degrees, the anorthosite specimens usually acquired consistent full TRMs.
-The most significant variations observed in acquired full TRMs reach to only XXX \%, much less than that has been reported to occurred within layerd anorthositic rocks in Stillwater (up to $\sim$60 \%, \cite{Selkin2000a}).
-
-% The reason that the Beaver River anorthosite xenoliths do not display strong remanence anisotropy, nor do they display nonlinear TRM acquisition problem, could be that they formed in a magmatic setting distinct from that of a layered mafic complex such as the nearby Duluth Complex, where igneous foliation can occur (Fig.
-\ref{fig:Geologic_map}).
-Petrographic images show that an equigranular texture where equidimensional, randomly oriented plagioclase often develop in the anorthosite xenoliths (Fig.
-\ref{fig:Petro_QDM}; \cite{Zhang2021b}).
-This is in contrast to the interlocking igneous foliation texture within the Anorthositic Series rocks of the \textit{ca.} 1096 Ma Duluth Complex (Fig.
-\ref{fig:Petro_QDM}).
-Further, petrographic and magnetic imaging also shows the Beaver River anorthosite xenoliths do not develop abundant exsolved magnetite and ilmenite needles within plagioclase crystals as the slowly-cooled anorthositic intrusions do, in which exsolved magnetite needles can be the dominate remanence carriers (Fig.
-\ref{fig:Petro_QDM}; \cite{Scofield1986a, Feinberg2006a}).
-The fact that all dipole features within a same Duluth Complex plagioclase crystal have indistinguishable orientations is consistent with the interpretation of there being a magnetic remanence anisotropy resulting from preferred petrofabric alignment, which could give rise to the remanence anisotropy and nonlinear TRM acquisition of layered anorthositic intrusions (Fig.
-\ref{fig:Petro_QDM}).
-
-\subsection*{Averaging secular variations and applying cooling rate correction}
-To best characterize the geomagnetic axial dipole field intensity during a certain time period requires a paleointensity dataset to cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged and the geocentric axial dipole hypothesis can be assumed.
-Thermal modeling results from ref.
-\citealp{Zhang2021b} suggest that the Beaver River anorthosite xenoliths were likely heated to tholeiitic magma temperature and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a few thousand years.
-In addition, anorthosite xenoliths have paleointensity results that are consistent within small regions but vary between localities.
-Anorthosite AX12 and AX16 were emplaced only $\sim$450 meters apart and have very similar paleointensity estimates.
-$\sim$10 km to the north, anorthosite AX11 and AX13 were also emplaced closely at a distance of $\sim$125 meters but yielding lower paleointensity estimates from those of AX12 and AX16 (Fig.
-\ref{fig:Geologic_map}).
-This is consistent with the interpretation that differential cooling history due to heterogeneity of the timing and depths of the Beaver River diabase intrusions caused the anorthosites to capture paleosecular variations \textit{ca.} 1092 Ma.
-
-Another consequence of having the diabase and anorthosite cooled slowly during emplacement is its effect on biasing the paleointensity estimates toward higher values.
-It has been shown that large differences in cooling rates between acquisition of an NRM in nature versus a pTRM in the lab can result in overestimated paleointensities for SD grains \cite{Dodson1980a, Halgedahl1980a, Nagy2021a}.
+Another consequence of the anorthosite xenoliths having cooled in the interior of thick diabase is and anorthosite is that slow cooling rates can bias paleointensity estimates toward higher values.
+Large differences in cooling rates between acquisition of an NRM in nature versus a pTRM in the lab can result in overestimated paleointensities for SD grains \cite{Dodson1980a, Halgedahl1980a, Nagy2021a}.
 From the thermal history model of ref.
@@
 We find the cooling time to be on the order of $\sim$1 kyr, which corresponds to a cooling rate of $\sim 1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
-In contrast, the lab cooling rate is much faster, at $\sim 1.3\times10^{-1}$ $^\circ$C s$^{-1}$, assuming that it takes $\sim$1000s in lab condition to cool through the same temperature interval.
+In contrast, the lab cooling rate is much faster through the same temperature interval with an estimated cooling rate of $\sim 1.3\times10^{-1}$ $^\circ$C s$^{-1}$.
 This significant cooling rate difference leads to a predicted $\sim$36\% overestimate of true ancient field following the model of ref.
@@
 
-The long-term evolution of the core's thermal regime and geodynamo has been an outstanding question in the study of Earth history.
-Progress has been made in several fronts including constraining iron and its alloy's thermal conductivity values in high-temperature, high-pressure conditions similar to that in Earth's core; constructing numerical geodynamo models to simulate geomagnetic field behaviors under various thermal regimes; and obtaining and compiling observational paleomagnetic intensity records to test hypotheses put forward by numerical models.
+%I think this section could be best arranged as:
+%- Discussion of mineral physics constraints (or lack thereof)
+%- Predictions for paleointensity signal associated with inner core nucleation (or lack thereof)
+%- Interpretations of paleomagnetic record (i.e. Bono model of monotonic decrease and then rapid increase;. Cite Biggin et al. 2015) related to inner core nucleation. 
+%- Challenges posed to this model based on our data and other low to strong field transitions.
 
-Some studies interpret that the records of the Proterozoic paleointensities hint a progressive, monotonic decay trend, indicating that the geodynamo had a progressively waning power source \cite[e.g.]{Bono2019a}.
-This trend was interpreted to support the hypothesis that Earth's inner core had not solidified and thermal convection was dissipating away its rigor during the Proterozoic.
+The crystallization of the solid inner core is an important event in the long-term evolution of Earth's core and in sustaining the geodynamo \cite{Buffett2000a}.
+The age of the inner core in thermal evolution models relies on estimates for the thermal conductivity of iron alloys at the temperatures and pressures of the core \cite{Ohta2021a}.
+Prior to studies in the last 10 years, an accepted value for this conductivity of $\sim$30 W m$^{-1}$ K$^{-1}$ was consistent with nucleation of the inner core age during the first half of Earth history \cite{Stacey2007a, Konopkova2016a}.
+Subsequently, experimental data and \textit{ab intio} simulations were interpreted to imply higher thermal conductivity values \cite{Pozzo2012a, Ohta2016a} which in turn imply a younger age of the inner core ($<$700 Ma; \cite{Labrosse2015a}).
+However, other experimental studies continue to indicate lower thermal conductivity values consistent with prior estimates \cite{Konopkova2016a, Hseih2020a} with no consensus yet emerging \cite{Williams2018a, Ohta2021a}.
+These experiments are challenging with issues such as constraining the sample thickness under high pressure and temperature conditions, the validity of applying the Wiedemann-Franz law to extrapolate thermal conductivity values based on electrical resistivity measurements \cite{Ohta2016a}, and modeling uncertainties associated with parameters used in the direct thermal conduction measurement experiment \citea{Konopkova2016a}.
+Further experiments and theory are needed to explain these contrasting results which at present leave open very different trajectories for Earth's thermal evolution.
+As a result, the age of the inner core is relatively unconstrained from a theoretical perspective.
+
+The other data type that can provide insight into the long-term history of the core's thermal regime and geodynamo are paleomagnetic data---both paleodirectional data that indicate the presence of a geomagnetic field and paleointensity data that constrain the field's strength.
+Inner core nucleation would have increased the power to the geodynamo which has the potential to manifest as an increase in Earth's surface field \cite{Davies2021a}.
+An approach combining dynamo simulations and power-based theoretical scaling relationships has predicted that progressive decaying of the field's dipole moment would be followed by a rapid increase in geomagnetic field intensity soon after the onset of inner core nucleation such that there is a minimum in dipole moment just before inner core nucleation \cite{Davies2021a}.
+Other scenarios are possible, however, such as the prediction that while power increases associated with inner core nucleation would result in increases in Earth's internal magnetic field strength, the dynamo would also become more deeply seated in the core, obscuring the increase in observed magnetic field at Earth's surface \cite{Aubert2009a, Landeau2017a}.
+Observational paleomagnetic records bears the potential for testing the different model predictions.
+
+Ref.
+\citealp{Bono2019a} based on an updated paleointensity data compilation used by ref.
+\citealp{Biggin2015a} to interpret that the records of the Proterozoic paleointensities hint a progressive, monotonic decay trend (Fig.
+\ref{fig:PINT_compilation}).
+It is interpreted that this decay is consistent with there being an absence of inner core and only a thermal dynamo driven by secular cooling was in function.
 With ultra-low paleointensity observations obtained from the ca.
-565 Ma Sept-$\hat{I}$les intrusions, ref.
-\citealp{Bono2019a} hypothesized that the inner core did not begin to nucleate before at least the late Neoproterozoic and the thermal conductivity of the core must be near the high end of viable values obtained through experiments \cite{Ohta2016a}.
-Following this hypothesis, numerical modeling with particularly high core thermal conductivity values show that a decaying geodynamo through the Proterozoic would reach its lowest strength before the emergence of the solid inner core during the Ediacaran-Cambrian \cite{Davies2021a}.
-Modeling also predicts that a rapid (within a few 10s of Myr) recovery in geomagnetic field intensity occurred soon after the onset of inner core nucleation, as compositional convection started to provide additional power to drive the geodynamo \cite{Driscoll2016a}.
-
-However, the high-quality, consistent paleointensity estimates from the well-preserved ca.
-1092 Ma anorthosite xenoliths of the intracontinental Midcontinent Rift confirm the existence of a high geomagnetic field in the late Mesoproterozoic.
+565 Ma Sept-$\hat{I}$les intrusions (Fig.
+\ref{fig:PINT_compilation}), ref.
+\citealp{Bono2019a} hypothesized that the inner core did not begin to nucleate before the Ediacaran and favor a high core thermal conductivity similar to those derived through some experimental studies \cite[][e.g.]{Ohta2016a}.
+Anomalous paleodirectional data including large directional excursions \cite{Halls2015a} and frequent polarity reversals \cite{Kodama2021a} obtained from this time are also interpreted to be consistent with numerical models that employ high core thermal conductivity values and predict waning dominance of the dipole component of the geomagnetic field in the Ediacaran-Cambrian \cite{Driscoll2016a}.
+However, the high-quality, consistent paleointensity estimates from the intracontinental Midcontinent Rift rocks challenge the hypothesized monotonic decay of geodynamo throughout the Proterozoic.
+The well-preserved ca.
+1092 Ma anorthosite xenoliths of the Beaver River diabase confirm the existence of a high geomagnetic field in the late Mesoproterozoic (Fig.
+\ref{fig:PINT_compilation}).
 Together with previous records obtained from the ca.
@@
 Besides the high geomagnetic fields recorded by the Midcontinent Rift rocks, the ca.
-755 Ma Mundine well dikes also record a strong geomagnetic field in the Neoproterozoic with an average virtual dipole field of $\sim$60 ZAm$^2$ (Fig.
+755 Ma Mundine well dikes also record a strong geomagnetic field in the Neoproterozoic with an average virtual dipole field of $\sim$60 ZAm$^2$ (ref.
+\citealp{Lloyd2021b}; Fig.
 \ref{fig:PINT_compilation}).
-These observations are inconsistent with there being a monotonic decay during the Proterozoic as predicted by magnetic scaling and thermal evolution models using high core thermal conductivity values \cite{Aubert2009a, Davies2021a}.
+These observed values are inconsistent with predictions of models based on simple thermal evolution histories \cite{Bono2019a, Davies2021a}.
 
-Numerical simulations that apply magnetic scaling law also predict that there was a strong increase in true geomagnetic dipole moment during the early stages of inner core growth \cite{Davies2021a}.
+The observed paleointensity variations from the Midcontinent Rift also question whether a rapid geomagnetic field intensity increase is unambiguously correlated to the onset of inner core nucleation.
 The updated paleointensity data compilation shows that rapid transitions from low to high field intensities do not occur on an unique case during Earth history (Fig.
@@
 550 Ma volcanics rocks of the Skinner Cove Formation \cite{Thallner2021a} (Fig.
-\ref{fig:PINT_compilation}).
-That transition has been interpreted to be caused by the initiation of the nucleation of the solid inner core \cite{Thallner2021a} .
+\ref{fig:PINT_compilation}), That transition has been interpreted to be caused by the initiation of the nucleation of the solid inner core \cite{Thallner2021a} .
 However, in the Mesoproterozoic, a six-fold increase in geomagnetic dipole moment is observed within a short period ($\sim$40 Myr) from a very low average virtual dipole moment of $\sim$13 ZAm$^2$ recorded by the ca.
@@
 On the other hand, particularly low paleointensity values similar to those observed in the Ediacaran has been obtained at multiple times before and after the Ediacaran-Cambrian.
-In the late Mesoproterozoic, besides the the \textit{ca.} low intensities recorded by the ca.
+In the late Mesoproterozoic, besides the low intensities recorded by the ca.
 1140 Ma Abitibi dikes \cite{Macouin2003a}, one ca.
@@
 If the weak paleointensities in the Ediacaran or the Devonian are attributed to the geodynamo being on the brink of collapse just prior to the nucleation of the solid inner core, such data would suggest that the thermal conductivity of the core must be extremely high \cite{Davies2015a, Ohta2016a}.
-Alternatively, these Proterozoic and Paleozoic records could be a manifestation of cyclic forcing of the geodynamo across the core–mantle boundary with a time scale on the order of 150 to 200 Myr \cite{Courtillot1987a, Hawkins2021a}, a feature that has been proposed to be associated with the Wilson cycle and magnetic superchrons, and perhaps postdate the emergence of the solid inner core \cite{Courtillot1987a, Courtillot2007a}.
 
-Another alternative explanation for the low paleointensities in the Paleozoic and Proterozoic is that they could be a result of short-term geomagnetic field fluctuations.
-A five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years \cite{Bono2021a}, when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed.
+% Alternatively, these Proterozoic and Paleozoic records could be a manifestation of stochastic forcing of the core and mantle with the presence of an inner core.
+
+An alternative explanation for the low paleointensities in the Paleozoic and Proterozoic is that they could be a result of short-term (perhaps $<$10 Myr) geomagnetic field fluctuations.
+A five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years \cite{Bono2021a}, when we know a present-day-like geodynamo likely powered by both thermal and compositional convection existed.
 The observation of weak fields before and after the Cambrian indicate that the geomagnetic field may excurse toward particularly low strengths on narrow time scales regardless of the existence of a solid inner core.
-Therefore, low paleointensity values or short-term fluctuations may not be sufficient for interpreting Earth's inner core status through history.
+Therefore, low paleointensity values or short-term fluctuations may not be sufficient evidence for interpreting Earth's inner core status through history.
 
@@
 The multiple occurrences of low geomagnetic field intensities during the Proterozoic and the Paleozoic indicate that a waning thermal convection regime in Earth's before the emergence of the solid inner core may not be the solely possible mechanism for triggering an weak dipole field.
-
-Constraining the age of the inner core remains to be an open question given the uncertainties in measuring the thermal conductivity values of iron alloys in lab using different approaches \cite{Konopkova2016a, Ohta2016a}.
-Experimental challenges such as measuring sample thicknesses under high-pressure, high-temperature conditions, together with theoretical difficulties of the validity of applying the Wiedemann-Franz law to extrapolate thermal conductivity values based on electrical resistivity measurements \cite{Ohta2016a}, and modeling uncertainties associated with parameters used in the direct thermal conduction measurement experiment \citea{Konopkova2016a} cast uncertainty in the age of the inner core by hundreds of millions of years.
-Until further experiments and theory that can explain these contrasting results are developed, our knowledge on the age of Earth's thermal evolution remain largely uncertain.
 
@@
 Sun compass orientations were preferentially used for determining the sample azimuth.
-Sister specimens underwent step-wise alternating field (AF) or thermal demagnetization at the UC Berkeley Paleomagnetism Lab to isolate paleomagnetic directions (data presented in \citeA{Zhang2021b}).
+Sister specimens underwent step-wise alternating field (AF) or thermal demagnetization at the UC Berkeley Paleomagnetism Lab to isolate paleomagnetic directions (data presented in \cite{Zhang2021b}).
 Based on the anorthosite thermal demagnetization results, we selected sites whose unblocking temperature ranges are narrow and near 580\textdegree C for paleointensity experiments.
@@
 All heating steps were performed in air.
-The temperature increments for the experiments were chosen to isolate magnetizations held by (titano)magnetite informed by the previous demagnetization data, with smaller increments performed close to the expected unblocking temperature of $\sim$580\textdegree C \cite{Zhang2021b}.
+The temperature increments for the experiments were chosen to isolate magnetizations held by (titano)magnetite informed by the previous demagnetization data, with smaller increments performed close to the expected unblocking temperature of $\sim$580\textdegree C.
 
@@
 \citealp{Kirschvink1980a}) of $<$20\textdegree; (2) scatter parameter ($\beta$; ref.
-\citealp{Coe1978a}) values of $<$15 per cent; (3) a deviation angle (DANG; ref.
+\citealp{Coe1978a}) values of $<$15$\%$; (3) a deviation angle (DANG; ref.
 \citealp{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC; ref.
@@
 
-\acknow{Project research was supported by NSF CAREER grant EAR-1847277 to N.L.S.-H.
-and an Institute on Lake Superior Geology Student Research Fund grant to Y.Z.
+\acknow{Project research was supported by NSF CAREER grant EAR-1847277 to N.L.S.-H., an Institute on Lake Superior Geology Student Research Fund grant to Y.Z., and an IRM Visiting Student Fellowship to Y.Z.
 Permits for fieldwork and sampling from the Minnesota Department of Natural Resources are gratefully acknowledged.
@@
 We thank Jim Miller for providing us with the Duluth Complex anorthosite thin section.
-We thank Dario Bilardello, Peat Solheid, Mike Jackson, Josh Feinberg, and Bruce Moskowitz for their tremendous help of instrumental operations, data interpretations, and research guidance.
-Y.Z.
-thanks the IRM for the U.S.
-Visiting Student Fellowship.
-Paleomagnetic data associated with this study are available within the MagIC database (\url{}; UPDATE WHEN DOI IS GENERATED) and all data are within a Github repository associated with this work (\url{}) that is also archived on Zenodo (INSERT URL AT TIME OF PROOFS).
+We thank Dario Bilardello, Peat Solheid, Mike Jackson, Josh Feinberg, and Bruce Moskowitz for their tremendous help of instrumental operations, data interpretations, and research guidance at the IRM.
+Conversations with Bruce Buffett and Zachary Geballe informed our perspectives on the geodynamic and mineral physics literature.
+Paleomagnetic data associated with this study are available within the MagIC database (\url{}; \textit{provide link when doi is generated at time of proofs}) and all data are within a Github repository associated with this work (\url{https://earthref.org/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}) that is also archived on Zenodo (\textit{provide link when doi is generated at time of proofs}).
 This repository also contains Python code related to calculations, visualizations and statistical tests discussed herein.}
@@
 % Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be published in raw format and will not be edited or composed.
+
+%EXTRA TEXT
+
+% Without an inner core, Earth's magnetic field would have been generated solely through CMB heat flow, with perhaps additional stirring by other power sources such as Mg precipitation \cite{ORourke2016a}, or by exotic processes such as a convecting basal magma ocean \cite{Ziegler2013a}.
+Long-term secular cooling of Earth has the potential to have decreased the power available to the geodynamo through time prior to the nucleation of the inner core \cite[e.g][]{Davies2021a}.
+As a result, it is possible that intensity of the geomagnetic axial dipole field would have experienced a long-term decay before the emergence of the solid inner core after which it would have strengthened \cite{Aubert2009a, Davies2021a}.
+However, the estimates for thermal conductivity values for the core continue to vary \cite{Konopkova2016a, Ohta2016a, Zhang2020a}.
+Obtaining an observational record of the timing of inner core formation would constrain the thermal evolution of the core and mantle, in addition to improving our understanding of the history of Earth’s dynamo.
+
+% Interpretations of selected records of the current Precambrian absolute paleointensity database (compiled in the PINT database; \url{http://earth.liv.ac.uk/pint/}) suggest that an appreciable geomagnetic field existed in the Archean but reached its weakest in the Ediacaran when very low virtual dipole moment (VDM) estimates have been obtained \cite{Bono2019a, Shcherbakova2019a, Thallner2021b, Lloyd2021a}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
+Ref.
+\citealp{Bono2019a} adopted a billion-year-scale view of such observations to suggest that these data are consistent with thermal evolution models that use high core thermal conductivity values, where the nucleation of the inner core did not begin to provide compositional convection power for driving the geodynamo until the Ediacaran-Cambrian, suggesting that the Proterozoic dynamo might have experienced a monotonic decay when compositional convection was lacking and thermal convection was progressively decreasing \cite{Driscoll2016a, Davies2021a}.
+
+% Other studies focusing on paleomagnetic directions have also found that the late Ediacaran might have been a period of time when the geomagnetic field were experiencing frequent reversals and waning of dipole dominance \cite{Bono2015a, Kodama2020a}.
+
+% However, previous paleointensity data from intrusive and extrusive rocks in the \textit{ca.} 1.1 Ga Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record VDM values similar to that of today, which are higher than the values predicted by the long term decay curve proposed by ref.
+\citealp{Bono2019a} (Fig.
+\ref{fig:PINT_compilation}).
+Ref.
+\citealp{Pesonen1983a} obtained an average virtual dipole moment of 93 $\times 10^{21}$ Am$^2$ (ZAm$^2$) after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using both extrusive and intrusive rocks from the Midcontinent Rift, including the \textit{ca.} 1108-1105 Ma Osler Volcanics, the \textit{ca.} 1100 Ma Maimainse Point volcanics, the \textit{ca.} 1107 Ma intrusive Logan Sills, Thunder Bay dikes, the Baraga and Marquette dikes, as well as lava flows of the \textit{ca.} 1096 Ma North Shore Volcanic Group (Fig.
+\ref{fig:Geologic_map}).
+Ref.
+\citealp{Kulakov2013a} obtained an average VDM of 59 $\pm$ 11 ZAm$^2$ using the \textit{ca.} 1085 Ma Lake Shore Traps volcanics and suggest that the present-day-like geomagnetic field strength likely existed through the late stage of Midcontinent Rift magmatism \cite{Vervoort2007a, Miller2013a}.
+However, non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results obtained by ref.
+\citealp{Pesonen1983a, Kulakov2013a}.
+One typical failure of paleointensity experiment is the double-slope behavior, as two distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
+Ref \citealp{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas ref.
+\citealp{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
+Results with such non-ideal behavior were rejected by ref.
+\citealp{Sprain2018a} who applied stricter paleointensity selection criteria on new paleointensity results and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
+A calculated mean VDM value of 56 $\pm$ 21 ZAm$^2$ is similar to the average value of the past 300 million years \cite{Sprain2018a}.
+
+% Although it is typical of paleointensity experiments to have low success rates, bulk plagioclase-rich rocks \cite{Selkin2000a} and single plagioclase crystals \cite{Tarduno2005a} are found to be able to produce high quality paleointensity results.
+It has been suggested that plagioclase crystals can protect magnetic minerals that exsolved within them against post-formation weathering and thermochemical alteration during paleointensity experiments \cite{Tarduno2005a}.
+In this study, we present new paleointensity results from IZZ modified absolute paleointensity experiments \cite{Yu2004a} on a suite of bulk samples from \textit{ca.} 1092 Ma anorthosite xenoliths and their Beaver River diabase host rocks of the Beaver Bay Complex within the Midcontinent Rift.
+Paleointensity experiments on the anorthosite xenoliths have a very high success rate.
+Their consistent specimen- and site-level paleointensity results together with rock magnetic data reveal that the anorthosite xenoliths have low anisotropy of thermal remanent magnetization (TRM) and can acquire TRM linearly.
+Magnetic imaging shows the anorthosite specimens have dominant magnetic carriers interstitial to and within plagioclase crystals and they do not display strong preferred orientations.
+Our single-slope, high quality paleointensity data add to the reliable paleointensity records during the $\sim$25 Myr of Midcontinent Rift magmatic activity.
+Our new data from the Midcontinent Rift intrusive rocks support previous data from the volcanics that there was a strong late Mesoproterozoic geodynamo.
+
+% The Mesoproterozoic North American Midcontinent Rift is an intracontinental rift that did not result in the break-up of Laurentia.
+Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; refs.
+\citealp{Vervoort2007a, Miller2013a}) was characterized by rapid and voluminous emplacement of intrusive rocks and extrusive volcanics.
+The $<$500 Kyr emplacement of the \textit{ca.} 1096 Ma Duluth Complex and the associated $\sim$8 km thick extrusive volcanics is one manifestation \cite{Swanson-Hysell2021a}.
+The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex in northern Minnesota punctuates another such period of magmatism during the Midcontinent Rift main stage activity.
+The hypabyssal intrusion of the 1091.7 $\pm$ 0.2 Ma Beaver River diabase dikes and sills of the Beaver Bay Complex transported numerous anorthosite xenoliths that have short-axis diameters up to 180 meters via wide conduits \cite{Boerboom2004a, Boerboom2006b}.
+Geochronology, geochemical, and paleomagnetic data support the hypothesis that the diabase extruded to the surface and fed the coeval voluminous Greenstone Flow of the Portage Lake volcanics \cite{Doyle2016a, Zhang2021b}.
+
+% The dominantly monomineralic anorthosite xenoliths of the Beaver River diabase have an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}, higher than the average of $\sim$60\% of the Anorthositic Series rocks of the Duluth Complex \cite{Miller1990a}.
+The plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture showing igneous foliations (Fig.
+\ref{fig:Petro_QDM}A), but the more calcium-rich plagioclase of the Beaver River anorthosites often develop granoblastic texture characterized by equigranular crystals with poor alignments (Fig.
+\ref{fig:Petro_QDM}B).
+Abundant magnetite and ilmenite needles that are typically 10 $\mu$m long can be found to have preferred orientations within Duluth Complex plagioclase.
+Interstitial titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex plagioclase \cite{Miller1990a}.
+In contrast, titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent but relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
+\ref{fig:Petro_QDM}).
+This reflects that the Beaver River diabase magma had a different oxidation history than that of the Duluth Complex.
+The plagioclase and oxide mineral textures of the Beaver River anorthosites also distinguish them from plagioclase cumulates of other layered mafic complex where magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
+
+% Such a high concentration of Fe oxides carrying magnetic remanence in the slowly cooled Duluth Complex anorthositic rocks reflect that iron in the plagioclase mush in the magma chamber acted as a means of buffering the melt with respect to oxygen, forming crystallites along preexisting plagioclase substrate or as iron oxides incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures \cite{Scofield1986a}.
+In contrast, titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent but relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
+\ref{fig:Petro_QDM}).
+% the Beaver River diabase rapidly ascended and entrained the anorthosite xenoliths directly into shallow chambers and likely did not reach  \cite{Flower1980a, Miller1997a, Zhang2021b}.
+Particularly, \cite{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
+This is consistent with the interpretation that these anorthosite were rapidly transported by upward currents into hypabyssal magma chambers without reaching structural equilibrium.
+Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Fig.
+\ref{Petro_QDM}).
+Instead, relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
+\ref{Petro_QDM}).
+
+% In the Beaver River anorthosite xenoliths, relatively sparse titano-magnetite occur within or interstitial to the plagioclase (Fig. \ref{Petro_QDM}).
+
+% Occurrences of magnetite and ilmenite symplectites and needles within anorthositic rocks have been described by \cite{Scofield1975a, Scofield1986a} in the Wichita Mountains and by \cite{Haggerty1976a} in other slowly cooled, mafic intrusions.
+\cite{Scofield1986a} hypothesized that the inclusions acted as a means of buffering the melt with respect to oxygen, forming either crystallites along preexisting plagioclase substrate or as iron incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures.
+Such a model requires that plagioclase, instead of other mafic minerals such as olivine, to be the first to crystallize from a magma and act as a sink for iron.
+The two-stage formation of the Duluth Complex Anorthositic Series fits this model \cite{Miller1990a}.
+During the first stage, high pressure at subcrustal depths (15-25 km depth) caused fractionation of Duluth Complex magma resulting in the formation of floating plagioclase mushes \cite{Kushiro1980a}.
+The second stage of crystallization then occurs at low pressures after the magma intruded into shallower chambers.
+
+% This proposed magmatic evolution of the Duluth Complex is in distinct contrast to the emplacement of the Beaver River diabase which rapidly ascended and entrained the anorthosite xenoliths directly into shallow chambers \cite{Flower1980a, Miller1997a, Zhang2021b}.
+Particularly, \cite{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
+This is consistent with the interpretation that these anorthosite were rapidly transported by upward currents into hypabyssal magma chambers without reaching structural equilibrium.
+Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Fig.
+\ref{Petro_QDM}).
+Instead, relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
+\ref{Petro_QDM}).
+
+% it is the most probable that the anorthosite xenoliths of the Beaver River diabase are cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
+Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths acquired thermal remanent magnetizations during cooling of the diabase \cite{Zhang2021b}.
+Step-wise thermal demagnetizations show many anorthosite xenoliths have minimal secondary remanence and single-component characteristic remanence magnetizations that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C.
+
+% Geochronological, paleomagnetic, geochemical, and petrologic data are consistent with the hypothesis that the Beaver River diabase magma extruded to the surface, feeding the massive Greenstone Flow - one of the largest lava flows on Earth \cite{Doyle2016a, Zhang2021b}.
+
+% Based on high-precision $^{206}$Pb/$^{238}$U geochronology and paleomagnetic data, \cite{Swanson-Hysell2021a} hypothesized that the massive Duluth Complex and the coeval North Shore Volcanic group (Fig.
+\ref{fig:Geologic_map}) formed as a result of upwelling mantled-derived plume encountering Laurentia's thinned lithosphere, following an ``upside-down” drainage" which rapidly funneled melt to today's northeastern Minnesota (Fig.
+\ref{fig:Geologic_map}) \cite{Sleep1997a, Swanson-Hysell2014a}).
+The plagioclase-phyric Anorthositic Series rocks, the crystal-poor Layered Series intrusions of the Duluth Complex, and the coeval volcanics of the North Shore Volcanic Group (NSVG) that outcrop along Northeastern Minnesota (Fig.
+\ref{fig:Geologic_map}) are interpreted to be a result of such vigorous magmatic activity.
+
+% Despite that the anorthosite xenoliths of the Beaver River diabase were entrained as xenoliths and emplaced only $\sim$4 Myr after the formation of the Anorthositic Series rocks of the Duluth Complex when massive plagioclase mushes were generated, $^{206}$Pb/$^{238}$U zircon geochronology data and zircon rare earth element diffusion modeling results suggest it is the most probable that the Beaver River diabase anorthosite xenoliths are entrained cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
+The anorthosite xenoliths of the \textit{ca.} 1092 Ma Beaver River diabase has distinct mineralogy from that of the Anorthositic Series rocks of the Duluth Complex.
+The dominantly monomineralic anorthosite xenoliths have an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}, higher than the average of $\sim$60\% of the Anorthositic Series rocks of the Duluth Complex \cite{Miller1990a}.
+In addition, petrographic imaging support that the plagioclase from the Beaver River anorthosite formed under a magmatic conditions different from the Duluth Complex anorthositic series magma.
+While the plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture (Fig.
+\ref{Petro_QDM}A), the more calcium-rich plagioclase of the Beaver River anorthosite often develop granoblastic texsture characterized by equigranular plagioclase crystals (Fig.
+\ref{Petro_QDM}B).
+Abundant magnetite and ilmenite needles that are typically 10 $\mu$m long can be found to have preferred orientations within Duluth Complex plagioclase.
+Interstitial vermicular titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex Anorthositic Series rocks \cite{Miller1990a}.
+
+% Paleointensity experiment results from the anorthosite xenoliths and their host diabase show three general types of dominant specimen behaviors on NRM/TRM plots (Fig.
+\ref{fig:IZZI_examples}): dominantly single-slope, double-slope or sagging, and zigzagging.
+
+% As is shown in Fig.
+\ref{fig:IZZI_examples}, specimen BD2-5b has a high-temperature slope with zigzagging behavior, poor pTRM checks, and low NRM/TRM ratio, in addition to a low-temperature slope with a steep NRM/TRM ratio.
+Although the BD2 specimens' lower temperature fraction of the plots are straight and the pTRM checks are consistent, lower temperature slopes from BD2 give consistently higher paleointensity estimates than those from site AX6, AX8, and AX10, which are anorthosite xenoliths included within BD2 (Fig.
+\ref{fig:Geologic_map}; ref.
+\citealp{Zhang2021a}.
+This higher estimate from the diabase lithology could be attributed to the systematic bias of using the lower temperature slope of the NRM/TRM plots demonstrated in Fig.
+\ref{fig:IZZI_examples} and specimens from site BD2 all fail the criteria of having a maximum magnetization gap between adjacent steps of 25\% of the total NRM.
+Eventually, we only accept the single-slope anorthosite specimens results for further analyses.
+
+% \begin{figure}[h!]
+% \noindent\includegraphics[width=\textwidth]{Cooling_rate_correction.pdf}
+% \centering
+% \caption{\small{}}
+% \label{fig:cooling_correction}
+% \end{figure}
+
+% A 20 mT alternating field treatment after in-field heating steps were applied to some specimens in an attempt to mitigate non-ideal behavior contributed by multidomain or non-uniformly magnetized magnetic carriers.
+The treatment did not make any diabase specimen to pass the selection criteria.
+On the other hand, 27 out of 58 ($\sim$46.6\%) anorthosite specimens that had AF treatment passed selection and 13 out of 28 ($\sim$46.4\%) specimens without AF treatment passed selection.
+Paleointensity estimates from specimens with or without AF treatment in anorthosite AX10, AX11, AX12, AX13, AX16 do not show significant difference (Fig.
+\ref{fig:PINT_cooling_corrected}).
+Specimen results for AX11 particularly illustrates a case where all specimens within the site yield almost the same paleointensity estimate, regardless of the AF treatment.
+
+% Therefore, the alternating field demagnetization did not result in significant increase in specimen success rate for the anorthosites although it might have improved the specimen behavior for those anorthosites that were subjected to this treatment.
+Together with the narrow range of unblocking temperatures of their sister specimens \cite{Zhang2021b}, these data are consistent with the interpretation that the anorthosites with successful paleointensity results have a magnetic mineral assemblage that is likely dominated by single domain, low-Ti magnetite.
+Some anorthosite specimens and all diabase specimens still failed during result selections even though AF treatment was applied.
+It could be that they have a very high dominance of multidomain (MD) or vortex state grains with relatively moderate coercivity such that an alternating field with 20 mT peak field did not significantly mitigate their non-ideal paleointensity behaviors.
+Another cause for their failure could be the thermochemical alterations that happened during heating in air.
+As can be exampled by the zigzagging high temperature slope with poor pTRM checks in paleointensity plot for specimen BD2-5b (Fig.
+\ref{fig:IZZI_examples}), the decrease in NRM/TRM ratios during higher temperature steps likely resulted from the formation of new ferromagnetic material within the specimen due to oxidation.
+Such alterations can allow specimens to acquire additional remanence during in-field steps that may not be fully removed by an AF treatment.
+
+% Further, we applied full TRMs to the anorthosite specimens that pass the paleointensity selection criteria after paleointensity experiment.
+With an applied field of 30 $\mu$T oriented toward the specimen at 0, 30, 60, 90, 120, and 150 degrees, the anorthosite specimens usually acquired consistent full TRMs.
+The most significant variations observed in acquired full TRMs reach to only XXX \%, much less than that has been reported to occurred within layerd anorthositic rocks in Stillwater (up to $\sim$60 \%, \cite{Selkin2000a}).
+
+% Ref.
+\citealp{Selkin2007a} also noted that vortex state magnetic grains and certain single domain particles can produce biased paleointensity estimates due to their acquisition of saturation remanence during laboratory in-field steps.
+According to the numerical modeling of ref.
+\citealp{Selkin2007a}, thermoremanent magnetization acquired by a population of randomly-oriented, elongate, single-domain magnetite particles are unlikely to saturate upon our applied field of 30 $\mu$T.
+Therefore, in this study we do not correct for anisotropy or nonlinear TRM to our specimens.
+
+% We applied full TRMs to our anorthosite specimens with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axis.
+The results show that our anorthosite specimens did not saturate upon an applied field of 30 $\mu$T and there is thus no need for correcting nonlinear thermal remanence acquisitions.
+% Further, we applied full TRMs to the anorthosite specimens that pass the paleointensity selection criteria after paleointensity experiment.
+With an applied field of 30 $\mu$T oriented toward the specimen at 0, 30, 60, 90, 120, and 150 degrees, the anorthosite specimens usually acquired consistent full TRMs.
+The most significant variations observed in acquired full TRMs reach to only XXX \%, much less than that has been reported to occurred within layerd anorthositic rocks in Stillwater (up to $\sim$60 \%, \cite{Selkin2000a}).
+
+% The reason that the Beaver River anorthosite xenoliths do not display strong remanence anisotropy, nor do they display nonlinear TRM acquisition problem, could be that they formed in a magmatic setting distinct from that of a layered mafic complex such as the nearby Duluth Complex, where igneous foliation can occur (Fig.
+\ref{fig:Geologic_map}).
+Petrographic images show that an equigranular texture where equidimensional, randomly oriented plagioclase often develop in the anorthosite xenoliths (Fig.
+\ref{fig:Petro_QDM}; \cite{Zhang2021b}).
+This is in contrast to the interlocking igneous foliation texture within the Anorthositic Series rocks of the \textit{ca.} 1096 Ma Duluth Complex (Fig.
+\ref{fig:Petro_QDM}).
+Further, petrographic and magnetic imaging also shows the Beaver River anorthosite xenoliths do not develop abundant exsolved magnetite and ilmenite needles within plagioclase crystals as the slowly-cooled anorthositic intrusions do, in which exsolved magnetite needles can be the dominate remanence carriers (Fig.
+\ref{fig:Petro_QDM}; \cite{Scofield1986a, Feinberg2006a}).
+The fact that all dipole features within a same Duluth Complex plagioclase crystal have indistinguishable orientations is consistent with the interpretation of there being a magnetic remanence anisotropy resulting from preferred petrofabric alignment, which could give rise to the remanence anisotropy and nonlinear TRM acquisition of layered anorthositic intrusions (Fig.
+\ref{fig:Petro_QDM}).

### 2022-01-27 PNAS_manuscript/PNAS_manuscript.tex
Updates from Overleaf

@@
 
-\title{High geomagnetic field intensity recorded by anorthosite xenoliths requires a vigorous late Mesoproterozoic geodynamo}
+\title{High geomagnetic field intensity recorded by anorthosite xenoliths requires a strongly powered late Mesoproterozoic geodynamo}
 
@@
 These anorthosite xenoliths from the North American Midcontinent Rift have experienced minimal alteration since their emplacement and yield high quality paleointensity estimates.
-These data record a strong present-day-like virtual dipole moment 1.1 billion years ago.
-A strong field persisted over a 25 million year or greater interval indicating the existence of appreciable power sources for Earth's dynamo during the late Mesoproterozoic.}
+In contrast to predictions of a progressively decaying field, these data record a strong virtual dipole moment 1.1 billion years ago.
+A strong field persisted over at least a 20-million-year interval indicating the existence of appreciable power sources for Earth's dynamo during the late Mesoproterozoic.}
 
@@
 and N.L.S.-H.
-wrote the paper.}
+wrote the paper with input from M.S.A.
+and R.R.F.}
 
@@
 % \equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: yimingzhang\@berkeley.edu.}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: yimingzhang\@berkeley.edu}
 
@@
 Obtaining estimates of Earth's magnetic field strength in deep time is complicated by non-ideal rock magnetic behavior in many igneous rocks.
-In this study, we target anorthosite xenoliths that cooled and acquired their magnetization within \textit{ca.} 1092 Ma shallowly emplaced diabase intrusions of the North American Midcontinent Rift.
+In this study, we target anorthosite xenoliths that cooled and acquired their magnetization within ca.
+1092 Ma shallowly emplaced diabase intrusions of the North American Midcontinent Rift.
 In contrast to the diabase which fails to provide reliable paleointensity estimates, the anorthosite xenoliths are unusually high-fidelity recorders yielding high-quality, single-slope paleointensity results that are consistent at specimen and site levels.
-An average value of $\sim$89 ZAm$^2$ for the virtual dipole moment from the anorthosites is higher than that of the dipole component of Earth's magnetic field today.
-Such a strong late Mesoproterozoic dipole recorded by the anorthosite xenoliths requires the existence of a vigorous geodynamo at the time.
-Together with previous paleointensity data from other Midcontinent Rift rocks, our new results indicate that a dynamo with strong power sources persisted for more than 20 million years \textit{ca.} 1.1 Ga.
+Consistent and high site-mean values (including one site-mean of 54.33$\pm$0.04 $\mu$T) and an average value of $\sim$89 ZAm$^2$ for the virtual dipole moment from the anorthosites is higher than that of the dipole component of Earth's magnetic field today.
+Such high intensities recorded by the anorthosite xenoliths requires the existence of a strongly powered geodynamo at the time.
+Together with previous paleointensity data from other Midcontinent Rift rocks, our new results indicate that a dynamo with strong power sources persisted for more than 20 million years ca.
+1.1 Ga.
 These data are inconsistent with there being a progressive monotonic decay of Earth's dynamo strength through the Proterozoic Eon.
@@
 Higher conductivity values imply faster cooling rates, which in turn imply that the threshold for the freezing of the inner core happened more recently \cite{Davies2015a}.
-While some estimates of these values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates have implied higher thermal conductivity values and an age for the inner core that is less than 1.3 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b}, with some suggesting even younger ages ($<$ 700 Ma; ref.
-\citealp{Labrosse2015a, Ohta2016a}).
+While some estimates of these values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates have implied higher thermal conductivity values and an age for the inner core that is less than 1.3 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b}, with some suggesting even younger ages ($<$ 700 Ma;  \citealp{Labrosse2015a, Ohta2016a}).
 Given that estimates for the core's thermal conductivity continue to be debated, it is crucial to use observational records as an independent constraint on the thermal evolution of Earth's core and mantle.
@@
 Paleomagnetic records from ancient rocks are one of the few types of observational data that have the potential to provide constraints on the thermal evolution of Earth’s core.
-Evidence for a persistent magnetic field through the Proterozoic, for example, necessitates the existence of plate tectonics that sustained core-mantle boundary heat flow \cite{Swanson-Hysell2021c}.
-However, strikingly low estimates of geomagnetic field strengths have been obtained during the 635 to 539 Ma Ediacaran Period \cite{Bono2019a, Shcherbakova2019a,Thallner2021b} and the 419 to 359 Ma Devonian Period \cite{Shcherbakova2017a, Shcherbakova2021a}, potentially indicating unusual periods of core dynamo regime at those times.
+Evidence for a persistent magnetic field through the Proterozoic, for example, likely necessitates the existence of plate tectonics that sustained core-mantle boundary heat flow \cite{Swanson-Hysell2021c}.
+However, strikingly low estimates of geomagnetic field strengths have been obtained ca.
+565 Ma during the Ediacaran Period \cite{Bono2019a, Shcherbakova2019a,Thallner2021b} and ca.
+400 Ma during the Devonian Period \cite{Shcherbakova2017a, Shcherbakova2021a, Hawkins2021a}, potentially indicating unusual periods of core dynamo regime at those times.
 The Ediacaran data have been interpreted to indicate that there was a progressively decaying field up to that time that was followed soon after by initial crystallization of the inner core \cite{Bono2019a}.
@@
 Purple stars mark sites with paleointensity results that passed the selection criteria from this study.
-Paleomagnetic sites from ref.
-\citealp{Pesonen1983a} are categorized by lithology.
-All sites from ref.
-\citealp{Pesonen1983a, Kulakov2013a, Sprain2018a} are color-coded by their ages.}}
+Paleomagnetic sites from  \citealp{Pesonen1983a} are categorized by lithology.
+All sites from  \citealp{Pesonen1983a, Kulakov2013a, Sprain2018a} are color-coded by their ages.}}
 \label{fig:Geologic_map}
@@
 
-Determinations of the absolute value of the ancient geomagnetic field strength rely on igneous rocks that acquire thermal remanent magnetizations as they cool.
+Determinations of the absolute value of ancient geomagnetic field strength rely on igneous rocks that acquire thermal remanent magnetizations as they cool.
 These magnetizations need to be unmodified by subsequent heating or chemical alteration in order to maintain a record of the ancient geomagnetic field from the time of cooling.
@@
 
-Despite the excellent preservation of the rocks, non-ideal paleointensity behaviors have challenged the interpretation of many previous paleointensity results from the Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a}.
-The most trusted type of paleointensity estimate is that obtained through experiments in which the primary natural thermal remanent magnetization is progressively replaced by a laboratory magnetization that is imparted in a known field with internal consistency checks (such as in IZZI-style Thellier experiments; ref.
-\citealp{Yu2004a}).
+Despite the excellent preservation of the rocks, non-ideal paleointensity behaviors have challenged the interpretation of many previous paleointensity results from the Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a}.
+The most trusted type of paleointensity estimate is that obtained through experiments in which the primary natural thermal remanent magnetization is progressively replaced by a laboratory magnetization that is imparted in a known field with internal consistency checks (such as in IZZI-style Thellier experiments;  \citealp{Yu2004a}).
 In such Thellier paleointensity experiments, one typical departure from ideal behavior is sagging or double-slopes as visualized in plots that show thermal remanent magnetization (TRM) acquired versus natural remanent magnetization (NRM) lost.
@@
 Typically, such non-ideal behavior would result in a higher paleointensity estimate from the steeper-sloped low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
-For example, in data from the Midcontinent Rift, ref.
-\citealp{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas ref.
-\citealp{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
-Results with such non-ideal behavior were rejected by ref.
-\citealp{Sprain2018a} who applied stricter paleointensity selection criteria, but as a result had few accepted sites.
+For example, in data from the Midcontinent Rift,  \citealp{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas  \citealp{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
+Results with such non-ideal behavior were rejected by \citealp{Sprain2018a} who applied stricter paleointensity selection criteria, but as a result had few accepted sites.
 
@@
 They cooled and acquired their magnetizations in conjunction with the host diabase.
-Paleointensity experiments on the anorthosite xenoliths have a very high success rate, yielding consistent specimen- and site-level paleointensity results.
+Paleointensity experiments on the anorthosite xenoliths have a high success rate, yielding consistent specimen- and site-level paleointensity results.
 Rock magnetic data reveal that the anorthosite xenoliths have low anisotropy of thermal remanent magnetization (TRM) and can acquire TRM linearly within relevant field strengths.
-Magnetic imaging shows the anorthosite specimens have dominant magnetic carriers interstitial to and within plagioclase crystals and they do not display strong preferred orientations.
-These single-slope, high quality paleointensity data, together with other high-quality paleointensity records during $\sim$25 Myr of Midcontinent Rift magmatic activity, require that there was a strong late Mesoproterozoic geodynamo.
+Magnetic imaging shows the anorthosite specimens have dominant magnetic carriers within and interstitial to plagioclase crystals, and they do not display strong preferred orientations.
+These single-slope, high quality paleointensity data, together with other high-quality paleointensity records during the protracted Midcontinent Rift magmatic activity, require that there was a strong late Mesoproterozoic geodynamo.
 
@@
 While magmatic activity within the Midcontinent Rift was protracted, there were intervals of particularly rapid volcanism and voluminous emplacement of intrusions.
-The \textit{ca.} 1092 Ma Beaver Bay Complex in northern Minnesota punctuates one such period of magmatism during the main stage of Midcontinent Rift activity.
+The ca.
+1092 Ma Beaver Bay Complex in northern Minnesota punctuates one such period of magmatism during the main stage of Midcontinent Rift activity.
 The magma that formed the 1091.7 $\pm$ 0.2 Ma Beaver River diabase dikes and sills of the Beaver Bay Complex transported numerous anorthosite xenoliths that have short-axis diameters up to 180 meters via wide conduits \cite{Boerboom2004a, Boerboom2006b}.
@@
 Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths to have acquired thermal remanent magnetizations while cooling with the Beaver River diabase \cite{Zhang2021b}.
-Step-wise thermal demagnetization data show the anorthosite xenoliths have dominantly single-component magnetization that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C, consistent with them being held by low titanium titanomagnetite \cite{Zhang2021b}.
+Step-wise thermal demagnetization data show the anorthosite xenoliths to have dominantly single-component magnetization that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C, consistent with magnetization held by low-titanium titanomagnetite \cite{Zhang2021b}.
 
@@
 \noindent\includegraphics[width=13 cm]{Petro_QDM.png}
-\caption{\footnotesize{Thin section petrographic images (A,B,C,D) and magnetic maps (E,F) of anorthosite from the Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (left column; ref.
-\citealp{Zhang2021b}) and a distinct anorthosite from the Duluth Complex Anorthositic Series (right column).
+\caption{\footnotesize{Thin section petrographic images (A,B,C,D) and magnetic maps (E,F) of anorthosite from the Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (left column;  \citealp{Zhang2021b}) and a distinct anorthosite from the Duluth Complex Anorthositic Series (right column).
 The Duluth Complex anorthosites were not targeted for paleointensity experiments in this study given the complexities associated with more pronounced fabrics.
-Cross-polarized petrographic images of the Beaver River anorthosite (A,C) reveal plagioclase with a granoblastic texture with platioclase crystals to are largely free of large opaque inclusions.
-In contrast, plagioclase crystals in the Duluth Complex anorthosite have euhedral, interlocking crystals with an igneous foliation (B) and the plagioclase crystals contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the [010] planes of the crystals.
-Magnetic field maps acquired by the quantum diamond microscope (QDM) show that the relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite in (E) while large oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the overlapping magnetic fields above the oxides in (F).
+Cross-polarized petrographic images of the Beaver River anorthosite (A,C) reveal plagioclase with a granoblastic texture with crystals that are largely free of large opaque inclusions.
+In contrast, plagioclase crystals in the Duluth Complex anorthosite have euhedral, interlocking crystals with an igneous foliation (B) and the plagioclase crystals contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the [001] axis of the plagioclase.
+Magnetic field maps acquired with a quantum diamond microscope (QDM) show that the relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite in (E) while large oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the overlapping magnetic fields above the oxides in (F).
 Magnetic field maps of (E) and (F) share the same spatial and magnetic scale.
@@
 The dominantly monomineralic anorthosite xenoliths within the Beaver River diabase often have granoblastic texture characterized by equigranular crystals with weakly developed petrofabrics (Fig.
-\ref{fig:Petro_QDM}A,C).
-In addition, the plagioclase crystals of the anorthosite xenoliths are typically free of large oxide inclusions.
-Magnetic imaging using the quantum diamond microscope (QDM) shows that the dominant magnetic remanence carriers often exist within, and sometimes interstitial to, plagioclase crystals (Fig.
+\ref{fig:Petro_QDM}A).
+In addition, the plagioclase crystals of the anorthosite xenoliths are typically free of large oxide inclusions (Fig.
+\ref{fig:Petro_QDM}C).
+Magnetic imaging using a quantum diamond microscope (QDM) shows that the dominant magnetic remanence carriers often exist within, and sometimes interstitial to, plagioclase crystals (Fig.
 \ref{fig:Petro_QDM}E).
 
-These features of the Beaver River anorthosite xenoliths are distinct from other anorthosites such as the anorthosite xenoliths within the Duluth Complex Anorthositic Series rocks---older intrusions within the Midcontinent Rift.
+These features of the Beaver River anorthosite xenoliths are distinct from many other anorthosites such as the anorthosite xenoliths within the Duluth Complex Anorthositic Series rocks---older intrusions within the Midcontinent Rift.
 To illustrate these difference, we present petrographic and magnetic imaging data from a sample of Duluth Complex anorthosite that was not targeted for paleointensity experiments.
@@
 \ref{fig:Petro_QDM}B).
-In addition, there are abundant Fe-Ti oxide needles that are typically aligned with the [010] planes of the plagioclase crystals (Fig.
+In addition, there are abundant Fe-Ti oxide needles that are typically aligned with the [001] axes of the plagioclase crystals (Fig.
 \ref{fig:Petro_QDM}D).
-Magnetic imaging confirms that those strongly magnetic needles within the plagioclase crystals often have magnetic moments oriented along their long axes (Fig.
+Magnetic imaging confirms that these strongly magnetic needles within the plagioclase crystals often have magnetic moments oriented along their long axes (Fig.
 \ref{fig:Petro_QDM}D,F).
-Titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex plagioclase \cite{Miller1990a}.
+Titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex magma \cite{Miller1990a}.
 
@@
 \ref{fig:PINT_cooling_corrected} (and provided in Table S1).
-The paleointensity quality index (Q$_{PI}$; ref.
-\citealp{Biggin2014a}) for the anorthosite xenoliths are all above 4 (Table S2).
-The cooling rate-corrected absolute paleointensity estimates from all anorthosite sites have a weighted mean of 52.72 $\pm$ 5.72 $\mu$T.
-This corresponds to a calculated virtual dipole moment of $\sim$89 ZAm$^2$ ($10^{21} Am^2$) \textit{ca.} 1092 Ma.
-All measurement-level paleointensity experiment data is available within the MagIC database (\url{https://earthref.org/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}; \textit{this private link is provided for the purpose of review; the url will be updated when a doi is generated for this manuscript}).
+The paleointensity quality index (Q$_{PI}$;  \citealp{Biggin2014a}) for the anorthosite xenoliths are all 5 or 6 (Table S2).
+The cooling rate-corrected absolute paleointensity estimates from all anorthosite sites have a mean of 38.28 $\pm$ 11.92 $\mu$T.
+This paleointensity value corresponds to a calculated virtual dipole moment of $\sim$89 ZAm$^2$ ($10^{21} Am^2$) ca.
+1092 Ma.
+All measurement-level paleointensity experiment data are available within the MagIC database (\url{https://earthorg/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}; \textit{this private link is provided for the purpose of review; the url will be updated when a doi is generated for this manuscript}).
 
@@
 \centering
-\caption{\footnotesize{Example results of paleointensity experiments are displayed on NRM/TRM plots and zero-field heating results are shown on inset orthogonal plots (Zijderveld plots) for anorthosite and diabase specimens.
+\caption{\footnotesize{Example results of paleointensity experiments are displayed on Arai plots and zero-field heating results are shown on inset orthogonal plots (Zijderveld plots) for anorthosite and diabase specimens.
 Red (blue) circles indicate zero-field/in-field (in-field/zero-field) steps `ZI’ (`IZ’).
-Triangles mark partial thermal remanent (pTRM) checks.
+Triangles mark partial thermal remanent magnetization (pTRM) checks.
 Blue and red squares in the Zijderveld plots are X–Y and X–Z projections, respectively, of the NRMs in specimen coordinates.
@@
 The green lines represent fits for the dominant single-slope component that passes the acceptance criteria and gives an estimate of the ancient field strength (B$_{anc}$).
-The plots for anorthosite specimens AX1-1a and diabase BD5-1b on the bottom row show typical non-ideal sagging and zigzagging behaviour that fails our acceptance criteria.
-Specimen BD2-5a is an example where the data appear linear with distinct slopes in the low and high temperature ranges such that it could pass less restrictive selection criteria, particularly if a more narrow temperature range was used for the experiment.}}
+The plots for anorthosite specimens AX1-1a and diabase BD5-1b on the bottom row show typical non-ideal sagging behaviour that fails our acceptance criteria.
+Specimen BD2-5a is an example where the data appear linear with distinct slopes in the low and high temperature ranges such that it could pass less restrictive selection criteria, particularly if a narrower temperature range was used for the experiment.}}
 \label{fig:IZZI_examples}
@@
 We accept specimen- and site-level absolute paleointensity results from those anorthosite xenoliths that pass the selection criteria.
-Other anorthosite xenoliths and diabase specimens failed the selection criteria largely because of double-slope or sagging behavior (fail FRAC selection; see methods), poor pTRM checks, and sometimes zigzagging behaviors superimposed on top of sagging behavior (fail SCAT, DANG selection; see methods; Fig.
+Other anorthosite xenoliths and diabase specimens failed the selection criteria largely because of double-slope or sagging behavior (fail FRAC selection; see Materials and Methods section), poor pTRM checks, and sometimes zigzagging behaviors superimposed on top of sagging behavior (fail SCAT, DANG selection; see Materials and Methods section; Fig.
 \ref{fig:IZZI_examples}).
@@
 \centering
-\caption{\small{Summary plot of individual specimen absolute paleointensity results (square and diamond symbols) from this study and their averages at site (grey bar) and locality level (weighted mean by site-level standard deviations; orange bar).
+\caption{\small{Summary plot of individual specimen absolute paleointensity results (square symbols) from this study and their averages at site (black bars with grey uncertainty boxes) and locality level (weighted mean by site-level standard deviations; orange bar).
 All results are corrected for cooling rate with a factor of 0.74.
@@
 
-In addition to estimating paleointensity values by introducing a set of selection criteria to filter our experiment results, we apply an independent method of ref.
-\citealp{Cych2021a} which uses all experimental data regardless of their NRM/TRM plot statistics to perform bias corrected estimation of paleointensity (BiCEP).
+In addition to estimating paleointensity values by introducing a set of selection criteria to filter our experiment results, we apply an independent method of \citealp{Cych2021a} which uses all experimental data regardless of their NRM/TRM plot statistics to perform bias corrected estimation of paleointensity (BiCEP).
 This Bayesian probabilistic method is based on an assumption that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Paterson2011a}.
@@
 
-Overall, the high-quality experiment results from the anorthosite xenoliths of the Beaver River diabase and the consistent paleointensity estimates based on independent methods confirm that the anorthosite xenoliths record a high geomagnetic field \textit{ca.} 1092 Ma.
+Overall, the high-quality results from the anorthosite xenoliths of the Beaver River diabase and the consistent paleointensity estimates based on independent methods confirm that the anorthosite xenoliths record a high geomagnetic field ca.
+1092 Ma.
 
@@
 Magnetic property measurement system (MPMS) data show both diabase and anorthosite contain (titano)magnetite as evidenced through the presence of the Verwey transition (Fig.
-S2; ref.
-\citealp{Verwey1939a, Feinberg2015a}).
-Anorthosite chips from sites that yield successful paleointensity results have Verwey transition temperatures near 120 K as expected for stochiometric magnetite with minimal Ti \cite{Ozdemir1993a}.
+S2;  \citealp{Verwey1939a, Feinberg2015a}).
+Anorthosite specimens from sites that yield successful paleointensity results have Verwey transition temperatures near 120 K as expected for stochiometric magnetite with minimal Ti \cite{Ozdemir1993a}.
 However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have Verwey transitions that are suppressed toward lower temperatures (Fig.
-S2), indicating that the magnetite in the specimens either has relatively higher Ti content or has been partially oxidized \cite{Ozdemir1993a}.
+S2), indicating that magnetite grains in the specimens either have relatively higher Ti content or have been partially oxidized \cite{Ozdemir1993a}.
 
@@
 A compilation of median destructive field (MDF) values in Fig.
-\ref{fig:coercivity} shows that specimens from anorthosites that pass paleointensity selections have a distinctly higher average MDF values than other anorthosites and diabase.
+\ref{fig:coercivity} shows that specimens from anorthosites that pass paleointensity selection criteria have distinctly higher average MDF values than other anorthosite and diabase specimens.
 Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity around 80 mT (Fig.
@@
 Significant remanence anisotropy has been documented to exist within certain anorthositic rocks that form in layered intrusive complexes \cite{Selkin2000a, Feinberg2006a}.
-Strong remanence anisotropy associated with the igneous foliation developed within anorthosite from the Stillwater Complex has been shown to lead to significant overestimate or underestimate of paleointensity values depending on the relative orientations between the fabrics and an applied magnetic field \cite{Selkin2000a}.
+Strong remanence anisotropy associated with the igneous foliation developed within anorthosite from the Stillwater Complex has been shown to lead to significant overestimation or underestimation of paleointensity values depending on the relative orientations between the fabrics and an applied magnetic field \cite{Selkin2000a}.
 To assess whether our paleointensity estimates are biased by remanence anisotropy, we calculated the gamma statistic, which is the angular difference between the last pTRM step of paleointensity experiment and the applied field direction.
@@
 We applied full TRMs to a set of anorthosite specimens with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axes.
-The results show that our anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions (ref.
-\citealp{Selkin2007a}; Fig.
+The results show that our anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions ( \citealp{Selkin2007a}; Fig.
 S4).
@@
 To best characterize the geomagnetic axial dipole field intensity during a certain time period a paleointensity dataset should cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged.
-Thermal modeling results from ref.
-\citealp{Zhang2021b} suggest that the Beaver River anorthosite xenoliths were likely heated to tholeiitic magma temperature ($\sim$1100\textdegree C) and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a thousand years, partially averaging secular variation within single samples.
+Thermal modeling results from \citealp{Zhang2021b} suggest that the Beaver River anorthosite xenoliths were heated to tholeiitic magma temperature ($\sim$1100\textdegree C) and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a thousand years, partially averaging secular variation within single samples.
 Anorthosite xenoliths have paleointensity results that are consistent within small regions, but vary between localities.
-Anorthosite AX12 and AX16 were emplaced $\sim$450 meters apart and have very similar paleointensity estimates.
-$\sim$10 km to the north, anorthosite xenoliths AX11 and AX13 were also emplaced closely at a distance of $\sim$125 meters and yield consistent and distinct paleointensity estimates from those of AX12 and AX16 (Fig.
-\ref{fig:Geologic_map}).
-The data therefore capture different intervals of time during the emplacement and cooling of the Beaver River diabase sills \textit{ca.} 1092 Ma.
+Anorthosite AX12 and AX16 were emplaced $\sim$450 meters apart and have very similar paleointensity estimates (Fig.
+\ref{fig:Geologic_map, fig:PINT_cooling_corrected}).
+$\sim$10 km to the north, anorthosite xenoliths AX11 and AX13 were also emplaced closely at a distance of $\sim$125 meters and yield similar but distinct paleointensity estimates from those of AX12 and AX16 (Fig.
+\ref{fig:Geologic_map, fig:PINT_cooling_corrected}).
+The data therefore capture different intervals of time during the emplacement and cooling of the Beaver River diabase sills ca.
+1092 Ma.
 
-Another consequence of the anorthosite xenoliths having cooled in the interior of thick diabase is and anorthosite is that slow cooling rates can bias paleointensity estimates toward higher values.
-Large differences in cooling rates between acquisition of an NRM in nature versus a pTRM in the lab can result in overestimated paleointensities for SD grains \cite{Dodson1980a, Halgedahl1980a, Nagy2021a}.
-From the thermal history model of ref.
-\citealp{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully blocked in their characteristic natural remanence magnetization ($\sim$500\textdegree C; ref.
-\citealp{Zhang2021b}).
-We find the cooling time to be on the order of $\sim$1 kyr, which corresponds to a cooling rate of $\sim 1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
-In contrast, the lab cooling rate is much faster through the same temperature interval with an estimated cooling rate of $\sim 1.3\times10^{-1}$ $^\circ$C s$^{-1}$.
-This significant cooling rate difference leads to a predicted $\sim$36\% overestimate of true ancient field following the model of ref.
-\citealp{Halgedahl1980a} (Fig.
+Another consequence of the anorthosite xenoliths having cooled in the interior of thick diabase intrusions is that slow cooling rates can bias paleointensity estimates toward higher values.
+Large differences in cooling rates between acquisition of an NRM in nature versus a TRM in the lab can result in overestimated paleointensities for SD grains \cite{Dodson1980a, Halgedahl1980a, Nagy2021a}.
+From the thermal history model of  \citealp{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully blocked in their characteristic natural remanence magnetization ($\sim$500\textdegree C;  \citealp{Zhang2021b}).
+We find the cooling time to be on the order of $\sim$1 kyr, which corresponds to a cooling rate of $\sim1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
+In contrast, the lab cooling rate is much faster through the same temperature interval with an estimated cooling rate of $\sim1.3\times10^{-1}$ $^\circ$C s$^{-1}$.
+The significant cooling rate difference leads to a predicted $\sim$36\% overestimate of true ancient field following the model of \citealp{Halgedahl1980a} (Fig.
 S3).
-This estimate on cooling rate effect is similar to the value of $\sim$35\% overestimate derived from the model of ref.
-\citealp{Nagy2021a}.
+This estimate on cooling rate effect is similar to the value of $\sim$35\% overestimate derived from the model of \citealp{Nagy2021a}.
 We therefore correct our paleointensity results by a factor of 0.74.
@@
 \centering
-\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment values from the PINT database (PINT v8.0.0; \url{http://www.pintdb.org/}; ref.
-\citealp{Bono2021a}) with additional Neoproterozoic data from refs.
+\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment values from the PINT database (PINT v8.0.0; \url{http://www.pintdb.org/};  \citealp{Bono2021a}) with additional Neoproterozoic data from refs.
 \citealp{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}.
@@
 
-\section*{Proterozoic geodynamo}
+\section*{Discussion}
 
-%I think this section could be best arranged as:
+%The outline for this section is::
 %- Discussion of mineral physics constraints (or lack thereof)
@@
 The age of the inner core in thermal evolution models relies on estimates for the thermal conductivity of iron alloys at the temperatures and pressures of the core \cite{Ohta2021a}.
-Prior to studies in the last 10 years, an accepted value for this conductivity of $\sim$30 W m$^{-1}$ K$^{-1}$ was consistent with nucleation of the inner core age during the first half of Earth history \cite{Stacey2007a, Konopkova2016a}.
-Subsequently, experimental data and \textit{ab intio} simulations were interpreted to imply higher thermal conductivity values \cite{Pozzo2012a, Ohta2016a} which in turn imply a younger age of the inner core ($<$700 Ma; \cite{Labrosse2015a}).
-However, other experimental studies continue to indicate lower thermal conductivity values consistent with prior estimates \cite{Konopkova2016a, Hseih2020a} with no consensus yet emerging \cite{Williams2018a, Ohta2021a}.
-These experiments are challenging with issues such as constraining the sample thickness under high pressure and temperature conditions, the validity of applying the Wiedemann-Franz law to extrapolate thermal conductivity values based on electrical resistivity measurements \cite{Ohta2016a}, and modeling uncertainties associated with parameters used in the direct thermal conduction measurement experiment \citea{Konopkova2016a}.
+Prior to studies in the last 10 years, an accepted value of $\sim$30 W m$^{-1}$ K$^{-1}$ for this thermal conductivity was used to constrain the timing of inner core nucleation to be during the first half of Earth history \cite{Stacey2007a, Konopkova2016a}.
+Subsequently, experimental data and \textit{ab intio} simulations were interpreted to imply higher thermal conductivity values \cite{Pozzo2012a, Ohta2016a} which in turn imply a younger age of the inner core ($<$700 Ma; \citealp{Labrosse2015a}).
+However, other experimental studies continue to indicate lower thermal conductivity values consistent with prior estimates \cite{Konopkova2016a, Hsieh2020a} with no consensus yet emerging \cite{Williams2018a, Ohta2021a}.
+These experiments are challenging with issues such as constraining the sample thickness under high pressure and temperature conditions, the validity of applying the Wiedemann-Franz law to extrapolate thermal conductivity values based on electrical resistivity measurements \cite{Ohta2016a}, and modeling uncertainties associated with parameters used in direct thermal conduction measurement experiments \cite{Konopkova2016a}.
 Further experiments and theory are needed to explain these contrasting results which at present leave open very different trajectories for Earth's thermal evolution.
@@
 
-The other data type that can provide insight into the long-term history of the core's thermal regime and geodynamo are paleomagnetic data---both paleodirectional data that indicate the presence of a geomagnetic field and paleointensity data that constrain the field's strength.
+The other data type that can provide insight into the long-term history of the core's thermal regime and geodynamo is paleomagnetic data---both paleodirectional data that indicate the presence of a geomagnetic field and paleointensity data that constrain the field's strength.
 Inner core nucleation would have increased the power to the geodynamo which has the potential to manifest as an increase in Earth's surface field \cite{Davies2021a}.
-An approach combining dynamo simulations and power-based theoretical scaling relationships has predicted that progressive decaying of the field's dipole moment would be followed by a rapid increase in geomagnetic field intensity soon after the onset of inner core nucleation such that there is a minimum in dipole moment just before inner core nucleation \cite{Davies2021a}.
-Other scenarios are possible, however, such as the prediction that while power increases associated with inner core nucleation would result in increases in Earth's internal magnetic field strength, the dynamo would also become more deeply seated in the core, obscuring the increase in observed magnetic field at Earth's surface \cite{Aubert2009a, Landeau2017a}.
-Observational paleomagnetic records bears the potential for testing the different model predictions.
+An approach combining dynamo simulations and power-based theoretical scaling relationships has predicted that progressive decay of the field's dipole moment would be followed by a rapid increase in geomagnetic field intensity soon after the onset of inner core nucleation such that there is a minimum in dipole moment just before inner core nucleation \cite{Davies2021a}.
+Other scenarios are possible, however, such as the prediction that while power increases associated with inner core nucleation would result in increases in Earth's internal magnetic field strength, the dynamo would also become more deeply seated in the core, diminishing the increase in observed magnetic field at Earth's surface \cite{Aubert2009a, Landeau2017a}.
+Observational paleomagnetic records hold the potential for testing different model predictions and identifying transitions in ancient field strength.
 
-Ref.
-\citealp{Bono2019a} based on an updated paleointensity data compilation used by ref.
-\citealp{Biggin2015a} to interpret that the records of the Proterozoic paleointensities hint a progressive, monotonic decay trend (Fig.
-\ref{fig:PINT_compilation}).
-It is interpreted that this decay is consistent with there being an absence of inner core and only a thermal dynamo driven by secular cooling was in function.
-With ultra-low paleointensity observations obtained from the ca.
-565 Ma Sept-$\hat{I}$les intrusions (Fig.
-\ref{fig:PINT_compilation}), ref.
-\citealp{Bono2019a} hypothesized that the inner core did not begin to nucleate before the Ediacaran and favor a high core thermal conductivity similar to those derived through some experimental studies \cite[][e.g.]{Ohta2016a}.
-Anomalous paleodirectional data including large directional excursions \cite{Halls2015a} and frequent polarity reversals \cite{Kodama2021a} obtained from this time are also interpreted to be consistent with numerical models that employ high core thermal conductivity values and predict waning dominance of the dipole component of the geomagnetic field in the Ediacaran-Cambrian \cite{Driscoll2016a}.
-However, the high-quality, consistent paleointensity estimates from the intracontinental Midcontinent Rift rocks challenge the hypothesized monotonic decay of geodynamo throughout the Proterozoic.
-The well-preserved ca.
-1092 Ma anorthosite xenoliths of the Beaver River diabase confirm the existence of a high geomagnetic field in the late Mesoproterozoic (Fig.
+It has been proposed that Proterozoic paleointensity data are consistent with a progressive monotonic decay leading up to ca.
+565 Ma in the Ediacaran Period (Fig.
+\ref{fig:PINT_compilation}, \citealp{Bono2019a}).
+This interpretation was motivated by paleointensity estimates developed from the ca.
+565 Ma Sept-$\hat{I}$les layered mafic intrusive complex of $\sim$7 ZAm$^2$ that are among the lowest values in the paleointensity database (Fig.
+\ref{fig:PINT_compilation}), \citealp{Bono2019a}.
+A decay in the lead-up to this time was argued to be consistent with an absence of an inner core and a dynamo to which progressively less power was available through secular cooling \cite{Bono2019a, Davies2021a}.
+This timing of inner core formation would favor a high core thermal conductivity \cite[e.g.][]{Ohta2016a}.
+Paleomagnetic directional excursions \cite{Halls2015a} and frequent polarity reversals \cite{Kodama2021a} during this time are interpreted to be consistent with numerical simulations \cite{Driscoll2016a} associated with a weak dipole field.
+
+The high paleointensity estimates from the Midcontinent Rift rocks challenge the hypothesized monotonic decay of the strength of the geomagnetic field throughout the Proterozoic Era.
+Instead, the well-preserved ca.
+1092 Ma anorthosite xenoliths of the Beaver River diabase record a strong geomagnetic field in the late Mesoproterozoic that exceeds the strength of the modern-day field for which crystallization of the inner core is a power source (Fig.
 \ref{fig:PINT_compilation}).
 Together with previous records obtained from the ca.
-1106 Ma Osler Volcanics of the Midcontinent Rift \cite{Sprain2018a}, these data indicate an appreciable power source for Earth's dynamo could have persisted through at least 14 Myr during the Mesoproterozoic (Fig.
+1106 Ma Osler Volcanics of the Midcontinent Rift \cite{Sprain2018a}, these data indicate that appreciable power to Earth's dynamo persisted through at least 14 Myr during the late Mesoproterozoic to maintain a strong surface field (Fig.
 \ref{fig:PINT_compilation}).
-Besides the high geomagnetic fields recorded by the Midcontinent Rift rocks, the ca.
-755 Ma Mundine well dikes also record a strong geomagnetic field in the Neoproterozoic with an average virtual dipole field of $\sim$60 ZAm$^2$ (ref.
-\citealp{Lloyd2021b}; Fig.
+In addition to these high geomagnetic fields recorded by Midcontinent Rift rocks, the ca.
+755 Ma Mundine Well dikes \cite{Lloyd2021b} also require a stronger geomagnetic field in the Neoproterozoic than would be predict by a progressive Proterozoic decline (Fig.
 \ref{fig:PINT_compilation}).
-These observed values are inconsistent with predictions of models based on simple thermal evolution histories \cite{Bono2019a, Davies2021a}.
 
-The observed paleointensity variations from the Midcontinent Rift also question whether a rapid geomagnetic field intensity increase is unambiguously correlated to the onset of inner core nucleation.
-The updated paleointensity data compilation shows that rapid transitions from low to high field intensities do not occur on an unique case during Earth history (Fig.
+The hypothesis that a weak Ediacaran geomagnetic field is a telltale sign of the lack of an inner core with core nucleation following shortly thereafter may predict that it is the most significant weak to strong field transition in the paleointensity record.
+However, Fig.
+\ref{fig:PINT_compilation} shows that transitions from low to high field intensities occurred before, during, and after the Ediacaran.
+In the Ediacaran record developed to date, there is a two-fold increase in Earth's virtual dipole moment when comparing estimates from the ca.
+565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a} to those from ca.
+550 Ma volcanics of the Skinner Cove Formation \cite{Thallner2021a} (Fig.
 \ref{fig:PINT_compilation}).
-A two-fold increase in Earth's virtual dipole moment in the Ediacaran has been recorded by the ca.
-565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a} and the ca.
-550 Ma volcanics rocks of the Skinner Cove Formation \cite{Thallner2021a} (Fig.
-\ref{fig:PINT_compilation}), That transition has been interpreted to be caused by the initiation of the nucleation of the solid inner core \cite{Thallner2021a} .
-However, in the Mesoproterozoic, a six-fold increase in geomagnetic dipole moment is observed within a short period ($\sim$40 Myr) from a very low average virtual dipole moment of $\sim$13 ZAm$^2$ recorded by the ca.
+In the late Mesoproterozoic, there is at least a six-fold increase within a period of $\sim$40 Myr from a low average virtual dipole moment of $\sim$13 ZAm$^2$ recorded by the ca.
 1140 Ma Abitibi dikes which yielded straight Arai plots \cite{Macouin2003a}, to a high moment of $\sim$70 ZAm$^2$ recorded by the ca.
-1106 Ma Osler Volcanics, which might have lasted through ca.
-1092 Ma when the Beaver River anorthosite xenoliths recorded a virtual dipole moment of $\sim$80 ZAm$^2$.
-This increase in absolute value of virtual dipole moment dwarfs all other existing records of geomagnetic field fluctuations in the Precambrian, including those in the Ediacaran (Fig.
+1106 Ma Osler Volcanics, with even stronger values recorded ca.
+1092 Ma by Beaver River anorthosite xenoliths that record virtual dipole moments up to $\sim$140 ZAm$^2$.
+While the Abitibi dikes paleointensity estimates do not go to values as low as the Sept-$\hat{I}$les intrusions, this virtual dipole moment increase in the Mesoproterozoic is the largest yet documented in the Precambrian (Fig.
 \ref{fig:PINT_compilation}).
-The tempo and scale of this field intensity transition may also be a better match with that predicted by numerical simulations upon the onset of inner core nucleation \cite{Davies2021a} but this similarity could be coincidental.
+The tempo and scale of this field intensity transition could match with model-based predictions associated with the onset of inner core nucleation \cite{Davies2021a}.
+This timing would be broadly consistent with that proposed by \citealp{Biggin2015a}.
+However, the model prediction would be of sustained strong field values following inner core nucleation which is challenged by data from the ca.
+1070 Ma Bangemall Sills which include a sill with a low virtual dipole moment of $\sim$6.4 ZAm$^2$ (Fig.
+\ref{fig:PINT_compilation};  \citealp{Lloyd2021b}).
+Following the Ediacaran, there are low paleointensity estimates from the ca.
+414 Ma Strathmore lava flows \cite{Hawkins2021a}, the ca.
+410-380 Ma lava flows of the Minusa Basinand the Kola Peninsula \cite{Shcherbakova2017a}, the ca.
+408–393 Ma Buribay volcanic complex \cite{Shcherbakova2021a}, as well as the ca.
+370 Ma dikes and lavas of the Siberian Viluy Traps \cite{Hawkins2019a} (Fig.
+\ref{fig:PINT_compilation}).
+This interval of time with low paleointensity estimates has been termed the ``Mid-Paleozoic Dipole Low'' \cite{Hawkins2021a} and is followed by high paleointensity values afterwards (Fig.
+\ref{fig:PINT_compilation}).
 
-On the other hand, particularly low paleointensity values similar to those observed in the Ediacaran has been obtained at multiple times before and after the Ediacaran-Cambrian.
-In the late Mesoproterozoic, besides the low intensities recorded by the ca.
-1140 Ma Abitibi dikes \cite{Macouin2003a}, one ca.
-1070 Ma Bangemall Sill records an extremely low virtual dipole moment of $\sim$6.4 ZAm$^2$ (Fig.
-\ref{fig:PINT_compilation}; ref.
-\citealp{Lloyd2021b}).
-In the Neoproterozoic, ca.
-720 Ma dolerite dikes of the Franklin Large Igneous Province record a mean virtual dipole moment of 11 ZAm$^2$ (Fig.
-\ref{fig:PINT_compilation}; ref.
-\citealp{Lloyd2021a}).
-In the Phanerazoic, Devonian-aged rocks from the Strathmore lava flows \cite{Hawkins2021a}, Minusa Basin, Kola Peninsula \cite{Shcherbakova2017a}, volcanic rocks from the East European Craton \cite{Shcherbakova2019a}, as well as the Buribay volcanic complex \cite{Shcherbakova2021a} show low paleointensities through an extended period of time (Fig.
+% that represent a transition to field strengths greater than the present
+
+% changes in core-mantle boundary heat flow related to plate tectonics - more general refrence to core-mantle heat flow pattern not necessarily telling about inner core nucleation.
+
+An alternative hypothesis for the multiple transitions from low to high field intensities in the middle Paleozoic and the Proterozoic is that they are manifestations of secularly changing mantle forcing of the geodynamo across the core–mantle boundary which may be unrelated to the growth of the inner core.
+The time scale over which these changes happen may be on the order of less than one million years.
+During the Pleistocene, when we known a present-day-like geodynamo existed, paleointensity records show at least a six-fold fluctuation of the geomagnetic dipole moment can happen (Fig.
 \ref{fig:PINT_compilation}).
-Accompanying the low paleointensity values, paleodirectional data from the Ediacaran-Cambrian \cite{Bono2015a, Halls2015b, Kodama2021a} and the Devonian show anomalous directional fluctuations \cite{Shcherbakova2019a} which could be consistent with there being a largely diminished dipole component of Earth's magnetic field.
-If the weak paleointensities in the Ediacaran or the Devonian are attributed to the geodynamo being on the brink of collapse just prior to the nucleation of the solid inner core, such data would suggest that the thermal conductivity of the core must be extremely high \cite{Davies2015a, Ohta2016a}.
+If the mantle forcing that drove such transitions exists regardless of the present of inner core, it could have contributed to the anomalously low or high paleointensity values observed during the Phanerozoic and Proterozoic.
+The low paleointensity records from the ca.
+720 Ma dolerite dikes of the rapidly emplaced Franklin Large Igneous Province could be one example of capturing these short-term variations and not fully averaging the secular changes of the geomagnetic field (Fig.
+\ref{fig:PINT_compilation}; \citealp{Lloyd2021a, Lloyd2021b}).
+The unknown magnitude of the forcing presents challenges for detecting the increase in dynamo strength caused by the kick-start of compositional convection driven by the crystallization of the inner core based on paleointensity data alone.
 
-% Alternatively, these Proterozoic and Paleozoic records could be a manifestation of stochastic forcing of the core and mantle with the presence of an inner core.
+% This interpretation may explain the low virtual dipole moment of 11 ZAm$^2$ recorded by the ca.
 
-An alternative explanation for the low paleointensities in the Paleozoic and Proterozoic is that they could be a result of short-term (perhaps $<$10 Myr) geomagnetic field fluctuations.
-A five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years \cite{Bono2021a}, when we know a present-day-like geodynamo likely powered by both thermal and compositional convection existed.
-The observation of weak fields before and after the Cambrian indicate that the geomagnetic field may excurse toward particularly low strengths on narrow time scales regardless of the existence of a solid inner core.
+Overall, the high-fidelity paleointensity recorders of the Beaver River anorthosite xenoliths in the well-preserved Midcontinent Rift record high paleointensities values and necessitates that there was a vigorous late Mesoproterozoic geodynamo.
+
+% In the Neoproterozoic, ca.
+720 Ma dolerite dikes of the Franklin Large Igneous Province record a mean virtual dipole moment of  In the Phanerozoic, multiple low paleointensities has been observed in Devonian-aged volcanic units (Fig.
+\ref{fig:PINT_compilation}; \citealp{Shcherbakova2017a, Shcherbakova2021a, Hawkins2021a}).
+
+%The observed paleointensity variations from the Midcontinent Rift also question the attribution to the onset of inner core nucleation as the unique cause for driving a significant paleointensity increase.
+Besides the
+
+% Accompanying the low paleointensity values, paleodirectional data from the Ediacaran-Cambrian \cite{Bono2015a, Halls2015b, Kodama2021a} and the Devonian show anomalous directional fluctuations \cite{Shcherbakova2019a} which could be consistent with there being a largely diminished dipole component of Earth's magnetic field.
+If the weak paleointensities in the Ediacaran or the Devonian are attributed to the geodynamo being on the brink of collapse just prior to the nucleation of the solid inner core.
+
+% rocks from
+
+% if one invokes inner core nucleation for all these low to high transitions, the data would suggest the inner core nucleated multiple times.
+Alternatively, it could that all except one are not related to the inner core nucleation, but merely a manifestation of short term dynamo variations.
+% n narrow time scales regardless of the existence of a solid inner core.
 Therefore, low paleointensity values or short-term fluctuations may not be sufficient evidence for interpreting Earth's inner core status through history.
@@
 % The high geomagnetic field estimates from the Midcontinent Rift volcanics and intrusives are not the only suite of high paleointensity records at odds with the hypothesis of a monotonically decaying geodynamo through the entire Proterozoic Eon \cite{Biggin2009a, Bono2019a}.
-The $\sim$60 ZAm$^2$ virtual dipole moment estimates from the \textit{ca.} 755 Ma Mundine Well dikes are much higher than the values predicted by the monotonic decrease model of \cite{Bono2019a}.
+The $\sim$60 ZAm$^2$ virtual dipole moment estimates from the ca.
+755 Ma Mundine Well dikes are much higher than the values predicted by the monotonic decrease model of \cite{Bono2019a}.
 Although one may consider these data capturing relatively short term fluctuations of the geomagnetic field, they nevertheless indicate that the geodynamo had strong power sources during the Proterozoic.
-On the other hand, besides the low paleointensity values recorded by the \textit{ca.} 1140 Ma Abitibi dikes, low estimates have been obtained from the \textit{ca.} 1070 Ma Bangemall Sills \cite{Lloyd2021b}, the \textit{ca.} 720 Ma dolerite dikes of the Franklin Large Igneous Province \cite{Lloyd2021a}, as well as Devonian-aged rocks from the Minusa Basin, Kola Peninsula, Buribay volcanic complex and volcanic rocks from the East European Craton (Fig.
-\ref{fig:PINT_compilation}; ref.
-\citealp{Shcherbakova2017a, Shcherbakova2019a, Shcherbakova2021a}).
+On the other hand, besides the low paleointensity values recorded by the ca.
+1140 Ma Abitibi dikes, low estimates have been obtained from the ca.
+1070 Ma Bangemall Sills \cite{Lloyd2021b}, the ca.
+720 Ma dolerite dikes of the Franklin Large Igneous Province \cite{Lloyd2021a}, as well as Devonian-aged rocks from the Minusa Basin, Kola Peninsula, Buribay volcanic complex and volcanic rocks from the East European Craton (Fig.
+\ref{fig:PINT_compilation};  \citealp{Shcherbakova2017a, Shcherbakova2019a, Shcherbakova2021a}).
 The multiple occurrences of low geomagnetic field intensities during the Proterozoic and the Paleozoic indicate that a waning thermal convection regime in Earth's before the emergence of the solid inner core may not be the solely possible mechanism for triggering an weak dipole field.
@@
 
-Overall, the high-fidelity paleointensity recorders of the Beaver River anorthosite xenoliths in the well-preserved Midcontinent Rift record high paleointensities values and necessitates that there was a vigorous late Mesoproterozoic geodynamo.
-
 % conclude on whether an apparent rise or decay within a narrow time period ($\sim$35 Ma) during the Proterozoic is associated with transitions in the core phase transition or is a result of short-term excursion of the geomagnetic field that may be common through much of the evolution of Earth's core.
 
-% an apparent rapid decrease in geomagnetic dipole moment from \textit{ca.} 1092 Ma to \textit{ca.} 1070 Ma can be evidenced by the contrasting high values from the Midcontinent Rift anorthosite xenoliths and the low values from the Bangemall Sills (Fig.
-\ref{fig:PINT_compilation}; ref.
-\citealp{Lloyd2021b}).
+% an apparent rapid decrease in geomagnetic dipole moment from ca.
+1092 Ma to ca.
+1070 Ma can be evidenced by the contrasting high values from the Midcontinent Rift anorthosite xenoliths and the low values from the Bangemall Sills (Fig.
+\ref{fig:PINT_compilation};  \citealp{Lloyd2021b}).
 These data suggest that an apparent transition of virtual dipole moment similar in scale with what is predicted for a dynamo powered with or without compositional convection in the core may not be unique during the Precambrian.
 In addition, given that a five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years, when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed, we cannot conclude on whether an apparent rise or decay within a narrow time period ($\sim$35 Ma) during the Proterozoic is associated with transitions in the core phase transition or is a result of short-term excursion of the geomagnetic field that may be common through much of the evolution of Earth's core.
-Recently, ultra-low paleointensities (2.01–7.07) $\mu$T from the Devonian has been found by ref.
-\cite{Shcherbakova2021a}.
+Recently, ultra-low paleointensities (2.01–7.07) $\mu$T from the Devonian has been found by  \cite{Shcherbakova2021a}.
 These data suggest that significant excursion of low geomagnetic field intensities may occur regardless of the inner core formation.
@@
 \ref{fig:PINT_compilation}).
-The \textit{ca.} 1140 Ma Abitibi dikes \cite{Macouin2003a} record an average virtual dipole moment of $\sim$13 ZAm$^2$, similar to the low values of $\sim$7 ZAm$^2$ obtained from the single silicate crystals of the \textit{ca.} 565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a}.
-However, the relatively high moments of $\sim$72 ZAm$^2$ from the \textit{ca.} 1105 Ma Osler Volcanic Group \cite{Sprain2018a} and $\sim$89 ZAm$^2$ from the \textit{ca.} 1092 Ma Beaver River anorthosite xenoliths (Fig.
+The ca.
+1140 Ma Abitibi dikes \cite{Macouin2003a} record an average virtual dipole moment of $\sim$13 ZAm$^2$, similar to the low values of $\sim$7 ZAm$^2$ obtained from the single silicate crystals of the ca.
+565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a}.
+However, the relatively high moments of $\sim$72 ZAm$^2$ from the ca.
+1105 Ma Osler Volcanic Group \cite{Sprain2018a} and $\sim$89 ZAm$^2$ from the ca.
+1092 Ma Beaver River anorthosite xenoliths (Fig.
 \ref{fig:PINT_compilation}) are about six times the values from the Abitibi dikes.
@@
 The Midcontinent Rift of North America presents an opportunity to develop robust paleointensity estimates spanning a $\sim$25 Myr interval of the late Mesoproterozoic.
-Rock magnetic data support that the \textit{ca.} 1092 Ma anorthosite xenoliths of the Midcontinent Rift diabase can have dominantly high coercivity magnetic mineralogy with minimal anisotropic or nonlinear thermal remanence acquisition, producing high quality paleointensity results.
+Rock magnetic data support that the ca.
+1092 Ma anorthosite xenoliths of the Midcontinent Rift diabase can have dominantly high coercivity magnetic mineralogy with minimal anisotropic or nonlinear thermal remanence acquisition, producing high quality paleointensity results.
 Their overall high estimates of the late Mesoproterozoic virtual dipole moment together with previous Midcontinent Rift records suggest that a strong geomagnetic field like today's might have persisted through the rifting.
@@
 A total of 86 specimens from 14 anorthosite xenoliths and a total of 69 specimens from 7 diabase sites underwent paleointensity experiments that followed the step-wise double-heating Thellier method \cite{Thellier1959a} using the IZZI protocol \cite{Yu2004a} with heating steps up to 585 \textdegree C.
-Partial thermal remanent magnetization (pTRM) checks were performed systematically throughout the experiment to test whether there was significant mineralogical alteration due to heating and were assessed using the SCAT parameter of ref.
-\citealp{Shaar2013a}.
+Partial thermal remanent magnetization (pTRM) checks were performed systematically throughout the experiment to test whether there was significant mineralogical alteration due to heating and were assessed using the SCAT parameter of  \citealp{Shaar2013a}.
 On top of the IZZI-Thellier experiment protocol, we also performed a comparative study where we added an extra step of 20 mT alternating field (AF) cleaning on some of the specimens after each in-field step.
@@
 \subsection*{Paleointensity result selection}
-The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD; ref.
-\citealp{Kirschvink1980a}) of $<$20\textdegree; (2) scatter parameter ($\beta$; ref.
-\citealp{Coe1978a}) values of $<$15$\%$; (3) a deviation angle (DANG; ref.
-\citealp{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC; ref.
-\cite{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT; ref.
-\citealp{Shaar2013a}) = TRUE; (6) a maximum magnetic moment difference between adjacent zero-field steps (GAP-Max; ref.
-\citealp{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
+The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD;  \citealp{Kirschvink1980a}) of $<$20\textdegree; (2) scatter parameter ($\beta$;  \citealp{Coe1978a}) values of $<$15$\%$; (3) a deviation angle (DANG;  \citealp{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC;  \cite{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT;  \citealp{Shaar2013a}) = TRUE; (6) a maximum magnetic moment difference between adjacent zero-field steps (GAP-Max;  \citealp{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
 The MAD measures the scatter about the best-fit line through the natural remanent magnetization (NRM) steps in the selected interval for which the intensity is defined.
@@
 Both MAD and DANG assess the directional variation of the NRM, with MAD measuring the scatter in the NRM directions and DANG assessing whether the component is trending toward the origin of the Zijderveld plot.
-$\beta$ is the ``scatter" parameter of ref.
-\cite{Coe1978a} and is the ratio of the standard error of the slope of the best-fit line of the selected NRM and pTRM points on an NRM/TRM plot to the absolute value of the slope.
+$\beta$ is the ``scatter" parameter of  \cite{Coe1978a} and is the ratio of the standard error of the slope of the best-fit line of the selected NRM and pTRM points on an NRM/TRM plot to the absolute value of the slope.
 FRAC is the fraction of the NRM that is used in the best-fit line \cite{Shaar2013a}.
@@
 $\beta$, FRAC, GAP-Max and SCAT are all statistics to assess the behavior of NRM/TRM plots.
-See the Standard Paleointensity Definitions (ref.
-\citealp{Paterson2014a}; \url{https://earthref.org/PmagPy/SPD/home.html}) for more details.
+See the Standard Paleointensity Definitions ( \citealp{Paterson2014a}; \url{https://earthorg/PmagPy/SPD/home.html}) for more details.
 Data analysis was conducted using Thellier GUI \cite{Shaar2013a} within the PmagPy software package \cite{Tauxe2016a}.
@@
 Specimen median destructive fields (MDF) are calculated based on the backfield curves.
-The calculated coercivity spectra were subsequently decomposed into one or more components using skew-normal distributions following the method of ref.
-\citealp{Maxbauer2016a} examples of which are shown in Fig.
+The calculated coercivity spectra were subsequently decomposed into one or more components using skew-normal distributions following the method of  \citealp{Maxbauer2016a} examples of which are shown in Fig.
 \ref{fig:coercivity}.
@@
 
-To further identify the magnetic carriers within the Beaver River anorthosite xenoliths and compare them with the anorthosites of the Duluth Complex Anorthositic Series rocks, we used the quantum diamond microscope (QDM) at the UC Berkeley Paleomagnetism laboratory to image a thin section of sample MS99033 from anorthosite xenolith AX16 (which yielded a $^{206}$Pb/$^{238}$U zircon date of 1091.83 $\pm$ 0.21 Ma; ref.
-\citealp{Zhang2021b}), and a thin section of a Duluth Complex anorthosite (Fig.
+To further identify the magnetic carriers within the Beaver River anorthosite xenoliths and compare them with the anorthosites of the Duluth Complex Anorthositic Series rocks, we used the quantum diamond microscope (QDM) at the UC Berkeley Paleomagnetism laboratory to image a thin section of sample MS99033 from anorthosite xenolith AX16 (which yielded a $^{206}$Pb/$^{238}$U zircon date of 1091.83 $\pm$ 0.21 Ma;  \citealp{Zhang2021b}), and a thin section of a Duluth Complex anorthosite (Fig.
 \ref{fig:Petro_QDM}).
@@
 Conversations with Bruce Buffett and Zachary Geballe informed our perspectives on the geodynamic and mineral physics literature.
-Paleomagnetic data associated with this study are available within the MagIC database (\url{}; \textit{provide link when doi is generated at time of proofs}) and all data are within a Github repository associated with this work (\url{https://earthref.org/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}) that is also archived on Zenodo (\textit{provide link when doi is generated at time of proofs}).
+Paleomagnetic data associated with this study are available within the MagIC database (\url{}; \textit{provide link when doi is generated at time of proofs}) and all data are within a Github repository associated with this work (\url{https://earthorg/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}) that is also archived on Zenodo (\textit{provide link when doi is generated at time of proofs}).
 This repository also contains Python code related to calculations, visualizations and statistical tests discussed herein.}
@@
 
-% However, previous paleointensity data from intrusive and extrusive rocks in the \textit{ca.} 1.1 Ga Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record VDM values similar to that of today, which are higher than the values predicted by the long term decay curve proposed by ref.
-\citealp{Bono2019a} (Fig.
+% However, previous paleointensity data from intrusive and extrusive rocks in the ca.
+1.1 Ga Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record VDM values similar to that of today, which are higher than the values predicted by the long term decay curve proposed by  \citealp{Bono2019a} (Fig.
 \ref{fig:PINT_compilation}).
 Ref.
-\citealp{Pesonen1983a} obtained an average virtual dipole moment of 93 $\times 10^{21}$ Am$^2$ (ZAm$^2$) after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using both extrusive and intrusive rocks from the Midcontinent Rift, including the \textit{ca.} 1108-1105 Ma Osler Volcanics, the \textit{ca.} 1100 Ma Maimainse Point volcanics, the \textit{ca.} 1107 Ma intrusive Logan Sills, Thunder Bay dikes, the Baraga and Marquette dikes, as well as lava flows of the \textit{ca.} 1096 Ma North Shore Volcanic Group (Fig.
+\citealp{Pesonen1983a} obtained an average virtual dipole moment of 93 $\times 10^{21}$ Am$^2$ (ZAm$^2$) after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using both extrusive and intrusive rocks from the Midcontinent Rift, including the ca.
+1108-1105 Ma Osler Volcanics, the ca.
+1100 Ma Maimainse Point volcanics, the ca.
+1107 Ma intrusive Logan Sills, Thunder Bay dikes, the Baraga and Marquette dikes, as well as lava flows of the ca.
+1096 Ma North Shore Volcanic Group (Fig.
 \ref{fig:Geologic_map}).
 Ref.
-\citealp{Kulakov2013a} obtained an average VDM of 59 $\pm$ 11 ZAm$^2$ using the \textit{ca.} 1085 Ma Lake Shore Traps volcanics and suggest that the present-day-like geomagnetic field strength likely existed through the late stage of Midcontinent Rift magmatism \cite{Vervoort2007a, Miller2013a}.
-However, non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results obtained by ref.
-\citealp{Pesonen1983a, Kulakov2013a}.
+\citealp{Kulakov2013a} obtained an average VDM of 59 $\pm$ 11 ZAm$^2$ using the ca.
+1085 Ma Lake Shore Traps volcanics and suggest that the present-day-like geomagnetic field strength likely existed through the late stage of Midcontinent Rift magmatism \cite{Vervoort2007a, Miller2013a}.
+However, non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results obtained by  \citealp{Pesonen1983a, Kulakov2013a}.
 One typical failure of paleointensity experiment is the double-slope behavior, as two distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
-Ref \citealp{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas ref.
-\citealp{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
-Results with such non-ideal behavior were rejected by ref.
-\citealp{Sprain2018a} who applied stricter paleointensity selection criteria on new paleointensity results and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
+Ref \citealp{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas  \citealp{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
+Results with such non-ideal behavior were rejected by  \citealp{Sprain2018a} who applied stricter paleointensity selection criteria on new paleointensity results and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
 A calculated mean VDM value of 56 $\pm$ 21 ZAm$^2$ is similar to the average value of the past 300 million years \cite{Sprain2018a}.
@@
 It has been suggested that plagioclase crystals can protect magnetic minerals that exsolved within them against post-formation weathering and thermochemical alteration during paleointensity experiments \cite{Tarduno2005a}.
-In this study, we present new paleointensity results from IZZ modified absolute paleointensity experiments \cite{Yu2004a} on a suite of bulk samples from \textit{ca.} 1092 Ma anorthosite xenoliths and their Beaver River diabase host rocks of the Beaver Bay Complex within the Midcontinent Rift.
+In this study, we present new paleointensity results from IZZ modified absolute paleointensity experiments \cite{Yu2004a} on a suite of bulk samples from ca.
+1092 Ma anorthosite xenoliths and their Beaver River diabase host rocks of the Beaver Bay Complex within the Midcontinent Rift.
 Paleointensity experiments on the anorthosite xenoliths have a very high success rate.
@@
 % The Mesoproterozoic North American Midcontinent Rift is an intracontinental rift that did not result in the break-up of Laurentia.
-Although the protracted magmatic activity in the rift lasted form \textit{ca.} 1109 Ma to \textit{ca.} 1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; refs.
+Although the protracted magmatic activity in the rift lasted form ca.
+1109 Ma to ca.
+1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; refs.
 \citealp{Vervoort2007a, Miller2013a}) was characterized by rapid and voluminous emplacement of intrusive rocks and extrusive volcanics.
-The $<$500 Kyr emplacement of the \textit{ca.} 1096 Ma Duluth Complex and the associated $\sim$8 km thick extrusive volcanics is one manifestation \cite{Swanson-Hysell2021a}.
-The emplacement of the \textit{ca.} 1092 Ma Beaver Bay Complex in northern Minnesota punctuates another such period of magmatism during the Midcontinent Rift main stage activity.
+The $<$500 Kyr emplacement of the ca.
+1096 Ma Duluth Complex and the associated $\sim$8 km thick extrusive volcanics is one manifestation \cite{Swanson-Hysell2021a}.
+The emplacement of the ca.
+1092 Ma Beaver Bay Complex in northern Minnesota punctuates another such period of magmatism during the Midcontinent Rift main stage activity.
 The hypabyssal intrusion of the 1091.7 $\pm$ 0.2 Ma Beaver River diabase dikes and sills of the Beaver Bay Complex transported numerous anorthosite xenoliths that have short-axis diameters up to 180 meters via wide conduits \cite{Boerboom2004a, Boerboom2006b}.
@@
 % Despite that the anorthosite xenoliths of the Beaver River diabase were entrained as xenoliths and emplaced only $\sim$4 Myr after the formation of the Anorthositic Series rocks of the Duluth Complex when massive plagioclase mushes were generated, $^{206}$Pb/$^{238}$U zircon geochronology data and zircon rare earth element diffusion modeling results suggest it is the most probable that the Beaver River diabase anorthosite xenoliths are entrained cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
-The anorthosite xenoliths of the \textit{ca.} 1092 Ma Beaver River diabase has distinct mineralogy from that of the Anorthositic Series rocks of the Duluth Complex.
+The anorthosite xenoliths of the ca.
+1092 Ma Beaver River diabase has distinct mineralogy from that of the Anorthositic Series rocks of the Duluth Complex.
 The dominantly monomineralic anorthosite xenoliths have an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}, higher than the average of $\sim$60\% of the Anorthositic Series rocks of the Duluth Complex \cite{Miller1990a}.
@@
 Although the BD2 specimens' lower temperature fraction of the plots are straight and the pTRM checks are consistent, lower temperature slopes from BD2 give consistently higher paleointensity estimates than those from site AX6, AX8, and AX10, which are anorthosite xenoliths included within BD2 (Fig.
-\ref{fig:Geologic_map}; ref.
-\citealp{Zhang2021a}.
+\ref{fig:Geologic_map};  \citealp{Zhang2021a}.
 This higher estimate from the diabase lithology could be attributed to the systematic bias of using the lower temperature slope of the NRM/TRM plots demonstrated in Fig.
@@
 \citealp{Selkin2007a} also noted that vortex state magnetic grains and certain single domain particles can produce biased paleointensity estimates due to their acquisition of saturation remanence during laboratory in-field steps.
-According to the numerical modeling of ref.
-\citealp{Selkin2007a}, thermoremanent magnetization acquired by a population of randomly-oriented, elongate, single-domain magnetite particles are unlikely to saturate upon our applied field of 30 $\mu$T.
+According to the numerical modeling of  \citealp{Selkin2007a}, thermoremanent magnetization acquired by a population of randomly-oriented, elongate, single-domain magnetite particles are unlikely to saturate upon our applied field of 30 $\mu$T.
 Therefore, in this study we do not correct for anisotropy or nonlinear TRM to our specimens.
@@
 \ref{fig:Petro_QDM}; \cite{Zhang2021b}).
-This is in contrast to the interlocking igneous foliation texture within the Anorthositic Series rocks of the \textit{ca.} 1096 Ma Duluth Complex (Fig.
+This is in contrast to the interlocking igneous foliation texture within the Anorthositic Series rocks of the ca.
+1096 Ma Duluth Complex (Fig.
 \ref{fig:Petro_QDM}).

### 2022-01-27 PNAS_manuscript/PNAS_manuscript.tex
update manuscript

@@
 In contrast to the diabase which fails to provide reliable paleointensity estimates, the anorthosite xenoliths are unusually high-fidelity recorders yielding high-quality, single-slope paleointensity results that are consistent at specimen and site levels.
-Consistent and high site-mean values (including one site-mean of 54.33$\pm$0.04 $\mu$T) and an average value of $\sim$89 ZAm$^2$ for the virtual dipole moment from the anorthosites is higher than that of the dipole component of Earth's magnetic field today.
+An average value of $\sim$89 ZAm$^2$ for the virtual dipole moment from the anorthosites with the highest site-level values up to $\sim$140 ZAm$^2$ are higher than that of the dipole component of Earth's magnetic field today.
 Such high intensities recorded by the anorthosite xenoliths requires the existence of a strongly powered geodynamo at the time.
@@
 Higher conductivity values imply faster cooling rates, which in turn imply that the threshold for the freezing of the inner core happened more recently \cite{Davies2015a}.
-While some estimates of these values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates have implied higher thermal conductivity values and an age for the inner core that is less than 1.3 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b}, with some suggesting even younger ages ($<$ 700 Ma;  \citealp{Labrosse2015a, Ohta2016a}).
+While some estimates of these values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates have implied higher thermal conductivity values and an age for the inner core that is less than 1.3 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b}, with some suggesting even younger ages ($<$700 Ma;  \citealp{Labrosse2015a, Ohta2016a}).
 Given that estimates for the core's thermal conductivity continue to be debated, it is crucial to use observational records as an independent constraint on the thermal evolution of Earth's core and mantle.
@@
 The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal intrusions of the Beaver River diabase \cite{Zhang2021b}.
-They cooled and acquired their magnetizations in conjunction with the host diabase.
+They cooled and acquired their magnetizations in conjunction with the host diabase at a paleolatitude of 22\textdegree $\pm$2 \textdegree (calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a,Zhang2021b}).
 Paleointensity experiments on the anorthosite xenoliths have a high success rate, yielding consistent specimen- and site-level paleointensity results.
@@
 \ref{fig:IZZI_examples}.
-Summary specimen absolute paleointensity estimates and site-level weighted mean paleointensity values are plotted in Fig.
+Summary specimen absolute paleointensity estimates and site-level mean paleointensity values are plotted in Fig.
 \ref{fig:PINT_cooling_corrected} (and provided in Table S1).
@@
 \centering
-\caption{\small{Summary plot of individual specimen absolute paleointensity results (square symbols) from this study and their averages at site (black bars with grey uncertainty boxes) and locality level (weighted mean by site-level standard deviations; orange bar).
+\caption{\small{Summary plot of individual specimen absolute paleointensity results (square symbols) from this study and their averages and standard deviations at site (black bars with grey uncertainty boxes) and locality level (orange dashed line and bar).
 All results are corrected for cooling rate with a factor of 0.74.
@@
 This Bayesian probabilistic method is based on an assumption that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Paterson2011a}.
-The posterior paleointensity distributions from these sites with high-quality specimen-level data are in agreement with the site-level weighted-mean results developed using the selection criteria approach (Fig.
+The posterior paleointensity distributions from these sites with high-quality specimen-level data are in agreement with the site-level averages developed using the selection criteria approach (Fig.
 \ref{fig:PINT_cooling_corrected}; Fig.
@@
 We applied full TRMs to a set of anorthosite specimens with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axes.
-The results show that our anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions ( \citealp{Selkin2007a}; Fig.
+The results show that this set of anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions ( \citealp{Selkin2007a}; Fig.
 S4).
@@
 We therefore correct our paleointensity results by a factor of 0.74.
-The cooling rate-corrected specimen paleointensity estimates together with specimen- and site-level weighted means are shown in Fig.
+The cooling rate-corrected specimen paleointensity estimates together with specimen- and site-level means are shown in Fig.
 \ref{fig:PINT_cooling_corrected}.
@@
 Paleointensity estimates from \cite{Pesonen1983a} and \cite{Kulakov2013a} are not included in the compilation due to the specimen-level double-slope behavior as discussed in the text.
-Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \cite{Bono2019a} which is based on an interpretation of there being a monotonic decay of the geodynamo through the Proterozoic.}}
+Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \cite{Bono2019a} which is based on an interpretation of there being a monotonic decay of the geodynamo through the Proterozoic.
+The highest site-mean virtual dipole moment of the anorthosites are similar to the highest moment amongst the selected Phanerozoic records with Q$_{PI}$ values $>$3.
+}}
 \label{fig:PINT_compilation}
@@
 However, other experimental studies continue to indicate lower thermal conductivity values consistent with prior estimates \cite{Konopkova2016a, Hsieh2020a} with no consensus yet emerging \cite{Williams2018a, Ohta2021a}.
-These experiments are challenging with issues such as constraining the sample thickness under high pressure and temperature conditions, the validity of applying the Wiedemann-Franz law to extrapolate thermal conductivity values based on electrical resistivity measurements \cite{Ohta2016a}, and modeling uncertainties associated with parameters used in direct thermal conduction measurement experiments \cite{Konopkova2016a}.
+These experiments are challenging to conduct and interpret given complexities such as constraining the sample thickness under high pressure and temperature conditions, the validity of applying the Wiedemann-Franz law to extrapolate thermal conductivity values based on electrical resistivity measurements \cite{Ohta2016a}, and modeling uncertainties associated with parameters used in direct thermal conduction measurement experiments \cite{Konopkova2016a}.
 Further experiments and theory are needed to explain these contrasting results which at present leave open very different trajectories for Earth's thermal evolution.
@@
 Other scenarios are possible, however, such as the prediction that while power increases associated with inner core nucleation would result in increases in Earth's internal magnetic field strength, the dynamo would also become more deeply seated in the core, diminishing the increase in observed magnetic field at Earth's surface \cite{Aubert2009a, Landeau2017a}.
-Observational paleomagnetic records hold the potential for testing different model predictions and identifying transitions in ancient field strength.
+Observational paleomagnetic records hold the potential for testing different model predictions and identifying transitions in ancient field strength \cite{Biggin2015a, Bono2019a}.
 
@@
 
-The high paleointensity estimates from the Midcontinent Rift rocks challenge the hypothesized monotonic decay of the strength of the geomagnetic field throughout the Proterozoic Era.
+The high paleointensity estimates from the 1.1 billion-year-old Midcontinent Rift rocks challenge the hypothesized monotonic decay of the strength of the geomagnetic field throughout the Proterozoic Era.
 Instead, the well-preserved ca.
@@
 1140 Ma Abitibi dikes which yielded straight Arai plots \cite{Macouin2003a}, to a high moment of $\sim$70 ZAm$^2$ recorded by the ca.
-1106 Ma Osler Volcanics, with even stronger values recorded ca.
+1106 Ma Osler Volcanics, with even stronger values from ca.
 1092 Ma by Beaver River anorthosite xenoliths that record virtual dipole moments up to $\sim$140 ZAm$^2$.
@@
 \ref{fig:PINT_compilation};  \citealp{Lloyd2021b}).
-Following the Ediacaran, there are low paleointensity estimates from the ca.
+Following the Ediacaran, there are also low paleointensity estimates from the ca.
 414 Ma Strathmore lava flows \cite{Hawkins2021a}, the ca.
-410-380 Ma lava flows of the Minusa Basinand the Kola Peninsula \cite{Shcherbakova2017a}, the ca.
+410-380 Ma lava flows of the Minusa Basin and the Kola Peninsula \cite{Shcherbakova2017a}, the ca.
 408–393 Ma Buribay volcanic complex \cite{Shcherbakova2021a}, as well as the ca.
@@
 \ref{fig:PINT_compilation}).
-This interval of time with low paleointensity estimates has been termed the ``Mid-Paleozoic Dipole Low'' \cite{Hawkins2021a} and is followed by high paleointensity values afterwards (Fig.
+This interval of time with low paleointensity estimates has been termed the ``Mid-Paleozoic Dipole Low'' \cite{Hawkins2021a}.
+This  ``Mid-Paleozoic Dipole Low'' is followed by high paleointensity values such that there is at least a four-fold increase from a virtual dipole moment of $\sim$23 ZAm$^2$ at ca.
+338 to $\sim$93 ZAm$^2$ at ca.
+302 Ma (Fig.
 \ref{fig:PINT_compilation}).
 
-% that represent a transition to field strengths greater than the present
+Numerical models have shown that the stability of the geomagnetic axial dipole moment is sensitive to the total heat flux and the pattern of the heat flow distribution through the core-mantle boundary when there are strong available power sources to the geodynamo \cite{Glatzmaier1999a, Olson2007a, Olson2010a}.
+Modeling results form \cite{Olson2010a} show that a low total heat flux through the core-mantle boundary can prevent the axial dipole from reversing whereas a high heat flux through the boundary can result in increase in reversal frequency and decrease in the axial dipole intensity.
+The ``Mid-Paleozoic Dipole Low'' has been hypothesized to be the result of a high core-mantle boundary heat flux condition in a framework that includes available power from a crystallizing inner core \cite{Hawkins2019a}, thereby explaining the observed low paleointensities which are nearly as weak as those of the Sept-$\hat{I}$les intrusions \cite{Bono2019a} (Fig.
+\ref{fig:PINT_compilation}).
+There can also be large changes in the strength of the geomagnetic axial dipole due to changes in the geometry of core-mantle boundary heat flux.
+Such spatial heat flow variations could be associated with changes in the structure of the deep mantle due to plumes \cite{Larson1991a, Courtillot2007a}, subduction slabs \cite{Tan2002b, Hounslow2018a}, and distribution of dense mantle basal piles \cite{Biggin2012a}.
 
-% changes in core-mantle boundary heat flow related to plate tectonics - more general refrence to core-mantle heat flow pattern not necessarily telling about inner core nucleation.
+Could core-mantle interactions similar to that today exist before the nucleation of the inner core and modulated the geodynamo?
+With emerging paleomagnetic evidence supporting the existence of modern-day-like plate tectonics that extends through the Proterozoic \cite{Swanson-Hysell2021c} and to the Archean \cite{Mitchell2014a}, kinetic and thermal energy exchange between deep and shallow Earth likely existed since the early time of Earth's history.
+If the mantle forcing associated with these activities exists regardless of the present of inner core, it could have contributed to the anomalously low or high paleointensity values observed during the Phanerozoic and Proterozoic (Fig.
+\ref{fig:PINT_compilation}).
+The unknown magnitude of such forcing may challenge our ability to detect the increase in surface geomagnetic field strength predicted to have happened at the onset of the crystallization of the inner core \cite{Davies2021a} based on paleointensity data alone.
 
-An alternative hypothesis for the multiple transitions from low to high field intensities in the middle Paleozoic and the Proterozoic is that they are manifestations of secularly changing mantle forcing of the geodynamo across the core–mantle boundary which may be unrelated to the growth of the inner core.
+Regardless, the high-fidelity paleointensity recorders of the Beaver River anorthosite xenoliths in the well-preserved Midcontinent Rift record high paleointensities values and necessitates that there was a strongly powered late Mesoproterozoic geodynamo.
+
+% heterogeneity through the lower mantle can weaken and destabilize the geomagnetic axial dipole, reduce its moment, and increase reversal frequency.
+with non-reversing behavior at low heat flow and frequent reversals at high heat flow conditions.
+% We need to say models indicate that there can be weak fields when there are strong available power sources to the geodynamo.
+Cite Olson 2007 and Olson et al.
+2010.
+The ``Mid-Paleozoic Dipole Low'' has been interpreted to be the result of high total core-mantle boundary heat flux thereby explaining field nearly as weak as those of the Sept-Iles data in a framework that includes available power from a crystallizing inner core (can cite Hawkins et al.
+2019 as well as other Biggin papers that make this point).
+There can also be large changes in the strength of the dipole due to changes in the geometry of core-mantle boundary heat flux (these can be connected to subducted slabs and upwelling).
+
+% An alternative hypothesis for the multiple transitions from low to high field intensities in the middle Paleozoic and the Proterozoic is that they are manifestations of secularly changing mantle forcing of the geodynamo across the core–mantle boundary which may be unrelated to the growth of the inner core.
 The time scale over which these changes happen may be on the order of less than one million years.
@@
 \ref{fig:PINT_compilation}).
-If the mantle forcing that drove such transitions exists regardless of the present of inner core, it could have contributed to the anomalously low or high paleointensity values observed during the Phanerozoic and Proterozoic.
-The low paleointensity records from the ca.
+%
+
+%The low paleointensity records from the ca.
 720 Ma dolerite dikes of the rapidly emplaced Franklin Large Igneous Province could be one example of capturing these short-term variations and not fully averaging the secular changes of the geomagnetic field (Fig.
 \ref{fig:PINT_compilation}; \citealp{Lloyd2021a, Lloyd2021b}).
-The unknown magnitude of the forcing presents challenges for detecting the increase in dynamo strength caused by the kick-start of compositional convection driven by the crystallization of the inner core based on paleointensity data alone.
-
-% This interpretation may explain the low virtual dipole moment of 11 ZAm$^2$ recorded by the ca.
-
-Overall, the high-fidelity paleointensity recorders of the Beaver River anorthosite xenoliths in the well-preserved Midcontinent Rift record high paleointensities values and necessitates that there was a vigorous late Mesoproterozoic geodynamo.
 
@@
 If the weak paleointensities in the Ediacaran or the Devonian are attributed to the geodynamo being on the brink of collapse just prior to the nucleation of the solid inner core.
-
-% rocks from
 

### 2022-01-27 PNAS_manuscript/PNAS_manuscript.tex

@@
 \centering
-\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment values from the PINT database (PINT v8.0.0; \url{http://www.pintdb.org/};  \citealp{Bono2021a}) with additional Neoproterozoic data from refs.
+\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment values from the PINT database (PINT v8.0.0; \url{http://www.pintdb.org/};  \citealp{Bono2021a}), including all Phanerozoic records with Q$_{PI}$ values $>$3 and additional Neoproterozoic data from refs.
 \citealp{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}.
@@
 Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \cite{Bono2019a} which is based on an interpretation of there being a monotonic decay of the geodynamo through the Proterozoic.
-The highest site-mean virtual dipole moment of the anorthosites are similar to the highest moment amongst the selected Phanerozoic records with Q$_{PI}$ values $>$3.
+The highest site-mean virtual dipole moment of the anorthosites are similar to the highest moment amongst the selected Phanerozoic data.
 }}

### 2022-01-28 PNAS_manuscript/PNAS_manuscript.tex
Updates from Overleaf

@@
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{Acquiring high-fidelity observational records of ancient magnetic field intensity from the remanent magnetization of rocks is crucial for constraining the evolution of Earth's core.
+\significancestatement{Acquiring high-fidelity observational records of ancient magnetic field intensity from the remanent magnetization of rocks is crucial for constraining the long-term evolution of Earth's core.
 However, robust estimates of ancient field strengths are often difficult to recover due to alteration or non-ideal rock magnetic behavior.
 In this study, we use plagioclase cumulates (anorthosite) that formed in the deep crust and were brought to the near surface by magma where they cooled and acquired thermal remanent magnetizations.
-These anorthosite xenoliths from the North American Midcontinent Rift have experienced minimal alteration since their emplacement and yield high quality paleointensity estimates.
-In contrast to predictions of a progressively decaying field, these data record a strong virtual dipole moment 1.1 billion years ago.
+These anorthosite xenoliths have experienced minimal alteration since their emplacement and yield high quality paleointensity estimates.
+In contrast to predictions of a progressively decaying field leading up to late nucleation of Earth's inner core, these data record a strong virtual dipole moment 1.1 billion years ago.
 A strong field persisted over at least a 20-million-year interval indicating the existence of appreciable power sources for Earth's dynamo during the late Mesoproterozoic.}
@@
 1.1 Ga.
-These data are inconsistent with there being a progressive monotonic decay of Earth's dynamo strength through the Proterozoic Eon.
+These data are inconsistent with there being a progressive monotonic decay of Earth's dynamo strength through the Proterozoic Eon and could challenge the hypothesis of a young inner core.
 \end{abstract}
@@
 565 Ma during the Ediacaran Period \cite{Bono2019a, Shcherbakova2019a,Thallner2021b} and ca.
-400 Ma during the Devonian Period \cite{Shcherbakova2017a, Shcherbakova2021a, Hawkins2021a}, potentially indicating unusual periods of core dynamo regime at those times.
+370 Ma during the Devonian Period \cite{Shcherbakova2017a, Shcherbakova2021a, Hawkins2021a}, potentially indicating unusual periods of core dynamo regime at those times.
 The Ediacaran data have been interpreted to indicate that there was a progressively decaying field up to that time that was followed soon after by initial crystallization of the inner core \cite{Bono2019a}.
@@
 1092 Ma by Beaver River anorthosite xenoliths that record virtual dipole moments up to $\sim$140 ZAm$^2$.
-While the Abitibi dikes paleointensity estimates do not go to values as low as the Sept-$\hat{I}$les intrusions, this virtual dipole moment increase in the Mesoproterozoic is the largest yet documented in the Precambrian (Fig.
+While the Abitibi dikes paleointensity estimates do not go to values as low as the Sept-$\hat{I}$les intrusions, this virtual dipole moment increase in the Mesoproterozoic is the largest yet documented in the Precambrian on a 10s of millions of year timescale (Fig.
 \ref{fig:PINT_compilation}).
@@
 This timing would be broadly consistent with that proposed by \citealp{Biggin2015a}.
-However, the model prediction would be of sustained strong field values following inner core nucleation which is challenged by data from the ca.
+However, a model prediction could be of sustained strong field values following inner core nucleation which is challenged by data from the ca.
 1070 Ma Bangemall Sills which include a sill with a low virtual dipole moment of $\sim$6.4 ZAm$^2$ (Fig.
 \ref{fig:PINT_compilation};  \citealp{Lloyd2021b}).
-Following the Ediacaran, there are also low paleointensity estimates from the ca.
+Following the Ediacaran, there are also low paleointensity estimates from Devonian rocks such as ca.
+370 Ma dikes and lavas of the Siberian Viluy Traps that give virtual dipole moment of 4.3 to 14.9 ZAm$^2$ (Fig.
+\ref{fig:PINT_compilation}; \citealp{Hawkins2019a}).
+These low values and as well as data from the ca.
 414 Ma Strathmore lava flows \cite{Hawkins2021a}, the ca.
 410-380 Ma lava flows of the Minusa Basin and the Kola Peninsula \cite{Shcherbakova2017a}, the ca.
-408–393 Ma Buribay volcanic complex \cite{Shcherbakova2021a}, as well as the ca.
-370 Ma dikes and lavas of the Siberian Viluy Traps \cite{Hawkins2019a} (Fig.
+408–393 Ma Buribay volcanics  volcanic complex \cite{Shcherbakova2021a} and the ca.
+332 Ma Kinghorn volcanics \cite{Hawkins2021a} has led to the proposal of this interval as the ``Mid-Paleozoic Dipole Low'' \cite{Hawkins2021a}.
+This  ``Mid-Paleozoic Dipole Low'' is followed by high paleointensity values such that there is a six-fold increase in the database from virtual dipole moment of $\sim$16 ZAm$^2$ at ca.
+332 Ma to $\sim$99 ZAm$^2$ ca.
+308 Ma in the late Carboniferous \cite{Hawkins2021a}.
+
+Given that not each record of a weak field can be a minimum prior to inner core nucleation, what processes could lead to a weak dipole at Earth's surface even in the presence of a crystallizing inner core?
+Numerical models have shown that the dipole moment is sensitive to both the magnitude and spatial pattern of heat flow across the core-mantle boundary when there are strong available power sources to the geodynamo \cite{Olson2007a, Olson2010a}.
+In such models, relatively low total heat flux across the core-mantle boundary can prevent the axial dipole from reversing whereas a high heat flux through the boundary can result in an increase in reversal frequency and decrease in the axial dipole intensity.
+The ``Mid-Paleozoic Dipole Low'' has been hypothesized to be the result of elevated core-mantle boundary heat flux condition at a time when there was also available power from a crystallizing inner core \cite{Hawkins2019a}, thereby explaining the observed low paleointensities which include values as weak as those of the ca.
+565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a} (Fig.
 \ref{fig:PINT_compilation}).
-This interval of time with low paleointensity estimates has been termed the ``Mid-Paleozoic Dipole Low'' \cite{Hawkins2021a}.
-This  ``Mid-Paleozoic Dipole Low'' is followed by high paleointensity values such that there is at least a four-fold increase from a virtual dipole moment of $\sim$23 ZAm$^2$ at ca.
-338 to $\sim$93 ZAm$^2$ at ca.
-302 Ma (Fig.
-\ref{fig:PINT_compilation}).
+Mantle convection can modulate core mantle boundary heat flow through changes in the structure of the deep mantle associated with upwelling plumes \cite{Larson1991a, Courtillot2007a} and subducted slabs \cite{Tan2002b, Biggin2012a, Hounslow2018a}.
+Strong evidence for differential plate tectonic motion extends back to ca.
+2.2 Ga in the Paleoproterozoic \cite{Mitchell2014a,Swanson-Hysell2021b} and potentially back to ca.
+3.2 Ga in the Archean \cite{Brenner2020a}.
+Plate tectonic modulations of core mantle boundary heat flow are therefore expected throughout the Proterozoic.
+Such changes may explain large variability in values as in the Phanerozoic \cite{Lloyd2021a} and may challenge our ability to detect the increase in surface geomagnetic field strength predicted to have happened at the onset of inner core crystallization.
 
-Numerical models have shown that the stability of the geomagnetic axial dipole moment is sensitive to the total heat flux and the pattern of the heat flow distribution through the core-mantle boundary when there are strong available power sources to the geodynamo \cite{Glatzmaier1999a, Olson2007a, Olson2010a}.
-Modeling results form \cite{Olson2010a} show that a low total heat flux through the core-mantle boundary can prevent the axial dipole from reversing whereas a high heat flux through the boundary can result in increase in reversal frequency and decrease in the axial dipole intensity.
-The ``Mid-Paleozoic Dipole Low'' has been hypothesized to be the result of a high core-mantle boundary heat flux condition in a framework that includes available power from a crystallizing inner core \cite{Hawkins2019a}, thereby explaining the observed low paleointensities which are nearly as weak as those of the Sept-$\hat{I}$les intrusions \cite{Bono2019a} (Fig.
-\ref{fig:PINT_compilation}).
-There can also be large changes in the strength of the geomagnetic axial dipole due to changes in the geometry of core-mantle boundary heat flux.
-Such spatial heat flow variations could be associated with changes in the structure of the deep mantle due to plumes \cite{Larson1991a, Courtillot2007a}, subduction slabs \cite{Tan2002b, Hounslow2018a}, and distribution of dense mantle basal piles \cite{Biggin2012a}.
+Overall, the high-fidelity paleointensity recorders of the Beaver River anorthosite xenoliths in the well-preserved Midcontinent Rift record strong field strengths 1.1 billion years ago.
+The highest site level field strength would rank in the xx percentile of the paleointensity database for the Cenozoic Era when there was unequivocally a crystallizing inner core.
+These high surface field strengths necessitate appreciable power to the late Mesoproterozoic geodynamo.
 
-Could core-mantle interactions similar to that today exist before the nucleation of the inner core and modulated the geodynamo?
+% Could core-mantle interactions similar to that today exist before the nucleation of the inner core and modulated the geodynamo?
 With emerging paleomagnetic evidence supporting the existence of modern-day-like plate tectonics that extends through the Proterozoic \cite{Swanson-Hysell2021c} and to the Archean \cite{Mitchell2014a}, kinetic and thermal energy exchange between deep and shallow Earth likely existed since the early time of Earth's history.
@@
 \ref{fig:PINT_compilation}).
-The unknown magnitude of such forcing may challenge our ability to detect the increase in surface geomagnetic field strength predicted to have happened at the onset of the crystallization of the inner core \cite{Davies2021a} based on paleointensity data alone.
-
-Regardless, the high-fidelity paleointensity recorders of the Beaver River anorthosite xenoliths in the well-preserved Midcontinent Rift record high paleointensities values and necessitates that there was a strongly powered late Mesoproterozoic geodynamo.
 

### 2022-01-28 PNAS_manuscript/Supporting Info.tex
Updates from Overleaf

@@
 
-% \begin{figure*}[h!]
-% \noindent\includegraphics[width=17.8 cm]{manuscript/nonlinear.pdf}
-% \centering
-% \caption{{}}
-% \label{fig:nonlinear_check}
-% \end{figure*}
+\begin{figure*}[h!]
+\noindent\includegraphics[width=0.9\textwidth]{Linear_TRM_test.pdf}
+\centering
+\caption{{Plot of linear thermal remanent magnetization acquisition experiments.
+After IZZI-type Thellier paleointensity experiments, full TRMs were imparted on the same specimens in known lab fields of 30, 50, 70, and 90 $\mu$T.
+The results show that the anorthosite xenoliths do not acquire saturation remanence or display non-linear remanence acquisition under fields relevant to this study.
+Therefore, non-linear acquisition correction is not needed for our paleointensity results.
+}}
+\label{fig:nonlinear_check}
+\end{figure*}
 

### 2022-01-28 PNAS_manuscript/PNAS_manuscript.tex

@@
 In contrast to the diabase which fails to provide reliable paleointensity estimates, the anorthosite xenoliths are unusually high-fidelity recorders yielding high-quality, single-slope paleointensity results that are consistent at specimen and site levels.
-An average value of $\sim$89 ZAm$^2$ for the virtual dipole moment from the anorthosites with the highest site-level values up to $\sim$140 ZAm$^2$ are higher than that of the dipole component of Earth's magnetic field today.
+An average value of $\sim$89 ZAm$^2$ for the virtual dipole moment from the anorthosites with the highest site-level values up to $\sim$140 ZAm$^2$ are higher than that of the dipole component of Earth's magnetic field today and rival the highest values in the paleointensity database.
 Such high intensities recorded by the anorthosite xenoliths requires the existence of a strongly powered geodynamo at the time.
@@
 \noindent\includegraphics[width=11.4 cm]{Geologic_map.pdf}
-\caption{\footnotesize{Simplified geologic map of the Lake Superior region showing the distribution of rift and post-rift-related rocks including extrusive, intrusive, and late/post-rift sediments around Lake Superior.
+\caption{\footnotesize{Simplified geologic map of the Lake Superior region showing the distribution of rocks associated with the late Mesoproterozoic Midcontinent Rift.
 Purple stars mark sites with paleointensity results that passed the selection criteria from this study.
@@
 However, intraplate magmatism associated with large igneous provinces is typically of geologically short duration with the bulk of magmatic products emplaced within 1 Myr or less \cite{Kasbohm2021a}.
-The Midcontinent Rift is an exception as it is a large igneous province where magmatism lasted $\sim$25 Myr from \texit{ca.} 1109 Ma to 1084 Ma \cite{Swanson-Hysell2019a}.
+The Midcontinent Rift (Fig.
+\ref{fig:Geologic_map}) is an exception as it is a large igneous province where magmatism lasted $\sim$25 Myr from \texit{ca.} 1109 Ma to 1084 Ma \cite{Swanson-Hysell2019a}.
 Additionally, extension ceased in the Midcontinent Rift prior to lithospheric separation, preserving volcanic, intrusive, and sedimentary rocks of the rift within the continental interior.
@@
 Typically, such non-ideal behavior would result in a higher paleointensity estimate from the steeper-sloped low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
-For example, in data from the Midcontinent Rift,  \citealp{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas  \citealp{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
+For example, in data from the Midcontinent Rift, ref.
+\citealp{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas ref.
+\citealp{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
 Results with such non-ideal behavior were rejected by \citealp{Sprain2018a} who applied stricter paleointensity selection criteria, but as a result had few accepted sites.
@@
 Anorthosites are cumulate rocks composed dominantly of the mineral plagioclase.
-They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud minute magnetic inclusions from alteration with minor alteration of the plagioclase crystals not resulting in the formation of secondary iron oxides in contrast with mafic minerals such as olivine.
+They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud magnetic inclusions from alteration with alteration of the plagioclase crystals themselves not resulting in the formation of secondary iron oxides in contrast with mafic minerals such as olivine.
 The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal intrusions of the Beaver River diabase \cite{Zhang2021b}.
-They cooled and acquired their magnetizations in conjunction with the host diabase at a paleolatitude of 22\textdegree $\pm$2 \textdegree (calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a,Zhang2021b}).
+They cooled and acquired their magnetizations in conjunction with the host diabase at a paleolatitude of 22\textdegree $\pm$2 \textdegree (calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a, Zhang2021b}).
 Paleointensity experiments on the anorthosite xenoliths have a high success rate, yielding consistent specimen- and site-level paleointensity results.
@@
 
-\section*{Anorthosite xenoliths of the Midcontinent diabase}
+\section*{Background on anorthosite xenoliths of the Beaver River diabase}
 
@@
 In contrast, plagioclase crystals in the Duluth Complex anorthosite have euhedral, interlocking crystals with an igneous foliation (B) and the plagioclase crystals contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the [001] axis of the plagioclase.
-Magnetic field maps acquired with a quantum diamond microscope (QDM) show that the relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite in (E) while large oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the overlapping magnetic fields above the oxides in (F).
+Magnetic field maps acquired with a quantum diamond microscope (QDM) on samples to which an IRM has been applied show that the relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite in (E) while large oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the overlapping magnetic fields above the oxides in (F).
 Magnetic field maps of (E) and (F) share the same spatial and magnetic scale.
@@
 
-The lack of symplecites and the lack of large plagioclase-hosted titanomagnetite needles within the Beaver River anorthosite xenoliths distinguish them from plagioclase cumulates of the Duluth Complex and other layered mafic complex where anisotropic magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
+The lack of symplecites and the lack of large plagioclase-hosted titanomagnetite needles within the Beaver River anorthosite xenoliths targetted in this study distinguish them from plagioclase cumulates of the Duluth Complex and other layered mafic complexes where anisotropic magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
 The relative lack of fabric and associated preferred orientation makes the Beaver River anorthosite xenoliths a particularly compelling target for paleointensity experiments.
@@
 7 anorthosite sites and no diabase site-level results pass these selection criteria.
-Example NRM/TRM (Arai) plots are shown in Fig.
-\ref{fig:IZZI_examples}.
-Summary specimen absolute paleointensity estimates and site-level mean paleointensity values are plotted in Fig.
-\ref{fig:PINT_cooling_corrected} (and provided in Table S1).
+Example NRM/TRM (Arai) plots are shown in Figure \ref{fig:IZZI_examples}.
+Summary specimen absolute paleointensity estimates and site-level mean paleointensity values are plotted in Figure \ref{fig:PINT_cooling_corrected} (and provided in Table S1).
 The paleointensity quality index (Q$_{PI}$;  \citealp{Biggin2014a}) for the anorthosite xenoliths are all 5 or 6 (Table S2).
@@
 S1).
-
-Overall, the high-quality results from the anorthosite xenoliths of the Beaver River diabase and the consistent paleointensity estimates based on independent methods confirm that the anorthosite xenoliths record a high geomagnetic field ca.
+Overall, the high-quality results from the anorthosite xenoliths of the Beaver River diabase indiacte that the anorthosite xenoliths record a high geomagnetic field ca.
 1092 Ma.
@@
 To assess whether our paleointensity estimates are biased by remanence anisotropy, we calculated the gamma statistic, which is the angular difference between the last pTRM step of paleointensity experiment and the applied field direction.
-The results show that the anorthosite specimens used for paleointensity experiment have low gamma values ranging from $\sim$0.9\textdegree to $\sim$11.9\textdegree, with a median value of 4.2\textdegree$\;$(Table S1).
+The results show that the anorthosite specimens used for paleointensity experiment have low gamma values ranging from 0.9\textdegree$\;$ to 11.9\textdegree, with a median value of 4.2\textdegree$\;$(Table S1).
 These gamma values are similar to those of Midcontinent Rift volcanics \cite{Sprain2018a}.
-Therefore, the Beaver River anorthosite xenolith bulk samples do not have a significant remanence anisotropy.
+Therefore, the Beaver River anorthosite xenolith bulk samples do not have significant remanence anisotropy.
 Paleodirectional data from our anorthosite xenoliths further support that they have minimal remanence anisotropy as their site mean directions closely match those of the Beaver River diabase hosts without deviating due to a fabric \cite{Zhang2021b}.
@@
 
-\subsection*{Averaging secular variations and applying cooling rate correction}
+\subsection*{Considering secular variation and cooling rate}
 To best characterize the geomagnetic axial dipole field intensity during a certain time period a paleointensity dataset should cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged.
-Thermal modeling results from \citealp{Zhang2021b} suggest that the Beaver River anorthosite xenoliths were heated to tholeiitic magma temperature ($\sim$1100\textdegree C) and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a thousand years, partially averaging secular variation within single samples.
+Thermal modeling results from ref.
+\citealp{Zhang2021b} suggest that the Beaver River anorthosite xenoliths were heated to tholeiitic magma temperature ($\sim$1100\textdegree C) and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a thousand years, partially averaging secular variation within single sites.
 Anorthosite xenoliths have paleointensity results that are consistent within small regions, but vary between localities.
-Anorthosite AX12 and AX16 were emplaced $\sim$450 meters apart and have very similar paleointensity estimates (Fig.
-\ref{fig:Geologic_map, fig:PINT_cooling_corrected}).
-$\sim$10 km to the north, anorthosite xenoliths AX11 and AX13 were also emplaced closely at a distance of $\sim$125 meters and yield similar but distinct paleointensity estimates from those of AX12 and AX16 (Fig.
-\ref{fig:Geologic_map, fig:PINT_cooling_corrected}).
+Anorthosite AX12 and AX16 were emplaced $\sim$450 meters apart and have very similar paleointensity estimates (Figs.
+\ref{fig:Geologic_map} and \ref{fig:PINT_cooling_corrected}).
+$\sim$10 km to the north, anorthosite xenoliths AX11 and AX13 were also emplaced closely (seperated by $\sim$125 meters) and yield similar values to one another but distinct paleointensity estimates from those of AX12 and AX16 (Figs.
+\ref{fig:Geologic_map} and \ref{fig:PINT_cooling_corrected}).
 The data therefore capture different intervals of time during the emplacement and cooling of the Beaver River diabase sills ca.
@@
 Another consequence of the anorthosite xenoliths having cooled in the interior of thick diabase intrusions is that slow cooling rates can bias paleointensity estimates toward higher values.
-Large differences in cooling rates between acquisition of an NRM in nature versus a TRM in the lab can result in overestimated paleointensities for SD grains \cite{Dodson1980a, Halgedahl1980a, Nagy2021a}.
-From the thermal history model of  \citealp{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully blocked in their characteristic natural remanence magnetization ($\sim$500\textdegree C;  \citealp{Zhang2021b}).
+Large differences in cooling rates between acquisition of an NRM in nature versus a TRM in the lab can result in overestimated paleointensities for single domain grains \cite{Dodson1980a, Halgedahl1980a, Nagy2021a}.
+From the thermal history model of ref.
+\citealp{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully blocked in their characteristic natural remanence magnetization ($\sim$500\textdegree C;  \citealp{Zhang2021b}).
 We find the cooling time to be on the order of $\sim$1 kyr, which corresponds to a cooling rate of $\sim1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
@@
 S3).
-This estimate on cooling rate effect is similar to the value of $\sim$35\% overestimate derived from the model of \citealp{Nagy2021a}.
+This estimate on cooling rate effect is similar to the value of $\sim$35\% overestimate derived from the model of ref.
+\citealp{Nagy2021a}.
 We therefore correct our paleointensity results by a factor of 0.74.
@@
 \centering
-\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment values from the PINT database (PINT v8.0.0; \url{http://www.pintdb.org/};  \citealp{Bono2021a}), including all Phanerozoic records with Q$_{PI}$ values $>$3 and additional Neoproterozoic data from refs.
+\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment values from the PINT database (PINT v8.0.0; \url{http://www.pintdb.org/};  \citealp{Bono2021a}), including all Phanerozoic VDM and VADM records with Q$_{PI}$ values $>$3 and additional Neoproterozoic data from refs.
 \citealp{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}.
@@
 Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \cite{Bono2019a} which is based on an interpretation of there being a monotonic decay of the geodynamo through the Proterozoic.
-The highest site-mean virtual dipole moment of the anorthosites are similar to the highest moment amongst the selected Phanerozoic data.
+The highest site-mean virtual dipole moment of the anorthosites are similar to the 99.5 percentile of the selected Phanerozoic data.
 }}
@@
 Inner core nucleation would have increased the power to the geodynamo which has the potential to manifest as an increase in Earth's surface field \cite{Davies2021a}.
-An approach combining dynamo simulations and power-based theoretical scaling relationships has predicted that progressive decay of the field's dipole moment would be followed by a rapid increase in geomagnetic field intensity soon after the onset of inner core nucleation such that there is a minimum in dipole moment just before inner core nucleation \cite{Davies2021a}.
+An approach combining dynamo simulations and theoretical scaling relationships has predicted that progressive decay of the field's dipole moment would be followed by a rapid increase in geomagnetic field intensity soon after the onset of inner core nucleation such that there is a minimum in dipole moment just before inner core nucleation \cite{Davies2021a}.
 Other scenarios are possible, however, such as the prediction that while power increases associated with inner core nucleation would result in increases in Earth's internal magnetic field strength, the dynamo would also become more deeply seated in the core, diminishing the increase in observed magnetic field at Earth's surface \cite{Aubert2009a, Landeau2017a}.
@@
 565 Ma in the Ediacaran Period (Fig.
-\ref{fig:PINT_compilation}, \citealp{Bono2019a}).
+\ref{fig:PINT_compilation}; \citealp{Bono2019a}).
 This interpretation was motivated by paleointensity estimates developed from the ca.
 565 Ma Sept-$\hat{I}$les layered mafic intrusive complex of $\sim$7 ZAm$^2$ that are among the lowest values in the paleointensity database (Fig.
-\ref{fig:PINT_compilation}), \citealp{Bono2019a}.
+\ref{fig:PINT_compilation}; \citealp{Bono2019a}).
 A decay in the lead-up to this time was argued to be consistent with an absence of an inner core and a dynamo to which progressively less power was available through secular cooling \cite{Bono2019a, Davies2021a}.
 This timing of inner core formation would favor a high core thermal conductivity \cite[e.g.][]{Ohta2016a}.
-Paleomagnetic directional excursions \cite{Halls2015a} and frequent polarity reversals \cite{Kodama2021a} during this time are interpreted to be consistent with numerical simulations \cite{Driscoll2016a} associated with a weak dipole field.
+Paleomagnetic directional excursions \cite{Halls2015a}, other weak paleointensity estimates \cite{Thallner2021b}, and frequent polarity reversals \cite{Kodama2021a} in rocks of similar age are interpreted to be consistent with numerical simulations \cite{Driscoll2016a} associated with a weak dipole field.
 
@@
 \ref{fig:PINT_compilation}).
-In the late Mesoproterozoic, there is at least a six-fold increase within a period of $\sim$40 Myr from a low average virtual dipole moment of $\sim$13 ZAm$^2$ recorded by the ca.
+In the late Mesoproterozoic, there is at least a six-fold increase within a period of $\sim$35 Myr from a low average virtual dipole moment of $\sim$13 ZAm$^2$ recorded by the ca.
 1140 Ma Abitibi dikes which yielded straight Arai plots \cite{Macouin2003a}, to a high moment of $\sim$70 ZAm$^2$ recorded by the ca.
@@
 
-Given that not each record of a weak field can be a minimum prior to inner core nucleation, what processes could lead to a weak dipole at Earth's surface even in the presence of a crystallizing inner core?
+Given that each record of a weak field can not be a minimum prior to inner core nucleation, what processes could lead to a weak dipole at Earth's surface even in the presence of a crystallizing inner core?
 Numerical models have shown that the dipole moment is sensitive to both the magnitude and spatial pattern of heat flow across the core-mantle boundary when there are strong available power sources to the geodynamo \cite{Olson2007a, Olson2010a}.
@@
 Plate tectonic modulations of core mantle boundary heat flow are therefore expected throughout the Proterozoic.
-Such changes may explain large variability in values as in the Phanerozoic \cite{Lloyd2021a} and may challenge our ability to detect the increase in surface geomagnetic field strength predicted to have happened at the onset of inner core crystallization.
+Such changes may explain large variability in Proterozoic paleointensity values similar to those seen in the Phanerozoic \cite{Lloyd2021a} and may challenge our ability to detect the increase in surface geomagnetic field strength predicted to have happened at the onset of inner core crystallization.
 
 Overall, the high-fidelity paleointensity recorders of the Beaver River anorthosite xenoliths in the well-preserved Midcontinent Rift record strong field strengths 1.1 billion years ago.
-The highest site level field strength would rank in the xx percentile of the paleointensity database for the Cenozoic Era when there was unequivocally a crystallizing inner core.
+The highest site level field strength would rank in the 99 percentile of the paleointensity database for the Cenozoic Era (the last 66 million years) when there was unequivocally a crystallizing inner core.
 These high surface field strengths necessitate appreciable power to the late Mesoproterozoic geodynamo.
@@
 
-\acknow{Project research was supported by NSF CAREER grant EAR-1847277 to N.L.S.-H., an Institute on Lake Superior Geology Student Research Fund grant to Y.Z., and an IRM Visiting Student Fellowship to Y.Z.
+\acknow{Project research was supported by NSF CAREER grant EAR-1847277 to N.L.S.-H., an Institute on Lake Superior Geology Student Research Fund grant to Y.Z., and an Institute for Rock Magnetism Visiting Student Fellowship to Y.Z.
+The Institute for Rock Magnetism is supported by the Instrumentation and Facilities program of the NSF Earth Science Division.
 Permits for fieldwork and sampling from the Minnesota Department of Natural Resources are gratefully acknowledged.
@@
 We thank Jim Miller for providing us with the Duluth Complex anorthosite thin section.
-We thank Dario Bilardello, Peat Solheid, Mike Jackson, Josh Feinberg, and Bruce Moskowitz for their tremendous help of instrumental operations, data interpretations, and research guidance at the IRM.
+We thank Dario Bilardello, Peat Solheid, Mike Jackson, Josh Feinberg, and Bruce Moskowitz for their tremendous help associated with instrument operation, data interpretations, and research guidance at the IRM.
 Conversations with Bruce Buffett and Zachary Geballe informed our perspectives on the geodynamic and mineral physics literature.
-Paleomagnetic data associated with this study are available within the MagIC database (\url{}; \textit{provide link when doi is generated at time of proofs}) and all data are within a Github repository associated with this work (\url{https://earthorg/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}) that is also archived on Zenodo (\textit{provide link when doi is generated at time of proofs}).
-This repository also contains Python code related to calculations, visualizations and statistical tests discussed herein.}
+Paleomagnetic data associated with this study are available within the MagIC database (\url{https://earthorg/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}; \textit{provide link when doi is generated at time of proofs}) and all data are within a Github repository associated with this work along with Python code related to calculations and visualizations associated with the study that is also archived on Zenodo (\textit{provide link when doi is generated at time of proofs}).}
 

### 2022-01-28 PNAS_manuscript/PNAS_manuscript.tex

@@
 
-The lack of symplecites and the lack of large plagioclase-hosted titanomagnetite needles within the Beaver River anorthosite xenoliths targetted in this study distinguish them from plagioclase cumulates of the Duluth Complex and other layered mafic complexes where anisotropic magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
+The lack of symplecites and the lack of large plagioclase-hosted titanomagnetite needles within the Beaver River anorthosite xenoliths targeted in this study distinguish them from plagioclase cumulates of the Duluth Complex and other layered mafic complexes where anisotropic magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
 The relative lack of fabric and associated preferred orientation makes the Beaver River anorthosite xenoliths a particularly compelling target for paleointensity experiments.
@@
 S1).
-Overall, the high-quality results from the anorthosite xenoliths of the Beaver River diabase indiacte that the anorthosite xenoliths record a high geomagnetic field ca.
+Overall, the high-quality results from the anorthosite xenoliths of the Beaver River diabase indicate that the anorthosite xenoliths record a high geomagnetic field ca.
 1092 Ma.
@@
 1092 Ma by Beaver River anorthosite xenoliths that record virtual dipole moments up to $\sim$140 ZAm$^2$.
-While the Abitibi dikes paleointensity estimates do not go to values as low as the Sept-$\hat{I}$les intrusions, this virtual dipole moment increase in the Mesoproterozoic is the largest yet documented in the Precambrian on a 10s of millions of year timescale (Fig.
+While the Abitibi dikes paleointensity estimates do not go to values as low as the Sept-$\hat{I}$les intrusions, this virtual dipole moment increase in the Mesoproterozoic is the largest yet documented in the Precambrian on a 10s of millions of years timescale (Fig.
 \ref{fig:PINT_compilation}).
@@
 Following the Ediacaran, there are also low paleointensity estimates from Devonian rocks such as ca.
-370 Ma dikes and lavas of the Siberian Viluy Traps that give virtual dipole moment of 4.3 to 14.9 ZAm$^2$ (Fig.
+370 Ma dikes and lavas of the Siberian Viluy Traps that give virtual dipole moment estimates of 4.3 to 14.9 ZAm$^2$ (Fig.
 \ref{fig:PINT_compilation}; \citealp{Hawkins2019a}).
-These low values and as well as data from the ca.
+These low values as well as data from the ca.
 414 Ma Strathmore lava flows \cite{Hawkins2021a}, the ca.
-410-380 Ma lava flows of the Minusa Basin and the Kola Peninsula \cite{Shcherbakova2017a}, the ca.
-408–393 Ma Buribay volcanics  volcanic complex \cite{Shcherbakova2021a} and the ca.
+410-380 Ma lava flows of Siberia and the Kola Peninsula \cite{Shcherbakova2017a}, the ca.
+408–393 Ma Buribay volcanics \cite{Shcherbakova2021a} and the ca.
 332 Ma Kinghorn volcanics \cite{Hawkins2021a} has led to the proposal of this interval as the ``Mid-Paleozoic Dipole Low'' \cite{Hawkins2021a}.
@@
 332 Ma to $\sim$99 ZAm$^2$ ca.
-308 Ma in the late Carboniferous \cite{Hawkins2021a}.
+308 Ma in the late Carboniferous (Fig.
+\ref{fig:PINT_compilation}; \citealp{Hawkins2021a}).
 
-Given that each record of a weak field can not be a minimum prior to inner core nucleation, what processes could lead to a weak dipole at Earth's surface even in the presence of a crystallizing inner core?
+Given that the multiple records of a weak field can not all be a minimum prior to inner core nucleation, what processes could lead to a weak dipole at Earth's surface even in the presence of a crystallizing inner core?
 Numerical models have shown that the dipole moment is sensitive to both the magnitude and spatial pattern of heat flow across the core-mantle boundary when there are strong available power sources to the geodynamo \cite{Olson2007a, Olson2010a}.
-In such models, relatively low total heat flux across the core-mantle boundary can prevent the axial dipole from reversing whereas a high heat flux through the boundary can result in an increase in reversal frequency and decrease in the axial dipole intensity.
-The ``Mid-Paleozoic Dipole Low'' has been hypothesized to be the result of elevated core-mantle boundary heat flux condition at a time when there was also available power from a crystallizing inner core \cite{Hawkins2019a}, thereby explaining the observed low paleointensities which include values as weak as those of the ca.
+In such models, relatively low total heat flux across the core-mantle boundary can prevent the axial dipole from reversing whereas a high heat flux through the boundary can result in an increase in reversal frequency and decrease in dipole intensity.
+The ``Mid-Paleozoic Dipole Low'' has been hypothesized to be the result of such elevated core-mantle boundary heat flux conditions at a time when there was also available power from a crystallizing inner core \cite{Hawkins2019a}, thereby explaining the observed low paleointensities which include values as weak as those of the ca.
 565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a} (Fig.
@@
 Overall, the high-fidelity paleointensity recorders of the Beaver River anorthosite xenoliths in the well-preserved Midcontinent Rift record strong field strengths 1.1 billion years ago.
-The highest site level field strength would rank in the 99 percentile of the paleointensity database for the Cenozoic Era (the last 66 million years) when there was unequivocally a crystallizing inner core.
+The highest site-level value of the virtual dipole moment would rank in the top 1$\%$ of those in the database for the Cenozoic Era (the last 66 million years) when there was unequivocally a crystallizing inner core.
 These high surface field strengths necessitate appreciable power to the late Mesoproterozoic geodynamo.
-
-% Could core-mantle interactions similar to that today exist before the nucleation of the inner core and modulated the geodynamo?
-With emerging paleomagnetic evidence supporting the existence of modern-day-like plate tectonics that extends through the Proterozoic \cite{Swanson-Hysell2021c} and to the Archean \cite{Mitchell2014a}, kinetic and thermal energy exchange between deep and shallow Earth likely existed since the early time of Earth's history.
-If the mantle forcing associated with these activities exists regardless of the present of inner core, it could have contributed to the anomalously low or high paleointensity values observed during the Phanerozoic and Proterozoic (Fig.
-\ref{fig:PINT_compilation}).
-
-% heterogeneity through the lower mantle can weaken and destabilize the geomagnetic axial dipole, reduce its moment, and increase reversal frequency.
-with non-reversing behavior at low heat flow and frequent reversals at high heat flow conditions.
-% We need to say models indicate that there can be weak fields when there are strong available power sources to the geodynamo.
-Cite Olson 2007 and Olson et al.
-2010.
-The ``Mid-Paleozoic Dipole Low'' has been interpreted to be the result of high total core-mantle boundary heat flux thereby explaining field nearly as weak as those of the Sept-Iles data in a framework that includes available power from a crystallizing inner core (can cite Hawkins et al.
-2019 as well as other Biggin papers that make this point).
-There can also be large changes in the strength of the dipole due to changes in the geometry of core-mantle boundary heat flux (these can be connected to subducted slabs and upwelling).
-
-% An alternative hypothesis for the multiple transitions from low to high field intensities in the middle Paleozoic and the Proterozoic is that they are manifestations of secularly changing mantle forcing of the geodynamo across the core–mantle boundary which may be unrelated to the growth of the inner core.
-The time scale over which these changes happen may be on the order of less than one million years.
-During the Pleistocene, when we known a present-day-like geodynamo existed, paleointensity records show at least a six-fold fluctuation of the geomagnetic dipole moment can happen (Fig.
-\ref{fig:PINT_compilation}).
-%
-
-%The low paleointensity records from the ca.
-720 Ma dolerite dikes of the rapidly emplaced Franklin Large Igneous Province could be one example of capturing these short-term variations and not fully averaging the secular changes of the geomagnetic field (Fig.
-\ref{fig:PINT_compilation}; \citealp{Lloyd2021a, Lloyd2021b}).
-
-% In the Neoproterozoic, ca.
-720 Ma dolerite dikes of the Franklin Large Igneous Province record a mean virtual dipole moment of  In the Phanerozoic, multiple low paleointensities has been observed in Devonian-aged volcanic units (Fig.
-\ref{fig:PINT_compilation}; \citealp{Shcherbakova2017a, Shcherbakova2021a, Hawkins2021a}).
-
-%The observed paleointensity variations from the Midcontinent Rift also question the attribution to the onset of inner core nucleation as the unique cause for driving a significant paleointensity increase.
-Besides the
-
-% Accompanying the low paleointensity values, paleodirectional data from the Ediacaran-Cambrian \cite{Bono2015a, Halls2015b, Kodama2021a} and the Devonian show anomalous directional fluctuations \cite{Shcherbakova2019a} which could be consistent with there being a largely diminished dipole component of Earth's magnetic field.
-If the weak paleointensities in the Ediacaran or the Devonian are attributed to the geodynamo being on the brink of collapse just prior to the nucleation of the solid inner core.
-
-% if one invokes inner core nucleation for all these low to high transitions, the data would suggest the inner core nucleated multiple times.
-Alternatively, it could that all except one are not related to the inner core nucleation, but merely a manifestation of short term dynamo variations.
-% n narrow time scales regardless of the existence of a solid inner core.
-Therefore, low paleointensity values or short-term fluctuations may not be sufficient evidence for interpreting Earth's inner core status through history.
-
-% The high geomagnetic field estimates from the Midcontinent Rift volcanics and intrusives are not the only suite of high paleointensity records at odds with the hypothesis of a monotonically decaying geodynamo through the entire Proterozoic Eon \cite{Biggin2009a, Bono2019a}.
-The $\sim$60 ZAm$^2$ virtual dipole moment estimates from the ca.
-755 Ma Mundine Well dikes are much higher than the values predicted by the monotonic decrease model of \cite{Bono2019a}.
-Although one may consider these data capturing relatively short term fluctuations of the geomagnetic field, they nevertheless indicate that the geodynamo had strong power sources during the Proterozoic.
-On the other hand, besides the low paleointensity values recorded by the ca.
-1140 Ma Abitibi dikes, low estimates have been obtained from the ca.
-1070 Ma Bangemall Sills \cite{Lloyd2021b}, the ca.
-720 Ma dolerite dikes of the Franklin Large Igneous Province \cite{Lloyd2021a}, as well as Devonian-aged rocks from the Minusa Basin, Kola Peninsula, Buribay volcanic complex and volcanic rocks from the East European Craton (Fig.
-\ref{fig:PINT_compilation};  \citealp{Shcherbakova2017a, Shcherbakova2019a, Shcherbakova2021a}).
-The multiple occurrences of low geomagnetic field intensities during the Proterozoic and the Paleozoic indicate that a waning thermal convection regime in Earth's before the emergence of the solid inner core may not be the solely possible mechanism for triggering an weak dipole field.
-
-% In the meantime, paleomagnetic directional anomalies may be another suite of intriguing observational records that is worth investigating.
-The anomalous directional fluctuations during the Ediacaran and Devonian \cite{Bono2015a, Kodama2021a, Shcherbakova2019a} perhaps can be consistent with there being a weak geodynamo regime when dipole component of Earth's field was largely diminished.
-
-% conclude on whether an apparent rise or decay within a narrow time period ($\sim$35 Ma) during the Proterozoic is associated with transitions in the core phase transition or is a result of short-term excursion of the geomagnetic field that may be common through much of the evolution of Earth's core.
-
-% an apparent rapid decrease in geomagnetic dipole moment from ca.
-1092 Ma to ca.
-1070 Ma can be evidenced by the contrasting high values from the Midcontinent Rift anorthosite xenoliths and the low values from the Bangemall Sills (Fig.
-\ref{fig:PINT_compilation};  \citealp{Lloyd2021b}).
-These data suggest that an apparent transition of virtual dipole moment similar in scale with what is predicted for a dynamo powered with or without compositional convection in the core may not be unique during the Precambrian.
-In addition, given that a five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years, when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed, we cannot conclude on whether an apparent rise or decay within a narrow time period ($\sim$35 Ma) during the Proterozoic is associated with transitions in the core phase transition or is a result of short-term excursion of the geomagnetic field that may be common through much of the evolution of Earth's core.
-Recently, ultra-low paleointensities (2.01–7.07) $\mu$T from the Devonian has been found by  \cite{Shcherbakova2021a}.
-These data suggest that significant excursion of low geomagnetic field intensities may occur regardless of the inner core formation.
-
-% These data suggest that an apparent transition of virtual dipole moment similar in scale with what is predicted for a dynamo powered with or without compositional convection in the core may not be unique during the Precambrian.
-In addition, given that a five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years, when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed, we cannot conclude on whether an apparent rise or decay within a narrow time period ($\sim$35 Ma) during the Proterozoic is associated with transitions in the core phase transition or is a result of short-term excursion of the geomagnetic field that may be common through much of the evolution of Earth's core.
-
-% Paleomagnetic observational records near the Ediacaran-Cambrian boundary have been interpreted to be consistent with a trend of decreasing axial dipole strength toward a very low value \cite{Lloyd2021a, Bono2019a}, followed by a rapid recovery (a few tens of million years) toward higher values soon after \cite{Thallner2021a}.
-Such a decrease-increase transition can be consistent with dynamo numerical simulations which adopt a rather high thermal conductivity of the core and projects an inner core birth near this time \cite{Gomi2013a, Davies2015a, Labrosse2015a, Ohta2016a}.
-In particular, some models have suggested that the kick-start of rigorous compositional convection between the inner core and the outer core due to the onset of inner core nucleation can result in almost an order of magnitude increase in Earth's axial dipole moment \cite[e.g.][]{Davies2021a}.
-The enriching observational records from the Ediacaran and Cambrian have been suggested to indicate that a geodynamo driven solely by thermal convection was waning to its brink of collapse near the end of Ediacaran \cite{Bono2019a, Thallner2021b} and was recovering toward modern-day strengths in the Cambrian \cite{Thallner2021a}.
-
-% However, the updated paleointensity records from the Midcontinent Rift together with previous results from other Mesoproterozoic rocks suggest that the a dramatic increase in geomagnetic dipole moment similar in scale to that modeled for after the onset of inner core nucleation may not be unique to the Ediacaran-Cambrian boundary (Fig.
-\ref{fig:PINT_compilation}).
-The ca.
-1140 Ma Abitibi dikes \cite{Macouin2003a} record an average virtual dipole moment of $\sim$13 ZAm$^2$, similar to the low values of $\sim$7 ZAm$^2$ obtained from the single silicate crystals of the ca.
-565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a}.
-However, the relatively high moments of $\sim$72 ZAm$^2$ from the ca.
-1105 Ma Osler Volcanic Group \cite{Sprain2018a} and $\sim$89 ZAm$^2$ from the ca.
-1092 Ma Beaver River anorthosite xenoliths (Fig.
-\ref{fig:PINT_compilation}) are about six times the values from the Abitibi dikes.
-This increase in absolute dipole moment estimates during $\sim$50 Myr recorded by the Midcontinent Rift rocks are much higher than that in the Ediacaran (Fig.
-\ref{fig:PINT_compilation}).
-
-% observation of an increase or a decrease in virtual dipole moment may not be unique during the Precambrian.
-A sparse spatial and temporal coverage of the Precambrian paleointensity records may not be sufficient for further analyses of the long-term evolution of the geodynamo.
-Given that a five-fold fluctuation of the geomagnetic dipole moment exist in the paleointensity records for the past 10 million years (5 and 95 percentiles), when we know a modern-day-like geodynamo likely powered by both thermal and compositional convection existed, we cannot conclude on whether an apparent rise or decay within a narrow time period ($\sim$35 Ma) during the Proterozoic is associated with transitions in the core fluid regime or is a result of short-term variations of the geomagnetic field that may be common through much of the evolution of Earth's core.
-Regardless, the high geomagnetic dipole field moment estimate from the Beaver River anorthosite xenoliths necessitates that there being a vigorous late Mesoproterozoic geodynamo.
-
-% Overall, the existing absolute paleointensity records in the PINT database remains sparse in terms of spatial and temporal coverage as compared to more geologically recent times.
-(Fig.
-\ref{fig:PINT_compilation}).
-More data are needed in order to better characterize the evolution of Earth's dynamo and thermal evolution.
-
-% \section*{Conclusion}
-
-% The paucity of Precambrian paleointensity estimates, as well as non-ideal results within the existing database, present challenges when seeking to interpret long-term trends in the strength of the geomagnetic field.
-The Midcontinent Rift of North America presents an opportunity to develop robust paleointensity estimates spanning a $\sim$25 Myr interval of the late Mesoproterozoic.
-Rock magnetic data support that the ca.
-1092 Ma anorthosite xenoliths of the Midcontinent Rift diabase can have dominantly high coercivity magnetic mineralogy with minimal anisotropic or nonlinear thermal remanence acquisition, producing high quality paleointensity results.
-Their overall high estimates of the late Mesoproterozoic virtual dipole moment together with previous Midcontinent Rift records suggest that a strong geomagnetic field like today's might have persisted through the rifting.
-
-% The new data in context of the current Precambrian paleointensity database suggest that an observation of a strong increase in geomagnetic dipole moment may not be unique during the Precambrian.
-More spatial and temporal coverage of the observational records is needed for further analyses of the long-term evolution of the geodynamo.
 
@@
 
-% After the paleointensity experiment were finished, the same specimens underwent additional TRM anisotropy checks with full TRMs given at 0, 30, 60, 90, 120, 150 degrees in the horizontal plane in specimen coordinates.
-After that, full TRM acquisition experiments were performed on the same specimens with applied fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axis.
-
 \subsection*{Paleointensity result selection}
@@
 Data analysis was conducted using Thellier GUI \cite{Shaar2013a} within the PmagPy software package \cite{Tauxe2016a}.
-
-% \begin{table*}
-% \caption{Quality criteria used for selecting paleointensity data. See the text for more detail.} 
-% \centering
-% \begin{tabular}{ccccccccc}
-% \hline
-% MAD (\textdegree) & Beta (per cent) & DANG (\textdegree) & FRAC& SCAT & GAP-MAX & NpTRM & N Arai\\
-% \hline
-% 20 & 15 & 5 & 0.6 & TRUE & 0.25 & 2 & 4 \\
-% \hline
-
-% \end{tabular}
-% \label{tab:criteria}
-% \end{table*}
 
@@
 \end{document}
-
-% \subsection*{Manuscript Length}
-
-% A standard 6-page article is approximately 4,000 words, 50 references, and 4 medium-size graphical elements (i.e., figures and tables).
-The preferred length of articles remains at 6 pages, but PNAS will allow articles up to a maximum of 12 pages.
-
-% \subsection*{References}
-
-% References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g.
-\cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric}.
-All references cited in the main text should be included in the main manuscript file.
-
-% \subsection*{Data Archival}
-
-% PNAS must be able to archive the data essential to a published article.
-Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the \href{https://www.pnas.org/page/authors/journal-policies#xi}{Information for Authors}, is acceptable.
-
-% \begin{SCfigure*}[\sidecaptionrelwidth][t]
-% \centering
-% \includegraphics[width=11.4cm,height=11.4cm]{frog}
-% \caption{This legend would be placed at the side of the figure, rather than below it.}
-% \label{fig:side}
-% \end{SCfigure*}
-
-% \subsection*{Digital Figures}
-
-% EPS, high-resolution PDF, and PowerPoint are preferred formats for figures that will be used in the main manuscript.
-Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.
-Color images must be in RGB (red, green, blue) mode.
-Include the font files for any text.
-
-% Images must be provided at final size, preferably 1 column width (8.7cm).
-Figures wider than 1 column should be sized to 11.4cm or 17.8cm wide.
-Numbers, letters, and symbols should be no smaller than 6 points (2mm) and no larger than 12 points (6mm) after reduction and must be consistent.
-
-% Figures and tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.
-
-% Fig.
-\ref{fig:frog} shows an example of how to insert a column-wide figure.
-To insert a Fig.
-wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment.
-Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
-Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide Fig.
-with side legends.
-
-% \subsection*{Tables}
-% Tables should be included in the main manuscript file and should not be uploaded separately.
-
-% \begin{table}%[tbhp]
-% \centering
-% \caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
-% \begin{tabular}{lrrr}
-% Species & CBS & CV & G3 \\
-% \midrule
-% 1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
-% 2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
-% 3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
-% \bottomrule
-% \end{tabular}
-
-% \addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
-% \end{table}
-
-% \subsection*{Supporting Information Appendix (SI)}
-
-% Authors should submit SI as a single separate SI Appendix PDF file, combining all text, figures, tables, movie legends, and SI references.
-SI will be published as provided by the authors; it will not be edited or composed.
-Additional details can be found in the \href{https://www.pnas.org/authors/submitting-your-manuscript#manuscript-formatting-guidelines}{PNAS Author Center}.
-The PNAS Overleaf SI template can be found \href{https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd}{here}.
-Refer to the SI Appendix in the manuscript at an appropriate point in the text.
-Number supporting figures and tables starting with S1, S2, etc.
-
-% Authors who place detailed materials and methods in an SI Appendix must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the SI methods.
-If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.
-
-% \subsubsection*{SI Datasets}
-
-% Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be published in raw format and will not be edited or composed.
-
-%EXTRA TEXT
-
-% Without an inner core, Earth's magnetic field would have been generated solely through CMB heat flow, with perhaps additional stirring by other power sources such as Mg precipitation \cite{ORourke2016a}, or by exotic processes such as a convecting basal magma ocean \cite{Ziegler2013a}.
-Long-term secular cooling of Earth has the potential to have decreased the power available to the geodynamo through time prior to the nucleation of the inner core \cite[e.g][]{Davies2021a}.
-As a result, it is possible that intensity of the geomagnetic axial dipole field would have experienced a long-term decay before the emergence of the solid inner core after which it would have strengthened \cite{Aubert2009a, Davies2021a}.
-However, the estimates for thermal conductivity values for the core continue to vary \cite{Konopkova2016a, Ohta2016a, Zhang2020a}.
-Obtaining an observational record of the timing of inner core formation would constrain the thermal evolution of the core and mantle, in addition to improving our understanding of the history of Earth’s dynamo.
-
-% Interpretations of selected records of the current Precambrian absolute paleointensity database (compiled in the PINT database; \url{http://earth.liv.ac.uk/pint/}) suggest that an appreciable geomagnetic field existed in the Archean but reached its weakest in the Ediacaran when very low virtual dipole moment (VDM) estimates have been obtained \cite{Bono2019a, Shcherbakova2019a, Thallner2021b, Lloyd2021a}, followed by a recovery near the Ediacaran-Cambrian boundary \cite{Thallner2021a}.
-Ref.
-\citealp{Bono2019a} adopted a billion-year-scale view of such observations to suggest that these data are consistent with thermal evolution models that use high core thermal conductivity values, where the nucleation of the inner core did not begin to provide compositional convection power for driving the geodynamo until the Ediacaran-Cambrian, suggesting that the Proterozoic dynamo might have experienced a monotonic decay when compositional convection was lacking and thermal convection was progressively decreasing \cite{Driscoll2016a, Davies2021a}.
-
-% Other studies focusing on paleomagnetic directions have also found that the late Ediacaran might have been a period of time when the geomagnetic field were experiencing frequent reversals and waning of dipole dominance \cite{Bono2015a, Kodama2020a}.
-
-% However, previous paleointensity data from intrusive and extrusive rocks in the ca.
-1.1 Ga Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a} record VDM values similar to that of today, which are higher than the values predicted by the long term decay curve proposed by  \citealp{Bono2019a} (Fig.
-\ref{fig:PINT_compilation}).
-Ref.
-\citealp{Pesonen1983a} obtained an average virtual dipole moment of 93 $\times 10^{21}$ Am$^2$ (ZAm$^2$) after applying a 13\% cooling rate correction (114 ZAm$^2$ before correction) using both extrusive and intrusive rocks from the Midcontinent Rift, including the ca.
-1108-1105 Ma Osler Volcanics, the ca.
-1100 Ma Maimainse Point volcanics, the ca.
-1107 Ma intrusive Logan Sills, Thunder Bay dikes, the Baraga and Marquette dikes, as well as lava flows of the ca.
-1096 Ma North Shore Volcanic Group (Fig.
-\ref{fig:Geologic_map}).
-Ref.
-\citealp{Kulakov2013a} obtained an average VDM of 59 $\pm$ 11 ZAm$^2$ using the ca.
-1085 Ma Lake Shore Traps volcanics and suggest that the present-day-like geomagnetic field strength likely existed through the late stage of Midcontinent Rift magmatism \cite{Vervoort2007a, Miller2013a}.
-However, non-ideal paleointensity behaviors at specimen level and overall low success rates have challenged the interpretation of many previous paleointensity results obtained by  \citealp{Pesonen1983a, Kulakov2013a}.
-One typical failure of paleointensity experiment is the double-slope behavior, as two distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope: a higher paleointensity estimate from the low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
-Ref \citealp{Pesonen1983a} used the low-temperature slope as the best representation of the past magnetic field strength (likely overestimating the field strength) whereas  \citealp{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
-Results with such non-ideal behavior were rejected by  \citealp{Sprain2018a} who applied stricter paleointensity selection criteria on new paleointensity results and accepted only 23 out of 133 specimens (5 out of 21 lava flows) from the Osler Volcanics and the Mamainse Point volcanics.
-A calculated mean VDM value of 56 $\pm$ 21 ZAm$^2$ is similar to the average value of the past 300 million years \cite{Sprain2018a}.
-
-% Although it is typical of paleointensity experiments to have low success rates, bulk plagioclase-rich rocks \cite{Selkin2000a} and single plagioclase crystals \cite{Tarduno2005a} are found to be able to produce high quality paleointensity results.
-It has been suggested that plagioclase crystals can protect magnetic minerals that exsolved within them against post-formation weathering and thermochemical alteration during paleointensity experiments \cite{Tarduno2005a}.
-In this study, we present new paleointensity results from IZZ modified absolute paleointensity experiments \cite{Yu2004a} on a suite of bulk samples from ca.
-1092 Ma anorthosite xenoliths and their Beaver River diabase host rocks of the Beaver Bay Complex within the Midcontinent Rift.
-Paleointensity experiments on the anorthosite xenoliths have a very high success rate.
-Their consistent specimen- and site-level paleointensity results together with rock magnetic data reveal that the anorthosite xenoliths have low anisotropy of thermal remanent magnetization (TRM) and can acquire TRM linearly.
-Magnetic imaging shows the anorthosite specimens have dominant magnetic carriers interstitial to and within plagioclase crystals and they do not display strong preferred orientations.
-Our single-slope, high quality paleointensity data add to the reliable paleointensity records during the $\sim$25 Myr of Midcontinent Rift magmatic activity.
-Our new data from the Midcontinent Rift intrusive rocks support previous data from the volcanics that there was a strong late Mesoproterozoic geodynamo.
-
-% The Mesoproterozoic North American Midcontinent Rift is an intracontinental rift that did not result in the break-up of Laurentia.
-Although the protracted magmatic activity in the rift lasted form ca.
-1109 Ma to ca.
-1084 Ma \cite{Swanson-Hysell2019a}, its activity during the main magmatic stage ($\sim$1098–1090 Ma; refs.
-\citealp{Vervoort2007a, Miller2013a}) was characterized by rapid and voluminous emplacement of intrusive rocks and extrusive volcanics.
-The $<$500 Kyr emplacement of the ca.
-1096 Ma Duluth Complex and the associated $\sim$8 km thick extrusive volcanics is one manifestation \cite{Swanson-Hysell2021a}.
-The emplacement of the ca.
-1092 Ma Beaver Bay Complex in northern Minnesota punctuates another such period of magmatism during the Midcontinent Rift main stage activity.
-The hypabyssal intrusion of the 1091.7 $\pm$ 0.2 Ma Beaver River diabase dikes and sills of the Beaver Bay Complex transported numerous anorthosite xenoliths that have short-axis diameters up to 180 meters via wide conduits \cite{Boerboom2004a, Boerboom2006b}.
-Geochronology, geochemical, and paleomagnetic data support the hypothesis that the diabase extruded to the surface and fed the coeval voluminous Greenstone Flow of the Portage Lake volcanics \cite{Doyle2016a, Zhang2021b}.
-
-% The dominantly monomineralic anorthosite xenoliths of the Beaver River diabase have an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}, higher than the average of $\sim$60\% of the Anorthositic Series rocks of the Duluth Complex \cite{Miller1990a}.
-The plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture showing igneous foliations (Fig.
-\ref{fig:Petro_QDM}A), but the more calcium-rich plagioclase of the Beaver River anorthosites often develop granoblastic texture characterized by equigranular crystals with poor alignments (Fig.
-\ref{fig:Petro_QDM}B).
-Abundant magnetite and ilmenite needles that are typically 10 $\mu$m long can be found to have preferred orientations within Duluth Complex plagioclase.
-Interstitial titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex plagioclase \cite{Miller1990a}.
-In contrast, titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent but relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
-\ref{fig:Petro_QDM}).
-This reflects that the Beaver River diabase magma had a different oxidation history than that of the Duluth Complex.
-The plagioclase and oxide mineral textures of the Beaver River anorthosites also distinguish them from plagioclase cumulates of other layered mafic complex where magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
-
-% Such a high concentration of Fe oxides carrying magnetic remanence in the slowly cooled Duluth Complex anorthositic rocks reflect that iron in the plagioclase mush in the magma chamber acted as a means of buffering the melt with respect to oxygen, forming crystallites along preexisting plagioclase substrate or as iron oxides incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures \cite{Scofield1986a}.
-In contrast, titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent but relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
-\ref{fig:Petro_QDM}).
-% the Beaver River diabase rapidly ascended and entrained the anorthosite xenoliths directly into shallow chambers and likely did not reach  \cite{Flower1980a, Miller1997a, Zhang2021b}.
-Particularly, \cite{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
-This is consistent with the interpretation that these anorthosite were rapidly transported by upward currents into hypabyssal magma chambers without reaching structural equilibrium.
-Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Fig.
-\ref{Petro_QDM}).
-Instead, relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
-\ref{Petro_QDM}).
-
-% In the Beaver River anorthosite xenoliths, relatively sparse titano-magnetite occur within or interstitial to the plagioclase (Fig. \ref{Petro_QDM}).
-
-% Occurrences of magnetite and ilmenite symplectites and needles within anorthositic rocks have been described by \cite{Scofield1975a, Scofield1986a} in the Wichita Mountains and by \cite{Haggerty1976a} in other slowly cooled, mafic intrusions.
-\cite{Scofield1986a} hypothesized that the inclusions acted as a means of buffering the melt with respect to oxygen, forming either crystallites along preexisting plagioclase substrate or as iron incorporated into the plagioclase that subsequently exsolved at subsolidus temperatures.
-Such a model requires that plagioclase, instead of other mafic minerals such as olivine, to be the first to crystallize from a magma and act as a sink for iron.
-The two-stage formation of the Duluth Complex Anorthositic Series fits this model \cite{Miller1990a}.
-During the first stage, high pressure at subcrustal depths (15-25 km depth) caused fractionation of Duluth Complex magma resulting in the formation of floating plagioclase mushes \cite{Kushiro1980a}.
-The second stage of crystallization then occurs at low pressures after the magma intruded into shallower chambers.
-
-% This proposed magmatic evolution of the Duluth Complex is in distinct contrast to the emplacement of the Beaver River diabase which rapidly ascended and entrained the anorthosite xenoliths directly into shallow chambers \cite{Flower1980a, Miller1997a, Zhang2021b}.
-Particularly, \cite{Miller1997a} observed a high state of structural disorder within plagioclase of the Beaver River anorthosite using X-ray diffraction analyses.
-This is consistent with the interpretation that these anorthosite were rapidly transported by upward currents into hypabyssal magma chambers without reaching structural equilibrium.
-Within the Beaver River anorthosite xenoliths, the titanomagnetite needles and the symplectic magnetite-ilmenite intergrowths are absent (Fig.
-\ref{Petro_QDM}).
-Instead, relatively sparse titano-magnetite within or interstitial to the plagioclase are typical of Beaver River anorthosites (Fig.
-\ref{Petro_QDM}).
-
-% it is the most probable that the anorthosite xenoliths of the Beaver River diabase are cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
-Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths acquired thermal remanent magnetizations during cooling of the diabase \cite{Zhang2021b}.
-Step-wise thermal demagnetizations show many anorthosite xenoliths have minimal secondary remanence and single-component characteristic remanence magnetizations that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C.
-
-% Geochronological, paleomagnetic, geochemical, and petrologic data are consistent with the hypothesis that the Beaver River diabase magma extruded to the surface, feeding the massive Greenstone Flow - one of the largest lava flows on Earth \cite{Doyle2016a, Zhang2021b}.
-
-% Based on high-precision $^{206}$Pb/$^{238}$U geochronology and paleomagnetic data, \cite{Swanson-Hysell2021a} hypothesized that the massive Duluth Complex and the coeval North Shore Volcanic group (Fig.
-\ref{fig:Geologic_map}) formed as a result of upwelling mantled-derived plume encountering Laurentia's thinned lithosphere, following an ``upside-down” drainage" which rapidly funneled melt to today's northeastern Minnesota (Fig.
-\ref{fig:Geologic_map}) \cite{Sleep1997a, Swanson-Hysell2014a}).
-The plagioclase-phyric Anorthositic Series rocks, the crystal-poor Layered Series intrusions of the Duluth Complex, and the coeval volcanics of the North Shore Volcanic Group (NSVG) that outcrop along Northeastern Minnesota (Fig.
-\ref{fig:Geologic_map}) are interpreted to be a result of such vigorous magmatic activity.
-
-% Despite that the anorthosite xenoliths of the Beaver River diabase were entrained as xenoliths and emplaced only $\sim$4 Myr after the formation of the Anorthositic Series rocks of the Duluth Complex when massive plagioclase mushes were generated, $^{206}$Pb/$^{238}$U zircon geochronology data and zircon rare earth element diffusion modeling results suggest it is the most probable that the Beaver River diabase anorthosite xenoliths are entrained cumulate enclaves that formed at the time of Beaver Bay Complex magmatism.
-The anorthosite xenoliths of the ca.
-1092 Ma Beaver River diabase has distinct mineralogy from that of the Anorthositic Series rocks of the Duluth Complex.
-The dominantly monomineralic anorthosite xenoliths have an average anorthite content of $\sim$70\% \cite{Morrison1983a, Doyle2016a}, higher than the average of $\sim$60\% of the Anorthositic Series rocks of the Duluth Complex \cite{Miller1990a}.
-In addition, petrographic imaging support that the plagioclase from the Beaver River anorthosite formed under a magmatic conditions different from the Duluth Complex anorthositic series magma.
-While the plagioclase in the Duluth Complex Anorthositic Series typically are euhedral with an interlocking texture (Fig.
-\ref{Petro_QDM}A), the more calcium-rich plagioclase of the Beaver River anorthosite often develop granoblastic texsture characterized by equigranular plagioclase crystals (Fig.
-\ref{Petro_QDM}B).
-Abundant magnetite and ilmenite needles that are typically 10 $\mu$m long can be found to have preferred orientations within Duluth Complex plagioclase.
-Interstitial vermicular titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex Anorthositic Series rocks \cite{Miller1990a}.
-
-% Paleointensity experiment results from the anorthosite xenoliths and their host diabase show three general types of dominant specimen behaviors on NRM/TRM plots (Fig.
-\ref{fig:IZZI_examples}): dominantly single-slope, double-slope or sagging, and zigzagging.
-
-% As is shown in Fig.
-\ref{fig:IZZI_examples}, specimen BD2-5b has a high-temperature slope with zigzagging behavior, poor pTRM checks, and low NRM/TRM ratio, in addition to a low-temperature slope with a steep NRM/TRM ratio.
-Although the BD2 specimens' lower temperature fraction of the plots are straight and the pTRM checks are consistent, lower temperature slopes from BD2 give consistently higher paleointensity estimates than those from site AX6, AX8, and AX10, which are anorthosite xenoliths included within BD2 (Fig.
-\ref{fig:Geologic_map};  \citealp{Zhang2021a}.
-This higher estimate from the diabase lithology could be attributed to the systematic bias of using the lower temperature slope of the NRM/TRM plots demonstrated in Fig.
-\ref{fig:IZZI_examples} and specimens from site BD2 all fail the criteria of having a maximum magnetization gap between adjacent steps of 25\% of the total NRM.
-Eventually, we only accept the single-slope anorthosite specimens results for further analyses.
-
-% \begin{figure}[h!]
-% \noindent\includegraphics[width=\textwidth]{Cooling_rate_correction.pdf}
-% \centering
-% \caption{\small{}}
-% \label{fig:cooling_correction}
-% \end{figure}
-
-% A 20 mT alternating field treatment after in-field heating steps were applied to some specimens in an attempt to mitigate non-ideal behavior contributed by multidomain or non-uniformly magnetized magnetic carriers.
-The treatment did not make any diabase specimen to pass the selection criteria.
-On the other hand, 27 out of 58 ($\sim$46.6\%) anorthosite specimens that had AF treatment passed selection and 13 out of 28 ($\sim$46.4\%) specimens without AF treatment passed selection.
-Paleointensity estimates from specimens with or without AF treatment in anorthosite AX10, AX11, AX12, AX13, AX16 do not show significant difference (Fig.
-\ref{fig:PINT_cooling_corrected}).
-Specimen results for AX11 particularly illustrates a case where all specimens within the site yield almost the same paleointensity estimate, regardless of the AF treatment.
-
-% Therefore, the alternating field demagnetization did not result in significant increase in specimen success rate for the anorthosites although it might have improved the specimen behavior for those anorthosites that were subjected to this treatment.
-Together with the narrow range of unblocking temperatures of their sister specimens \cite{Zhang2021b}, these data are consistent with the interpretation that the anorthosites with successful paleointensity results have a magnetic mineral assemblage that is likely dominated by single domain, low-Ti magnetite.
-Some anorthosite specimens and all diabase specimens still failed during result selections even though AF treatment was applied.
-It could be that they have a very high dominance of multidomain (MD) or vortex state grains with relatively moderate coercivity such that an alternating field with 20 mT peak field did not significantly mitigate their non-ideal paleointensity behaviors.
-Another cause for their failure could be the thermochemical alterations that happened during heating in air.
-As can be exampled by the zigzagging high temperature slope with poor pTRM checks in paleointensity plot for specimen BD2-5b (Fig.
-\ref{fig:IZZI_examples}), the decrease in NRM/TRM ratios during higher temperature steps likely resulted from the formation of new ferromagnetic material within the specimen due to oxidation.
-Such alterations can allow specimens to acquire additional remanence during in-field steps that may not be fully removed by an AF treatment.
-
-% Further, we applied full TRMs to the anorthosite specimens that pass the paleointensity selection criteria after paleointensity experiment.
-With an applied field of 30 $\mu$T oriented toward the specimen at 0, 30, 60, 90, 120, and 150 degrees, the anorthosite specimens usually acquired consistent full TRMs.
-The most significant variations observed in acquired full TRMs reach to only XXX \%, much less than that has been reported to occurred within layerd anorthositic rocks in Stillwater (up to $\sim$60 \%, \cite{Selkin2000a}).
-
-% Ref.
-\citealp{Selkin2007a} also noted that vortex state magnetic grains and certain single domain particles can produce biased paleointensity estimates due to their acquisition of saturation remanence during laboratory in-field steps.
-According to the numerical modeling of  \citealp{Selkin2007a}, thermoremanent magnetization acquired by a population of randomly-oriented, elongate, single-domain magnetite particles are unlikely to saturate upon our applied field of 30 $\mu$T.
-Therefore, in this study we do not correct for anisotropy or nonlinear TRM to our specimens.
-
-% We applied full TRMs to our anorthosite specimens with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axis.
-The results show that our anorthosite specimens did not saturate upon an applied field of 30 $\mu$T and there is thus no need for correcting nonlinear thermal remanence acquisitions.
-% Further, we applied full TRMs to the anorthosite specimens that pass the paleointensity selection criteria after paleointensity experiment.
-With an applied field of 30 $\mu$T oriented toward the specimen at 0, 30, 60, 90, 120, and 150 degrees, the anorthosite specimens usually acquired consistent full TRMs.
-The most significant variations observed in acquired full TRMs reach to only XXX \%, much less than that has been reported to occurred within layerd anorthositic rocks in Stillwater (up to $\sim$60 \%, \cite{Selkin2000a}).
-
-% The reason that the Beaver River anorthosite xenoliths do not display strong remanence anisotropy, nor do they display nonlinear TRM acquisition problem, could be that they formed in a magmatic setting distinct from that of a layered mafic complex such as the nearby Duluth Complex, where igneous foliation can occur (Fig.
-\ref{fig:Geologic_map}).
-Petrographic images show that an equigranular texture where equidimensional, randomly oriented plagioclase often develop in the anorthosite xenoliths (Fig.
-\ref{fig:Petro_QDM}; \cite{Zhang2021b}).
-This is in contrast to the interlocking igneous foliation texture within the Anorthositic Series rocks of the ca.
-1096 Ma Duluth Complex (Fig.
-\ref{fig:Petro_QDM}).
-Further, petrographic and magnetic imaging also shows the Beaver River anorthosite xenoliths do not develop abundant exsolved magnetite and ilmenite needles within plagioclase crystals as the slowly-cooled anorthositic intrusions do, in which exsolved magnetite needles can be the dominate remanence carriers (Fig.
-\ref{fig:Petro_QDM}; \cite{Scofield1986a, Feinberg2006a}).
-The fact that all dipole features within a same Duluth Complex plagioclase crystal have indistinguishable orientations is consistent with the interpretation of there being a magnetic remanence anisotropy resulting from preferred petrofabric alignment, which could give rise to the remanence anisotropy and nonlinear TRM acquisition of layered anorthositic intrusions (Fig.
-\ref{fig:Petro_QDM}).

### 2022-01-28 PNAS_manuscript/PNAS_manuscript.tex

@@
 \documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
-% Use the lineno option to display guide line numbers if required.
 
 \templatetype{pnasresearcharticle}
-% {pnasresearcharticle} = Template for a two-column research article
-% {pnasmathematics} %= Template for a one-column mathematics article
-% {pnasinvited} %= Template for a PNAS invited submission
 
@@
 
-% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
 \author[a,1]{Yiming Zhang}
@@
 
-% Please give the surname of the lead author for the running footer
 \leadauthor{Zhang}
 
-% Please add a significance statement to explain the relevance of your work
 \significancestatement{Acquiring high-fidelity observational records of ancient magnetic field intensity from the remanent magnetization of rocks is crucial for constraining the long-term evolution of Earth's core.
@@
 An average value of $\sim$89 ZAm$^2$ for the virtual dipole moment from the anorthosites with the highest site-level values up to $\sim$140 ZAm$^2$ are higher than that of the dipole component of Earth's magnetic field today and rival the highest values in the paleointensity database.
-Such high intensities recorded by the anorthosite xenoliths requires the existence of a strongly powered geodynamo at the time.
-Together with previous paleointensity data from other Midcontinent Rift rocks, our new results indicate that a dynamo with strong power sources persisted for more than 20 million years ca.
+Such high intensities recorded by the anorthosite xenoliths require the existence of a strongly powered geodynamo at the time.
+Together with previous paleointensity data from other Midcontinent Rift rocks, our new results indicate that a dynamo with strong power sources persisted for more than 14 million years ca.
 1.1 Ga.
@@
 \citealp{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
-Results with such non-ideal behavior were rejected by \citealp{Sprain2018a} who applied stricter paleointensity selection criteria, but as a result had few accepted sites.
+Such non-ideal results were rejected by \citealp{Sprain2018a} who applied stricter paleointensity selection criteria, but as a result had few accepted sites.
 
@@
 
-Following IZZI-modified paleointensity experiments \cite{Yu2004a}, 40 from a total of 86 anorthosite specimens and 0 out of a total of 69 diabase specimens passed our paleointensity result selection criteria (see Materials and Methods section).
+Following IZZI-style paleointensity experiments \cite{Yu2004a}, 40 from a total of 86 anorthosite specimens and 0 out of a total of 69 diabase specimens passed our paleointensity result selection criteria (see Materials and Methods section).
 7 anorthosite sites and no diabase site-level results pass these selection criteria.
@@
 \ref{fig:coercivity}).
-A compilation of median destructive field (MDF) values in Fig.
+A compilation of median destructive field (MDF) values in Figure.
 \ref{fig:coercivity} shows that specimens from anorthosites that pass paleointensity selection criteria have distinctly higher average MDF values than other anorthosite and diabase specimens.
@@
 We applied full TRMs to a set of anorthosite specimens with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axes.
-The results show that this set of anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions ( \citealp{Selkin2007a}; Fig.
+The results show that this set of anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions (\citealp{Selkin2007a}; Fig.
 S4).
@@
 In addition to these high geomagnetic fields recorded by Midcontinent Rift rocks, the ca.
-755 Ma Mundine Well dikes \cite{Lloyd2021b} also require a stronger geomagnetic field in the Neoproterozoic than would be predict by a progressive Proterozoic decline (Fig.
+755 Ma Mundine Well dikes \cite{Lloyd2021b} also require a stronger geomagnetic field in the Neoproterozoic than would be predicted by a progressive Proterozoic decline (Fig.
 \ref{fig:PINT_compilation}).
@@
 This timing would be broadly consistent with that proposed by \citealp{Biggin2015a}.
-However, a model prediction could be of sustained strong field values following inner core nucleation which is challenged by data from the ca.
+However, a model prediction of sustained strong field values following inner core nucleation is challenged by data from the ca.
 1070 Ma Bangemall Sills which include a sill with a low virtual dipole moment of $\sim$6.4 ZAm$^2$ (Fig.
@@
 In such models, relatively low total heat flux across the core-mantle boundary can prevent the axial dipole from reversing whereas a high heat flux through the boundary can result in an increase in reversal frequency and decrease in dipole intensity.
-The ``Mid-Paleozoic Dipole Low'' has been hypothesized to be the result of such elevated core-mantle boundary heat flux conditions at a time when there was also available power from a crystallizing inner core \cite{Hawkins2019a}, thereby explaining the observed low paleointensities which include values as weak as those of the ca.
-565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a} (Fig.
-\ref{fig:PINT_compilation}).
+The ``Mid-Paleozoic Dipole Low'' has been hypothesized to be the result of such elevated core-mantle boundary heat flux conditions at a time when there was also available power from a crystallizing inner core \cite{Hawkins2019a}, thereby also explaining the observed low paleointensities which include values as weak as those of the ca.
+565 Ma Sept-$\hat{I}$les intrusions (Fig.
+\ref{fig:PINT_compilation}; \citealp{Bono2019a}).
 Mantle convection can modulate core mantle boundary heat flow through changes in the structure of the deep mantle associated with upwelling plumes \cite{Larson1991a, Courtillot2007a} and subducted slabs \cite{Tan2002b, Biggin2012a, Hounslow2018a}.
@@
 Conversations with Bruce Buffett and Zachary Geballe informed our perspectives on the geodynamic and mineral physics literature.
-Paleomagnetic data associated with this study are available within the MagIC database (\url{https://earthorg/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}; \textit{provide link when doi is generated at time of proofs}) and all data are within a Github repository associated with this work along with Python code related to calculations and visualizations associated with the study that is also archived on Zenodo (\textit{provide link when doi is generated at time of proofs}).}
+Paleomagnetic data associated with this study are available within the MagIC database (\url{https://earthorg/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}; \textit{provide link when doi is generated at time of proofs}) and all data are within a Github repository along with Python code related to calculations and visualizations (\url{https://github.com/Swanson-Hysell-Group/AX_BD_PINT}) that are also archived on Zenodo (\textit{provide link when doi is generated at time of proofs}).}
 

### 2022-02-04 PNAS_manuscript/PNAS_manuscript.tex

@@
 \affil[a]{Department of Earth and Planetary Science, University of California, Berkeley, CA, 94720}
-\affil[b]{Geology, Minerals, Energy, and Geophysics Science Center, U.S. Geological Survey, Moffett Field, CA, 94025}
+\affil[b]{U.S. Geological Survey, Moffett Field, CA, 94025}
 \affil[c]{Department of Earth and Planetary Sciences, Harvard University, Cambridge, MA, 02138}
@@
 The most trusted type of paleointensity estimate is that obtained through experiments in which the primary natural thermal remanent magnetization is progressively replaced by a laboratory magnetization that is imparted in a known field with internal consistency checks (such as in IZZI-style Thellier experiments;  \citealp{Yu2004a}).
-In such Thellier paleointensity experiments, one typical departure from ideal behavior is sagging or double-slopes as visualized in plots that show thermal remanent magnetization (TRM) acquired versus natural remanent magnetization (NRM) lost.
+In such Thellier paleointensity experiments, one typical departure from ideal behavior due to the presence of nonuniformly magnetized grains and alteration during experiments is sagging or double-slopes as visualized in plots that show thermal remanent magnetization (TRM) acquired versus natural remanent magnetization (NRM) lost.
 For such data, distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope.
@@
 Magnetic field maps of (E) and (F) share the same spatial and magnetic scale.
+B_z denotes magnetic field intensity in the vertical direction with positive values corresponding to the out of page direction.
 The unit of the axes are in $\mu$m.}}
@@
 \ref{fig:Petro_QDM}A).
-In addition, the plagioclase crystals of the anorthosite xenoliths are typically free of large oxide inclusions (Fig.
+In addition, the plagioclase crystals of the anorthosite xenoliths are typically free of large oxide inclusions with high aspect ratios (Fig.
 \ref{fig:Petro_QDM}C).
@@
 \ref{fig:coercivity}).
-A compilation of median destructive field (MDF) values in Figure.
-\ref{fig:coercivity} shows that specimens from anorthosites that pass paleointensity selection criteria have distinctly higher average MDF values than other anorthosite and diabase specimens.
+A compilation of median destructive field (MDF) values in Figure \ref{fig:coercivity} shows that specimens from anorthosites that pass paleointensity selection criteria have distinctly higher average MDF values than other anorthosite and diabase specimens.
 Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity around 80 mT (Fig.
@@
 In contrast, other anorthosite and diabase tend to have lower peak coercivities ($\sim$30 mT).
-This result is consistent with an interpretation that a population of magnetic grains with more multidomain-like behavior is responsible for the non-ideal paleointensity behaviors during experiments on such specimens.
+This result is consistent with an interpretation that a population of magnetic grains with more multidomain-like behavior is responsible for the non-ideal paleointensity behaviors during experiments on such specimens \cite{Xu2004a}.
 
@@
 The results show that the anorthosite specimens used for paleointensity experiment have low gamma values ranging from 0.9\textdegree$\;$ to 11.9\textdegree, with a median value of 4.2\textdegree$\;$(Table S1).
-These gamma values are similar to those of Midcontinent Rift volcanics \cite{Sprain2018a}.
+Because these anorthosite specimens were collected with various orientations with respect to the outcrops, the angle between the applied lab field direction during paleointensity experiments with respect to any fabrics within the anorthosites specimens are expected to cover a range of angles.
+In addition, these gamma values are similar to those of Midcontinent Rift volcanics \cite{Sprain2018a}.
 Therefore, the Beaver River anorthosite xenolith bulk samples do not have significant remanence anisotropy.
@@
 Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \cite{Bono2019a} which is based on an interpretation of there being a monotonic decay of the geodynamo through the Proterozoic.
-The highest site-mean virtual dipole moment of the anorthosites are similar to the 99.5 percentile of the selected Phanerozoic data.
+The highest site-mean virtual dipole moment of the anorthosites would rank in the top 1$\%$ of those in the database for the Cenozoic Era (the last 66 million years) when there was unequivocally a crystallizing inner core.
 }}
@@
 \section*{Discussion}
-
-%The outline for this section is::
-%- Discussion of mineral physics constraints (or lack thereof)
-%- Predictions for paleointensity signal associated with inner core nucleation (or lack thereof)
-%- Interpretations of paleomagnetic record (i.e. Bono model of monotonic decrease and then rapid increase;. Cite Biggin et al. 2015) related to inner core nucleation. 
-%- Challenges posed to this model based on our data and other low to strong field transitions.
 
@@
 An approach combining dynamo simulations and theoretical scaling relationships has predicted that progressive decay of the field's dipole moment would be followed by a rapid increase in geomagnetic field intensity soon after the onset of inner core nucleation such that there is a minimum in dipole moment just before inner core nucleation \cite{Davies2021a}.
-Other scenarios are possible, however, such as the prediction that while power increases associated with inner core nucleation would result in increases in Earth's internal magnetic field strength, the dynamo would also become more deeply seated in the core, diminishing the increase in observed magnetic field at Earth's surface \cite{Aubert2009a, Landeau2017a}.
-Observational paleomagnetic records hold the potential for testing different model predictions and identifying transitions in ancient field strength \cite{Biggin2015a, Bono2019a}.
+Other scenarios are possible, however, such as the model-based prediction that while power increases associated with inner core nucleation strengthen Earth's internal magnetic field, that the dynamo becomes more deeply seated in the core diminishes the increase in magnetic field strength at Earth's surface \cite{Aubert2009a, Landeau2017a}.
+Such a shift of the dynamo to a greater depth associated with inner core nucleation led \cite{Aubert2009a} to conclude that the increase in power to the dynamo would be difficult to detect with paleointensity data.
+Ultimately, further observational paleomagnetic records is key as they hold the potential for testing different model predictions and identifying transitions in ancient field strength \cite{Biggin2015a, Bono2019a}.
 
@@
 
-The high paleointensity estimates from the 1.1 billion-year-old Midcontinent Rift rocks challenge the hypothesized monotonic decay of the strength of the geomagnetic field throughout the Proterozoic Era.
-Instead, the well-preserved ca.
+The high paleointensity estimates from the 1.1 billion-year-old Midcontinent Rift rocks challenge the hypothesized monotonic decay of the strength of the geomagnetic field throughout the Proterozoic Era (Fig.
+\ref{fig:PINT_compilation}).
+The well-preserved ca.
 1092 Ma anorthosite xenoliths of the Beaver River diabase record a strong geomagnetic field in the late Mesoproterozoic that exceeds the strength of the modern-day field for which crystallization of the inner core is a power source (Fig.
@@
 1092 Ma by Beaver River anorthosite xenoliths that record virtual dipole moments up to $\sim$140 ZAm$^2$.
-While the Abitibi dikes paleointensity estimates do not go to values as low as the Sept-$\hat{I}$les intrusions, this virtual dipole moment increase in the Mesoproterozoic is the largest yet documented in the Precambrian on a 10s of millions of years timescale (Fig.
+While the ca.
+1140 Abitibi dikes paleointensity estimates do not go to values as low as the ca.
+565 Ma Sept-$\hat{I}$les intrusions, this virtual dipole moment increase in the Mesoproterozoic from the ca.
+1140 Ma data to the ca.
+1100 Ma data is the largest yet documented in the Precambrian on a 10s of millions of years timescale (Fig.
 \ref{fig:PINT_compilation}).
 The tempo and scale of this field intensity transition could match with model-based predictions associated with the onset of inner core nucleation \cite{Davies2021a}.
-This timing would be broadly consistent with that proposed by \citealp{Biggin2015a}.
+This ca.
+1.1 Ga timing would be broadly consistent with the ca.
+1.3 Ga onset proposed by \citealp{Biggin2015a} albeit later given the exclusion of previous overestimated paleointensity values from the ca.
+1.3 Ga Gardar basalts that are superseded by data in \cite{Kodama2019}.
 However, a model prediction of sustained strong field values following inner core nucleation is challenged by data from the ca.

### 2022-02-04 PNAS_manuscript/PNAS_manuscript.tex
Updates from Overleaf

@@
 In contrast to the diabase which fails to provide reliable paleointensity estimates, the anorthosite xenoliths are unusually high-fidelity recorders yielding high-quality, single-slope paleointensity results that are consistent at specimen and site levels.
-An average value of $\sim$89 ZAm$^2$ for the virtual dipole moment from the anorthosites with the highest site-level values up to $\sim$140 ZAm$^2$ are higher than that of the dipole component of Earth's magnetic field today and rival the highest values in the paleointensity database.
+An average value of $\sim$82 ZAm$^2$ for the virtual dipole moment from the anorthosites with the highest site-level values up to $\sim$127 ZAm$^2$ are higher than that of the dipole component of Earth's magnetic field today and rival the highest values in the paleointensity database.
 Such high intensities recorded by the anorthosite xenoliths require the existence of a strongly powered geodynamo at the time.
@@
 1.1 Ga.
-These data are inconsistent with there being a progressive monotonic decay of Earth's dynamo strength through the Proterozoic Eon and could challenge the hypothesis of a young inner core.
+These data are inconsistent with there being a progressive monotonic decay of Earth's dynamo strength through the Proterozoic Eon could challenge the hypothesis of a young inner core.
+The multiple observed paleointensity transitions from weak to strong in the Paleozoic and the Proterozoic present challenges on the identification of the timing of inner core nucleation based on paleointensity records alone.
 \end{abstract}
@@
 The Ediacaran data have been interpreted to indicate that there was a progressively decaying field up to that time that was followed soon after by initial crystallization of the inner core \cite{Bono2019a}.
+However, the ability for the geomagnetic dipole field to reach similar low values in rather recent times when the inner core unambiguously exists (Fig.
+\ref{fig:PINT_compilation}) makes it challenging to unambiguously determine the timing of the initiation of inner core nucleation based on only a few records that span a relatively short time period.
 
-%\begin{figure*}
 \begin{SCfigure*}
@@
 \label{fig:Geologic_map}
-%\end{figure*}
 \end{SCfigure*}
@@
 Magnetic field maps acquired with a quantum diamond microscope (QDM) on samples to which an IRM has been applied show that the relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite in (E) while large oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the overlapping magnetic fields above the oxides in (F).
-Magnetic field maps of (E) and (F) share the same spatial and magnetic scale.
-B_z denotes magnetic field intensity in the vertical direction with positive values corresponding to the out of page direction.
+Magnetic field maps of (E) and (F) share the same spatial scale.
+Note the large contrast in color scale of panels E and F.
+B$_z$ denotes magnetic field intensity in the vertical direction of the field of view with positive values corresponding to the out of page direction.
 The unit of the axes are in $\mu$m.}}
@@
 \ref{fig:Petro_QDM}D,F).
+Magnetic imaging also finds that orientations of magnetic needles within the Duluth plagioclase crystals depict the orientation of remanence magnetizations as is shown by the oblique acquired remanence to the direction of an applied IRM in Figure \ref{fig:Petro_QDM}F.
+Upon applying a second IRM of 300 mT in the orthogonal direction on the Duluth anorthosite, the applied field failed to rotate the remanence directions in plagioclase 1 but resulted in the reversal of remanence in plagioclase 2 which is still oblique to the applied field direction (Fig.
+S1).
+In contrast, the remanence magnetizations imaged in the Beaver River anorthosite xenoliths align parallel to the applied field directions upon applying a set of orthogonal external fields (Fig.
+\ref{fig:Petro_QDM}, S1).
 Titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex magma \cite{Miller1990a}.
@@
 The cooling rate-corrected absolute paleointensity estimates from all anorthosite sites have a mean of 38.28 $\pm$ 11.92 $\mu$T.
-This paleointensity value corresponds to a calculated virtual dipole moment of $\sim$89 ZAm$^2$ ($10^{21} Am^2$) ca.
+This paleointensity value corresponds to a calculated virtual dipole moment of $\sim$82 ZAm$^2$ ($10^{21} Am^2$) ca.
 1092 Ma.
@@
 \ref{fig:PINT_cooling_corrected}; Fig.
-S1).
+S2).
 Overall, the high-quality results from the anorthosite xenoliths of the Beaver River diabase indicate that the anorthosite xenoliths record a high geomagnetic field ca.
@@
 Magnetic property measurement system (MPMS) data show both diabase and anorthosite contain (titano)magnetite as evidenced through the presence of the Verwey transition (Fig.
-S2;  \citealp{Verwey1939a, Feinberg2015a}).
+S3;  \citealp{Verwey1939a, Feinberg2015a}).
 Anorthosite specimens from sites that yield successful paleointensity results have Verwey transition temperatures near 120 K as expected for stochiometric magnetite with minimal Ti \cite{Ozdemir1993a}.
 However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have Verwey transitions that are suppressed toward lower temperatures (Fig.
-S2), indicating that magnetite grains in the specimens either have relatively higher Ti content or have been partially oxidized \cite{Ozdemir1993a}.
+S3), indicating that magnetite grains in the specimens either have relatively higher Ti content or have been partially oxidized \cite{Ozdemir1993a}.
 
@@
 The results show that this set of anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions (\citealp{Selkin2007a}; Fig.
-S4).
+S5).
 
@@
 The significant cooling rate difference leads to a predicted $\sim$36\% overestimate of true ancient field following the model of \citealp{Halgedahl1980a} (Fig.
-S3).
+S4).
 This estimate on cooling rate effect is similar to the value of $\sim$35\% overestimate derived from the model of ref.
@@
 1106 Ma Osler Volcanics, with even stronger values from ca.
-1092 Ma by Beaver River anorthosite xenoliths that record virtual dipole moments up to $\sim$140 ZAm$^2$.
+1092 Ma by Beaver River anorthosite xenoliths that record virtual dipole moments up to $\sim$127 ZAm$^2$.
 While the ca.
@@
 Overall, the high-fidelity paleointensity recorders of the Beaver River anorthosite xenoliths in the well-preserved Midcontinent Rift record strong field strengths 1.1 billion years ago.
-The highest site-level value of the virtual dipole moment would rank in the top 1$\%$ of those in the database for the Cenozoic Era (the last 66 million years) when there was unequivocally a crystallizing inner core.
+The highest site-level value of the virtual dipole moment would rank in the top 2$\%$ of those in the database for the Cenozoic Era (the last 66 million years) when there was unequivocally a crystallizing inner core.
 These high surface field strengths necessitate appreciable power to the late Mesoproterozoic geodynamo.

### 2022-02-04 PNAS_manuscript/Supporting Info.tex

@@
 \documentclass[9pt,twoside,lineno]{pnas-new}
-% Use the lineno option to display guide line numbers if required.
 \usepackage{rotating}
@@
 
-%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
-% \SItext
-
-% \subsection*{Subhead}
-% Type or paste text here. This should be additional explanatory text such as an extended technical description of results, full details of mathematical models, etc.
-
-% \section*{Heading}
-% \subsection*{Subhead}
-% Type or paste text here. You may break this section up into subheads as needed (e.g., one section on ``Materials'' and one on ``Methods'').
-
-% \subsection*{Materials}
-% Add a materials subsection if you need to.
-
-% \subsection*{Methods}
-% Add a methods subsection if you need to.
+\begin{figure*}[h!]
+\noindent\includegraphics[width=0.8\textwidth]{Petro_QDM_SI.png}
+\centering
+\caption{{Petrographic images and magnetic images from the main manuscript with additional anisotropy experiment results.
+In A-F, we show that the Beaver River anorthosite often lack large oxide needles within plagioclase crystals carrying strong magnetic remanence  differences of magnetic carriers which is in contrast to what is seen in Duluth Complex anorthosites.
+E-H show an experiment performed on both anorthosite xenoliths where we apply a first field along the y axis of the field of view and then apply a second field of 300 mT orthogonal to the first field direction.
+The isothermal remanent magnetizations acquired by both anorthosites were imaged with a quantum diamond microscope after the application of each field.
+The magnetic images show that dipole-like remanent magnetizations of Beaver River anorthosite align well with the first applied field direction and rotated to align with the second applied field direction with little anisotropic behavior.
+In contrast, the remanence magnetizations carried by the magnetic needles in plagioclase 1 of the Duluth anorthosite xenolith align well with the first applied field but those in plagioclase 2 tend to acquire an oblique remanence direction with respect to the field direction (F).
+After the application of a 300 mT external field, magnetization of those needles in plagioclase 1 did not reverse due to strong shape anisotropy whereas those needles in plagioclase 2 reversed but still acquired remanence oblique to the field direction.
+This experiments illustrates that the Beaver River anorthosite xenolith has much lower degrees of magnetic anisotropy than the Duluth anorthosite xenolith.
+B$_z$ denotes magnetic field intensity in the vertical direction with positive values corresponding to the out of page direction.
+}}
+\label{fig:Petro_QDM_SI}
+\end{figure*}
+\clearpage
 
 \begin{figure*}[h!]
-\noindent\includegraphics[width=17.8 cm]{manuscript/PINT_BiCEP.pdf}
+\noindent\includegraphics[width=17.8 cm]{PINT_BiCEP.pdf}
 \centering
-\caption{{Violin plots of site-level posterior paleointensity distributions estimated using the bias corrected estimation of paleointensity (BiCEP) method developed by ref.
-\ref{Cych2021a}.
+\caption{{Violin plots of site-level posterior paleointensity distributions estimated using the bias corrected estimation of paleointensity (BiCEP) method developed by \citealp{Cych2021a}.
 Assuming that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Arai1963a, Paterson2011a}, the method uses all paleointensity measurement-level data without applying selection criteria.
@@
 \centering
-\caption{{Graph of predicted paleointensity overestimate due to slow cooling of the intrusive Beaver River diabase and anorthosite xenoliths relative to the cooling rate in laboratory following the model of \citeA{Halgedahl1980a}.
+\caption{{Graph of predicted paleointensity overestimate due to slow cooling of the intrusive Beaver River diabase and anorthosite xenoliths relative to the cooling rate in laboratory following the model of \citealp{Halgedahl1980a}.
 Because the majority of the anorthosites have unblocking temperatures between 500\textdegree C and 580\textdegree C, we estimate that the slow cooling during natural remanence acquisition could have resulted in a 36\% overestimate.
@@
 \centering
-\caption{{Plot of linear thermal remanent magnetization acquisition experiments.
+\caption{{Plot of linear thermal remanent magnetization acquisition experiment results.
 After IZZI-type Thellier paleointensity experiments, full TRMs were imparted on the same specimens in known lab fields of 30, 50, 70, and 90 $\mu$T.
+The red dashed line shows a linear fit through the data points.
 The results show that the anorthosite xenoliths do not acquire saturation remanence or display non-linear remanence acquisition under fields relevant to this study.
@@
 \hline
-AX6  & AX6-2a   & 18.25     & 400       & 585       & 18 & 0.7  & 10.0  & 0.04    & 0.12    & 3.44           & 3.43            & PASS & 15.61         & 2.7      & 31.69          \\
-AX6  & AX6-3a   & 25.39     & 400       & 585       & 18 & 0.76 & 10.0  & 0.04    & 0.1     & 4.28           & 2.88            & PASS & 15.61         & 3.2      & 44.08          \\
-AX6  & AX6-1a   & 49.13     & 425       & 585       & 17 & 0.71 & 10.0  & 0.03    & 0.14    & 4.45           & 0.96            & PASS & 15.61         & 2.0      & 85.3           \\
-AX8  & AX8-3a   & 41.3      & 400       & 580       & 17 & 0.75 & 9.0   & 0.03    & 0.14    & 4.38           & 2.22            & PASS & 17.16         & 11.2     & 70.45          \\
-AX8  & AX8-1a   & 66.18     & 450       & 570       & 13 & 0.68 & 8.0   & 0.06    & 0.18    & 4.54           & 2.37            & PASS & 17.16         & 7.0      & 112.88         \\
-AX10 & AX10-1a  & 45.84     & 425       & 585       & 17 & 0.78 & 10.0  & 0.06    & 0.24    & 5.6            & 2.62            & PASS & 14.41         & 4.7      & 80.64          \\
-AX10 & AX10-3a  & 48.01     & 425       & 585       & 17 & 0.69 & 10.0  & 0.04    & 0.24    & 4.02           & 1.56            & PASS & 14.41         & 5.8      & 84.45          \\
-AX10 & AX10-2a  & 48.23     & 425       & 585       & 17 & 0.67 & 10.0  & 0.05    & 0.19    & 5.3            & 1.93            & PASS & 14.41         & 7.8      & 84.84          \\
-AX11 & AX11-1a  & 73.29     & 425       & 560       & 10 & 0.67 & 6.0   & 0.08    & 0.21    & 5.28           & 2.54            & PASS & 12.27         & 5.9      & 131.75         \\
-AX11 & AX11-2a  & 73.29     & 400       & 560       & 11 & 0.65 & 6.0   & 0.07    & 0.21    & 3.94           & 1.51            & PASS & 12.27         & 9.9      & 131.75         \\
-AX11 & AX11-4a  & 73.31     & 500       & 570       & 11 & 0.66 & 8.0   & 0.09    & 0.2     & 1.53           & 3.27            & PASS & 12.27         & 5.6      & 131.78         \\
-AX11 & AX11-6a  & 73.32     & 200       & 562       & 14 & 0.69 & 7.0   & 0.07    & 0.17    & 4.0            & 1.33            & PASS & 12.27         & 3.0      & 131.8          \\
-AX11 & AX11-9a  & 73.34     & 100       & 562       & 15 & 0.67 & 7.0   & 0.06    & 0.17    & 5.34           & 1.13            & PASS & 12.27         & 11.9     & 131.84         \\
-AX11 & AX11-3a  & 73.41     & 400       & 560       & 11 & 0.66 & 6.0   & 0.08    & 0.21    & 3.95           & 2.11            & PASS & 12.27         & 8.2      & 131.96         \\
-AX11 & AX11-5a  & 73.44     & 400       & 566       & 14 & 0.73 & 8.0   & 0.05    & 0.18    & 2.93           & 0.85            & PASS & 12.27         & 1.9      & 132.02         \\
-AX12 & AX12-14a & 30.34     & 475       & 585       & 16 & 0.75 & 10.0  & 0.06    & 0.17    & 1.74           & 0.55            & PASS & 24.54         & 0.9      & 47.18          \\
-AX12 & AX12-1a  & 33.89     & 0         & 580       & 21 & 0.97 & 9.0   & 0.03    & 0.17    & 5.47           & 4.73            & PASS & 24.54         & 10.0     & 52.7           \\
-AX12 & AX12-6a  & 37.18     & 425       & 564       & 12 & 0.69 & 7.0   & 0.08    & 0.22    & 3.64           & 2.05            & PASS & 24.54         & 5.0      & 57.81          \\
-AX12 & AX12-8a  & 38.34     & 475       & 565       & 12 & 0.75 & 8.0   & 0.06    & 0.25    & 1.46           & 2.29            & PASS & 24.54         & 1.6      & 59.62          \\
-AX12 & AX12-4a  & 41.89     & 500       & 585       & 14 & 0.66 & 10.0  & 0.05    & 0.24    & 3.85           & 3.19            & PASS & 24.54         & 11.4     & 65.14          \\
-AX12 & AX12-2a  & 46.94     & 425       & 570       & 14 & 0.7  & 8.0   & 0.05    & 0.24    & 3.39           & 2.38            & PASS & 24.54         & 7.0      & 72.99          \\
-AX13 & AX13-3a  & 71.68     & 100       & 550       & 12 & 0.71 & 5.0   & 0.08    & 0.22    & 5.92           & 1.88            & PASS & 11.25         & 3.4      & 130.08         \\
-AX13 & AX13-4a  & 75.42     & 200       & 555       & 12 & 0.7  & 6.0   & 0.03    & 0.17    & 7.28           & 0.41            & PASS & 11.25         & 3.7      & 136.87         \\
-AX13 & AX13-6a  & 76.8      & 200       & 555       & 12 & 0.7  & 6.0   & 0.06    & 0.22    & 3.82           & 1.53            & PASS & 11.25         & 4.5      & 139.37         \\
-AX13 & AX13-2a  & 76.96     & 300       & 555       & 11 & 0.69 & 6.0   & 0.05    & 0.2     & 7.87           & 2.73            & PASS & 11.25         & 4.1      & 139.66         \\
-AX13 & AX13-9a  & 77.48     & 0         & 564       & 17 & 0.94 & 7.0   & 0.03    & 0.16    & 5.96           & 1.07            & PASS & 11.25         & 2.6      & 140.61         \\
-AX13 & AX13-8a  & 84.83     & 450       & 570       & 13 & 0.81 & 8.0   & 0.02    & 0.24    & 3.69           & 1.82            & PASS & 11.25         & 5.1      & 153.94   \\
-AX16 & AX16-13a & 41.68     & 400       & 570       & 15 & 0.72 & 8.0   & 0.07    & 0.2     & 3.47           & 1.85            & PASS & 22.11         & 1.1      & 66.88          \\
-AX16 & AX16-16a & 42.2      & 400       & 570       & 15 & 0.77 & 8.0   & 0.06    & 0.15    & 3.9            & 1.81            & PASS & 22.11         & 2.0      & 67.72          \\
-AX16 & AX16-14a & 43.57     & 400       & 570       & 15 & 0.71 & 8.0   & 0.06    & 0.13    & 3.25           & 1.79            & PASS & 22.11         & 6.0      & 69.91          \\
-AX16 & AX16-11a & 44.17     & 450       & 570       & 14 & 0.7  & 8.0   & 0.06    & 0.15    & 2.86           & 2.07            & PASS & 22.11         & 4.6      & 70.88          \\
-AX16 & AX16-5a  & 45.75     & 400       & 560       & 11 & 0.69 & 6.0   & 0.09    & 0.22    & 7.98           & 4.32            & PASS & 22.11         & 4.0      & 73.41          \\
-AX16 & AX16-4a  & 45.96     & 425       & 570       & 14 & 0.75 & 8.0   & 0.05    & 0.21    & 5.01           & 1.18            & PASS & 22.11         & 4.2      & 73.75          \\
-AX16 & AX16-2a  & 45.98     & 400       & 564       & 13 & 0.68 & 7.0   & 0.06    & 0.21    & 4.37           & 1.57            & PASS & 22.11         & 4.9      & 73.78          \\
-AX16 & AX16-1a  & 46.25     & 200       & 562       & 14 & 0.85 & 7.0   & 0.07    & 0.23    & 6.04           & 2.87            & PASS & 22.11         & 4.5      & 74.22          \\
-AX16 & AX16-9a  & 48.9      & 475       & 585       & 16 & 0.66 & 10.0  & 0.04    & 0.16    & 2.92           & 1.4             & PASS & 22.11         & 4.3      & 78.47          \\
-AX16 & AX16-10a & 54.02     & 500       & 585       & 15 & 0.66 & 10.0  & 0.04    & 0.18    & 2.78           & 1.15            & PASS & 22.11         & 5.5      & 86.68          \\
+AX6  & AX6-2a   & 18.25     & 400       & 585       & 18 & 0.7  & 10    & 0.04    & 0.12    & 3.44           & 3.43            & PASS & 21.97         & 2.7      & 29.34         \\
+AX6  & AX6-3a   & 25.39     & 400       & 585       & 18 & 0.76 & 10    & 0.04    & 0.1     & 4.28           & 2.88            & PASS & 21.97         & 3.2      & 40.82         \\
+AX6  & AX6-1a   & 47.36     & 475       & 585       & 15 & 0.6  & 10    & 0.02    & 0.16    & 2.92           & 1.67            & PASS & 21.97         & 2        & 76.13         \\
+AX8  & AX8-3a   & 41.3      & 400       & 580       & 17 & 0.75 & 9     & 0.03    & 0.14    & 4.38           & 2.22            & PASS & 22.98         & 11.2     & 65.53         \\
+AX8  & AX8-2a   & 50.51     & 400       & 580       & 17 & 0.63 & 9     & 0.04    & 0.16    & 3.19           & 1.29            & PASS & 22.98         & 3.7      & 80.15         \\
+AX8  & AX8-1a   & 65.38     & 425       & 566       & 13 & 0.6  & 8     & 0.07    & 0.2     & 5.28           & 2.95            & PASS & 22.98         & 7        & 103.74        \\
+AX10 & AX10-1a  & 45.84     & 425       & 585       & 17 & 0.78 & 10    & 0.06    & 0.24    & 5.6            & 2.62            & PASS & 20.37         & 4.7      & 75.2          \\
+AX10 & AX10-2a  & 46.61     & 450       & 585       & 16 & 0.62 & 10    & 0.04    & 0.2     & 5.32           & 2.24            & PASS & 20.37         & 7.8      & 76.46         \\
+AX10 & AX10-3a  & 48.01     & 425       & 585       & 17 & 0.69 & 10    & 0.04    & 0.24    & 4.02           & 1.56            & PASS & 20.37         & 5.8      & 78.76         \\
+AX11 & AX11-1a  & 73.29     & 425       & 560       & 10 & 0.67 & 6     & 0.08    & 0.21    & 5.28           & 2.54            & PASS & 19.43         & 5.9      & 121.64        \\
+AX11 & AX11-2a  & 73.29     & 400       & 560       & 11 & 0.65 & 6     & 0.07    & 0.21    & 3.94           & 1.51            & PASS & 19.43         & 9.9      & 121.64        \\
+AX11 & AX11-4a  & 73.31     & 500       & 570       & 11 & 0.66 & 8     & 0.09    & 0.2     & 1.53           & 3.27            & PASS & 19.43         & 5.6      & 121.68        \\
+AX11 & AX11-6a  & 73.32     & 200       & 562       & 14 & 0.69 & 7     & 0.07    & 0.17    & 4              & 1.33            & PASS & 19.43         & 3        & 121.69        \\
+AX11 & AX11-9a  & 73.34     & 100       & 562       & 15 & 0.67 & 7     & 0.06    & 0.17    & 5.34           & 1.13            & PASS & 19.43         & 11.9     & 121.73        \\
+AX11 & AX11-3a  & 73.41     & 400       & 560       & 11 & 0.66 & 6     & 0.08    & 0.21    & 3.95           & 2.11            & PASS & 19.43         & 8.2      & 121.84        \\
+AX11 & AX11-5a  & 73.44     & 400       & 566       & 14 & 0.73 & 8     & 0.05    & 0.18    & 2.93           & 0.85            & PASS & 19.43         & 1.9      & 121.89        \\
+AX12 & AX12-14a & 30.34     & 475       & 585       & 16 & 0.75 & 10    & 0.06    & 0.17    & 1.74           & 0.55            & PASS & 35.73         & 0.9      & 40.86         \\
+AX12 & AX12-1a  & 33.89     & 0         & 580       & 21 & 0.97 & 9     & 0.03    & 0.17    & 5.47           & 4.73            & PASS & 35.73         & 10       & 45.64         \\
+AX12 & AX12-6a  & 37.18     & 425       & 564       & 12 & 0.69 & 7     & 0.08    & 0.22    & 3.64           & 2.05            & PASS & 35.73         & 5        & 50.07         \\
+AX12 & AX12-8a  & 38.34     & 475       & 565       & 12 & 0.75 & 8     & 0.06    & 0.25    & 1.46           & 2.29            & PASS & 35.73         & 1.6      & 51.63         \\
+AX12 & AX12-4a  & 41.89     & 500       & 585       & 14 & 0.66 & 10    & 0.05    & 0.24    & 3.85           & 3.19            & PASS & 35.73         & 11.4     & 56.42         \\
+AX12 & AX12-2a  & 46.94     & 425       & 570       & 14 & 0.7  & 8     & 0.05    & 0.24    & 3.39           & 2.38            & PASS & 35.73         & 7        & 63.22         \\
+AX13 & AX13-3a  & 71.68     & 100       & 550       & 12 & 0.71 & 5     & 0.08    & 0.22    & 5.92           & 1.88            & PASS & 19.36         & 3.4      & 119.07        \\
+AX13 & AX13-7a  & 72.6      & 475       & 585       & 15 & 0.61 & 10    & 0.03    & 0.2     & 3.62           & 2.09            & PASS & 19.36         & 11.6     & 120.6         \\
+AX13 & AX13-4a  & 76.13     & 200       & 550       & 11 & 0.65 & 5     & 0.04    & 0.18    & 8.82           & 1.07            & PASS & 19.36         & 3.7      & 126.47        \\
+AX13 & AX13-6a  & 76.8      & 200       & 555       & 12 & 0.7  & 6     & 0.06    & 0.22    & 3.82           & 1.53            & PASS & 19.36         & 4.5      & 127.58        \\
+AX13 & AX13-2a  & 76.86     & 200       & 555       & 12 & 0.71 & 6     & 0.05    & 0.2     & 8.08           & 2.23            & PASS & 19.36         & 4.1      & 127.68        \\
+AX13 & AX13-9a  & 77.48     & 0         & 564       & 17 & 0.94 & 7     & 0.03    & 0.16    & 5.96           & 1.07            & PASS & 19.36         & 2.6      & 128.71        \\
+AX13 & AX13-8a  & 84.83     & 450       & 570       & 13 & 0.81 & 8     & 0.02    & 0.24    & 3.69           & 1.82            & PASS & 19.36         & 5.1      & 140.92        \\
+AX16 & AX16-15a & 31.91     & 475       & 570       & 13 & 0.61 & 8     & 0.07    & 0.17    & 3.07           & 1.41            & PASS & 30.08         & 2.6      & 46.16         \\
+AX16 & AX16-13a & 41.68     & 400       & 570       & 15 & 0.72 & 8     & 0.07    & 0.2     & 3.47           & 1.85            & PASS & 30.08         & 1.1      & 60.29         \\
+AX16 & AX16-16a & 43.3      & 400       & 562       & 13 & 0.65 & 7     & 0.09    & 0.18    & 5.14           & 2.94            & PASS & 30.08         & 2        & 62.63         \\
+AX16 & AX16-14a & 43.64     & 400       & 565       & 14 & 0.63 & 8     & 0.08    & 0.14    & 3.84           & 2.12            & PASS & 30.08         & 6        & 63.13         \\
+AX16 & AX16-11a & 44.17     & 450       & 570       & 14 & 0.7  & 8     & 0.06    & 0.15    & 2.86           & 2.07            & PASS & 30.08         & 4.6      & 63.89         \\
+AX16 & AX16-4a  & 44.82     & 425       & 575       & 15 & 0.76 & 9     & 0.05    & 0.21    & 4.77           & 1.08            & PASS & 30.08         & 4.2      & 64.83         \\
+AX16 & AX16-1a  & 44.9      & 200       & 564       & 15 & 0.87 & 7     & 0.06    & 0.22    & 5.56           & 2.4             & PASS & 30.08         & 4.5      & 64.95         \\
+AX16 & AX16-5a  & 44.91     & 425       & 560       & 10 & 0.65 & 6     & 0.1     & 0.24    & 7.42           & 3.9             & PASS & 30.08         & 4        & 64.96         \\
+AX16 & AX16-2a  & 45.05     & 425       & 564       & 12 & 0.65 & 7     & 0.07    & 0.22    & 4.25           & 1.44            & PASS & 30.08         & 4.9      & 65.17         \\
+AX16 & AX16-9a  & 48.04     & 500       & 585       & 15 & 0.61 & 10    & 0.04    & 0.17    & 2.98           & 1.23            & PASS & 30.08         & 4.3      & 69.49         \\
+AX16 & AX16-10a & 53.27     & 510       & 585       & 14 & 0.62 & 10    & 0.04    & 0.2     & 2.75           & 1.1             & PASS & 30.08         & 5.5      & 77.06          \\
 \hline
@@
 
-% \begin{table}\centering
-% \caption{This is a table}
-
-% \begin{tabular}{lrrr}s
-% Species & CBS & CV & G3 \\
-% \midrule
-% 1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
-% 2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
-% 3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
-% \bottomrule
-% \end{tabular}
-% \end{table}
-
-%%% Add this line AFTER all your figures and tables
 \FloatBarrier
-
-% \movie{Type legend for the movie here.}
-
-% \movie{Type legend for the other movie here. Adding longer text to show what happens, to decide on alignment and/or indentations.}
-
-% \movie{A third movie, just for kicks.}
-
-% \dataset{dataset_one.txt}{Type or paste legend here.}
-
-% \dataset{dataset_two.txt}{Type or paste legend here. Adding longer text to show what happens, to decide on alignment and/or indentations for multi-line or paragraph captions.}
 

### 2022-02-08 PNAS_manuscript/PNAS_manuscript.tex
Updates from Overleaf

@@
 However, the ability for the geomagnetic dipole field to reach similar low values in rather recent times when the inner core unambiguously exists (Fig.
-\ref{fig:PINT_compilation}) makes it challenging to unambiguously determine the timing of the initiation of inner core nucleation based on only a few records that span a relatively short time period.
+\ref{fig:PINT_compilation}) makes it challenging to unambiguously settle the timing of the initiation of inner core nucleation based on such few records that span a relatively short time period.
 
@@
 \centering
-\noindent\includegraphics[width=13 cm]{Petro_QDM.png}
+\noindent\includegraphics[width=13 cm]{Petro_QDM_full.png}
 \caption{\footnotesize{Thin section petrographic images (A,B,C,D) and magnetic maps (E,F) of anorthosite from the Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (left column;  \citealp{Zhang2021b}) and a distinct anorthosite from the Duluth Complex Anorthositic Series (right column).
@@
 Magnetic field maps acquired with a quantum diamond microscope (QDM) on samples to which an IRM has been applied show that the relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite in (E) while large oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the overlapping magnetic fields above the oxides in (F).
-Magnetic field maps of (E) and (F) share the same spatial scale.
-Note the large contrast in color scale of panels E and F.
-B$_z$ denotes magnetic field intensity in the vertical direction of the field of view with positive values corresponding to the out of page direction.
+(E)-(H) show an experiment performed on both samples where we apply a first field along the y axis of the field of view and then apply a second field of 300 mT orthogonal to the first field direction.
+The isothermal remanent magnetizations acquired by both anorthosites were mapped after the application of each field.
+The magnetic images show that dipole-like remanent magnetizations of Beaver River anorthosite align well with the first applied field direction (E) and rotated to align with the second applied field direction with little anisotropic behavior (G).
+In contrast, the remanence magnetizations carried by the magnetic needles in plagioclase 1 of the Duluth anorthosite xenolith align well with the first applied field but those in plagioclase 2 tend to acquire an oblique remanence direction with respect to the field direction (F).
+After the application of a 300 mT external field, magnetization of those needles in plagioclase 1 did not reverse due to strong shape anisotropy whereas those needles in plagioclase 2 reversed but still acquired remanence oblique to the field direction (H).
+This experiment illustrates that the Beaver River anorthosite xenolith has much lower degrees of magnetic anisotropy than the Duluth anorthosite xenolith.
+B$_z$ denotes magnetic field intensity in the vertical direction with positive values corresponding to the out of page direction.
+Note the large contrast in magnetic field scale of (E) and (G) with those in (F) and (G).
 The unit of the axes are in $\mu$m.}}
@@
 Upon applying a second IRM of 300 mT in the orthogonal direction on the Duluth anorthosite, the applied field failed to rotate the remanence directions in plagioclase 1 but resulted in the reversal of remanence in plagioclase 2 which is still oblique to the applied field direction (Fig.
-S1).
+\ref{fig:Petro_QDM}).
 In contrast, the remanence magnetizations imaged in the Beaver River anorthosite xenoliths align parallel to the applied field directions upon applying a set of orthogonal external fields (Fig.
-\ref{fig:Petro_QDM}, S1).
+\ref{fig:Petro_QDM}).
 Titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex magma \cite{Miller1990a}.
@@
 \ref{fig:PINT_cooling_corrected}; Fig.
-S2).
+S1).
 Overall, the high-quality results from the anorthosite xenoliths of the Beaver River diabase indicate that the anorthosite xenoliths record a high geomagnetic field ca.
@@
 Magnetic property measurement system (MPMS) data show both diabase and anorthosite contain (titano)magnetite as evidenced through the presence of the Verwey transition (Fig.
-S3;  \citealp{Verwey1939a, Feinberg2015a}).
+S2;  \citealp{Verwey1939a, Feinberg2015a}).
 Anorthosite specimens from sites that yield successful paleointensity results have Verwey transition temperatures near 120 K as expected for stochiometric magnetite with minimal Ti \cite{Ozdemir1993a}.
 However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have Verwey transitions that are suppressed toward lower temperatures (Fig.
-S3), indicating that magnetite grains in the specimens either have relatively higher Ti content or have been partially oxidized \cite{Ozdemir1993a}.
+S2), indicating that magnetite grains in the specimens either have relatively higher Ti content or have been partially oxidized \cite{Ozdemir1993a}.
 
@@
 The results show that this set of anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions (\citealp{Selkin2007a}; Fig.
-S5).
+S4).
 
@@
 The significant cooling rate difference leads to a predicted $\sim$36\% overestimate of true ancient field following the model of \citealp{Halgedahl1980a} (Fig.
-S4).
+S3).
 This estimate on cooling rate effect is similar to the value of $\sim$35\% overestimate derived from the model of ref.

### 2022-02-08 PNAS_manuscript/Supporting Info.tex
Updates from Overleaf

@@
 
-\begin{figure*}[h!]
-\noindent\includegraphics[width=0.8\textwidth]{Petro_QDM_SI.png}
-\centering
-\caption{{Petrographic images and magnetic images from the main manuscript with additional anisotropy experiment results.
+% \begin{figure*}[h!]
+% \noindent\includegraphics[width=0.8\textwidth]{Petro_QDM_SI.png}
+% \centering
+% \caption{{Petrographic images and magnetic images from the main manuscript with additional anisotropy experiment results.
 In A-F, we show that the Beaver River anorthosite often lack large oxide needles within plagioclase crystals carrying strong magnetic remanence  differences of magnetic carriers which is in contrast to what is seen in Duluth Complex anorthosites.
@@
 }}
-\label{fig:Petro_QDM_SI}
-\end{figure*}
-\clearpage
+% \label{fig:Petro_QDM_SI}
+% \end{figure*}
+% \clearpage
 

### 2022-02-09 PNAS_manuscript/PNAS_manuscript.tex

@@
 All results are corrected for cooling rate with a factor of 0.74.
+As discussed in the text, the Beaver River anorthosite xenoliths likely record geomagnetic field intensities at similar but distinct times during the cooling of the Beaver River diabase.
+To compare these observations with more recent paleointensity variability, we took random draws of 3 values from the PADM2M model \cite{Ziegler2011a} for the VADM during the past 2 million years for a total of 10000 iterations and show the distribution of the calculated paleointensity average value of each draw at a latitude of 22\textdegree N.
 }}
@@
 Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \cite{Bono2019a} which is based on an interpretation of there being a monotonic decay of the geodynamo through the Proterozoic.
-The highest site-mean virtual dipole moment of the anorthosites would rank in the top 1$\%$ of those in the database for the Cenozoic Era (the last 66 million years) when there was unequivocally a crystallizing inner core.
-}}
+The highest site-mean virtual dipole moment of the anorthosites would rank in the top 2$\%$ of those in the database for the Cenozoic Era (the last 66 million years) when there was unequivocally a crystallizing inner core.
+The y axis range is set such that the top 1\% of the Cenozoic paleointensity data is cut off.}}
 \label{fig:PINT_compilation}

### 2022-02-09 PNAS_manuscript/PNAS_manuscript.tex

@@
 1.1 Ga.
-These data are inconsistent with there being a progressive monotonic decay of Earth's dynamo strength through the Proterozoic Eon could challenge the hypothesis of a young inner core.
-The multiple observed paleointensity transitions from weak to strong in the Paleozoic and the Proterozoic present challenges on the identification of the timing of inner core nucleation based on paleointensity records alone.
+These data are inconsistent with there being a progressive monotonic decay of Earth's dynamo strength through the Proterozoic Eon and could challenge the hypothesis of a young inner core.
+The multiple observed paleointensity transitions from weak to strong in the Paleozoic and the Proterozoic present challenges on identifying the onset of inner core nucleation based on paleointensity records alone.
 \end{abstract}
@@
 370 Ma during the Devonian Period \cite{Shcherbakova2017a, Shcherbakova2021a, Hawkins2021a}, potentially indicating unusual periods of core dynamo regime at those times.
-The Ediacaran data have been interpreted to indicate that there was a progressively decaying field up to that time that was followed soon after by initial crystallization of the inner core \cite{Bono2019a}.
-However, the ability for the geomagnetic dipole field to reach similar low values in rather recent times when the inner core unambiguously exists (Fig.
-\ref{fig:PINT_compilation}) makes it challenging to unambiguously settle the timing of the initiation of inner core nucleation based on such few records that span a relatively short time period.
+The Ediacaran data have been interpreted to indicate that there was a progressively decaying field up to that time that was soon followed by initial crystallization of the inner core \cite{Bono2019a}.
+However, the ability for the geomagnetic dipole field to reach similar low values in rather recent times when we known the inner core exists (Fig.
+\ref{fig:PINT_compilation}) makes it challenging to unambiguously settle the timing of the initiation of inner core nucleation based on the current few paleointensity records.
 
@@
 Magnetic imaging confirms that these strongly magnetic needles within the plagioclase crystals often have magnetic moments oriented along their long axes (Fig.
-\ref{fig:Petro_QDM}D,F).
-Magnetic imaging also finds that orientations of magnetic needles within the Duluth plagioclase crystals depict the orientation of remanence magnetizations as is shown by the oblique acquired remanence to the direction of an applied IRM in Figure \ref{fig:Petro_QDM}F.
-Upon applying a second IRM of 300 mT in the orthogonal direction on the Duluth anorthosite, the applied field failed to rotate the remanence directions in plagioclase 1 but resulted in the reversal of remanence in plagioclase 2 which is still oblique to the applied field direction (Fig.
-\ref{fig:Petro_QDM}).
-In contrast, the remanence magnetizations imaged in the Beaver River anorthosite xenoliths align parallel to the applied field directions upon applying a set of orthogonal external fields (Fig.
-\ref{fig:Petro_QDM}).
+\ref{fig:Petro_QDM}D, F, H).
+Magnetic imaging also finds that orientations of magnetic needles within the Duluth plagioclase crystals depict the orientation of remanence magnetizations as is shown by the oblique acquired remanence to the direction of a first applied isothermal remanent magnetization (IRM) in Figure \ref{fig:Petro_QDM}F.
+Upon applying a second IRM with a field of 300 mT in the orthogonal direction, remanent magnetization moments of the Duluth anorthosite failed to rotate in plagioclase 1 but reversed in plagioclase 2 but still oblique to the applied field direction (Fig.
+\ref{fig:Petro_QDM}F, H).
+In contrast, the remanence magnetizations imaged in the Beaver River anorthosite xenoliths align more parallel to the applied field directions upon applying a set of orthogonal external fields (Fig.
+\ref{fig:Petro_QDM}E, G).
 Titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex magma \cite{Miller1990a}.
@@
 All results are corrected for cooling rate with a factor of 0.74.
-As discussed in the text, the Beaver River anorthosite xenoliths likely record geomagnetic field intensities at similar but distinct times during the cooling of the Beaver River diabase.
-To compare these observations with more recent paleointensity variability, we took random draws of 3 values from the PADM2M model \cite{Ziegler2011a} for the VADM during the past 2 million years for a total of 10000 iterations and show the distribution of the calculated paleointensity average value of each draw at a latitude of 22\textdegree N.
+The Beaver River anorthosite xenoliths likely record geomagnetic field intensities at similar but distinct times during cooling of the Beaver River diabase, capturing secular variations.
+To compare these variable observations with more recent paleointensity variability, we took random draws of 3 values from the PADM2M model \cite{Ziegler2011a} for the virtual axial dipole moment values during the past 2 million years for a total of 10000 iterations and show the distribution of average paleointensity values of each draw calculated for a latitude of 22\textdegree N.
 }}
@@
 The results show that the anorthosite specimens used for paleointensity experiment have low gamma values ranging from 0.9\textdegree$\;$ to 11.9\textdegree, with a median value of 4.2\textdegree$\;$(Table S1).
-Because these anorthosite specimens were collected with various orientations with respect to the outcrops, the angle between the applied lab field direction during paleointensity experiments with respect to any fabrics within the anorthosites specimens are expected to cover a range of angles.
-In addition, these gamma values are similar to those of Midcontinent Rift volcanics \cite{Sprain2018a}.
-Therefore, the Beaver River anorthosite xenolith bulk samples do not have significant remanence anisotropy.
+Because these anorthosite specimens were oriented at various directions with respect to the outcrops, the angle between the applied lab field direction during paleointensity experiments with respect to any fabrics within the anorthosites specimens are expected to cover a wide range of angles.
+These gamma values of the anorthosite xenolith bulk specimens are similar to those of Midcontinent Rift volcanics \cite{Sprain2018a}.
+Therefore, the bulk Beaver River anorthosite xenoliths do not have significant remanence anisotropy.
 Paleodirectional data from our anorthosite xenoliths further support that they have minimal remanence anisotropy as their site mean directions closely match those of the Beaver River diabase hosts without deviating due to a fabric \cite{Zhang2021b}.
@@
 1.3 Ga onset proposed by \citealp{Biggin2015a} albeit later given the exclusion of previous overestimated paleointensity values from the ca.
-1.3 Ga Gardar basalts that are superseded by data in \cite{Kodama2019}.
+1.3 Ga Gardar basalts that are superseded by data in \cite{Kodama2019a}.
 However, a model prediction of sustained strong field values following inner core nucleation is challenged by data from the ca.
@@
 
-Given that the multiple records of a weak field can not all be a minimum prior to inner core nucleation, what processes could lead to a weak dipole at Earth's surface even in the presence of a crystallizing inner core?
+Given that the multiple records of a weak field cannot all be the minimum prior to inner core nucleation, what processes could lead to a weak dipole at Earth's surface even in the presence of a crystallizing inner core?
 Numerical models have shown that the dipole moment is sensitive to both the magnitude and spatial pattern of heat flow across the core-mantle boundary when there are strong available power sources to the geodynamo \cite{Olson2007a, Olson2010a}.
@@
 $\beta$, FRAC, GAP-Max and SCAT are all statistics to assess the behavior of NRM/TRM plots.
-See the Standard Paleointensity Definitions ( \citealp{Paterson2014a}; \url{https://earthorg/PmagPy/SPD/home.html}) for more details.
+See the Standard Paleointensity Definitions (\citealp{Paterson2014a}; \url{https://earthorg/PmagPy/SPD/home.html}) for more details.
 Data analysis was conducted using Thellier GUI \cite{Shaar2013a} within the PmagPy software package \cite{Tauxe2016a}.

### 2022-02-09 PNAS_manuscript/cover_letter.tex

\documentclass[11pt, letterpaper]{article}
\usepackage{xcolor}
\usepackage{textcomp,marvosym}
\usepackage{amsmath,amssymb}
\usepackage[left]{lineno}
\usepackage{changepage}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{sidecap}
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\usepackage{titlesec}

\raggedright
\textwidth = 6.5 in
\textheight = 8.25 in
\oddsidemargin = 0.0 in
\evensidemargin = 0.0 in
\topmargin = 0.0 in
\headheight = .5 in
\headsep = 0.5 in
\parskip = 0.1 in
\parindent = 0.2in
\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}

\thispagestyle{myheadings}
\thispagestyle{fancy}
\fancyhf{}
%\lhead{\Large \textbf {Summer School Scholarship Application} \noindent\small\textit{Yiming Zhang, UC Berkeley}}
\lhead{\noindent\small{Yiming Zhang \\ PhD candidate \\ 307 McCone Hall, Berkeley, CA 94720 \\ yimingzhang@berkeley.edu}}
%\rhead{\thepage}

%\author{Yiming Zhang}
%\date{University of California, Berkeley}
\begin{document}

%\maketitle
\begin{flushleft}

\hfill 2022
%{\noindent\textit {Yiming Zhang, UC Berkeley}}

Dear \textit{Proceedings of the National Academy of Sciences} editors,

 This letter accompanies our submission of a manuscript entitled ``High geomagnetic field intensity recorded by anorthosite xenoliths requires a strongly powered late Mesoproterozoic geodynamo".

 Acquiring high-fidelity observational records of ancient magnetic field intensity from the remanent magnetization of rocks is crucial for constraining the long-term evolution of Earth's core. However, robust estimates of ancient field strengths are often difficult to recover due to alteration or non-ideal rock magnetic behavior. In this contribution, we develop high-quality paleointensity data with the best site-level standard deviation of 0.04 $\mu$T from well-preserved anorthosite xenoliths of the Beaver River diabase which were rapidly emplaced during the ca. 1.1 Ga North America Midcontinent Rift. The primary interpretation for the high paleointensity values gains strong support for there being a strongly powered geodynamo during the late Mesoproterozoic. These observations agree with previous results from Midcontinent Rift volcanic rocks and support such a strong field could have lasted at least 14 Myr. Together, paleointensity records from the Midcontinent Rift are inconsistent with there being a progressive monotonic decay of Earth's dynamo strength through the Proterozoic Eon.

  Our updated paleointensity data compilation show multiple observed paleointensity transitions from weak to strong in the Phanerozoic and the Proterozoic. Given that these records cannot all be the minimum prior to the beginning of inner core nucleation, they instead could indicate that processes such as plate tectonics could have modulated the core-mantle heat flow pattern in the Proterozoic. Regardless of when the initiation of inner core nucleation occurred, the observation of large variability in paleointensity in the Precambrian may present challenges in detecting the increase in surface geomagnetic field strength predicted by numerical models to have happened at the onset of inner core crystallization.

We are excited about the results presented in this manuscript and believe that the Earth science community and readers of \textit{Proceedings of the National Academy of Sciences} will find them compelling and a significant advance.

Sincerely,

Yiming Zhang\\

PhD student of Earth and Planetary Science\\

University of California, Berkeley

\clearpage

\textbf{FAIR data availability:}\\
The measurement level paleomagnetic data are in the FAIR data repository MagIC and a link within the manuscript makes these data available for reviewers.

\textbf{Suggested editor reviewers:}\

Lisa Tauxe (ltauxe\@ucsd.edu)\

\textit{expertise in paleointensity and particularly in the behavior of the ancient geomagnetic field}

Joshua Feinberg (feinberg\@umn.edu)\

\textit{expertise in rock magnetism and paleomagnetism}

Andrew Biggin (A.Biggin\@liverpool.ac.uk)\

\textit{expertise in paleomagnetism and particularly in the long term evolution of Earth interior}

Peter Selkin (paselkin\@uw.edu)\

\textit{expertise in paleomagnetism and particularly in the  magnetic properties of magnetic minerals in feldpsar}

Peter Driscoll (pdriscoll\@carnegiescience.edu)\

\textit{expertise in numerical modeling of Earth's thermal and magnetic evolution}
\textit{}
\end{flushleft}
\end{document}

### 2022-02-10 PNAS_manuscript/PNAS_manuscript.tex

@@
 Following IZZI-style paleointensity experiments \cite{Yu2004a}, 40 from a total of 86 anorthosite specimens and 0 out of a total of 69 diabase specimens passed our paleointensity result selection criteria (see Materials and Methods section).
-7 anorthosite sites and no diabase site-level results pass these selection criteria.
+7 anorthosite sites and no diabase sites have specimen results that pass these selection criteria.
 Example NRM/TRM (Arai) plots are shown in Figure \ref{fig:IZZI_examples}.
-Summary specimen absolute paleointensity estimates and site-level mean paleointensity values are plotted in Figure \ref{fig:PINT_cooling_corrected} (and provided in Table S1).
-The paleointensity quality index (Q$_{PI}$; \citealp{Biggin2014a}) for the anorthosite xenoliths are all 5 or 6 (Table S2).
+Summary specimen absolute paleointensity estimates and site-level mean paleointensity values are plotted in Figure \ref{fig:PINT_cooling_corrected} (and provided in Table S1) where each site is a individual anorthosite xenolith.
+The paleointensity quality index (Q$_{PI}$; \citealp{Biggin2014a}) for the anorthosite xenolith sites are all 5 or 6 (Table S2).
 The cooling rate-corrected absolute paleointensity estimates from all anorthosite sites have a mean of 38.28 $\pm$ 11.92 $\mu$T.
-This paleointensity value corresponds to a calculated virtual dipole moment of $\sim$82 ZAm$^2$ ($10^{21} Am^2$) ca.
+This mean of the site virtual dipole moments is $\sim$82 ZAm$^2$ ($10^{21} Am^2$) ca.
 1092 Ma.
@@
 \centering
-\caption{\small{Summary plot of individual specimen absolute paleointensity results (square symbols) from this study and their averages and standard deviations at site (black bars with grey uncertainty boxes) and locality level (orange dashed line and bar).
+\caption{\footnotesize{A) Summary plot of individual specimen absolute paleointensity results (square symbols) from this study and their averages and standard deviations at site level (black bars with grey uncertainty boxes).
 All results are corrected for cooling rate with a factor of 0.74.
-The Beaver River anorthosite xenoliths likely record geomagnetic field intensities at similar but distinct times during cooling of the Beaver River diabase, capturing secular variations.
-To compare these variable observations with more recent paleointensity variability, we took random draws of 3 values from the PADM2M model \cite{Ziegler2011a} for the virtual axial dipole moment values during the past 2 million years for a total of 10000 iterations and show the distribution of average paleointensity values of each draw calculated for a latitude of 22\textdegree N.
+Each `AX' site is an individual anorthosite xenolith within the Beaver River diabase.
+The sites with successful experiments come from three regions which would have cooled at distinct times yielding similar estimates within the region with differences between regions.
+B) Specimen level means calculated for these regions are compared to the distribution of intensities calculated from the paleomagnetic axial dipole moment model for the past 2 million years (PADM2M; \citealp{Ziegler2011a}) at the latitude corresponding to the paleolatitude of study region (22\textdegree N).
+C) The mean of the three regional means is compared to means calculated from 3 random values drawn from the PADM2M model \cite{Ziegler2011a}.
+The distribution represents a total of 10,000 iterations of taking 3 random draws and calculating the mean.
 }}
@@
 \citealp{Zhang2021b} suggest that the Beaver River anorthosite xenoliths were heated to tholeiitic magma temperature ($\sim$1100\textdegree C) and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a thousand years, partially averaging secular variation within single sites.
-Anorthosite xenoliths have paleointensity results that are consistent within small regions, but vary between localities.
-Anorthosite AX12 and AX16 were emplaced $\sim$450 meters apart and have very similar paleointensity estimates (Figs.
+Anorthosite xenoliths have paleointensity results that are consistent within small regions, but vary between regions.
+Anorthosite AX12 and AX16 in the Silver Bay area were emplaced $\sim$450 meters apart and have indistinguishable paleointensity estimates (Figs.
 \ref{fig:Geologic_map} and \ref{fig:PINT_cooling_corrected}).
-$\sim$10 km to the north, anorthosite xenoliths AX11 and AX13 were also emplaced closely (seperated by $\sim$125 meters) and yield similar values to one another but distinct paleointensity estimates from those of AX12 and AX16 (Figs.
+$\sim$10 km to the north, anorthosite xenoliths AX11 and AX13 in the Kennedy Creek area were also emplaced closely (separated by $\sim$125 meters) and yield similar values to one another but distinct paleointensity estimates from those of AX12 and AX16 (Figs.
 \ref{fig:Geologic_map} and \ref{fig:PINT_cooling_corrected}).
-The data therefore capture different intervals of time during the emplacement and cooling of the Beaver River diabase sills ca.
+$\sim$30 km further to the north, anorthosite In the Carleton Peak region, AX6, AX8, and AX10 are anorthosite xenoliths within 10 meters and also yield similar paleointensity estimates albeit with relatively large uncertainties.
+The data therefore could capture at least three different intervals of time during the emplacement and cooling of the Beaver River diabase sills ca.
 1092 Ma.

### 2022-02-11 PNAS_manuscript/PNAS_manuscript.tex

@@
 Higher conductivity values imply faster cooling rates, which in turn imply that the threshold for the freezing of the inner core happened more recently \cite{Davies2015a}.
-While some estimates of these values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates have implied higher thermal conductivity values and an age for the inner core that is less than 1.3 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b}, with some suggesting even younger ages ($<$700 Ma;  \citealp{Labrosse2015a, Ohta2016a}).
+While some estimates of these values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates have implied higher thermal conductivity values and an age for the inner core that is less than 1.5 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b, Frost2022a}, with some suggesting even younger ages ($<$700 Ma;  \citealp{Labrosse2015a, Ohta2016a}).
 Given that estimates for the core's thermal conductivity continue to be debated, it is crucial to use observational records as an independent constraint on the thermal evolution of Earth's core and mantle.
@@
 The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal intrusions of the Beaver River diabase \cite{Zhang2021b}.
-They cooled and acquired their magnetizations in conjunction with the host diabase at a paleolatitude of 22\textdegree $\pm$2 \textdegree (calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a, Zhang2021b}).
+They cooled and acquired their magnetizations in conjunction with the host diabase at a paleolatitude of 22\textdegree $\pm$2\textdegree (calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a, Zhang2021b}).
 Paleointensity experiments on the anorthosite xenoliths have a high success rate, yielding consistent specimen- and site-level paleointensity results.
-Rock magnetic data reveal that the anorthosite xenoliths have low anisotropy of thermal remanent magnetization (TRM) and can acquire TRM linearly within relevant field strengths.
+Rock magnetic data reveal that the anorthosite xenoliths have low anisotropy of thermal remanent magnetization (TRM) and can acquire TRM linearly within the range of relevant field strengths.
 Magnetic imaging shows the anorthosite specimens have dominant magnetic carriers within and interstitial to plagioclase crystals, and they do not display strong preferred orientations.
@@
 1092 Ma Beaver Bay Complex in northern Minnesota punctuates one such period of magmatism during the main stage of Midcontinent Rift activity.
-The magma that formed the 1091.7 $\pm$ 0.2 Ma Beaver River diabase dikes and sills of the Beaver Bay Complex transported numerous anorthosite xenoliths that have short-axis diameters up to 180 meters via wide conduits \cite{Boerboom2004a, Boerboom2006b}.
+The magma that formed the 1091.7$\pm$0.2 Ma Beaver River diabase dikes and sills of the Beaver Bay Complex transported numerous anorthosite xenoliths that have short-axis diameters up to 180 meters via wide conduits \cite{Boerboom2004a, Boerboom2006b}.
 These anorthosite xenoliths are plagioclase cumulates that formed comagmatically with the host diabase in the lower crust---an interpretation confirmed by U-Pb zircon geochronology \cite{Zhang2021b}.
@@
 Magnetic imaging using a quantum diamond microscope (QDM) shows that the dominant magnetic remanence carriers often exist within, and sometimes interstitial to, plagioclase crystals (Fig.
-\ref{fig:Petro_QDM}E).
+\ref{fig:Petro_QDM}E, G).
 
-These features of the Beaver River anorthosite xenoliths are distinct from many other anorthosites such as the anorthosite xenoliths within the Duluth Complex Anorthositic Series rocks---older intrusions within the Midcontinent Rift.
+The features of the Beaver River anorthosite xenoliths are distinct from many other anorthosites such as the anorthosite xenoliths within the Duluth Complex Anorthositic Series rocks---older intrusions within the Midcontinent Rift.
 To illustrate these difference, we present petrographic and magnetic imaging data from a sample of Duluth Complex anorthosite that was not targeted for paleointensity experiments.
-The plagioclase of the Duluth Complex anorthosites typically develop interlocking textures that display strong igneous foliation (Fig.
+The plagioclase of the Duluth Complex anorthosites typically develop interlocking textures that display pronounced igneous foliation (Fig.
 \ref{fig:Petro_QDM}B).
@@
 Magnetic imaging also finds that orientations of magnetic needles within the Duluth plagioclase crystals depict the orientation of remanence magnetizations as is shown by the oblique acquired remanence to the direction of a first applied isothermal remanent magnetization (IRM) in Figure \ref{fig:Petro_QDM}F.
-Upon applying a second IRM with a field of 300 mT in the orthogonal direction, remanent magnetization moments of the Duluth anorthosite failed to rotate in plagioclase 1 but reversed in plagioclase 2 but still oblique to the applied field direction (Fig.
+Upon applying a second IRM with a field of 300 mT in the orthogonal direction, the remanent magnetic moments failed to rotate in plagioclase 1, reversed in plagioclase 2 but still along the easy axis depicted by the needle orientations that are oblique to the applied field direction (Fig.
 \ref{fig:Petro_QDM}F, H).
-In contrast, the remanence magnetizations imaged in the Beaver River anorthosite xenoliths align more parallel to the applied field directions upon applying a set of orthogonal external fields (Fig.
+In contrast, the remanent magnetizations imaged in the Beaver River anorthosite xenoliths align much more parallel to the applied field directions upon applying a set of orthogonal external fields (Fig.
 \ref{fig:Petro_QDM}E, G).
+To our knowledge, this is the first time that the magnetic anisotropy associated with plagioclase petrofabrics is captured by magnetic imaging at such scale.
+This set of orthogonal IRM experiment clearly shows that the elongated (titano)magnetite exsolutions within plagioclase feldspar are responsible for the magnetic anisotropy of plagioclase which have been observed before at bulk sample scale \cite{Selkin2000a, Feinberg2006a}.
+
 Titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex magma \cite{Miller1990a}.
-
 The lack of symplecites and the lack of large plagioclase-hosted titanomagnetite needles within the Beaver River anorthosite xenoliths targeted in this study distinguish them from plagioclase cumulates of the Duluth Complex and other layered mafic complexes where anisotropic magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
-The relative lack of fabric and associated preferred orientation makes the Beaver River anorthosite xenoliths a particularly compelling target for paleointensity experiments.
+The relative lack of petrofabrics and associated preferred orientation makes the Beaver River anorthosite xenoliths a particularly compelling target for paleointensity experiments.
 
@@
 Example NRM/TRM (Arai) plots are shown in Figure \ref{fig:IZZI_examples}.
-Summary specimen absolute paleointensity estimates and site-level mean paleointensity values are plotted in Figure \ref{fig:PINT_cooling_corrected} (and provided in Table S1) where each site is a individual anorthosite xenolith.
+Summary specimen absolute paleointensity estimates and site-level mean paleointensity values are plotted in Figure \ref{fig:PINT_cooling_corrected} (and provided in Table S1) where each site is an individual anorthosite xenolith.
 The paleointensity quality index (Q$_{PI}$; \citealp{Biggin2014a}) for the anorthosite xenolith sites are all 5 or 6 (Table S2).
-The cooling rate-corrected absolute paleointensity estimates from all anorthosite sites have a mean of 38.28 $\pm$ 11.92 $\mu$T.
-This mean of the site virtual dipole moments is $\sim$82 ZAm$^2$ ($10^{21} Am^2$) ca.
+The cooling rate-corrected absolute paleointensity estimates from the anorthosite xenoliths have a mean of 38.28$\pm$11.92 $\mu$T.
+The site-mean virtual dipole moment is $\sim$82 ZAm$^2$ ($10^{21} Am^2$) ca.
 1092 Ma.
@@
 \centering
-\caption{\footnotesize{A) Summary plot of individual specimen absolute paleointensity results (square symbols) from this study and their averages and standard deviations at site level (black bars with grey uncertainty boxes).
+\caption{\footnotesize{A) Summary plot of individual specimen absolute paleointensity results (square symbols) and their averages and standard deviations at site level (black bars with grey uncertainty boxes).
 All results are corrected for cooling rate with a factor of 0.74.
 Each `AX' site is an individual anorthosite xenolith within the Beaver River diabase.
-The sites with successful experiments come from three regions which would have cooled at distinct times yielding similar estimates within the region with differences between regions.
-B) Specimen level means calculated for these regions are compared to the distribution of intensities calculated from the paleomagnetic axial dipole moment model for the past 2 million years (PADM2M; \citealp{Ziegler2011a}) at the latitude corresponding to the paleolatitude of study region (22\textdegree N).
-C) The mean of the three regional means is compared to means calculated from 3 random values drawn from the PADM2M model \cite{Ziegler2011a}.
-The distribution represents a total of 10,000 iterations of taking 3 random draws and calculating the mean.
-}}
+The sites with successful experiments come from 3 regions which would have cooled at distinct times yielding similar estimates within the each region with differences between regions.
+B) Specimen level means calculated for these regions are compared to the distribution of intensities calculated from a paleomagnetic axial dipole moment model for the past 2 million years (PADM2M; \citealp{Ziegler2011a}) at the latitude corresponding to the paleolatitude of study region (22\textdegree N).
+C) The mean of the 3 regional means is compared to means calculated from 3 random values drawn from the PADM2M model \cite{Ziegler2011a}.
+The distribution represents a total of 10,000 iterations of taking 3 random draws and calculating the mean.}}
 \label{fig:PINT_cooling_corrected}
@@
 We applied full TRMs to a set of anorthosite specimens with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axes.
-The results show that this set of anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions (\citealp{Selkin2007a}; Fig.
+The results show that these anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions (\citealp{Selkin2007a}; Fig.
 S4).
@@
 \ref{fig:Geologic_map} and \ref{fig:PINT_cooling_corrected}).
-$\sim$30 km further to the north, anorthosite In the Carleton Peak region, AX6, AX8, and AX10 are anorthosite xenoliths within 10 meters and also yield similar paleointensity estimates albeit with relatively large uncertainties.
-The data therefore could capture at least three different intervals of time during the emplacement and cooling of the Beaver River diabase sills ca.
+$\sim$30 km further to the north in the Carlton Peak region, AX6, AX8, and AX10 are anorthosite xenoliths within 10 meters and also yield similar paleointensity estimates albeit with relatively large uncertainties (Fig.
+\ref{fig:PINT_cooling_corrected}).
+Therefore, one interpretation could consider the anorthosites of the three distinct regions to have captured at least three different records of geomagnetic field intensities during the emplacement and cooling of the Beaver River diabase sills ca.
+1092 Ma.
+To contextualize our high paleointensity estimates, we plot the regional mean values based on specimen-level paleointensity results together with with the calculated paleointensity values at 22\textdegree N latitude based on the VADM values of model PADM2M \cite{Ziegler2011a} in Figure \ref{fig:PINT_cooling_corrected}B.
+The locality-mean values of these three regions are plotted in Figure \ref{fig:PINT_cooling_corrected}C where we also present the distribution of means calculated from 3 random values from the PADM2M model for a total of 10000 iterations.
+In addition, we perform the same comparison with the existing paleointensity data for the Cenozoic Era (the last 66 million years) from the PINT database (filtered by Q$_{PI}\geq$3) given that the observational dataset could record more geomagnetic field excursions or regional high flux patches (Fig.
+S5).
+The anorthosite xenoliths 3 region mean intensity is higher than all values from the resample results of the PADM2M model, and rival the highest values of the resample averages from the Cenozoic data (Fig.
+S5).
+These data supports that the anorthosite xenoliths likely record an exceptionally strong geomagnetic field ca.
 1092 Ma.
@@
 Overall, the high-fidelity paleointensity recorders of the Beaver River anorthosite xenoliths in the well-preserved Midcontinent Rift record strong field strengths 1.1 billion years ago.
-The highest site-level value of the virtual dipole moment would rank in the top 2$\%$ of those in the database for the Cenozoic Era (the last 66 million years) when there was unequivocally a crystallizing inner core.
+The highest site-level value of the virtual dipole moment would rank in the top 2$\%$ of those in the database for the Cenozoic Era when there was unequivocally a crystallizing inner core.
 These high surface field strengths necessitate appreciable power to the late Mesoproterozoic geodynamo.

### 2022-02-11 PNAS_manuscript/Supporting Info.tex
Updates from Overleaf

@@
 
-% \clearpage
+\clearpage
+
+\begin{figure*}[h!]
+\noindent\includegraphics[width=0.9\textwidth]{Cenozoic_resample.pdf}
+\centering
+\caption{\footnotesize{A) Summary plot of individual specimen absolute paleointensity results (square symbols) and their averages and standard deviations at site level (black bars with grey uncertainty boxes).
+All results are corrected for cooling rate with a factor of 0.74.
+Each `AX' site is an individual anorthosite xenolith within the Beaver River diabase.
+The sites with successful experiments come from 3 regions which would have cooled at distinct times yielding similar estimates within the each region with differences between regions.
+B) Specimen level means calculated for these regions are compared to the distribution of intensities calculated from a paleomagnetic axial dipole moment model for the past 2 million years (PADM2M; \citealp{Ziegler2011a}) at the latitude corresponding to the paleolatitude of study region (22\textdegree N).
+C) The mean of the 3 regional means is compared to means calculated from 3 random values drawn from the PADM2M model \cite{Ziegler2011a}.
+The distribution represents a total of 10,000 iterations of taking 3 random draws and calculating the mean.}}
+\label{fig:Cenozoic_PINT}
+\end{figure*}
+
+\clearpage
 
 \begin{sidewaystable}
-\caption{\footnotesize{Summary paleointensity results for specimens that passed our selection.
+\caption{\footnotesize{Specimen paleointensity results that passed our selection.
 Paleointensity results for specimens that passed quality criteria.
@@
 
+\clearpage
+
+\begin{table}[]
+\centering
+\caption{\footnotesize{Summary paleointensity result statistics at site-level, region-level, and overall mean.
+The statistics names are the same as in Table \ref{tab:PINT_result}}.
+n represents the total number of specimen- or site-level results used for calculating mean values.
+Site-level means for AX6, AX8, AX10, AX11, AX12, AX13, AX16 are calculated based on specimen results.
+Regional means for Carlton Peak, Kennedy Creek, and Silver Bay are calculated by grouping specimen results from AX6 and AX8 and AX10, AX11 and AX13, AX12 and AX16, respectively.
+Overall site mean values are calculated using all site level results in this table.
+Overall region mean values are calculated using all regional mean results in this table.
+}
+\begin{tabular}{cccccccccc}
+\hline
+Group                & B$_{anc}$ & n  & FRAC & GAP-MAX & MAD ($^\circ$) & DANG ($^\circ$) & Paleolatitude & $\gamma$ & VDM (ZAm$^2$) \\
+\hline
+AX6                 & 30.33     & 3  & 0.69 & 0.13    & 3.55           & 2.66            & 21.97         & 2.63     & 48.76         \\
+AX8                 & 52.4      & 3  & 0.66 & 0.17    & 4.28           & 2.15            & 22.98         & 7.3      & 83.14         \\
+AX10                & 46.82     & 3  & 0.7  & 0.23    & 4.98           & 2.14            & 20.37         & 6.1      & 76.81         \\
+AX12                & 38.1      & 6  & 0.75 & 0.22    & 3.26           & 2.53            & 35.73         & 5.98     & 51.31         \\
+AX16                & 44.15     & 11 & 0.68 & 0.19    & 4.19           & 1.96            & 30.08         & 3.97     & 63.87         \\
+AX11                & 73.34     & 7  & 0.68 & 0.19    & 3.85           & 1.82            & 19.43         & 6.63     & 121.73        \\
+AX13                & 76.63     & 7  & 0.73 & 0.2     & 5.7            & 1.67            & 19.36         & 5        & 127.29        \\
+Carlton Peak        & 43.18     & 9  & 0.68 & 0.17    & 4.27           & 2.32            & 21.77         & 5.34     & 69.57         \\
+Kennedy Creek       & 74.98     & 14 & 0.7  & 0.2     & 4.78           & 1.74            & 19.4          & 5.81     & 124.51        \\
+Silver Bay          & 42.02     & 17 & 0.71 & 0.2     & 3.86           & 2.16            & 32.07         & 4.68     & 59.44         \\
+overall site mean   & 51.68     & 40 & 0.7  & 0.19    & 4.26           & 2.13            & 24.27         & 5.37     & 81.84         \\
+overall region mean & 53.39     & 40 & 0.7  & 0.19    & 4.3            & 2.07            & 24.41         & 5.28     & 84.51        \\
+\hline
+\end{tabular}
+\end{table}
+
+\clearpage
+
 \begin{table}

### 2022-02-12 PNAS_manuscript/Supporting Info.tex
Updates from Overleaf

--- 2022-02-08 PNAS_manuscript/Supporting Info.tex
@@
 Because the majority of the anorthosites have unblocking temperatures between 500\textdegree C and 580\textdegree C, we estimate that the slow cooling during natural remanence acquisition could have resulted in a 36\% overestimate.
-Thus, a correction factor of 0.74 is applied at specimen level for the paleointensity summary plot.
-}}
+Thus, a correction factor of 0.74 is applied at specimen level for the paleointensity summary plot.}}
 \label{fig:PINT_cooling_corrected}
@@
 
-% \clearpage
+\clearpage
+
+\begin{figure*}[h!]
+\noindent\includegraphics[width=0.9\textwidth]{Cenozoic_resample_SI.pdf}
+\centering
+\caption{\footnotesize{A) Summary plot of individual specimen absolute paleointensity results (square symbols) and their averages and standard deviations at site level (black bars with grey uncertainty boxes) from this study.
+All results are corrected for cooling rate with a factor of 0.74.
+Each `AX' site is an individual anorthosite xenolith within the Beaver River diabase.
+The sites with successful experiments come from 3 regions which would have cooled at distinct times yielding similar estimates within the each region with differences between regions.
+B) Specimen level means calculated for these regions are compared to the distribution of intensities calculated from the existing paleointensity data in the PINT database (PINT v8.0.0; \url{http://www.pintdb.org/}; \citealp{Bono2021a}) at the latitude corresponding to the paleolatitude of study region (22\textdegree N).
+C) The mean of the 3 regional means is compared to means calculated from 3 random values drawn from the Cenozoic data.
+The paleointensity records used for resampling are filtered for Q$_{PI} \geq$3 \cite{Biggin2014a}.
+The distribution represents a total of 10,000 iterations of taking 3 random draws and calculating the mean.}}
+\label{fig:Cenozoic_PINT}
+\end{figure*}
+
+\clearpage
 
 \begin{sidewaystable}
-\caption{\footnotesize{Summary paleointensity results for specimens that passed our selection.
+\caption{\footnotesize{Specimen paleointensity results that passed our selection.
 Paleointensity results for specimens that passed quality criteria.
@@
 
+\clearpage
+
 \begin{table}
@@
 
+\clearpage
+
+\begin{table}[]
+\centering
+\caption{\footnotesize{Summary paleointensity result statistics at site-level, region-level, and overall arithmetic means.
+The statistics names are the same as in Table \ref{tab:PINT_result}}.
+n represents the total number of specimen- or site-level results used for calculating mean values.
+inc$_{tc}$ is the tilt corrected mean inclination.
+Site-level means for AX6, AX8, AX10, AX11, AX12, AX13, AX16 are calculated based on specimen results.
+Regional means for Carlton Peak, Kennedy Creek, and Silver Bay are calculated by grouping specimen results from AX6 and AX8 and AX10, AX11 and AX13, AX12 and AX16, respectively.
+Overall site mean values are calculated using all site level results in this table.
+Overall region mean values are calculated using all regional mean results in this table.
+}
+\begin{tabular}{cccccccccc}
+Group               & B$_{anc}$ & n  & FRAC & GAP-MAX & MAD ($^\circ$) & DANG ($^\circ$) & inc$_{tc}$ & $\gamma$ & VDM (ZAm$^2$) \\
+AX6                 & 30.33     & 3  & 0.69 & 0.13    & 3.55           & 2.66            & 38.9       & 2.63     & 48.76         \\
+AX8                 & 52.4      & 3  & 0.66 & 0.17    & 4.28           & 2.15            & 40.3       & 7.3      & 83.14         \\
+AX10                & 46.82     & 3  & 0.7  & 0.23    & 4.98           & 2.14            & 36.6       & 6.1      & 76.81         \\
+AX12                & 38.1      & 6  & 0.75 & 0.22    & 3.26           & 2.53            & 55.2       & 5.98     & 51.31         \\
+AX16                & 44.15     & 11 & 0.68 & 0.19    & 4.19           & 1.96            & 49.2       & 3.97     & 63.87         \\
+AX11                & 73.34     & 7  & 0.68 & 0.19    & 3.85           & 1.82            & 35.2       & 6.63     & 121.73        \\
+AX13                & 76.63     & 7  & 0.73 & 0.2     & 5.7            & 1.67            & 35.1       & 5        & 127.29        \\
+Carlton Peak        & 43.18     & 9  & 0.68 & 0.17    & 4.27           & 2.32            & 38.6       & 5.34     & 69.57         \\
+Kennedy Creek       & 74.98     & 14 & 0.7  & 0.2     & 4.78           & 1.74            & 35.15      & 5.81     & 124.51        \\
+Silver Bay          & 42.02     & 17 & 0.71 & 0.2     & 3.86           & 2.16            & 51.32      & 4.68     & 59.43         \\
+overall site mean   & 51.68     &    &      &         &                &                 &            &          & 81.84         \\
+overall region mean & 53.39     &    &      &         &                &                 &            &          & 84.5         
+\end{tabular}
+\end{table}
+
+\clearpage
+
 \FloatBarrier

### 2022-02-12 PNAS_manuscript/PNAS_manuscript.tex

@@
 In contrast to the diabase which fails to provide reliable paleointensity estimates, the anorthosite xenoliths are unusually high-fidelity recorders yielding high-quality, single-slope paleointensity results that are consistent at specimen and site levels.
-An average value of $\sim$82 ZAm$^2$ for the virtual dipole moment from the anorthosites with the highest site-level values up to $\sim$127 ZAm$^2$ are higher than that of the dipole component of Earth's magnetic field today and rival the highest values in the paleointensity database.
+An average value of $\sim$82 ZAm$^2$ for the virtual dipole moment from the anorthosite xenoliths, with the highest site-level values up to $\sim$127 ZAm$^2$, are higher than that of the dipole component of Earth's magnetic field today and rival the highest values in the paleointensity database.
 Such high intensities recorded by the anorthosite xenoliths require the existence of a strongly powered geodynamo at the time.
-Together with previous paleointensity data from other Midcontinent Rift rocks, our new results indicate that a dynamo with strong power sources persisted for more than 14 million years ca.
+Together with previous paleointensity data from other Midcontinent Rift rocks, these results indicate that a dynamo with strong power sources persisted for more than 14 million years ca.
 1.1 Ga.
 These data are inconsistent with there being a progressive monotonic decay of Earth's dynamo strength through the Proterozoic Eon and could challenge the hypothesis of a young inner core.
-The multiple observed paleointensity transitions from weak to strong in the Paleozoic and the Proterozoic present challenges on identifying the onset of inner core nucleation based on paleointensity records alone.
+The multiple observed paleointensity transitions from weak to strong in the Paleozoic and the Proterozoic present challenges in identifying the onset of inner core nucleation based on paleointensity records alone.
 \end{abstract}
@@
 565 Ma during the Ediacaran Period \cite{Bono2019a, Shcherbakova2019a,Thallner2021b} and ca.
-370 Ma during the Devonian Period \cite{Shcherbakova2017a, Shcherbakova2021a, Hawkins2021a}, potentially indicating unusual periods of core dynamo regime at those times.
+370 Ma during the Devonian Period \cite{Shcherbakova2017a, Shcherbakova2021a, Hawkins2021a}, potentially indicating unusual periods of core dynamo activity at those times.
 The Ediacaran data have been interpreted to indicate that there was a progressively decaying field up to that time that was soon followed by initial crystallization of the inner core \cite{Bono2019a}.
-However, the ability for the geomagnetic dipole field to reach similar low values in rather recent times when we known the inner core exists (Fig.
+Sparse data in the Proterozoic Era (2500 to 539 Ma) and the reality of high variability in Phanerozoic Era records (539 to 0 Ma) necessitate additional data to evaluate interpreted trends.
+
+% However, the ability for the geomagnetic dipole field to reach similar low values in rather recent times when we known the inner core exists (Fig.
 \ref{fig:PINT_compilation}) makes it challenging to unambiguously settle the timing of the initiation of inner core nucleation based on the current few paleointensity records.
@@
 These magnetizations need to be unmodified by subsequent heating or chemical alteration in order to maintain a record of the ancient geomagnetic field from the time of cooling.
-Intracontinental magmatic events are therefore an important target for determination of ancient paleointensity as they can be well-preserved within continental interiors.
+Intracontinental magmatic events are an important target for determination of ancient paleointensity as they can be well-preserved within continental interiors.
 This interior position results in them typically being distant from tectonic events along continental margins that can drive alteration through heat and fluids.
@@
 The Midcontinent Rift (Fig.
-\ref{fig:Geologic_map}) is an exception as it is a large igneous province where magmatism lasted $\sim$25 Myr from \texit{ca.} 1109 Ma to 1084 Ma \cite{Swanson-Hysell2019a}.
+\ref{fig:Geologic_map}) is an exception as it is a large igneous province where magmatism lasted $\sim$25 Myr from \texit{ca.} 1109 Ma to 1084 Ma during which there were pulsed intervals of more rapid activity \cite{Swanson-Hysell2021a}.
 Additionally, extension ceased in the Midcontinent Rift prior to lithospheric separation, preserving volcanic, intrusive, and sedimentary rocks of the rift within the continental interior.
@@
 The most trusted type of paleointensity estimate is that obtained through experiments in which the primary natural thermal remanent magnetization is progressively replaced by a laboratory magnetization that is imparted in a known field with internal consistency checks (such as in IZZI-style Thellier experiments;  \citealp{Yu2004a}).
-In such Thellier paleointensity experiments, one typical departure from ideal behavior due to the presence of nonuniformly magnetized grains and alteration during experiments is sagging or double-slopes as visualized in plots that show thermal remanent magnetization (TRM) acquired versus natural remanent magnetization (NRM) lost.
+In such Thellier paleointensity experiments, one typical departure from ideal behavior due to the presence of nonuniformly magnetized grains (with either multidomain \cite{Dunlop2001a} or vortex states \cite{Tauxe2020a}) is sagging or double-slopes as visualized in Arai plots that show thermal remanent magnetization (TRM) acquired versus natural remanent magnetization (NRM) lost.
 For such data, distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope.
@@
 They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud magnetic inclusions from alteration with alteration of the plagioclase crystals themselves not resulting in the formation of secondary iron oxides in contrast with mafic minerals such as olivine.
-The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal intrusions of the Beaver River diabase \cite{Zhang2021b}.
+The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal (shallowly emplaced) intrusions of the Beaver River diabase \cite{Zhang2021b}.
 They cooled and acquired their magnetizations in conjunction with the host diabase at a paleolatitude of 22\textdegree $\pm$2\textdegree (calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a, Zhang2021b}).
@@
 Magnetic imaging shows the anorthosite specimens have dominant magnetic carriers within and interstitial to plagioclase crystals, and they do not display strong preferred orientations.
-These single-slope, high quality paleointensity data, together with other high-quality paleointensity records during the protracted Midcontinent Rift magmatic activity, require that there was a strong late Mesoproterozoic geodynamo.
+These single-slope, high-quality paleointensity data, together with other high-quality paleointensity records during the protracted Midcontinent Rift magmatic activity, require that there was a strong late Mesoproterozoic geodynamo.
 
@@
 
-While magmatic activity within the Midcontinent Rift was protracted, there were intervals of particularly rapid volcanism and voluminous emplacement of intrusions.
+While magmatic activity within the Midcontinent Rift was protracted, there were intervals of particularly rapid volcanism and voluminous emplacement of intrusions \cite{Swanson-Hysell2021a}.
 The ca.
@@
 In contrast, plagioclase crystals in the Duluth Complex anorthosite have euhedral, interlocking crystals with an igneous foliation (B) and the plagioclase crystals contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the [001] axis of the plagioclase.
-Magnetic field maps acquired with a quantum diamond microscope (QDM) on samples to which an IRM has been applied show that the relatively weak magnetic sources within and interstitial to plagioclase crystals exist in the Beaver River anorthosite in (E) while large oxide needles within Duluth Complex plagioclase are strongly magnetic as evidenced through the overlapping magnetic fields above the oxides in (F).
-(E)-(H) show an experiment performed on both samples where we apply a first field along the y axis of the field of view and then apply a second field of 300 mT orthogonal to the first field direction.
-The isothermal remanent magnetizations acquired by both anorthosites were mapped after the application of each field.
-The magnetic images show that dipole-like remanent magnetizations of Beaver River anorthosite align well with the first applied field direction (E) and rotated to align with the second applied field direction with little anisotropic behavior (G).
-In contrast, the remanence magnetizations carried by the magnetic needles in plagioclase 1 of the Duluth anorthosite xenolith align well with the first applied field but those in plagioclase 2 tend to acquire an oblique remanence direction with respect to the field direction (F).
-After the application of a 300 mT external field, magnetization of those needles in plagioclase 1 did not reverse due to strong shape anisotropy whereas those needles in plagioclase 2 reversed but still acquired remanence oblique to the field direction (H).
-This experiment illustrates that the Beaver River anorthosite xenolith has much lower degrees of magnetic anisotropy than the Duluth anorthosite xenolith.
-B$_z$ denotes magnetic field intensity in the vertical direction with positive values corresponding to the out of page direction.
-Note the large contrast in magnetic field scale of (E) and (G) with those in (F) and (G).
-The unit of the axes are in $\mu$m.}}
+Maps of the vertical component of magnetic field (B$_z$) developed with a quantum diamond microscope (QDM) show relatively weak magnetic sources within and interstitial to plagioclase crystals in the Beaver River anorthosite (in E) relative to the strongly magnetic large oxide needles within Duluth Complex plagioclase (in F).
+The B$_z$ color scale is an order of magnitude greater in the Duluth Complex anorthosite maps (F, H) than the maps for the Beaver River anorthosite (E, G).
+(E to G) and (F to H) show experiments performed on both samples where we apply a first field along the y axis of the field of view and then apply a second field of 300 mT orthogonal to the first field direction.
+The remanent magnetizations acquired by both anorthosites were mapped after the application of each field.
+The magnetic images show that remanent magnetizations of Beaver River anorthosite (i.e.
+the individual dipoles visible with paired red +B$_z$ and blue -B$_z$ lobes) align well with the first applied field direction (E) and then rotated to align with the second applied field direction indicating minimal anisotropic behavior (G).
+In contrast, the remanence magnetizations carried by the magnetic needles in plagioclase 1 of the Duluth anorthosite xenolith align well with the first applied field but those in plagioclase 2 acquired an oblique remanence direction with respect to the field direction (F).
+After the application of an orthogonal 300 mT external field, magnetization of those needles in plagioclase 1 did not change direction due to strong shape anisotropy whereas the magnetization of needles in plagioclase 2 flipped, but with the acquired remanence remaining oblique to the field direction (H).
+This experiment illustrates that the Beaver River anorthosite xenolith has much lower magnetic anisotropy than the Duluth anorthosite xenolith.
+The unit of the axes in the QDM maps are in $\mu$m.}}
 \label{fig:Petro_QDM}
@@
 \ref{fig:Petro_QDM}A).
-In addition, the plagioclase crystals of the anorthosite xenoliths are typically free of large oxide inclusions with high aspect ratios (Fig.
+In addition, the plagioclase crystals of the anorthosite xenoliths are typically free of oxide inclusions 10s of microns or greater in length.
+with high aspect ratios (Fig.
 \ref{fig:Petro_QDM}C).
-Magnetic imaging using a quantum diamond microscope (QDM) shows that the dominant magnetic remanence carriers often exist within, and sometimes interstitial to, plagioclase crystals (Fig.
+Magnetic imaging using a quantum diamond microscope (QDM) shows that the dominant magnetic remanence carriers exist within, and sometimes interstitial to, plagioclase crystals (Fig.
 \ref{fig:Petro_QDM}E, G).
@@
 The features of the Beaver River anorthosite xenoliths are distinct from many other anorthosites such as the anorthosite xenoliths within the Duluth Complex Anorthositic Series rocks---older intrusions within the Midcontinent Rift.
-To illustrate these difference, we present petrographic and magnetic imaging data from a sample of Duluth Complex anorthosite that was not targeted for paleointensity experiments.
+To illustrate these differences, we present petrographic and magnetic imaging data from a sample of Duluth Complex anorthosite that was not targeted for paleointensity experiments.
 The plagioclase of the Duluth Complex anorthosites typically develop interlocking textures that display pronounced igneous foliation (Fig.
@@
 \ref{fig:Petro_QDM}D, F, H).
-Magnetic imaging also finds that orientations of magnetic needles within the Duluth plagioclase crystals depict the orientation of remanence magnetizations as is shown by the oblique acquired remanence to the direction of a first applied isothermal remanent magnetization (IRM) in Figure \ref{fig:Petro_QDM}F.
-Upon applying a second IRM with a field of 300 mT in the orthogonal direction, the remanent magnetic moments failed to rotate in plagioclase 1, reversed in plagioclase 2 but still along the easy axis depicted by the needle orientations that are oblique to the applied field direction (Fig.
+Magnetic imaging also finds that orientations of magnetite needles within the Duluth plagioclase crystals dictate the orientation of remanence magnetizations as is shown by magnetization being acquired at angles highly oblique to applied isothermal remanent magnetization (IRM) directions (Figure \ref{fig:Petro_QDM}F).
+In an experiment where a  second orthogonal IRM is applied, the remanent magnetic moments either fail to rotate the magnetization of the ferromagnetic inclusions in a given plagioclase grain or flip the magnetization by 180º remaining in a direction that is oblique to the applied field direction dictated by shape anisotropy (Fig.
 \ref{fig:Petro_QDM}F, H).
-In contrast, the remanent magnetizations imaged in the Beaver River anorthosite xenoliths align much more parallel to the applied field directions upon applying a set of orthogonal external fields (Fig.
+These experiments enable novel visualization of the grain-scale magnetic anisotropy of elongated exsolved (titano)magnetite inclusions within plagioclase that leads to magnetic anisotropy observed at the bulk sample scale \cite{Selkin2000a, Feinberg2006a}.
+In contrast, the remanent magnetizations of ferromagnetic grains imaged in the Beaver River anorthosite xenoliths, targeted for paleointensity in this study, align with the applied field directions and rotate in the experiment where an orthogonal field is applied (Fig.
 \ref{fig:Petro_QDM}E, G).
-To our knowledge, this is the first time that the magnetic anisotropy associated with plagioclase petrofabrics is captured by magnetic imaging at such scale.
-This set of orthogonal IRM experiment clearly shows that the elongated (titano)magnetite exsolutions within plagioclase feldspar are responsible for the magnetic anisotropy of plagioclase which have been observed before at bulk sample scale \cite{Selkin2000a, Feinberg2006a}.
+These results highlight that the population
+
+%To our knowledge, this is the first time that the magnetic anisotropy associated with plagioclase petrofabrics is captured by magnetic imaging at such scale.
+
+%Those new IRM maps look really spectacular- extremely convincing that your anorthites are very different and much less anisotropic than the usual ones!
+I would just add some text to the paper (and maybe the supplement if you run out of space) explaining these data and saying that this really shows what anisotropy looks like on the grain scale, i.e.
+dipoles oriented towards the easy axis most aligned to the IRM direction.
+I don't think I've seen such clear grain-scale B field maps of extreme anisotropy published elsewhere, so it would be good to emphasize this is a really novel measurement and technique for seeing the origin of anisotropy.
 
@@
 1092 Ma.
-To contextualize our high paleointensity estimates, we plot the regional mean values based on specimen-level paleointensity results together with with the calculated paleointensity values at 22\textdegree N latitude based on the VADM values of model PADM2M \cite{Ziegler2011a} in Figure \ref{fig:PINT_cooling_corrected}B.
-The locality-mean values of these three regions are plotted in Figure \ref{fig:PINT_cooling_corrected}C where we also present the distribution of means calculated from 3 random values from the PADM2M model for a total of 10000 iterations.
-In addition, we perform the same comparison with the existing paleointensity data for the Cenozoic Era (the last 66 million years) from the PINT database (filtered by Q$_{PI}\geq$3) given that the observational dataset could record more geomagnetic field excursions or regional high flux patches (Fig.
-S5).
+To contextualize our high paleointensity estimates, we plot the regional mean values (Table.
+S3) based on specimen-level paleointensity results together with with the calculated paleointensity values at 22\textdegree N latitude based on the VADM values of model PADM2M \cite{Ziegler2011a} in Figure \ref{fig:PINT_cooling_corrected}B.
+The locality-mean values of these three regions are plotted in Figure \ref{fig:PINT_cooling_corrected}C (and reported in Table.
+S3) where we also present the distribution of means calculated from 3 random values from the PADM2M model for a total of 10000 iterations.
+In addition, we perform the same comparison with the existing paleointensity data for the Cenozoic Era (the last 66 million years) from the PINT database (filtered by Q$_{PI}\geq$3; Fig.
+S5) given that the observational dataset could record more geomagnetic field excursions or regional high flux patches.
 The anorthosite xenoliths 3 region mean intensity is higher than all values from the resample results of the PADM2M model, and rival the highest values of the resample averages from the Cenozoic data (Fig.

### 2022-02-18 PNAS_manuscript/cover_letter.tex
Updates from Overleaf

@@
 
-\hfill 2022
+\hfill \today
 %{\noindent\textit {Yiming Zhang, UC Berkeley}}
@@
 
- This letter accompanies our submission of a manuscript entitled ``High geomagnetic field intensity recorded by anorthosite xenoliths requires a strongly powered late Mesoproterozoic geodynamo".
+This letter accompanies our submission of a manuscript entitled ``High geomagnetic field intensity recorded by anorthosite xenoliths requires a strongly powered late Mesoproterozoic geodynamo."
 
- Acquiring high-fidelity observational records of ancient magnetic field intensity from the remanent magnetization of rocks is crucial for constraining the long-term evolution of Earth's core.
-However, robust estimates of ancient field strengths are often difficult to recover due to alteration or non-ideal rock magnetic behavior.
-In this contribution, we develop high-quality paleointensity data with the best site-level standard deviation of 0.04 $\mu$T from well-preserved anorthosite xenoliths of the Beaver River diabase which were rapidly emplaced during the ca.
+Earth's long-term thermal evolution and the timing of inner core nucleation is currently very difficult to solve from a theoretical perspective as there continue to be conflicting estimates from the mineral physics community on relevant thermal conductivity values.
+While some in the paleomagnetic community have embraced a late timing of inner core nucleation associated with high thermal conductivity values, such an interpretation is reliant on the interpretation that weak fields ca.
+565 Ma capture a near demise of a thermally driven dynamo that was followed by inner core nucleation.
+Given the sparse paleointensity database of the Proterozoic, acquiring new high-fidelity observational records of ancient magnetic field intensity from the remanent magnetization of rocks is crucial for constraining the long-term evolution of Earth's core from an observation perspective.
+However, robust estimates of ancient field strengths are often difficult to recover due to alteration or non-ideal rock magnetic behavior that is found in many igneous rocks.
+
+In this study, we target a quite unique rock type which are lower crustal anorthosite cumulates that were brought to the near surface where they cooled and acquired their magnetization 1.1 billion years ago.
+In contrast to the diabase that hosts the xenoliths which exhibits non-ideal behavior, the anorthosites are high fidelity paleointensity recorders.
+These anorthosites are different than those typical of layered mafic intrusions as they formed deeper in the crust.
+We are able to use $\mu$m-scale magnetic maps generated with a quantum diamond microscope to show that in contrast to typical anorthosites with exsolved magnetite needles there is minimal grain-scale anisotropy associated with the small magnetite hosted within the plagioclase crystals.
+The data that emerges from this unique archive shows that the geomagnetic field was quite strong 1.1 billion years ago --- stronger than any time in the past 2 million years and rivaling some of the strongest dipole moments of the entire database.
+Rather than there being a progressive monotonic decay of Earth's dynamo strength through the Proterozoic Eon, there was a strongly powered dynamo in the late Mesoproterozoic.
+
+Multiple observed transitions from weak to strong paleointensity values in the Phanerozoic and the Proterozoic cannot all be the minimum prior to the beginning of inner core nucleation.
+Instead they likely indicate are the result of plate tectonics modulating core-mantle heat flow patterns in the Proterozoic as in the Phanerozoic (giving rise to intervals such as the mid-Paleozoic dipole low).
+We are excited about the results presented in this manuscript and believe that the readers of \textit{Proceedings of the National Academy of Sciences} will find them compelling and a significant advance.
+
+Sincerely,
+
+Yiming Zhang\\
+
+% PhD student of Earth and Planetary Science\\
+
+% University of California, Berkeley
+
+% In this contribution, we develop high-quality paleointensity data with the best site-level standard deviation of 0.04 $\mu$T from well-preserved anorthosite xenoliths of the Beaver River diabase which were rapidly emplaced during the ca.
 1.1 Ga North America Midcontinent Rift.
@@
 
-  Our updated paleointensity data compilation show multiple observed paleointensity transitions from weak to strong in the Phanerozoic and the Proterozoic.
+%   Our updated paleointensity data compilation show multiple observed paleointensity transitions from weak to strong in the Phanerozoic and the Proterozoic.
 Given that these records cannot all be the minimum prior to the beginning of inner core nucleation, they instead could indicate that processes such as plate tectonics could have modulated the core-mantle heat flow pattern in the Proterozoic.
@@
 
-We are excited about the results presented in this manuscript and believe that the Earth science community and readers of \textit{Proceedings of the National Academy of Sciences} will find them compelling and a significant advance.
+%
 
-Sincerely,
+% \clearpage
 
-Yiming Zhang\\
+% \textbf{FAIR data availability:}\\
+% The measurement level paleomagnetic data are in the FAIR data repository MagIC and a link within the manuscript makes these data available for reviewers.
 
-PhD student of Earth and Planetary Science\\
+% \textbf{Suggested editor reviewers:}\
 
-University of California, Berkeley
+% Lisa Tauxe (ltauxe\@ucsd.edu)\
 
-\clearpage
+% \textit{expertise in paleointensity and particularly in the behavior of the ancient geomagnetic field}
 
-\textbf{FAIR data availability:}\\
-The measurement level paleomagnetic data are in the FAIR data repository MagIC and a link within the manuscript makes these data available for reviewers.
+% Joshua Feinberg (feinberg\@umn.edu)\
 
-\textbf{Suggested editor reviewers:}\
+% \textit{expertise in rock magnetism and paleomagnetism}
 
-Lisa Tauxe (ltauxe\@ucsd.edu)\
+% Andrew Biggin (A.Biggin\@liverpool.ac.uk)\
 
-\textit{expertise in paleointensity and particularly in the behavior of the ancient geomagnetic field}
+% \textit{expertise in paleomagnetism and particularly in the long term evolution of Earth interior}
 
-Joshua Feinberg (feinberg\@umn.edu)\
+% Peter Selkin (paselkin\@uw.edu)\
 
-\textit{expertise in rock magnetism and paleomagnetism}
+% \textit{expertise in paleomagnetism and particularly in the  magnetic properties of magnetic minerals in feldpsar}
 
-Andrew Biggin (A.Biggin\@liverpool.ac.uk)\
+% Peter Driscoll (pdriscoll\@carnegiescience.edu)\
 
-\textit{expertise in paleomagnetism and particularly in the long term evolution of Earth interior}
-
-Peter Selkin (paselkin\@uw.edu)\
-
-\textit{expertise in paleomagnetism and particularly in the  magnetic properties of magnetic minerals in feldpsar}
-
-Peter Driscoll (pdriscoll\@carnegiescience.edu)\
-
-\textit{expertise in numerical modeling of Earth's thermal and magnetic evolution}
-\textit{}
+% \textit{expertise in numerical modeling of Earth's thermal and magnetic evolution}
+% \textit{}
 \end{flushleft}

### 2022-02-18 PNAS_manuscript/PNAS_manuscript.tex

@@
 In this study, we use plagioclase cumulates (anorthosite) that formed in the deep crust and were brought to the near surface by magma where they cooled and acquired thermal remanent magnetizations.
-These anorthosite xenoliths have experienced minimal alteration since their emplacement and yield high quality paleointensity estimates.
-In contrast to predictions of a progressively decaying field leading up to late nucleation of Earth's inner core, these data record a strong virtual dipole moment 1.1 billion years ago.
-A strong field persisted over at least a 20-million-year interval indicating the existence of appreciable power sources for Earth's dynamo during the late Mesoproterozoic.}
+These anorthosite xenoliths have experienced minimal alteration since their emplacement and yield high-quality paleointensity estimates.
+In contrast to scenarios of a progressively decaying field leading up to late nucleation of Earth's inner core, these data record a strong virtual dipole moment 1.1 billion years ago.
+A strong field persisted over at least a 14-million-year interval indicating the existence of appreciable power sources for Earth's dynamo during the late Mesoproterozoic.}
 
@@
 % At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
-\keywords{absolute paleointensity $|$ Midcontinent Rift $|$ anorthosite $|$ geodynamo $|$ Mesoproterozoic $|$ inner core}
+\keywords{absolute paleointensity $|$ anorthosite $|$ geodynamo $|$ Proterozoic $|$ inner core}
 
@@
 
-% However, the ability for the geomagnetic dipole field to reach similar low values in rather recent times when we known the inner core exists (Fig.
-\ref{fig:PINT_compilation}) makes it challenging to unambiguously settle the timing of the initiation of inner core nucleation based on the current few paleointensity records.
-
 \begin{SCfigure*}
@@
 For such data, distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope.
-Typically, such non-ideal behavior would result in a higher paleointensity estimate from the steeper-sloped low-temperature portion and a lower paleointensity estimate from the high-temperature portion.
+Typically, such non-ideal behavior would result in a higher paleointensity estimate from the steeper-sloped low-temperature portion of the experiment and a lower paleointensity estimate from the high-temperature portion.
 For example, in data from the Midcontinent Rift, ref.
@@
 \citealp{Kulakov2013a} used the high-temperature slope (likely underestimating the field strength).
-Such non-ideal results were rejected by \citealp{Sprain2018a} who applied stricter paleointensity selection criteria, but as a result had few accepted sites.
+Such non-ideal results were rejected by ref.
+\citealp{Sprain2018a} who applied stricter paleointensity selection criteria, but as a result had few accepted sites.
 
@@
 Rock magnetic data reveal that the anorthosite xenoliths have low anisotropy of thermal remanent magnetization (TRM) and can acquire TRM linearly within the range of relevant field strengths.
-Magnetic imaging shows the anorthosite specimens have dominant magnetic carriers within and interstitial to plagioclase crystals, and they do not display strong preferred orientations.
+Magnetic imaging shows that the anorthosite specimens have dominant magnetic carriers within and interstitial to plagioclase crystals without strong preferred orientations.
 These single-slope, high-quality paleointensity data, together with other high-quality paleointensity records during the protracted Midcontinent Rift magmatic activity, require that there was a strong late Mesoproterozoic geodynamo.
@@
 \noindent\includegraphics[width=13 cm]{Petro_QDM_full.png}
-\caption{\footnotesize{Thin section petrographic images (A,B,C,D) and magnetic maps (E,F) of anorthosite from the Beaver River anorthosite xenolith from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (left column;  \citealp{Zhang2021b}) and a distinct anorthosite from the Duluth Complex Anorthositic Series (right column).
-The Duluth Complex anorthosites were not targeted for paleointensity experiments in this study given the complexities associated with more pronounced fabrics.
-Cross-polarized petrographic images of the Beaver River anorthosite (A,C) reveal plagioclase with a granoblastic texture with crystals that are largely free of large opaque inclusions.
+\caption{\footnotesize{Thin section petrographic images (A,B,C,D) and magnetic maps (E,F,G,H) of anorthosite from the Beaver River anorthosite xenolith in the Silver Bay region from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (left column;  \citealp{Zhang2021b}) and a distinct anorthosite from the Duluth Complex Anorthositic Series (right column).
+The Duluth Complex anorthosites were not targeted for paleointensity experiments in this study given complexities associated with more pronounced fabrics.
+Cross-polarized petrographic images of the Beaver River anorthosite (A,C) reveal plagioclase with a granoblastic texture of crystals that are largely free of large opaque inclusions.
 In contrast, plagioclase crystals in the Duluth Complex anorthosite have euhedral, interlocking crystals with an igneous foliation (B) and the plagioclase crystals contain abundant Fe-Ti oxide needles that have preferred orientations that are often parallel with the [001] axis of the plagioclase.
-Maps of the vertical component of magnetic field (B$_z$) developed with a quantum diamond microscope (QDM) show relatively weak magnetic sources within and interstitial to plagioclase crystals in the Beaver River anorthosite (in E) relative to the strongly magnetic large oxide needles within Duluth Complex plagioclase (in F).
+Maps of the vertical component of magnetic field (B$_z$) developed with a quantum diamond microscope (QDM) show relatively weak magnetic sources within plagioclase crystals in the Beaver River anorthosite (in E) relative to the strongly magnetic large oxide needles within Duluth Complex plagioclase (in F).
 The B$_z$ color scale is an order of magnitude greater in the Duluth Complex anorthosite maps (F, H) than the maps for the Beaver River anorthosite (E, G).
@@
 \ref{fig:Petro_QDM}A).
-In addition, the plagioclase crystals of the anorthosite xenoliths are typically free of oxide inclusions 10s of microns or greater in length.
-with high aspect ratios (Fig.
+Plagioclase crystals within coarse-grained intrusions often contain abundant elongate Fe-Ti oxide inclusions that are visible through optical microscopy \cite{Feinberg2006a}.
+The Beaver River anorthosite xenoliths lack such large (10s of $\mu$m) oxide inclusions such that oxides are not visible within the plagioclase crystals through optical microscopy (Fig.
 \ref{fig:Petro_QDM}C).
-Magnetic imaging using a quantum diamond microscope (QDM) shows that the dominant magnetic remanence carriers exist within, and sometimes interstitial to, plagioclase crystals (Fig.
+However, magnetic imaging using a quantum diamond microscope (QDM) shows that there are magnetic remanence carriers within the plagioclase crystals (Fig.
 \ref{fig:Petro_QDM}E, G).
-
-The features of the Beaver River anorthosite xenoliths are distinct from many other anorthosites such as the anorthosite xenoliths within the Duluth Complex Anorthositic Series rocks---older intrusions within the Midcontinent Rift.
-To illustrate these differences, we present petrographic and magnetic imaging data from a sample of Duluth Complex anorthosite that was not targeted for paleointensity experiments.
-The plagioclase of the Duluth Complex anorthosites typically develop interlocking textures that display pronounced igneous foliation (Fig.
+To highlight these distinct aspects of the Beaver River anorthosite xenoliths, we present petrographic and magnetic imaging data from the Beaver River anorthosite as well as a sample of an anorthosite from the Duluth Complex Anorthositic Series---an older intrusive complex within the Midcontinent Rift that was not targeted for paleointensity experiments in this study (Fig.
+\ref{fig:Petro_QDM}).
+In contrast to the Beaver River anorthosite, the plagioclase of the Duluth Complex anorthosites typically develop interlocking textures that display pronounced igneous foliation (Fig.
 \ref{fig:Petro_QDM}B).
-In addition, there are abundant Fe-Ti oxide needles that are typically aligned with the [001] axes of the plagioclase crystals (Fig.
+In addition, there are abundant Fe-Ti oxide needles within the Duluth Complex plagioclase grains that are typically aligned with the [001] axes of the crystals (Fig.
 \ref{fig:Petro_QDM}D).
-Magnetic imaging confirms that these strongly magnetic needles within the plagioclase crystals often have magnetic moments oriented along their long axes (Fig.
-\ref{fig:Petro_QDM}D, F, H).
-Magnetic imaging also finds that orientations of magnetite needles within the Duluth plagioclase crystals dictate the orientation of remanence magnetizations as is shown by magnetization being acquired at angles highly oblique to applied isothermal remanent magnetization (IRM) directions (Figure \ref{fig:Petro_QDM}F).
-In an experiment where a  second orthogonal IRM is applied, the remanent magnetic moments either fail to rotate the magnetization of the ferromagnetic inclusions in a given plagioclase grain or flip the magnetization by 180º remaining in a direction that is oblique to the applied field direction dictated by shape anisotropy (Fig.
+Magnetic imaging confirms that these magnetite needles have magnetic moments oriented along their long axes (Fig.
+\ref{fig:Petro_QDM}F, H).
+As a result of this shape anisotropy, magnetization is acquired at angles highly oblique to applied fields (Figure \ref{fig:Petro_QDM}F).
+In an experiment where a second orthogonal field was applied, the first isothermal remanent magnetizations (IRM) of a set of magnetite needles within a plagioclase grain either fail to rotate or flip by 180\textdegree.
+When they flip, they remain in a direction that is oblique to the applied field direction dictated by shape anisotropy (Fig.
 \ref{fig:Petro_QDM}F, H).
 These experiments enable novel visualization of the grain-scale magnetic anisotropy of elongated exsolved (titano)magnetite inclusions within plagioclase that leads to magnetic anisotropy observed at the bulk sample scale \cite{Selkin2000a, Feinberg2006a}.
-In contrast, the remanent magnetizations of ferromagnetic grains imaged in the Beaver River anorthosite xenoliths, targeted for paleointensity in this study, align with the applied field directions and rotate in the experiment where an orthogonal field is applied (Fig.
+In contrast, the remanent magnetizations of the ferromagnetic grains imaged in the Beaver River anorthosite xenoliths, targeted for paleointensity in this study, align with the applied field directions indicating minimal remanence anisotropy  (Fig.
 \ref{fig:Petro_QDM}E, G).
-These results highlight that the population
-
-%To our knowledge, this is the first time that the magnetic anisotropy associated with plagioclase petrofabrics is captured by magnetic imaging at such scale.
+The relative lack of petrofabrics and minimal grain-scale magnetic anisotropy make the Beaver River anorthosite xenoliths a particularly compelling target for paleointensity experiments.
 
@@
 
-Titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex magma \cite{Miller1990a}.
+%Titanomagnetite and ilmenite symplectic intergrowths also coexist with pyroxene and relict olivine as a product of olivine oxidation in the Duluth Complex magma \cite{Miller1990a}.
 The lack of symplecites and the lack of large plagioclase-hosted titanomagnetite needles within the Beaver River anorthosite xenoliths targeted in this study distinguish them from plagioclase cumulates of the Duluth Complex and other layered mafic complexes where anisotropic magnetic mineral fabrics associated with igneous foliation often occur \cite{Scofield1986a, Selkin2000a, Feinberg2006a}.
-The relative lack of petrofabrics and associated preferred orientation makes the Beaver River anorthosite xenoliths a particularly compelling target for paleointensity experiments.
 
@@
 The green lines represent fits for the dominant single-slope component that passes the acceptance criteria and gives an estimate of the ancient field strength (B$_{anc}$).
-The plots for anorthosite specimens AX1-1a and diabase BD5-1b on the bottom row show typical non-ideal sagging behaviour that fails our acceptance criteria.
-Specimen BD2-5a is an example where the data appear linear with distinct slopes in the low and high temperature ranges such that it could pass less restrictive selection criteria, particularly if a narrower temperature range was used for the experiment.}}
+The plots for anorthosite specimens AX1-1a and diabase BD5-1b on the bottom row show non-ideal sagging behaviour that fails our acceptance criteria.
+Specimen BD2-5a is an example where the data appear linear with distinct slopes in the low and high temperature ranges such that it could pass less restrictive selection criteria, particularly if a narrower temperature range was used for the experiment.
+Data analysis and visualization was conducted using PmagPy \cite{Tauxe2016a}}}
 \label{fig:IZZI_examples}
@@
 \centering
-\caption{\footnotesize{A) Summary plot of individual specimen absolute paleointensity results (square symbols) and their averages and standard deviations at site level (black bars with grey uncertainty boxes).
+\caption{\footnotesize{A) Summary plot of individual specimen absolute paleointensity results (square symbols) and their averages and standard deviations at site level (black bars with grey uncertainty boxes) where each `AX' site is an individual anorthosite xenolith within the Beaver River diabase.
 All results are corrected for cooling rate with a factor of 0.74.
-Each `AX' site is an individual anorthosite xenolith within the Beaver River diabase.
-The sites with successful experiments come from 3 regions which would have cooled at distinct times yielding similar estimates within the each region with differences between regions.
-B) Specimen level means calculated for these regions are compared to the distribution of intensities calculated from a paleomagnetic axial dipole moment model for the past 2 million years (PADM2M; \citealp{Ziegler2011a}) at the latitude corresponding to the paleolatitude of study region (22\textdegree N).
+The sites with successful experiments come from 3 regions (Carlton Peak, Silver Bay, Kennedy Creek) which would have cooled at distinct times yielding similar estimates within the each region with differences between regions.
+B) Regional means calculated from the specimen-level data are compared to the distribution of intensities calculated from a paleomagnetic axial dipole moment model for the past 2 million years (PADM2M; \citealp{Ziegler2011a}) at the latitude corresponding to the paleolatitude of study region (22\textdegree N).
 C) The mean of the 3 regional means is compared to means calculated from 3 random values drawn from the PADM2M model \cite{Ziegler2011a}.
-The distribution represents a total of 10,000 iterations of taking 3 random draws and calculating the mean.}}
+The distribution represents a total of 10,000 iterations of taking 3 random draws and calculating the mean.
+These comparisons highlight that these paleointensity values are strong in relation to the geomagnetic field over the past 2 million years.}}
 \label{fig:PINT_cooling_corrected}
@@
 
-In addition to estimating paleointensity values by introducing a set of selection criteria to filter our experiment results, we apply an independent method of \citealp{Cych2021a} which uses all experimental data regardless of their NRM/TRM plot statistics to perform bias corrected estimation of paleointensity (BiCEP).
+In addition to estimating paleointensity values by introducing a set of selection criteria to filter our experiment results, we apply an independent method from ref.
+\citealp{Cych2021a} with all experimental data regardless of their NRM/TRM plot statistics to perform bias corrected estimation of paleointensity (BiCEP).
 This Bayesian probabilistic method is based on an assumption that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Paterson2011a}.
@@
 \subsection*{Rock magnetism}
-
-Rock magnetic data support that the anorthosite specimens that pass the paleointensity selection criteria have dominant magnetic remanence carriers with magnetic properties similar to stochiometric, non-interacting, single domain magnetite, whereas anorthosite samples that failed the paleointensity result selection and all diabase samples have more pronounced populations of non-ideal carriers.
+\subsubsection*{Coercivity}
+Additional rock magnetic data support that the anorthosite specimens that pass the paleointensity selection criteria have dominant magnetic remanence carriers with magnetic properties similar to stochiometric, non-interacting, single domain magnetite, whereas anorthosite samples that failed the paleointensity result selection and all diabase samples have more pronounced populations of non-ideal carriers.
 Magnetic property measurement system (MPMS) data show both diabase and anorthosite contain (titano)magnetite as evidenced through the presence of the Verwey transition (Fig.
@@
 S2), indicating that magnetite grains in the specimens either have relatively higher Ti content or have been partially oxidized \cite{Ozdemir1993a}.
-
-Coercivity spectra from backfield experiments further support the interpretation that rocks that pass or fail the selection criteria have distinct magnetic mineralogy (Fig.
+Another difference is that samples which pass paleointensity selection criteria have distinctly higher average median destructive field values than other anorthosite and diabase specimens (Fig.
 \ref{fig:coercivity}).
-A compilation of median destructive field (MDF) values in Figure \ref{fig:coercivity} shows that specimens from anorthosites that pass paleointensity selection criteria have distinctly higher average MDF values than other anorthosite and diabase specimens.
 Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity around 80 mT (Fig.
@@
 
-\subsection*{TRM anisotropy and nonlinear acquisition check}
+\subsubsection*{Bulk TRM anisotropy}
 Significant remanence anisotropy has been documented to exist within certain anorthositic rocks that form in layered intrusive complexes \cite{Selkin2000a, Feinberg2006a}.
 Strong remanence anisotropy associated with the igneous foliation developed within anorthosite from the Stillwater Complex has been shown to lead to significant overestimation or underestimation of paleointensity values depending on the relative orientations between the fabrics and an applied magnetic field \cite{Selkin2000a}.
-To assess whether our paleointensity estimates are biased by remanence anisotropy, we calculated the gamma statistic, which is the angular difference between the last pTRM step of paleointensity experiment and the applied field direction.
+To assess whether our paleointensity estimates are biased by bulk remanence anisotropy, we calculated the gamma statistic, which is the angular difference between the last pTRM step of paleointensity experiment and the applied field direction.
 The results show that the anorthosite specimens used for paleointensity experiment have low gamma values ranging from 0.9\textdegree$\;$ to 11.9\textdegree, with a median value of 4.2\textdegree$\;$(Table S1).
@@
 Paleodirectional data from our anorthosite xenoliths further support that they have minimal remanence anisotropy as their site mean directions closely match those of the Beaver River diabase hosts without deviating due to a fabric \cite{Zhang2021b}.
-In addition, no resolvable magnetic fabric was observed in the anorthosite xenoliths through petrographic and magnetic imaging (Fig.
+These bulk results are consistent with the data from the QDM magnetic map experiments that reveal minimal remenance anisotropy (Fig.
 \ref{fig:Petro_QDM}).
-
-We applied full TRMs to a set of anorthosite specimens with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axes.
-The results show that these anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions (\citealp{Selkin2007a}; Fig.
-S4).
 
@@
 Thermal modeling results from ref.
-\citealp{Zhang2021b} suggest that the Beaver River anorthosite xenoliths were heated to tholeiitic magma temperature ($\sim$1100\textdegree C) and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a thousand years, partially averaging secular variation within single sites.
-Anorthosite xenoliths have paleointensity results that are consistent within small regions, but vary between regions.
-Anorthosite AX12 and AX16 in the Silver Bay area were emplaced $\sim$450 meters apart and have indistinguishable paleointensity estimates (Figs.
-\ref{fig:Geologic_map} and \ref{fig:PINT_cooling_corrected}).
-$\sim$10 km to the north, anorthosite xenoliths AX11 and AX13 in the Kennedy Creek area were also emplaced closely (separated by $\sim$125 meters) and yield similar values to one another but distinct paleointensity estimates from those of AX12 and AX16 (Figs.
-\ref{fig:Geologic_map} and \ref{fig:PINT_cooling_corrected}).
-$\sim$30 km further to the north in the Carlton Peak region, AX6, AX8, and AX10 are anorthosite xenoliths within 10 meters and also yield similar paleointensity estimates albeit with relatively large uncertainties (Fig.
-\ref{fig:PINT_cooling_corrected}).
-Therefore, one interpretation could consider the anorthosites of the three distinct regions to have captured at least three different records of geomagnetic field intensities during the emplacement and cooling of the Beaver River diabase sills ca.
-1092 Ma.
-To contextualize our high paleointensity estimates, we plot the regional mean values (Table.
-S3) based on specimen-level paleointensity results together with with the calculated paleointensity values at 22\textdegree N latitude based on the VADM values of model PADM2M \cite{Ziegler2011a} in Figure \ref{fig:PINT_cooling_corrected}B.
-The locality-mean values of these three regions are plotted in Figure \ref{fig:PINT_cooling_corrected}C (and reported in Table.
-S3) where we also present the distribution of means calculated from 3 random values from the PADM2M model for a total of 10000 iterations.
-In addition, we perform the same comparison with the existing paleointensity data for the Cenozoic Era (the last 66 million years) from the PINT database (filtered by Q$_{PI}\geq$3; Fig.
-S5) given that the observational dataset could record more geomagnetic field excursions or regional high flux patches.
-The anorthosite xenoliths 3 region mean intensity is higher than all values from the resample results of the PADM2M model, and rival the highest values of the resample averages from the Cenozoic data (Fig.
-S5).
-These data supports that the anorthosite xenoliths likely record an exceptionally strong geomagnetic field ca.
-1092 Ma.
-
+\citealp{Zhang2021b} indicate that the Beaver River anorthosite xenoliths were heated to tholeiitic magma temperature ($\sim$1100\textdegree C) and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a thousand years, partially averaging secular variation within single sites.
 Another consequence of the anorthosite xenoliths having cooled in the interior of thick diabase intrusions is that slow cooling rates can bias paleointensity estimates toward higher values.
@@
 
+Anorthosite xenoliths have paleointensity results that are consistent within small regions, but vary between regions.
+Anorthosite AX12 and AX16 in the Silver Bay area were emplaced $\sim$450 meters apart and have indistinguishable paleointensity estimates (Figs.
+\ref{fig:Geologic_map} and \ref{fig:PINT_cooling_corrected}).
+$\sim$10 km to the north, anorthosite xenoliths AX11 and AX13 in the Kennedy Creek area were also emplaced closely (separated by $\sim$125 meters) and yield similar values to one another, but distinct paleointensity estimates from those of AX12 and AX16 (Figs.
+\ref{fig:Geologic_map} and \ref{fig:PINT_cooling_corrected}).
+$\sim$30 km further to the north in the Carlton Peak region, AX6, AX8, and AX10 are anorthosite xenoliths within 10 meters of one another and also yield similar paleointensity estimates albeit with relatively large uncertainties (Fig.
+\ref{fig:PINT_cooling_corrected}).
+The anorthosites of the three distinct regions captured three different intervals of geomagnetic field intensities during the emplacement and cooling of the Beaver River diabase sills ca.
+1092 Ma.
+To contextualize the high paleointensity values, we plot the regional mean values (Table.
+S3) based on specimen-level paleointensity results together with calculated paleointensity values at 22\textdegree N latitude based on the PADM2M model of the time-varying geomagnetic field over the past 2 million years \cite{Ziegler2011a} in Figure \ref{fig:PINT_cooling_corrected}B.
+The locality-mean values of these three regions are plotted in Figure \ref{fig:PINT_cooling_corrected}C (and reported in Table.
+S3) where we also present the distribution of means calculated from 3 random values from the PADM2M model for a total of 10000 iterations.
+In addition, we perform the same comparison with the existing paleointensity data for the Cenozoic Era (the last 66 million years) from the PINT database (filtered by Q$_{PI}\geq$3; Fig.
+S5) given that the observational dataset could record more geomagnetic field excursions or regional high flux patches.
+The anorthosite xenoliths 3 region mean intensity is higher than all values from the resample results of the PADM2M model, and rival the top 3\% values of the resample averages from the Cenozoic data (Fig.
+S5).
+These data supports that the anorthosite xenoliths likely record an exceptionally strong geomagnetic field ca.
+1092 Ma.
+
 \begin{figure*}[h!]
@@
 The highest site-mean virtual dipole moment of the anorthosites would rank in the top 2$\%$ of those in the database for the Cenozoic Era (the last 66 million years) when there was unequivocally a crystallizing inner core.
-The y axis range is set such that the top 1\% of the Cenozoic paleointensity data is cut off.}}
+The y axis maximum is set at the 99 percentile of the compiled Cenozoic paleointensity data.}}
 \label{fig:PINT_compilation}
@@
 However, other experimental studies continue to indicate lower thermal conductivity values consistent with prior estimates \cite{Konopkova2016a, Hsieh2020a} with no consensus yet emerging \cite{Williams2018a, Ohta2021a}.
-These experiments are challenging to conduct and interpret given complexities such as constraining the sample thickness under high pressure and temperature conditions, the validity of applying the Wiedemann-Franz law to extrapolate thermal conductivity values based on electrical resistivity measurements \cite{Ohta2016a}, and modeling uncertainties associated with parameters used in direct thermal conduction measurement experiments \cite{Konopkova2016a}.
+These experiments are challenging to conduct and interpret given complexities such as constraining the sample thickness under high pressure and temperature conditions, the validity of applying the Wiedemann-Franz law to extrapolate thermal conductivity values based on electrical resistivity measurements \cite{Ohta2016a}, and propagating uncertainties from free parameters used in finite element modeling of direct thermal conduction experiments \cite{Konopkova2016a}.
 Further experiments and theory are needed to explain these contrasting results which at present leave open very different trajectories for Earth's thermal evolution.
@@
 Other scenarios are possible, however, such as the model-based prediction that while power increases associated with inner core nucleation strengthen Earth's internal magnetic field, that the dynamo becomes more deeply seated in the core diminishes the increase in magnetic field strength at Earth's surface \cite{Aubert2009a, Landeau2017a}.
-Such a shift of the dynamo to a greater depth associated with inner core nucleation led \cite{Aubert2009a} to conclude that the increase in power to the dynamo would be difficult to detect with paleointensity data.
+Such a scenario where the dynamo shifts to a greater depth associated with inner core nucleation led \cite{Aubert2009a} to conclude that the increase in power to the dynamo would be difficult to detect with paleointensity data.
 Ultimately, further observational paleomagnetic records is key as they hold the potential for testing different model predictions and identifying transitions in ancient field strength \cite{Biggin2015a, Bono2019a}.
@@
 The well-preserved ca.
-1092 Ma anorthosite xenoliths of the Beaver River diabase record a strong geomagnetic field in the late Mesoproterozoic that exceeds the strength of the modern-day field for which crystallization of the inner core is a power source (Fig.
-\ref{fig:PINT_compilation}).
+1092 Ma anorthosite xenoliths of the Beaver River diabase record a strong geomagnetic field in the late Mesoproterozoic that exceeds the strength of the modern-day field for which crystallization of the inner core is a power source (Figs.
+\ref{fig:PINT_cooling_corrected} and \ref{fig:PINT_compilation}).
 Together with previous records obtained from the ca.
@@
 However, Fig.
-\ref{fig:PINT_compilation} shows that transitions from low to high field intensities occurred before, during, and after the Ediacaran.
+\ref{fig:PINT_compilation} shows that transitions from low to high field intensities occurred before, during, and after the Ediacaran Period.
 In the Ediacaran record developed to date, there is a two-fold increase in Earth's virtual dipole moment when comparing estimates from the ca.
@@
 1.1 Ga timing would be broadly consistent with the ca.
-1.3 Ga onset proposed by \citealp{Biggin2015a} albeit later given the exclusion of previous overestimated paleointensity values from the ca.
-1.3 Ga Gardar basalts that are superseded by data in \cite{Kodama2019a}.
+1.3 Ga onset proposed by ref.
+\citealp{Biggin2015a} albeit later given the exclusion of previous overestimated paleointensity values from the ca.
+1.3 Ga Gardar basalts that are superseded by data in ref.
+\citealp{Kodama2019a}.
 However, a model prediction of sustained strong field values following inner core nucleation is challenged by data from the ca.

### 2022-02-18 PNAS_manuscript/Supporting Info.tex

@@
 \usepackage{rotating}
+\usepackage{multirow}
 \templatetype{pnassupportinginfo}
@@
 \end{figure*}
+
+%To check for nonlinear acqusition, we applied full TRMs to a set of anorthosite specimens with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axes.
+The results show that these anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions (\citealp{Selkin2007a}; Fig.
+S4).
 
@@
 The statistics names are the same as in Table \ref{tab:PINT_result}}.
-n represents the total number of specimen- or site-level results used for calculating mean values.
-inc$_{tc}$ is the tilt corrected mean inclination.
+n represents the total number of specimen- or site- or region-level results used for calculating mean values.
+inc$_{tc}$ is the tilt-corrected mean inclination.
 Site-level means for AX6, AX8, AX10, AX11, AX12, AX13, AX16 are calculated based on specimen results.
 Regional means for Carlton Peak, Kennedy Creek, and Silver Bay are calculated by grouping specimen results from AX6 and AX8 and AX10, AX11 and AX13, AX12 and AX16, respectively.
-Overall site mean values are calculated using all site level results in this table.
-Overall region mean values are calculated using all regional mean results in this table.
+Overall site mean values are calculated using all site-level results in this table.
+Overall region-mean values are calculated using all regional mean results in this table.
 }
-\begin{tabular}{cccccccccc}
-Group               & B$_{anc}$ & n  & FRAC & GAP-MAX & MAD ($^\circ$) & DANG ($^\circ$) & inc$_{tc}$ & $\gamma$ & VDM (ZAm$^2$) \\
-AX6                 & 30.33     & 3  & 0.69 & 0.13    & 3.55           & 2.66            & 38.9       & 2.63     & 48.76         \\
-AX8                 & 52.4      & 3  & 0.66 & 0.17    & 4.28           & 2.15            & 40.3       & 7.3      & 83.14         \\
-AX10                & 46.82     & 3  & 0.7  & 0.23    & 4.98           & 2.14            & 36.6       & 6.1      & 76.81         \\
-AX12                & 38.1      & 6  & 0.75 & 0.22    & 3.26           & 2.53            & 55.2       & 5.98     & 51.31         \\
-AX16                & 44.15     & 11 & 0.68 & 0.19    & 4.19           & 1.96            & 49.2       & 3.97     & 63.87         \\
-AX11                & 73.34     & 7  & 0.68 & 0.19    & 3.85           & 1.82            & 35.2       & 6.63     & 121.73        \\
-AX13                & 76.63     & 7  & 0.73 & 0.2     & 5.7            & 1.67            & 35.1       & 5        & 127.29        \\
-Carlton Peak        & 43.18     & 9  & 0.68 & 0.17    & 4.27           & 2.32            & 38.6       & 5.34     & 69.57         \\
-Kennedy Creek       & 74.98     & 14 & 0.7  & 0.2     & 4.78           & 1.74            & 35.15      & 5.81     & 124.51        \\
-Silver Bay          & 42.02     & 17 & 0.71 & 0.2     & 3.86           & 2.16            & 51.32      & 4.68     & 59.43         \\
-overall site mean   & 51.68     &    &      &         &                &                 &            &          & 81.84         \\
-overall region mean & 53.39     &    &      &         &                &                 &            &          & 84.5         
+\begin{tabular}{ccccccccccc}
+\hline
+\multicolumn{1}{l}{}                       &               & B$_{anc}$ & n  & FRAC & GAP-MAX & MAD ($^\circ$) & DANG ($^\circ$) & inc$_{tc}$ & $\gamma$ & VDM (ZAm$^2$) \\
+\hline
+\multirow{7}{*}{specimen mean for sites}   & AX6           & 30.33     & 3  & 0.69 & 0.13    & 3.55           & 2.66            & 38.9       & 2.63     & 48.76         \\
+                                           & AX8           & 52.4      & 3  & 0.66 & 0.17    & 4.28           & 2.15            & 40.3       & 7.3      & 83.14         \\
+                                           & AX10          & 46.82     & 3  & 0.7  & 0.23    & 4.98           & 2.14            & 36.6       & 6.1      & 76.81         \\
+                                           & AX12          & 38.1      & 6  & 0.75 & 0.22    & 3.26           & 2.53            & 55.2       & 5.98     & 51.31         \\
+                                           & AX16          & 44.15     & 11 & 0.68 & 0.19    & 4.19           & 1.96            & 49.2       & 3.97     & 63.87         \\
+                                           & AX11          & 73.34     & 7  & 0.68 & 0.19    & 3.85           & 1.82            & 35.2       & 6.63     & 121.73        \\
+                                           & AX13          & 76.63     & 7  & 0.73 & 0.2     & 5.7            & 1.67            & 35.1       & 5        & 127.29        \\ 
+                                           \hline
+\multirow{3}{*}{specimen mean for regions} & Carlton Peak  & 43.18     & 9  & 0.68 & 0.17    & 4.27           & 2.32            & 38.6       & 5.34     & 69.57         \\
+                                           & Kennedy Creek & 74.98     & 14 & 0.7  & 0.2     & 4.78           & 1.74            & 35.15      & 5.81     & 124.51        \\
+                                           & Silver Bay    & 42.02     & 17 & 0.71 & 0.2     & 3.86           & 2.16            & 51.32      & 4.68     & 59.43         \\
+\hline
+overall site mean                          &               & 51.68     & 7  &      &         &                &                 &            &          & 81.84         \\
+overall region mean                        &               & 53.39     & 3  &      &         &                &                 &            &          & 84.5         
 \end{tabular}

### 2022-02-25 PNAS_manuscript/PNAS_manuscript.tex
Updates from Overleaf

@@
 \centering
-\caption{\footnotesize{A) Summary plot of individual specimen absolute paleointensity results (square symbols) and their averages and standard deviations at site level (black bars with grey uncertainty boxes) where each `AX' site is an individual anorthosite xenolith within the Beaver River diabase.
+\caption{\footnotesize{A) Summary plot of individual specimen absolute paleointensity results (square symbols) and their averages and standard deviations at site level (black bars with grey 1$\sigma$ uncertainty boxes) where each `AX' site is an individual anorthosite xenolith within the Beaver River diabase.
 All results are corrected for cooling rate with a factor of 0.74.
@@
 
-Given that the multiple records of a weak field cannot all be the minimum prior to inner core nucleation, what processes could lead to a weak dipole at Earth's surface even in the presence of a crystallizing inner core?
+Given that the multiple records of a weak field in the Proterozoic and Paleozoic cannot all be the minimum prior to the singular event of the onset of inner core nucleation, what processes could lead to a weak dipole at Earth's surface even in the presence of a crystallizing inner core?
 Numerical models have shown that the dipole moment is sensitive to both the magnitude and spatial pattern of heat flow across the core-mantle boundary when there are strong available power sources to the geodynamo \cite{Olson2007a, Olson2010a}.
@@
 Sun compass orientations were preferentially used for determining the sample azimuth.
-Sister specimens underwent step-wise alternating field (AF) or thermal demagnetization at the UC Berkeley Paleomagnetism Lab to isolate paleomagnetic directions (data presented in \cite{Zhang2021b}).
+Sister specimens underwent step-wise alternating field (AF) or thermal demagnetization at the UC Berkeley Paleomagnetism Lab to isolate paleomagnetic directions (data presented in \citealp{Zhang2021b}).
 Based on the anorthosite thermal demagnetization results, we selected sites whose unblocking temperature ranges are narrow and near 580\textdegree C for paleointensity experiments.

### 2022-02-25 PNAS_manuscript/cover_letter.tex
Updates from Overleaf

@@
 
-Dear \textit{Proceedings of the National Academy of Sciences} editors,
+Dear \textit{PNAS} editors,
 
@@
 In contrast to the diabase that hosts the xenoliths which exhibits non-ideal behavior, the anorthosites are high fidelity paleointensity recorders.
-These anorthosites are different than those typical of layered mafic intrusions as they formed deeper in the crust.
+These anorthosite xenoliths are different than those typical of layered mafic intrusions as they formed deeper in the crust, but then acquired their magnetization near to the surface.
 We are able to use $\mu$m-scale magnetic maps generated with a quantum diamond microscope to show that in contrast to typical anorthosites with exsolved magnetite needles there is minimal grain-scale anisotropy associated with the small magnetite hosted within the plagioclase crystals.
@@
 Multiple observed transitions from weak to strong paleointensity values in the Phanerozoic and the Proterozoic cannot all be the minimum prior to the beginning of inner core nucleation.
-Instead they likely indicate are the result of plate tectonics modulating core-mantle heat flow patterns in the Proterozoic as in the Phanerozoic (giving rise to intervals such as the mid-Paleozoic dipole low).
-We are excited about the results presented in this manuscript and believe that the readers of \textit{Proceedings of the National Academy of Sciences} will find them compelling and a significant advance.
+Instead, they likely are the result of plate tectonics modulating core-mantle heat flow patterns in the Proterozoic as in the Phanerozoic (giving rise to intervals such as the mid-Paleozoic dipole low).
+We are excited about the results presented in this manuscript and believe that the readers of \textit{PNAS} will find them compelling and a significant advance.
 

### 2022-02-25 PNAS_manuscript/PNAS_manuscript.tex

@@
 
-\significancestatement{Acquiring high-fidelity observational records of ancient magnetic field intensity from the remanent magnetization of rocks is crucial for constraining the long-term evolution of Earth's core.
-However, robust estimates of ancient field strengths are often difficult to recover due to alteration or non-ideal rock magnetic behavior.
-In this study, we use plagioclase cumulates (anorthosite) that formed in the deep crust and were brought to the near surface by magma where they cooled and acquired thermal remanent magnetizations.
-These anorthosite xenoliths have experienced minimal alteration since their emplacement and yield high-quality paleointensity estimates.
-In contrast to scenarios of a progressively decaying field leading up to late nucleation of Earth's inner core, these data record a strong virtual dipole moment 1.1 billion years ago.
-A strong field persisted over at least a 14-million-year interval indicating the existence of appreciable power sources for Earth's dynamo during the late Mesoproterozoic.}
+\significancestatement{Acquiring high-fidelity ancient magnetic field intensity records from rocks is crucial for constraining the long-term evolution of Earth's core.
+However, robust estimates of ancient field strengths are often difficult to recover due to alteration or non-ideal behavior.
+We use anorthosites that formed in the deep crust and were brought to the near surface where they acquired thermal remanent magnetizations.
+These anorthosites have experienced minimal post-formation alteration and yield high-quality paleointensity estimates.
+In contrast to scenarios of a progressively decaying field leading up to late nucleation of Earth's inner core, these data record a strong field 1.1 Ga.
+A strong field that persisted over a 14-million-year interval indicates the existence of appreciable power sources for Earth's dynamo during the late Mesoproterozoic.}
 

### 2022-04-03 PNAS_manuscript/PNAS_Revision_Letter.tex

\documentclass[11pt, letterpaper]{article}
\usepackage{xcolor}
\usepackage{textcomp,marvosym}
\usepackage{amsmath,amssymb}
\usepackage[left]{lineno}
\usepackage{changepage}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{sidecap}
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\usepackage{titlesec}

\raggedright
\textwidth = 6.5 in
\textheight = 8.25 in
\oddsidemargin = 0.0 in
\evensidemargin = 0.0 in
\topmargin = 0.0 in
\headheight = .5 in
\headsep = 0.5 in
\parskip = 0.2 in
\parindent = 0.2in
\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}

\thispagestyle{myheadings}
\thispagestyle{fancy}
%\fancyhf{}
%\lhead{\Large \textbf {Summer School Scholarship Application} \noindent\small\textit{Yiming Zhang, UC Berkeley}}
\lhead{\noindent\small{Yiming Zhang \\ PhD candidate \\ 307 McCone Hall, Berkeley, CA 94720 \\ yimingzhang@berkeley.edu}}
%\rhead{\thepage}

%\author{Yiming Zhang}
%\date{University of California, Berkeley}
\begin{document}

%\maketitle
\begin{flushleft}

\hfill 2021
%{\noindent\textit {Yiming Zhang, UC Berkeley}}

Dear \textit{PNAS} editors,

We appreciate the positive assessment that our manuscript received through review. We are also grateful for the suggestions made by the reviewers that have guided revisions. The implementation of these revisions is detailed below and we have attached a track changes PDF with the submission of the revised manuscript. All reviewer's comments below are in italics and our responses are in regular fonts.

 \textbf{\Large Editor Remarks to Author:}

\textit{This manuscript has now been favorably reviewed by two experts in the field. Both suggested minor to moderate revisions which will improve the manuscript. I agree.}

Thank you! We appreciate the positive assessment of this work.

\textbf{\Large Reviewer 1 comments}

\itshape
This well-written manuscript presents new paleointensity estimates for the Mesoproterozoic which are unusually high, providing a compelling argument against a systematic decline in geomagnetic field strength throughout the Neoproterozoic. This is a significant result, since the previously claimed decline in field strength has been used to argue for young inner core nucleation $\sim$500 Ma. These new results convincingly show that there are several substantial changes in field strength throughout the time period 1500 - 500 Ma, and therefore paleointensity trends may not be a good approach for identifying the onset of core solidification.

The authors have clearly explained their approach, justified the interpretation of their results and the reliability of the chosen samples for paleomagnetic analysis. I recommend this study for publication and have included some suggestions and comments below which I hope the authors will find helpful.

\upshape

Thank you! We appreciate the positive assessment of this work. The comments have helped improve the manuscript. Below are our point-by-point responses.

\itshape
Major comments:

The results presented highlight two particularly interesting points: first, that the geodynamo did not lack power at ~1 Ga at odds with a dwindling thermal dynamo and second that tectonic processes may be responsible for the transitions in field strength around this time. Both of these points could be discussed in more detail, particularly alternative dynamo mechanisms e.g., Mg and Si precipitation and the likely efficiency of such mechanisms to generate the observed intensities observed here. Additionally, the role of tectonic processes in driving transitions in field strength is only mentioned at the end of the manuscript, it would be useful to describe the influence of tectonics on CMB heat flux in more detail earlier in the paper, and in particular any constrains from previous studies on how large these transitions in heat flux (and therefore field strength) are likely to be.
\upshape

These are good points. We have added a sentence to the introduction that briefly describes that there are proposals for additional power sources to sustain the geomagnetic field including exsolution of Mg and Si-bearing minerals at the core-mantle boundary. We think that the discussion of plate tectonic driven modulation to field strength is better positioned within the discussion than the introduction. We have added additional detail regarding such hypotheses.

%Studies have proposed various additional power sources that could have existed regardless of the inner core that could help drive an active geodynamo. \cite{Driscoll2014a} proposed that additional radioactive energy could have been supplied to the core by radioactive isotopes such as $^{40}$K; \cite{Mittal2020a} suggested that the additional chemical flux in the core produced by the precipitation of light elements could have helped maintain a dynamo, and studies such as \cite{Crowley2012a} explored theoretically that the core-mantle boundary modulation on the core heat flux which can allow for the maintenance of an active geodynamo without the existence of the inner core. We have added text in the introduction part of the manuscript to echo with our discussion.

\itshape
Figure 6 presents a nice compilation of paleointensity estimates from 1.5 Ga to the present day, including the previously proposed decline in field strength put forward by Bono et al., 2019 which is clearly at odds with the strong intensities which form the main result of this work. Based on all the data now available, it would be interesting to also include a simple statistical test to see if, given the sparsity of data and uncertainties, whether all results to-date can be distinguished from a transitional field oscillating around a constant value? I have never been convinced that the polynomial fit put forward by Bono et al., was statistically robust but it would be nice to quantitatively show it isn't with the new data. Additionally, would it be possible to estimate the timescale over which the transitions between high/low fields occur? Is this consistent with what we know about mantle dynamics and the rate at which subducted material should arrive at the CMB? These questions may be beyond the scope of this paper but may be worth further thought down the line.
\upshape

We agree with the overall take-away that our new data are clearly at odds with the polynomial fit of Bono et al. (2019). While additional statistical treatments of compilations are available (e.g. Biggin et al., 2015), in the present manuscript we think it best to focus on these new paleointensity estimates and their relative strength to highlight this main result. With these questions being well-positioned for further thought down the line as the reviewer writes.

We have added additional details on the timescale of plate tectonic drive fluctuations in heat flows with connections to mantle dynamics timescales.

\itshape
Minor comments:
\upshape

\textit{Line 6 - compositional buoyancy due to the ejection of light elements from the growing solid inner core}

The text ``due to the exclusion of light elements'' is added to be more specific about the mechanism behind compositional bouynacy.

\textit{Line 8 - also cite (Biggin et al., 2011) for evidence of a field 3.5 Ga}

Citation added.

\textit{Line 8-18 - may also be worth including the following paper which has just come out (Pozzo et al., 2022). Aside from the age of the inner core, the paleomagnetic record may also allow us to determine the mechanism that drove the dynamo prior to inner core nucleation. It would be good to add a couple of sentences discussing the paradox that if the inner core is young and the thermal conductivity is high thermal convection will be inhibited and unable to drive a dynamo prior to nucleation. In this case other mechanisms need to be considered e.g., Mg or Si precipitation (Badro et al., 2016; Hirose et al., 2016; O'Rourke \& Stevenson, 2016)}

Pozzo et al. 2022 is now cited.

The sentence `` The possibility of late inner core nucleation has motivated proposals of novel power sources to sustain the geomagnetic field through Earth history including precipitation of light element bearing minerals at the core-mantle boundary including MgO \cite{Badro2016a, ORourke2016b} and SiO$_2$ \cite{Mittal2020a}.'' has been added which refers to the reviewers point here.

\textit{Figure 1 - in this context I think providing a simplified geological map is fine, however it would be beneficial to include slightly more information in the legend. E.g., rather than just referring to them as intrusive igneous rocks/extrusive volcanic rocks (also being pedantic, both are igneous!), can you list the range of lithologies denoted by each colour? Similarly, for Pesonen and Halls include the lithologies. Also colour bar with ages is currently over the map, it should be moved to the side.}

In order to provide more details on lithology, we have added descriptive text to the legend associated with each study. At this scale of a geologic map, making further division of the polygons beyond the extrusive vs intrusive categorization would not be legible.

We have moved the color bar so that it does not overlap with any map units.

\textit{Line 57-59 - why does the lack of extension allow these rocks to be preserved?}

Were the Midcontinent Rift extension to have continued, it is very likely that it would have separated Laurentia into two, exposing the rift rocks at a continental margin setting, where subsequent collisions and associated orogenesis would have metamorphosed or completely eroded away the rocks through exhumation. However, the extension failed, preserving such rocks in the interior of Laurentia, where subsequent sediments covered them and resulted in great preservation until the recent exposure due to glaciation events.

We have added ``far from the continental margin and subsequent orogenesis.''

\textit{Line 97 - in what way is this rock type unique?}

The paragraph following line 97 explains the uniqueness in terms of its distinct petrologic and magnetic properties from other anorthosite cumulus rocks. To avoid overstating these properties we changed this sentence to ``In this study, we target the high-purity anorthosite xenoliths of the Beaver River diabase in the Midcontinent Rift."

\textit{Line 99-104 - the wording here could be clearer. Because plagioclase isn't Fe-bearing it cannot alter to form Fe-oxides, whereas minerals like olivine and pyroxenes are Fe-bearing so more likely to alter in a way that influenced the magnetic mineralogy.}

In fact, plagioclase often bears small amount of Fe but weathering under surface conditions are unlikely to produce iron oxides that affect the magnetic minerals significantly. We have split the sentence into two: ``They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud magnetic inclusions from alteration. In addition, the weathering of the plagioclase crystals do not resulting in the formation of secondary iron oxides, whereas other minerals such as olivine and pyroxene are more likely to alter and influence the magnetic mineralogy."

\textit{Line 107 - regarding the xenolith cooling, can you clarify whether they were partially molten or completely solid when they were brought to the near-surface? And specify the temperatures they cooled from, the phases acquiring magnetization and the Curie/blocking temperatures of these phases.}

These are good points and are in detail discussed in \cite{Zhang2021b} which is cited in the sentence. These anorthosites were likely heated to tholeiitic magma temperature and partially recrystallized. Thus they cooled above the Curie temperature of Fe-Ti oxides which are shown to be the dominant remanence carriers by \cite{Zhang2021a} as well as the rock magnetic data in this study.

We added ``from tholeiitic magma temperatures" to this sentence and added ``The details regarding the thermal history of these anorthosite xenoliths are discussed in \cite{Zhang2021b}" to help refer readers to our previous work.

\textit{Line 109 - missing space and only need degree symbol once 22 {plus minus} 2{degree sign}}

Text changed to ``22 {plus minus} 2{degree sign}."

\textit{Line 119-122 - this seems a little out-of-place here to conclude the Mesoproterozoic geodynamo must have been strong before presenting paleointensity results.}

\textit{Line 123-145 - this is a little repetitive with the introduction and I think could be combined with the previous section. When discussing geological setting and rock magnetism, respectively.}

\textit{Line 188 - 189 - can you clarify whether remanence anisotropy has also been measured by applying ARMs and demagnetizing them on a 2G? If possible, it would be good to quantify the degree of anisotropy (or lack thereof) using this method for comparison with other anisotropy studies.}

No we did not measure remanence anisotropy by the AARM method. Applying such study would require that we use sister specimens to those used for paleointensity. We prefer to use the gamma statistics to demonstrate the low degree of TRM anisotropy of our specimens.

We have changed the sentence to ``The anorthosite xenoliths show low anisotropy of thermal remanent magnetization (TRM) acquisition and can acquire TRM linearly within the range of relevant field strengths."

\textit{Figure 2 - the QDM maps for the Duluth anorthosite look significantly stronger than for Beaver River because of the choice of Bz scale range. Is it possible to make them the same so that it's a more direct comparison between the two?}

We decide to keep the Bz scale of the QDM maps as is. Bringing them to the same scale will either diminish the dipoles in the map of the Beaver River anorthosite xenolith or saturate the color scales in the Duluth anorthosite map. For clarification we noted in the figure caption that the two maps have Bz scales an order of magnitude different.

\textit{Figure 3 - many of the pTRM checks appear to have failed. Can you include some discussion of the degree of similarity between the first and second in-field step required for each pTRM check to pass? Similarly, how much sagging is allowed before a specimen fails? AX16-9a for example, looks like it shows a small amount of sagging - how much more would be tolerated before this specimen would fail? Also include in the figure caption a description of what the dashed blue/yellow lines represent.}

We choose to use the SCAT parameter of \cite{Tauxe2004a} which uses a threshold value to indicate whether the data points (including the pTRM checks and tail checks) in the chosen segment of the Arai plot are too scattered. The combination of the SCAT parameter and the FRAC parameter (which is the fraction of remanence used for determining paleointensity estimate) determines the amount of sagging allowed.

We have added ``Dashed red and blue lines show bounding regions calculated for determine the \textit{SCAT} statistic \cite{Shaar2013a}." to the figure caption.

\textit{Figure 4A - are all uncertainties 1 standard deviation?}

Yes. We edited text of ``(black bars with grey one standard deviation uncertainty boxes)" in the caption.

\textit{Figure 4C - what is the uncertainty on the 3 site mean? This should be included on the figure.}

The 1 standard deviation uncertainty is added to Figure 4C as well as Figure S5.

\textit{Line 311-316 - again can you clarify whether the anorthosite xenoliths are partially/fully molten at this temperature? Is there any mixing/mingling with the surrounding host rock that could influence the composition of the magnetic carriers?}

Again thermal modeling suggests that the anorthosite and diabase cooled together from tholeiitic temperature. For this clarification reason, we would like to keep the short section of ``Background on anorthosite xenoliths of the Beaver River diabase" which summarizes previous geochronology, paleomagnetic direction, and thermal modeling results in \cite{Zhang2021b}.

\textit{Line 327 - how do you know 500 (degree sign)C is the characteristic blocking temperature for your samples? Or are you just assuming that by the time they have cooled to this temperature all remanence will have been acquired?}

This is a good point. We did not make it clear in the text. 500\textdegree C is an estimated lab unblocking temperature for the anorthosites. Considering the cooling model of \cite{Zhang2021a} and the effect of lowering unblocking temperature due to slow cooling \cite{Halgedahl1980a}, a conservative estimate for the cooling time for the anorthosite from the Curie temperature of magnetite to the natural unblocking temperature should be about 10 kyr, leading to a cooling rate of 2.5\times10^{-10}. This is the value we used for calculating the paleointensity estimate value of 36\%. This was not made clear in the original manuscript.

\textit{Line 329 - what temperature range was this cooling estimate calculated for? For cooling from 580 to 500 over 1000 years I get a cooling rate of 2.5e-9 C\/s, although I suspect this doesn't have any influence on your cooling rate correction.}

Thanks for doing this calculation. We have changed the text to ``From the thermal history model of ref.  \citealp{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they blocked in the majority of their characteristic natural remanence magnetization (lab unblocking temperatures $\sim$500\textdegree C;  \citealp{Zhang2021b}). We find the cooling time to be on the order of $\sim$10 kyr considering the effect of slow cooling on lowering the natural remanence unblocking temperatures \cite{Halgedahl1980a}. This range of cooling time corresponds to a cooling rate of $\sim2.5\times10^{-10}$ $^\circ$C s$^{-1}$. In contrast, the lab cooling rate is much faster through the same temperature interval with an estimated cooling rate of $\sim1.3\times10^{-1}$ $^\circ$C s$^{-1}$. The significant cooling rate difference leads to a predicted $\sim$36\% overestimate of true ancient field following the model of \citealp{Halgedahl1980a} (Fig. S3)."

\textit{Lines 340-352 - are the anorthosites from similar regions in agreement with statistical significance - e.g., can you show they pass a student-T test? Similarly, are those from different regions in agreement\/ different with any statistical significance? It would also be help to state the recovered paleointensities\/ uncertainties in the text here.}

\textit{Line 425 - 'is key' should be 'are key'}

`is key' has been changed to `are key'.

\textit{Line 434-443 - in addition it's unclear whether a core with a high thermal conductivity could drive a thermally convecting dynamo at all. If something like Mg precipitation or Si precipitation at the CMB is at play instead then might not expect to see the decrease in power anyway (in fact this may make a lot more sense in light of your results)}

\textit{Lines 465-484 - this section is quite hard to follow, is there a simpler way to describe these transitions? Perhaps they could be labelled on Figure 6?}

\textit{Lines 526-538 - the implications for a role in tectonic processes in modulating transitions in intensity\/ reversal frequency of the geodynamo is important and feels a little understated given this is the first time it is really mentioned. Can these ideas be introduced sooner in the abstract/introduction to explain why you have recovered such high paleointensities?}

\textit{Line 605 - a MAD of $<$20 seems like quite a high cut-off given the apparent lack of anisotropy and stability of the anorthosite samples. Is there a reason why the data have so much scatter?}

Thanks for pointing this out. This MAD selection criteria needed to be updated. The data in fact have much smaller scatter, it is just the selection criteria we implemented were more generous than necessary. A MAD selection of $<$10 would result in the same selection as the highest specimen MAD is around 8.

\textit{Lines 619-620 - DANG on its own doesn't tell you if components are origin trending, did you use MAD $>$ DANG to identify origin-trending components?}

According to \cite{Tauxe2004a} DANG tests whether the component selected is actually trend- ing toward the origin as it measures the angle between the best-fit direction that is free floating and the direction between the centre of mass of the data and the origin of the vector component diagram. We did not use MAD $>$ DANG as an additional selection criteria and prefer to use MAD $<$ 5 and DANG $<$ 10.

\clearpage

\textbf{\Large Reviewer 2 comments}

\itshape
This well-written manuscript describes new geomagnetic paleointensity data from the Proterozoic. These data add to the very few existing paleointensity data prior to the mid-Paleozoic that have been used by other authors to infer the onset of inner core nucleation. In particular, recent paleointensity data from the Ediacaran was used to argue for a late inner core. The data presented here are inconsistent with that interpretation, and the authors argue (rightly, I believe) that the currently-available data make it challenging to use paleointensity alone to determine inner core formation. The issue of Earth evolution and inner core nucleation should be of wide interest and is appropriate for publication in PNAS.

The study appears to be well executed and the arguments in the manuscript are mostly well formed. I recommend publication, but have a few mostly minor comments detailed below.
\upshape

Thank you! We appreciate the positive assessment of this work. The comments have helped improve the manuscript. Below are our point-by-point responses.

\textit{General comments:}

\textit{I'm curious why you think the Beaver River magnetite inclusions are apparently not crystallographically controlled within the plagioclase - in contrast to what is observed in other intrusions. Do you think the magnetite was formed via a different process or at a different temperature? Related, I think at least a brief comment/discussion on the likely temperature of magnetite formation is warranted. (i.e., formed above/below the Curie temperature)}

These are good questions that we also wonder about and would like to explore further. \cite{Ageeva2016a, Ageeva2017a, Ageeva2020a, Bian2021, Ageeva2022a} provides a a series of detailed petrologic and crystallographic investigation into the origin of the oxide needles that exsolve within plagioclase. \cite{Bian2021a} concludes that these oxides form above 600\textdegree C, thus above the Curie temperature of magnetite.

It has been found that partitioning of Fe into plagioclase is enhanced by high anorthite content of the plagioclase, by high oxygen fugacity, and by high silica content of the melt \cite{Phinney1992a, Longhi1976a}. Our Beaver River anorthosite xenoliths do have higher anorthite content ($\sim$70 An; bytownite) as compared to the plagioclase of the Duluth Complex ($\sim$60 An, labradorite). We suspect that the anorthosite xenoliths form during the early stage of the magmatic evolution of the Beaver River diabase when anorthite content, and likely the oxygen fugacity and silica content in the magma was high, leading to more partition of Fe into plagioclase crystal lattice, leading to minor amount of Fe to exsolve and form oxide needles (although those that did could still be crystallographically controlled), instead of forming abundant, large Fe-oxide exsolutions like those in the Duluth anorthosite.

While discussion regarding the details about the petrologic origin of the Fe-Ti oxides is beyond the scope of this manuscript, we added a reference to \cite{Bian2021a} in our text of section ``Petrography and magnetic imaging of anorthosite xenoliths" to help clarify the formation of these oxides are above the Curie temperature of magnetite.

\textit{Line 14-15: ``these values" is a little vague. Are you referring just to conductivity, or something else too?}

The text has been changed to ``While some estimated thermal conductivity values are consistent with an inner core age $>$3 Ga..."

\textit{Lines 101-102: For clarity, would suggest changing to ``...from alteration. Alteration of the plagioclase crystals themselves does not result in ..."}

This is a good suggestion. This sentence has been split into two: `` They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud magnetic inclusions from alteration. In addition, the weathering of the plagioclase crystals do not resulting in the formation of secondary iron oxides, whereas other minerals such as olivine and pyroxene are more likely to alter and influence the magnetic mineralogy."

\textit{Line 143: ``unblock" should be ``unblocks"}

``Unblock" has been changed to ``unblocks".

\textit{Line 151-153. If you're going to say that these crystals ``often" contain oxide inclusions, you should probably cite more than one example.}

Additional references to \cite{Wenk2011a, Ageeva2016a} have been added.

\textit{Lines 176-181 and Fig. 2: For the PNAS audience, I don't think most people will be able to ``see" this magnetization rotation in Figure 2. Even for those who could recognize and interpret the dipole signature, the relevant panels in Figure 2 are a bit hard to make out. Can you enlarge a region with one or two dipoles so it is easier to see the change/deviation from the applied field direction?}

\textit{Figure 5: I think it's a bit deceptive to just include the smoothed fit and not show what are actually quite noisy data (based on what's in your Jupyter notebook). I think the interpretation is fine, but it would be nice to show the original data - either here or in the supplementary materials. (Also, the font size is quite small and difficult to read.)}

An additional supplementary figure S6 which plots the fitted coercivity curve and the smoothed data points has been added. The font sizes have been increased.

\textit{Lines 268-271: Also a slightly misleading statement. Passing samples ``can" have peak coercivity distributions around 80 mT, but based on your box plot (Fig5), only one sample has an MDF >80 mT, one has an MDF of about 70 mT, and the rest are <60 mT.}

The sentence has been changed to ``Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivities ranging from $\sim$44-144 mT with a median of 53 mT."

\textit{Line 328: cooling time on the order of ~1 kyr - That's not what is shown in Figure S3. There, you show 10 kyr.}

Thanks for catching this inconsistency. Considering the effect of lowering unblocking temperature due to slow cooling, the cooling timescale should be closer to 10 kyr. We have changed the text to ``We find the cooling time to be on the order of $\sim$10 kyr considering the effect of slow cooling on lowering the natural remanence unblocking temperatures \cite{Halgedahl1980a}."

\textit{Lines 331-333. Both methods of assessing cooling rate bias assume uniaxial SD, yet the zig-zag behavior shown in at least 2 out of 3 of your passing Arai plots (Fig. 3) suggest something larger than SD. How might that affect your cooling rate calculations?}

This is a good point. Experimental study by \cite{Biggin2013a} shows that PSD and MD grains may be less biased toward recording higher paleointensities than SD grains due to slow cooling. Therefore our cooling rate correction may be an over-correction given the evidence of some MD grains as pointed out by the reviewer. This does not affect our interpretations, as it means the true paleointensity values could be even higher than currently estimated based on the correction factor of 0.74.

We have added text ``This factor may result in an over-correction, as some multidomain grains that do not tend to severely bias paleointensity estimates toward higher \cite{Biggin2013a}, may exist in the anorthosite xenoliths as is evidenced through the slightly zigzagging Arai plots" to the related paragraph.

\textit{Lines 356-360: Do you think any of the intensity difference between your samples and the PADM2M model can be explained by the fact that the model only reflects the axial dipole contribution to the field? Also, I believe the necessarily smoothing in most of these models will likely underestimate field variability. I don't know if that will necessarily lead to a lower average, but it probably underrepresents the high field values.}

Figure S5 which compares our samples and resampled Cenozoic paleointensity records illustrates that our values are also near the high end.

\textit{Figure 6: Some of the pale colors are very hard to see.}

We have decreased transparency on the data points from the compilation so they are better illustrated.

\textit{Supplementary material: Assuming this material will not be reformatted, please consider enlarging the font size in the figure captions and within many of the figures. It is quite difficult to read without constantly zooming in and out.}

The fonts have all been enlarged.

\textit{Figure S1. What does the yellow dashed line represent? Please explain red vs. blue and ``A" vs ``B" sites.}

The caption has been updated as the following: `Violin plots of site-level posterior paleointensity distributions estimated using the bias corrected estimation of paleointensity (BiCEP) method developed by \citealp{Cych2021a}. Blue bars of each violin plot represents the median of the distributions. The yellow dashed line is the mean paleointensity of the posterior estimates across all sites. Assuming that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Arai1963a, Paterson2011a}, the method uses all paleointensity measurement-level data without applying selection criteria. For comparison of results from this independent method with those based on our selection (as show in Fig. 4 in manuscript), we highlight the anorthosite sites that pass our paleointensity selection and make other anorthosite and diabase transparent. The results from the BiCEP method address the uncertainties associated with anorthosite AX6 and AX8. This is associated with the relatively variable specimen behaviors within these two sites. But for sites AX10, AX11, AX13, AX12, and AX16, the posterior probability distributions have very narrow bounds, consistent with the interpretation that these anorthosites are faithful paleointensity recorders that have high-quality paleointensity behaviors. AX-anorthosite xenolith site; BD-Beaver River diabase site. The grouping of cooling unit is based on the spacial proximity between site locations and the inclusion relationship of host diabase and anorthosite xenoliths reported by \citealp{Zhang2021b}."

\textit{Figure S2. Can you indicate which of these specimens come from ``passing" samples and which are from ``failing" samples? Also, I'm pretty sure the ZFC>FC being an indication of MD material has been demonstrated to be unreliable. (Sorry I can't find the relevant reference.)}

Labels for whether the sister specimen to the rock chips used for MPMS measurements have been added. Figure caption has been updated with the following sentence: ``Pass" or ``Fail" represents whether their sister specimens passed our paleointensity selection or not.

\textit{Data availability: The link to the MagIC database did not work, so I was unable to verify how much of the data were uploaded. It's unclear if it is just the paleointensity data or if it includes all the related rock magnetic data.}

The link has been updated: \url{https://earthref.org/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}.

Sincerely,

Yiming Zhang\\

PhD candidate of Earth and Planetary Science\\

University of California, Berkeley

\end{flushleft}
\end{document}

### 2022-04-03 PNAS_manuscript/PNAS_manuscript.tex

@@
 \affil[a]{Department of Earth and Planetary Science, University of California, Berkeley, CA, 94720}
-\affil[b]{U.S. Geological Survey, Moffett Field, CA, 94025}
+\affil[b]{U.S. Geological Survey, Moffett Field, CA, 94035}
 \affil[c]{Department of Earth and Planetary Sciences, Harvard University, Cambridge, MA, 02138}
@@
 \dropcap{E}arth's magnetic field is the result of convective flow of liquid iron-alloy in Earth's outer core.
-At present day, the geodynamo is collectively powered by heat flow across the core-mantle boundary (CMB) and from the crystallization of the solid inner core from the liquid outer core which provides latent heat and compositional buoyancy \cite{Buffett2000a}.
-However, while paleomagnetic studies have found that a dynamo field has existed since at least 3.4 billion years ago \cite{Selkin2007a, Tarduno2014a, Brenner2020a}, Earth's inner core likely crystallized more recently.
+At present day, the geodynamo is collectively powered by heat flow across the core-mantle boundary (CMB) and from the crystallization of the solid inner core from the liquid outer core which provides latent heat and compositional buoyancy due to the exclusion of light elements \cite{Buffett2000a}.
+However, while paleomagnetic studies have found that a dynamo field has existed since at least 3.4 billion years ago \cite{Selkin2007a, Biggin2011a, Tarduno2014a, Brenner2020a}, Earth's inner core likely crystallized more recently.
 Estimates of the timing of the initial crystallization of Earth's inner core are interconnected with estimates for the core's thermal conductivity.
 Higher conductivity values imply faster cooling rates, which in turn imply that the threshold for the freezing of the inner core happened more recently \cite{Davies2015a}.
-While some estimates of these values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates have implied higher thermal conductivity values and an age for the inner core that is less than 1.5 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b, Frost2022a}, with some suggesting even younger ages ($<$700 Ma;  \citealp{Labrosse2015a, Ohta2016a}).
+While some estimated thermal conductivity values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, others have implied higher thermal conductivity values and an age for the inner core that is less than 1.5 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b, Frost2022a}, with some suggesting even younger ages ($<$700 Ma;  \citealp{Labrosse2015a, Ohta2016a, Pozzo2022a}).
+The possibility of late inner core nucleation has motivated proposals of novel power sources to sustain the geomagnetic field through early Earth history including precipitation of light materials such as MgO \cite{Badro2016a, ORourke2016a, ORourke2016b} and SiO$_2$ \cite{Mittal2020a} at the core-mantle boundary.
 Given that estimates for the core's thermal conductivity continue to be debated, it is crucial to use observational records as an independent constraint on the thermal evolution of Earth's core and mantle.
@@
 \ref{fig:Geologic_map}) is an exception as it is a large igneous province where magmatism lasted $\sim$25 Myr from \texit{ca.} 1109 Ma to 1084 Ma during which there were pulsed intervals of more rapid activity \cite{Swanson-Hysell2021a}.
-Additionally, extension ceased in the Midcontinent Rift prior to lithospheric separation, preserving volcanic, intrusive, and sedimentary rocks of the rift within the continental interior.
+Additionally, extension ceased in the Midcontinent Rift prior to lithospheric separation, preserving volcanic, intrusive, and sedimentary rocks of the rift within the continental interior far from the continental margin and subsequent orogenesis.
 As a result, rocks of the rift have unusually simple paleomagnetic behavior for their greater than one billion-year-old age and paleomagnetic data from rift rocks forms a central record of Mesoproterozoic paleogeography \cite{Swanson-Hysell2021c}.
@@
 
-In this study, we target a unique rock type---high-purity anorthosite xenoliths.
+In this study, we target the high-purity anorthosite xenoliths of the Beaver River diabase in the Midcontinent Rift.
 Anorthosites are cumulate rocks composed dominantly of the mineral plagioclase.
-They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud magnetic inclusions from alteration with alteration of the plagioclase crystals themselves not resulting in the formation of secondary iron oxides in contrast with mafic minerals such as olivine.
+They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud magnetic inclusions from alteration.
+In addition, the weathering of the plagioclase crystals do not resulting in the formation of secondary iron oxides, whereas other minerals such as olivine and pyroxene are more likely to alter and influence the magnetic mineralogy.
 The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal (shallowly emplaced) intrusions of the Beaver River diabase \cite{Zhang2021b}.
-They cooled and acquired their magnetizations in conjunction with the host diabase at a paleolatitude of 22\textdegree $\pm$2\textdegree (calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a, Zhang2021b}).
+They likely cooled from tholeiitic magma temperatures and acquired their magnetizations in conjunction with the host diabase at a paleolatitude of 22 $\pm$ 2\textdegree (calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a}).
+The details regarding the thermal history of these anorthosite xenoliths are discussed in \cite{Zhang2021b}.
 Paleointensity experiments on the anorthosite xenoliths have a high success rate, yielding consistent specimen- and site-level paleointensity results.
-Rock magnetic data reveal that the anorthosite xenoliths have low anisotropy of thermal remanent magnetization (TRM) and can acquire TRM linearly within the range of relevant field strengths.
+The anorthosite xenoliths show low anisotropy of thermal remanent magnetization (TRM) acquisition and can acquire TRM linearly within the range of relevant field strengths.
 Magnetic imaging shows that the anorthosite specimens have dominant magnetic carriers within and interstitial to plagioclase crystals without strong preferred orientations.
@@
 Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths to have acquired thermal remanent magnetizations while cooling with the Beaver River diabase \cite{Zhang2021b}.
-Step-wise thermal demagnetization data show the anorthosite xenoliths to have dominantly single-component magnetization that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C, consistent with magnetization held by low-titanium titanomagnetite \cite{Zhang2021b}.
+Step-wise thermal demagnetization data show the anorthosite xenoliths to have dominantly single-component magnetization that often unblocks sharply within temperature ranges between 500\textdegree C and 580\textdegree C, consistent with magnetization held by low-titanium titanomagnetite \cite{Zhang2021b}.
 
@@
 \ref{fig:Petro_QDM}A).
-Plagioclase crystals within coarse-grained intrusions often contain abundant elongate Fe-Ti oxide inclusions that are visible through optical microscopy \cite{Feinberg2006a}.
+Plagioclase crystals within coarse-grained intrusions often contain abundant elongate Fe-Ti oxide inclusions that are visible through optical microscopy \cite{Feinberg2006a, Wenk2011a, Ageeva2016a}.
 The Beaver River anorthosite xenoliths lack such large (10s of $\mu$m) oxide inclusions such that oxides are not visible within the plagioclase crystals through optical microscopy (Fig.
@@
 \ref{fig:Petro_QDM}E, G).
+The sources of these magnetic remanence carriers are likely to be Fe-Ti oxides that exsolved within plagioclase crystals above the Curie temperature of magnetite \cite{Zhang2021b, Bian2021a}.
 To highlight these distinct aspects of the Beaver River anorthosite xenoliths, we present petrographic and magnetic imaging data from the Beaver River anorthosite as well as a sample of an anorthosite from the Duluth Complex Anorthositic Series---an older intrusive complex within the Midcontinent Rift that was not targeted for paleointensity experiments in this study (Fig.
@@
 1092 Ma.
-All measurement-level paleointensity experiment data are available within the MagIC database (\url{https://earthorg/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}; \textit{this private link is provided for the purpose of review; the url will be updated when a doi is generated for this manuscript}).
+All measurement-level paleointensity experiment data are available within the MagIC database (\url{https://earthref.org/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}; \textit{this private link is provided for the purpose of review; the url will be updated when a doi is generated for this manuscript}).
 
@@
 Blue and red squares in the Zijderveld plots are X–Y and X–Z projections, respectively, of the NRMs in specimen coordinates.
+Dashed red and blue lines show bounding regions calculated for determine the \textit{SCAT} statistic \cite{Shaar2013a}.
 Plots on the top row show successful specimen paleointensity results with straight, single-slope behaviors that pass our selection criteria.
@@
 \centering
-\caption{\footnotesize{A) Summary plot of individual specimen absolute paleointensity results (square symbols) and their averages and standard deviations at site level (black bars with grey 1$\sigma$ uncertainty boxes) where each `AX' site is an individual anorthosite xenolith within the Beaver River diabase.
+\caption{\footnotesize{A) Summary plot of individual specimen absolute paleointensity results (square symbols) and their averages and standard deviations at site level (black bars with grey one standard deviation uncertainty boxes) where each `AX' site is an individual anorthosite xenolith within the Beaver River diabase.
 All results are corrected for cooling rate with a factor of 0.74.
@@
 The distribution represents a total of 10,000 iterations of taking 3 random draws and calculating the mean.
-These comparisons highlight that these paleointensity values are strong in relation to the geomagnetic field over the past 2 million years.}}
+These comparisons highlight that these paleointensity values are strong in relation to the geomagnetic field over the past 2 million years.
+All shaded regions in B) and C) represent one standard deviation uncertainties.}}
 \label{fig:PINT_cooling_corrected}
@@
 \ref{fig:coercivity}).
-Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivity around 80 mT (Fig.
+Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivities ranging from $\sim$44-144 mT with a median of 53 mT (Fig.
 \ref{fig:coercivity}).
@@
 From the thermal history model of ref.
-\citealp{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they fully blocked in their characteristic natural remanence magnetization ($\sim$500\textdegree C;  \citealp{Zhang2021b}).
-We find the cooling time to be on the order of $\sim$1 kyr, which corresponds to a cooling rate of $\sim1.6\times10^{-9}$ $^\circ$C s$^{-1}$.
+\citealp{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they blocked in the majority of their characteristic natural remanence magnetization (lab unblocking temperatures $\sim$500\textdegree C;  \citealp{Zhang2021b}).
+We find the cooling time to be on the order of $\sim$10 kyr considering the effect of slow cooling on lowering the natural remanence unblocking temperatures \cite{Halgedahl1980a}.
+This range of cooling time corresponds to a cooling rate of $\sim2.5\times10^{-10}$ $^\circ$C s$^{-1}$.
 In contrast, the lab cooling rate is much faster through the same temperature interval with an estimated cooling rate of $\sim1.3\times10^{-1}$ $^\circ$C s$^{-1}$.
@@
 We therefore correct our paleointensity results by a factor of 0.74.
+This factor may result in an over-correction, as some multidomain grains that do not tend to severely bias paleointensity estimates toward higher \cite{Biggin2013a}, may exist in the anorthosite xenoliths as is evidenced through the slightly zigzagging Arai plots (Figure \ref{fig:IZZI_examples}).
 The cooling rate-corrected specimen paleointensity estimates together with specimen- and site-level means are shown in Fig.
@@
 Such a scenario where the dynamo shifts to a greater depth associated with inner core nucleation led \cite{Aubert2009a} to conclude that the increase in power to the dynamo would be difficult to detect with paleointensity data.
-Ultimately, further observational paleomagnetic records is key as they hold the potential for testing different model predictions and identifying transitions in ancient field strength \cite{Biggin2015a, Bono2019a}.
+Ultimately, further observational paleomagnetic records are key as they hold the potential for testing different model predictions and identifying transitions in ancient field strength \cite{Biggin2015a, Bono2019a}.
 
@@
 \subsection*{Paleointensity result selection}
-The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD;  \citealp{Kirschvink1980a}) of $<$20\textdegree; (2) scatter parameter ($\beta$;  \citealp{Coe1978a}) values of $<$15$\%$; (3) a deviation angle (DANG;  \citealp{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC;  \cite{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT;  \citealp{Shaar2013a}) = TRUE; (6) a maximum magnetic moment difference between adjacent zero-field steps (GAP-Max;  \citealp{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
+The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD;  \citealp{Kirschvink1980a}) of $<$10\textdegree; (2) scatter parameter ($\beta$;  \citealp{Coe1978a}) values of $<$15$\%$; (3) a deviation angle (DANG;  \citealp{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC;  \cite{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT;  \citealp{Shaar2013a}) = TRUE; (6) a maximum magnetic moment difference between adjacent zero-field steps (GAP-Max;  \citealp{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
 The MAD measures the scatter about the best-fit line through the natural remanent magnetization (NRM) steps in the selected interval for which the intensity is defined.
@@
 Conversations with Bruce Buffett and Zachary Geballe informed our perspectives on the geodynamic and mineral physics literature.
-Paleomagnetic data associated with this study are available within the MagIC database (\url{https://earthorg/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}; \textit{provide link when doi is generated at time of proofs}) and all data are within a Github repository along with Python code related to calculations and visualizations (\url{https://github.com/Swanson-Hysell-Group/AX_BD_PINT}) that are also archived on Zenodo (\textit{provide link when doi is generated at time of proofs}).}
+Paleomagnetic data associated with this study are available within the MagIC database (\url{https://earthref.org/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}; \textit{provide link when doi is generated at time of proofs}) and all data are within a Github repository along with Python code related to calculations and visualizations (\url{https://github.com/Swanson-Hysell-Group/AX_BD_PINT}) that are also archived on Zenodo (\textit{provide link when doi is generated at time of proofs}).}
 

### 2022-04-03 PNAS_manuscript/Supporting Info.tex

@@
 
-% \begin{figure*}[h!]
-% \noindent\includegraphics[width=0.8\textwidth]{Petro_QDM_SI.png}
-% \centering
-% \caption{{Petrographic images and magnetic images from the main manuscript with additional anisotropy experiment results.
-In A-F, we show that the Beaver River anorthosite often lack large oxide needles within plagioclase crystals carrying strong magnetic remanence  differences of magnetic carriers which is in contrast to what is seen in Duluth Complex anorthosites.
-E-H show an experiment performed on both anorthosite xenoliths where we apply a first field along the y axis of the field of view and then apply a second field of 300 mT orthogonal to the first field direction.
-The isothermal remanent magnetizations acquired by both anorthosites were imaged with a quantum diamond microscope after the application of each field.
-The magnetic images show that dipole-like remanent magnetizations of Beaver River anorthosite align well with the first applied field direction and rotated to align with the second applied field direction with little anisotropic behavior.
-In contrast, the remanence magnetizations carried by the magnetic needles in plagioclase 1 of the Duluth anorthosite xenolith align well with the first applied field but those in plagioclase 2 tend to acquire an oblique remanence direction with respect to the field direction (F).
-After the application of a 300 mT external field, magnetization of those needles in plagioclase 1 did not reverse due to strong shape anisotropy whereas those needles in plagioclase 2 reversed but still acquired remanence oblique to the field direction.
-This experiments illustrates that the Beaver River anorthosite xenolith has much lower degrees of magnetic anisotropy than the Duluth anorthosite xenolith.
-B$_z$ denotes magnetic field intensity in the vertical direction with positive values corresponding to the out of page direction.
-}}
-% \label{fig:Petro_QDM_SI}
-% \end{figure*}
-% \clearpage
-
 \begin{figure*}[h!]
@@
 \caption{{Violin plots of site-level posterior paleointensity distributions estimated using the bias corrected estimation of paleointensity (BiCEP) method developed by \citealp{Cych2021a}.
+Blue bars of each violin plot represents the median of the distributions.
+The yellow dashed line is the mean paleointensity of the posterior estimates across all sites.
 Assuming that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Arai1963a, Paterson2011a}, the method uses all paleointensity measurement-level data without applying selection criteria.
@@
 But for sites AX10, AX11, AX13, AX12, and AX16, the posterior probability distributions have very narrow bounds, consistent with the interpretation that these anorthosites are faithful paleointensity recorders that have high-quality paleointensity behaviors.
-}}
+AX-anorthosite xenolith site; BD-Beaver River diabase site.
+The grouping of cooling unit is based on the spacial proximity between site locations and the inclusion relationship of host diabase and anorthosite xenoliths reported by \citealp{Zhang2021b}.}}
 \label{fig:PINT_BiCEP}
@@
 \caption{{Low-temperature magnetic property measurement system (MPMS) experiment results.
+``Pass" or ``Fail" represents whether their sister specimens passed our paleointensity selection or not.
 In the field-cooled (FC) experiments, the magnetization was measured upon warming following the specimen having cooled in an applied field of 2.5 T from 300 to 10 K.
@@
 Specimen BD2-2r has a consistently higher moment during the zero-field-cooled step than during the field-cooled step.
-This is consistent with the interpretation that multidomian magnetic carriers exist in significant quantity in this specimen.}}
+This is consistent with the interpretation that multidomian magnetic carriers exist in significant quantity in this specimen \cite{Carter-Stiglitz2006a}.}}
 \label{fig:MPMS}

### 2022-04-03 PNAS_manuscript/Supporting Info.tex
Updates from Overleaf

@@
 The distribution represents a total of 10,000 iterations of taking 3 random draws and calculating the mean.}}
+\label{fig:Cenozoic_PINT}
+\end{figure*}
+
+\clearpage
+
+\begin{figure*}[h!]
+\noindent\includegraphics[width=0.9\textwidth]{example_unmix_plot_with_data.pdf}
+\centering
+\caption{\footnotesize{A) Example coercivity spectra of anorthosite and diabase specimens from sites that pass or fail our paleointensity selection criteria.
+The specimens and fits are the same as those shown in Figure 5 in the manuscript.
+Data points used for fitting the coercivity curves are shown in this figure.
+}}
 \label{fig:Cenozoic_PINT}

### 2022-04-04 PNAS_manuscript/Supporting Info.tex
Updates from Overleaf

@@
 Overall region-mean values are calculated using all regional mean results in this table.
-}
+Independent two-sample t-test results show that site AX12 and AX16 pass the t-test at 1\% significance level; site AX11 and AX13 pass the test at 5\% significance level; region Kennedy Creek have distinguishable distributions from the other two regions; Silver Bay region and Carlton Peak region are indistinguishable at 5\% significance level.}
 \begin{tabular}{ccccccccccc}
@@
 \hline
-\multirow{7}{*}{specimen mean for sites}   & AX6           & 30.33     & 3  & 0.69 & 0.13    & 3.55           & 2.66            & 38.9       & 2.63     & 48.76         \\
-                                           & AX8           & 52.4      & 3  & 0.66 & 0.17    & 4.28           & 2.15            & 40.3       & 7.3      & 83.14         \\
-                                           & AX10          & 46.82     & 3  & 0.7  & 0.23    & 4.98           & 2.14            & 36.6       & 6.1      & 76.81         \\
-                                           & AX12          & 38.1      & 6  & 0.75 & 0.22    & 3.26           & 2.53            & 55.2       & 5.98     & 51.31         \\
-                                           & AX16          & 44.15     & 11 & 0.68 & 0.19    & 4.19           & 1.96            & 49.2       & 3.97     & 63.87         \\
-                                           & AX11          & 73.34     & 7  & 0.68 & 0.19    & 3.85           & 1.82            & 35.2       & 6.63     & 121.73        \\
-                                           & AX13          & 76.63     & 7  & 0.73 & 0.2     & 5.7            & 1.67            & 35.1       & 5        & 127.29        \\ 
+\multirow{7}{*}{specimen mean for sites}   & AX6           & 30.3     & 3  & 0.69 & 0.13    & 3.6           & 2.7            & 38.9       & 2.6     & 48.8         \\
+                                           & AX8           & 52.4      & 3  & 0.66 & 0.17    & 4.3           & 2.2            & 40.3       & 7.3      & 83.1         \\
+                                           & AX10          & 46.8     & 3  & 0.70  & 0.23    & 5.0           & 2.1            & 36.6       & 6.1      & 76.8         \\
+                                           & AX12          & 38.1      & 6  & 0.75 & 0.22    & 3.3           & 2.5            & 55.2       & 6.0     & 51.3         \\
+                                           & AX16          & 44.2     & 11 & 0.68 & 0.19    & 4.2           & 2.0            & 49.2       & 4.0     & 63.9         \\
+                                           & AX11          & 73.3     & 7  & 0.68 & 0.19    & 3.9           & 1.8            & 35.2       & 6.6     & 121.7        \\
+                                           & AX13          & 76.6     & 7  & 0.73 & 0.20     & 5.7            & 1.7            & 35.1       & 5        & 127.3        \\ 
                                            \hline
-\multirow{3}{*}{specimen mean for regions} & Carlton Peak  & 43.18     & 9  & 0.68 & 0.17    & 4.27           & 2.32            & 38.6       & 5.34     & 69.57         \\
-                                           & Kennedy Creek & 74.98     & 14 & 0.7  & 0.2     & 4.78           & 1.74            & 35.15      & 5.81     & 124.51        \\
-                                           & Silver Bay    & 42.02     & 17 & 0.71 & 0.2     & 3.86           & 2.16            & 51.32      & 4.68     & 59.43         \\
+\multirow{3}{*}{specimen mean for regions} & Carlton Peak  & 43.2     & 9  & 0.68 & 0.17    & 4.3          & 2.3            & 38.6       & 5.3     & 69.6         \\
+                                           & Kennedy Creek & 75.0     & 14 & 0.70  & 0.20     & 4.8           & 1.7            & 35.2      & 5.8     & 124.5        \\
+                                           & Silver Bay    & 42.0     & 17 & 0.71 & 0.20     & 3.9           & 2.2            & 51.3      & 4.7     & 59.4         \\
 \hline
-overall site mean                          &               & 51.68     & 7  &      &         &                &                 &            &          & 81.84         \\
-overall region mean                        &               & 53.39     & 3  &      &         &                &                 &            &          & 84.5         
+overall site mean                          &               & 51.7     & 7  &      &         &                &                 &            &          & 81.8         \\
+overall region mean                        &               & 53.4     & 3  &      &         &                &                 &            &          & 84.5         
 \end{tabular}

### 2022-04-04 PNAS_manuscript/PNAS_Revision_Letter.tex

@@
 \usepackage{titlesec}
+\usepackage{url}
 
@@
 We think that the discussion of plate tectonic driven modulation to field strength is better positioned within the discussion than the introduction.
-We have added additional detail regarding such hypotheses.
 
@@
 
-We agree with the overall take-away that our new data are clearly at odds with the polynomial fit of Bono et al.
-(2019).
-While additional statistical treatments of compilations are available (e.g.
-Biggin et al., 2015), in the present manuscript we think it best to focus on these new paleointensity estimates and their relative strength to highlight this main result.
-With these questions being well-positioned for further thought down the line as the reviewer writes.
-
-We have added additional details on the timescale of plate tectonic drive fluctuations in heat flows with connections to mantle dynamics timescales.
+We agree with the overall take-away that our new data are clearly at odds with the polynomial fit of \cite{Bono2019a}.
+While additional statistical treatments of compilations are available \cite[e.g.][]{Biggin2015a}, in the present manuscript we think it best to focus on these new paleointensity estimates and their relative strength to highlight this main result.
+These questions are well-positioned for further analysis down the line as the reviewer writes.
 
@@
 
-The text ``due to the exclusion of light elements'' is added to be more specific about the mechanism behind compositional bouynacy.
+The text ``due to the exclusion of light elements'' is added to be more specific about the mechanism behind compositional buoyancy.
 
@@
 
-We have added ``far from the continental margin and subsequent orogenesis.''
+We have added ``far from the continental margin and subsequent orogenesis."
 
@@
 
-The paragraph following line 97 explains the uniqueness in terms of its distinct petrologic and magnetic properties from other anorthosite cumulus rocks.
-To avoid overstating these properties we changed this sentence to ``In this study, we target the high-purity anorthosite xenoliths of the Beaver River diabase in the Midcontinent Rift."
+The paragraph following line 97 explains the distinct petrologic and magnetic properties of the anorthosite xenoliths.
+To avoid overstating we changed this sentence to ``In this study, we target the high-purity anorthosite xenoliths of the Beaver River diabase in the Midcontinent Rift."
 
@@
 
-In fact, plagioclase often bears small amount of Fe but weathering under surface conditions are unlikely to produce iron oxides that affect the magnetic minerals significantly.
 We have split the sentence into two: ``They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud magnetic inclusions from alteration.
@@
 
-We added ``from tholeiitic magma temperatures" to this sentence and added ``The details regarding the thermal history of these anorthosite xenoliths are discussed in \cite{Zhang2021b}" to help refer readers to our previous work.
+We have integrated this paragraph with the section of ``Background on anorthosite xenoliths of the Beaver River diabase" and the paragraph now reads ``In this study, we target the high-purity anorthosite xenoliths of the Beaver River diabase in the Midcontinent Rift.
+While magmatic activity within the Midcontinent Rift was protracted, there were intervals of particularly rapid volcanism and voluminous emplacement of intrusions \cite{Swanson-Hysell2021a}.
+The ca.
+1092 Ma Beaver Bay Complex in northern Minnesota punctuates one such period of magmatism during the main stage of Midcontinent Rift activity.
+The magma that formed the 1091.7$\pm$0.2 Ma Beaver River diabase dikes and sills of the Beaver Bay Complex transported numerous anorthosite xenoliths that have short-axis diameters up to 180 meters via wide conduits \cite{Boerboom2004a, Boerboom2006b}.
+These anorthosite xenoliths are plagioclase cumulates that formed comagmatically with the host diabase in the lower crust---an interpretation confirmed by U-Pb zircon geochronology \cite{Zhang2021b}.
+Anorthosites are cumulate rocks composed dominantly of the mineral plagioclase.
+They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud magnetic inclusions from alteration.
+In addition, the weathering of the plagioclase crystals do not resulting in the formation of secondary iron oxides, whereas other minerals such as olivine and pyroxene are more likely to alter and influence the magnetic mineralogy.
+The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal (shallowly emplaced) intrusions of the Beaver River diabase \cite{Zhang2021b}.
+They were likely heated to tholeiitic magma temperatures and partially recrystallized \cite{Zhang2021b}.
+Fe-Ti oxides within the plagioclase exsolved above magnetite Curie temperature ($\sim$580\textdegree C; \cite{Bian2021a}) and subsequently cooled and acquired thermal remanent magnetizations in conjunction with the host diabase at a paleolatitude of 22 $\pm$ 2\textdegree (calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a}).
+The details regarding the thermal history of these anorthosite xenoliths are discussed in \cite{Zhang2021b}.
+Paleointensity experiments on the anorthosite xenoliths have a high success rate, yielding consistent specimen- and site-level paleointensity results.
+The anorthosite xenoliths show low anisotropy of thermal remanent magnetization (TRM) acquisition and can acquire TRM linearly within the range of relevant field strengths.
+Magnetic imaging shows that the anorthosite specimens have dominant magnetic carriers within and interstitial to plagioclase crystals without strong preferred orientations.
+Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths to have acquired thermal remanent magnetizations while cooling with the Beaver River diabase \cite{Zhang2021b}.
+Step-wise thermal demagnetization data show the anorthosite xenoliths to have dominantly single-component magnetization that often unblocks sharply within temperature ranges between 500\textdegree C and 580\textdegree C, consistent with magnetization held by low-titanium titanomagnetite \cite{Zhang2021b}.
+"
 
@@
 
-Text changed to ``22 {plus minus} 2{degree sign}."
+Text changed to ``22 $\pm$ 2\textdegree."
 
@@
 
+We removed this sentence.
+
 \textit{Line 123-145 - this is a little repetitive with the introduction and I think could be combined with the previous section. When discussing geological setting and rock magnetism, respectively.}
+
+We have integrated this section into the introduction.
 
@@
 Bringing them to the same scale will either diminish the dipoles in the map of the Beaver River anorthosite xenolith or saturate the color scales in the Duluth anorthosite map.
-For clarification we noted in the figure caption that the two maps have Bz scales an order of magnitude different.
+For clarification, we noted in the figure caption that the two maps have Bz scales that are an order of magnitude different.
 
@@
 
-Again thermal modeling suggests that the anorthosite and diabase cooled together from tholeiitic temperature.
-For this clarification reason, we would like to keep the short section of ``Background on anorthosite xenoliths of the Beaver River diabase" which summarizes previous geochronology, paleomagnetic direction, and thermal modeling results in \cite{Zhang2021b}.
+We clarify this in the introduction.
 
@@
 We did not make it clear in the text.
-500\textdegree C is an estimated lab unblocking temperature for the anorthosites.
-Considering the cooling model of \cite{Zhang2021a} and the effect of lowering unblocking temperature due to slow cooling \cite{Halgedahl1980a}, a conservative estimate for the cooling time for the anorthosite from the Curie temperature of magnetite to the natural unblocking temperature should be about 10 kyr, leading to a cooling rate of 2.5\times10^{-10}.
-This is the value we used for calculating the paleointensity estimate value of 36\%.
-This was not made clear in the original manuscript.
+500\textdegree C is an estimated lab unblocking temperature for the anorthosites based on thermal demagnetization of \cite{Zhang2021b} and the zero-field heating steps of this study.
+Considering the cooling model of \cite{Zhang2021a} and the effect of lowering unblocking temperature due to slow cooling \cite{Pullaiah1975a}, an estimate for the cooling time for the anorthosite from the Curie temperature of magnetite to the natural unblocking temperature could be longer than 1 kyr.
+To fully account for the range of various lower temperature bounds used for optimize the paleointensity estimate, we use 10 kyr for calculating the cooling rate correction factor.
+This leads to a cooling rate of 2.5\times10$^{-10}$.
+This is the value we use for calculating the paleointensity estimate value of 36\%.
 
@@
 Thanks for doing this calculation.
-We have changed the text to ``From the thermal history model of ref.
-\citealp{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they blocked in the majority of their characteristic natural remanence magnetization (lab unblocking temperatures $\sim$500\textdegree C;  \citealp{Zhang2021b}).
-We find the cooling time to be on the order of $\sim$10 kyr considering the effect of slow cooling on lowering the natural remanence unblocking temperatures \cite{Halgedahl1980a}.
-This range of cooling time corresponds to a cooling rate of $\sim2.5\times10^{-10}$ $^\circ$C s$^{-1}$.
-In contrast, the lab cooling rate is much faster through the same temperature interval with an estimated cooling rate of $\sim1.3\times10^{-1}$ $^\circ$C s$^{-1}$.
-The significant cooling rate difference leads to a predicted $\sim$36\% overestimate of true ancient field following the model of \citealp{Halgedahl1980a} (Fig.
-S3)."
+We have changed the text to ``To fully account for the range of various lower temperature bounds used for optimize the paleointensity estimate and the effect of lowering unblocking temperature due to slow cooling, we estimate the anorthosite cooled over a time period on the order of $\sim$10 kyr (from $\sim$580\textdegree C to as low as $\sim$400\textdegree C) based the thermal history model of ref.
+\citealp{Zhang2021b} and cooling rate correction model of \cite{Pullaiah1975a}.
+This cooling time corresponds to a cooling rate of $\sim2.5\times10^{-10}$ $^\circ$C s$^{-1}$.
+In contrast, the lab cooling rate is much faster through the same temperature interval with an estimated cooling rate of $\sim1.3\times10^{-1}$ $^\circ$C s$^{-1}$."
 
@@
 It would also be help to state the recovered paleointensities\/ uncertainties in the text here.}
+
+While the independent two-sample t-test can provide statistical assessment for the similarity of site- and region-level results, the difference between high values from the Kennedy Creek region and the other two regions are quite significant, and the overlapping uncertainties between site AX12 and AX16, and AX11 and AX13 are indicative of their distribution similarities.
+
+We have added reference to our table S3 and added statement of t-test results within and between regions in the caption for table S3.
+Independent two-sample t-test results show that site AX12 and AX16 pass the t-test at 1\% significance level; site AX11 and AX13 pass the test at 5\% significance level; region Kennedy Creek have distinguishable distributions from the other two regions; Silver Bay region and Carlton Peak region are indistinguishable at 5\% significance level.
 
@@
 
+We agree with the comments and have added additional text to the introduction section regarding the light element precipitation aspect.
+
 \textit{Lines 465-484 - this section is quite hard to follow, is there a simpler way to describe these transitions? Perhaps they could be labelled on Figure 6?}
+
+To aid the reader to identify the timing of these transitions, we have added more ticks on the x axis of figure 6. This should make the figure and related text easier to follow.
 
@@
 Can these ideas be introduced sooner in the abstract/introduction to explain why you have recovered such high paleointensities?}
+
+We think that the discussion of plate tectonic driven modulation to field strength is better positioned within the discussion than the introduction.
 
@@
 A MAD selection of $<$10 would result in the same selection as the highest specimen MAD is around 8.
+Our table Figure S3 also presents the small scatter of MAD.
 
@@
 
+While additional inset figures can help illustrating the dipole directions they would add too much to the magnetic field maps and complicate the figure.
+In addition, it would make more sense to show a population of magnetic sources all rotate as the external field changes direction Our figure caption should help guide readers to understand these maps better.
+We hope our caption  ``The magnetic images show that remanent magnetizations of Beaver River anorthosite (i.e.
+the individual dipoles visible with paired red +B$_z$ and blue -B$_z$ lobes) align well with the first applied field direction (E) and then rotated to align with the second applied field direction indicating minimal anisotropic behavior (G)" describes the rotation of the dipoles in detail.
+
 \textit{Figure 5: I think it's a bit deceptive to just include the smoothed fit and not show what are actually quite noisy data (based on what's in your Jupyter notebook).
@@
 
-An additional supplementary figure S6 which plots the fitted coercivity curve and the smoothed data points has been added. The font sizes have been increased.
+The three example specimen shown in this figure actually have quite nice data that are not very noisy.
+To make the main figure clear and highlight the difference in peak coercivities, we decide to keep the presentation of this figure as is.
+The font sizes have been increased.
+
+An additional supplementary figure S6 which plots the fitted coercivity curve and the smoothed data points has been added.
 
@@
 Considering the effect of lowering unblocking temperature due to slow cooling, the cooling timescale should be closer to 10 kyr.
-We have changed the text to ``We find the cooling time to be on the order of $\sim$10 kyr considering the effect of slow cooling on lowering the natural remanence unblocking temperatures \cite{Halgedahl1980a}."
+We have changed related text to ``To fully account for the range of various lower temperature bounds used for optimize the paleointensity estimate and the effect of lowering unblocking temperature due to slow cooling, we estimate the anorthosite cooled over a time period on the order of $\sim$10 kyr (from $\sim$580\textdegree C to as low as $\sim$400\textdegree C) based the thermal history model of ref.
+\citealp{Zhang2021b} and cooling rate correction model of \cite{Pullaiah1975a}.
+This cooling time corresponds to a cooling rate of $\sim2.5\times10^{-10}$ $^\circ$C s$^{-1}$.
+In contrast, the lab cooling rate is much faster through the same temperature interval with an estimated cooling rate of $\sim1.3\times10^{-1}$ $^\circ$C s$^{-1}$.
+The significant cooling rate difference leads to a predicted $\sim$36\% overestimate of true ancient field following the model of \citealp{Pullaiah1975a} (Fig.
+S3)."
 
@@
 
-We have added text ``This factor may result in an over-correction, as some multidomain grains that do not tend to severely bias paleointensity estimates toward higher \cite{Biggin2013a}, may exist in the anorthosite xenoliths as is evidenced through the slightly zigzagging Arai plots" to the related paragraph.
+We have added text ``This factor may result in an over-correction, as some multidomain and pseudo-single domain grains that do not tend to severely bias paleointensity estimates toward higher \cite{Biggin2013a}, may exist in the anorthosite xenoliths as is evidenced through the slightly zigzagging Arai plots" to the related paragraph.
 
@@
 Figure S5 which compares our samples and resampled Cenozoic paleointensity records illustrates that our values are also near the high end.
+We discuss that ``we perform the same comparison with the existing paleointensity data for the Cenozoic Era (the last 66 million years) from the PINT database (filtered by Q$_{PI}\geq$3; Fig.
+S5) given that the observational dataset could record more geomagnetic field excursions or regional high flux patches.
+The anorthosite xenoliths 3 region-mean intensity is higher than all values from the resample results of the PADM2M model, and rival the top 3\% values of the resample averages from the Cenozoic data (Fig.
+S5)."
 
@@
 
+\bibliographystyle{gsabull}
+\bibliography{YZ_ref}
+
 \end{flushleft}

### 2022-04-04 PNAS_manuscript/PNAS_manuscript.tex

@@
 The Ediacaran data have been interpreted to indicate that there was a progressively decaying field up to that time that was soon followed by initial crystallization of the inner core \cite{Bono2019a}.
-Sparse data in the Proterozoic Era (2500 to 539 Ma) and the reality of high variability in Phanerozoic Era records (539 to 0 Ma) necessitate additional data to evaluate interpreted trends.
+Sparse paleointensity data in the Proterozoic Era (2500 to 539 Ma) and the reality of high variability in Phanerozoic Era records (539 to 0 Ma) necessitate additional data to evaluate interpreted trends.
 
@@
 Determinations of the absolute value of ancient geomagnetic field strength rely on igneous rocks that acquire thermal remanent magnetizations as they cool.
-These magnetizations need to be unmodified by subsequent heating or chemical alteration in order to maintain a record of the ancient geomagnetic field from the time of cooling.
+These magnetizations need to be unmodified by subsequent heating or chemical alteration in order to maintain the record of the ancient geomagnetic field from the time of cooling.
 Intracontinental magmatic events are an important target for determination of ancient paleointensity as they can be well-preserved within continental interiors.
-This interior position results in them typically being distant from tectonic events along continental margins that can drive alteration through heat and fluids.
+This interior position results in them typically being distant from tectonic events along continental margins that can drive alteration through heat and fluid flow.
 However, intraplate magmatism associated with large igneous provinces is typically of geologically short duration with the bulk of magmatic products emplaced within 1 Myr or less \cite{Kasbohm2021a}.
@@
 Additionally, extension ceased in the Midcontinent Rift prior to lithospheric separation, preserving volcanic, intrusive, and sedimentary rocks of the rift within the continental interior far from the continental margin and subsequent orogenesis.
-As a result, rocks of the rift have unusually simple paleomagnetic behavior for their greater than one billion-year-old age and paleomagnetic data from rift rocks forms a central record of Mesoproterozoic paleogeography \cite{Swanson-Hysell2021c}.
-The duration of magmatic activity within the Midcontinent Rift is longer than the entire 20.4 Myr long Neogene Period such that it enables an extended well-preserved window into the intensity of Earth's magnetic field in the late Mesoproterozoic.
+As a result, rocks of the rift have unusually simple paleomagnetic behavior for their greater than one billion-year-old age and paleomagnetic data from rift rocks form a central record of Mesoproterozoic paleogeography \cite{Swanson-Hysell2021c}.
+The duration of magmatic activity within the Midcontinent Rift is longer than the entire 20.4 Myr long Neogene Period such that it provides an extended well-preserved window into the intensity of Earth's magnetic field in the late Mesoproterozoic.
 
-Despite the excellent preservation of the rocks, non-ideal paleointensity behaviors have challenged the interpretation of many previous paleointensity results from the Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a}.
-The most trusted type of paleointensity estimate is that obtained through experiments in which the primary natural thermal remanent magnetization is progressively replaced by a laboratory magnetization that is imparted in a known field with internal consistency checks (such as in IZZI-style Thellier experiments;  \citealp{Yu2004a}).
-In such Thellier paleointensity experiments, one typical departure from ideal behavior due to the presence of nonuniformly magnetized grains (with either multidomain \cite{Dunlop2001a} or vortex states \cite{Tauxe2020a}) is sagging or double-slopes as visualized in Arai plots that show thermal remanent magnetization (TRM) acquired versus natural remanent magnetization (NRM) lost.
+Despite the excellent preservation of the rocks, non-ideal paleointensity behaviors have thwarted the interpretation of many previous paleointensity results from the Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a}.
+The most trusted type of paleointensity estimate is that obtained through experiments in which the primary natural remanent magnetization (NRM) is progressively replaced by a laboratory magnetization that is imparted in a known field with internal consistency checks (such as in IZZI-style Thellier experiments;  \citealp{Yu2004a}).
+In such Thellier paleointensity experiments, one typical departure from ideal behavior due to the presence of nonuniformly magnetized grains (with either multidomain \cite{Dunlop2001a} or vortex states \cite{Tauxe2020a}) is sagging or double-slopes as visualized in Arai plots that show thermal remanent magnetization (TRM) acquired versus NRM lost.
 For such data, distinct paleointensity estimates may be calculated depending on the interpreter's choice of slope.
@@
 In this study, we target the high-purity anorthosite xenoliths of the Beaver River diabase in the Midcontinent Rift.
-Anorthosites are cumulate rocks composed dominantly of the mineral plagioclase.
-They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud magnetic inclusions from alteration.
-In addition, the weathering of the plagioclase crystals do not resulting in the formation of secondary iron oxides, whereas other minerals such as olivine and pyroxene are more likely to alter and influence the magnetic mineralogy.
-The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal (shallowly emplaced) intrusions of the Beaver River diabase \cite{Zhang2021b}.
-They likely cooled from tholeiitic magma temperatures and acquired their magnetizations in conjunction with the host diabase at a paleolatitude of 22 $\pm$ 2\textdegree (calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a}).
-The details regarding the thermal history of these anorthosite xenoliths are discussed in \cite{Zhang2021b}.
-Paleointensity experiments on the anorthosite xenoliths have a high success rate, yielding consistent specimen- and site-level paleointensity results.
-The anorthosite xenoliths show low anisotropy of thermal remanent magnetization (TRM) acquisition and can acquire TRM linearly within the range of relevant field strengths.
-Magnetic imaging shows that the anorthosite specimens have dominant magnetic carriers within and interstitial to plagioclase crystals without strong preferred orientations.
-These single-slope, high-quality paleointensity data, together with other high-quality paleointensity records during the protracted Midcontinent Rift magmatic activity, require that there was a strong late Mesoproterozoic geodynamo.
-
-\section*{Background on anorthosite xenoliths of the Beaver River diabase}
-
 While magmatic activity within the Midcontinent Rift was protracted, there were intervals of particularly rapid volcanism and voluminous emplacement of intrusions \cite{Swanson-Hysell2021a}.
@@
 These anorthosite xenoliths are plagioclase cumulates that formed comagmatically with the host diabase in the lower crust---an interpretation confirmed by U-Pb zircon geochronology \cite{Zhang2021b}.
-
-Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths to have acquired thermal remanent magnetizations while cooling with the Beaver River diabase \cite{Zhang2021b}.
-Step-wise thermal demagnetization data show the anorthosite xenoliths to have dominantly single-component magnetization that often unblocks sharply within temperature ranges between 500\textdegree C and 580\textdegree C, consistent with magnetization held by low-titanium titanomagnetite \cite{Zhang2021b}.
+Anorthosites are cumulate rocks composed dominantly of the mineral plagioclase.
+They are attractive targets for paleomagnetic study as plagioclase crystals can protect magnetic inclusions from alteration.
+In addition, the weathering of the plagioclase crystals do not resulting in the formation of secondary iron oxides, whereas other minerals such as olivine and pyroxene are more likely to alter and influence the magnetic mineralogy.
+The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal (shallowly emplaced) intrusions of the Beaver River diabase \cite{Zhang2021b}.
+They were likely heated to tholeiitic magma temperatures and partially recrystallized \cite{Zhang2021b}.
+Fe-Ti oxides within the plagioclase exsolved above magnetite Curie temperature ($\sim$580\textdegree C; \citealp{Bian2021a}) and subsequently cooled and acquired thermal remanent magnetizations in conjunction with the host diabase at a paleolatitude of 22 $\pm$ 2\textdegree (calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a}).
+The details regarding the thermal history of these anorthosite xenoliths are discussed in \cite{Zhang2021b}.
+Paleointensity experiments on the anorthosite xenoliths have a high success rate, yielding consistent specimen- and site-level paleointensity results.
+The anorthosite xenoliths show low anisotropy of thermal remanent magnetization (TRM) acquisition and can acquire TRM linearly within the range of relevant field strengths.
+Magnetic imaging shows that the anorthosite specimens have dominant magnetic carriers within and interstitial to plagioclase crystals without strong preferred orientations.
+Thermal modeling results and paleomagnetic directional data show that the anorthosite xenoliths to have acquired thermal remanent magnetizations while cooling with the Beaver River diabase \cite{Zhang2021b}.
+Step-wise thermal demagnetization data show the anorthosite xenoliths to have dominantly single-component magnetizations that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C, consistent with remanence being held by low-titanium titanomagnetite \cite{Zhang2021b}.
 
@@
 \noindent\includegraphics[width=13 cm]{Petro_QDM_full.png}
-\caption{\footnotesize{Thin section petrographic images (A,B,C,D) and magnetic maps (E,F,G,H) of anorthosite from the Beaver River anorthosite xenolith in the Silver Bay region from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (left column;  \citealp{Zhang2021b}) and a distinct anorthosite from the Duluth Complex Anorthositic Series (right column).
+\caption{\footnotesize{Thin section petrographic images (A,B,C,D) and magnetic maps (E,F,G,H) of an anorthosite sample from the Beaver River anorthosite xenolith in the Silver Bay region from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (left column;  \citealp{Zhang2021b}) and from a distinct anorthosite within the Duluth Complex Anorthositic Series (right column).
 The Duluth Complex anorthosites were not targeted for paleointensity experiments in this study given complexities associated with more pronounced fabrics.
@@
 The B$_z$ color scale is an order of magnitude greater in the Duluth Complex anorthosite maps (F, H) than the maps for the Beaver River anorthosite (E, G).
-(E to G) and (F to H) show experiments performed on both samples where we apply a first field along the y axis of the field of view and then apply a second field of 300 mT orthogonal to the first field direction.
+Panels (E) to (G) and (F) to (H) show experiments performed on both samples where we apply a first field along the y axis of the field of view and then apply a second field of 300 mT orthogonal to the first field direction.
 The remanent magnetizations acquired by both anorthosites were mapped after the application of each field.
@@
 \ref{fig:Petro_QDM}A).
-Plagioclase crystals within coarse-grained intrusions often contain abundant elongate Fe-Ti oxide inclusions that are visible through optical microscopy \cite{Feinberg2006a, Wenk2011a, Ageeva2016a}.
-The Beaver River anorthosite xenoliths lack such large (10s of $\mu$m) oxide inclusions such that oxides are not visible within the plagioclase crystals through optical microscopy (Fig.
+Plagioclase crystals within coarse-grained intrusions often contain abundant elongate Fe-Ti oxide inclusions that are visible with optical microscopy \cite{Feinberg2006a, Wenk2011a, Ageeva2016a}.
+The Beaver River anorthosite xenoliths lack such large (10s of $\mu$m) oxide inclusions such that they are not visible within the plagioclase crystals using optical microscopy (Fig.
 \ref{fig:Petro_QDM}C).
-However, magnetic imaging using a quantum diamond microscope (QDM) shows that there are magnetic remanence carriers within the plagioclase crystals (Fig.
+Magnetic imaging using a quantum diamond microscope (QDM), however, shows that there are magnetic remanence carriers within the plagioclase crystals (Fig.
 \ref{fig:Petro_QDM}E, G).
 The sources of these magnetic remanence carriers are likely to be Fe-Ti oxides that exsolved within plagioclase crystals above the Curie temperature of magnetite \cite{Zhang2021b, Bian2021a}.
-To highlight these distinct aspects of the Beaver River anorthosite xenoliths, we present petrographic and magnetic imaging data from the Beaver River anorthosite as well as a sample of an anorthosite from the Duluth Complex Anorthositic Series---an older intrusive complex within the Midcontinent Rift that was not targeted for paleointensity experiments in this study (Fig.
+To highlight these distinctive aspects of the Beaver River anorthosite xenoliths, we present petrographic and magnetic imaging data from the Beaver River anorthosite as well as an anorthosite sample from the Duluth Complex Anorthositic Series---an older intrusive complex within the Midcontinent Rift that was not targeted for paleointensity experiments in this study (Fig.
 \ref{fig:Petro_QDM}).
-In contrast to the Beaver River anorthosite, the plagioclase of the Duluth Complex anorthosites typically develop interlocking textures that display pronounced igneous foliation (Fig.
+In contrast to the Beaver River anorthosite, the plagioclase crystals of the Duluth Complex anorthosites typically develop interlocking textures that display pronounced igneous foliation (Fig.
 \ref{fig:Petro_QDM}B).
@@
 \ref{fig:Petro_QDM}F, H).
-As a result of this shape anisotropy, magnetization is acquired at angles highly oblique to applied fields (Figure \ref{fig:Petro_QDM}F).
+As a result of this shape anisotropy, remanent magnetizations are acquired at angles highly oblique to applied fields (Figure \ref{fig:Petro_QDM}F).
 In an experiment where a second orthogonal field was applied, the first isothermal remanent magnetizations (IRM) of a set of magnetite needles within a plagioclase grain either fail to rotate or flip by 180\textdegree.
@@
 \ref{fig:Petro_QDM}F, H).
-These experiments enable novel visualization of the grain-scale magnetic anisotropy of elongated exsolved (titano)magnetite inclusions within plagioclase that leads to magnetic anisotropy observed at the bulk sample scale \cite{Selkin2000a, Feinberg2006a}.
+These experiments enable novel visualization of the grain-scale magnetic anisotropy of elongated exsolved (titano)magnetite inclusions within plagioclase grains that leads to magnetic anisotropy observed at the bulk sample scale \cite{Selkin2000a, Feinberg2006a}.
 In contrast, the remanent magnetizations of the ferromagnetic grains imaged in the Beaver River anorthosite xenoliths, targeted for paleointensity in this study, align with the applied field directions indicating minimal remanence anisotropy  (Fig.
@@
 Following IZZI-style paleointensity experiments \cite{Yu2004a}, 40 from a total of 86 anorthosite specimens and 0 out of a total of 69 diabase specimens passed our paleointensity result selection criteria (see Materials and Methods section).
-7 anorthosite sites and no diabase sites have specimen results that pass these selection criteria.
+Seven anorthosite sites and no diabase sites have specimen results that pass these selection criteria.
 Example NRM/TRM (Arai) plots are shown in Figure \ref{fig:IZZI_examples}.
-Summary specimen absolute paleointensity estimates and site-level mean paleointensity values are plotted in Figure \ref{fig:PINT_cooling_corrected} (and provided in Table S1) where each site is an individual anorthosite xenolith.
+Summary specimen absolute paleointensity estimates and site-level mean paleointensity values are plotted in Figure \ref{fig:PINT_cooling_corrected} (and provided in Table S1) where each site represents an individual anorthosite xenolith.
 The paleointensity quality index (Q$_{PI}$; \citealp{Biggin2014a}) for the anorthosite xenolith sites are all 5 or 6 (Table S2).
 The cooling rate-corrected absolute paleointensity estimates from the anorthosite xenoliths have a mean of 38.28$\pm$11.92 $\mu$T.
-The site-mean virtual dipole moment is $\sim$82 ZAm$^2$ ($10^{21} Am^2$) ca.
+The site-mean virtual dipole moment is $\sim$82 ZAm$^2$ ($10^{21} Am^2$) at ca.
 1092 Ma.
@@
 \centering
-\caption{\footnotesize{Example results of paleointensity experiments are displayed on Arai plots and zero-field heating results are shown on inset orthogonal plots (Zijderveld plots) for anorthosite and diabase specimens.
+\caption{\footnotesize{Example results of paleointensity experiments are displayed on Arai plots and zero-field heating results are shown as inset orthogonal plots (Zijderveld plots) for anorthosite and diabase specimens.
 Red (blue) circles indicate zero-field/in-field (in-field/zero-field) steps `ZI’ (`IZ’).
@@
 We accept specimen- and site-level absolute paleointensity results from those anorthosite xenoliths that pass the selection criteria.
-Other anorthosite xenoliths and diabase specimens failed the selection criteria largely because of double-slope or sagging behavior (fail FRAC selection; see Materials and Methods section), poor pTRM checks, and sometimes zigzagging behaviors superimposed on top of sagging behavior (fail SCAT, DANG selection; see Materials and Methods section; Fig.
+Other anorthosite xenoliths and diabase specimens failed the selection criteria largely because of double-slope or sagging behavior (fail FRAC selection; see Materials and Methods section), poor pTRM checks, and zigzagging behaviors occasionally superimposed on top of sagging behavior (fail SCAT, DANG selection; see Materials and Methods section; Fig.
 \ref{fig:IZZI_examples}).
-A 20 mT alternating field treatment after in-field heating steps was applied to some specimens, but it did not result in significant changes in the experimental results for the anorthosite xenolith or diabase specimens (Fig.
+A 20 mT alternating field treatment after in-field heating steps was applied to some specimens, but this treatment did not result in significant changes in the experimental results for the anorthosite xenolith or diabase specimens (Fig.
 \ref{fig:PINT_cooling_corrected}; Supporting Information).
@@
 
-In addition to estimating paleointensity values by introducing a set of selection criteria to filter our experiment results, we apply an independent method from ref.
-\citealp{Cych2021a} with all experimental data regardless of their NRM/TRM plot statistics to perform bias corrected estimation of paleointensity (BiCEP).
-This Bayesian probabilistic method is based on an assumption that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Paterson2011a}.
+In addition to estimating paleointensity values by introducing a set of selection criteria to filter our experiment results, we applied an independent statistical method from ref.
+\citealp{Cych2021a} to all experimental data regardless of their NRM/TRM plot statistics to make a bias-corrected estimation of paleointensity (BiCEP).
+This Bayesian probabilistic method is based on the assumption that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Paterson2011a}.
 The posterior paleointensity distributions from these sites with high-quality specimen-level data are in agreement with the site-level averages developed using the selection criteria approach (Fig.
@@
 S1).
-Overall, the high-quality results from the anorthosite xenoliths of the Beaver River diabase indicate that the anorthosite xenoliths record a high geomagnetic field ca.
+Overall, the high-quality results from the anorthosite xenoliths of the Beaver River diabase indicate that the anorthosite xenoliths record a high geomagnetic field at ca.
 1092 Ma.
@@
 \subsubsection*{Coercivity}
-Additional rock magnetic data support that the anorthosite specimens that pass the paleointensity selection criteria have dominant magnetic remanence carriers with magnetic properties similar to stochiometric, non-interacting, single domain magnetite, whereas anorthosite samples that failed the paleointensity result selection and all diabase samples have more pronounced populations of non-ideal carriers.
-Magnetic property measurement system (MPMS) data show both diabase and anorthosite contain (titano)magnetite as evidenced through the presence of the Verwey transition (Fig.
+Additional rock magnetic data indicate that those anorthosite specimens which pass the paleointensity selection criteria contain dominant magnetic remanence carriers with magnetic properties similar to stochiometric, non-interacting, single domain magnetite, whereas anorthosite samples that failed the paleointensity result selection along with all diabase samples have more pronounced populations of non-ideal carriers.
+Magnetic property measurement system (MPMS) data show that both the diabase and anorthosite contain (titano)magnetite as evidenced by the presence of the Verwey transition (Fig.
 S2;  \citealp{Verwey1939a, Feinberg2015a}).
-Anorthosite specimens from sites that yield successful paleointensity results have Verwey transition temperatures near 120 K as expected for stochiometric magnetite with minimal Ti \cite{Ozdemir1993a}.
-However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have Verwey transitions that are suppressed toward lower temperatures (Fig.
-S2), indicating that magnetite grains in the specimens either have relatively higher Ti content or have been partially oxidized \cite{Ozdemir1993a}.
-Another difference is that samples which pass paleointensity selection criteria have distinctly higher average median destructive field values than other anorthosite and diabase specimens (Fig.
+Anorthosite specimens from sites that yield successful paleointensity results have Verwey transition temperatures near 120 K as expected for stochiometric magnetite with minimal Ti contents \cite{Ozdemir1993a}.
+However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have Verwey transitions that are trending toward lower temperatures (Fig.
+S2), indicating that their magnetic grains either have relatively higher Ti content or have been partially oxidized \cite{Ozdemir1993a}.
+Another difference is that samples which pass paleointensity selection have distinctly higher average median destructive field values than other anorthosite and diabase specimens (Fig.
 \ref{fig:coercivity}).
-Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivities ranging from $\sim$44-144 mT with a median of 53 mT (Fig.
+Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites having successful paleointensity results can contain magnetic grain populations with peak coercivities ranging from $\sim$44-144 mT with a median of 53 mT (Fig.
 \ref{fig:coercivity}).
-In contrast, other anorthosite and diabase tend to have lower peak coercivities ($\sim$30 mT).
-This result is consistent with an interpretation that a population of magnetic grains with more multidomain-like behavior is responsible for the non-ideal paleointensity behaviors during experiments on such specimens \cite{Xu2004a}.
+In contrast, other anorthosite and diabase specimens tend to have lower peak coercivities ($\sim$30 mT).
+This result is consistent with an interpretation that magnetic grain populations with more multidomain-like behavior are responsible for the non-ideal paleointensity behaviors during experiments of such specimens \cite{Xu2004a}.
 
@@
 \subsubsection*{Bulk TRM anisotropy}
-Significant remanence anisotropy has been documented to exist within certain anorthositic rocks that form in layered intrusive complexes \cite{Selkin2000a, Feinberg2006a}.
-Strong remanence anisotropy associated with the igneous foliation developed within anorthosite from the Stillwater Complex has been shown to lead to significant overestimation or underestimation of paleointensity values depending on the relative orientations between the fabrics and an applied magnetic field \cite{Selkin2000a}.
+Significant remanence anisotropy has been documented to exist within certain anorthositic rocks that formed in layered intrusive complexes \cite{Selkin2000a, Feinberg2006a}.
+Strong remanence anisotropy associated with the igneous foliation developed within anorthosite from the Stillwater Complex has been shown to cause significant overestimation or underestimation of paleointensity values depending on the relative orientations between the fabrics and an applied magnetic field \cite{Selkin2000a}.
 To assess whether our paleointensity estimates are biased by bulk remanence anisotropy, we calculated the gamma statistic, which is the angular difference between the last pTRM step of paleointensity experiment and the applied field direction.
-The results show that the anorthosite specimens used for paleointensity experiment have low gamma values ranging from 0.9\textdegree$\;$ to 11.9\textdegree, with a median value of 4.2\textdegree$\;$(Table S1).
-Because these anorthosite specimens were oriented at various directions with respect to the outcrops, the angle between the applied lab field direction during paleointensity experiments with respect to any fabrics within the anorthosites specimens are expected to cover a wide range of angles.
-These gamma values of the anorthosite xenolith bulk specimens are similar to those of Midcontinent Rift volcanics \cite{Sprain2018a}.
+The results show that the anorthosite specimens used in our paleointensity experiment have low gamma values ranging from 0.9\textdegree$\;$ to 11.9\textdegree, with a median value of 4.2\textdegree$\;$(Table S1).
+Because these anorthosite specimens were oriented at various directions with respect to the outcrops, the angles between the applied lab field direction during paleointensity experiments with respect and any fabrics within the anorthosites specimens are expected to cover a wide range of angles.
+These gamma values of the anorthosite xenolith bulk specimens are similar to those of the Midcontinent Rift volcanics studied by \citealp{Sprain2018a}.
 Therefore, the bulk Beaver River anorthosite xenoliths do not have significant remanence anisotropy.
-Paleodirectional data from our anorthosite xenoliths further support that they have minimal remanence anisotropy as their site mean directions closely match those of the Beaver River diabase hosts without deviating due to a fabric \cite{Zhang2021b}.
-These bulk results are consistent with the data from the QDM magnetic map experiments that reveal minimal remenance anisotropy (Fig.
-\ref{fig:Petro_QDM}).
+Paleodirectional data from our anorthosite xenoliths indicate that they have minimal remanence anisotropy as their site mean directions closely match those of the Beaver River diabase hosts without deviating due to a fabric \cite{Zhang2021b}.
+These bulk results are consistent with the data from the QDM magnetic map experiments that also show minimal remenance anisotropy (Fig.
+\ref{fig:Petro_QDM}E, G).
 
@@
 Thermal modeling results from ref.
-\citealp{Zhang2021b} indicate that the Beaver River anorthosite xenoliths were heated to tholeiitic magma temperature ($\sim$1100\textdegree C) and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a thousand years, partially averaging secular variation within single sites.
-Another consequence of the anorthosite xenoliths having cooled in the interior of thick diabase intrusions is that slow cooling rates can bias paleointensity estimates toward higher values.
+\citealp{Zhang2021b} indicate that the Beaver River anorthosite xenoliths were heated to tholeiitic magma temperature ($\sim$1100\textdegree C)---lower than the melting temperature of plagioclase of An$_70$ \cite{Morrison1983a, Doyle2016a} and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a thousand years, partially averaging secular variation within single sites.
+Another consequence of the anorthosite xenoliths having slowly cooled in the interior of thick diabase intrusions is the bias of paleointensity estimates toward higher values.
 Large differences in cooling rates between acquisition of an NRM in nature versus a TRM in the lab can result in overestimated paleointensities for single domain grains \cite{Dodson1980a, Halgedahl1980a, Nagy2021a}.
-From the thermal history model of ref.
-\citealp{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they blocked in the majority of their characteristic natural remanence magnetization (lab unblocking temperatures $\sim$500\textdegree C;  \citealp{Zhang2021b}).
-We find the cooling time to be on the order of $\sim$10 kyr considering the effect of slow cooling on lowering the natural remanence unblocking temperatures \cite{Halgedahl1980a}.
-This range of cooling time corresponds to a cooling rate of $\sim2.5\times10^{-10}$ $^\circ$C s$^{-1}$.
+To fully account for the range of various lower temperature bounds used for optimize the paleointensity estimate and the effect of lowering unblocking temperature due to slow cooling, we estimate the anorthosite cooled over a time period on the order of $\sim$10 kyr (from $\sim$580\textdegree C to as low as $\sim$400\textdegree C) based the thermal history model of ref.
+\citealp{Zhang2021b} and cooling rate correction model of \cite{Pullaiah1975a}.
+This cooling time corresponds to a cooling rate of $\sim2.5\times10^{-10}$ $^\circ$C s$^{-1}$.
 In contrast, the lab cooling rate is much faster through the same temperature interval with an estimated cooling rate of $\sim1.3\times10^{-1}$ $^\circ$C s$^{-1}$.
-The significant cooling rate difference leads to a predicted $\sim$36\% overestimate of true ancient field following the model of \citealp{Halgedahl1980a} (Fig.
+The significant cooling rate difference leads to a predicted $\sim$36\% overestimate of true ancient field following the model of \citealp{Pullaiah1975a} (Fig.
 S3).
@@
 We therefore correct our paleointensity results by a factor of 0.74.
-This factor may result in an over-correction, as some multidomain grains that do not tend to severely bias paleointensity estimates toward higher \cite{Biggin2013a}, may exist in the anorthosite xenoliths as is evidenced through the slightly zigzagging Arai plots (Figure \ref{fig:IZZI_examples}).
+This factor may result in an over-correction, as some multidomain grains and pseudo-single domain grains that do not tend to severely bias paleointensity estimates toward higher \cite{Biggin2013a}, may exist in the anorthosite xenoliths as is evidenced through the slightly zigzagging Arai plots (Figure \ref{fig:IZZI_examples}).
 The cooling rate-corrected specimen paleointensity estimates together with specimen- and site-level means are shown in Fig.
@@
 \ref{fig:Geologic_map} and \ref{fig:PINT_cooling_corrected}).
-$\sim$10 km to the north, anorthosite xenoliths AX11 and AX13 in the Kennedy Creek area were also emplaced closely (separated by $\sim$125 meters) and yield similar values to one another, but distinct paleointensity estimates from those of AX12 and AX16 (Figs.
+$\sim$10 km to the north, anorthosite xenoliths AX11 (54.33 $\pm$ 0.04 $\mu$T) and AX13 (56.76 $\pm$ 3.16 $\mu$T) in the Kennedy Creek area were also emplaced closely ($\sim$125 meters apart) and yield similar values to one another, but distinct paleointensity estimates from those of AX12 (28.22 $\pm$ 4.33 $\mu$T) and AX16 (32.70 $\pm$ 3.76 $\mu$T) (Figs.
 \ref{fig:Geologic_map} and \ref{fig:PINT_cooling_corrected}).
-$\sim$30 km further to the north in the Carlton Peak region, AX6, AX8, and AX10 are anorthosite xenoliths within 10 meters of one another and also yield similar paleointensity estimates albeit with relatively large uncertainties (Fig.
+$\sim$30 km further to the north in the Carlton Peak region, AX6 (22.47 $\pm$ 11.24 $\mu$T), AX8 (38.81 $\pm$ 9.00 $\mu$T), and AX10 (34.68 $\pm$ 0.81 $\mu$T) are anorthosite xenoliths within 10 meters of one another and also yield similar paleointensity estimates albeit with relatively large uncertainties (Fig.
 \ref{fig:PINT_cooling_corrected}).
-The anorthosites of the three distinct regions captured three different intervals of geomagnetic field intensities during the emplacement and cooling of the Beaver River diabase sills ca.
+The anorthosites of the three distinct regions captured three different intervals of geomagnetic field intensities during the emplacement and cooling of the Beaver River diabase sills at ca.
 1092 Ma.
@@
 The locality-mean values of these three regions are plotted in Figure \ref{fig:PINT_cooling_corrected}C (and reported in Table.
-S3) where we also present the distribution of means calculated from 3 random values from the PADM2M model for a total of 10000 iterations.
+S3) where we also present the distribution of means calculated from 3 random values taken from the PADM2M model for a total of 10000 iterations.
 In addition, we perform the same comparison with the existing paleointensity data for the Cenozoic Era (the last 66 million years) from the PINT database (filtered by Q$_{PI}\geq$3; Fig.
 S5) given that the observational dataset could record more geomagnetic field excursions or regional high flux patches.
-The anorthosite xenoliths 3 region mean intensity is higher than all values from the resample results of the PADM2M model, and rival the top 3\% values of the resample averages from the Cenozoic data (Fig.
+The anorthosite xenoliths 3 region-mean intensity is higher than all values from the resample results of the PADM2M model, and rival the top 3\% values of the resample averages from the Cenozoic data (Fig.
 S5).
@@
 The age of the inner core in thermal evolution models relies on estimates for the thermal conductivity of iron alloys at the temperatures and pressures of the core \cite{Ohta2021a}.
-Prior to studies in the last 10 years, an accepted value of $\sim$30 W m$^{-1}$ K$^{-1}$ for this thermal conductivity was used to constrain the timing of inner core nucleation to be during the first half of Earth history \cite{Stacey2007a, Konopkova2016a}.
-Subsequently, experimental data and \textit{ab intio} simulations were interpreted to imply higher thermal conductivity values \cite{Pozzo2012a, Ohta2016a} which in turn imply a younger age of the inner core ($<$700 Ma; \citealp{Labrosse2015a}).
+Prior to studies of the last 10 years, an accepted value of $\sim$30 W m$^{-1}$ K$^{-1}$ for this thermal conductivity constrained the timing of inner core nucleation to be during the first half of Earth's history \cite{Stacey2007a, Konopkova2016a}.
+Subsequently, experimental data and \textit{ab intio} simulations were interpreted to imply higher thermal conductivity values \cite{Pozzo2012a, Ohta2016a} which in turn implied a younger age for the inner core ($<$700 Ma; \citealp{Labrosse2015a}).
 However, other experimental studies continue to indicate lower thermal conductivity values consistent with prior estimates \cite{Konopkova2016a, Hsieh2020a} with no consensus yet emerging \cite{Williams2018a, Ohta2021a}.
@@
 The other data type that can provide insight into the long-term history of the core's thermal regime and geodynamo is paleomagnetic data---both paleodirectional data that indicate the presence of a geomagnetic field and paleointensity data that constrain the field's strength.
-Inner core nucleation would have increased the power to the geodynamo which has the potential to manifest as an increase in Earth's surface field \cite{Davies2021a}.
-An approach combining dynamo simulations and theoretical scaling relationships has predicted that progressive decay of the field's dipole moment would be followed by a rapid increase in geomagnetic field intensity soon after the onset of inner core nucleation such that there is a minimum in dipole moment just before inner core nucleation \cite{Davies2021a}.
+Inner core nucleation would have increased the power of the geodynamo which has the potential to manifest as an increase in Earth's surface field \cite{Davies2021a}.
+An approach combining dynamo simulations and theoretical scaling relationships has predicted that progressive decay of the field's dipole moment would be followed by a rapid increase in geomagnetic field intensity soon after the onset of inner core nucleation such that a minimum in dipole moment would occur just before inner core nucleation \cite{Davies2021a}.
 Other scenarios are possible, however, such as the model-based prediction that while power increases associated with inner core nucleation strengthen Earth's internal magnetic field, that the dynamo becomes more deeply seated in the core diminishes the increase in magnetic field strength at Earth's surface \cite{Aubert2009a, Landeau2017a}.
@@
 
-It has been proposed that Proterozoic paleointensity data are consistent with a progressive monotonic decay leading up to ca.
+It has been proposed that Proterozoic paleointensity data are consistent with a progressive monotonic decay leading to a minimum ca.
 565 Ma in the Ediacaran Period (Fig.
@@
 
-The high paleointensity estimates from the 1.1 billion-year-old Midcontinent Rift rocks challenge the hypothesized monotonic decay of the strength of the geomagnetic field throughout the Proterozoic Era (Fig.
+The high paleointensity estimates from the 1.1 billion-year-old Midcontinent Rift rocks challenge the hypothesized monotonic decay of the geomagnetic field strength throughout the Proterozoic Era (Fig.
 \ref{fig:PINT_compilation}).
 The well-preserved ca.
-1092 Ma anorthosite xenoliths of the Beaver River diabase record a strong geomagnetic field in the late Mesoproterozoic that exceeds the strength of the modern-day field for which crystallization of the inner core is a power source (Figs.
+1092 Ma anorthosite xenoliths of the Beaver River diabase record a strong geomagnetic field in the late Mesoproterozoic that exceeds the modern-day field strength for which crystallization of the inner core is a power source (Figs.
 \ref{fig:PINT_cooling_corrected} and \ref{fig:PINT_compilation}).
@@
 While the ca.
-1140 Abitibi dikes paleointensity estimates do not go to values as low as the ca.
+1140 Abitibi dikes' paleointensity estimates are not as low as those of the ca.
 565 Ma Sept-$\hat{I}$les intrusions, this virtual dipole moment increase in the Mesoproterozoic from the ca.
@@
 \ref{fig:PINT_compilation};  \citealp{Lloyd2021b}).
-Following the Ediacaran, there are also low paleointensity estimates from Devonian rocks such as ca.
+Following the Ediacaran, there are also low paleointensity estimates from Devonian rocks such as the ca.
 370 Ma dikes and lavas of the Siberian Viluy Traps that give virtual dipole moment estimates of 4.3 to 14.9 ZAm$^2$ (Fig.
@@
 The highest site-level value of the virtual dipole moment would rank in the top 2$\%$ of those in the database for the Cenozoic Era when there was unequivocally a crystallizing inner core.
-These high surface field strengths necessitate appreciable power to the late Mesoproterozoic geodynamo.
+These high surface field strengths necessitate that appreciable power was provided to the late Mesoproterozoic geodynamo.
 
@@
 On top of the IZZI-Thellier experiment protocol, we also performed a comparative study where we added an extra step of 20 mT alternating field (AF) cleaning on some of the specimens after each in-field step.
-The purpose is to study whether the AF cleaning could help improve experiment success rate by removing the remanence component carried by materials such as multi-domain (MD) grains that contribute to non-ideal paleointensity behaviors.
+The purpose was to study whether the AF cleaning could help improve experiment success rate by removing the remanence component carried by materials such as multi-domain (MD) grains that contribute to non-ideal paleointensity behaviors.
 The results were similar when this step was applied without an observed change in experimental success rate.
@@
 The following criteria were used as quality filters on the paleointensity results: (1) a maximum angular deviation (MAD;  \citealp{Kirschvink1980a}) of $<$10\textdegree; (2) scatter parameter ($\beta$;  \citealp{Coe1978a}) values of $<$15$\%$; (3) a deviation angle (DANG;  \citealp{Tauxe2004a}) of $<$5\textdegree; (4) fraction of remanence fitted for paleointensity estimate (FRAC;  \cite{Shaar2013a}) $>$0.6; (5) scatter statistic (SCAT;  \citealp{Shaar2013a}) = TRUE; (6) a maximum magnetic moment difference between adjacent zero-field steps (GAP-Max;  \citealp{Shaar2013a}) $<$ 0.25; (7) number of pTRM checks $>$ 2; (8) and number of measurements used for paleointensity determination $\geq$ 4.
-The MAD measures the scatter about the best-fit line through the natural remanent magnetization (NRM) steps in the selected interval for which the intensity is defined.
-DANG, the deviation angle, is the angle between the best-fit direction that is free floating and the direction between the centre of mass of the data and the origin of the vector component diagram \cite{Tauxe2004a}.
+MAD measures the scatter about the best-fit line through the natural remanent magnetization (NRM) steps in the selected interval for which the intensity is defined.
+DANG, the deviation angle, is the angle between the best-fit direction that is free floating and the direction between the center of mass of the data and the origin of the vector component diagram \cite{Tauxe2004a}.
 Both MAD and DANG assess the directional variation of the NRM, with MAD measuring the scatter in the NRM directions and DANG assessing whether the component is trending toward the origin of the Zijderveld plot.
@@
 
-We conducted rock magnetic experiments with a purpose of gaining magnetic mineralogy insights into the paleointensity results of the anorthosite and diabase.
+We conducted rock magnetic experiments to characterize the magnetic mineralogy and gain insights into the paleointensity results of the anorthosite and diabase.
 Backfield curves were measured at room temperature using a Micromag Princeton Measurements vibrating sample magnetometer (VSM) and a Lake Shore 8600 series VSM at the Institute for Rock Magnetism.
-Specimen median destructive fields (MDF) are calculated based on the backfield curves.
+Specimen median destructive fields (MDF) were calculated based on the backfield curves.
 The calculated coercivity spectra were subsequently decomposed into one or more components using skew-normal distributions following the method of  \citealp{Maxbauer2016a} examples of which are shown in Fig.
@@
 In the room-temperature saturation isothermal remanence (RTSIRM) experiment, the sample was pulsed with a 2.5 T field at room temperature ($\sim$300 K) and then cooled to 10 K and warmed back to room temperature in a (near-)zero field.
-The magnetic moment transitions at critical temperatures revealed through MPMS experiments are used to identify magnetic minerals such as magnetite within specimens \cite{Feinberg2015a}.
+The magnetic moment transitions at critical temperatures shown through MPMS experiments were used to identify magnetic minerals such as magnetite within specimens \cite{Feinberg2015a}.
 
@@
 Conversations with Bruce Buffett and Zachary Geballe informed our perspectives on the geodynamic and mineral physics literature.
+We thank Jon Hagstrum at USGS and two anonymous reviewers for their helpful comments.
 Paleomagnetic data associated with this study are available within the MagIC database (\url{https://earthref.org/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}; \textit{provide link when doi is generated at time of proofs}) and all data are within a Github repository along with Python code related to calculations and visualizations (\url{https://github.com/Swanson-Hysell-Group/AX_BD_PINT}) that are also archived on Zenodo (\textit{provide link when doi is generated at time of proofs}).}

### 2022-04-05 PNAS_manuscript/PNAS_Revision_Letter.tex
Updates from Overleaf

@@
 To fully account for the range of various lower temperature bounds used for optimize the paleointensity estimate, we use 10 kyr for calculating the cooling rate correction factor.
-This leads to a cooling rate of 2.5\times10$^{-10}$.
+This leads to a cooling rate of $2.5\times10^{-10}$.
 This is the value we use for calculating the paleointensity estimate value of 36\%.
@@
 A MAD selection of $<$10 would result in the same selection as the highest specimen MAD is around 8.
-Our table Figure S3 also presents the small scatter of MAD.
+Our table Figure S3 presents these MAD values.
 
@@
 
-According to \cite{Tauxe2004a} DANG tests whether the component selected is actually trend- ing toward the origin as it measures the angle between the best-fit direction that is free floating and the direction between the centre of mass of the data and the origin of the vector component diagram.
-We did not use MAD $>$ DANG as an additional selection criteria and prefer to use MAD $<$ 5 and DANG $<$ 10.
+DANG is the Deviation ANGle which is the angle between the free-floating best-fit direction and the direction between data center of mass and the origin of the vector component diagram  \cite{Tauxe2004a}.
+As a result, DANG is considered to assess whether the component selected is actually trending toward the origin and we consider our usage to be appropriate.
 
@@
 Can you indicate which of these specimens come from ``passing" samples and which are from ``failing" samples?
-Also, I'm pretty sure the ZFC>FC being an indication of MD material has been demonstrated to be unreliable.
+Also, I'm pretty sure the ZFC$>$FC being an indication of MD material has been demonstrated to be unreliable.
 (Sorry I can't find the relevant reference.)}
@@
 Labels for whether the sister specimen to the rock chips used for MPMS measurements have been added.
-Figure caption has been updated with the following sentence: ``Pass" or ``Fail" represents whether their sister specimens passed our paleointensity selection or not.
+Figure caption has been updated with the following sentence: ``Pass" or ``Fail" represents whether their sister specimens passed our paleointensity selection criteria or not.
+We are not aware of the demonstration of ZFC$>$FC as being unreliable.
+The papers of Carter-Stiglitz et al.
+nicely demonstrate the effect for magnetite and titanomagnetite.
+Jackson and Moskowitz (2021) cite this effect as giving insight into MD behavior which indicates that these experts still consider it to be a reliable diagnostic.
 

### 2022-04-05 PNAS_manuscript/PNAS_manuscript.tex

@@
 In contrast to scenarios of a progressively decaying field leading up to late nucleation of Earth's inner core, these data record a strong field 1.1 Ga.
-A strong field that persisted over a 14-million-year interval indicates the existence of appreciable power sources for Earth's dynamo during the late Mesoproterozoic.}
+A strong field that persisted over a 14-million-year interval indicates the existence of appreciable power sources for Earth's dynamo during the late Mesoproterozoic Era.}
 
@@
 Estimates of the timing of the initial crystallization of Earth's inner core are interconnected with estimates for the core's thermal conductivity.
-Higher conductivity values imply faster cooling rates, which in turn imply that the threshold for the freezing of the inner core happened more recently \cite{Davies2015a}.
-While some estimated thermal conductivity values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, others have implied higher thermal conductivity values and an age for the inner core that is less than 1.5 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b, Frost2022a}, with some suggesting even younger ages ($<$700 Ma; \citealp{Labrosse2015a, Ohta2016a, Pozzo2022a}).
-The possibility of late inner core nucleation has motivated proposals of novel power sources to sustain the geomagnetic field through early Earth history including precipitation of light materials such as MgO \cite{Badro2016a, ORourke2016a, ORourke2016b} and SiO$_2$ \cite{Mittal2020a} at the core-mantle boundary.
+Higher conductivity values imply faster cooling rates to maintain the geomagnetic field, which in turn imply that the threshold for the freezing of the inner core happened more recently \cite{Davies2015a}.
+While some estimated thermal conductivity values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates have implied higher thermal conductivity values and an age for the inner core that is less than 1.5 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b, Frost2022a}, with some suggesting even younger ages ($<$700 Ma; \citealp{Labrosse2015a, Ohta2016a, Pozzo2022a}).
+The possibility of late inner core nucleation has motivated proposals of novel power sources to sustain the geomagnetic field through early Earth history including precipitation of light-element minerals such as MgO \cite{Badro2016a, ORourke2016a, ORourke2016b} and SiO$_2$ \cite{Mittal2020a} at the core-mantle boundary.
 Given that estimates for the core's thermal conductivity continue to be debated, it is crucial to use observational records as an independent constraint on the thermal evolution of Earth's core and mantle.
@@
 The Midcontinent Rift (Fig.
-\ref{fig:Geologic_map}) is an exception as it is a large igneous province where magmatism lasted $\sim$25 Myr from \texit{ca.} 1109 Ma to 1084 Ma during which there were pulsed intervals of more rapid activity \cite{Swanson-Hysell2021a}.
+\ref{fig:Geologic_map}) is an exception as it is a large igneous province where magmatism lasted $\sim$25 Myr from \texit{ca.} 1109 Ma to 1084 Ma during which there were pulsed intervals of more rapid magmatic activity \cite{Swanson-Hysell2021a}.
 Additionally, extension ceased in the Midcontinent Rift prior to lithospheric separation, preserving volcanic, intrusive, and sedimentary rocks of the rift within the continental interior far from the continental margin and subsequent orogenesis.
@@
 
-Despite the excellent preservation of the rocks, non-ideal paleointensity behaviors have thwarted the interpretation of many previous paleointensity results from the Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a}.
+Despite the excellent preservation of the rocks, non-ideal paleointensity behaviors have posed challenges for the interpretation of many previous paleointensity results from the Midcontinent Rift \cite{Pesonen1983a, Kulakov2013a, Sprain2018a}.
 The most trusted type of paleointensity estimate is that obtained through experiments in which the primary natural remanent magnetization (NRM) is progressively replaced by a laboratory magnetization that is imparted in a known field with internal consistency checks (such as in IZZI-style Thellier experiments;  \citealp{Yu2004a}).
@@
 They are attractive targets for paleomagnetic study as plagioclase crystals can protect magnetic inclusions from alteration.
-In addition, the weathering of the plagioclase crystals do not resulting in the formation of secondary iron oxides, whereas other minerals such as olivine and pyroxene are more likely to alter and influence the magnetic mineralogy.
+In addition, the alteration of the plagioclase crystals does not readily result in the formation of secondary iron oxides in contrast with Fe-silicates minerals such as olivine and pyroxene.
 The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal (shallowly emplaced) intrusions of the Beaver River diabase \cite{Zhang2021b}.
-They were likely heated to tholeiitic magma temperatures and partially recrystallized \cite{Zhang2021b}.
-Fe-Ti oxides within the plagioclase exsolved above magnetite Curie temperature ($\sim$580\textdegree C; \citealp{Bian2021a}) and subsequently cooled and acquired thermal remanent magnetizations in conjunction with the host diabase at a paleolatitude of 22 $\pm$ 2\textdegree (calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a}).
-The details regarding the thermal history of these anorthosite xenoliths are discussed in \cite{Zhang2021b}.
+They were heated to tholeiitic magma temperatures and partially recrystallized \cite{Zhang2021b}.
+Fe-Ti oxides within the plagioclase exsolved above magnetite Curie temperature ($\sim$580\textdegree C; \citealp{Bian2021a}) and subsequently cooled and acquired thermal remanent magnetizations in conjunction with the host diabase at a paleolatitude of 22 $\pm$ 2\textdegree$\;$(calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a}).
 Paleointensity experiments on the anorthosite xenoliths have a high success rate, yielding consistent specimen- and site-level paleointensity results.
@@
 Plagioclase crystals within coarse-grained intrusions often contain abundant elongate Fe-Ti oxide inclusions that are visible with optical microscopy \cite{Feinberg2006a, Wenk2011a, Ageeva2016a}.
-The Beaver River anorthosite xenoliths lack such large (10s of $\mu$m) oxide inclusions such that they are not visible within the plagioclase crystals using optical microscopy (Fig.
+The Beaver River anorthosite xenoliths lack such large (10s of $\mu$m) oxide inclusions such that the oxides are not visible within the plagioclase crystals using optical microscopy (Fig.
 \ref{fig:Petro_QDM}C).
@@
 \ref{fig:Petro_QDM}).
-In contrast to the Beaver River anorthosite, the plagioclase crystals of the Duluth Complex anorthosites typically develop interlocking textures that display pronounced igneous foliation (Fig.
+In contrast to the Beaver River anorthosite, the plagioclase crystals of this Duluth Complex anorthosite sample has a pronounced igneous foliation (Fig.
 \ref{fig:Petro_QDM}B).
@@
 The cooling rate-corrected absolute paleointensity estimates from the anorthosite xenoliths have a mean of 38.28$\pm$11.92 $\mu$T.
-The site-mean virtual dipole moment is $\sim$82 ZAm$^2$ ($10^{21} Am^2$) at ca.
+The site-mean virtual dipole moment is $\sim$82 ZAm$^2$ ($10^{21} Am^2$) ca.
 1092 Ma.
@@
 S1).
-Overall, the high-quality results from the anorthosite xenoliths of the Beaver River diabase indicate that the anorthosite xenoliths record a high geomagnetic field at ca.
+Overall, the high-quality results from the anorthosite xenoliths of the Beaver River diabase indicate that the anorthosite xenoliths record a high geomagnetic field ca.
 1092 Ma.
@@
 Anorthosite specimens from sites that yield successful paleointensity results have Verwey transition temperatures near 120 K as expected for stochiometric magnetite with minimal Ti contents \cite{Ozdemir1993a}.
-However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have Verwey transitions that are trending toward lower temperatures (Fig.
+However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have Verwey transitions that are supressed toward lower temperatures (Fig.
 S2), indicating that their magnetic grains either have relatively higher Ti content or have been partially oxidized \cite{Ozdemir1993a}.
@@
 Thermal modeling results from ref.
-\citealp{Zhang2021b} indicate that the Beaver River anorthosite xenoliths were heated to tholeiitic magma temperature ($\sim$1100\textdegree C)---lower than the melting temperature of plagioclase of An$_70$ \cite{Morrison1983a, Doyle2016a} and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a thousand years, partially averaging secular variation within single sites.
-Another consequence of the anorthosite xenoliths having slowly cooled in the interior of thick diabase intrusions is the bias of paleointensity estimates toward higher values.
+\citealp{Zhang2021b} indicate that the Beaver River anorthosite xenoliths were heated to tholeiitic magma temperature ($\sim$1100\textdegree C)---lower than the melting temperature of the anorthosite plagioclase (given its composition of An$_{70}$; \citealp{Morrison1983a, Doyle2016a}) and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a thousand years, partially averaging secular variation within single sites.
+Another consequence of the anorthosite xenoliths having slowly cooled in the interior of thick diabase intrusions is that slow cooling rates can bias paleointensity estimates toward higher values \cite{Halgedahl1980a}.
 Large differences in cooling rates between acquisition of an NRM in nature versus a TRM in the lab can result in overestimated paleointensities for single domain grains \cite{Dodson1980a, Halgedahl1980a, Nagy2021a}.
@@
 \ref{fig:PINT_cooling_corrected}).
-The anorthosites of the three distinct regions captured three different intervals of geomagnetic field intensities during the emplacement and cooling of the Beaver River diabase sills at ca.
+The anorthosites of the three distinct regions captured three different intervals of geomagnetic field intensities during the emplacement and cooling of the Beaver River diabase sills ca.
 1092 Ma.
@@
 332 Ma Kinghorn volcanics \cite{Hawkins2021a} has led to the proposal of this interval as the ``Mid-Paleozoic Dipole Low'' \cite{Hawkins2021a}.
-This ``Mid-Paleozoic Dipole Low'' is followed by high paleointensity values such that there is a six-fold increase in the database from virtual dipole moment of $\sim$16 ZAm$^2$ at ca.
+This ``Mid-Paleozoic Dipole Low'' is followed by high paleointensity values such that there is a six-fold increase in the database from virtual dipole moment of $\sim$16 ZAm$^2$ ca.
 332 Ma to $\sim$99 ZAm$^2$ ca.

### 2022-04-06 PNAS_manuscript/PNAS_Revision_Letter.tex
Updates from Overleaf

@@
 We agree with the overall take-away that our new data are clearly at odds with the polynomial fit of \cite{Bono2019a}.
-While additional statistical treatments of compilations are available \cite[e.g.][]{Biggin2015a}, in the present manuscript we think it best to focus on these new paleointensity estimates and their relative strength to highlight this main result.
+While additional statistical treatments of compilations are available \cite[e.g.][]{Biggin2015a}, in the present manuscript we think it is the best to focus on these new paleointensity estimates and their relative strength to highlight this main result.
 These questions are well-positioned for further analysis down the line as the reviewer writes.
@@
 
-The sentence `` The possibility of late inner core nucleation has motivated proposals of novel power sources to sustain the geomagnetic field through Earth history including precipitation of light element bearing minerals at the core-mantle boundary including MgO \cite{Badro2016a, ORourke2016b} and SiO$_2$ \cite{Mittal2020a}.'' has been added which refers to the reviewers point here.
+The sentence ``The possibility of late inner core nucleation has motivated proposals of novel power sources to sustain the geomagnetic field through early Earth history including precipitation of light-element minerals such as MgO \cite{Badro2016a, ORourke2016a, ORourke2016b} and SiO$_2$ \cite{Mittal2020a} at the core-mantle boundary.'' has been added which refers to the reviewers point here.
 
@@
 These are good points and are in detail discussed in \cite{Zhang2021b} which is cited in the sentence.
-These anorthosites were likely heated to tholeiitic magma temperature and partially recrystallized.
+These anorthosites were likely heated to tholeiitic magma temperatures which are below the melting point of plagioclase.
 Thus they cooled above the Curie temperature of Fe-Ti oxides which are shown to be the dominant remanence carriers by \cite{Zhang2021a} as well as the rock magnetic data in this study.
@@
 Anorthosites are cumulate rocks composed dominantly of the mineral plagioclase.
-They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud magnetic inclusions from alteration.
-In addition, the weathering of the plagioclase crystals do not resulting in the formation of secondary iron oxides, whereas other minerals such as olivine and pyroxene are more likely to alter and influence the magnetic mineralogy.
+They are attractive targets for paleomagnetic study as plagioclase crystals can protect magnetic inclusions from alteration.
+In addition, the alteration of the plagioclase crystals does not readily result in the formation of secondary iron oxides in contrast with Fe-silicate minerals such as olivine and pyroxene.
 The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal (shallowly emplaced) intrusions of the Beaver River diabase \cite{Zhang2021b}.
-They were likely heated to tholeiitic magma temperatures and partially recrystallized \cite{Zhang2021b}.
-Fe-Ti oxides within the plagioclase exsolved above magnetite Curie temperature ($\sim$580\textdegree C; \cite{Bian2021a}) and subsequently cooled and acquired thermal remanent magnetizations in conjunction with the host diabase at a paleolatitude of 22 $\pm$ 2\textdegree (calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a}).
-The details regarding the thermal history of these anorthosite xenoliths are discussed in \cite{Zhang2021b}.
+They were heated to tholeiitic magma temperatures which are below the melting point of plagioclase (An$_{70}$) \cite{Zhang2021b}.
+Fe-Ti oxides within the plagioclase likely exsolved above magnetite Curie temperature \cite{Bian2021a} and subsequently cooled and acquired thermal remanent magnetizations in conjunction with the host diabase at a paleolatitude of 22 $\pm$ 2\textdegree$\;$(calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a}).
 Paleointensity experiments on the anorthosite xenoliths have a high success rate, yielding consistent specimen- and site-level paleointensity results.
@@
 Magnetic imaging shows that the anorthosite specimens have dominant magnetic carriers within and interstitial to plagioclase crystals without strong preferred orientations.
-Thermal modeling results and paleomagnetic directional data show the anorthosite xenoliths to have acquired thermal remanent magnetizations while cooling with the Beaver River diabase \cite{Zhang2021b}.
-Step-wise thermal demagnetization data show the anorthosite xenoliths to have dominantly single-component magnetization that often unblocks sharply within temperature ranges between 500\textdegree C and 580\textdegree C, consistent with magnetization held by low-titanium titanomagnetite \cite{Zhang2021b}.
-"
+Thermal modeling results and paleomagnetic directional data show that the anorthosite xenoliths to have acquired thermal remanent magnetizations while cooling with the Beaver River diabase \cite{Zhang2021b}.
+Step-wise thermal demagnetization data show the anorthosite xenoliths to have dominantly single-component magnetizations that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C, consistent with remanence being held by low-titanium titanomagnetite \cite{Zhang2021b}."
 
@@
 
-We decide to keep the Bz scale of the QDM maps as is.
-Bringing them to the same scale will either diminish the dipoles in the map of the Beaver River anorthosite xenolith or saturate the color scales in the Duluth anorthosite map.
+We have increased the scale of the magnetic maps for the Beaver River anorthosite xenoliths.
+However, bringing all maps to the same scale will either diminish the dipoles in the map of the Beaver River anorthosite xenolith or saturate the color scales in the Duluth anorthosite map.
 For clarification, we noted in the figure caption that the two maps have Bz scales that are an order of magnitude different.
@@
 500\textdegree C is an estimated lab unblocking temperature for the anorthosites based on thermal demagnetization of \cite{Zhang2021b} and the zero-field heating steps of this study.
-Considering the cooling model of \cite{Zhang2021a} and the effect of lowering unblocking temperature due to slow cooling \cite{Pullaiah1975a}, an estimate for the cooling time for the anorthosite from the Curie temperature of magnetite to the natural unblocking temperature could be longer than 1 kyr.
-To fully account for the range of various lower temperature bounds used for optimize the paleointensity estimate, we use 10 kyr for calculating the cooling rate correction factor.
-This leads to a cooling rate of $2.5\times10^{-10}$.
-This is the value we use for calculating the paleointensity estimate value of 36\%.
+An estimated cooling time for the anorthosite from the Curie temperature of magnetite to 500\textdegree C is about 1.5 kyr.
+This corresponds to a cooling rate of $1.7\times10^{-9}$.
 
@@
 Thanks for doing this calculation.
-We have changed the text to ``To fully account for the range of various lower temperature bounds used for optimize the paleointensity estimate and the effect of lowering unblocking temperature due to slow cooling, we estimate the anorthosite cooled over a time period on the order of $\sim$10 kyr (from $\sim$580\textdegree C to as low as $\sim$400\textdegree C) based the thermal history model of ref.
-\citealp{Zhang2021b} and cooling rate correction model of \cite{Pullaiah1975a}.
-This cooling time corresponds to a cooling rate of $\sim2.5\times10^{-10}$ $^\circ$C s$^{-1}$.
-In contrast, the lab cooling rate is much faster through the same temperature interval with an estimated cooling rate of $\sim1.3\times10^{-1}$ $^\circ$C s$^{-1}$."
+We have changed the text to ``From the thermal history model of \cite{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they blocked in the majority of their characteristic natural remanence magnetization ($\sim$500\textdegree  C; Fig.
+\ref{fig:IZZI_examples}; \citealp{Zhang2021b}).
+We find the cooling time to be on the order of $\sim$1.5 kyr, which corresponds to a cooling rate of $\sim1.7\times10^{-9}$ $^\circ$C s$^{-1}$."
 
@@
 
-We have added reference to our table S3 and added statement of t-test results within and between regions in the caption for table S3.
-Independent two-sample t-test results show that site AX12 and AX16 pass the t-test at 1\% significance level; site AX11 and AX13 pass the test at 5\% significance level; region Kennedy Creek have distinguishable distributions from the other two regions; Silver Bay region and Carlton Peak region are indistinguishable at 5\% significance level.
+We have added reference to our table S3 and added statement of t-test results within and between regions in the caption for table S3: `` Taking a threshold p value of 0.01, the independent two-sample t-test results show that specimen paleointensity distributions of site AX12 and AX16, and those of site AX11 and AX13 are indistinguishable; region Kennedy Creek have distinguishable distributions from the other two regions; Silver Bay region and Carlton Peak region are indistinguishable."
 
@@
 A MAD selection of $<$10 would result in the same selection as the highest specimen MAD is around 8.
-Our table Figure S3 presents these MAD values.
+Our table S1 presents these MAD values.
 
@@
 Our Beaver River anorthosite xenoliths do have higher anorthite content ($\sim$70 An; bytownite) as compared to the plagioclase of the Duluth Complex ($\sim$60 An, labradorite).
-We suspect that the anorthosite xenoliths form during the early stage of the magmatic evolution of the Beaver River diabase when anorthite content, and likely the oxygen fugacity and silica content in the magma was high, leading to more partition of Fe into plagioclase crystal lattice, leading to minor amount of Fe to exsolve and form oxide needles (although those that did could still be crystallographically controlled), instead of forming abundant, large Fe-oxide exsolutions like those in the Duluth anorthosite.
+We suspect that the anorthosite xenoliths form during the early stage of the magmatic evolution of the Beaver River diabase when anorthite content, and likely the oxygen fugacity and silica content in the magma was high, leading to more partition of Fe into plagioclase crystal lattice, and minor amount of Fe to exsolve and form oxide needles (although those that did could still be crystallographically controlled), instead of forming abundant, large Fe-oxide exsolutions like those in the Duluth anorthosite.
 
@@
 This is a good suggestion.
-This sentence has been split into two: `` They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud magnetic inclusions from alteration.
-In addition, the weathering of the plagioclase crystals do not resulting in the formation of secondary iron oxides, whereas other minerals such as olivine and pyroxene are more likely to alter and influence the magnetic mineralogy."
+This sentence has been split into two: `` They are attractive targets for paleomagnetic study as plagioclase crystals can protect magnetic inclusions from alteration.
+In addition, the alteration of the plagioclase crystals does not readily result in the formation of secondary iron oxides in contrast with Fe-silicate minerals such as olivine and pyroxene."
 
@@
 
-While additional inset figures can help illustrating the dipole directions they would add too much to the magnetic field maps and complicate the figure.
-In addition, it would make more sense to show a population of magnetic sources all rotate as the external field changes direction Our figure caption should help guide readers to understand these maps better.
+This is a good point. We added enlarged representative magnetic dipoles maps from both samples and the insets are now included in Fig. 2.
+
 We hope our caption  ``The magnetic images show that remanent magnetizations of Beaver River anorthosite (i.e.
@@
 The font sizes have been increased.
-
 An additional supplementary figure S6 which plots the fitted coercivity curve and the smoothed data points has been added.
@@
 
-\textit{Line 328: cooling time on the order of ~1 kyr - That's not what is shown in Figure S3. There, you show 10 kyr.}
+\textit{Line 328: cooling time on the order of $\sim$1 kyr - That's not what is shown in Figure S3. There, you show 10 kyr.}
 
 Thanks for catching this inconsistency.
-Considering the effect of lowering unblocking temperature due to slow cooling, the cooling timescale should be closer to 10 kyr.
-We have changed related text to ``To fully account for the range of various lower temperature bounds used for optimize the paleointensity estimate and the effect of lowering unblocking temperature due to slow cooling, we estimate the anorthosite cooled over a time period on the order of $\sim$10 kyr (from $\sim$580\textdegree C to as low as $\sim$400\textdegree C) based the thermal history model of ref.
-\citealp{Zhang2021b} and cooling rate correction model of \cite{Pullaiah1975a}.
-This cooling time corresponds to a cooling rate of $\sim2.5\times10^{-10}$ $^\circ$C s$^{-1}$.
+We have updated the text and Fig.
+S3 so that they are consistent.
+The text now reads ``From the thermal history model of \cite{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they blocked in the majority of their characteristic natural remanence magnetization ($\sim$500\textdegree  C; Fig.
+\ref{fig:IZZI_examples}; \citealp{Zhang2021b}).
+We find the cooling time to be on the order of $\sim$1.5 kyr, which corresponds to a cooling rate of $\sim1.7\times10^{-9}$ $^\circ$C s$^{-1}$.
 In contrast, the lab cooling rate is much faster through the same temperature interval with an estimated cooling rate of $\sim1.3\times10^{-1}$ $^\circ$C s$^{-1}$.
-The significant cooling rate difference leads to a predicted $\sim$36\% overestimate of true ancient field following the model of \citealp{Pullaiah1975a} (Fig.
-S3)."
+The significant cooling rate difference leads to a predicted $\sim$33\% overestimate of true ancient field following the model of \citealp{Halgedahl1980a} (Fig.
+S3).
+This estimate on cooling rate effect is similar to the value of $\sim$30\% overestimate derived from the model of ref.
+\citealp{Nagy2021a}.
+We therefore correct our paleointensity results by a factor of 0.75."
 

### 2022-04-06 PNAS_manuscript/PNAS_manuscript.tex
Updates from Overleaf

@@
 In contrast to the diabase which fails to provide reliable paleointensity estimates, the anorthosite xenoliths are unusually high-fidelity recorders yielding high-quality, single-slope paleointensity results that are consistent at specimen and site levels.
-An average value of $\sim$82 ZAm$^2$ for the virtual dipole moment from the anorthosite xenoliths, with the highest site-level values up to $\sim$127 ZAm$^2$, are higher than that of the dipole component of Earth's magnetic field today and rival the highest values in the paleointensity database.
+An average value of $\sim$83 ZAm$^2$ for the virtual dipole moment from the anorthosite xenoliths, with the highest site-level values up to $\sim$129 ZAm$^2$, are higher than that of the dipole component of Earth's magnetic field today and rival the highest values in the paleointensity database.
 Such high intensities recorded by the anorthosite xenoliths require the existence of a strongly powered geodynamo at the time.
@@
 Estimates of the timing of the initial crystallization of Earth's inner core are interconnected with estimates for the core's thermal conductivity.
-Higher conductivity values imply faster cooling rates to maintain the geomagnetic field, which in turn imply that the threshold for the freezing of the inner core happened more recently \cite{Davies2015a}.
+Higher conductivity values imply faster cooling rates to maintain the geomagnetic field, which in turn imply that the threshold for the crystallization of the inner core happened more recently \cite{Davies2015a}.
 While some estimated thermal conductivity values are consistent with an inner core age $>$3 Ga \cite{Gubbins2004a, Konopkova2016a}, other estimates have implied higher thermal conductivity values and an age for the inner core that is less than 1.5 Ga \cite{Pozzo2012a, Koker2012a, Gomi2013a, Zhang2020b, Frost2022a}, with some suggesting even younger ages ($<$700 Ma; \citealp{Labrosse2015a, Ohta2016a, Pozzo2022a}).
@@
 They are attractive targets for paleomagnetic study as plagioclase crystals can protect magnetic inclusions from alteration.
-In addition, the alteration of the plagioclase crystals does not readily result in the formation of secondary iron oxides in contrast with Fe-silicates minerals such as olivine and pyroxene.
+In addition, the alteration of the plagioclase crystals does not readily result in the formation of secondary iron oxides in contrast with Fe-silicate minerals such as olivine and pyroxene.
 The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal (shallowly emplaced) intrusions of the Beaver River diabase \cite{Zhang2021b}.
-They were heated to tholeiitic magma temperatures and partially recrystallized \cite{Zhang2021b}.
-Fe-Ti oxides within the plagioclase exsolved above magnetite Curie temperature ($\sim$580\textdegree C; \citealp{Bian2021a}) and subsequently cooled and acquired thermal remanent magnetizations in conjunction with the host diabase at a paleolatitude of 22 $\pm$ 2\textdegree$\;$(calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a}).
+They were heated to tholeiitic magma temperatures which are below the melting point of plagioclase (An$_{70}$) \cite{Zhang2021b}.
+Fe-Ti oxides within the plagioclase likely exsolved above magnetite Curie temperature \cite{Bian2021a} and subsequently cooled and acquired thermal remanent magnetizations in conjunction with the host diabase at a paleolatitude of 22 $\pm$ 2\textdegree$\;$(calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a}).
 Paleointensity experiments on the anorthosite xenoliths have a high success rate, yielding consistent specimen- and site-level paleointensity results.
@@
 The cooling rate-corrected absolute paleointensity estimates from the anorthosite xenoliths have a mean of 38.28$\pm$11.92 $\mu$T.
-The site-mean virtual dipole moment is $\sim$82 ZAm$^2$ ($10^{21} Am^2$) ca.
+The site-mean virtual dipole moment is $\sim$83 ZAm$^2$ ($10^{21} Am^2$) ca.
 1092 Ma.
@@
 Anorthosite specimens from sites that yield successful paleointensity results have Verwey transition temperatures near 120 K as expected for stochiometric magnetite with minimal Ti contents \cite{Ozdemir1993a}.
-However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have Verwey transitions that are supressed toward lower temperatures (Fig.
+However, diabase and anorthosite specimens that did not pass our paleointensity selection typically have Verwey transitions that are suppressed toward lower temperatures (Fig.
 S2), indicating that their magnetic grains either have relatively higher Ti content or have been partially oxidized \cite{Ozdemir1993a}.
@@
 \ref{fig:coercivity}).
-In contrast, other anorthosite and diabase specimens tend to have lower peak coercivities ($\sim$30 mT).
+In contrast, other anorthosite and diabase specimens tend to have lower peak coercivities ($\sim$18-35 mT with a median of 22 mT).
 This result is consistent with an interpretation that magnetic grain populations with more multidomain-like behavior are responsible for the non-ideal paleointensity behaviors during experiments of such specimens \cite{Xu2004a}.
@@
 Large differences in cooling rates between acquisition of an NRM in nature versus a TRM in the lab can result in overestimated paleointensities for single domain grains \cite{Dodson1980a, Halgedahl1980a, Nagy2021a}.
-To fully account for the range of various lower temperature bounds used for optimize the paleointensity estimate and the effect of lowering unblocking temperature due to slow cooling, we estimate the anorthosite cooled over a time period on the order of $\sim$10 kyr (from $\sim$580\textdegree C to as low as $\sim$400\textdegree C) based the thermal history model of ref.
-\citealp{Zhang2021b} and cooling rate correction model of \cite{Pullaiah1975a}.
-This cooling time corresponds to a cooling rate of $\sim2.5\times10^{-10}$ $^\circ$C s$^{-1}$.
+From the thermal history model of \cite{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they blocked in the majority of their characteristic natural remanence magnetization ($\sim$500\textdegree  C; Fig.
+\ref{fig:IZZI_examples}; \citealp{Zhang2021b}).
+We find the cooling time to be on the order of $\sim$1.5 kyr, which corresponds to a cooling rate of $\sim1.7\times10^{-9}$ $^\circ$C s$^{-1}$.
 In contrast, the lab cooling rate is much faster through the same temperature interval with an estimated cooling rate of $\sim1.3\times10^{-1}$ $^\circ$C s$^{-1}$.
-The significant cooling rate difference leads to a predicted $\sim$36\% overestimate of true ancient field following the model of \citealp{Pullaiah1975a} (Fig.
+The significant cooling rate difference leads to a predicted $\sim$33\% overestimate of true ancient field following the model of \citealp{Halgedahl1980a} (Fig.
 S3).
-This estimate on cooling rate effect is similar to the value of $\sim$35\% overestimate derived from the model of ref.
+This estimate on cooling rate effect is similar to the value of $\sim$30\% overestimate derived from the model of ref.
 \citealp{Nagy2021a}.
-We therefore correct our paleointensity results by a factor of 0.74.
+We therefore correct our paleointensity results by a factor of 0.75.
 This factor may result in an over-correction, as some multidomain grains and pseudo-single domain grains that do not tend to severely bias paleointensity estimates toward higher \cite{Biggin2013a}, may exist in the anorthosite xenoliths as is evidenced through the slightly zigzagging Arai plots (Figure \ref{fig:IZZI_examples}).
@@
 \ref{fig:Geologic_map} and \ref{fig:PINT_cooling_corrected}).
-$\sim$10 km to the north, anorthosite xenoliths AX11 (54.33 $\pm$ 0.04 $\mu$T) and AX13 (56.76 $\pm$ 3.16 $\mu$T) in the Kennedy Creek area were also emplaced closely ($\sim$125 meters apart) and yield similar values to one another, but distinct paleointensity estimates from those of AX12 (28.22 $\pm$ 4.33 $\mu$T) and AX16 (32.70 $\pm$ 3.76 $\mu$T) (Figs.
+$\sim$10 km to the north, anorthosite xenoliths AX11 (55.14 $\pm$ 0.05 $\mu$T) and AX13 (57.62 $\pm$ 3.20 $\mu$T) in the Kennedy Creek area were also emplaced closely ($\sim$125 meters apart) and yield similar values to one another, but distinct paleointensity estimates from those of AX12 (28.65 $\pm$ 4.40 $\mu$T) and AX16 (33.20 $\pm$ 3.82 $\mu$T) (Figs.
 \ref{fig:Geologic_map} and \ref{fig:PINT_cooling_corrected}).
-$\sim$30 km further to the north in the Carlton Peak region, AX6 (22.47 $\pm$ 11.24 $\mu$T), AX8 (38.81 $\pm$ 9.00 $\mu$T), and AX10 (34.68 $\pm$ 0.81 $\mu$T) are anorthosite xenoliths within 10 meters of one another and also yield similar paleointensity estimates albeit with relatively large uncertainties (Fig.
+$\sim$30 km further to the north in the Carlton Peak region, AX6 (22.80 $\pm$ 11.41 $\mu$T), AX8 (39.40 $\pm$ 9.14 $\mu$T), and AX10 (35.20 $\pm$ 0.83 $\mu$T) are anorthosite xenoliths within 10 meters of one another and also yield similar paleointensity estimates albeit with relatively large uncertainties (Fig.
 \ref{fig:PINT_cooling_corrected}).
@@
 1106 Ma Osler Volcanics, with even stronger values from ca.
-1092 Ma by Beaver River anorthosite xenoliths that record virtual dipole moments up to $\sim$127 ZAm$^2$.
+1092 Ma by Beaver River anorthosite xenoliths that record virtual dipole moments up to $\sim$129 ZAm$^2$.
 While the ca.

### 2022-04-06 PNAS_manuscript/Supporting Info.tex

--- 2022-04-03 PNAS_manuscript/Supporting Info.tex
@@
 \end{figure*}
-
-%To check for nonlinear acqusition, we applied full TRMs to a set of anorthosite specimens with lab fields of 30, 50, 70, and 90 $\mu$T along specimen vertical axes.
-The results show that these anorthosite specimens did not saturate upon the applied fields and there is thus no need for correcting nonlinear thermal remanence acquisitions (\citealp{Selkin2007a}; Fig.
-S4).
 
@@
 Overall region-mean values are calculated using all regional mean results in this table.
-}
-\begin{tabular}{ccccccccccc}
-\hline
-\multicolumn{1}{l}{}                       &               & B$_{anc}$ & n  & FRAC & GAP-MAX & MAD ($^\circ$) & DANG ($^\circ$) & inc$_{tc}$ & $\gamma$ & VDM (ZAm$^2$) \\
-\hline
-\multirow{7}{*}{specimen mean for sites}   & AX6           & 30.33     & 3  & 0.69 & 0.13    & 3.55           & 2.66            & 38.9       & 2.63     & 48.76         \\
-                                           & AX8           & 52.4      & 3  & 0.66 & 0.17    & 4.28           & 2.15            & 40.3       & 7.3      & 83.14         \\
-                                           & AX10          & 46.82     & 3  & 0.7  & 0.23    & 4.98           & 2.14            & 36.6       & 6.1      & 76.81         \\
-                                           & AX12          & 38.1      & 6  & 0.75 & 0.22    & 3.26           & 2.53            & 55.2       & 5.98     & 51.31         \\
-                                           & AX16          & 44.15     & 11 & 0.68 & 0.19    & 4.19           & 1.96            & 49.2       & 3.97     & 63.87         \\
-                                           & AX11          & 73.34     & 7  & 0.68 & 0.19    & 3.85           & 1.82            & 35.2       & 6.63     & 121.73        \\
-                                           & AX13          & 76.63     & 7  & 0.73 & 0.2     & 5.7            & 1.67            & 35.1       & 5        & 127.29        \\ 
-                                           \hline
-\multirow{3}{*}{specimen mean for regions} & Carlton Peak  & 43.18     & 9  & 0.68 & 0.17    & 4.27           & 2.32            & 38.6       & 5.34     & 69.57         \\
-                                           & Kennedy Creek & 74.98     & 14 & 0.7  & 0.2     & 4.78           & 1.74            & 35.15      & 5.81     & 124.51        \\
-                                           & Silver Bay    & 42.02     & 17 & 0.71 & 0.2     & 3.86           & 2.16            & 51.32      & 4.68     & 59.43         \\
-\hline
-overall site mean                          &               & 51.68     & 7  &      &         &                &                 &            &          & 81.84         \\
-overall region mean                        &               & 53.39     & 3  &      &         &                &                 &            &          & 84.5         
+All paleointensity values and corresponding virtual dipole moments are corrected for cooling rate by a factor of 0.75.
+Taking a threshold p value of 0.01, the independent two-sample t-test results show that specimen paleointensity distributions of site AX12 and AX16, and those of site AX11 and AX13 are indistinguishable; region Kennedy Creek have distinguishable distributions from the other two regions; Silver Bay region and Carlton Peak region are indistinguishable.}
+\begin{tabular}{cccccccccccc}
+                                           &               & B$_{anc}$ & n  & FRAC & GAP-MAX & MAD ($^\circ$) & DANG ($^\circ$) & inc$_{tc}$ & Paleolatitude & $\gamma$ & VDM (ZAm$^2$) \\ \hline
+\multirow{7}{*}{specimen mean for sites}   & AX6           & 22.8      & 3  & 0.69 & 0.13    & 3.6            & 2.7             & 38.9       & 22.0          & 2.6      & 49.5          \\
+                                           & AX8           & 39.4      & 3  & 0.66 & 0.17    & 4.3            & 2.2             & 40.3       & 23.0          & 7.3      & 84.4          \\
+                                           & AX10          & 35.2      & 3  & 0.70 & 0.23    & 5.0            & 2.1             & 36.6       & 20.4          & 6.1      & 77.9          \\
+                                           & AX11          & 55.2      & 7  & 0.68 & 0.19    & 3.8            & 1.8             & 35.2       & 19.4          & 6.6      & 123.6         \\
+                                           & AX12          & 28.6      & 6  & 0.75 & 0.22    & 3.3            & 2.5             & 55.2       & 35.7          & 6.0      & 52.1          \\
+                                           & AX13          & 57.6      & 7  & 0.73 & 0.2     & 5.7            & 1.7             & 35.1       & 19.4          & 5.0      & 129.2         \\
+                                           & AX16          & 33.2      & 11 & 0.68 & 0.19    & 4.2            & 2.0             & 49.2       & 30.1          & 4.0      & 64.8          \\ \hline
+\multirow{3}{*}{specimen mean for regions} & Carlton Peak  & 32.5      & 9  & 0.68 & 0.17    & 4.3            & 2.3             & 38.6       & 21.8          & 5.3      & 70.6          \\
+                                           & Kennedy Creek & 56.4      & 14 & 0.70 & 0.20    & 4.8            & 1.7             & 35.2       & 19.4          & 5.8      & 126.4         \\
+                                           & Silver Bay    & 31.6      & 17 & 0.71 & 0.20    & 3.9            & 2.2             & 51.3       & 32.1          & 4.7      & 60.3          \\ \hline
+overall site mean                          &               & 38.9      & 7  &      &         &                &                 &            &               &          & 83.1          \\
+overall region mean                        &               & 40.2      & 3  &      &         &                &                 &            &               &          & 85.8         
 \end{tabular}

### 2022-04-07 PNAS_manuscript/PNAS_Revision_Letter.tex
Updates from Overleaf

@@
 
-We agree with the overall take-away that our new data are clearly at odds with the polynomial fit of \cite{Bono2019a}.
-While additional statistical treatments of compilations are available \cite[e.g.][]{Biggin2015a}, in the present manuscript we think it is the best to focus on these new paleointensity estimates and their relative strength to highlight this main result.
+We agree with the overall take-away that our new data are at odds with the polynomial fit of \cite{Bono2019a}.
+While statistical treatments of compiled data \cite[e.g.][]{Biggin2015a} are worth revisiting, in the present manuscript we think it is the best to focus on these new paleointensity estimates and their relative strength to highlight this main result.
 These questions are well-positioned for further analysis down the line as the reviewer writes.
@@
 
-The sentence ``The possibility of late inner core nucleation has motivated proposals of novel power sources to sustain the geomagnetic field through early Earth history including precipitation of light-element minerals such as MgO \cite{Badro2016a, ORourke2016a, ORourke2016b} and SiO$_2$ \cite{Mittal2020a} at the core-mantle boundary.'' has been added which refers to the reviewers point here.
+The sentence ``The possibility of late inner core nucleation has motivated proposals of novel power sources to sustain the geomagnetic field through early Earth history including precipitation of light-element minerals such as MgO \cite{Badro2016a, ORourke2016a, ORourke2016b} and SiO$_2$ \cite{Mittal2020a} at the core-mantle boundary.'' has been added which refers to the reviewer's point here.
 
@@
 Were the Midcontinent Rift extension to have continued, it is very likely that it would have separated Laurentia into two, exposing the rift rocks at a continental margin setting, where subsequent collisions and associated orogenesis would have metamorphosed or completely eroded away the rocks through exhumation.
-However, the extension failed, preserving such rocks in the interior of Laurentia, where subsequent sediments covered them and resulted in great preservation until the recent exposure due to glaciation events.
+However, the extension failed thereby preserving such rocks in the interior of Laurentia.
 
-We have added ``far from the continental margin and subsequent orogenesis."
+We have added ``far from the continental margin and subsequent orogenesis" to succinctly make this point.
 
@@
 
-We have split the sentence into two: ``They are attractive targets for paleomagnetic study as plagioclase crystals can enshroud magnetic inclusions from alteration.
-In addition, the weathering of the plagioclase crystals do not resulting in the formation of secondary iron oxides, whereas other minerals such as olivine and pyroxene are more likely to alter and influence the magnetic mineralogy."
+We have split the sentence into two: ``They are attractive targets for paleomagnetic study as plagioclase crystals can protect magnetic inclusions from alteration.
+In addition, the alteration of the plagioclase crystals does not readily result in the formation of secondary iron oxides in contrast with Fe-silicate minerals such as olivine and pyroxene."
 
@@
 
-These are good points and are in detail discussed in \cite{Zhang2021b} which is cited in the sentence.
-These anorthosites were likely heated to tholeiitic magma temperatures which are below the melting point of plagioclase.
-Thus they cooled above the Curie temperature of Fe-Ti oxides which are shown to be the dominant remanence carriers by \cite{Zhang2021a} as well as the rock magnetic data in this study.
-
-We have integrated this paragraph with the section of ``Background on anorthosite xenoliths of the Beaver River diabase" and the paragraph now reads ``In this study, we target the high-purity anorthosite xenoliths of the Beaver River diabase in the Midcontinent Rift.
-While magmatic activity within the Midcontinent Rift was protracted, there were intervals of particularly rapid volcanism and voluminous emplacement of intrusions \cite{Swanson-Hysell2021a}.
-The ca.
-1092 Ma Beaver Bay Complex in northern Minnesota punctuates one such period of magmatism during the main stage of Midcontinent Rift activity.
-The magma that formed the 1091.7$\pm$0.2 Ma Beaver River diabase dikes and sills of the Beaver Bay Complex transported numerous anorthosite xenoliths that have short-axis diameters up to 180 meters via wide conduits \cite{Boerboom2004a, Boerboom2006b}.
-These anorthosite xenoliths are plagioclase cumulates that formed comagmatically with the host diabase in the lower crust---an interpretation confirmed by U-Pb zircon geochronology \cite{Zhang2021b}.
-Anorthosites are cumulate rocks composed dominantly of the mineral plagioclase.
-They are attractive targets for paleomagnetic study as plagioclase crystals can protect magnetic inclusions from alteration.
-In addition, the alteration of the plagioclase crystals does not readily result in the formation of secondary iron oxides in contrast with Fe-silicate minerals such as olivine and pyroxene.
-The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal (shallowly emplaced) intrusions of the Beaver River diabase \cite{Zhang2021b}.
-They were heated to tholeiitic magma temperatures which are below the melting point of plagioclase (An$_{70}$) \cite{Zhang2021b}.
-Fe-Ti oxides within the plagioclase likely exsolved above magnetite Curie temperature \cite{Bian2021a} and subsequently cooled and acquired thermal remanent magnetizations in conjunction with the host diabase at a paleolatitude of 22 $\pm$ 2\textdegree$\;$(calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a}).
-Paleointensity experiments on the anorthosite xenoliths have a high success rate, yielding consistent specimen- and site-level paleointensity results.
-The anorthosite xenoliths show low anisotropy of thermal remanent magnetization (TRM) acquisition and can acquire TRM linearly within the range of relevant field strengths.
-Magnetic imaging shows that the anorthosite specimens have dominant magnetic carriers within and interstitial to plagioclase crystals without strong preferred orientations.
-Thermal modeling results and paleomagnetic directional data show that the anorthosite xenoliths to have acquired thermal remanent magnetizations while cooling with the Beaver River diabase \cite{Zhang2021b}.
-Step-wise thermal demagnetization data show the anorthosite xenoliths to have dominantly single-component magnetizations that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C, consistent with remanence being held by low-titanium titanomagnetite \cite{Zhang2021b}."
+We have clarified these points to describe that the anorthosites were heated to tholeiitic magma temperatures below the melting point of plagioclase.
+We also clarify how these means that they cooled from temperatures above the Curie temperature of the magnetite remanence carriers.
 
@@
 
-We have integrated this section into the introduction.
+We have integrated this section into the Introduction to avoid repetition.
 
@@
 
-No we did not measure remanence anisotropy by the AARM method.
-Applying such study would require that we use sister specimens to those used for paleointensity.
-We prefer to use the gamma statistics to demonstrate the low degree of TRM anisotropy of our specimens.
+We did not measure remanence anisotropy by the AARM method which would require experiments on distinct sister specimens.
+Given that it is a metric associated with the specimen on which the paleointensity experiment is being conducted, we prefer to use the gamma statistic which is consistent  with a low degree of TRM anisotropy of our specimens.
+Note that directional data also show the same mean directions acquired by the anorthosite xenoliths and host diabase consistent with minimal remanence anisotropy.
 
@@
 
-We have increased the scale of the magnetic maps for the Beaver River anorthosite xenoliths.
+We have decreased the range of the scale for the magnetic maps for the Beaver River anorthosite xenoliths to increase contrast and make the sources more visible.
 However, bringing all maps to the same scale will either diminish the dipoles in the map of the Beaver River anorthosite xenolith or saturate the color scales in the Duluth anorthosite map.
@@
 
-We have added ``Dashed red and blue lines show bounding regions calculated for determine the \textit{SCAT} statistic \cite{Shaar2013a}." to the figure caption.
+We have added ``Dashed red and blue lines show bounding regions associated with the \textit{SCAT} statistic \cite{Shaar2013a}." to the figure caption.
 
@@
 
-We clarify this in the introduction.
+We clarify this point in the introduction.
 
@@
 
-Thanks for doing this calculation.
-We have changed the text to ``From the thermal history model of \cite{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they blocked in the majority of their characteristic natural remanence magnetization ($\sim$500\textdegree  C; Fig.
+We have changed the text and the calculation to be more specific.
+Before we were using a less precise and more conservative order of magnitude estimate.
+With the increase in specificity associated with our cooling model the text now reads: ``From the thermal history model of \cite{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they blocked in the majority of their characteristic natural remanence magnetization ($\sim$500\textdegree  C; Fig.
 \ref{fig:IZZI_examples}; \citealp{Zhang2021b}).
-We find the cooling time to be on the order of $\sim$1.5 kyr, which corresponds to a cooling rate of $\sim1.7\times10^{-9}$ $^\circ$C s$^{-1}$."
+We find the cooling time to be $\sim$1.5 kyr, which corresponds to a cooling rate of $\sim1.7\times10^{-9}$ $^\circ$C s$^{-1}$."
 
@@
 
-While the independent two-sample t-test can provide statistical assessment for the similarity of site- and region-level results, the difference between high values from the Kennedy Creek region and the other two regions are quite significant, and the overlapping uncertainties between site AX12 and AX16, and AX11 and AX13 are indicative of their distribution similarities.
-
-We have added reference to our table S3 and added statement of t-test results within and between regions in the caption for table S3: `` Taking a threshold p value of 0.01, the independent two-sample t-test results show that specimen paleointensity distributions of site AX12 and AX16, and those of site AX11 and AX13 are indistinguishable; region Kennedy Creek have distinguishable distributions from the other two regions; Silver Bay region and Carlton Peak region are indistinguishable."
+We have added reference to our table S3 and added statement of t-test results within and between regions in the caption for table S3: `` Taking a threshold p value of 0.01, the independent two-sample t-test results show that specimen paleointensity distributions of site AX12 and AX16, and those of site AX11 and AX13 are indistinguishable; region Kennedy Creek have distinguishable distributions from the other two regions; Silver Bay region and Carlton Peak region values are indistinguishable."
 
@@
 
-To aid the reader to identify the timing of these transitions, we have added more ticks on the x axis of figure 6. This should make the figure and related text easier to follow.
+To aid the reader to identify the timing of these transitions, we have added more ticks on the x axis (Age [Ma]) of figure 6. This edit should make the figure and related text easier to follow.
 
@@
 
-DANG is the Deviation ANGle which is the angle between the free-floating best-fit direction and the direction between data center of mass and the origin of the vector component diagram  \cite{Tauxe2004a}.
+DANG is the Deviation ANGle which is the angle between the free-floating best-fit direction and the direction between data center of mass and the origin of the vector component diagram  \citep{Tauxe2004a}.
 As a result, DANG is considered to assess whether the component selected is actually trending toward the origin and we consider our usage to be appropriate.
@@
 
-It has been found that partitioning of Fe into plagioclase is enhanced by high anorthite content of the plagioclase, by high oxygen fugacity, and by high silica content of the melt \cite{Phinney1992a, Longhi1976a}.
+It has been found that partitioning of Fe into plagioclase is enhanced by high anorthite content of the plagioclase, by high oxygen fugacity, and by high silica content of the melt \citep{Phinney1992a, Longhi1976a}.
 Our Beaver River anorthosite xenoliths do have higher anorthite content ($\sim$70 An; bytownite) as compared to the plagioclase of the Duluth Complex ($\sim$60 An, labradorite).
@@
 
-The sentence has been changed to ``Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites that yielded successful paleointensity results can have magnetic grain populations with peak coercivities ranging from $\sim$44-144 mT with a median of 53 mT."
+The sentence has been changed to more fully describe the result of higher coercivity for these samples ``Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites having successful paleointensity results can contain magnetic grain populations that have higher coercivities.
+For these successful samples, median destructive destructive (MDF) field values associated with back-field demagnetization experiments range from 44 to 144 mT with a median of 53 mT (Fig.
+\ref{fig:coercivity}).
+In contrast, other anorthosite and diabase specimens tend to have lower peak coercivities (MDF range of 18-35 mT with a median of 22 mT; Fig.
+\ref{fig:coercivity})."
 
@@
 
-This is a good point.
-Experimental study by \cite{Biggin2013a} shows that PSD and MD grains may be less biased toward recording higher paleointensities than SD grains due to slow cooling.
-Therefore our cooling rate correction may be an over-correction given the evidence of some MD grains as pointed out by the reviewer.
-This does not affect our interpretations, as it means the true paleointensity values could be even higher than currently estimated based on the correction factor of 0.74.
-
-We have added text ``This factor may result in an over-correction, as some multidomain and pseudo-single domain grains that do not tend to severely bias paleointensity estimates toward higher \cite{Biggin2013a}, may exist in the anorthosite xenoliths as is evidenced through the slightly zigzagging Arai plots" to the related paragraph.
+This point is a good one given that PSD and MD grains may be less biased toward recording higher paleointensities than SD grains associated with to slow cooling.
+We have added text ``Remanence held by vortex state (pseudo-single domain) and multidomain grains is not as biased by cooling rate \cite{Biggin2013a}.
+The potential for some remanence to be held by these grains, as suggested by slightly zigzagging Arai plots (Figure \ref{fig:IZZI_examples}), could mean that this factor is an over-correction and true paleointensity values could be slightly higher than those reported here.
+The cooling rate-corrected specimen paleointensity estimates together with specimen- and site-level means are shown in Fig.
+\ref{fig:PINT_cooling_corrected}." to the related paragraph.
 
@@
 
-Figure S5 which compares our samples and resampled Cenozoic paleointensity records illustrates that our values are also near the high end.
-We discuss that ``we perform the same comparison with the existing paleointensity data for the Cenozoic Era (the last 66 million years) from the PINT database (filtered by Q$_{PI}\geq$3; Fig.
-S5) given that the observational dataset could record more geomagnetic field excursions or regional high flux patches.
-The anorthosite xenoliths 3 region-mean intensity is higher than all values from the resample results of the PADM2M model, and rival the top 3\% values of the resample averages from the Cenozoic data (Fig.
-S5)."
+This point is an interesting one as local high flux patches are not represented in an axial dipole mode.
+While we think the comparison to PADM2M is valuable, this aspect of the comparison is what also motivated us to compare to the Cenozoic paleointensity database (where the values from this study are in the 97 percentile).
+This comparison is described in the text and illustrated in Figure S5.
 
@@
 
-We have decreased transparency on the data points from the compilation so they are better illustrated.
+We have decreased transparency on the data points from the compilation. Now they can be seen more clearly.
 
@@
 
-The caption has been updated as the following: `Violin plots of site-level posterior paleointensity distributions estimated using the bias corrected estimation of paleointensity (BiCEP) method developed by \citealp{Cych2021a}.
-Blue bars of each violin plot represents the median of the distributions.
-The yellow dashed line is the mean paleointensity of the posterior estimates across all sites.
-Assuming that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Arai1963a, Paterson2011a}, the method uses all paleointensity measurement-level data without applying selection criteria.
-For comparison of results from this independent method with those based on our selection (as show in Fig.
-4 in manuscript), we highlight the anorthosite sites that pass our paleointensity selection and make other anorthosite and diabase transparent.
-The results from the BiCEP method address the uncertainties associated with anorthosite AX6 and AX8.
-This is associated with the relatively variable specimen behaviors within these two sites.
-But for sites AX10, AX11, AX13, AX12, and AX16, the posterior probability distributions have very narrow bounds, consistent with the interpretation that these anorthosites are faithful paleointensity recorders that have high-quality paleointensity behaviors.
-AX-anorthosite xenolith site; BD-Beaver River diabase site.
-The grouping of cooling unit is based on the spacial proximity between site locations and the inclusion relationship of host diabase and anorthosite xenoliths reported by \citealp{Zhang2021b}."
+The caption has been updated to clairfy the symbology.
 

### 2022-04-07 PNAS_manuscript/PNAS_manuscript.tex

@@
 conducted and analyzed rock magnetic experiments; Y.Z.
-performed petrography and electron microscopy; R.R.F.
+performed petrography and magnetic microscopy; R.R.F.
 contributed to magnetic imaging data analyses; Y.Z.
@@
 These anorthosite xenoliths are plagioclase cumulates that formed comagmatically with the host diabase in the lower crust---an interpretation confirmed by U-Pb zircon geochronology \cite{Zhang2021b}.
-Anorthosites are cumulate rocks composed dominantly of the mineral plagioclase.
 They are attractive targets for paleomagnetic study as plagioclase crystals can protect magnetic inclusions from alteration.
@@
 The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal (shallowly emplaced) intrusions of the Beaver River diabase \cite{Zhang2021b}.
-They were heated to tholeiitic magma temperatures which are below the melting point of plagioclase (An$_{70}$) \cite{Zhang2021b}.
+They would have been heated to tholeiitic magma temperature which at $\sim$1100\textdegree C is below the melting point of the xenolith plagioclase (70$\%$ anorthite), but well above the Curie temperature of magnetite \cite{Zhang2021b}.
 Fe-Ti oxides within the plagioclase likely exsolved above magnetite Curie temperature \cite{Bian2021a} and subsequently cooled and acquired thermal remanent magnetizations in conjunction with the host diabase at a paleolatitude of 22 $\pm$ 2\textdegree$\;$(calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a}).
 Paleointensity experiments on the anorthosite xenoliths have a high success rate, yielding consistent specimen- and site-level paleointensity results.
-The anorthosite xenoliths show low anisotropy of thermal remanent magnetization (TRM) acquisition and can acquire TRM linearly within the range of relevant field strengths.
+The anorthosite xenoliths show low anisotropy of thermal remanent magnetization acquisition and can acquire TRM linearly within the range of relevant field strengths.
 Magnetic imaging shows that the anorthosite specimens have dominant magnetic carriers within and interstitial to plagioclase crystals without strong preferred orientations.
-Thermal modeling results and paleomagnetic directional data show that the anorthosite xenoliths to have acquired thermal remanent magnetizations while cooling with the Beaver River diabase \cite{Zhang2021b}.
+Thermal modeling results and paleomagnetic directional data show that the anorthosite xenoliths acquired thermal remanent magnetizations while cooling with the Beaver River diabase \cite{Zhang2021b}.
 Step-wise thermal demagnetization data show the anorthosite xenoliths to have dominantly single-component magnetizations that often unblock sharply within temperature ranges between 500\textdegree C and 580\textdegree C, consistent with remanence being held by low-titanium titanomagnetite \cite{Zhang2021b}.
@@
 After the application of an orthogonal 300 mT external field, magnetization of those needles in plagioclase 1 did not change direction due to strong shape anisotropy whereas the magnetization of needles in plagioclase 2 flipped, but with the acquired remanence remaining oblique to the field direction (H).
+Insets in (E) to (H) show example dipoles direction responses to the application of orthogonal IRMs.
 This experiment illustrates that the Beaver River anorthosite xenolith has much lower magnetic anisotropy than the Duluth anorthosite xenolith.
@@
 \ref{fig:Petro_QDM}D).
-Magnetic imaging confirms that these magnetite needles have magnetic moments oriented along their long axes (Fig.
+Magnetic imaging confirms that these needles have magnetic moments oriented along their long axes (Fig.
 \ref{fig:Petro_QDM}F, H).
@@
 The paleointensity quality index (Q$_{PI}$; \citealp{Biggin2014a}) for the anorthosite xenolith sites are all 5 or 6 (Table S2).
-The cooling rate-corrected absolute paleointensity estimates from the anorthosite xenoliths have a mean of 38.28$\pm$11.92 $\mu$T.
+The cooling rate-corrected absolute paleointensity estimates from the anorthosite xenoliths have a mean of 38.86 $\pm$ 12.10 $\mu$T.
 The site-mean virtual dipole moment is $\sim$83 ZAm$^2$ ($10^{21} Am^2$) ca.
@@
 Blue and red squares in the Zijderveld plots are X–Y and X–Z projections, respectively, of the NRMs in specimen coordinates.
-Dashed red and blue lines show bounding regions calculated for determine the \textit{SCAT} statistic \cite{Shaar2013a}.
+Dashed red and blue lines show bounding regions associated with the \textit{SCAT} statistic \cite{Shaar2013a}.
 Plots on the top row show successful specimen paleointensity results with straight, single-slope behaviors that pass our selection criteria.
@@
 Specimen BD2-5a is an example where the data appear linear with distinct slopes in the low and high temperature ranges such that it could pass less restrictive selection criteria, particularly if a narrower temperature range was used for the experiment.
-Data analysis and visualization was conducted using PmagPy \cite{Tauxe2016a}}}
+Data analysis and visualization was conducted using PmagPy \cite{Tauxe2016a}.}}
 \label{fig:IZZI_examples}
@@
 \centering
-\caption{\footnotesize{A) Summary plot of individual specimen absolute paleointensity results (square symbols) and their averages and standard deviations at site level (black bars with grey one standard deviation uncertainty boxes) where each `AX' site is an individual anorthosite xenolith within the Beaver River diabase.
-All results are corrected for cooling rate with a factor of 0.74.
-The sites with successful experiments come from 3 regions (Carlton Peak, Silver Bay, Kennedy Creek) which would have cooled at distinct times yielding similar estimates within the each region with differences between regions.
+\caption{\footnotesize{A) Summary plot of individual specimen absolute paleointensity results (square symbols) and their averages and standard deviations at site level (black bars with grey one standard deviation uncertainty envelopes) where each `AX' site is an individual anorthosite xenolith within the Beaver River diabase.
+All results are corrected for cooling rate bias with a factor of 0.75.
+The sites with successful experiments come from 3 regions (Carlton Peak, Silver Bay, Kennedy Creek) which would have cooled at distinct times yielding similar estimates within each region with differences between regions.
 B) Regional means calculated from the specimen-level data are compared to the distribution of intensities calculated from a paleomagnetic axial dipole moment model for the past 2 million years (PADM2M; \citealp{Ziegler2011a}) at the latitude corresponding to the paleolatitude of study region (22\textdegree N).
@@
 The distribution represents a total of 10,000 iterations of taking 3 random draws and calculating the mean.
-These comparisons highlight that these paleointensity values are strong in relation to the geomagnetic field over the past 2 million years.
+These comparisons highlight that these anorthosites' paleointensity values are strong in relation to the geomagnetic field over the past 2 million years.
 All shaded regions in B) and C) represent one standard deviation uncertainties.}}
@@
 \ref{fig:coercivity}).
-Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites having successful paleointensity results can contain magnetic grain populations with peak coercivities ranging from $\sim$44-144 mT with a median of 53 mT (Fig.
+Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites having successful paleointensity results can contain magnetic grain populations that have higher coercivities.
+For these successful samples, median destructive destructive (MDF) field values associated with back-field demagnetization experiments range from 44 to 144 mT with a median of 53 mT (Fig.
 \ref{fig:coercivity}).
-In contrast, other anorthosite and diabase specimens tend to have lower peak coercivities ($\sim$18-35 mT with a median of 22 mT).
+In contrast, other anorthosite and diabase specimens tend to have lower peak coercivities (MDF range of 18-35 mT with a median of 22 mT; Fig.
+\ref{fig:coercivity}).
 This result is consistent with an interpretation that magnetic grain populations with more multidomain-like behavior are responsible for the non-ideal paleointensity behaviors during experiments of such specimens \cite{Xu2004a}.
@@
 \centering
-\caption{\footnotesize{Top: Example coercivity spectra of anorthosite and diabase specimens from sites that pass or fail our paleointensity selection criteria.
-Bottom: Box plots of median destructive field (MDF) values for all anorthosite and diabase specimens with single-component coercivity unmixing results.
+\caption{\footnotesize{Top: Example coercivity spectra of anorthosite and diabase chips from sites that pass or fail our paleointensity selection criteria.
+Bottom: Box plots of median destructive field (MDF) values associated with back-field demagnetization for all anorthosite and diabase specimens with single-component coercivity unmixing results.
 Both plots show that anorthosite specimens that pass paleointensity selection criteria have higher coercivities consistent with a higher portion of single-domain-like magnetite grains than the other anorthosite specimens and the diabase.}}
@@
 The results show that the anorthosite specimens used in our paleointensity experiment have low gamma values ranging from 0.9\textdegree$\;$ to 11.9\textdegree, with a median value of 4.2\textdegree$\;$(Table S1).
-Because these anorthosite specimens were oriented at various directions with respect to the outcrops, the angles between the applied lab field direction during paleointensity experiments with respect and any fabrics within the anorthosites specimens are expected to cover a wide range of angles.
+Because these anorthosite specimens were oriented at various directions with respect to the outcrops, the angles between the applied lab field direction during paleointensity experiments with respect to any fabrics within the anorthosite specimens are expected to cover a wide range of angles.
 These gamma values of the anorthosite xenolith bulk specimens are similar to those of the Midcontinent Rift volcanics studied by \citealp{Sprain2018a}.
@@
 Paleodirectional data from our anorthosite xenoliths indicate that they have minimal remanence anisotropy as their site mean directions closely match those of the Beaver River diabase hosts without deviating due to a fabric \cite{Zhang2021b}.
-These bulk results are consistent with the data from the QDM magnetic map experiments that also show minimal remenance anisotropy (Fig.
+These bulk results are consistent with the data from the grain-scale magnetic anisotropy experiments that also show minimal remenance anisotropy (Fig.
 \ref{fig:Petro_QDM}E, G).
@@
 Thermal modeling results from ref.
-\citealp{Zhang2021b} indicate that the Beaver River anorthosite xenoliths were heated to tholeiitic magma temperature ($\sim$1100\textdegree C)---lower than the melting temperature of the anorthosite plagioclase (given its composition of An$_{70}$; \citealp{Morrison1983a, Doyle2016a}) and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a thousand years, partially averaging secular variation within single sites.
+\citealp{Zhang2021b} indicate that the Beaver River anorthosite xenoliths were heated to tholeiitic magma temperature ($\sim$1100\textdegree C)---lower than the melting temperature of the anorthosite plagioclase (given its composition of An$_{70}$; \citealp{Morrison1983a, Doyle2016a}) and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a few thousand years, partially averaging secular variation within single sites.
 Another consequence of the anorthosite xenoliths having slowly cooled in the interior of thick diabase intrusions is that slow cooling rates can bias paleointensity estimates toward higher values \cite{Halgedahl1980a}.
@@
 \ref{fig:IZZI_examples}; \citealp{Zhang2021b}).
-We find the cooling time to be on the order of $\sim$1.5 kyr, which corresponds to a cooling rate of $\sim1.7\times10^{-9}$ $^\circ$C s$^{-1}$.
+We find the cooling time to $\sim$500\textdegree C to be $\sim$1.5 kyr, which corresponds to a cooling rate of $\sim1.7\times10^{-9}$ $^\circ$C s$^{-1}$.
 In contrast, the lab cooling rate is much faster through the same temperature interval with an estimated cooling rate of $\sim1.3\times10^{-1}$ $^\circ$C s$^{-1}$.
@@
 We therefore correct our paleointensity results by a factor of 0.75.
-This factor may result in an over-correction, as some multidomain grains and pseudo-single domain grains that do not tend to severely bias paleointensity estimates toward higher \cite{Biggin2013a}, may exist in the anorthosite xenoliths as is evidenced through the slightly zigzagging Arai plots (Figure \ref{fig:IZZI_examples}).
+Remanence held by vortex state (pseudo-single domain) and multidomain grains is not as biased by cooling rate \cite{Biggin2013a}.
+The potential for some remanence to be held by these grains, as suggested by slightly zigzagging Arai plots (Figure \ref{fig:IZZI_examples}), could mean that this factor is an over-correction and true paleointensity values could be slightly higher than those reported here.
 The cooling rate-corrected specimen paleointensity estimates together with specimen- and site-level means are shown in Fig.
@@
 Anorthosite xenoliths have paleointensity results that are consistent within small regions, but vary between regions.
-Anorthosite AX12 and AX16 in the Silver Bay area were emplaced $\sim$450 meters apart and have indistinguishable paleointensity estimates (Figs.
+Anorthosite AX12 (28.65 $\pm$ 4.40 $\mu$T) and AX16 (33.20 $\pm$ 3.82 $\mu$T) in the Silver Bay area were emplaced $\sim$450 meters apart and have indistinguishable paleointensity estimates (Figs.
 \ref{fig:Geologic_map} and \ref{fig:PINT_cooling_corrected}).
-$\sim$10 km to the north, anorthosite xenoliths AX11 (55.14 $\pm$ 0.05 $\mu$T) and AX13 (57.62 $\pm$ 3.20 $\mu$T) in the Kennedy Creek area were also emplaced closely ($\sim$125 meters apart) and yield similar values to one another, but distinct paleointensity estimates from those of AX12 (28.65 $\pm$ 4.40 $\mu$T) and AX16 (33.20 $\pm$ 3.82 $\mu$T) (Figs.
+$\sim$10 km to the north, anorthosite xenoliths AX11 (55.14 $\pm$ 0.05 $\mu$T) and AX13 (57.62 $\pm$ 3.20 $\mu$T) in the Kennedy Creek area were also emplaced closely ($\sim$125 meters apart) and yield similar values to one another, but distinct paleointensity estimates from those of AX12 and AX16 (Figs.
 \ref{fig:Geologic_map} and \ref{fig:PINT_cooling_corrected}).
@@
 S5) given that the observational dataset could record more geomagnetic field excursions or regional high flux patches.
-The anorthosite xenoliths 3 region-mean intensity is higher than all values from the resample results of the PADM2M model, and rival the top 3\% values of the resample averages from the Cenozoic data (Fig.
+The anorthosite xenoliths three region-mean intensity is higher than all values from the resample results of the PADM2M model, and rival the top 3\% values of the resample averages from the Cenozoic data (Fig.
 S5).
-These data supports that the anorthosite xenoliths likely record an exceptionally strong geomagnetic field ca.
+This comparison supports that the anorthosite xenoliths record an exceptionally strong geomagnetic field ca.
 1092 Ma.
@@
 \citealp{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}.
-Paleointensity estimates from \cite{Pesonen1983a} and \cite{Kulakov2013a} are not included in the compilation due to the specimen-level double-slope behavior as discussed in the text.
-Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \cite{Bono2019a} which is based on an interpretation of there being a monotonic decay of the geodynamo through the Proterozoic.
+Paleointensity estimates from refs.
+\citealp{Pesonen1983a} and \citealp{Kulakov2013a} are not included in the compilation due to the specimen-level double-slope behavior as discussed in the text.
+Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from ref.
+\citealp{Bono2019a} which is based on an interpretation of there being a monotonic decay of the geodynamo through the Proterozoic.
 The highest site-mean virtual dipole moment of the anorthosites would rank in the top 2$\%$ of those in the database for the Cenozoic Era (the last 66 million years) when there was unequivocally a crystallizing inner core.
@@
 The other data type that can provide insight into the long-term history of the core's thermal regime and geodynamo is paleomagnetic data---both paleodirectional data that indicate the presence of a geomagnetic field and paleointensity data that constrain the field's strength.
-Inner core nucleation would have increased the power of the geodynamo which has the potential to manifest as an increase in Earth's surface field \cite{Davies2021a}.
+Inner core nucleation would have increased the power to the geodynamo which has the potential to manifest as an increase in Earth's surface field \cite{Davies2021a}.
 An approach combining dynamo simulations and theoretical scaling relationships has predicted that progressive decay of the field's dipole moment would be followed by a rapid increase in geomagnetic field intensity soon after the onset of inner core nucleation such that a minimum in dipole moment would occur just before inner core nucleation \cite{Davies2021a}.
 Other scenarios are possible, however, such as the model-based prediction that while power increases associated with inner core nucleation strengthen Earth's internal magnetic field, that the dynamo becomes more deeply seated in the core diminishes the increase in magnetic field strength at Earth's surface \cite{Aubert2009a, Landeau2017a}.
-Such a scenario where the dynamo shifts to a greater depth associated with inner core nucleation led \cite{Aubert2009a} to conclude that the increase in power to the dynamo would be difficult to detect with paleointensity data.
+Such a scenario where the dynamo shifts to a greater depth associated with inner core nucleation led ref.
+\citealp{Aubert2009a} to conclude that the increase in power to the dynamo would be difficult to detect with paleointensity data.
 Ultimately, further observational paleomagnetic records are key as they hold the potential for testing different model predictions and identifying transitions in ancient field strength \cite{Biggin2015a, Bono2019a}.
@@
 565 Ma Sept-$\hat{I}$les intrusions \cite{Bono2019a} to those from ca.
-550 Ma volcanics of the Skinner Cove Formation \cite{Thallner2021a} (Fig.
-\ref{fig:PINT_compilation}).
+550 Ma volcanics of the Skinner Cove Formation (Fig.
+\ref{fig:PINT_compilation}; \citealp{Thallner2021a}).
 In the late Mesoproterozoic, there is at least a six-fold increase within a period of $\sim$35 Myr from a low average virtual dipole moment of $\sim$13 ZAm$^2$ recorded by the ca.
@@
 408–393 Ma Buribay volcanics \cite{Shcherbakova2021a} and the ca.
-332 Ma Kinghorn volcanics \cite{Hawkins2021a} has led to the proposal of this interval as the ``Mid-Paleozoic Dipole Low'' \cite{Hawkins2021a}.
+332 Ma Kinghorn volcanics \cite{Hawkins2021a} has led to the proposal of this interval being the ``Mid-Paleozoic Dipole Low'' \cite{Hawkins2021a}.
 This ``Mid-Paleozoic Dipole Low'' is followed by high paleointensity values such that there is a six-fold increase in the database from virtual dipole moment of $\sim$16 ZAm$^2$ ca.
@@
 All heating steps were performed in air.
-The temperature increments for the experiments were chosen to isolate magnetizations held by (titano)magnetite informed by the previous demagnetization data, with smaller increments performed close to the expected unblocking temperature of $\sim$580\textdegree C.
+The temperature increments for the experiments were chosen to isolate magnetizations held by (titano)magnetite informed by the previous demagnetization data, with smaller increments performed close to $\sim$580\textdegree C.
 
@@
 We conducted rock magnetic experiments to characterize the magnetic mineralogy and gain insights into the paleointensity results of the anorthosite and diabase.
-Backfield curves were measured at room temperature using a Micromag Princeton Measurements vibrating sample magnetometer (VSM) and a Lake Shore 8600 series VSM at the Institute for Rock Magnetism.
-Specimen median destructive fields (MDF) were calculated based on the backfield curves.
+Back-field curves were measured at room temperature using a Micromag Princeton Measurements vibrating sample magnetometer (VSM) and a Lake Shore 8600 series VSM at the Institute for Rock Magnetism.
+Specimen median destructive fields (MDF) were calculated based on the back-field experiments.
 The calculated coercivity spectra were subsequently decomposed into one or more components using skew-normal distributions following the method of  \citealp{Maxbauer2016a} examples of which are shown in Fig.
 \ref{fig:coercivity}.
-We also used a magnetic property measurement system (MPMS) at the Institute for Rock Magnetism to aid in the identification of magnetic minerals.
+We also used a magnetic property measurement system (MPMS) at the Institute for Rock Magnetism to aid the identification of magnetic minerals.
 In the field-cooled (FC) experiments, specimen magnetizations were measured upon warming following the specimen having cooled in an applied field of 2.5 T from 300 to 10 K.
@@
 \ref{fig:Petro_QDM}).
-We use the QDM to image the magnetic field over the polished thin section surfaces with a sample-sensor distance of 5 $\mu$m in projective magnetic microscopy (PMM) mode with a spatial resolution of 4.7 μm per pixel and an instantaneous 0.9 mT bias field that is canceled during the course of measurement \cite{Glenn2017a}.
+We use the QDM to image the magnetic field over the polished thin section surfaces with a sample-sensor distance of $\sim$5 $\mu$m in projective magnetic microscopy (PMM) mode with a spatial resolution of 4.7 μm per pixel and an instantaneous 0.9 mT bias field that is canceled during the course of measurement \cite{Glenn2017a}.
 }
@@
 Conversations with Bruce Buffett and Zachary Geballe informed our perspectives on the geodynamic and mineral physics literature.
-We thank Jon Hagstrum at USGS and two anonymous reviewers for their helpful comments.
+We thank Jon Hagstrum at USGS and two anonymous reviewers for their helpful comments on the manuscript.
 Paleomagnetic data associated with this study are available within the MagIC database (\url{https://earthref.org/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}; \textit{provide link when doi is generated at time of proofs}) and all data are within a Github repository along with Python code related to calculations and visualizations (\url{https://github.com/Swanson-Hysell-Group/AX_BD_PINT}) that are also archived on Zenodo (\textit{provide link when doi is generated at time of proofs}).}

### 2022-04-07 PNAS_manuscript/Supporting Info.tex

@@
 \centering
-\caption{{Violin plots of site-level posterior paleointensity distributions estimated using the bias corrected estimation of paleointensity (BiCEP) method developed by \citealp{Cych2021a}.
-Blue bars of each violin plot represents the median of the distributions.
-The yellow dashed line is the mean paleointensity of the posterior estimates across all sites.
+\caption{{Violin plots of site-level posterior paleointensity distributions estimated using the bias corrected estimation of paleointensity (BiCEP) method developed by ref.
+\citealp{Cych2021a}.
+Blue bars of each violin plot represent the median of the distributions.
+The yellow dashed line is the mean paleointensity of the posterior estimates across all sites estimated using BiCEP (38.82 $\pm$ 4.60 mT).
 Assuming that paleointensity estimates from specimens that come from a same cooling unit are distributed around a true paleointensity value with the various deflections being expressed as the curvature parameter of the NRM/TRM plot \cite{Arai1963a, Paterson2011a}, the method uses all paleointensity measurement-level data without applying selection criteria.
 For comparison of results from this independent method with those based on our selection (as show in Fig.
-4 in manuscript), we highlight the anorthosite sites that pass our paleointensity selection and make other anorthosite and diabase transparent.
+4 in manuscript), we highlight the anorthosite sites that pass our paleointensity selection and make other anorthosite and diabase more transparent.
 The results from the BiCEP method address the uncertainties associated with anorthosite AX6 and AX8.
 This is associated with the relatively variable specimen behaviors within these two sites.
-But for sites AX10, AX11, AX13, AX12, and AX16, the posterior probability distributions have very narrow bounds, consistent with the interpretation that these anorthosites are faithful paleointensity recorders that have high-quality paleointensity behaviors.
-AX-anorthosite xenolith site; BD-Beaver River diabase site.
-The grouping of cooling unit is based on the spacial proximity between site locations and the inclusion relationship of host diabase and anorthosite xenoliths reported by \citealp{Zhang2021b}.}}
+For sites AX10, AX11, AX13, AX12, and AX16, the posterior probability distributions have very narrow bounds, consistent with the interpretation that these anorthosites are faithful paleointensity recorders that have high-quality paleointensity behaviors.
+AX--anorthosite xenolith site; BD--Beaver River diabase site.
+The grouping of cooling units is based on the spatial proximity between site locations and the inclusion relationship of host diabase and anorthosite xenoliths reported by ref.
+\citealp{Zhang2021b}.}}
 \label{fig:PINT_BiCEP}
@@
 Specimen BD2-2r has a consistently higher moment during the zero-field-cooled step than during the field-cooled step.
-This is consistent with the interpretation that multidomian magnetic carriers exist in significant quantity in this specimen \cite{Carter-Stiglitz2006a}.}}
+This is consistent with the interpretation that multidomain (MD) magnetic carriers exist in significant quantity in this specimen \cite{Carter-Stiglitz2006a}.}}
 \label{fig:MPMS}
@@
 \caption{{Graph of predicted paleointensity overestimate due to slow cooling of the intrusive Beaver River diabase and anorthosite xenoliths relative to the cooling rate in laboratory following the model of \citealp{Halgedahl1980a}.
-Because the majority of the anorthosites have unblocking temperatures between 500\textdegree C and 580\textdegree C, we estimate that the slow cooling during natural remanence acquisition could have resulted in a 36\% overestimate.
-Thus, a correction factor of 0.74 is applied at specimen level for the paleointensity summary plot.}}
+Because the majority of the anorthosites have unblocking temperatures between 500\textdegree C and 580\textdegree C, we estimate that the slow cooling during natural remanence acquisition could have resulted in a 33\% overestimate.
+Thus, a correction factor of 0.75 is applied at specimen level for the paleointensity summary plot.}}
 \label{fig:PINT_cooling_corrected}
@@
 \centering
-\caption{\footnotesize{A) Summary plot of individual specimen absolute paleointensity results (square symbols) and their averages and standard deviations at site level (black bars with grey uncertainty boxes) from this study.
-All results are corrected for cooling rate with a factor of 0.74.
+\caption{\footnotesize{Data from this study as in Figure 4 of the main text, but compared to the Cenozoic paleointensity database rather than the PADM2M model.
+A) Summary plot of individual specimen absolute paleointensity results (square symbols) and their averages and standard deviations at site level (black bars with grey uncertainty boxes) from this study.
+All results are corrected for cooling rate with a factor of 0.75.
 Each `AX' site is an individual anorthosite xenolith within the Beaver River diabase.
 The sites with successful experiments come from 3 regions which would have cooled at distinct times yielding similar estimates within the each region with differences between regions.
-B) Specimen level means calculated for these regions are compared to the distribution of intensities calculated from the existing paleointensity data in the PINT database (PINT v8.0.0; \url{http://www.pintdb.org/}; \citealp{Bono2021a}) at the latitude corresponding to the paleolatitude of study region (22\textdegree N).
+B) Specimen level means calculated for these regions are compared to the distribution of intensities calculated from the existing paleointensity data in the Cenozoic (past 66 Myr) in the PINT database (PINT v8.0.0; \url{http://www.pintdb.org/}; \citealp{Bono2021a}) at the latitude corresponding to the paleolatitude of study region (22\textdegree N).
 C) The mean of the 3 regional means is compared to means calculated from 3 random values drawn from the Cenozoic data.
@@
 \centering
-\caption{\footnotesize{A) Example coercivity spectra of anorthosite and diabase specimens from sites that pass or fail our paleointensity selection criteria.
+\caption{\footnotesize{Example coercivity spectra of anorthosite and diabase specimens from sites that pass or fail our paleointensity selection criteria.
 The specimens and fits are the same as those shown in Figure 5 in the manuscript.
@@
 \caption{\footnotesize{Specimen paleointensity results that passed our selection.
-Paleointensity results for specimens that passed quality criteria.
 B$_{anc}$ is the calculated ancient field intensity over the chosen temperature interval in $\mu$T.
@@
 N is the number of steps used within the selected interval for paleointensity determination.
-FRAC is the fraction of remanence.
+FRAC is the fraction of remanence used for fitting.
 NpTRM shows the number of pTRM checks within the selected interval for paleointensity determination.
@@
 SCAT is the scatter parameter.
+inc$_{tc}$ is the tilt-corrected inclination.
 Paleolatitude is calculated from the inclination values reported in \cite{Zhang2021b}.
@@
 \centering
-\begin{tabular}{cccccccccccccccc}
+\begin{tabular}{ccccccccccccccccc}
 \hline
-Site & Specimen & B$_{anc}$ & T$_{min}$ & T$_{max}$ & N    & FRAC & NpTRM & $\beta$ & GAP-MAX & MAD ($^\circ$) & DANG ($^\circ$) & SCAT & Paleolatitude & $\gamma$ & VADM (ZAm$^2$) \\
+Site & Specimen & B$_{anc}$ & T$_{min}$ & T$_{max}$ & N  & FRAC & NpTRM & $\beta$ & GAP-MAX & MAD ($^\circ$) & DANG ($^\circ$) & SCAT & inc$_{tc}$ & Paleolatitude & $\gamma$ & VDM (ZAm$^2$) \\
 \hline
-AX6  & AX6-2a   & 18.25     & 400       & 585       & 18 & 0.7  & 10    & 0.04    & 0.12    & 3.44           & 3.43            & PASS & 21.97         & 2.7      & 29.34         \\
-AX6  & AX6-3a   & 25.39     & 400       & 585       & 18 & 0.76 & 10    & 0.04    & 0.1     & 4.28           & 2.88            & PASS & 21.97         & 3.2      & 40.82         \\
-AX6  & AX6-1a   & 47.36     & 475       & 585       & 15 & 0.6  & 10    & 0.02    & 0.16    & 2.92           & 1.67            & PASS & 21.97         & 2        & 76.13         \\
-AX8  & AX8-3a   & 41.3      & 400       & 580       & 17 & 0.75 & 9     & 0.03    & 0.14    & 4.38           & 2.22            & PASS & 22.98         & 11.2     & 65.53         \\
-AX8  & AX8-2a   & 50.51     & 400       & 580       & 17 & 0.63 & 9     & 0.04    & 0.16    & 3.19           & 1.29            & PASS & 22.98         & 3.7      & 80.15         \\
-AX8  & AX8-1a   & 65.38     & 425       & 566       & 13 & 0.6  & 8     & 0.07    & 0.2     & 5.28           & 2.95            & PASS & 22.98         & 7        & 103.74        \\
-AX10 & AX10-1a  & 45.84     & 425       & 585       & 17 & 0.78 & 10    & 0.06    & 0.24    & 5.6            & 2.62            & PASS & 20.37         & 4.7      & 75.2          \\
-AX10 & AX10-2a  & 46.61     & 450       & 585       & 16 & 0.62 & 10    & 0.04    & 0.2     & 5.32           & 2.24            & PASS & 20.37         & 7.8      & 76.46         \\
-AX10 & AX10-3a  & 48.01     & 425       & 585       & 17 & 0.69 & 10    & 0.04    & 0.24    & 4.02           & 1.56            & PASS & 20.37         & 5.8      & 78.76         \\
-AX11 & AX11-1a  & 73.29     & 425       & 560       & 10 & 0.67 & 6     & 0.08    & 0.21    & 5.28           & 2.54            & PASS & 19.43         & 5.9      & 121.64        \\
-AX11 & AX11-2a  & 73.29     & 400       & 560       & 11 & 0.65 & 6     & 0.07    & 0.21    & 3.94           & 1.51            & PASS & 19.43         & 9.9      & 121.64        \\
-AX11 & AX11-4a  & 73.31     & 500       & 570       & 11 & 0.66 & 8     & 0.09    & 0.2     & 1.53           & 3.27            & PASS & 19.43         & 5.6      & 121.68        \\
-AX11 & AX11-6a  & 73.32     & 200       & 562       & 14 & 0.69 & 7     & 0.07    & 0.17    & 4              & 1.33            & PASS & 19.43         & 3        & 121.69        \\
-AX11 & AX11-9a  & 73.34     & 100       & 562       & 15 & 0.67 & 7     & 0.06    & 0.17    & 5.34           & 1.13            & PASS & 19.43         & 11.9     & 121.73        \\
-AX11 & AX11-3a  & 73.41     & 400       & 560       & 11 & 0.66 & 6     & 0.08    & 0.21    & 3.95           & 2.11            & PASS & 19.43         & 8.2      & 121.84        \\
-AX11 & AX11-5a  & 73.44     & 400       & 566       & 14 & 0.73 & 8     & 0.05    & 0.18    & 2.93           & 0.85            & PASS & 19.43         & 1.9      & 121.89        \\
-AX12 & AX12-14a & 30.34     & 475       & 585       & 16 & 0.75 & 10    & 0.06    & 0.17    & 1.74           & 0.55            & PASS & 35.73         & 0.9      & 40.86         \\
-AX12 & AX12-1a  & 33.89     & 0         & 580       & 21 & 0.97 & 9     & 0.03    & 0.17    & 5.47           & 4.73            & PASS & 35.73         & 10       & 45.64         \\
-AX12 & AX12-6a  & 37.18     & 425       & 564       & 12 & 0.69 & 7     & 0.08    & 0.22    & 3.64           & 2.05            & PASS & 35.73         & 5        & 50.07         \\
-AX12 & AX12-8a  & 38.34     & 475       & 565       & 12 & 0.75 & 8     & 0.06    & 0.25    & 1.46           & 2.29            & PASS & 35.73         & 1.6      & 51.63         \\
-AX12 & AX12-4a  & 41.89     & 500       & 585       & 14 & 0.66 & 10    & 0.05    & 0.24    & 3.85           & 3.19            & PASS & 35.73         & 11.4     & 56.42         \\
-AX12 & AX12-2a  & 46.94     & 425       & 570       & 14 & 0.7  & 8     & 0.05    & 0.24    & 3.39           & 2.38            & PASS & 35.73         & 7        & 63.22         \\
-AX13 & AX13-3a  & 71.68     & 100       & 550       & 12 & 0.71 & 5     & 0.08    & 0.22    & 5.92           & 1.88            & PASS & 19.36         & 3.4      & 119.07        \\
-AX13 & AX13-7a  & 72.6      & 475       & 585       & 15 & 0.61 & 10    & 0.03    & 0.2     & 3.62           & 2.09            & PASS & 19.36         & 11.6     & 120.6         \\
-AX13 & AX13-4a  & 76.13     & 200       & 550       & 11 & 0.65 & 5     & 0.04    & 0.18    & 8.82           & 1.07            & PASS & 19.36         & 3.7      & 126.47        \\
-AX13 & AX13-6a  & 76.8      & 200       & 555       & 12 & 0.7  & 6     & 0.06    & 0.22    & 3.82           & 1.53            & PASS & 19.36         & 4.5      & 127.58        \\
-AX13 & AX13-2a  & 76.86     & 200       & 555       & 12 & 0.71 & 6     & 0.05    & 0.2     & 8.08           & 2.23            & PASS & 19.36         & 4.1      & 127.68        \\
-AX13 & AX13-9a  & 77.48     & 0         & 564       & 17 & 0.94 & 7     & 0.03    & 0.16    & 5.96           & 1.07            & PASS & 19.36         & 2.6      & 128.71        \\
-AX13 & AX13-8a  & 84.83     & 450       & 570       & 13 & 0.81 & 8     & 0.02    & 0.24    & 3.69           & 1.82            & PASS & 19.36         & 5.1      & 140.92        \\
-AX16 & AX16-15a & 31.91     & 475       & 570       & 13 & 0.61 & 8     & 0.07    & 0.17    & 3.07           & 1.41            & PASS & 30.08         & 2.6      & 46.16         \\
-AX16 & AX16-13a & 41.68     & 400       & 570       & 15 & 0.72 & 8     & 0.07    & 0.2     & 3.47           & 1.85            & PASS & 30.08         & 1.1      & 60.29         \\
-AX16 & AX16-16a & 43.3      & 400       & 562       & 13 & 0.65 & 7     & 0.09    & 0.18    & 5.14           & 2.94            & PASS & 30.08         & 2        & 62.63         \\
-AX16 & AX16-14a & 43.64     & 400       & 565       & 14 & 0.63 & 8     & 0.08    & 0.14    & 3.84           & 2.12            & PASS & 30.08         & 6        & 63.13         \\
-AX16 & AX16-11a & 44.17     & 450       & 570       & 14 & 0.7  & 8     & 0.06    & 0.15    & 2.86           & 2.07            & PASS & 30.08         & 4.6      & 63.89         \\
-AX16 & AX16-4a  & 44.82     & 425       & 575       & 15 & 0.76 & 9     & 0.05    & 0.21    & 4.77           & 1.08            & PASS & 30.08         & 4.2      & 64.83         \\
-AX16 & AX16-1a  & 44.9      & 200       & 564       & 15 & 0.87 & 7     & 0.06    & 0.22    & 5.56           & 2.4             & PASS & 30.08         & 4.5      & 64.95         \\
-AX16 & AX16-5a  & 44.91     & 425       & 560       & 10 & 0.65 & 6     & 0.1     & 0.24    & 7.42           & 3.9             & PASS & 30.08         & 4        & 64.96         \\
-AX16 & AX16-2a  & 45.05     & 425       & 564       & 12 & 0.65 & 7     & 0.07    & 0.22    & 4.25           & 1.44            & PASS & 30.08         & 4.9      & 65.17         \\
-AX16 & AX16-9a  & 48.04     & 500       & 585       & 15 & 0.61 & 10    & 0.04    & 0.17    & 2.98           & 1.23            & PASS & 30.08         & 4.3      & 69.49         \\
-AX16 & AX16-10a & 53.27     & 510       & 585       & 14 & 0.62 & 10    & 0.04    & 0.2     & 2.75           & 1.1             & PASS & 30.08         & 5.5      & 77.06          \\
+AX6  & AX6-2a   & 13.72     & 400       & 585       & 18 & 0.70 & 10    & 0.04    & 0.12    & 3.4            & 3.4             & PASS & 38.9       & 22.0          & 2.7      & 29.8          \\
+AX6  & AX6-3a   & 19.09     & 400       & 585       & 18 & 0.76 & 10    & 0.04    & 0.10    & 4.3            & 2.9             & PASS & 38.9       & 22.0          & 3.2      & 41.4          \\
+AX6  & AX6-1a   & 35.61     & 475       & 585       & 15 & 0.60 & 10    & 0.02    & 0.16    & 2.9            & 1.7             & PASS & 38.9       & 22.0          & 2.0      & 77.3          \\
+AX8  & AX8-3a   & 31.05     & 400       & 580       & 17 & 0.75 & 9     & 0.03    & 0.14    & 4.4            & 2.2             & PASS & 40.3       & 23.0          & 11.2     & 66.5          \\
+AX8  & AX8-2a   & 37.98     & 400       & 580       & 17 & 0.63 & 9     & 0.04    & 0.16    & 3.2            & 1.3             & PASS & 40.3       & 23.0          & 3.7      & 81.4          \\
+AX8  & AX8-1a   & 49.16     & 425       & 566       & 13 & 0.60 & 8     & 0.07    & 0.20    & 5.3            & 3.0             & PASS & 40.3       & 23.0          & 7.0      & 105.3         \\
+AX10 & AX10-1a  & 34.47     & 425       & 585       & 17 & 0.78 & 10    & 0.06    & 0.24    & 5.6            & 2.6             & PASS & 36.6       & 20.4          & 4.7      & 76.3          \\
+AX10 & AX10-2a  & 35.05     & 450       & 585       & 16 & 0.62 & 10    & 0.04    & 0.20    & 5.3            & 2.2             & PASS & 36.6       & 20.4          & 7.8      & 77.6          \\
+AX10 & AX10-3a  & 36.10     & 425       & 585       & 17 & 0.69 & 10    & 0.04    & 0.24    & 4.0            & 1.6             & PASS & 36.6       & 20.4          & 5.8      & 79.9          \\
+AX11 & AX11-1a  & 55.11     & 425       & 560       & 10 & 0.67 & 6     & 0.08    & 0.21    & 5.3            & 2.5             & PASS & 35.2       & 19.4          & 5.9      & 123.5         \\
+AX11 & AX11-2a  & 55.11     & 400       & 560       & 11 & 0.65 & 6     & 0.07    & 0.21    & 3.9            & 1.5             & PASS & 35.2       & 19.4          & 9.9      & 123.5         \\
+AX11 & AX11-4a  & 55.12     & 500       & 570       & 11 & 0.66 & 8     & 0.09    & 0.20    & 1.5            & 3.3             & PASS & 35.2       & 19.4          & 5.6      & 123.5         \\
+AX11 & AX11-6a  & 55.13     & 200       & 562       & 14 & 0.69 & 7     & 0.07    & 0.17    & 4.0            & 1.3             & PASS & 35.2       & 19.4          & 3.0      & 123.5         \\
+AX11 & AX11-9a  & 55.14     & 100       & 562       & 15 & 0.67 & 7     & 0.06    & 0.17    & 5.3            & 1.1             & PASS & 35.2       & 19.4          & 11.9     & 123.5         \\
+AX11 & AX11-3a  & 55.20     & 400       & 560       & 11 & 0.66 & 6     & 0.08    & 0.21    & 4.0            & 2.1             & PASS & 35.2       & 19.4          & 8.2      & 123.7         \\
+AX11 & AX11-5a  & 55.22     & 400       & 566       & 14 & 0.73 & 8     & 0.05    & 0.18    & 2.9            & 0.9             & PASS & 35.2       & 19.4          & 1.9      & 123.7         \\
+AX12 & AX12-14a & 22.81     & 475       & 585       & 16 & 0.75 & 10    & 0.06    & 0.17    & 1.7            & 0.6             & PASS & 55.2       & 35.7          & 0.9      & 41.5          \\
+AX12 & AX12-1a  & 25.48     & 0         & 580       & 21 & 0.97 & 9     & 0.03    & 0.17    & 5.5            & 4.7             & PASS & 55.2       & 35.7          & 10.0     & 46.3          \\
+AX12 & AX12-6a  & 27.95     & 425       & 564       & 12 & 0.69 & 7     & 0.08    & 0.22    & 3.6            & 2.1             & PASS & 55.2       & 35.7          & 5.0      & 50.8          \\
+AX12 & AX12-8a  & 28.83     & 475       & 565       & 12 & 0.75 & 8     & 0.06    & 0.25    & 1.5            & 2.3             & PASS & 55.2       & 35.7          & 1.6      & 52.4          \\
+AX12 & AX12-4a  & 31.50     & 500       & 585       & 14 & 0.66 & 10    & 0.05    & 0.24    & 3.9            & 3.2             & PASS & 55.2       & 35.7          & 11.4     & 57.3          \\
+AX12 & AX12-2a  & 35.29     & 425       & 570       & 14 & 0.70 & 8     & 0.05    & 0.24    & 3.4            & 2.4             & PASS & 55.2       & 35.7          & 7.0      & 64.2          \\
+AX13 & AX13-3a  & 53.89     & 100       & 550       & 12 & 0.71 & 5     & 0.08    & 0.22    & 5.9            & 1.9             & PASS & 35.1       & 19.4          & 3.4      & 120.9         \\
+AX13 & AX13-7a  & 54.59     & 475       & 585       & 15 & 0.61 & 10    & 0.03    & 0.20    & 3.6            & 2.1             & PASS & 35.1       & 19.4          & 11.6     & 122.4         \\
+AX13 & AX13-4a  & 57.24     & 200       & 550       & 11 & 0.65 & 5     & 0.04    & 0.18    & 8.8            & 1.1             & PASS & 35.1       & 19.4          & 3.7      & 128.4         \\
+AX13 & AX13-6a  & 57.74     & 200       & 555       & 12 & 0.70 & 6     & 0.06    & 0.22    & 3.8            & 1.5             & PASS & 35.1       & 19.4          & 4.5      & 129.5         \\
+AX13 & AX13-2a  & 57.79     & 200       & 555       & 12 & 0.71 & 6     & 0.05    & 0.20    & 8.1            & 2.2             & PASS & 35.1       & 19.4          & 4.1      & 129.6         \\
+AX13 & AX13-9a  & 58.26     & 0         & 564       & 17 & 0.94 & 7     & 0.03    & 0.16    & 6.0            & 1.1             & PASS & 35.1       & 19.4          & 2.6      & 130.7         \\
+AX13 & AX13-8a  & 63.78     & 450       & 570       & 13 & 0.81 & 8     & 0.02    & 0.24    & 3.7            & 1.8             & PASS & 35.1       & 19.4          & 5.1      & 143.0         \\
+AX16 & AX16-15a & 23.99     & 475       & 570       & 13 & 0.61 & 8     & 0.07    & 0.17    & 3.1            & 1.4             & PASS & 49.2       & 30.1          & 2.6      & 46.8          \\
+AX16 & AX16-13a & 31.34     & 400       & 570       & 15 & 0.72 & 8     & 0.07    & 0.20    & 3.5            & 1.9             & PASS & 49.2       & 30.1          & 1.1      & 61.2          \\
+AX16 & AX16-16a & 32.56     & 400       & 562       & 13 & 0.65 & 7     & 0.09    & 0.18    & 5.1            & 2.9             & PASS & 49.2       & 30.1          & 2.0      & 63.6          \\
+AX16 & AX16-14a & 32.81     & 400       & 565       & 14 & 0.63 & 8     & 0.08    & 0.14    & 3.8            & 2.1             & PASS & 49.2       & 30.1          & 6.0      & 64.1          \\
+AX16 & AX16-11a & 33.21     & 450       & 570       & 14 & 0.70 & 8     & 0.06    & 0.15    & 2.9            & 2.1             & PASS & 49.2       & 30.1          & 4.6      & 64.9          \\
+AX16 & AX16-4a  & 33.70     & 425       & 575       & 15 & 0.76 & 9     & 0.05    & 0.21    & 4.8            & 1.1             & PASS & 49.2       & 30.1          & 4.2      & 65.8          \\
+AX16 & AX16-1a  & 33.76     & 200       & 564       & 15 & 0.87 & 7     & 0.06    & 0.22    & 5.6            & 2.4             & PASS & 49.2       & 30.1          & 4.5      & 65.9          \\
+AX16 & AX16-5a  & 33.77     & 425       & 560       & 10 & 0.65 & 6     & 0.10    & 0.24    & 7.4            & 3.9             & PASS & 49.2       & 30.1          & 4.0      & 65.9          \\
+AX16 & AX16-2a  & 33.87     & 425       & 564       & 12 & 0.65 & 7     & 0.07    & 0.22    & 4.3            & 1.4             & PASS & 49.2       & 30.1          & 4.9      & 66.1          \\
+AX16 & AX16-9a  & 36.12     & 500       & 585       & 15 & 0.61 & 10    & 0.04    & 0.17    & 3.0            & 1.2             & PASS & 49.2       & 30.1          & 4.3      & 70.5          \\
+AX16 & AX16-10a & 40.05     & 510       & 585       & 14 & 0.62 & 10    & 0.04    & 0.20    & 2.8            & 1.1             & PASS & 49.2       & 30.1          & 5.5      & 78.2         \\
 \hline
@@
 All paleointensity values and corresponding virtual dipole moments are corrected for cooling rate by a factor of 0.75.
-Taking a threshold p value of 0.01, the independent two-sample t-test results show that specimen paleointensity distributions of site AX12 and AX16, and those of site AX11 and AX13 are indistinguishable; region Kennedy Creek have distinguishable distributions from the other two regions; Silver Bay region and Carlton Peak region are indistinguishable.}
+Taking a threshold p value of 0.01, the independent two-sample t-test results show that specimen paleointensity distributions of site AX12 and AX16, and those of site AX11 and AX13 are indistinguishable; region Kennedy Creek have distinguishable distributions from the other two regions; Silver Bay region and Carlton Peak region values are indistinguishable.}
 \begin{tabular}{cccccccccccc}

### 2022-04-13 PNAS_manuscript/Supporting Info.tex
Updates from Overleaf

@@
 \centering
-\caption{{Graph of predicted paleointensity overestimate due to slow cooling of the intrusive Beaver River diabase and anorthosite xenoliths relative to the cooling rate in laboratory following the model of \citealp{Halgedahl1980a}.
+\caption{{Graph of predicted paleointensity overestimate due to slow cooling of the intrusive Beaver River diabase and anorthosite xenoliths relative to the cooling rate in laboratory following the model of ref.
+\citealp{Halgedahl1980a}.
 Because the majority of the anorthosites have unblocking temperatures between 500\textdegree C and 580\textdegree C, we estimate that the slow cooling during natural remanence acquisition could have resulted in a 33\% overestimate.

### 2022-04-13 PNAS_manuscript/PNAS_Revision_Letter.tex

@@
 
-\hfill 2021
+\hfill 2022
 %{\noindent\textit {Yiming Zhang, UC Berkeley}}
@@
 
-The sentence ``The possibility of late inner core nucleation has motivated proposals of novel power sources to sustain the geomagnetic field through early Earth history including precipitation of light-element minerals such as MgO \cite{Badro2016a, ORourke2016a, ORourke2016b} and SiO$_2$ \cite{Mittal2020a} at the core-mantle boundary.'' has been added which refers to the reviewer's point here.
+The sentence ``The possibility of late inner core nucleation has motivated proposals of novel power sources to sustain the geomagnetic field through early Earth history including precipitation of light-element minerals such as MgO \citep{Badro2016a, ORourke2016a, ORourke2016b} and SiO$_2$ \citep{Mittal2020a} at the core-mantle boundary.'' has been added which refers to the reviewer's point here.
 
@@
 
-We have added ``Dashed red and blue lines show bounding regions associated with the \textit{SCAT} statistic \cite{Shaar2013a}." to the figure caption.
+We have added ``Dashed red and blue lines show bounding regions associated with the \textit{SCAT} statistic \citep{Shaar2013a}." to the figure caption.
 
@@
 An estimated cooling time for the anorthosite from the Curie temperature of magnetite to 500\textdegree C is about 1.5 kyr.
-This corresponds to a cooling rate of $1.7\times10^{-9}$.
+This corresponds to a cooling rate of $1.7\times10^{-9}$ $^\circ$C s$^{-1}$.
 
@@
 With the increase in specificity associated with our cooling model the text now reads: ``From the thermal history model of \cite{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they blocked in the majority of their characteristic natural remanence magnetization ($\sim$500\textdegree  C; Fig.
-\ref{fig:IZZI_examples}; \citealp{Zhang2021b}).
+3; \citealp{Zhang2021b}).
 We find the cooling time to be $\sim$1.5 kyr, which corresponds to a cooling rate of $\sim1.7\times10^{-9}$ $^\circ$C s$^{-1}$."
@@
 
-We have added reference to our table S3 and added statement of t-test results within and between regions in the caption for table S3: `` Taking a threshold p value of 0.01, the independent two-sample t-test results show that specimen paleointensity distributions of site AX12 and AX16, and those of site AX11 and AX13 are indistinguishable; region Kennedy Creek have distinguishable distributions from the other two regions; Silver Bay region and Carlton Peak region values are indistinguishable."
+We have added reference to our table S3 and added statement of t-test results within and between regions in the caption for table S3: ``Taking a threshold p value of 0.01, the independent two-sample t-test results show that specimen paleointensity distributions of site AX12 and AX16, and those of site AX11 and AX13 are indistinguishable; region Kennedy Creek have distinguishable distributions from the other two regions; Silver Bay region and Carlton Peak region values are indistinguishable."
 
@@
 These are good questions that we also wonder about and would like to explore further.
-\cite{Ageeva2016a, Ageeva2017a, Ageeva2020a, Bian2021, Ageeva2022a} provides a a series of detailed petrologic and crystallographic investigation into the origin of the oxide needles that exsolve within plagioclase.
+\cite{Ageeva2016a, Ageeva2017a, Ageeva2020a, Ageeva2022a,Bian2021a} provides a a series of detailed petrologic and crystallographic investigation into the origin of the oxide needles that exsolve within plagioclase.
 \cite{Bian2021a} concludes that these oxides form above 600\textdegree C, thus above the Curie temperature of magnetite.
@@
 It has been found that partitioning of Fe into plagioclase is enhanced by high anorthite content of the plagioclase, by high oxygen fugacity, and by high silica content of the melt \citep{Phinney1992a, Longhi1976a}.
-Our Beaver River anorthosite xenoliths do have higher anorthite content ($\sim$70 An; bytownite) as compared to the plagioclase of the Duluth Complex ($\sim$60 An, labradorite).
-We suspect that the anorthosite xenoliths form during the early stage of the magmatic evolution of the Beaver River diabase when anorthite content, and likely the oxygen fugacity and silica content in the magma was high, leading to more partition of Fe into plagioclase crystal lattice, and minor amount of Fe to exsolve and form oxide needles (although those that did could still be crystallographically controlled), instead of forming abundant, large Fe-oxide exsolutions like those in the Duluth anorthosite.
+Our Beaver River anorthosite xenoliths do have higher anorthite content ($\sim$70\% An; bytownite) as compared to the plagioclase of the Duluth Complex ($\sim$60\% An; labradorite).
+We suspect that the anorthosite xenoliths form during the early stage of the magmatic evolution of the Beaver River diabase when anorthite content, and likely the oxygen fugacity and silica content in the magma were high, leading to more partition of Fe into plagioclase crystal lattice, and minor amount of Fe to exsolve and form oxide needles (although those that did could still be crystallographically controlled), instead of forming abundant, large Fe-oxide exsolutions like those in the Duluth anorthosite.
 
@@
 
-The sentence has been changed to more fully describe the result of higher coercivity for these samples ``Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites having successful paleointensity results can contain magnetic grain populations that have higher coercivities.
+The sentence has been changed to more fully describe the result of higher coercivity for these samples ``Single-component fits for coercivity spectra \citep{Maxbauer2016a} show that anorthosites having successful paleointensity results can contain magnetic grain populations that have higher coercivities.
 For these successful samples, median destructive destructive (MDF) field values associated with back-field demagnetization experiments range from 44 to 144 mT with a median of 53 mT (Fig.
-\ref{fig:coercivity}).
+5).
 In contrast, other anorthosite and diabase specimens tend to have lower peak coercivities (MDF range of 18-35 mT with a median of 22 mT; Fig.
-\ref{fig:coercivity})."
+5)."
 
@@
 The text now reads ``From the thermal history model of \cite{Zhang2021b}, we can estimate the duration over which the diabase and anorthosite cooled from the Curie temperature of magnetite ($\sim$580\textdegree C) to the time when they blocked in the majority of their characteristic natural remanence magnetization ($\sim$500\textdegree  C; Fig.
-\ref{fig:IZZI_examples}; \citealp{Zhang2021b}).
+3; \citealp{Zhang2021b}).
 We find the cooling time to be on the order of $\sim$1.5 kyr, which corresponds to a cooling rate of $\sim1.7\times10^{-9}$ $^\circ$C s$^{-1}$.
@@
 This point is a good one given that PSD and MD grains may be less biased toward recording higher paleointensities than SD grains associated with to slow cooling.
-We have added text ``Remanence held by vortex state (pseudo-single domain) and multidomain grains is not as biased by cooling rate \cite{Biggin2013a}.
-The potential for some remanence to be held by these grains, as suggested by slightly zigzagging Arai plots (Figure \ref{fig:IZZI_examples}), could mean that this factor is an over-correction and true paleointensity values could be slightly higher than those reported here.
+We have added text ``Remanence held by vortex state (pseudo-single domain) and multidomain grains is not as biased by cooling rate \citep{Biggin2013a}.
+The potential for some remanence to be held by these grains, as suggested by slightly zigzagging Arai plots (Fig.
+3), could mean that this factor is an over-correction and true paleointensity values could be slightly higher than those reported here.
 The cooling rate-corrected specimen paleointensity estimates together with specimen- and site-level means are shown in Fig.
-\ref{fig:PINT_cooling_corrected}." to the related paragraph.
+4." to the related paragraph.
 

### 2022-04-13 PNAS_manuscript/PNAS_manuscript.tex

@@
 Purple stars mark sites with paleointensity results that passed the selection criteria from this study.
-Paleomagnetic sites from \citealp{Pesonen1983a} are categorized by lithology.
-All sites from  \citealp{Pesonen1983a, Kulakov2013a, Sprain2018a} are color-coded by their ages.}}
+Paleomagnetic sites from ref.
+\citealp{Pesonen1983a} are categorized by lithology.
+All sites from refs.
+\citealp{Pesonen1983a, Kulakov2013a, Sprain2018a} are color-coded by their ages.}}
 \label{fig:Geologic_map}
@@
 \ref{fig:Petro_QDM}F, H).
-As a result of this shape anisotropy, remanent magnetizations are acquired at angles highly oblique to applied fields (Figure \ref{fig:Petro_QDM}F).
+As a result of this shape anisotropy, remanent magnetizations are acquired at angles highly oblique to applied fields (Fig.
+\ref{fig:Petro_QDM}F).
 In an experiment where a second orthogonal field was applied, the first isothermal remanent magnetizations (IRM) of a set of magnetite needles within a plagioclase grain either fail to rotate or flip by 180\textdegree.
@@
 A 20 mT alternating field treatment after in-field heating steps was applied to some specimens, but this treatment did not result in significant changes in the experimental results for the anorthosite xenolith or diabase specimens (Fig.
-\ref{fig:PINT_cooling_corrected}; Supporting Information).
+\ref{fig:PINT_cooling_corrected}).
 
@@
 Thermal modeling results from ref.
-\citealp{Zhang2021b} indicate that the Beaver River anorthosite xenoliths were heated to tholeiitic magma temperature ($\sim$1100\textdegree C)---lower than the melting temperature of the anorthosite plagioclase (given its composition of An$_{70}$; \citealp{Morrison1983a, Doyle2016a}) and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a few thousand years, partially averaging secular variation within single sites.
+\citealp{Zhang2021b} indicate that the Beaver River anorthosite xenoliths were heated to tholeiitic magma temperature ($\sim$1100\textdegree C)---lower than the melting temperature of the anorthosite plagioclase (given its composition of 70\% anorthite; \citealp{Morrison1983a, Doyle2016a}) and acquired thermal remanent magnetization during cooling with their diabase host on a time scale of a few thousand years, partially averaging secular variation within single sites.
 Another consequence of the anorthosite xenoliths having slowly cooled in the interior of thick diabase intrusions is that slow cooling rates can bias paleointensity estimates toward higher values \cite{Halgedahl1980a}.
@@
 Remanence held by vortex state (pseudo-single domain) and multidomain grains is not as biased by cooling rate \cite{Biggin2013a}.
-The potential for some remanence to be held by these grains, as suggested by slightly zigzagging Arai plots (Figure \ref{fig:IZZI_examples}), could mean that this factor is an over-correction and true paleointensity values could be slightly higher than those reported here.
+The potential for some remanence to be held by these grains, as suggested by slightly zigzagging Arai plots (Fig.
+\ref{fig:IZZI_examples}), could mean that this factor is an over-correction and true paleointensity values could be slightly higher than those reported here.
 The cooling rate-corrected specimen paleointensity estimates together with specimen- and site-level means are shown in Fig.
@@
 \citealp{Pesonen1983a} and \citealp{Kulakov2013a} are not included in the compilation due to the specimen-level double-slope behavior as discussed in the text.
-Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from ref.
+Overall, the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from ref.
 \citealp{Bono2019a} which is based on an interpretation of there being a monotonic decay of the geodynamo through the Proterozoic.
@@
 In addition to these high geomagnetic fields recorded by Midcontinent Rift rocks, the ca.
-755 Ma Mundine Well dikes \cite{Lloyd2021b} also require a stronger geomagnetic field in the Neoproterozoic than would be predicted by a progressive Proterozoic decline (Fig.
+755 Ma Mundine Well dikes \cite{Lloyd2021b} also require a stronger geomagnetic field in the Neoproterozoic than would be predicted by a progressive Proterozoic decline.
+Taken together, these results call into question whether the progressively decaying polynomial fit implemented by ref.
+\cite{Bono2019a} is an accurate representation for the evolution of the Proterozoic geomagnetic field (Fig.
 \ref{fig:PINT_compilation}).
@@
 $\beta$, FRAC, GAP-Max and SCAT are all statistics to assess the behavior of NRM/TRM plots.
-See the Standard Paleointensity Definitions (\citealp{Paterson2014a}; \url{https://earthorg/PmagPy/SPD/home.html}) for more details.
+See the Standard Paleointensity Definitions (\citealp{Paterson2014a}; \url{https://earthref.org/PmagPy/SPD/home.html}) for more details.
 Data analysis was conducted using Thellier GUI \cite{Shaar2013a} within the PmagPy software package \cite{Tauxe2016a}.

### 2022-04-14 PNAS_manuscript/PNAS_Revision_Letter.tex
update manuscript

@@
 
-We have changed the sentence to ``The anorthosite xenoliths show low anisotropy of thermal remanent magnetization (TRM) acquisition and can acquire TRM linearly within the range of relevant field strengths."
+We have changed the sentence to ``The anorthosite xenoliths show low anisotropy of thermal remanent magnetization acquisition and can acquire TRM linearly within the range of relevant field strengths."
 

### 2022-04-14 PNAS_manuscript/PNAS_manuscript.tex

@@
 The anorthosite xenoliths targeted in this study were brought to the near surface in magma that formed hypabyssal (shallowly emplaced) intrusions of the Beaver River diabase \cite{Zhang2021b}.
-They would have been heated to tholeiitic magma temperature which at $\sim$1100\textdegree C is below the melting point of the xenolith plagioclase (70$\%$ anorthite), but well above the Curie temperature of magnetite \cite{Zhang2021b}.
+They would have been heated to tholeiitic magma temperature ($\sim$1100\textdegree C) which is below the melting point of the xenolith plagioclase (70$\%$ anorthite), but well above the Curie temperature of magnetite \cite{Zhang2021b}.
 Fe-Ti oxides within the plagioclase likely exsolved above magnetite Curie temperature \cite{Bian2021a} and subsequently cooled and acquired thermal remanent magnetizations in conjunction with the host diabase at a paleolatitude of 22 $\pm$ 2\textdegree$\;$(calculated from the paleomagnetic pole of the coeval Portage Lake Volcanics; \citealp{Swanson-Hysell2019a}).
@@
 \centering
-\noindent\includegraphics[width=13 cm]{Petro_QDM_full.png}
+\noindent\includegraphics[width=12.4 cm]{Petro_QDM_full.png}
 \caption{\footnotesize{Thin section petrographic images (A,B,C,D) and magnetic maps (E,F,G,H) of an anorthosite sample from the Beaver River anorthosite xenolith in the Silver Bay region from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (left column;  \citealp{Zhang2021b}) and from a distinct anorthosite within the Duluth Complex Anorthositic Series (right column).
@@
 the individual dipoles visible with paired red +B$_z$ and blue -B$_z$ lobes) align well with the first applied field direction (E) and then rotated to align with the second applied field direction indicating minimal anisotropic behavior (G).
-In contrast, the remanence magnetizations carried by the magnetic needles in plagioclase 1 of the Duluth anorthosite xenolith align well with the first applied field but those in plagioclase 2 acquired an oblique remanence direction with respect to the field direction (F).
+In contrast, the remanent magnetizations carried by the magnetic needles in plagioclase 1 of the Duluth anorthosite xenolith align well with the first applied field but those in plagioclase 2 acquired an oblique remanence direction with respect to the field direction (F).
 After the application of an orthogonal 300 mT external field, magnetization of those needles in plagioclase 1 did not change direction due to strong shape anisotropy whereas the magnetization of needles in plagioclase 2 flipped, but with the acquired remanence remaining oblique to the field direction (H).
-Insets in (E) to (H) show example dipoles direction responses to the application of orthogonal IRMs.
-This experiment illustrates that the Beaver River anorthosite xenolith has much lower magnetic anisotropy than the Duluth anorthosite xenolith.
+Insets in (E) to (H) show example dipole directions (fit using the algorithm of \cite{Lima2016a}) in response to the application of orthogonal IRMs.
+The directional changes of the Beaver River anorthosite xenolith sources indicate a lower magnetic anisotropy compared to the relative lack of change for the Duluth Complex anorthosite xenolith sources.
 The unit of the axes in the QDM maps are in $\mu$m.}}

### 2022-05-06 PNAS_manuscript/PNAS_manuscript.tex
Updates from Overleaf
Implement edits according to USGS comments

@@
 \centering
-\noindent\includegraphics[width=12.4 cm]{Petro_QDM_full.png}
+\noindent\includegraphics[width=12.4 cm]{Petro_QDM_full.pdf}
 \caption{\footnotesize{Thin section petrographic images (A,B,C,D) and magnetic maps (E,F,G,H) of an anorthosite sample from the Beaver River anorthosite xenolith in the Silver Bay region from which paleomagnetic site AX16 and geochronology sample MS99033 were collected (left column;  \citealp{Zhang2021b}) and from a distinct anorthosite within the Duluth Complex Anorthositic Series (right column).
@@
 The B$_z$ color scale is an order of magnitude greater in the Duluth Complex anorthosite maps (F, H) than the maps for the Beaver River anorthosite (E, G).
-Panels (E) to (G) and (F) to (H) show experiments performed on both samples where we apply a first field along the y axis of the field of view and then apply a second field of 300 mT orthogonal to the first field direction.
+Panels (E) to (G) and (F) to (H) show experiments performed on both samples where we apply a first field along the y-axis of the field of view and then apply a second field of 300 mT orthogonal to the first field direction.
 The remanent magnetizations acquired by both anorthosites were mapped after the application of each field.
@@
 \caption{\footnotesize{Example results of paleointensity experiments are displayed on Arai plots and zero-field heating results are shown as inset orthogonal plots (Zijderveld plots) for anorthosite and diabase specimens.
+Green squares and lines in Arai plots and green squares in orthogonal plots show the range of data points used for fitting.
 Red (blue) circles indicate zero-field/in-field (in-field/zero-field) steps `ZI’ (`IZ’).
@@
 \subsubsection*{Coercivity}
-Additional rock magnetic data indicate that those anorthosite specimens which pass the paleointensity selection criteria contain dominant magnetic remanence carriers with magnetic properties similar to stochiometric, non-interacting, single domain magnetite, whereas anorthosite samples that failed the paleointensity result selection along with all diabase samples have more pronounced populations of non-ideal carriers.
+Additional rock magnetic data indicate that those anorthosite specimens which pass the paleointensity selection criteria contain dominant magnetic remanence carriers with magnetic properties similar to stoichiometric, non-interacting, single domain magnetite, whereas anorthosite samples that failed the paleointensity result selection along with all diabase samples have more pronounced populations of non-ideal carriers.
 Magnetic property measurement system (MPMS) data show that both the diabase and anorthosite contain (titano)magnetite as evidenced by the presence of the Verwey transition (Fig.
@@
 Single-component fits for coercivity spectra \cite{Maxbauer2016a} show that anorthosites having successful paleointensity results can contain magnetic grain populations that have higher coercivities.
-For these successful samples, median destructive destructive (MDF) field values associated with back-field demagnetization experiments range from 44 to 144 mT with a median of 53 mT (Fig.
+For these successful samples, median destructive (MDF) field values associated with back-field demagnetization experiments range from 44 to 144 mT with a median of 53 mT (Fig.
 \ref{fig:coercivity}).
@@
 Bottom: Box plots of median destructive field (MDF) values associated with back-field demagnetization for all anorthosite and diabase specimens with single-component coercivity unmixing results.
+Bars in the boxes show median values and the whiskers show the minimum and the maximum value of the datasets.
 Both plots show that anorthosite specimens that pass paleointensity selection criteria have higher coercivities consistent with a higher portion of single-domain-like magnetite grains than the other anorthosite specimens and the diabase.}}
@@
 \subsection*{Considering secular variation and cooling rate}
-To best characterize the geomagnetic axial dipole field intensity during a certain time period a paleointensity dataset should cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged.
+To best characterize the geomagnetic axial dipole field intensity during a certain time period, a paleointensity dataset should cover a sufficient amount of time such that paleosecular variations of the geomagnetic field are averaged.
 Thermal modeling results from ref.
@@
 1092 Ma.
-To contextualize the high paleointensity values, we plot the regional mean values (Table.
-S3) based on specimen-level paleointensity results together with calculated paleointensity values at 22\textdegree N latitude based on the PADM2M model of the time-varying geomagnetic field over the past 2 million years \cite{Ziegler2011a} in Figure \ref{fig:PINT_cooling_corrected}B.
-The locality-mean values of these three regions are plotted in Figure \ref{fig:PINT_cooling_corrected}C (and reported in Table.
-S3) where we also present the distribution of means calculated from 3 random values taken from the PADM2M model for a total of 10000 iterations.
+To contextualize the high paleointensity values, we plot the regional mean values (Table S3) based on specimen-level paleointensity results together with calculated paleointensity values at 22\textdegree N latitude based on the PADM2M model of the time-varying geomagnetic field over the past 2 million years \cite{Ziegler2011a} in Figure \ref{fig:PINT_cooling_corrected}B.
+The locality-mean values of these three regions are plotted in Figure \ref{fig:PINT_cooling_corrected}C (and reported in Table S3) where we also present the distribution of means calculated from 3 random values taken from the PADM2M model for a total of 10000 iterations.
 In addition, we perform the same comparison with the existing paleointensity data for the Cenozoic Era (the last 66 million years) from the PINT database (filtered by Q$_{PI}\geq$3; Fig.
@@
 The highest site-mean virtual dipole moment of the anorthosites would rank in the top 2$\%$ of those in the database for the Cenozoic Era (the last 66 million years) when there was unequivocally a crystallizing inner core.
-The y axis maximum is set at the 99 percentile of the compiled Cenozoic paleointensity data.}}
+The y-axis maximum is set at the 99th percentile of the compiled Cenozoic paleointensity data.}}
 \label{fig:PINT_compilation}
@@
 408–393 Ma Buribay volcanics \cite{Shcherbakova2021a} and the ca.
-332 Ma Kinghorn volcanics \cite{Hawkins2021a} has led to the proposal of this interval being the ``Mid-Paleozoic Dipole Low'' \cite{Hawkins2021a}.
-This ``Mid-Paleozoic Dipole Low'' is followed by high paleointensity values such that there is a six-fold increase in the database from virtual dipole moment of $\sim$16 ZAm$^2$ ca.
+332 Ma Kinghorn volcanics \cite{Hawkins2021a} have led to the proposal of this interval being the ``Mid-Paleozoic Dipole Low'' \cite{Hawkins2021a}.
+This ``Mid-Paleozoic Dipole Low'' is followed by high paleointensity values such that there is a six-fold increase from virtual dipole moments of $\sim$16 ZAm$^2$ ca.
 332 Ma to $\sim$99 ZAm$^2$ ca.
@@
 We thank Jim Miller for providing us with the Duluth Complex anorthosite thin section.
-We thank Dario Bilardello, Peat Solheid, Mike Jackson, Josh Feinberg, and Bruce Moskowitz for their tremendous help associated with instrument operation, data interpretations, and research guidance at the IRM.
+We thank Dario Bilardello, Peat Solheid, Mike Jackson, Josh Feinberg, and Bruce Moskowitz for their tremendous help with instrument operation, data interpretations, and research guidance at the IRM.
 Conversations with Bruce Buffett and Zachary Geballe informed our perspectives on the geodynamic and mineral physics literature.
 We thank Jon Hagstrum at USGS and two anonymous reviewers for their helpful comments on the manuscript.
-Paleomagnetic data associated with this study are available within the MagIC database (\url{https://earthref.org/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}; \textit{provide link when doi is generated at time of proofs}) and all data are within a Github repository along with Python code related to calculations and visualizations (\url{https://github.com/Swanson-Hysell-Group/AX_BD_PINT}) that are also archived on Zenodo (\textit{provide link when doi is generated at time of proofs}).}
+Paleomagnetic data associated with this study are available within the MagIC database (\url{https://earthref.org/MagIC/19462/8d3c2258-11ae-4830-b99f-3f6b02eceb7e}; \textit{provide link when doi is generated at time of proofs}) and all data are within a Github repository along with Python code related to calculations and visualizations (\url{https://github.com/Swanson-Hysell-Group/AX_BD_PINT}) that are also archived on Zenodo (\textit{provide link when doi is generated at time of proofs}).
+Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
+Government.}
 

### 2022-05-18 PNAS_manuscript/PNAS_manuscript.tex
edits for final files

@@
 However, robust estimates of ancient field strengths are often difficult to recover due to alteration or non-ideal behavior.
-We use anorthosites that formed in the deep crust and were brought to the near surface where they acquired thermal remanent magnetizations.
-These anorthosites have experienced minimal post-formation alteration and yield high-quality paleointensity estimates.
-In contrast to scenarios of a progressively decaying field leading up to late nucleation of Earth's inner core, these data record a strong field 1.1 Ga.
-A strong field that persisted over a 14-million-year interval indicates the existence of appreciable power sources for Earth's dynamo during the late Mesoproterozoic Era.}
+We use rocks known as anorthosite that formed in the deep crust and were brought to the near surface where they acquired thermal remanent magnetizations.
+These rocks have experienced minimal post-formation alteration and yield high-quality paleointensity estimates.
+In contrast to scenarios of a progressively decaying field leading up to proposed late nucleation of Earth's inner core, these data record a strong field 1.1 billion years ago.
+A strong field that persisted over a 14-million-year interval indicates the existence of appreciable power sources for Earth's dynamo at this time.}
 
